### 2013-07-22 text/ms.tex
some more text

\documentclass[12pt,preprint]{aastex}

\usepackage{amsmath}

\begin{document}

\title{Something about Lens Models}

\author{First Author,\altaffilmark{1}
Second Author,\altaffilmark{2} and
Third Author\altaffilmark{3}}
\altaffiltext{1}{First place}
\altaffiltext{2}{Second place}
\altaffiltext{3}{Third place}

\begin{abstract}
This is an abstract.
\end{abstract}

\keywords{}

\section{Introduction}

See Figure \ref{fig:ASW0000kad}.

\appendix

\section{Essentials of Lensing Theory}

First we define comoving distances, as follows.
\begin{equation}
\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
\end{equation}
\begin{equation}
\begin{aligned}
&D_S    &                                &\int_0^{a_S} \\
\noalign{\medskip}
&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
\noalign{\medskip}
&D_L    &                                &\int_0^{z_L}
\end{aligned}
\end{equation}

Then define the arrival time, following \cite{1986ApJ...310..568B}.
Let $x,y$ be physical distances on the lens plane and $D_L$ be
comoving distances.  Then
\begin{equation}
l_{\rm geom} = \frac1{2D_L} \frac{D_S}{D_{LS}}
\left( (x-s_x)^2 + (y-s_y)^2 \right)
\end{equation}
\begin{equation}
\nabla^2 l_{\rm grav} = \frac{8\pi G}{c^2} \Sigma(x,y)
\end{equation}
But $l_{\rm grav}$ expands by $(1+z)$.

It is usual to work in terms of
\begin{equation}
\kappa = \frac{4\pi G}{c^2} \frac{D_L}{1+z} \frac{D_{LS}}{D_S} \Sigma
\end{equation}
known as the convergence.

The Laplacian $\nabla^2 f(x,y)$ is
\begin{equation}
 \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
      {\Delta x \; \Delta y}
\end{equation}
in the limit of $\Delta x,\Delta y$ small.

\newpage

\bibliographystyle{apj}
\bibliography{ms}

\newpage

\begin{figure}
\includegraphics[width=.3\hsize]{ASW0000kadC_kappa.png}
\includegraphics[width=.3\hsize]{ASW0000kadC_arriv.png}
\includegraphics[width=.3\hsize]{ASW0000kadC_menc.png}
\caption{This is a figure.\label{fig:ASW0000kad}}
\end{figure}

\end{document}

### 2013-07-24 text/ms.tex
writing text

@@
 \usepackage{amsmath}
+\def\half{{\textstyle\frac12}}
 
@@
 
+\section{Fermat's Principle}
+
+Explain Fermat's principle
+\begin{equation}
+\hbox{arrival time} = A_{\rm geom} - A_{\rm grav}
+\end{equation}
+By convention, $A_{\rm grav}$ is defined in such a way that it is
+negative, hence the minus sign.
+
+Although $A_{\rm geom}$ and $A_{\rm grav}$ are times, we are going to
+write them as if they are areas.  In other words, we will suppress a
+constant factor in the $A$.  That factor is basically the lens
+distance times the speed of light, but the precise expression is given
+in the Appendix.
+
+Introduce the Laplacian $\nabla^2 f(x,y)$ is
+\begin{equation}
+ \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
+        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
+      {\Delta x \; \Delta y}
+\end{equation}
+in the limit of $\Delta x,\Delta y$ small.
+
+Let $x,y$ be coordinates on the lens, transverse to the line of sight.
+Let there be a point source behind $x=0,y=0$.  Then
+\begin{equation}
+A_{\rm geom}(x,y) = \half(x^2 + y^2)
+\end{equation}
+\begin{equation}
+\nabla^2 A_{\rm grav}(x,y) = 2\kappa(x,y)
+\end{equation}
+Here $\kappa$ is a dimensionless quantity with two distinct physical
+interpretations.  On the one hand, it is the sky projected density ---
+in special units, to make it dimensionless.  On the other hand,
+$\kappa$ is the {\em convergence\/} of a bundle of rays.
+
+\section{SpaghettiLens}
+
+Give a guess for the maximum, minimum and saddle points.  Program
+tries find $\kappa(x,y)$ that reproduces these properties exactly, and
+looks reasonably like a galaxy.  Solution not unique, an ensemble
+generated.
+
+\section{The simulated lenses}
+
+Three kinds of sims.
+
+\begin{itemize}
+\item {\em Quasar:\/} A singular elliptical isothermal plus constant
+  external shear, with a circular Gaussian source.
+\item {\em Galaxy:\/} Lens as with quasar, but elliptical de
+  Vaucouleurs source.
+\item {\em Galaxy:\/} Lens is circular NFW plus one dominant
+  elliptical SIE and perturbating elliptical SIEs, source as with
+  galaxy.
+\end{itemize}
+Lenses follow \cite{2001astro.ph..2341K,2001astro.ph..2340K}.
+Parameters in online supplement.
+
+\section{Results}
+
 \appendix
 
-\section{Essentials of Lensing Theory}
+\section{More on Lensing Theory}
 
@@
 
-Then define the arrival time, following \cite{1986ApJ...310..568B}.
+\begin{equation}
+\begin{aligned}
+A           &= cD_L \frac{D_{LS}}{D_S} \times t \\
+\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L}{1+z} \frac{D_{LS}}{D_S}
+               \times \Sigma(x,y)
+\end{aligned}
+\end{equation}
+
 Let $x,y$ be physical distances on the lens plane and $D_L$ be
@@
 \begin{equation}
-l_{\rm geom} = \frac1{2D_L} \frac{D_S}{D_{LS}}
+t_{\rm geom} = \frac{(1+z_L)^2}{2cD_L} \frac{D_S}{D_{LS}}
 \left( (x-s_x)^2 + (y-s_y)^2 \right)
@@
 \begin{equation}
-\nabla^2 l_{\rm grav} = \frac{8\pi G}{c^2} \Sigma(x,y)
+\nabla^2 t_{\rm grav} = (1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
 \end{equation}
-But $l_{\rm grav}$ expands by $(1+z)$.
 
-It is usual to work in terms of
+Can replace comoving distances $D$ with angular-diameter distances
+$d$, as follows
 \begin{equation}
-\kappa = \frac{4\pi G}{c^2} \frac{D_L}{1+z} \frac{D_{LS}}{D_S} \Sigma
+\begin{aligned}
+D_L    &= (1+z_L) d_{OL}
+D_{LS} &= \frac{1+z_S}{1+z_L} d_{LS}
+D_S    &= (1+z_L) d_{OS}
+\end{aligned}
 \end{equation}
-known as the convergence.
+and
+\begin{equation}
+(x,y) = d_{OL} (\theta_x,\theta_y)
+\end{equation}
 
-The Laplacian $\nabla^2 f(x,y)$ is
-\begin{equation}
- \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
-        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
-      {\Delta x \; \Delta y}
-\end{equation}
-in the limit of $\Delta x,\Delta y$ small.
+Recover equations (2.1) to (2.6) from \cite{1986ApJ...310..568B}.
 

### 2013-07-24 spaghetti/analysis.py
analysis plots

@@ -8,7 +8,6 @@ then:
 python glass.py analysis.py
 
 '''
-
 from __future__ import division
 
 import os
@@ -28,14 +27,17 @@ from matplotlib import rc
 rc('text', usetex=False)
 
 
+test = False
+
+
 #os.environ['LD_LIBRARY_PATH'] = '/win/proj/master/glassMint/build/glpk_build/lib'
 
 glass_basis('glass.basis.pixels', solver=None)
 exclude_all_priors()
 
 
-data_dir = '../../SimAnalysis/spaghetti/demodata'
 data_dir = '/srv/lmt/tmp_media'
+if test: data_dir = '../../SimAnalysis/spaghetti/demodata'
 data_dir = os.path.abspath(data_dir)
 
 
@@ -54,6 +56,9 @@ files  = [
   '003897','003916','003925','003926','003943'
   ]
 
+# local test run
+if test: files = ['003211']
+
 # i've gotten the filenames with:
 # re.findall('mite\.physik\.uzh\.ch\/data\/(\d+)', txt)
 # in the html source
@@ -61,6 +66,7 @@ files  = [
   
 auth_re = re.compile('meta\(author=\'(.*)\',')
 inf_re = re.compile('globject\(\'(\d+)__(.+)\'')
+nmod_re = re.compile('model\((\d+)')
 
   
 distance_factor = 0.428
@@ -78,7 +84,7 @@ csvfilen = csv.writer(filen)
 #csvfile2 = csv.writer(file0)
   
 for iii, fl in enumerate(files):
-  print 'file %3i / %3i' % (iii, len(files))
+  print 'file %3i / %3i' % (iii, len(files)),
 
   fstr = os.path.join(data_dir, fl, 'state.txt')
   g = loadstate(fstr)
@@ -105,8 +111,11 @@ for iii, fl in enumerate(files):
     pixPerRing[i] = len(obj.basis.rings[i])
     for j in range(i):
       pixEnc[i] += len(obj.basis.rings[j])
-
+  
+  print '  |',n_rings,
   for k in range(n_rings): #pixrad
+    print '*',
+    if (k+1)%3==0: print '\'',
     kappaRenc_k_all = np.zeros(0)
     for m in g.models:
       obj,ps = m['obj,data'][0]
@@ -126,13 +135,16 @@ for iii, fl in enumerate(files):
[…]

### 2013-07-24 spaghetti/generate_plots.py

@@ -9,39 +9,57 @@ import csv
 import numpy as np
 import matplotlib.pyplot as plt
 
+#from matplotlib import rc
+
+#rc('font',**{'family':'sans-serif','sans-serif':['Helvetica']})
+## for Palatino and other serif fonts use:
+#rc('font',**{'family':'serif','serif':['Palatino']})
+#rc('text', usetex=True)
+
 
 sims_dir = '../systems/figs'
-
 user_data_dir = './data.csv'
+save_fig_path = './figs'
+
+div_scale_factors = 440./500*100
 
 data = []
 
 # read the user generated data
 with open(user_data_dir, 'r') as f:
+
+  c = 6 # offset for real data
+
   csvfile = csv.reader(f)
   
   for row in csvfile:
     l = len(row)
 
-    if l-4 != int(row[3])*4:
-      print '!! strange1 error @', row[0], 4*int(row[3]), l-4
-      continue
+    dl = (l-c-int(row[4])*2) / 4  
     
-    dl = (l-4) / 4  
-    
-    if (l-4)%4 != 0:
+    #print l, dl, (l-5-int(row[4])*2) % 4
+    if (l-c-int(row[4])*2) % 4 != 0:
       print '!! strange2 error @', row[0]
       continue 
     
+    points = []
+    pntdata = row[c+4*dl:]
+    for ii in range(int(row[4])):
+      points.append({'d':float(pntdata[ii*2])*div_scale_factors,'t':pntdata[ii*2+1]})
+      
     o = {
       'id':     int(row[0]),
       'user':   row[1],
       'name':   row[2],
-      'nr':     row[3],
-      'x':      np.array(row[4:4+dl], dtype=float32),
-      'y':      np.array(row[4+dl: 4+2*dl], dtype=float32),
-      'err_p':  np.array(row[4+2*dl: 4+3*dl], dtype=float32),
-      'err_m':  np.array(row[4+3*dl: 4+4*dl], dtype=float32)
+      'nr':     int(row[3]),
+      'pxR':    int(row[3])-1,
+      'nPnt':   int(row[4]),
+      'nMod':   int(row[5]),
+      'x':      np.array(row[c:c+dl], dtype=np.float32),
+      'y':      np.array(row[c+dl: c+2*dl], dtype=np.float32),
[…]

### 2013-07-24 text/ms.tex
some text

@@
 
-\title{Something about Lens Models}
+\title{Lens Modeling in Spacewarps}
 
@@
 \begin{abstract}
-This is an abstract.
+In Spacewarps, volunteers are invited to search sky surveys for lens
+candidates.  Here we report on a system that allows experienced
+volunteers to model lens candidates.  A sample of 29 simulated lenses
+were modelled, each multiple times.  The quality of the models was
+then examined in two ways.  First, the image parities and time
+orderings: these were identified correctly in most cases, but not
+always.  Second, the mean convergence (equivalent to the enclosed
+mass): this was constrained consistently well in the image region.
+The performance of professionals and experienced volunteers was
+comparable.
 \end{abstract}
@@
 \section{Introduction}
-
-See Figure \ref{fig:ASW0000kad}.
 
@@
 
+Example Figure \ref{fig:ASW0000ar2}.
+
 \appendix
@@
 Can replace comoving distances $D$ with angular-diameter distances
-$d$, as follows
+$d$,
+\begin{equation}
+(x,y) = d_L (\theta_x,\theta_y)
+\end{equation}
+and so on.
 \begin{equation}
 \begin{aligned}
-D_L    &= (1+z_L) d_{OL}
-D_{LS} &= \frac{1+z_S}{1+z_L} d_{LS}
-D_S    &= (1+z_L) d_{OS}
+D_S &= (1+z_S) \, d_S \\
+D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
+D_L &= (1+z_L) \, d_L
 \end{aligned}
 \end{equation}
-and
-\begin{equation}
-(x,y) = d_{OL} (\theta_x,\theta_y)
-\end{equation}
+Angular-diameter distances are always smaller than comoving distances.
 
@@
 \begin{figure}
-\includegraphics[width=.3\hsize]{ASW0000kadC_kappa.png}
-\includegraphics[width=.3\hsize]{ASW0000kadC_arriv.png}
-\includegraphics[width=.3\hsize]{ASW0000kadC_menc.png}
-\caption{This is a figure.\label{fig:ASW0000kad}}
+\includegraphics[width=.32\hsize]{ASW0000ar2Q_kappa.png}
+\includegraphics[width=.32\hsize]{ASW0000ar2Q_arriv.png} \\
+\includegraphics[width=.32\hsize]{ASW0000ar2_kappa.png}
+\includegraphics[width=.32\hsize]{ASW0000ar2_arriv.png}
+\includegraphics[width=.32\hsize]{ASW0000ar2_menc.png}
+\caption{A simple two-image lensed quasar.  The minimum and saddle are
+  identified correctly.  The mean model shows two spurious images, but
+  the mean convergence in the region of images is good.
+\label{fig:ASW0000ar2}}
 \end{figure}

### 2013-07-24 spaghetti/generate_plots.py

@@ -187,5 +187,7 @@ def plot_all(show=True, cens=False):
     
 def plot_all1():
   plot_all(False, True)
-  
+
+# plot_all(False)
+
 #plott(56, False, True)

### 2013-08-26 text/ms.tex
added images
restructured tex files

--- 2013-07-22 text/ms.tex
@@
 
+%%% PACKAGES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \usepackage{amsmath}
+\usepackage{xspace}   % fixes spaces after custom commands
+\usepackage{float}    % exact placement of figures (using \begin{figure}[H])
+%\usepackage{caption}  % better captions for floats
+
+%%% DEFINITIONS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\def\half{{\textstyle\frac12}}
+\newcommand{\spl}{SpaghettiLens\xspace}
+\newcommand{\sw}{SpaceWarps\xspace}
+
+%%% DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\title{Something about Lens Models}
-
+\title{Lens Modeling in \sw}
 \author{First Author,\altaffilmark{1}
@@
 
-\begin{abstract}
-This is an abstract.
-\end{abstract}
+\input{tex/abstract}
+\input{tex/intro}
+\input{tex/theory}
+\input{tex/spaghetti}
+\input{tex/results}
+\input{tex/outlook}
 
-\keywords{}
-
-\section{Introduction}
-
-See Figure \ref{fig:ASW0000kad}.
-
-\appendix
-
-\section{Essentials of Lensing Theory}
-
-First we define comoving distances, as follows.
-\begin{equation}
-\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
-\end{equation}
-\begin{equation}
-\begin{aligned}
-&D_S    &                                &\int_0^{a_S} \\
-\noalign{\medskip}
-&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
-\noalign{\medskip}
-&D_L    &                                &\int_0^{z_L}
-\end{aligned}
-\end{equation}
-
-Then define the arrival time, following \cite{1986ApJ...310..568B}.
-Let $x,y$ be physical distances on the lens plane and $D_L$ be
-comoving distances.  Then
-\begin{equation}
-l_{\rm geom} = \frac1{2D_L} \frac{D_S}{D_{LS}}
-\left( (x-s_x)^2 + (y-s_y)^2 \right)
-\end{equation}
-\begin{equation}
-\nabla^2 l_{\rm grav} = \frac{8\pi G}{c^2} \Sigma(x,y)
-\end{equation}
-But $l_{\rm grav}$ expands by $(1+z)$.
-
-It is usual to work in terms of
-\begin{equation}
-\kappa = \frac{4\pi G}{c^2} \frac{D_L}{1+z} \frac{D_{LS}}{D_S} \Sigma
-\end{equation}
-known as the convergence.
-
-The Laplacian $\nabla^2 f(x,y)$ is
-\begin{equation}
- \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
-        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
-      {\Delta x \; \Delta y}
-\end{equation}
-in the limit of $\Delta x,\Delta y$ small.
+\input{tex/appendix}
 
@@
 \bibliographystyle{apj}
-\bibliography{ms}
+\bibliography{bib/ms}
 
@@
 
-\begin{figure}
-\includegraphics[width=.3\hsize]{ASW0000kadC_kappa.png}
-\includegraphics[width=.3\hsize]{ASW0000kadC_arriv.png}
-\includegraphics[width=.3\hsize]{ASW0000kadC_menc.png}
-\caption{This is a figure.\label{fig:ASW0000kad}}
-\end{figure}
+\input{tex/figures}
 

### 2013-08-26 text/tex/abstract.tex
restructured tex files

\begin{abstract}
In Spacewarps, volunteers are invited to search sky surveys for lens
candidates.  Here we report on a system that allows experienced
volunteers to model lens candidates.  A sample of 29 simulated lenses
were modelled, each multiple times.  The quality of the models was
then examined in two ways.  First, the image parities and time
orderings: these were identified correctly in most cases, but not
always.  Second, the mean convergence (equivalent to the enclosed
mass): this was constrained consistently well in the image region.
The performance of professionals and experienced volunteers was
comparable.
\end{abstract}

\keywords{}

### 2013-08-26 text/tex/appendix.tex
restructured tex files

\appendix

\section{More on Lensing Theory} \label{sec:more-theory}

First we define comoving distances, as follows.
\begin{equation}
\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
\end{equation}
\begin{equation}
\begin{aligned}
&D_S    &                                &\int_0^{a_S} \\
\noalign{\medskip}
&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
\noalign{\medskip}
&D_L    &                                &\int_0^{z_L}
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
A           &= cD_L \frac{D_{LS}}{D_S} \times t \\
\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L}{1+z} \frac{D_{LS}}{D_S}
               \times \Sigma(x,y)
\end{aligned}
\end{equation}

Let $x,y$ be physical distances on the lens plane and $D_L$ be
comoving distances.  Then
\begin{equation}
t_{\rm geom} = \frac{(1+z_L)^2}{2cD_L} \frac{D_S}{D_{LS}}
\left( (x-s_x)^2 + (y-s_y)^2 \right)
\end{equation}
\begin{equation}
\nabla^2 t_{\rm grav} = (1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
\end{equation}

Can replace comoving distances $D$ with angular-diameter distances
$d$,
\begin{equation}
(x,y) = d_L (\theta_x,\theta_y)
\end{equation}
and so on.
\begin{equation}
\begin{aligned}
D_S &= (1+z_S) \, d_S \\
D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
D_L &= (1+z_L) \, d_L
\end{aligned}
\end{equation}
Angular-diameter distances are always smaller than comoving distances.

Recover equations (2.1) to (2.6) from \cite{1986ApJ...310..568B}.

### 2013-08-26 text/tex/intro.tex
restructured tex files

\section{Introduction}

Spacewarps is the future.

Moreover, that lens modelling is an essential accompaniment to lens
searches has been clear since the early days \citep{1981ApJ...244..723Y}.

Why Fermat's principle?  Details in \ref{sec:Fermat} and
\ref{sec:more-theory}.

SpaghettiLens details in \ref{sec:SpaghettiLens}.

Sims in \ref{sec:sims}.

Results in \ref{sec:results}.

Things to do in \ref{sec:todo}.

### 2013-08-26 text/tex/results.tex
restructured tex files

\section{Results} \label{sec:results}

Example Figure \ref{fig:ASW0000ar2}.

### 2013-08-26 text/tex/spaghetti.tex
restructured tex files

\section{SpaghettiLens} \label{sec:SpaghettiLens}

Give a guess for the maximum, minimum and saddle points.  Program
tries find $\kappa(x,y)$ that reproduces these properties exactly, and
looks reasonably like a galaxy.  Solution not unique, an ensemble
generated.

\section{The simulated lenses} \label{sec:sims}

Three kinds of sims.

\begin{itemize}
\item {\em Quasar:\/} A singular elliptical isothermal plus constant
  external shear, with a circular Gaussian source.
\item {\em Galaxy:\/} Lens as with quasar, but elliptical de
  Vaucouleurs source.
\item {\em Galaxy:\/} Lens is circular NFW plus one dominant
  elliptical SIE and perturbating elliptical SIEs, source as with
  galaxy.
\end{itemize}
Lenses follow \cite{2001astro.ph..2341K,2001astro.ph..2340K}.
Parameters in online supplement.

### 2013-08-26 text/tex/theory.tex
restructured tex files

\section{Fermat's Principle} \label{sec:Fermat}

Explain Fermat's principle
\begin{equation}
\hbox{arrival time} = A_{\rm geom} - A_{\rm grav}
\end{equation}
By convention, $A_{\rm grav}$ is defined in such a way that it is
negative, hence the minus sign.

Although $A_{\rm geom}$ and $A_{\rm grav}$ are times, we are going to
write them as if they are areas.  In other words, we will suppress a
constant factor in the $A$.  That factor is basically the lens
distance times the speed of light, but the precise expression is given
in the Appendix.

Introduce the Laplacian $\nabla^2 f(x,y)$ is
\begin{equation}
 \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
      {\Delta x \; \Delta y}
\end{equation}
in the limit of $\Delta x,\Delta y$ small.

Let $x,y$ be coordinates on the lens, transverse to the line of sight.
Let there be a point source behind $x=0,y=0$.  Then
\begin{equation}
A_{\rm geom}(x,y) = \half(x^2 + y^2)
\end{equation}
\begin{equation}
\nabla^2 A_{\rm grav}(x,y) = 2\kappa(x,y)
\end{equation}
Here $\kappa$ is a dimensionless quantity with two distinct physical
interpretations.  On the one hand, it is the sky projected density ---
in special units, to make it dimensionless.  On the other hand,
$\kappa$ is the {\em convergence\/} of a bundle of rays.

### 2013-08-30 systems/many2.py
plaing around with contour plots

@@ -3,6 +3,9 @@ import numpy as np
 from pylab import contour, figure, savefig, show
 from poten import poten_SIE, poten_NFW, poten_shear
 from angdiam import dratio
+import matplotlib as mpl
+
+from scipy import ndimage
 
 from matplotlib import rc
 rc('text', usetex=True)
@@ -128,21 +131,86 @@ def draw(asw):
     savefig(folder+asw+flag+'_menc.png')
     
     
+    #arriv = upsample(x,y,arriv, upsample=10)
     x1, y1 = gradient(arriv)
-    x1, y1 = np.abs(x1), np.abs(y1)
+    #print 'x1',x1
+    #x1, y1 = np.abs(x1), np.abs(y1)
+    sig = 1
+    x1 = ndimage.gaussian_filter(x1, sigma=sig, mode='wrap')
+    y1 = ndimage.gaussian_filter(y1, sigma=sig, mode='wrap')
+    
+        
     
     z1 = x1*x1+y1*y1
-    zmin = amin(z1) * 42. # because it's 42
-    print 'zmin',zmin
-    mask = z1<zmin
-  
+    #z1 = ndimage.gaussian_filter(z1, sigma=3)
+
+    z1 = ndimage.gaussian_gradient_magnitude(arriv, sigma=sig, mode='nearest')
+    #z2 = ndimage.laplace(arriv)    
+    
+    #zmin = amin(z1) * 42. # because it's 42
+    #print 'zmin',zmin
+    #mask = z1<zmin
+    mask, ids = detect_local_minima(np.log(z1))
+    
+    typestr = {
+      -2: 'bg',
+      -1: 'udef',
+      0: 'sad',
+      1: 'min',
+      2: 'max',
+      3: 'canc'}
+      
+    types, s1, s2 = def_type(ids, x1,y1)
+    #print types
+    
+    ids, types, s1, s2 = filter_types(ids, types, s1, s2, thres=10)
+    #print types
+    
+    pnttype = np.ones(mask.shape) * -2
+    xids, yids = ids
+    #print ids
+    for i, xx in enumerate(xids):
+      yy=yids[i]
+      t=types[i]
+      print i, xx, yy, t, typestr[t], s1[i], s2[i]
+      
+      pnttype[xx,yy]=t
+    
+    #return mask
+    if False:
+      fig = figure()
+      panel = fig.add_subplot(2,2,1)
[…]

### 2013-08-31 text/tex/intro.tex
finished setup
plaing around with contour plots
layout of intro, changes to scirpts, made class copatible to setlength

@@
 
-Spacewarps is the future.
+Spacewarps is the future.\needcite
 
@@
 
+\needfig[testfig]
+
 Results in \ref{sec:results}.
@@
 Things to do in \ref{sec:todo}.
+
+This is a general test \secref{todo}\todo{remove this part again}
+
+\hr
+
+Gravitational Lensing, prediced bz ART and first confirmed by ... can be used as a tool by astronomers to examine structres far away.
+Allows to esimate masses, not only visual but total masses (inkl dark matter).
+Further more, allows to esimate cosmologial parameters, like Hubblekonst (Age of universe \needcite) and massdenity \needcite.
+To accheve this, one needs to find / identify lenses and in a second step model those lenses to get involved parameters.
+
+How to find gravitational lenses? Huge amount of image data from surveys that needs to be processed.
+Even more comming in the future with new surveys \needcite
+Robotic procession has been suggested \needcite, and tested \needcite, but has failed so far to be convincing \needcite.
+Another suggestion involves humans, volountiers \footnote{A volounteer is considered everybody that has no background in astro physics.}.
+\sw uses this approach with great succes so far.\needcite
+
+Next step involves modelling. that needs advanced knolegde and takes a lot of time.
+Too much time to be done by astronomers them selves as the results from new surveys and identifiers like \sw come in.
+So there is a demand of a means to modell a great amount of identified lens data, that scales with increasing data amount.
+
+The purpose of this study was to provide a means to model a large amount of gravitational lenses by showing that gravitational lens modelling can be learned / done by volunteers.
+We suggest, that volunteers will be as successful as professionals with modelling if provided with an easy to use tool with visual feedback (WYSIWYG) and a minimal set of instructions.
+Voulnteers will then croud work (using buzz word here ;) ) / work collarobartive on modelling lenses from several sources / groups at a central place.
+Since this is a iterative learning process, the more involved volunteers will quickly gain knoledge that can be passed down to new volunteers.
+That creates a social structre that scales well with the number of volunteers, as other projects\needcite have already shown.
+Finding people working as volunteers has been shown to be successful lst but not least by \sw and the whole galaxy zoo project.
+To test the peoples abilities, we invesigated the performance of a first set of volunteers modelling a set of simulated lenses.
+We tested the ability to correcly identify lensed images and reproduce simmilar mass maps of the lens.

### 2013-09-01 text/tex/spaghetti.tex
a few words on sl, added bib entries

@@
-\section{SpaghettiLens} \label{sec:SpaghettiLens}
+\section{\spl} \label{sec:SpaghettiLens}
 
@@
 generated.
+
+\hr
+
+SpaghettiLens is build on top of GLASS \citep{Lubini2012}, that builds and improves upon PixeLens \citep{Saha2004}.
+It provides a web based graphical user interface that allows to trace contours with the mouse and identify extremal points by clicking on the image.
+\spl then tries to find a mass distribution $\kappa(x,y)$ that reproduces the input parameters.
+It provides direct visual feedback by rendering a synthetic image side by side the original image, such that users can directly compare the predictions of their model to the survey image.
+It additinally provides plots of the modelled mass distribution and the contour lines of arrival time surface as feedback.
+
+Using this direct visual feedback, even unexperienced users can successfully model lenses using a iterative approach if they know what to look for.
+
+The modelling process consitis of three basic steps:
+
+\begin{enumerate}
+  \item identify lensed images and separate them from other background light sources
+  \item classify and order images accoring to arrival time. (local minima, maxima, saddlepoints)
+  \item fine tune the arrangement, identify addidional external point masses influencing the result
+\end{enumerate}
+
+\spl assists volunteers with several features in this process.
+Step 1 by supplying several images from several bands (not yet implemented in \sw).
+Step 2 by restricting the user, only valid configuration can be entered, the odd number theorem\needcite is taken care of.
+Step 3 by providing the visual feedback. Users can check the generated mass distribution and synthetic image.
+
+Explanation:
+Synthetic image: A rendering of the derivative of the arrival time for each pixel. This leads to a black and white image that reasonably looks like the visual appearace of the model.
+
+Volunteers enter an assumed lens configuration and get a rendered synthetic image along with a reconstructed / simulated mass map and reconstructed / simulated contour lines.
+This allows volunteers to directly see the effect of changing the parameters of the model and to iteratively approach a good solution.
+
+In the next step, several volunteers can work together / exchange their ideas and models and try to improve previous modelling attempts by others.
+This leads to a set of models organised in a branching, tree like structre for each model.
+Volunteers then organise them selves to narrow down the different branches and come up with a few consensus models in the end.
+
+\spl provides plug ins to support any data source. At the moment, a data link to \sw and to the MasterLens database\needcite are implemented.
 

### 2013-09-02 plots/gen_plots.py
figures scripts

@@ -1,0 +1,363 @@
+# -*- coding: utf-8 -*-
+"""
+collects all the data distributed all over this place and generates the
+final plots
+
+Created on Mon Sep 02 15:35:36 2013
+
+@author: RafiK
+"""
+
+
+import sys
+import os
+import shutil
+
+import requests as rq
+
+import numpy as np
+import matplotlib as mpl
+import matplotlib.pylab as pl
+
+from numpy import pi
+
+write_to_tex_folder = False
+
+# realtive to plots dir
+outdir = 'figs'
+
+simdir = 'sim'
+moddir = 'mod'
+
+
+
+resdir = 'res'
+# relative to root of git repro
+texdir = 'text/fig/sims'
+
+
+
+pardir = os.path.abspath(os.path.join(os.path.dirname(__file__), os.path.pardir))
+
+outdir = os.path.abspath(os.path.join(os.path.dirname(__file__), outdir))
+simdir = os.path.abspath(os.path.join(outdir, simdir))
+moddir = os.path.abspath(os.path.join(outdir, moddir))
+
+if write_to_tex_folder:
+  resdir = os.path.abspath(os.path.join(pardir, texdir))
+else:
+  resdir = os.path.abspath(os.path.join(outdir, resdir))
+
+
+sys.path.append(os.path.join(pardir, 'spaghetti'))
+sys.path.append(os.path.join(pardir, 'systems'))
+
+import gen_kappa_encl_plots as spg
+import many
+
+
+def tmp():
+  for asw in many.sim:
+    draw_sim(asw, many.sim)
+    break
+  
+  for idd, elem in enumerate(spg.data):
+    draw_mod(0, elem['id'], 0, spg.data, spg.sims)
+    break
+  
+  for idd, elem in enumerate(spg.data):
+    create_tex(elem['id'], None)
+    break
+    
+  
+  
+def create_tex(mid, asw):
+  try:
+    figpath = os.path.join(resdir, '%06i'%mid)
+    texpath = resdir
+    os.makedirs(figpath)
+  except OSError as e:
+    print 'error creating dir', e
+
+  rawtex = r"""
+\begin{figure}
+  \centering
+    \includegraphics[width=0.45\textwidth]{%s}
+  \caption{%s}
+  \label{fig:%s}
+\end{figure}
[…]

### 2013-09-02 spaghetti/gen_kappa_encl_plots.py

@@ -60,6 +60,7 @@ import os.path
 #rc('font',**{'family':'serif','serif':['Palatino']})
 #rc('text', usetex=True)
 
+datadir = os.path.abspath(os.path.join(os.path.dirname(__file__), datadir))
 
 user_data_dir = os.path.abspath(os.path.join(datadir, inpfilename))
 save_fig_path = os.path.abspath(os.path.join(datadir, fig_dir))
@@ -344,8 +345,3 @@ def getEinsteinR(x, y):
     else:
       return False
     
-
-
-
-
-

### 2013-09-02 spaghetti/gen_plots_for_paper.py

@@ -1,0 +1,87 @@
+# -*- coding: utf-8 -*-
+"""
+Creates Plots for the paper for the sim modelling challenge result ids
+
+those ids are read from the file generated by gen_kappa_encl_data.py
+
+Needs to be run on the server, in the glass environment!
+
+cd /srv/lmt/worker
+./run_glass this_file.py
+
+
+saves the new plots in outdir
+
+
+Created on Mon Sep 02 18:26:32 2013
+
+@author: RafiK
+"""
+
+
+
+import os, shutils
+import matplotlib.pylab as pl
+
+
+idfile = '/srv/lmt/scripts_output/incr_result_resolution/ids.txt'
+
+outdir = '/srv/lmt/scripts_output/gen_plots_for_paper/'
+
+datadir = '/srv/lmt/tmp_media'
+
+
+#which plots to draw
+plts = [True, True, True, False]
+
+
+
+with open(idfile, 'r') as f:
+  ids = f.readlines()
+  ids = [int(_) for _ in ids]
+  print ids
+
+
+for idd in ids:
+  print 'regenerating %i' % idd
+  idstr = '%06i' % idd
+  fstr = os.path.join(datadir, idstr, 'state.txt')
+  g = loadstate(fstr)
+  
+  outpath = os.path.join(outdir, idstr)  
+  
+  try:
+    shutils.makedirs(outpath)
+  except:
+    pass
+  
+  g.make_ensemble_average()
+  
+  if plts[0]:
+    g.arrival_plot(g.ensemble_average, only_contours=True, colors='magenta', clevels=40)
+    #pl.gca().axes.get_xaxis().set_visible(False)
+    #pl.gca().axes.get_yaxis().set_visible(False)
+    pl.savefig(outpath + '/img1.png')
+    pl.close()
+  
+  if plts[1]:
+    g.kappa_plot(g.ensemble_average, 0, with_contours=True, clevels=20, vmax=1, with_colorbar=False)
+    #pl.gca().axes.get_xaxis().set_visible(False)
+    #pl.gca().axes.get_yaxis().set_visible(False)
+    pl.savefig(outpath + '/img2.png')
+    pl.close()
+  
+  if plts[2]:
+    g.srcdiff_plot(g.ensemble_average)
+    #pl.gca().axes.get_xaxis().set_visible(False)
+    #pl.gca().axes.get_yaxis().set_visible(False)
+    pl.savefig(outpath + '/img3.png')
[…]

### 2013-09-02 systems/many.py

@@ -133,7 +133,11 @@ def draw(asw):
     panel.contour(x,y,arriv,lev)
     savefig('figs/'+asw+flag+'_arriv.png')
 
-for asw in sim:
+
+
+if __name__ == '__main__':
+  for asw in sim:
     draw(asw)
+
 
 

### 2013-09-02 spaghetti/gen_plots_for_paper.py
bugif

@@ -20,7 +20,7 @@ Created on Mon Sep 02 18:26:32 2013
 
 
 
-import os, shutils
+import os, shutil
 import matplotlib.pylab as pl
 
 
@@ -51,7 +51,7 @@ for idd in ids:
   outpath = os.path.join(outdir, idstr)  
   
   try:
-    shutils.makedirs(outpath)
+    shutil.makedirs(outpath)
   except:
     pass
   

### 2013-09-02 spaghetti/gen_plots_for_paper.py
bugxif

@@ -24,9 +24,9 @@ import os, shutil
 import matplotlib.pylab as pl
 
 
-idfile = '/srv/lmt/scripts_output/incr_result_resolution/ids.txt'
+idfile = '/srv/lmt/script_output/incr_result_resolution/ids.txt'
 
-outdir = '/srv/lmt/scripts_output/gen_plots_for_paper/'
+outdir = '/srv/lmt/script_output/gen_plots_for_paper/'
 
 datadir = '/srv/lmt/tmp_media'
 

### 2013-09-02 spaghetti/gen_plots_for_paper.py
bugxif

@@ -35,6 +35,10 @@ datadir = '/srv/lmt/tmp_media'
 plts = [True, True, True, False]
 
 
+glass_basis('glass.basis.pixels', solver=None)
+exclude_all_priors()
+
+
 
 with open(idfile, 'r') as f:
   ids = f.readlines()

### 2013-09-02 spaghetti/gen_plots_for_paper.py
bugxif

@@ -20,7 +20,7 @@ Created on Mon Sep 02 18:26:32 2013
 
 
 
-import os, shutil
+import os
 import matplotlib.pylab as pl
 
 
@@ -55,8 +55,9 @@ for idd in ids:
   outpath = os.path.join(outdir, idstr)  
   
   try:
-    shutil.makedirs(outpath)
+    os.makedirs(outpath)
   except:
+    print 'path creat failed'
     pass
   
   g.make_ensemble_average()

### 2013-09-03 plots/gen_plots.py
plot gen script working

@@ -17,11 +17,35 @@ import requests as rq
 
 import numpy as np
 import matplotlib as mpl
+
+mpl.rc('text', usetex=True)
+mpl.rc('font', family='serif')
+
+params = {
+# see http://matplotlib.org/users/customizing.html
+#  'backend': 'ps',
+  'text.latex.preamble': [
+    r"\usepackage{amsmath}",
+    r"\usepackage{textgreek}"
+    ],
+  'mathtext.default': 'regular',
+  'axes.labelsize': 14,
+  'text.fontsize': 14,
+  'legend.fontsize': 10,
+  'xtick.labelsize': 12,
+  'ytick.labelsize': 12,
+  'text.usetex': True,
+#  'figure.figsize': fig_size,
+#  'axes.unicode_minus': True
+  }
+mpl.rcParams.update(params)
+
 import matplotlib.pylab as pl
 
 from numpy import pi
 
-write_to_tex_folder = False
+
+write_to_tex_folder = True
 
 # realtive to plots dir
 outdir = 'figs'
@@ -35,6 +59,8 @@ resdir = 'res'
 # relative to root of git repro
 texdir = 'text/fig/sims'
 
+#image extension
+ext = 'png'
 
 
 pardir = os.path.abspath(os.path.join(os.path.dirname(__file__), os.path.pardir))
@@ -55,56 +81,251 @@ sys.path.append(os.path.join(pardir, 'systems'))
 import gen_kappa_encl_plots as spg
 import many
 
+debug = False
 
-def tmp():
+
+
+def test():
+#  for idd, elem in enumerate(spg.data):
+#    mid = elem['id']
+#    draw_mod(0, mid, 0, spg.data, spg.sims)
+#    break
   for asw in many.sim:
     draw_sim(asw, many.sim)
     break
-  
-  for idd, elem in enumerate(spg.data):
-    draw_mod(0, elem['id'], 0, spg.data, spg.sims)
-    break
-  
-  for idd, elem in enumerate(spg.data):
-    create_tex(elem['id'], None)
-    break
+
+
+
+def run():
+  all_sim_plots()
+  all_mod_plots()
+  all_tex()
+
+#==============================================================================
+#   # generate sim plots
+#==============================================================================
+def all_sim_plots():
+  print '\nSIMS'
+  print '============================================='  
+  for asw in many.sim:
+
[…]

### 2013-09-03 spaghetti/gen_plots_for_paper.py

@@ -19,6 +19,38 @@ Created on Mon Sep 02 18:26:32 2013
 """
 
 
+#==============================================================================
+# MATPLOTLIB SETTINGS
+#==============================================================================
+
+# make sure they are in sync with the local plots command
+import matplotlib as mpl
+
+mpl.rc('text', usetex=True)
+mpl.rc('font', family='serif')
+
+params = {
+# see http://matplotlib.org/users/customizing.html
+#  'backend': 'ps',
+  'text.latex.preamble': [
+    r"\usepackage{amsmath}",
+    r"\usepackage{textgreek}"
+    ],
+  'mathtext.default': 'regular',
+  'axes.labelsize': 14,
+  'text.fontsize': 14,
+  'legend.fontsize': 10,
+  'xtick.labelsize': 12,
+  'ytick.labelsize': 12,
+  'text.usetex': True,
+#  'figure.figsize': fig_size,
+#  'axes.unicode_minus': True
+  }
+mpl.rcParams.update(params)
+#==============================================================================
+# END MATPLOTLIB SETTINGS
+#==============================================================================
+
 
 import os
 import matplotlib.pylab as pl

### 2013-09-03 plots/gen_plots.py
plot gen script dok

@@ -1,7 +1,49 @@
 # -*- coding: utf-8 -*-
 """
 collects all the data distributed all over this place and generates the
-final plots
+final plots.
+
+make sure to have the following installed:
+- python module 'requests'
+- latex modules 'amsmath', 'textgreek'?
+
+should run in this folder /plots, doens't do much by default, you need
+to call some command after importing this module that anything happens.
+Needs the whole repro to be checked out. (folders /spaghetti and /systems)
+
+either:
+  use python in interactive mode:
+  python -i -c "import gen_plots"
+
+or (better):
+  use ipython:
+  ipython
+  %run gen_plots.py
+  
+available commands are:
+
+- run():
+  does all the work below
+  
+- all_sim_plot()
+  creates all the simulation data plots ('real' data)
+  
+- all_mod_plots()
+  creates all the results / modells plots
+  (server side created plots get grapped from internet)
+  
+- all_tex()
+  assumes the plots are already created (under /plots)
+  creates all the tex files and copies the relevant files
+  to the tex folder
+
+
+data is stored in:
+- many.sim
+- spg.data
+
+
+
 
 Created on Mon Sep 02 15:35:36 2013
 
@@ -44,8 +86,11 @@ import matplotlib.pylab as pl
 
 from numpy import pi
 
-
+# set to false to do a dry run in /plots/res
 write_to_tex_folder = True
+#image extension (png or pdf)
+ext = 'png'
+
 
 # realtive to plots dir
 outdir = 'figs'
@@ -53,14 +98,10 @@ outdir = 'figs'
 simdir = 'sim'
 moddir = 'mod'
 
-
-
 resdir = 'res'
 # relative to root of git repro
 texdir = 'text/fig/sims'
 
-#image extension
-ext = 'png'
 
 
 pardir = os.path.abspath(os.path.join(os.path.dirname(__file__), os.path.pardir))
@@ -205,6 +246,12 @@ def create_tex(mid, asw):
       sf_opt
     ),  
     tFig(
+      r'fig/sims/%s/kappa.%s'%(asw, ext),
+      r'[real mass distribution]',
+      r'%04i_sim_mass' % mid,
+      sf_opt
+    ),
+    tFig(
       r'fig/sims/%06i/spaghetti.%s'%(mid, ext),
       r'[modelled arrivaltime contour lines]',
       r'%04i_cont'%mid,
[…]

### 2013-09-04 plots/gen_plots.py
server side plots

@@ -154,6 +154,7 @@ def all_sim_plots():
     if not os.path.isdir(path):
       os.makedirs(path)    
     
+    get_sim_adv_data(asw)
     draw_sim(asw, many.sim)
     if debug: break
   
@@ -172,7 +173,7 @@ def all_mod_plots():
       os.makedirs(path)    
 
     get_mod_adv_data(mid) #download plots generated on server with glass
-    draw_mod(0, mid, 0, spg.data, spg.sims)
+    draw_mod(mid, elem, spg.data, spg.sims)
 
     if debug: break
   
@@ -411,9 +412,51 @@ def get_mod_adv_data(mid):
           f.write(chunk)
     else:
       print '!! file', baseurl, img, 'not found'
-          
+  
+  # get input image
+  r = rq.get('http://mite.physik.uzh.ch/result/%06i/input.png'%mid, stream=True)
+  if r.status_code == 200:
+    with open(os.path.join(path, 'input.png'), 'wb') as f:
+      for chunk in r.iter_content(1024):
+        f.write(chunk)
+  else:
+    print '!! file input.png not found'
+  
   print '...DONE.'
   for f in saveas: print '  - %s'%f
+  print '  - input.png'
+
+
+def get_sim_adv_data(asw):
+  """gets some online data"""
+  print '> getting sims data online', 
+
+  path = os.path.join(simdir, asw)
+
+  #get sw image
+  data = {
+    'action': 'datasourceApi',
+    'src_id':3,
+    'do':'fetch',
+    'swid':'ASW0001mze'
+    }
+  r=rq.post('http://mite.physik.uzh.ch/api', data)
+  if r.status_code == 200:
+    iurl = r.json()['list'][0]['url']
+    
+    r = rq.get(iurl, stream=True)
+    if r.status_code == 200:
+      with open(os.path.join(path, 'sw.png'), 'wb') as f:
+        for chunk in r.iter_content(1024):
+          f.write(chunk)
+    else:
+      print '!! file on spacewarps not found'        
+    
+  else:
+    pass
+
+  print '...DONE.'
+  print '  - sw.png'
 
 
 def draw_sim(asw, sim):
@@ -499,7 +542,7 @@ def draw_sim(asw, sim):
     for f in filenames: print '  - %s'%(f+'.'+ext)
 
 
-def draw_mod(idd, mid, count, data, sims):
[…]

### 2013-09-04 spaghetti/gen_plots_for_paper.py

@@ -96,6 +96,7 @@ for idd in ids:
   
   if plts[0]:
     g.arrival_plot(g.ensemble_average, only_contours=True, colors='magenta', clevels=40)
+    g.overlay_input_points(g.ensemble_average)
     #pl.gca().axes.get_xaxis().set_visible(False)
     #pl.gca().axes.get_yaxis().set_visible(False)
     pl.savefig(outpath + '/img1.png')
@@ -110,13 +111,15 @@ for idd in ids:
   
   if plts[2]:
     g.srcdiff_plot(g.ensemble_average)
+    g.overlay_input_points(g.ensemble_average)
     #pl.gca().axes.get_xaxis().set_visible(False)
     #pl.gca().axes.get_yaxis().set_visible(False)
     pl.savefig(outpath + '/img3.png')
     pl.close()
   
   if plts[3]:
-    g.srcdiff_plot_adv(g.ensemble_average, night=True, upsample=8)
+    g.srcdiff_plot_adv(g.ensemble_average, night=False, upsample=8)
+    g.overlay_input_points(g.ensemble_average)
     pl.savefig(outpath + '/img4.png', facecolor='black', edgecolor='none')
     pl.close()
     

### 2013-09-05 plots/gen_plots.py
edits

@@ -236,34 +236,34 @@ def create_tex(mid, asw):
     ),
     tFig(
       r'fig/sims/%s/arriv.%s'%(asw, ext),
-      r'[real arrival time surface]',
+      r'[real arrival-time surface]',
       r'%04i_sim_arr' % mid,
       sf_opt
     ),  
     tFig(
       r'fig/sims/%06i/mass.%s'%(mid, ext),
-      r'[modelled mass distribution]',
+      r'[model mass distribution]',
       r'%04i_mass'%mid,
       sf_opt
     ),  
     tFig(
       r'fig/sims/%06i/spaghetti.%s'%(mid, ext),
-      r'[modelled arrivaltime contour lines]',
+      r'[model arrival-time surface]',
       r'%04i_cont'%mid,
       sf_opt
     ),  
     tFig(
       r'fig/sims/%06i/kappa_encl.%s'%(mid, ext),
-      r'[modelled enclosed mass]',
+      r'[real vs model enclosed mass]',
       r'%04i_kappa'%mid,
       sf_opt
     ),  
     tFig(
       r'fig/sims/%06i/arr_time.%s'%(mid, ext),
-      r'[modelled arrival time surface]',
+      r'[model lensed image]',
       r'%04i_atime'%mid,
       sf_opt
-    ),  
+    ),
   ]
 
   #figpath = os.path.join(resdir, '%06i'%mid)

### 2013-09-06 text/tex/theory.tex
editing
editing Fermat-principle section

@@
+\begin{figure}
+\includegraphics[height=.3\vsize]{fig/arriv1.png}
+\includegraphics[height=.3\vsize]{fig/arriv2.png}
+\includegraphics[height=.3\vsize]{fig/arriv3.png}
+\caption{Arrival-time surfaces, with contour levels of equal arrival
+  time.  The upper surface is with no lens.  The middle surface is
+  when a circular lensing mass (offset from the source) is added; a
+  maximum, a minimum and a saddle point can be seen.  The last surface
+  is the result of an elongated lensing mass; a maximum, two minima
+  and two saddle points can be seen. \label{fig:arriv}}
+\end{figure}
+
 \section{Fermat's Principle} \label{sec:Fermat}
 
-Explain Fermat's principle
+There are several ways to understand the formation of arcs and
+multiple images in gravitational lensing.  We will follow some ideas
+originally introduced by \cite{1986ApJ...310..568B}, based on Fermat's
+principle.  They key to this approach is an abstract construct called
+the {\em arrival-time surface.}  This surface cannot itself be
+observed, but several observable quantities can be derived from it.
+
+Consider a gravitational lens.  As in most astrophysical lensing, this
+lens is `thin' along the line of sight, and effectively lies on a
+plane.  Let $x,y$ be coordinates on this plane.  These coordinates
+measure ordinary physical lengths (metres, etc) on the lens plane.
+Let there be a point source of light at $x=0,y=0$ but far behind the
+lens plane.  Now imagine a light ray coming from the source to the
+lens plane $(x,y)$ and then changing its path to reach the observer.
+The observer would see the source apparently behind $(x,y)$ on the
+lens.  If this happened, the light ray would have taken a longer path
+than having come directly to the observer.  The geometrical path
+difference would be $\propto x^2 + y^2$ for small angles.  Let us
+define
+\begin{equation} \label{eq:Ageom}
+A_{\rm geom}(x,y) = \half(x^2 + y^2) \,.
+\end{equation}
+As written, this looks is an area.  But we can also think of it as a
+time delay introduced by the light ray bending --- with an unwritten
+constant factor.  That factor is basically the lens distance times the
+speed of light; the precise expression has to take the expansion of
+the Universe into account, and is given in the Appendix.  The top part
+of Figure~\ref{fig:arriv} shows what $A_{\rm geom}$ looks like: it is
+just a paraboloid.  The minimum of the surface is at the source, and
+that is where, according to Fermat's principle, the image will be.
+The rest of the arrival-time surface is just an imaginary surface.
+
+The geometrical time delay \eqref{eq:Ageom} is not the only one.  The
+warping of space by a gravitational field introduces a further time
+delay $A_{\rm grav}$.  The arrival time of the light ray, compared to
+what it would have been with no lensing, is
 \begin{equation}
-\hbox{arrival time} = A_{\rm geom} - A_{\rm grav}
+\hbox{arrival time} = A_{\rm geom} + A_{\rm grav} \,.
 \end{equation}
-By convention, $A_{\rm grav}$ is defined in such a way that it is
-negative, hence the minus sign.
-
-Although $A_{\rm geom}$ and $A_{\rm grav}$ are times, we are going to
-write them as if they are areas.  In other words, we will suppress a
-constant factor in the $A$.  That factor is basically the lens
-distance times the speed of light, but the precise expression is given
-in the Appendix.
-
-Introduce the Laplacian $\nabla^2 f(x,y)$ is
+We can also express $A_{\rm grav}$ as an area, with the same implicit
+constant factor that makes it a time.  The expression, however, is
+more complicated.  We first introduce the notation $\nabla^2 f(x,y)$
+to denote
 \begin{equation}
@@
 \end{equation}
-in the limit of $\Delta x,\Delta y$ small.
+in the limit of $\Delta x,\Delta y$ small.  Then we write the
+projected mass density of the lens, which would be in $\rm
+kg\,m^{-2}$, in special units (given in the Appendix) such that the
+surface density becomes a dimensional field $\kappa(x,y)$.  Then
+\begin{equation} \label{eq:poisson}
+\nabla^2 A_{\rm grav}(x,y) = -2\kappa(x,y) \,.
+\end{equation}
+Note that $A_{\rm grav}(x,y)$ is not given explicitly in terms of
+$\kappa(x,y)$, but as a differential equation that must be solved.
+Equations of the type \eqref{eq:poisson} are, however, well known (as
+Poisson equations in two dimensions), and there are techniques for
+solving them.
 
-Let $x,y$ be coordinates on the lens, transverse to the line of sight.
-Let there be a point source behind $x=0,y=0$.  Then
-\begin{equation}
-A_{\rm geom}(x,y) = \half(x^2 + y^2)
-\end{equation}
-\begin{equation}
-\nabla^2 A_{\rm grav}(x,y) = 2\kappa(x,y)
-\end{equation}
+What is the effect of mass on the arrival-time surface?  If the mass
+distribution $\kappa(x,y)$ is circular, $A_{\rm grav}(x,y)$ will have
+a peak at the centre of that circle.  But if the mass is not precisely
+in front of the mass, the arrival-time surface will not be circular
+any more.  The result is illustrated in the middle part of
+Figure~\ref{fig:arriv}.  In addition to the maximum, there is now a
+minimum and a saddle point.\footnote{A saddle point is recognizable in
+  a contour map because contours do not loop around it, they tend to
+  approach and then pull away.  There is one contour ---the
+  saddle-point contour--- that makes an X on the saddle point itself.
+  A saddle point contour may not appear on a plot (and
+  Figure~\ref{fig:arriv} does not show one.  But we can tell from the
+  other contours where the saddle-point contour would be.}
+
 Here $\kappa$ is a dimensionless quantity with two distinct physical

### 2013-09-06 text/tex/spaghetti.tex
shaghetti add
edits

@@
+\begin{figure}
+\includegraphics[height=.3\vsize]{fig/spag6941.png} \\
+\includegraphics[height=.3\vsize]{fig/spag7022.png} \\
+\includegraphics[height=.3\vsize]{fig/spag6919.png}
+\caption{Examples of Spaghetti input.  The models from these appear
+  later in Figures \ref{fig:6941}, \ref{fig:7022} and
+  \ref{fig:6919}. \label{fig:input-spag}}
+\end{figure}
+
 \section{\spl} \label{sec:SpaghettiLens}
+
+A saddle point is recognizable in a contour map because contours do
+not loop around it, they tend to approach and then pull away.  There
+is one contour ---the saddle-point contour--- that makes an X on the
+saddle point itself.  A saddle point contour may not appear on a plot
+(and Figure~\ref{fig:arriv} does not show one.  But we can tell from
+the other contours where the saddle-point contour would be.
 
@@
 \spl then tries to find a mass distribution $\kappa(x,y)$ that reproduces the input parameters.
+Since there is no unique solution, \spl samples the solution space and produces an ensemble of solutions, as described in the paper \citep{Lubini2012}.
 It provides direct visual feedback by rendering a synthetic image side by side the original image, such that users can directly compare the predictions of their model to the survey image.

### 2013-09-06 text/tex/spaghetti.tex
some changes

--- 2013-09-01 text/tex/spaghetti.tex
@@
-\section{\spl} \label{sec:SpaghettiLens}
+\section {Setup} \label{sec:setup}
+
+\subsection{\sw} \label{sec:SpaceWarps}
+some words about \sw
+
+\subsection{\spl} \label{sec:SpaghettiLens}
 
@@
 
-\section{The simulated lenses} \label{sec:sims}
+\subsection{The simulated lenses} \label{sec:sims}
 
-Three kinds of sims.
+In order to estimate users performance, \sw included some simulated lensing configurations in their image set.
+
+There are three kinds of simulated lenses:
 
@@
 Lenses follow \cite{2001astro.ph..2341K,2001astro.ph..2340K}.
-Parameters in online supplement.
+Parameters can be looked up in online supplement.
+
+\subsection{The Test setup} \label{sec:testsetup}
+We made use of this simulated lenses and further tested the volunteers abilities to model those simulated lensing images.
+
+The testing was done with a small number of volunteers ($N_v=4$) and a set of simulated configurations (N_{sim}=30) chosen to represent some typical configurations.
+\todo{witch??}
+
+Two tests were done on the models created by the volunteers:
+
+\begin{enumerate}
+  \item Correct identification of extremal points
+  \item Reproduction of the mass distribution of the lens
+\end{enumerate}
+
+The first test required the volunteers to accomplish 2 tasks: First to identify and distinguish lensed images from background light sources.
+Second to come up with a logical ordering of the arrival time of those lensed images.
+
+The second test compared the enclosed mass $\kappa(r)$ of the models and the simulations.

### 2013-09-06 spaghetti/gen_plots_for_paper.py

@@ -34,7 +34,7 @@ params = {
 #  'backend': 'ps',
   'text.latex.preamble': [
     r"\usepackage{amsmath}",
-    r"\usepackage{textgreek}"
+    #r"\usepackage{textgreek}"
     ],
   'mathtext.default': 'regular',
   'axes.labelsize': 14,
@@ -111,7 +111,7 @@ for idd in ids:
   
   if plts[2]:
     g.srcdiff_plot(g.ensemble_average)
-    g.overlay_input_points(g.ensemble_average)
+    #g.overlay_input_points(g.ensemble_average)
     #pl.gca().axes.get_xaxis().set_visible(False)
     #pl.gca().axes.get_yaxis().set_visible(False)
     pl.savefig(outpath + '/img3.png')
@@ -119,7 +119,7 @@ for idd in ids:
   
   if plts[3]:
     g.srcdiff_plot_adv(g.ensemble_average, night=False, upsample=8)
-    g.overlay_input_points(g.ensemble_average)
+    #g.overlay_input_points(g.ensemble_average)
     pl.savefig(outpath + '/img4.png', facecolor='black', edgecolor='none')
     pl.close()
     

### 2013-09-06 text/tex/results.tex
choosing sample figures

@@
 
-Example Figure \ref{fig:ASW0000ar2}.
+Figure \ref{fig:6941} shows a simple example.
+
+Figure \ref{fig:6975} shows an example where substructure introduces a
+complication, but model ok.
+
+Figure \ref{fig:6937} shows a case where substructure leads to a
+poorer model.
+
+Figure \ref{fig:6975} shows a nice symmetric quad.
+
+Figure \ref{fig:6990} shows a long-axis quad.
+
+Figure \ref{fig:6919} shows a short-axis quad.
+
+Figure \ref{fig:6915} shows an inclined quad.
+
+Figure \ref{fig:6971} shows incorrect identification, but the enclosed
+mass still quite well recovered.

### 2013-09-06 text/tex/spaghetti.tex

@@
 
+\begin{figure}
+\includegraphics[height=.3\vsize]{fig/spag6941.png} \\
+\includegraphics[height=.3\vsize]{fig/spag7022.png} \\
+\includegraphics[height=.3\vsize]{fig/spag6919.png}
+\caption{Examples of Spaghetti input.  The models from these appear
+  later in Figures \ref{fig:6941}, \ref{fig:7022} and
+  \ref{fig:6919}. \label{fig:input-spag}}
+\end{figure}
+
 \subsection{\spl} \label{sec:SpaghettiLens}
+
+A saddle point is recognizable in a contour map because contours do
+not loop around it, they tend to approach and then pull away.  There
+is one contour ---the saddle-point contour--- that makes an X on the
+saddle point itself.  A saddle point contour may not appear on a plot
+(and Figure~\ref{fig:arriv} does not show one.  But we can tell from
+the other contours where the saddle-point contour would be.
 
@@
 Lenses follow \cite{2001astro.ph..2341K,2001astro.ph..2340K}.
+
 Parameters can be looked up in online supplement.

### 2013-09-06 text/tex/theory.tex

@@
 Poisson equations in two dimensions), and there are techniques for
-solving them.
+solving them.  Incidentally, $\kappa$ has a second meaning, as well as
+being a dimensionless form of the the projected density: it also
+measures how a bundle of light rays is brought together by
+gravitational lensing, and is called the {\em convergence}.
 
@@
 Figure~\ref{fig:arriv}.  In addition to the maximum, there is now a
-minimum and a saddle point.\footnote{A saddle point is recognizable in
-  a contour map because contours do not loop around it, they tend to
-  approach and then pull away.  There is one contour ---the
-  saddle-point contour--- that makes an X on the saddle point itself.
-  A saddle point contour may not appear on a plot (and
-  Figure~\ref{fig:arriv} does not show one.  But we can tell from the
-  other contours where the saddle-point contour would be.}
+minimum and a saddle point.  According to Fermat's principle, images
+appear at maxima and saddle points as well as minima.  If the mass
+distribution is non-circular, more images can appear.  The lower
+example in Figure~\ref{fig:arriv} shows the arrival-time surface due
+to an elongated mass.  There is still a maximum, part of an elongated
+hill, and around it there are two minima and two saddle points.
 
-Here $\kappa$ is a dimensionless quantity with two distinct physical
-interpretations.  On the one hand, it is the sky projected density ---
-in special units, to make it dimensionless.  On the other hand,
-$\kappa$ is the {\em convergence\/} of a bundle of rays.
+So far we have considered a single point source.  What if there is an
+extended source?  To understand this, let us consider what happens if
+we move the original point source slightly, or equivalently, keep the
+original point source behind $x=0,y=0$ and move the lens slightly.
+The contours of constant arrival time will, naturally, move slightly,
+and so will the images.  The movement of the contours will be most
+noticeable where the contours are far apart, that is, where the
+arrival-time surface is nearly flat.  As is evident from
+Figure~\ref{fig:arriv}, this is the region where the minima and saddle
+points lie, or near the images.  In this region, points on the source
+that are close together produce images that are comparatively far
+apart.  In other words, the image is highly magnified.  In summary,
+lower curvature in the arrival-time surface for a point source implied
+larger magnification of an extended source.  Conversely, where the
+arrival-time surface is strongly curved, the image will be
+demagnified.  We see from Figure~\ref{fig:arriv} that the arrival-time
+surface tends to be highly curved near the maximum.  Hence maxima tend
+to be demagnified.  In practice, maxima of the arrival time are nearly
+always too faint to see.  The minima and saddle points dominate.

### 2013-09-07 text/tex/spaghetti.tex
SHELL FILE to create the plots. not really tested

@@
 
-The testing was done with a small number of volunteers ($N_v=4$) and a set of simulated configurations (N_{sim}=30) chosen to represent some typical configurations.
-\todo{witch??}
+The testing was done with a small number of volunteers ($N_v=4$) and a set of simulated configurations ($N_{sim}=30$) chosen to represent some typical configurations.
 

### 2013-09-07 plots/gen_plots.py

@@ -520,6 +520,10 @@ def draw_sim(asw, sim):
     for k in range(len(rad)):
         fil.write('%9.2e %9.2e\n' % (rad[k],sum[k]))
     fil.close()
+    fil = open(os.path.join(simdir, asw+'.txt'),'w')
+    for k in range(len(rad)):
+        fil.write('%9.2e %9.2e\n' % (rad[k],sum[k]))
+    fil.close()
     panel.scatter(rad,sum)
     panel.set_xlabel('radius [pixels]')
     panel.set_ylabel('average interior \textkappa [1]')
@@ -575,7 +579,7 @@ def draw_mod(mid, elem, data, sims):
   try:
     sims[name]
   except KeyError:
-    print '!! missing sims data for', mid, name
+    print '\n!! missing sims data for', mid, name
     return
 
   #yerr=[elem['err_p'] - elem['y'], elem['y']- elem['err_m']]

### 2013-09-07 spaghetti/gen_kappa_encl_plots.py

@@ -35,6 +35,7 @@ datadir = './data/mod_chal/'
 
 #input files
 sims_dir = '../systems/figs' #rel to ./spaghetti
+sims_alt_dir = '../plots/figs/sims'
 inpfilename = 'kappa_encl.csv'
 
 #output
@@ -119,6 +120,10 @@ import glob
 import os
 
 paths = glob.glob(os.path.join(sims_dir, '*.txt' ))
+
+# fix for running gen_plots
+if len(paths)==0:
+  paths = glob.glob(os.path.join(sims_alt_dir, '*.txt' ))
 
 sims = {}
 

### 2013-09-08 plots/gen_plots.py
gen_plots: gets right org img, gens csv overview

@@ -205,11 +205,15 @@ def all_tex():
     #only create tex file if all files available and copied    
     if t1 and t2:
       create_tex(mid, asw)
+      create_alt_tex(mid, asw)
       mid_list.append(str(mid))
     if debug: break
   
   with open(os.path.join(resdir, '_all.tex'), 'w') as ff:
     ff.write('\n'.join(['\input{fig/sims/%s}\n\clearpage'%_ for _ in mid_list]))
+
+  with open(os.path.join(resdir, '_all_alt.tex'), 'w') as ff:
+    ff.write('\n'.join(['\input{fig/sims/%s_alt}\n\clearpage'%_ for _ in mid_list]))
 
 class tFig(dict):
   def __init__(self, sf_path, sf_capt, sf_label, sf_opt):
@@ -317,6 +321,112 @@ def create_tex(mid, asw):
   print '...DONE.'
   print '  - %04i.tex'%mid
 
+
+
+def create_alt_tex(mid, asw):
+  '''creates tex files for alternative arrangement (to compare point placement)'''
+  print '> generating alt tex file',mid,
+  
+  sf_opt = r'[height=0.4\vsize]'
+  
+  txFigs = [
+    tFig(
+      r'fig/sims/%s/arriv.%s'%(asw, ext),
+      r'[real arrival-time surface]',
+      r'%04i_sim_arr' % mid,
+      sf_opt
+    ),  
+    tFig(
+      r'fig/sims/%06i/spaghetti.%s'%(mid, ext),
+      r'[model arrival-time surface]',
+      r'%04i_cont'%mid,
+      sf_opt
+    ),  
+    tFig(
+      r'fig/sims/%s/sw.%s'%(asw, ext),
+      r'[sw image]',
+      r'%04i_sw'%mid,
+      sf_opt
+    ),  
+    tFig(
+      r'fig/sims/%06i/input.%s'%(mid, ext),
+      r'[input image]',
+      r'%04i_input'%mid,
+      sf_opt
+    ),
+  ]
+
+  #figpath = os.path.join(resdir, '%06i'%mid)
+  texpath = resdir
+
+#  try:
+#    figpath = os.path.join(resdir, '%06i'%mid)
+#    texpath = resdir
+#    #os.makedirs(figpath)
+#  except OSError as e:
+#    print 'error creating dir', e
+
+  tex_env = r"""
+\begin{figure}
+  \centering
+%(subfigtex)s
+  \caption%(env_capt_short)s{%(env_capt_long)s}
[…]

### 2013-09-08 plots/gen_plots.py
added results for test1, inkl autogenerated tables

@@ -324,7 +324,9 @@ def create_tex(mid, asw):
 
 
 def create_alt_tex(mid, asw):
-  '''creates tex files for alternative arrangement (to compare point placement)'''
+  '''creates tex files for alternative arrangement (to compare point placement)
+  you have to manually copy the output of many2 into the sims folder..  
+  '''
   print '> generating alt tex file',mid,
   
   sf_opt = r'[height=0.4\vsize]'
@@ -343,7 +345,7 @@ def create_alt_tex(mid, asw):
       sf_opt
     ),  
     tFig(
-      r'fig/sims/%s/sw.%s'%(asw, ext),
+      r'fig/sims/%s/extr_points.%s'%(asw, ext),
       r'[sw image]',
       r'%04i_sw'%mid,
       sf_opt

### 2013-09-08 plots/gen_table.py

@@ -1,0 +1,242 @@
+# -*- coding: utf-8 -*-
+"""
+generates the results table.
+
+dataorigin is the overview.csv, which is generated from
+overview.ods by saving as csv.
+
+Created on Sun Sep 08 17:05:05 2013
+
+@author: RafiK
+"""
+
+import os
+import csv
+from fractions import Fraction as Frac
+
+
+inp_file_name = 'overview.csv'
+inp_file_path = '../spaghetti/data/mod_chal'
+
+out_file_name = 'results.tex'
+out_file_path = '../text/tab/auto'
+
+inpfile = os.path.join(os.path.abspath(inp_file_path),inp_file_name)
+outfile = os.path.join(os.path.abspath(out_file_path),out_file_name)
+
+
+all_data = {}
+all_stats = []
+
+with open(inpfile, 'r') as f:
+  csvr = csv.reader(f)
+  
+  simname = ''
+
+  for i, row in enumerate(csvr):
+    #print row
+
+    if i in [156,159,162,165]:
+      
+      all_stats.append({
+        'n'          : int(row[ 5]),
+        'npnt'       : int(row[ 7]),
+        'approxPlace': int(row[ 8]),
+        'approx'     : int(row[ 9]),
+        'veryRight'  : int(row[10]),
+        'rightPlace' : int(row[12]),
+        'rightType'  : int(row[13]),
+        'rightOrder' : int(row[14]),
+        'err'        : [int(_) for _ in row[16:26]]
+      })
+  
+    if i<2 or i>155 or (row[0]=='' and row[1]==''):
+      continue
+    
+    if row[0]!='':
+      #print row[0]
+      simname=row[0]
+      all_data[simname]=[]
+      
+    else:
+      if simname=='ASW0001gae':
+        continue
+      #print row[1], row[29], row[30], row[31]
+
+      r12 = reduce(lambda x, y: x+y, [Frac(_) for _ in row[12].split(' ') if _!='']).limit_denominator()
+      r13 = reduce(lambda x, y: x+y, [Frac(_) for _ in row[13].split(' ') if _!='']).limit_denominator()
+
+      #true / false values / strings
+      t = "X"
+      f = 'O'
+
+      data = {
+        'id'         : row[1],
+        'npnt'       : t if row[ 7]=='x' else f,
[…]

### 2013-09-08 plots/gen_table.py
update tables, comment gen_tables

@@ -15,11 +15,21 @@ import csv
 from fractions import Fraction as Frac
 
 
+#==============================================================================
+# SETTINGS
+#==============================================================================
+
 inp_file_name = 'overview.csv'
 inp_file_path = '../spaghetti/data/mod_chal'
 
 out_file_name = 'results.tex'
 out_file_path = '../text/tab/auto'
+
+# for tex template settings see below
+
+#==============================================================================
+# PARSING
+#==============================================================================
 
 inpfile = os.path.join(os.path.abspath(inp_file_path),inp_file_name)
 outfile = os.path.join(os.path.abspath(out_file_path),out_file_name)
@@ -36,6 +46,7 @@ with open(inpfile, 'r') as f:
   for i, row in enumerate(csvr):
     #print row
 
+    # get the stats from the end of the file
     if i in [156,159,162,165]:
       
       all_stats.append({
@@ -49,27 +60,31 @@ with open(inpfile, 'r') as f:
         'rightOrder' : int(row[14]),
         'err'        : [int(_) for _ in row[16:26]]
       })
-  
+
+    # skip if not in data rows  
     if i<2 or i>155 or (row[0]=='' and row[1]==''):
       continue
     
-    if row[0]!='':
+    
+    if row[0]!='': # if header row containing sim name
       #print row[0]
       simname=row[0]
       all_data[simname]=[]
       
-    else:
-      if simname=='ASW0001gae':
+    else: # regular data
+      if simname=='ASW0001gae': # we don't like this sim
         continue
-      #print row[1], row[29], row[30], row[31]
 
+      # recover the fractions in row 12 and 13 as such..
+      # makes sure strings like '2/3', '1 3/5' ect are recovered as math. fractions/numbers
       r12 = reduce(lambda x, y: x+y, [Frac(_) for _ in row[12].split(' ') if _!='']).limit_denominator()
       r13 = reduce(lambda x, y: x+y, [Frac(_) for _ in row[13].split(' ') if _!='']).limit_denominator()
 
-      #true / false values / strings
+      #true / false values / strings in output
       t = "X"
       f = 'O'
 
+      # read rows
       data = {
         'id'         : row[1],
         'npnt'       : t if row[ 7]=='x' else f,
@@ -85,26 +100,25 @@ with open(inpfile, 'r') as f:
[…]

### 2013-09-08 text/tex/results.tex
ideas for results section

@@
 \section{Results} \label{sec:results}
+
+\subsection{First test} \label{sec:results.1}
+
+The evaluation of the volunteers performance was done manually / by hand/eye, comparing their input to the actual contour lines.
+The actual contour lines were generated from the original simulation parameters.
+
+For each model a set of binary tests was done:
+
+\begin{enumerate}
+  \item (approxPlace) have the images approximately been identified correctly (approx within of 10\% of img width)
+%  \item (rightPlace) are the images placed precisely (approx. within 5\% of image width)
+  \item (rightOrder) is the ordering of the images correct (sad / min / max detection and ordering)
+\end{enumerate}
+
+Additionally, ten categories of error that could happend were recorded, where each model could contain multiple of those:
+
+\begin{enumerate}
+  \item innacurate placement in an extended arc
+  \item wrongly identified sad and min in 3 image configuration
+  \item identified only 3 instead of 5 images
+  \item tried to model an arc with a min instead of min-sad-min
+  \item PI-err (rotation by 180 degrees; in 5 image configuration, exchanged the ordering of the two saddle points)
+  \item PI/2-err (rotation by 90 degree; sad->min->sad->min->sad)
+  \item missed faint image(s)
+  \item tried to model an arc with min-sad-min instead of only min
+  \item did identify two close by images as one
+  \item used 7 or more image to model a 5 image system.
+\end{enumerate}
+
+\input{tab/auto/_stats}
+
+By looking at the numbers in table \ref{tab:stats}, we conclude that the volunteers are performing very well identifying images and putting them approximately at the right spot (test 1), with a performance of 92%.
+Most of the problems where due to unclear arc-like structures. (Error 1, p=0.18)
+
+\todo{add more detail?} Add more details like total amount of images detected / tot am images (rightPlace fraction)??
+
+The ordering and assignment of minima / maxima / saddlepoint posed a more difficult task.
+In p=59\% (N=70) of the cases the volunteers succeeded to identify the right configuration.
+Most of the failures are due to error type 6 (PI/2), with N=38, p=32\%, followed by type 5 (PI) with N=7, p=6\%.
+For an example of type 6 error, see \figref{6971}.
+
+\subsection{Second test} \label{sec:results.2}
 
@@
 mass still quite well recovered.
+
+\clearpage

### 2013-09-08 plots/gen_table.py

@@ -123,12 +123,13 @@ ca = '''1. col: places of imgs idendited correctly; 2. col: more or less
 correct identification of extr points. 3.col: exact identification of extremal
 points. 4.col: type(s) of errors ocured....
 1) innacurate in arc
-2) min-sad conf in 3 img.
+2) min-sad conf in 3 img. conf
 3) only 3 imgs instead 5
-4) arc not expanded ()
+4) arc not expanded
 5) PI-err (rot by 180deg, 2 saddle points false)
 6) PI/2-err (rot by 90deg, sad->min->sad->min->sad)
 7) missed faint image(s)
+8) expanded point too mucj (min-sad-min instead of only min)
 8) missed double img(s)
 9) too many imgs
 '''
@@ -273,7 +274,7 @@ for key, val in all_stats[1].items():
 
 # table footer
 tex_stats += '\n'
-tex_stats += r'\end{tabular}\caption{Stats}\label{tab:stats}\end{table}'
+tex_stats += r'\end{tabular}\caption{Error Statistics}\label{tab:stats}\end{table}'
 
 # ... and save
 path = os.path.join(os.path.abspath(out_file_path), '_stats.tex')

### 2013-09-08 text/tex/outlook.tex
ideas for outlook, errornames in gen_table

\section{Extensions needed} \label{sec:todo}

As seen in \secref{results.1}, volunteers fail mostly in two situations:
when to identify an arclike structre while placing the points
and in 5 image configurations, that are quite symmetric, then to identify the correct ordering of the points.

The first can be improved by additional images from different filters (?)
Will be implemented with a next major \spl upgrade.

The second can be made less by better training (introduce the ruler), maybe more filters help too?

We hope that those problems are also taken care of when staring collaborative modelling. (Already in progress)
Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models ect...)

### 2013-09-08 plots/gen_table.py

@@ -122,17 +122,20 @@ with open(inpfile, 'r') as f:
 ca = '''1. col: places of imgs idendited correctly; 2. col: more or less
 correct identification of extr points. 3.col: exact identification of extremal
 points. 4.col: type(s) of errors ocured....
-1) innacurate in arc
-2) min-sad conf in 3 img. conf
-3) only 3 imgs instead 5
-4) arc not expanded
-5) PI-err (rot by 180deg, 2 saddle points false)
-6) PI/2-err (rot by 90deg, sad->min->sad->min->sad)
-7) missed faint image(s)
-8) expanded point too mucj (min-sad-min instead of only min)
-8) missed double img(s)
-9) too many imgs
 '''
+eee=[
+'inaccurate placement in an extended arc',
+'wrongly identified sad and min in 3 image configuration',
+'identified only 3 instead of 5 images',
+'tried to model an arc with a min instead of min-sad-min',
+'PI-err (rotation by 180 degrees; in 5 image configuration, exchanged the ordering of the two saddle points)',
+'PI/2-err (rotation by 90 degree; sad->min->sad->min->sad)',
+'missed faint image(s)',
+'tried to model an arc with min-sad-min instead of only min',
+'did identify two close by images as one',
+'used 7 or more image to model a 5 image system. ',
+]
+ca+='; '.join(['%i) %s'%(i+1,e) for i, e in enumerate(eee)])
 
 
 settings = {
@@ -261,7 +264,7 @@ nn=int(all_stats[1]['n'])
 for key, val in all_stats[1].items():
   if key=='err': # split up errors
     for m, v in enumerate(val):
-      tex_stats += r' %s & %s & %.2f\\' % (key+'%2i'%m, v, float(v)/nn)
+      tex_stats += r' %s & %s & %.2f\\' % (key+' %i'%(m+1), v, float(v)/nn)
       tex_stats += '\n'
 
   elif key=='': # should not happend...

### 2013-09-08 plots/gen_plots.py
added eR plots

@@ -822,3 +822,126 @@ def draw_mod(mid, elem, data, sims):
   
   print ' ... DONE.'
   print '  - %s' % imgname
+
+tmp={}
+def plotAllRE():
+  '''plots the final big scatter plot with all rE'''
+  print '> drawing einsteinR plot',
+
+  plots = [True, True, True, True]
+  
+  path = resdir
+
+  spg.genREData() # genreate the data
+
+  #collect data
+  sims = {}
+  lu = {}
+  for i, item in enumerate(spg.sims.items()):
+    key, val = item
+    sims[key] = val['rE']
+    lu[key] = i
+    
+  xi = []
+  re = []
+  ep = []
+  em = []
+  re_rel = []
+  ep_rel = []
+  em_rel = []
+  se = []
+  for dat in spg.data:
+    try:
+      re_sim = sims[dat['name']]
+    except KeyError:
+      continue
+    se.append(re_sim)
+    xi.append(dat['id'])
+    re.append(dat['rE_mean'])
+    ep.append(dat['rE_max'])
+    em.append(dat['rE_min'])
+    re_rel.append(dat['rE_mean']/re_sim)
+    ep_rel.append(dat['rE_max']/re_sim)
+    em_rel.append(dat['rE_min']/re_sim)
+    
+  xi = np.array(xi)
+  re = np.array(re)
+  ep = np.array(ep) - re
+  em = re - np.array(em)
+  ee = [ep, em]
+  re_rel = np.array(re_rel)
+  ep_rel = np.array(ep_rel) - re_rel
+  em_rel = re_rel - np.array(em_rel)
+  ee_rel = [ep_rel, em_rel]
+  
+  if plots[0]:
+    pl.figure()
+    pl.errorbar(xi, re, ee, marker='s', mfc='blue', ls='' ,ecolor='blue')
+    pl.plot(xi, se, 'rx')
+    
+    pl.savefig(os.path.join(path, 'eR_1.png'))
+    print '.',
+    #pl.show()
+  
+  if plots[1]:  
+    pl.figure()
+    pl.errorbar(xi, re_rel, ee_rel, marker='s', mfc='blue', ls='' ,ecolor='blue')
+    pl.plot(xi, xi*0+1, '-r')
+    
+    pl.savefig(os.path.join(path, 'eR_2.png'))
+    print '.',
+    #pl.show()
+  
+  if plots[2]:
+    pl.figure()
+    for i, dat in enumerate(spg.data):
+      simn = dat['name']    
[…]

### 2013-09-08 plots/gen_table.py

@@ -119,9 +119,9 @@ with open(inpfile, 'r') as f:
 #==============================================================================
 
 # caption
-ca = '''1. col: places of imgs idendited correctly; 2. col: more or less
-correct identification of extr points. 3.col: exact identification of extremal
-points. 4.col: type(s) of errors ocured....
+ca = '''1. col: id of model; 2.col: places of imgs idendited correctly; 3. col: more or less
+correct identification of extr points. 4.col: exact identification of extremal
+points. 5.col: type(s) of errors ocured....
 '''
 eee=[
 'inaccurate placement in an extended arc',
@@ -129,7 +129,7 @@ eee=[
 'identified only 3 instead of 5 images',
 'tried to model an arc with a min instead of min-sad-min',
 'PI-err (rotation by 180 degrees; in 5 image configuration, exchanged the ordering of the two saddle points)',
-'PI/2-err (rotation by 90 degree; sad->min->sad->min->sad)',
+'PI/2-err (rotation by 90 degree; sad-min-sad-min-sad)',
 'missed faint image(s)',
 'tried to model an arc with min-sad-min instead of only min',
 'did identify two close by images as one',

### 2013-09-08 spaghetti/gen_kappa_encl_plots.py

@@ -350,3 +350,15 @@ def getEinsteinR(x, y):
     else:
       return False
     
+def genREData():
+  '''generates the einsteinradius data for all the elemts'''
+  
+  # all models:
+  for elem in data:
+    elem['rE_mean'] = getEinsteinR(elem['x'], elem['y'])
+    elem['rE_max'] = getEinsteinR(elem['x'], elem['err_p'])
+    elem['rE_min'] = getEinsteinR(elem['x'], elem['err_m'])
+  
+  for key, val in sims.items():
+    val['rE'] = getEinsteinR(val['x'], val['y'])
+    

### 2013-09-08 text/tex/results.tex
ideas for outlook, errornames in gen_table
added eR plots
cleaned up make files

@@
 \begin{enumerate}
-  \item innacurate placement in an extended arc
+  \item inaccurate placement in an extended arc
   \item wrongly identified sad and min in 3 image configuration
@@
   \item PI-err (rotation by 180 degrees; in 5 image configuration, exchanged the ordering of the two saddle points)
-  \item PI/2-err (rotation by 90 degree; sad->min->sad->min->sad)
+  \item PI/2-err (rotation by 90 degree; sad$\longmapsto$min$\longmapsto$sad$\longmapsto$min$\longmapsto$sad)
   \item missed faint image(s)
@@
 \subsection{Second test} \label{sec:results.2}
+
+\begin{figure}[htbp]
+  \centering
+    \includegraphics[width=0.80\textwidth]{fig/sims/eR_4.png}
+  \caption{relative Einstein Radius for each model, per sim. }
+  \label{fig:eR_all}
+\end{figure}
+
+\hr
 

### 2013-09-08 make.sh

@@ -23,20 +23,28 @@ echo "or similar"
 echo "(but you should really use virtualenv and pip)"
 echo "and of course numpy and matplotlib"
 
+read -p "Press [Enter] key to start..."
+
 echo " > generating data (many.py)"
 cd systems
 python many.py
 cd ..
 
-echo " > generating plots (gen_plots.py)"
+echo " "
+echo "/===================================="
+echo "=> generating plots (gen_plots.py)"
 cd plots
-python -c "import gen_plots; run()"
+python -c "from gen_plots import *; run()"
 cd ..
 
-echo " > generating tables (gen_tables.py)"
+echo " "
+echo "/===================================="
+echo "=> generating tables (gen_tables.py)"
 cd plots
-python gen_tables.py
+python gen_table.py
 cd ..
 
+echo " "
+echo "/===================================="
 echo "finished. you can start working on the tex files in /text"
 echo "please only compile the master file <ms.tex>"

### 2013-09-08 plots/gen_plots.py

@@ -141,6 +141,7 @@ def run():
   all_sim_plots()
   all_mod_plots()
   all_tex()
+  plotAllRE()
 
 #==============================================================================
 #   # generate sim plots
@@ -826,7 +827,7 @@ def draw_mod(mid, elem, data, sims):
 tmp={}
 def plotAllRE():
   '''plots the final big scatter plot with all rE'''
-  print '> drawing einsteinR plot',
+  print '> drawing EinsteinR plots',
 
   plots = [True, True, True, True]
   
@@ -914,7 +915,7 @@ def plotAllRE():
     pl.xticks(range(i+1), lbls, rotation=90)
     pl.gcf().subplots_adjust(bottom=0.2)
     pl.xlim([-.5, i+0.5])
-    pl.ylabel(r'Einstein Radius $\Theta_{E}$')
+    pl.ylabel(r'Einstein Radius $\Theta_{\text{E}}$')
 
     pl.savefig(os.path.join(path, 'eR_3.png'))
     print '.',

### 2013-09-08 update.sh

@@ -1,0 +1,26 @@
+#!/bin/bash
+echo "this expects that you have already run make"
+echo "and no major changes happend"
+echo "will regenerate all tex files and copy over the plots"
+
+read -p "Press [Enter] key to start..."
+
+echo " "
+echo "/===================================="
+echo "=> generating plots (gen_plots.py)"
+pwd
+cd plots
+pwd
+python -c "from gen_plots import *; all_tex(); plotAllRE()"
+cd ..
+
+echo " "
+echo "/===================================="
+echo "=> generating tables (gen_tables.py)"
+cd plots
+python gen_table.py
+cd ..
+
+echo " "
+echo "/===================================="
+echo "finished update"

### 2013-09-09 update.sh
update.sh

@@ -8,9 +8,7 @@ read -p "Press [Enter] key to start..."
 echo " "
 echo "/===================================="
 echo "=> generating plots (gen_plots.py)"
-pwd
 cd plots
-pwd
 python -c "from gen_plots import *; all_tex(); plotAllRE()"
 cd ..
 

### 2013-09-09 text/tex/spaghetti.tex
ideas for results section
added eR plots
time to go to sleep

@@
 \begin{figure}
-\includegraphics[height=.3\vsize]{fig/spag6941.png} \\
-\includegraphics[height=.3\vsize]{fig/spag7022.png} \\
-\includegraphics[height=.3\vsize]{fig/spag6919.png}
+  \centering
+  \subfigure{\includegraphics[width=0.45\textwidth]{fig/sims/006941/input.png}}
+  \subfigure{\includegraphics[width=0.45\textwidth]{fig/sims/007022/input.png}}
+  \subfigure{\includegraphics[width=0.45\textwidth]{fig/sims/006919/input.png}}
+%  \includegraphics[width=.45\textsize]{fig/spag6941.png}
+%  \includegraphics[width=.45\textsize]{fig/spag7022.png}
+%  \includegraphics[width=.45\textsize]{fig/spag6919.png}
   \caption{Examples of Spaghetti input.  The models from these appear
-  later in Figures \ref{fig:6941}, \ref{fig:7022} and
-  \ref{fig:6919}. \label{fig:input-spag}}
+    later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
+  \label{fig:input-spag}}
 \end{figure}
@@
 
-The testing was done with a small number of volunteers ($N_v=4$) and a set of simulated configurations ($N_{sim}=30$) chosen to represent some typical configurations.
+The testing was done with a small number of volunteers ($N_v=4$) and a set of simulated configurations ($N_{sim}=29$) chosen to represent some typical configurations.
 
@@
 Second to come up with a logical ordering of the arrival time of those lensed images.
+It also tries to identify the problems volunteers encounter during modelling.
 

### 2013-09-09 plots/gen_plots.py

@@ -935,8 +935,10 @@ def plotAllRE():
     for i, item in enumerate(spg.sims.items()):
       key, val = item
     pl.plot([-0.5, i+0.5], [1,1], '--r')
-      
+    
     pl.xticks(range(i+1), lbls, rotation=90)
+    pl.yticks(np.linspace(0,2,(2/0.25)+1))
+    pl.grid(axis='y')
     pl.gcf().subplots_adjust(bottom=0.2)
     pl.xlim([-.5, i+0.5])
     pl.ylabel(r'rel Einstein Radius $\Theta_{\text{E}}$/$\Theta_{\text{E, sim}}$')

### 2013-09-09 text/tex/appendix.tex
edits

@@
 
-First we define comoving distances, as follows.
+In Section \ref{sec:Fermat}, for the sake of a more intuitive
+explanation, we suppressed some constant factors in
+equations \eqref{eq:Ageom}, \eqref{eq:Ageom} and \eqref{eq:Poisson}.
+Here we fill in the details.
+
+First we define comoving distances.
 \begin{equation}
@@
 \end{equation}
-
-\begin{equation}
-\begin{aligned}
-A           &= cD_L \frac{D_{LS}}{D_S} \times t \\
-\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L}{1+z} \frac{D_{LS}}{D_S}
-               \times \Sigma(x,y)
-\end{aligned}
-\end{equation}
-
-Let $x,y$ be physical distances on the lens plane and $D_L$ be
-comoving distances.  Then
-\begin{equation}
-t_{\rm geom} = \frac{(1+z_L)^2}{2cD_L} \frac{D_S}{D_{LS}}
-\left( (x-s_x)^2 + (y-s_y)^2 \right)
-\end{equation}
-\begin{equation}
-\nabla^2 t_{\rm grav} = (1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
-\end{equation}
-
 Can replace comoving distances $D$ with angular-diameter distances
-$d$,
-\begin{equation}
-(x,y) = d_L (\theta_x,\theta_y)
-\end{equation}
-and so on.
+$d$, and so on:
 \begin{equation}
@@
 Angular-diameter distances are always smaller than comoving distances.
+Locations on the lens plane can be replaced angular coordinates on the
+sky, as
+\begin{equation}
+(x,y) = d_L (\theta_x,\theta_y) \,.
+\end{equation}
 
-Recover equations (2.1) to (2.6) from \cite{1986ApJ...310..568B}.
+The $A$ times and the $\kappa$ density are related to physical arrival
+time $t$ and density $\Sigma$ as
+\begin{equation}
+\begin{aligned}
+A           &= \frac{cD_L}{(1+z_L)^2} \frac{D_{LS}}{D_S} \times t \\
+\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L}{1+z_L} \frac{D_{LS}}{D_S}
+               \times \Sigma(x,y)
+\end{aligned}
+\end{equation}
+Letting the source be behind $(s_x,s_y)$ rather than behind the
+origin, we have
+\begin{equation}
+\begin{aligned}
+t_{\rm geom} &= \frac{(1+z_L)^2}{2cD_L} \frac{D_S}{D_{LS}}
+\left( (x-s_x)^2 + (y-s_y)^2 \right)
+\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
+\end{aligned}
+\end{equation}
+We can now compare with equations (2.1) to (2.6)
+from \cite{1986ApJ...310..568B}.

### 2013-09-09 text/tex/spaghetti.tex
edits

--- 2013-09-07 text/tex/spaghetti.tex
@@
 
-There are three kinds of simulated lenses:
+There are three kinds of simulated lenses.
+
+Sims produced using {\tt gravlens}
+\citep{2001astro.ph..2341K,2001astro.ph..2340K}.
 
@@
 \end{itemize}
-Lenses follow \cite{2001astro.ph..2341K,2001astro.ph..2340K}.
 
-Parameters can be looked up in online supplement.
+Singular isothermal is in equations (33-35) of
+\cite{2001astro.ph..2341K} with core radius set to zero. NFW is
+equations (48,50) of the same paper, and shear is the $\gamma$ term in
+equation (76) of that work.
 
@@
 The testing was done with a small number of volunteers ($N_v=4$) and a set of simulated configurations ($N_{sim}=30$) chosen to represent some typical configurations.
+\todo{witch??}
 

### 2013-09-09 text/tex/theory.tex
edits

@@
 what it would have been with no lensing, is
-\begin{equation}
+\begin{equation} \label{eq:Aarriv}
 \hbox{arrival time} = A_{\rm geom} + A_{\rm grav} \,.
@@
 surface density becomes a dimensional field $\kappa(x,y)$.  Then
-\begin{equation} \label{eq:poisson}
+\begin{equation} \label{eq:Poisson}
 \nabla^2 A_{\rm grav}(x,y) = -2\kappa(x,y) \,.
@@
 $\kappa(x,y)$, but as a differential equation that must be solved.
-Equations of the type \eqref{eq:poisson} are, however, well known (as
+Equations of the type \eqref{eq:Poisson} are, however, well known (as
 Poisson equations in two dimensions), and there are techniques for

### 2013-09-09 systems/poten.py

@@ -1,6 +1,6 @@
 from numpy import sqrt,pi,cos,sin,arctan2,arctan,arctanh,log
 
-# See eqns (34,35) from Keeton astro-ph/0102341
+# See eqns (33-35) from Keeton astro-ph/0102341
 def poten_SIE(x,y,reinst,ell,ell_pa):
     pa = ell_pa*pi/180 + pi/2
     q = 1 - ell

### 2013-09-09 text/tex/results.tex

@@
 
+In \figref{eR_all} can be seen, that the calculated einstein radius $\Theta_E$ of the models tend to be too high.
+The overshoot varies from around 0.2 to 0.4 for for good models.
+One of the reasons for this is, that it's hard to get the center of the lens on spot.
+An offset leads to a a flatter mass profile for the model compared to the simulation.
+
 \begin{figure}[htbp]

### 2013-09-09 text/tex/results.tex
edits

@@
-\section{Results} \label{sec:results}
+\section{The lens modelling challenge} \label{sec:results}
+
+\subsection{\sw} \label{sec:SpaceWarps}
+some words about \sw
+
+\subsection{The simulated lenses} \label{sec:sims}
+
+In order to estimate users performance, \sw included some simulated lensing configurations in their image set.
+
+There are three kinds of simulated lenses.
+
+Sims produced using {\tt gravlens}
+\citep{2001astro.ph..2341K,2001astro.ph..2340K}.
+
+\begin{itemize}
+  \item {\em Quasar:\/} A singular elliptical isothermal plus constant
+    external shear, with a circular Gaussian source.
+  \item {\em Galaxy:\/} Lens as with quasar, but elliptical de
+    Vaucouleurs source.
+  \item {\em Galaxy:\/} Lens is circular NFW plus one dominant
+    elliptical SIE and perturbating elliptical SIEs, source as with
+    galaxy.
+\end{itemize}
+
+Singular isothermal is in equations (33-35) of
+\cite{2001astro.ph..2341K} with core radius set to zero. NFW is
+equations (48,50) of the same paper, and shear is the $\gamma$ term in
+equation (76) of that work.
+
+\subsection{The Test setup} \label{sec:testsetup}
+We made use of this simulated lenses and further tested the volunteers abilities to model those simulated lensing images.
+
+The testing was done with a small number of volunteers ($N_v=4$) and a set of simulated configurations ($N_{sim}=29$) chosen to represent some typical configurations.
+
+Two tests were done on the models created by the volunteers:
+
+\begin{enumerate}
+  \item Correct identification of extremal points
+  \item Reproduction of the mass distribution of the lens
+\end{enumerate}
+
+The first test required the volunteers to accomplish 2 tasks: First to identify and distinguish lensed images from background light sources.
+Second to come up with a logical ordering of the arrival time of those lensed images.
+It also tries to identify the problems volunteers encounter during modelling.
+
+The second test compared the enclosed mass $\kappa(r)$ of the models and the simulations.
 

### 2013-09-09 text/tex/spaghetti.tex
edits

--- 2013-09-09 text/tex/spaghetti.tex
@@
-\section {Setup} \label{sec:setup}
-
-\subsection{\sw} \label{sec:SpaceWarps}
-some words about \sw
-
 \begin{figure}
   \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]{fig/sims/006941/input.png}}
-  \subfigure{\includegraphics[width=0.45\textwidth]{fig/sims/007022/input.png}}
-  \subfigure{\includegraphics[width=0.45\textwidth]{fig/sims/006919/input.png}}
-%  \includegraphics[width=.45\textsize]{fig/spag6941.png}
-%  \includegraphics[width=.45\textsize]{fig/spag7022.png}
-%  \includegraphics[width=.45\textsize]{fig/spag6919.png}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/006941/input.png}} \\
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/007022/input.png}} \\
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/006919/input.png}}
   \caption{Examples of Spaghetti input.  The models from these appear
@@
 \spl provides plug ins to support any data source. At the moment, a data link to \sw and to the MasterLens database\needcite are implemented.
-
-\subsection{The simulated lenses} \label{sec:sims}
-
-In order to estimate users performance, \sw included some simulated lensing configurations in their image set.
-
-There are three kinds of simulated lenses:
-
-\begin{itemize}
-  \item {\em Quasar:\/} A singular elliptical isothermal plus constant
-    external shear, with a circular Gaussian source.
-  \item {\em Galaxy:\/} Lens as with quasar, but elliptical de
-    Vaucouleurs source.
-  \item {\em Galaxy:\/} Lens is circular NFW plus one dominant
-    elliptical SIE and perturbating elliptical SIEs, source as with
-    galaxy.
-\end{itemize}
-Lenses follow \cite{2001astro.ph..2341K,2001astro.ph..2340K}.
-
-Parameters can be looked up in online supplement.
-
-\subsection{The Test setup} \label{sec:testsetup}
-We made use of this simulated lenses and further tested the volunteers abilities to model those simulated lensing images.
-
-The testing was done with a small number of volunteers ($N_v=4$) and a set of simulated configurations ($N_{sim}=29$) chosen to represent some typical configurations.
-
-Two tests were done on the models created by the volunteers:
-
-\begin{enumerate}
-  \item Correct identification of extremal points
-  \item Reproduction of the mass distribution of the lens
-\end{enumerate}
-
-The first test required the volunteers to accomplish 2 tasks: First to identify and distinguish lensed images from background light sources.
-Second to come up with a logical ordering of the arrival time of those lensed images.
-It also tries to identify the problems volunteers encounter during modelling.
-
-The second test compared the enclosed mass $\kappa(r)$ of the models and the simulations.

### 2013-09-09 plots/gen_plots.py
tested repaired build queue on linux

@@ -68,7 +68,7 @@ params = {
 #  'backend': 'ps',
   'text.latex.preamble': [
     r"\usepackage{amsmath}",
-    r"\usepackage{textgreek}"
+    #r"\usepackage{textgreek}"
     ],
   'mathtext.default': 'regular',
   'axes.labelsize': 14,

### 2013-09-09 plots/gen_table.py

@@ -34,6 +34,11 @@ out_file_path = '../text/tab/auto'
 inpfile = os.path.join(os.path.abspath(inp_file_path),inp_file_name)
 outfile = os.path.join(os.path.abspath(out_file_path),out_file_name)
 
+try:
+  os.makedirs(os.path.abspath(out_file_path))
+except:
+  print 'error creating paths'
+  
 
 all_data = {}
 all_stats = []

### 2013-09-09 plots/gen_table.py
bugfix

@@ -36,8 +36,9 @@ outfile = os.path.join(os.path.abspath(out_file_path),out_file_name)
 
 try:
   os.makedirs(os.path.abspath(out_file_path))
-except:
-  print 'error creating paths'
+except OSError as e:
+  print 'error creating paths', e
+  print 'continuing'
   
 
 all_data = {}

### 2013-09-09 text/tex/intro.tex
commit to be merged

@@
 
-Spacewarps is the future.\needcite
+%Spacewarps is the future.\needcite
+%
+%Moreover, that lens modelling is an essential accompaniment to lens
+%searches has been clear since the early days \citep{1981ApJ...244..723Y}.
+%
+%Why Fermat's principle?  Details in \ref{sec:Fermat} and
+%\ref{sec:more-theory}.
+%
+%SpaghettiLens details in \ref{sec:SpaghettiLens}.
+%
+%Sims in \ref{sec:sims}.
+%
+%\needfig[testfig]
+%
+%Results in \ref{sec:results}.
+%
+%Things to do in \ref{sec:todo}.
+%
+%This is a general test \secref{todo}\todo{remove this part again}
+%
+%\hr
 
-Moreover, that lens modelling is an essential accompaniment to lens
-searches has been clear since the early days \citep{1981ApJ...244..723Y}.
+Gravitational lenses (GLs), predicted by ART, can be used as a tool by astronomers to examine many properties of the cosmos.
+They allow for example to estimate the masses and their distribution of galaxies, automatically including the hard to grasp dark matter.
+Further more, they allow an estimate on cosmological parameters like the Hubble constant \citep{Saha2006} and the mass density \needcite.
 
-Why Fermat's principle?  Details in \ref{sec:Fermat} and
-\ref{sec:more-theory}.
-
-SpaghettiLens details in \ref{sec:SpaghettiLens}.
-
-Sims in \ref{sec:sims}.
-
-\needfig[testfig]
-
-Results in \ref{sec:results}.
-
-Things to do in \ref{sec:todo}.
-
-This is a general test \secref{todo}\todo{remove this part again}
-
-\hr
-
-Gravitational Lensing, prediced bz ART and first confirmed by ... can be used as a tool by astronomers to examine structres far away.
-Allows to esimate masses, not only visual but total masses (inkl dark matter).
-Further more, allows to esimate cosmologial parameters, like Hubblekonst (Age of universe \needcite) and massdenity \needcite.
-To accheve this, one needs to find / identify lenses and in a second step model those lenses to get involved parameters.
+%To accheve this, one needs to find / identify lenses and in a second step model those lenses to get involved parameters.
 
@@
 
-Next step involves modelling. that needs advanced knolegde and takes a lot of time.
+Next step involves modeling. that needs advanced knolegde and takes a lot of time.
 Too much time to be done by astronomers them selves as the results from new surveys and identifiers like \sw come in.

### 2013-09-09 text/tex/results.tex
test setup

--- 2013-09-09 text/tex/results.tex
@@
-\section{Results} \label{sec:results}
+\section{The lens modeling challenge} \label{sec:results}
+
+\subsection{\sw} \label{sec:SpaceWarps}
+some words about \sw
+
+\subsection{The simulated lenses} \label{sec:sims}
+
+In order to estimate users performance, \sw included some simulated lensing configurations in their image set.
+
+There are three kinds of simulated lenses.
+
+Sims produced using {\tt gravlens}
+\citep{2001astro.ph..2341K,2001astro.ph..2340K}.
+
+\begin{itemize}
+  \item {\em Quasar:\/} A singular elliptical isothermal plus constant
+    external shear, with a circular Gaussian source.
+  \item {\em Galaxy:\/} Lens as with quasar, but elliptical de
+    Vaucouleurs source.
+  \item {\em Galaxy:\/} Lens is circular NFW plus one dominant
+    elliptical SIE and perturbating elliptical SIEs, source as with
+    galaxy.
+\end{itemize}
+
+Singular isothermal is in equations (33-35) of
+\cite{2001astro.ph..2341K} with core radius set to zero. NFW is
+equations (48,50) of the same paper, and shear is the $\gamma$ term in
+equation (76) of that work.
+
+\subsection{The test setup} \label{sec:testsetup}
+
+We made use of the simulated lenses (sims) to test the volunteers abilities to model those sims.
+
+The testing was done with a small number of volunteers ($N_\text{v}=4$) and a set of simulated configurations ($N_\text{sim}=29$) chosen to represent some typical configurations.
+The volunteers were given an introduction video explaining the theory behind gravitational lensing and the usage of \spl.
+Additionally the volunteers had the opportunity to pose general questions to the experts during several hangouts during the training phase that preceded the modeling challenge.\footnote{No questions about the to be modeled sims were answered.}
+
+To estimate the performance of the volunteers and the quality of the generated models, two test T1 and T2 were done.
+%\begin{enumerate}
+%  \item (T1) Correct identification of extremal points
+%  \item (T2) Reproduction of the mass distribution of the lens
+%\end{enumerate}
+
+T1 tested the volunteers ability to reconstruct the arrival time surface given a survey image containing a sim.
+This task consists of two parts.
+First, the correct identification and location of lensed images (T1a).
+Second to find the correct ordering in respect of the arrival time for the identified lensed images (T1b).
+
+While we expected T1a to be trivial, given the nature of the survey images and the success of \sw, we expect T1b to pose more problems.
+T1b tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
+While we can provide the volunteers with some general rule of thumbs, T1b involves a lot of imagination and guessing and therefore a training effect could be visible in this test.
+
+T1 was also designed to give some feedback on the difficulties volunteers encounter, to further imrove the tutorial materials.
+
+The second test T2 was to compare the desired results of a modeling process, the mass distribution of the lens $\kappa(x, y)$.
+To get a means of comparing the simulated data (sims) to the generated modeled data (models), the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ for both was calculated.
+The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between the sims and the models.
+
+\todo{!} Should I write down prior expectation for T2 too?
 

### 2013-09-09 text/tex/results.tex
tests.1

@@
 
-\subsection{First test} \label{sec:results.1}
+\subsection{T1} \label{sec:tests.t1}
 
-The evaluation of the volunteers performance was done manually / by hand/eye, comparing their input to the actual contour lines.
-The actual contour lines were generated from the original simulation parameters.
+The evaluation of the volunteers performance for T1 was done manually, comparing their input from \spl () and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot genreated using simulation parameters.
+\figref{output_compare} shows the setup used to evaluate the models.
+
+\begin{figure}
+  \centering
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/006917/input.png}}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/006917/spaghetti.png}} \\
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/ASW0001hpf/arriv.png}}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/ASW0001hpf/extr_points.png}}
+  \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom left), additionally using a extremal point detector (bottom right).
+  \label{fig:output_compare}}
+\end{figure}
 

### 2013-09-09 text/tex/outlook.tex
t1

@@
 
-As seen in \secref{results.1}, volunteers fail mostly in two situations:
+As seen in \secref{tests.1}, volunteers fail mostly in two situations:
 when to identify an arclike structre while placing the points

### 2013-09-09 text/tex/results.tex
t1

@@
 
-\subsection{T1} \label{sec:tests.t1}
+\subsection{First test T1} \label{sec:tests.t1}
 
-The evaluation of the volunteers performance for T1 was done manually, comparing their input from \spl () and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot genreated using simulation parameters.
-\figref{output_compare} shows the setup used to evaluate the models.
+The evaluation of the volunteers performance for T1 was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot genreated using simulation parameters.
+\Figref{output_compare} shows the setup used to evaluate the models.
 
@@
 
-For each model a set of binary tests was done:
+To evaluate T1a and T1b, each model was evaluated with two binary tests.
+The first is passed, if all the images have been identified and are approx. within $\pm0.05\cdot\text{imgwidth}$.
+The second test is passed, if those identified points have the right parity, ordering with respect to arrival time.
 
-\begin{enumerate}
-  \item (approxPlace) have the images approximately been identified correctly (approx within of 10\% of img width)
-%  \item (rightPlace) are the images placed precisely (approx. within 5\% of image width)
-  \item (rightOrder) is the ordering of the images correct (sad / min / max detection and ordering)
-\end{enumerate}
+Additionally, ten types of errors that could occur were analyzed, where each model could contain multiple of those.
+The complete table with the results for each model can be found in the appendix, \tabref{detail_results}.
 
-Additionally, ten categories of error that could happend were recorded, where each model could contain multiple of those:
+In \tabref{stats} a summary of this evaluation is presented.
 
-\begin{enumerate}
-  \item inaccurate placement in an extended arc
-  \item wrongly identified sad and min in 3 image configuration
-  \item identified only 3 instead of 5 images
-  \item tried to model an arc with a min instead of min-sad-min
-  \item PI-err (rotation by 180 degrees; in 5 image configuration, exchanged the ordering of the two saddle points)
-  \item PI/2-err (rotation by 90 degree; sad$\longmapsto$min$\longmapsto$sad$\longmapsto$min$\longmapsto$sad)
-  \item missed faint image(s)
-  \item tried to model an arc with min-sad-min instead of only min
-  \item did identify two close by images as one
-  \item used 7 or more image to model a 5 image system.
-\end{enumerate}
+%\begin{enumerate}
+%1  \item inaccurate placement in an extended arc
+%2  \item wrongly identified sad and min in 3 image configuration
+%3  \item identified only 3 instead of 5 images
+%4  \item tried to model an arc with a min instead of min-sad-min
+%5  \item PI-err (rotation by 180 degrees; in 5 image configuration, exchanged the ordering of the two saddle points)
+%6  \item PI/2-err (rotation by 90 degree; sad$\longmapsto$min$\longmapsto$sad$\longmapsto$min$\longmapsto$sad)
+%7  \item missed faint image(s)
+%8  \item tried to model an arc with min-sad-min instead of only min
+%9  \item did identify two close by images as one
+%10  \item used 7 or more image to model a 5 image system.
+%\end{enumerate}
 
@@
 
-By looking at the numbers in table \ref{tab:stats}, we conclude that the volunteers are performing very well identifying images and putting them approximately at the right spot (test 1), with a performance of 92%.
-Most of the problems where due to unclear arc-like structures. (Error 1, p=0.18)
+We conclude that the volunteers are performing very well identifying images and putting them approximately at the right spot (test 1), with a performance of 92\%.
+Most of the problems where due to unclear arc-like structures (Error 1, p=0.18; Error 4, p=0.03; Error 8, p=0.04).
+Critical failures like the failure to identify all fife images in an fife image system (Error 3, p=0.04) or to include too many images (Error 10, p=0.01) did almost never happend.
+From this we conclude that the introduction materials was adequate and the volunteers got the basics of gravitational lensing.
 
@@
 
-\subsection{Second test} \label{sec:results.2}
+\subsection{Second test T2} \label{sec:results.2}
 

### 2013-09-09 text/tex/results.tex
t1

@@
 
-We conclude that the volunteers are performing very well identifying images and putting them approximately at the right spot (test 1), with a performance of 92\%.
+We conclude that the volunteers are performing very well identifying and positioning images (T1a), with a performance of 92\% (approxPlace p=0.92).
 Most of the problems where due to unclear arc-like structures (Error 1, p=0.18; Error 4, p=0.03; Error 8, p=0.04).
@@
 
+Error 5 and 6 happen mostly in very symetric images, where it's hard to tell
+It changes the orientation of the mass distribution.
+But kappa encl and thus the einsteinradius is not changed, and thus the final result is still a valid model.
+
 \hr

### 2013-09-09 text/tex/spaghetti.tex
edis

@@
 
+In \spl the user input is an educated guess for the topography of the
+arrival-time surface.
+
 A saddle point is recognizable in a contour map because contours do
@@
 
-Give a guess for the maximum, minimum and saddle points.  Program
-tries find $\kappa(x,y)$ that reproduces these properties exactly, and
-looks reasonably like a galaxy.  Solution not unique, an ensemble
-generated.
+It provides a web based graphical user interface that allows to trace contours with the mouse and identify extremal points by clicking on the image.
 
-\hr
-
-SpaghettiLens is build on top of GLASS \citep{Lubini2012}, that builds and improves upon PixeLens \citep{Saha2004}.
-It provides a web based graphical user interface that allows to trace contours with the mouse and identify extremal points by clicking on the image.
-\spl then tries to find a mass distribution $\kappa(x,y)$ that reproduces the input parameters.
-Since there is no unique solution, \spl samples the solution space and produces an ensemble of solutions, as described in the paper \citep{Lubini2012}.
-It provides direct visual feedback by rendering a synthetic image side by side the original image, such that users can directly compare the predictions of their model to the survey image.
-It additinally provides plots of the modelled mass distribution and the contour lines of arrival time surface as feedback.
-
-Using this direct visual feedback, even unexperienced users can successfully model lenses using a iterative approach if they know what to look for.
+\spl provides plug ins to support any data source. At the moment, a data link to \sw and to the MasterLens database\needcite are implemented.
 
 The modelling process consitis of three basic steps:
-
 \begin{enumerate}
@@
 Step 2 by restricting the user, only valid configuration can be entered, the odd number theorem\needcite is taken care of.
+
+Give a guess for the maximum, minimum and saddle points.  Program
+tries find $\kappa(x,y)$ that reproduces these properties exactly, and
+looks reasonably like a galaxy.  Solution not unique, an ensemble
+generated.
+SpaghettiLens is build on top of GLASS \citep{Lubini2012}, that builds and improves upon PixeLens \citep{Saha2004}.
+
+\spl then tries to find a mass distribution $\kappa(x,y)$ that reproduces the input parameters.
+Since there is no unique solution, \spl samples the solution space and produces an ensemble of solutions, as described in the paper \citep{Lubini2012}.
+
 Step 3 by providing the visual feedback. Users can check the generated mass distribution and synthetic image.
 
-Explanation:
-Synthetic image: A rendering of the derivative of the arrival time for each pixel. This leads to a black and white image that reasonably looks like the visual appearace of the model.
+It additinally provides plots of the modelled mass distribution and the contour lines of arrival time surface as feedback.
 
-Volunteers enter an assumed lens configuration and get a rendered synthetic image along with a reconstructed / simulated mass map and reconstructed / simulated contour lines.
-This allows volunteers to directly see the effect of changing the parameters of the model and to iteratively approach a good solution.
+It provides direct visual feedback by rendering a synthetic image side
+by side the original image, such that users can directly compare the
+predictions of their model to the survey image. Explanation: Synthetic
+image: A rendering of the derivative of the arrival time for each
+pixel. This leads to a black and white image that reasonably looks
+like the visual appearace of the model.
+
+Using this direct visual feedback, even inexperienced users can
+successfully model lenses using a iterative approach if they know what
+to look for.
 
@@
 Volunteers then organise them selves to narrow down the different branches and come up with a few consensus models in the end.
-
-\spl provides plug ins to support any data source. At the moment, a data link to \sw and to the MasterLens database\needcite are implemented.

### 2013-09-09 text/tex/results.tex
further work

@@
 Most of the problems where due to unclear arc-like structures (Error 1, p=0.18; Error 4, p=0.03; Error 8, p=0.04).
-Critical failures like the failure to identify all fife images in an fife image system (Error 3, p=0.04) or to include too many images (Error 10, p=0.01) did almost never happend.
-From this we conclude that the introduction materials was adequate and the volunteers got the basics of gravitational lensing.
+Critical failures like the failure to identify all fife images in an fife image system (Error 3, p=0.04) or to include too many images (Error 10, p=0.01) did almost never happened.
+From this we conclude that the introduction materials was adequate and the volunteers understand the basics of gravitational lensing.
 
@@
 
-The ordering and assignment of minima / maxima / saddlepoint posed a more difficult task.
-In p=59\% (N=70) of the cases the volunteers succeeded to identify the right configuration.
-Most of the failures are due to error type 6 (PI/2), with N=38, p=32\%, followed by type 5 (PI) with N=7, p=6\%.
+The assignment of the parity of the images (T1b) posed a more difficult task.
+In 59\% (rightOrder, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
+Most of the failures are due to Error 6 (PI/2), with N=38, p=32\%, followed by Error 5 (PI) with N=7, p=6\%.
 For an example of type 6 error, see \figref{6971}.
+While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
+For more challenging cases, like very symmetrical distribution of the lensed images (for example model 6971, \figref{6971}), those errors should still produce plausible results, as will be explained in the next section.
 
@@
 
-In \figref{eR_all} can be seen, that the calculated einstein radius $\Theta_E$ of the models tend to be too high.
+To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] directly generated in the modeling process.
+From the ensemble of models generated by one modelling process, the mean is taken as the resulting $\kappa(r)$ to calculate $\theta_E$.
+To estimate the errors, also the extremal models are used to estimate a lower and upper limit for $\theta_E$.
+These results can be seen in \figref{eR_all_models}.
+This figure shows that this tecnique of esitmating the error is underestimating the error significantly and should be improved for further analysis.
+
+\begin{figure}[htbp]
+  \centering
+    \includegraphics[width=0.80\textwidth]{fig/sims/eR_1.png}
+  \caption{\tEr \gEr for all models with estimated erros in blue, \gEr of simulation in red}
+  \label{fig:eR_all_models}
+\end{figure}
+
+In \figref{eR_per_sim} can be seen, that the calculated \tEr \gEr of the models tend to be too high.
 The overshoot varies from around 0.2 to 0.4 for for good models.
@@
   \caption{relative Einstein Radius for each model, per sim. }
-  \label{fig:eR_all}
+  \label{fig:eR_per_sim}
 \end{figure}
 
-Error 5 and 6 happen mostly in very symetric images, where it's hard to tell
-It changes the orientation of the mass distribution.
-But kappa encl and thus the einsteinradius is not changed, and thus the final result is still a valid model.
+Error 5 and 6 happen mostly in very symmetric images, where it's hard to tell
+It changes the orientation of the mass distribution \kap[x,y].
+But \kenc and thus the \tEr \gEr is not changed, and thus the final result is still a valid model.
+This can be easily corrected, if further analysis is done for the modeled system and time delays are known.
 
@@
 
+\todo{incl this figs}
 Figure \ref{fig:6941} shows a simple example.

### 2013-09-09 text/tex/theory.tex
edits

@@
+\section{A lens modelling program}
+
+It is well known that making a scientific contribution is a central
+motivation cited by citizen-science volunteers.  So for a lens
+modeller in a citizen-science project, nobody wants that it should
+sacrifice scientific usefulness in order to be fun to use.  That said,
+being aesthetically pleasing and having a short initial learning curve
+are also essential qualities.  Also desirable is that the user is
+encouraged/challenged to go deeper and wider into the subject.
+Yet another aspect is enabling small incremental contributions
+by different people.  \spl tries to address all of these.
+
+\subsection{Theory: Fermat's Principle} \label{sec:Fermat}
+
+There are several ways to understand the formation of arcs and
+multiple images in gravitational lensing.  We will follow some ideas
+originally introduced by \cite{1986ApJ...310..568B}, based on Fermat's
+principle.  They key to this approach is an abstract construct called
+the {\em arrival-time surface.}  This surface cannot itself be
+observed, but several observable quantities can be derived from it.
+
 \begin{figure}
@@
 \end{figure}
-
-\section{Fermat's Principle} \label{sec:Fermat}
-
-There are several ways to understand the formation of arcs and
-multiple images in gravitational lensing.  We will follow some ideas
-originally introduced by \cite{1986ApJ...310..568B}, based on Fermat's
-principle.  They key to this approach is an abstract construct called
-the {\em arrival-time surface.}  This surface cannot itself be
-observed, but several observable quantities can be derived from it.
 

### 2013-09-09 text/tex/spaghetti.tex
more edits

@@
 
-In \spl the user input is an educated guess for the topography of the
-arrival-time surface.
+In \spl the user begins by making an educated guess for the topography
+of the arrival-time surface --- specifically, the locations of the
+maxima, minima and saddle points that correspond to images of the
+brightest part of the source.  Input is given by tracing a proposal
+for the saddle-point contours.  Figure \ref{fig:input-spag} shows some
+example inputs --- and incidentally also indicates the origin of the
+name \spl.
 
-A saddle point is recognizable in a contour map because contours do
-not loop around it, they tend to approach and then pull away.  There
-is one contour ---the saddle-point contour--- that makes an X on the
-saddle point itself.  A saddle point contour may not appear on a plot
-(and Figure~\ref{fig:arriv} does not show one.  But we can tell from
-the other contours where the saddle-point contour would be.
+\spl, as implemented so far, assumes that the lens is dominated by a
+single galaxy, and that the centre of that galaxy is the sole maximum.
+Once the maximum has been identified, we wish to characterize the rest
+of the arrival-time surface in relation to it.  From geometry, any
+surface that has a central maximum and goes high at the edges must
+have a lima\c con-like contour, meaning a two-looped curve with one
+loop inside the other.  The maximum lies within the inner loop, a
+minimum lies between the two loops, and a saddle point lies at the
+self-intersection of the contour.  The top panel of
+Figure~\ref{fig:input-spag} shows an example: red marks a suggested
+maximum, green a saddle point, and blue a minimum.  (The small pink
+dots are just help sketch the proposed contour.)  The middle and lower
+panels of Figure~\ref{fig:input-spag} show a more complex situation.
+The minimum has split into three images: two new minima with a saddle
+point in between.  In this situation the contour through the new
+saddle point forms a figure of eight, its two loops enclosing the two
+new minima.  The configuration of a lima\c con, possibly with a figure
+of eight inside, suffices to give a preliminary description of the
+arrival-time surface for nearly all cases where the lens is dominated
+by one galaxy, and the source is a single small object.
 

### 2013-09-09 text/tex/theory.tex
more edits

@@
 always too faint to see.  The minima and saddle points dominate.
+
+In the lensed examples in Figure~\ref{fig:arriv}, the maximum in each
+case is evident, but how can we distinguish the minima from the saddle
+points?  The key is to examine the contours of equal arrival time.
+Contour curves loop around a minimum or a maximum, but near a saddle
+point, the contours do not loop around, instead they approach and then
+pull away.  The contour curve precisely level with a saddle point
+forms an X at the saddle point.  In Figure~\ref{fig:arriv}, one such X
+is evident.  In other cases the X contour passing through the saddle
+point is not itself plotted, but we can tell the other contours where
+the X would be.  These self-crossing contours will play a special role
+in the following.

### 2013-09-09 text/tex/spaghetti.tex
still editing

@@
 
-It provides a web based graphical user interface that allows to trace contours with the mouse and identify extremal points by clicking on the image.
+The exact placement of the pink spaghetti loops shown in
+Figure~\ref{fig:input-spag} has no significance.  Their utility is
+just to help mark up the image as maximum, minima and saddle points.
+As the figure suggests, the marking up can be done very easily with a
+mouse.  Human intuition is required to (a)~identify lensed images and
+separate them from other background light sources, and (b)~classify
+and order images accoring to arrival time.  But the computer helps by
+allowing only valid lensing configurations to be entered, and ensuring
+that the odd image theorem\needcite is taken care of.
 
@@
 
-The modelling process consitis of three basic steps:
-\begin{enumerate}
-  \item identify lensed images and separate them from other background light sources
-  \item classify and order images accoring to arrival time. (local minima, maxima, saddlepoints)
-  \item fine tune the arrangement, identify addidional external point masses influencing the result
-\end{enumerate}
-
 \spl assists volunteers with several features in this process.
 Step 1 by supplying several images from several bands (not yet implemented in \sw).
-Step 2 by restricting the user, only valid configuration can be entered, the odd number theorem\needcite is taken care of.
 

### 2013-09-10 text/tex/results.tex
more edits
moar results

@@
 The volunteers were given an introduction video explaining the theory behind gravitational lensing and the usage of \spl.
-Additionally the volunteers had the opportunity to pose general questions to the experts during several hangouts during the training phase that preceded the modeling challenge.\footnote{No questions about the to be modeled sims were answered.}
+Additionally the volunteers had the opportunity to pose general questions to the experts during several video conferences during the training phase that preceded the modeling challenge.\footnote{No questions about the to be modeled sims were answered.}
 
@@
 T1b tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
-While we can provide the volunteers with some general rule of thumbs, T1b involves a lot of imagination and guessing and therefore a training effect could be visible in this test.
+While we can provide the volunteers with some general rule of thumbs, T1b involves imagination and guessing and therefore training could improve their skills in a later stage.
 
-T1 was also designed to give some feedback on the difficulties volunteers encounter, to further imrove the tutorial materials.
+T1 was also designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
 
@@
 The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between the sims and the models.
+We also let an expert model two selected systems to compare the results from volunteers to those of a professional.
 
@@
              {fig/sims/ASW0001hpf/extr_points.png}}
-  \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom left), additionally using a extremal point detector (bottom right).
-  \label{fig:output_compare}}
+  \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom left), additionally using a extremal point detector (bottom right).}
+  \label{fig:output_compare}
 \end{figure}
@@
 To evaluate T1a and T1b, each model was evaluated with two binary tests.
-The first is passed, if all the images have been identified and are approx. within $\pm0.05\cdot\text{imgwidth}$.
+The first is passed, if all the images have been identified and are approx. within $\pm0.05\cdot\text{imgage width}$.
 The second test is passed, if those identified points have the right parity, ordering with respect to arrival time.
@@
 But \kenc and thus the \tEr \gEr is not changed, and thus the final result is still a valid model.
+This can be seen by comparing the results for \asw{0h2m}, shown in \figref{kapenc_compare_faulty}: 
+The correct model 7022 ($\mEr=10.76$), which was done by an expert, and
+7020 ($\mEr=10.72$),
+7024 ($\mEr=10.80$),
+7025 ($\mEr=11.16$),
+7021 ($\mEr=11.04$).
+The first three show error 6, while the last is of error type 5.
 This can be easily corrected, if further analysis is done for the modeled system and time delays are known.
+
+\begin{figure}
+  \centering
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/007020/kappa_encl}}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/007021/kappa_encl}} \\
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/007024/kappa_encl}}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/007025/kappa_encl}}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/sims/007022/kappa_encl}}
+  \caption{\kenc for models of \asw{0vqg}: 7020 (top left), 7021 (top right), 7024 (mid left), 7025 (mid right) by volunteers and a correct model 7022 (bottom) by an expert.}
+  \label{fig:kapenc_compare_faulty}
+\end{figure}
+
+Comparing the models of volunteers and experts can be done in \figref{kapenc_compare_faulty}, where only the
+expert got the right configuration, but all the resulting models are comparable, besides rotation.
 

### 2013-09-10 text/tex/results.tex

@@
 
+\subsection{Some example models}
+
+Figure \ref{fig:6941} shows a simple example.
+
+Figure \ref{fig:6975} shows an example where substructure introduces a
+complication, but model ok.
+
+Figure \ref{fig:6937} shows a case where substructure leads to a
+poorer model.
+
+Figure \ref{fig:6975} shows a nice symmetric quad.
+
+Figure \ref{fig:6990} shows a long-axis quad.
+
+Figure \ref{fig:6919} shows a short-axis quad.
+
+Figure \ref{fig:6915} shows an inclined quad.
+
+Figure \ref{fig:6971} shows incorrect identification, but the enclosed
+mass still quite well recovered.
+
 \subsection{The test setup} \label{sec:testsetup}
@@
 
-T1 was also designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
+T1 was also designed to give some feedback on the difficulties volunteers encounter, to further imrove the tutorial materials.
 
@@
              {fig/sims/ASW0001hpf/arriv.png}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/ASW0001hpf/extr_points.png}}
+%  \subfigure{\includegraphics[width=0.45\textwidth]
+%             {fig/sims/ASW0001hpf/extr_points.png}}
   \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom left), additionally using a extremal point detector (bottom right).}
@@
 
-\hr
-
-\todo{incl this figs}
-Figure \ref{fig:6941} shows a simple example.
-
-Figure \ref{fig:6975} shows an example where substructure introduces a
-complication, but model ok.
-
-Figure \ref{fig:6937} shows a case where substructure leads to a
-poorer model.
-
-Figure \ref{fig:6975} shows a nice symmetric quad.
-
-Figure \ref{fig:6990} shows a long-axis quad.
-
-Figure \ref{fig:6919} shows a short-axis quad.
-
-Figure \ref{fig:6915} shows an inclined quad.
-
-Figure \ref{fig:6971} shows incorrect identification, but the enclosed
-mass still quite well recovered.
-
 \clearpage

### 2013-09-10 text/tex/spaghetti.tex

@@
 
-In \spl the user begins by making an educated guess for the topography
-of the arrival-time surface --- specifically, the locations of the
-maxima, minima and saddle points that correspond to images of the
-brightest part of the source.  Input is given by tracing a proposal
-for the saddle-point contours.  Figure \ref{fig:input-spag} shows some
-example inputs --- and incidentally also indicates the origin of the
-name \spl.
+\spl implements the ideas of the arrival-time surface into a modelling
+program.
+
+The first modelling step with \spl is for the user to call up the \sw
+of interest, and make an educated guess for the topography of the
+arrival-time surface --- specifically, the locations of the maxima,
+minima and saddle points that correspond to images of the brightest
+part of the source.  Input is given by tracing a proposal for the
+saddle-point contours.  Figure \ref{fig:input-spag} shows some example
+inputs --- and incidentally also indicates the origin of the name
+\spl.
 
@@
 
-\spl provides plug ins to support any data source. At the moment, a data link to \sw and to the MasterLens database\needcite are implemented.
+In the second modelling step, once the user has sketched a proposed
+spaghetti configurations, \spl sends the input to its server-side
+modelling engine, called GLASS.  The task of GLASS is to find a mass
+distribution $\kappa(x,y)$ that exactly reproduces the locations of
+the maximum, minima and saddle points.  Now, this criterion by itself
+is extremely underdetermined --- there are infinitely many mass
+distributions that will reproduce a given set of maxima, minima and
+saddle points, but typically they (a)~produce lots of extra images,
+and (b)~look very unlike galaxies.  Additional assumptions (a prior)
+are necessary.  GLASS uses a prior based on suggestions by by
+\cite{1997MNRAS.292..148S}.
+\begin{enumerate}
+\item The mass distribution is built out of non-negative tiles of
+  mass.  (Sometimes these tiles are called mass pixels, but we should
+  emphasize that they are unrelated to image pixels, and are much
+  larger.)
+\item There is a notional lens centre, say $(x_0,y_0)$ which is
+  identified with the maximum of the arrival time.  The source can
+  have an arbitrary offset with respect to the lens centre.
+\item The mass distribution must be centrally concentrated, in two
+  respects.  First, the circularly averaged density must fall away
+  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
+  Second, the direction of increasing density at any $(x,y)$ can point
+  at most $45^\circ$ away from from $(x_0,y_0)$.
+\item The lens must symmetric with respect to $180^\circ$ rotations
+  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
+  user wishes.
+\end{enumerate}
+There are still infinitely many models that satisfy both data and
+prior constraints, but now they are more credible as galaxy lenses.
+It is then possible to generate an ensemble of models.  The sampling
+technique used by GLASS is described in \citep{Lubini2012}, and
+improves upon earlier techniques \citep{2000AJ....119..439W,Saha2004}.
+Typically, ensembles of 200 models are used.  That is to say, what we
+call a \spl model is really the mean of an ensemble of 200 models, and
+its estimated uncertainty is the range covered by the whole ensemble.
 
-\spl assists volunteers with several features in this process.
-Step 1 by supplying several images from several bands (not yet implemented in \sw).
+In the third modelling step, having generated a model ensemble, \spl
+postprocesses it to present results and diagnostics to the user for
+inspection.  This takes the form of three figures.
+\begin{enumerate}
+\item A grayscale plus contour map of the mass distribution.
+\item A contour map of the arrival-time surface.
+\item A synthetic image of the lensed features.
+\end{enumerate}
+After examining this feedback, the user can archive the results for
+discussion, or modify their input and try again, or discard the
+attempt altogether.
 
-Give a guess for the maximum, minimum and saddle points.  Program
-tries find $\kappa(x,y)$ that reproduces these properties exactly, and
-looks reasonably like a galaxy.  Solution not unique, an ensemble
-generated.
-SpaghettiLens is build on top of GLASS \citep{Lubini2012}, that builds and improves upon PixeLens \citep{Saha2004}.
-
-\spl then tries to find a mass distribution $\kappa(x,y)$ that reproduces the input parameters.
-Since there is no unique solution, \spl samples the solution space and produces an ensemble of solutions, as described in the paper \citep{Lubini2012}.
-
-Step 3 by providing the visual feedback. Users can check the generated mass distribution and synthetic image.
-
-It additinally provides plots of the modelled mass distribution and the contour lines of arrival time surface as feedback.
-
-It provides direct visual feedback by rendering a synthetic image side
-by side the original image, such that users can directly compare the
-predictions of their model to the survey image. Explanation: Synthetic
-image: A rendering of the derivative of the arrival time for each
-pixel. This leads to a black and white image that reasonably looks
-like the visual appearace of the model.
-
-Using this direct visual feedback, even inexperienced users can
-successfully model lenses using a iterative approach if they know what
-to look for.
-
-In the next step, several volunteers can work together / exchange their ideas and models and try to improve previous modelling attempts by others.
-This leads to a set of models organised in a branching, tree like structre for each model.
-Volunteers then organise them selves to narrow down the different branches and come up with a few consensus models in the end.
+The fourth modelling step is discussion among modellers and iteration
+on the model.  Any archived model can be revised by another user: they
+can modify the spaghetti configuration slightly or drastically, or
+change options like the size of the mass tiles.  Particularly
+interesting lens candidates lead to trees of models in this way.
+Forum discussion then prunes the tree, focussing attention on one or a
+few models.

### 2013-09-10 text/tex/results.tex
more edits

--- 2013-09-09 text/tex/results.tex
@@
-\section{The lens modeling challenge} \label{sec:results}
+\section{A lens modeling challenge} \label{sec:results}
 
-\subsection{\sw} \label{sec:SpaceWarps}
-some words about \sw
+% \subsection{\sw} \label{sec:SpaceWarps} some words about \sw
+
+Interested volunteers from the \sw forum were initially introduced to
+\spl through a video tutorial and by videocon.  After this
+introductory stage, a modelling challenge was presented.  This
+consisted of 29 simulated lenses (sims) covering a range of lensing
+configurations.
 
@@
 
-In order to estimate users performance, \sw included some simulated lensing configurations in their image set.
+The sims were generated by AM, in consultation with PM and AV.  In the
+interest of blind testing, the information in this section was
+revealed neither to RK, while choosing the challenge set, nor to the
+people making models (EB, CC, CM, JO, PS and JW) until they were done.
 
-There are three kinds of simulated lenses.
-
-Sims produced using {\tt gravlens}
-\citep{2001astro.ph..2341K,2001astro.ph..2340K}.
+The sims produced using {\tt gravlens}
+\citep{2001astro.ph..2341K,2001astro.ph..2340K} and were of three
+kinds, as follows.
 
@@
 
+\subsection{Some example models}
+
+Figure \ref{fig:6941} shows a simple example.
+
+Figure \ref{fig:6975} shows an example where substructure introduces a
+complication, but model ok.
+
+Figure \ref{fig:6937} shows a case where substructure leads to a
+poorer model.
+
+Figure \ref{fig:6975} shows a nice symmetric quad.
+
+Figure \ref{fig:6990} shows a long-axis quad.
+
+Figure \ref{fig:6919} shows a short-axis quad.
+
+Figure \ref{fig:6915} shows an inclined quad.
+
+Figure \ref{fig:6971} shows incorrect identification, but the enclosed
+mass still quite well recovered.
+
 \subsection{The test setup} \label{sec:testsetup}
-
-We made use of the simulated lenses (sims) to test the volunteers abilities to model those sims.
-
-The testing was done with a small number of volunteers ($N_\text{v}=4$) and a set of simulated configurations ($N_\text{sim}=29$) chosen to represent some typical configurations.
-The volunteers were given an introduction video explaining the theory behind gravitational lensing and the usage of \spl.
-Additionally the volunteers had the opportunity to pose general questions to the experts during several hangouts during the training phase that preceded the modeling challenge.\footnote{No questions about the to be modeled sims were answered.}
 
@@
              {fig/sims/ASW0001hpf/arriv.png}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/ASW0001hpf/extr_points.png}}
+%  \subfigure{\includegraphics[width=0.45\textwidth]
+%             {fig/sims/ASW0001hpf/extr_points.png}}
   \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom left), additionally using a extremal point detector (bottom right).

### 2013-09-10 text/tex/results.tex

--- 2013-09-10 text/tex/results.tex
@@
 \subsection{The test setup} \label{sec:testsetup}
-
-We made use of the simulated lenses (sims) to test the volunteers abilities to model those sims.
-
-The testing was done with a small number of volunteers ($N_\text{v}=4$) and a set of simulated configurations ($N_\text{sim}=29$) chosen to represent some typical configurations.
-The volunteers were given an introduction video explaining the theory behind gravitational lensing and the usage of \spl.
-Additionally the volunteers had the opportunity to pose general questions to the experts during several video conferences during the training phase that preceded the modeling challenge.\footnote{No questions about the to be modeled sims were answered.}
 

### 2013-09-10 text/tex/results.tex
results

--- 2013-09-10 text/tex/results.tex
@@
 
+Two additional sims were modeled by an expert, \asw{1hpf} and \asw{0vqg}.
+Looking at the results for \gEr for those models in \figref{eR_per_sim}, we conclude that the performance of volunteers and experts is comparable.
+Please note that the models of \asw{0vqg} with \gEr{, rel} around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.
+
 \clearpage

### 2013-09-10 text/tex/results.tex
hilight experst model in result plot

@@
     \includegraphics[width=0.80\textwidth]{fig/sims/eR_4.png}
-  \caption{relative Einstein Radius for each model, per sim. }
+  \caption{relative Einstein Radius for each model, per sim. red crosses for models made by an expert.}
   \label{fig:eR_per_sim}
@@
 Two additional sims were modeled by an expert, \asw{1hpf} and \asw{0vqg}.
-Looking at the results for \gEr for those models in \figref{eR_per_sim}, we conclude that the performance of volunteers and experts is comparable.
+Looking at the results for \gEr for those models in \figref{eR_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses) is comparable.
 Please note that the models of \asw{0vqg} with \gEr{, rel} around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.

### 2013-09-10 plots/gen_plots.py

@@ -929,7 +929,8 @@ def plotAllRE():
         lu[simn]
       except KeyError:
         continue
-      pl.plot(lu[simn], dat['rE_mean']/sims[simn], 'bx')
+      style = 'rx' if dat['user']=='psaha' else 'bx'
+      pl.plot(lu[simn], dat['rE_mean']/sims[simn], style)
       
     lbls = [key for key, val in spg.sims.items()]
     for i, item in enumerate(spg.sims.items()):

### 2013-09-10 text/tex/results.tex
added note

@@
 Looking at the results for \gEr for those models in \figref{eR_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses) is comparable.
-Please note that the models of \asw{0vqg} with \gEr{, rel} around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.
+Please note that the models of \asw{0vqg} with \gEr{, rel} around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
 

### 2013-09-10 plots/gen_plots.py
improved plot

@@ -830,6 +830,7 @@ def plotAllRE():
   print '> drawing EinsteinR plots',
 
   plots = [True, True, True, True]
+  #plots = [0, 0, 0, True]
   
   path = resdir
 
@@ -930,7 +931,8 @@ def plotAllRE():
       except KeyError:
         continue
       style = 'rx' if dat['user']=='psaha' else 'bx'
-      pl.plot(lu[simn], dat['rE_mean']/sims[simn], style)
+      xofs = 0.15 if dat['user']=='psaha' else -0.15
+      pl.plot(lu[simn]+xofs, dat['rE_mean']/sims[simn], style)
       
     lbls = [key for key, val in spg.sims.items()]
     for i, item in enumerate(spg.sims.items()):

### 2013-09-10 plots/gen_plots.py
better table

@@ -37,11 +37,15 @@ available commands are:
   creates all the tex files and copies the relevant files
   to the tex folder
 
+by just calling the script from command line:
+  python gen_plots
+the run() function gets called automatically
+
+
 
 data is stored in:
 - many.sim
 - spg.data
-
 
 
 
@@ -951,3 +955,9 @@ def plotAllRE():
     #pl.show()
     
   print ' ... DONE'
+
+
+#if __name__ == "__main__":
+#  run()
+  
+  

### 2013-09-10 plots/gen_table.py

@@ -34,258 +34,327 @@ out_file_path = '../text/tab/auto'
 inpfile = os.path.join(os.path.abspath(inp_file_path),inp_file_name)
 outfile = os.path.join(os.path.abspath(out_file_path),out_file_name)
 
-try:
-  os.makedirs(os.path.abspath(out_file_path))
-except OSError as e:
-  print 'error creating paths', e
-  print 'continuing'
   
 
 all_data = {}
 all_stats = []
 
-with open(inpfile, 'r') as f:
-  csvr = csv.reader(f)
+
+def run():
+  read()
+  createDirs()
+  genDetailTables()
+  genStatsTable()
+
+
+def createDirs():
+  try:
+    os.makedirs(os.path.abspath(out_file_path))
+  except OSError as e:
+    print 'error creating paths', e
+    print 'continuing'
+
+
+#==============================================================================
+# READS DATA FROM CSV TABLE
+#==============================================================================
+def read():
+  with open(inpfile, 'r') as f:
+    csvr = csv.reader(f)
+    
+    simname = ''
   
-  simname = ''
-
-  for i, row in enumerate(csvr):
-    #print row
-
-    # get the stats from the end of the file
-    if i in [156,159,162,165]:
-      
-      all_stats.append({
-        'n'          : int(row[ 5]),
-        'npnt'       : int(row[ 7]),
-        'approxPlace': int(row[ 8]),
-        'approx'     : int(row[ 9]),
-        'veryRight'  : int(row[10]),
-        'rightPlace' : int(row[12]),
-        'rightType'  : int(row[13]),
-        'rightOrder' : int(row[14]),
-        'err'        : [int(_) for _ in row[16:26]]
-      })
-
-    # skip if not in data rows  
-    if i<2 or i>155 or (row[0]=='' and row[1]==''):
-      continue
-    
-    
-    if row[0]!='': # if header row containing sim name
-      #print row[0]
-      simname=row[0]
-      all_data[simname]=[]
-      
-    else: # regular data
-      if simname=='ASW0001gae': # we don't like this sim
+    for i, row in enumerate(csvr):
+      #print row
+  
+      # get the stats from the end of the file
+      if i in [156,159,162,165]:
+        
+        all_stats.append({
+          'n'          : int(row[ 5]),
+          'npnt'       : int(row[ 7]),
+          'approxPlace': int(row[ 8]),
[…]

### 2013-09-10 text/tex/results.tex
minor ajustments to plots and tables

@@
 %             {fig/sims/ASW0001hpf/extr_points.png}}
-  \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom left), additionally using a extremal point detector (bottom right).}
+  \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom).}
   \label{fig:output_compare}
@@
 The complete table with the results for each model can be found in the appendix, \tabref{detail_results}.
-
-In \tabref{stats} a summary of this evaluation is presented.
 
@@
 
-We conclude that the volunteers are performing very well identifying and positioning images (T1a), with a performance of 92\% (approxPlace p=0.92).
-Most of the problems where due to unclear arc-like structures (Error 1, p=0.18; Error 4, p=0.03; Error 8, p=0.04).
-Critical failures like the failure to identify all fife images in an fife image system (Error 3, p=0.04) or to include too many images (Error 10, p=0.01) did almost never happened.
+In \tabref{stats} a summary of this evaluation is presented.
+We conclude that the volunteers are performing very well identifying and positioning images (T1a), with a performance of 92\% (R1 p=0.92).
+Most of the problems where due to unclear arc-like structures (E01, p=0.18; E04, p=0.03; E08, p=0.04).
+Critical failures like the failure to identify all five images in an five image system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happened.
 From this we conclude that the introduction materials was adequate and the volunteers understand the basics of gravitational lensing.
@@
 The assignment of the parity of the images (T1b) posed a more difficult task.
-In 59\% (rightOrder, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
-Most of the failures are due to Error 6 (PI/2), with N=38, p=32\%, followed by Error 5 (PI) with N=7, p=6\%.
-For an example of type 6 error, see \figref{6971}.
+In 59\% (R2, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
+Most of the failures are due to E06, with N=38, p=32\%, followed by E05 (PI) with N=7, p=6\%.
+For an example of E06, see \figref{6971}, it basically describes a situation, where the minima and saddlepoints were exchanged (rotated by $\pm90\dgr$).
+E05 describes the situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$)
 While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
@@
 To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] directly generated in the modeling process.
-From the ensemble of models generated by one modelling process, the mean is taken as the resulting $\kappa(r)$ to calculate $\theta_E$.
+From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate $\theta_E$.
 To estimate the errors, also the extremal models are used to estimate a lower and upper limit for $\theta_E$.
 These results can be seen in \figref{eR_all_models}.
-This figure shows that this tecnique of esitmating the error is underestimating the error significantly and should be improved for further analysis.
+This figure shows that this technique of estimating the error is underestimating the error significantly and should be improved for further analysis.
 
@@
     \includegraphics[width=0.80\textwidth]{fig/sims/eR_1.png}
-  \caption{\tEr \gEr for all models with estimated erros in blue, \gEr of simulation in red}
+  \caption{\tEr \gEr for all models with estimated errors in blue squares, \gEr of simulation in red crosses}
   \label{fig:eR_all_models}
@@
     \includegraphics[width=0.80\textwidth]{fig/sims/eR_4.png}
-  \caption{relative Einstein Radius for each model, per sim. red crosses for models made by an expert.}
+  \caption{relative Einstein Radius \gEr / \gEr[, sim] for each model, per sim. red crosses with right offset for models made by an expert.}
   \label{fig:eR_per_sim}
@@
 Two additional sims were modeled by an expert, \asw{1hpf} and \asw{0vqg}.
-Looking at the results for \gEr for those models in \figref{eR_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses) is comparable.
+Looking at the results for \gEr for those models in \figref{eR_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses, offset) is comparable.
 Please note that the models of \asw{0vqg} with \gEr{, rel} around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}

### 2013-09-10 plots/gen_plots.py

@@ -30,7 +30,7 @@ available commands are:
   
 - all_mod_plots()
   creates all the results / modells plots
-  (server side created plots get grapped from internet)
+  (server side created plots get grabbed from internet)
   
 - all_tex()
   assumes the plots are already created (under /plots)
@@ -884,6 +884,9 @@ def plotAllRE():
     pl.figure()
     pl.errorbar(xi, re, ee, marker='s', mfc='blue', ls='' ,ecolor='blue')
     pl.plot(xi, se, 'rx')
+    pl.xlim([np.min(xi)-1, np.max(xi)+1])
+    pl.xlabel('model id')
+    pl.ylabel(r'Einstrin radius $\Theta_\text{E}$ [pixel]')
     
     pl.savefig(os.path.join(path, 'eR_1.png'))
     print '.',

### 2013-09-10 plots/gen_table.py

@@ -280,23 +280,23 @@ def genStatsTable():
   lue = {
     'n'          : (00, 'n points'),
     'npnt'       : (00, 'correct amount of images'),
-    'approxPlace': (01, 'images approx. on right location'),
+    'approxPlace': (01, 'R1: images approx. on right location'),
     'approx'     : (00, 'setup approx. correct'),
     'veryRight'  : (00, 'setup correct'),
     'rightPlace' : (00, 'fraction of images on right place'),
     'rightType'  : (00, 'fraction of images right type'),
-    'rightOrder' : (02, 'images with correct parity'),
+    'rightOrder' : (02, 'R2: images with correct parity'),
     'err'        : (10, r'\hline', (
-      (20 , 'Error 01'),
-      (21 , 'Error 02'),
-      (22 , 'Error 03'),
-      (23 , 'Error 04'),
-      (24 , 'Error 05'),
-      (25 , 'Error 06'),
-      (26 , 'Error 07'),
-      (27 , 'Error 08'),
-      (28 , 'Error 09'),
-      (29 , 'Error 10'),
+      (20 , 'E01: inaccurate in arc'),
+      (21 , 'E02: wrong parity in 3 lens conf.'),
+      (22 , 'E03: identified 3 of 5 imgs.'),
+      (23 , 'E04: modeled arc with single img.'),
+      (24 , 'E05: wrong order of sad. points'),
+      (25 , 'E06: $\pi$/2 rotated parity'),
+      (26 , 'E07: missed faint img.'),
+      (27 , 'E08: too many imgs in arc.'),
+      (28 , 'E09: missed double img'),
+      (29 , 'E10: too many imgs.'),
     )),
   }  
   
@@ -326,7 +326,7 @@ def genStatsTable():
 
   entries.sort()
   
-  for e in entries: print e
+  #for e in entries: print e
   
   tex_stats += '\n'.join([e[1] for e in entries if not e[0]==0])
   tex_stats += '\n'

### 2013-09-10 text/tex/outlook.tex
more edits
labels

@@
 
-As seen in \secref{tests.1}, volunteers fail mostly in two situations:
+As seen in \secref{tests.t1}, volunteers fail mostly in two situations:
 when to identify an arclike structre while placing the points
@@
 Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models ect...)
+
+Lenses with more than one maximum can be allowed for.  See Figure 5c
+in \citep{2001ApJ...557..594R} and Figure 4b in
+\cite{2003ApJ...590...39K}.
+
+Supply several images from several bands.

### 2013-09-10 text/tex/results.tex
labels
some figure changes

@@
 
+Figure \ref{fig:7022} shows incorrect identification, but the enclosed
+mass still quite well recovered.
+
 Figure \ref{fig:6975} shows a nice symmetric quad.
@@
 Figure \ref{fig:6915} shows an inclined quad.
-
-Figure \ref{fig:6971} shows incorrect identification, but the enclosed
-mass still quite well recovered.
 
@@
 Most of the failures are due to E06, with N=38, p=32\%, followed by E05 (PI) with N=7, p=6\%.
-For an example of E06, see \figref{6971}, it basically describes a situation, where the minima and saddlepoints were exchanged (rotated by $\pm90\dgr$).
+For an example of E06, see \figref{7022}, it basically describes a situation, where the minima and saddlepoints were exchanged (rotated by $\pm90\dgr$).
 E05 describes the situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$)
 While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
-For more challenging cases, like very symmetrical distribution of the lensed images (for example model 6971, \figref{6971}), those errors should still produce plausible results, as will be explained in the next section.
+For more challenging cases, like very symmetrical distribution of the lensed images (for example model 7022, \figref{7022}), those errors should still produce plausible results, as will be explained in the next section.
 
-\subsection{Second test T2} \label{sec:results.2}
+\subsection{Second test T2} \label{sec:tests.t2}
 

### 2013-09-10 text/tex/abstract.tex
a few words on sl, added bib entries
abstract edited

@@
 were modelled, each multiple times.  The quality of the models was
-then examined in two ways.  First, the image parities and time
-orderings: these were identified correctly in most cases, but not
-always.  Second, the mean convergence (equivalent to the enclosed
-mass): this was constrained consistently well in the image region.
-The performance of professionals and experienced volunteers was
-comparable.
+then examined in different ways.  In particular, (i)~the image
+parities and time orderings were identified correctly in most cases,
+but not always; (ii)~the mean convergence (equivalent to the enclosed
+mass) was quite well constrained in the image region; (iii)~the
+performance of experienced volunteers was comparable to professionals.
 \end{abstract}

### 2013-09-11 text/tex/appendix.tex
small fixes on text, tables, plots

@@
 
-In Section \ref{sec:Fermat}, for the sake of a more intuitive
+In \secref{Fermat}, for the sake of a more intuitive
 explanation, we suppressed some constant factors in

### 2013-09-11 text/tex/results.tex
small fixes on text, tables, plots

@@
-\section{A lens modeling challenge} \label{sec:results}
+\section{A lens modeling challenge} \label{sec:mod_challenge}
 
@@
 
-\subsection{Some example models}
+\subsection{Some example models} \label{sec:example_models}
 
-Figure \ref{fig:6941} shows a simple example.
+\Figref{6941} shows a simple example.
 
-Figure \ref{fig:6975} shows an example where substructure introduces a
+\Figref{6975} shows an example where substructure introduces a
 complication, but model ok.
 
-Figure \ref{fig:6937} shows a case where substructure leads to a
+\Figref{6937} shows a case where substructure leads to a
 poorer model.
 
-Figure \ref{fig:7022} shows incorrect identification, but the enclosed
+\Figref{7022} shows incorrect identification, but the enclosed
 mass still quite well recovered.
 
-Figure \ref{fig:6975} shows a nice symmetric quad.
+\Figref{6975} shows a nice symmetric quad.
 
-Figure \ref{fig:6990} shows a long-axis quad.
+\Figref{6990} shows a long-axis quad.
 
-Figure \ref{fig:6919} shows a short-axis quad.
+\Figref{6919} shows a short-axis quad.
 
-Figure \ref{fig:6915} shows an inclined quad.
+\Figref{6915} shows an inclined quad.
 
@@
 To estimate the performance of the volunteers and the quality of the generated models, two test T1 and T2 were done.
-%\begin{enumerate}
-%  \item (T1) Correct identification of extremal points
-%  \item (T2) Reproduction of the mass distribution of the lens
-%\end{enumerate}
 
@@
 
-While we expected T1a to be trivial, given the nature of the survey images and the success of \sw, we expect T1b to pose more problems.
+While we expected T1a to be trivial, given the nature of the survey images and the success of \sw, we expect T1b to be more difficult.
 T1b tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
-While we can provide the volunteers with some general rule of thumbs, T1b involves imagination and guessing and therefore training could improve their skills in a later stage.
+While we can provide the volunteers with some general rules of thumb, T1b involves imagination and guessing and therefore training could improve their skills in a later stage.
 
@@
 The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between the sims and the models.
-We also let an expert model two selected systems to compare the results from volunteers to those of a professional.
+We also let an expert model three selected systems to compare the results from volunteers to those of a professional.
 
@@
 
-The evaluation of the volunteers performance for T1 was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot genreated using simulation parameters.
+The evaluation of the volunteers performance for T1 was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot generated using simulation parameters.
 \Figref{output_compare} shows the setup used to evaluate the models.
@@
 
-Additionally, ten types of errors that could occur were analyzed, where each model could contain multiple of those.
+Additionally, ten types of errors (E01 -- E10, listed in \tabref{stats}) that could occur were analyzed, where each model could contain multiple of those.
 The complete table with the results for each model can be found in the appendix, \tabref{detail_results}.
@@
 In \tabref{stats} a summary of this evaluation is presented.
-We conclude that the volunteers are performing very well identifying and positioning images (T1a), with a performance of 92\% (R1 p=0.92).
+We conclude that the volunteers are performing very well identifying and positioning images (T1a), with a performance of 92\% (R1, p=0.92).
 Most of the problems where due to unclear arc-like structures (E01, p=0.18; E04, p=0.03; E08, p=0.04).
-Critical failures like the failure to identify all five images in an five image system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happened.
+Critical errors like the failure to identify all five images in an five image system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happened.
 From this we conclude that the introduction materials was adequate and the volunteers understand the basics of gravitational lensing.
@@
 
-The assignment of the parity of the images (T1b) posed a more difficult task.
+The assignment of the parity of the images (T1b) was a more difficult task.
 In 59\% (R2, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
 Most of the failures are due to E06, with N=38, p=32\%, followed by E05 (PI) with N=7, p=6\%.
-For an example of E06, see \figref{7022}, it basically describes a situation, where the minima and saddlepoints were exchanged (rotated by $\pm90\dgr$).
-E05 describes the situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$)
+For an example of E06, see \figref{7022}.
+It basically describes a situation, where the minima and saddle points of a five image configuration were exchanged (rotated by $\pm90\dgr$).
+E05 describes the situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$).
 While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
@@
 To estimate the errors, also the extremal models are used to estimate a lower and upper limit for $\theta_E$.
-These results can be seen in \figref{eR_all_models}.
+These results can be seen in \figref{ER_all_models}.
 This figure shows that this technique of estimating the error is underestimating the error significantly and should be improved for further analysis.
@@
     \includegraphics[width=0.80\textwidth]{fig/sims/eR_1.png}
-  \caption{\tEr \gEr for all models with estimated errors in blue squares, \gEr of simulation in red crosses}
-  \label{fig:eR_all_models}
+  \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
+  \label{fig:ER_all_models}
 \end{figure}
 
-In \figref{eR_per_sim} can be seen, that the calculated \tEr \gEr of the models tend to be too high.
+In \figref{ER_per_sim} can be seen, that the calculated \ERf of the models tend to be too high.
 The overshoot varies from around 0.2 to 0.4 for for good models.
@@
     \includegraphics[width=0.80\textwidth]{fig/sims/eR_4.png}
-  \caption{relative Einstein Radius \gEr / \gEr[, sim] for each model, per sim. red crosses with right offset for models made by an expert.}
-  \label{fig:eR_per_sim}
+  \caption{relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
+  \label{fig:ER_per_sim}
 \end{figure}
 
-Error 5 and 6 happen mostly in very symmetric images, where it's hard to tell
-It changes the orientation of the mass distribution \kap[x,y].
-But \kenc and thus the \tEr \gEr is not changed, and thus the final result is still a valid model.
+E05 and E06 happen mostly in very symmetric images, that are hard to come up with a unique valid solution.
+This errors change the orientation of the mass distribution \kap[x,y],
+but \kenc and thus the \ERf is not influence that much, and thus the final result is still a valid model.
 This can be seen by comparing the results for \asw{0h2m}, shown in \figref{kapenc_compare_faulty}: 
-The correct model 7022 ($\mEr=10.76$), which was done by an expert, and
-7020 ($\mEr=10.72$),
-7024 ($\mEr=10.80$),
-7025 ($\mEr=11.16$),
-7021 ($\mEr=11.04$).
-The first three show error 6, while the last is of error type 5.
+The correct model 7022 ($\ERm=10.76px$), which was done by an expert, and
+7020 ($\ERm=10.72px$),
+7024 ($\ERm=10.80px$),
+7025 ($\ERm=11.16px$),
+7021 ($\ERm=11.04px$).
+The first three are of type E06, while the last is of type E05.
 This can be easily corrected, if further analysis is done for the modeled system and time delays are known.
@@
              {fig/sims/007022/kappa_encl}}
-  \caption{\kenc for models of \asw{0vqg}: 7020 (top left), 7021 (top right), 7024 (mid left), 7025 (mid right) by volunteers and a correct model 7022 (bottom) by an expert.}
+  \caption{\kenc for models of \asw{0h2m}: 7020 (top left), 7021 (top right), 7024 (mid left), 7025 (mid right) by volunteers and a correct model 7022 (bottom) by an expert.}
   \label{fig:kapenc_compare_faulty}
@@
 Two additional sims were modeled by an expert, \asw{1hpf} and \asw{0vqg}.
-Looking at the results for \gEr for those models in \figref{eR_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses, offset) is comparable.
-Please note that the models of \asw{0vqg} with \gEr{, rel} around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
+Looking at the results for \ERg for those models in \figref{ER_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses, offset) is comparable.
+Note that the models of \asw{0vqg} with \ERg[, rel] around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
 

### 2013-09-11 text/tex/spaghetti.tex
more edits
small fixes on text, tables, plots

@@
     later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
-  \label{fig:input-spag}}
+  }
+  \label{fig:input-spag}
 \end{figure}
@@
 
-\spl implements the ideas of the arrival-time surface into a modelling
+\spl implements the ideas of the arrival-time surface into a modeling
 program.
 
-The first modelling step with \spl is for the user to call up the \sw
+The first modeling step with \spl is for the user to call up the \sw
 of interest, and make an educated guess for the topography of the
@@
 part of the source.  Input is given by tracing a proposal for the
-saddle-point contours.  Figure \ref{fig:input-spag} shows some example
+saddle-point contours.  \Figref{input-spag} shows some example
 inputs --- and incidentally also indicates the origin of the name
@@
 self-intersection of the contour.  The top panel of
-Figure~\ref{fig:input-spag} shows an example: red marks a suggested
+\figref{input-spag} shows an example: red marks a suggested
 maximum, green a saddle point, and blue a minimum.  (The small pink
 dots are just help sketch the proposed contour.)  The middle and lower
-panels of Figure~\ref{fig:input-spag} show a more complex situation.
+panels of \figref{input-spag} show a more complex situation.
 The minimum has split into three images: two new minima with a saddle
@@
 The exact placement of the pink spaghetti loops shown in
-Figure~\ref{fig:input-spag} has no significance.  Their utility is
+\Figref{input-spag} has no significance.  Their utility is
 just to help mark up the image as maximum, minima and saddle points.
@@
 and (b)~look very unlike galaxies.  Additional assumptions (a prior)
-are necessary.  GLASS uses a prior based on suggestions by by
+are necessary.  GLASS uses a prior based on suggestions by
 \cite{1997MNRAS.292..148S}.

### 2013-09-11 text/tex/theory.tex
small fixes on text, tables, plots

@@
 \begin{figure}
+\centering
 \includegraphics[height=.3\vsize]{fig/arriv1.png}
@@
   is the result of an elongated lensing mass; a maximum, two minima
-  and two saddle points can be seen. \label{fig:arriv}}
+  and two saddle points can be seen.
+}
+\label{fig:arriv}
 \end{figure}
@@
 the Universe into account, and is given in the Appendix.  The top part
-of Figure~\ref{fig:arriv} shows what $A_{\rm geom}$ looks like: it is
+of \figref{arriv} shows what $A_{\rm geom}$ looks like: it is
 just a paraboloid.  The minimum of the surface is at the source, and
@@
 any more.  The result is illustrated in the middle part of
-Figure~\ref{fig:arriv}.  In addition to the maximum, there is now a
+\figref{arriv}.  In addition to the maximum, there is now a
 minimum and a saddle point.  According to Fermat's principle, images
@@
 distribution is non-circular, more images can appear.  The lower
-example in Figure~\ref{fig:arriv} shows the arrival-time surface due
+example in \figref{arriv} shows the arrival-time surface due
 to an elongated mass.  There is still a maximum, part of an elongated
@@
 arrival-time surface is nearly flat.  As is evident from
-Figure~\ref{fig:arriv}, this is the region where the minima and saddle
+\figref{arriv}, this is the region where the minima and saddle
 points lie, or near the images.  In this region, points on the source
@@
 arrival-time surface is strongly curved, the image will be
-demagnified.  We see from Figure~\ref{fig:arriv} that the arrival-time
+demagnified.  We see from \figref{arriv} that the arrival-time
 surface tends to be highly curved near the maximum.  Hence maxima tend
@@
 
-In the lensed examples in Figure~\ref{fig:arriv}, the maximum in each
+In the lensed examples in \figref{arriv}, the maximum in each
 case is evident, but how can we distinguish the minima from the saddle
@@
 pull away.  The contour curve precisely level with a saddle point
-forms an X at the saddle point.  In Figure~\ref{fig:arriv}, one such X
+forms an X at the saddle point.  In \figref{arriv}, one such X
 is evident.  In other cases the X contour passing through the saddle

### 2013-09-11 plots/gen_plots.py

@@ -836,6 +836,10 @@ def plotAllRE():
   plots = [True, True, True, True]
   #plots = [0, 0, 0, True]
   
+  import gen_table as tab
+
+  tab.read()  
+  
   path = resdir
 
   spg.genREData() # genreate the data
@@ -856,6 +860,7 @@ def plotAllRE():
   ep_rel = []
   em_rel = []
   se = []
+  ac = [] # accepted and not rejected?
   for dat in spg.data:
     try:
       re_sim = sims[dat['name']]
@@ -869,6 +874,8 @@ def plotAllRE():
     re_rel.append(dat['rE_mean']/re_sim)
     ep_rel.append(dat['rE_max']/re_sim)
     em_rel.append(dat['rE_min']/re_sim)
+
+    ac.append(tab.all_mods[str(dat['id'])]['acc'])
     
   xi = np.array(xi)
   re = np.array(re)
@@ -879,12 +886,18 @@ def plotAllRE():
   ep_rel = np.array(ep_rel) - re_rel
   em_rel = re_rel - np.array(em_rel)
   ee_rel = [ep_rel, em_rel]
+
+  ac = np.array(ac, dtype=bool)
   
   if plots[0]:
     pl.figure()
-    pl.errorbar(xi, re, ee, marker='s', mfc='blue', ls='' ,ecolor='blue')
+    #paint accepted
+    pl.errorbar(xi[ac], re[ac], [ep[ac], em[ac]], marker='s', mfc='blue', ls='' ,ecolor='blue')
+    #paint rejected
+    pl.errorbar(xi[~ac], re[~ac], [ep[~ac], em[~ac]], marker='o', mfc='green', ls='' ,ecolor='green')
     pl.plot(xi, se, 'rx')
-    pl.xlim([np.min(xi)-1, np.max(xi)+1])
+    pl.xlim([np.min(xi), np.max(xi)])
+    pl.ylim([0., 35.])
     pl.xlabel('model id')
     pl.ylabel(r'Einstrin radius $\Theta_\text{E}$ [pixel]')
     
@@ -937,7 +950,12 @@ def plotAllRE():
         lu[simn]
       except KeyError:
         continue
-      style = 'rx' if dat['user']=='psaha' else 'bx'
+      if dat['user']=='psaha':
+        style = 'rx'
+      elif not tab.all_mods[str(dat['id'])]['acc']:
+        style = 'g+'
+      else:
[…]

### 2013-09-11 plots/gen_table.py

@@ -37,6 +37,7 @@ outfile = os.path.join(os.path.abspath(out_file_path),out_file_name)
   
 
 all_data = {}
+all_mods = {}
 all_stats = []
 
 
@@ -93,7 +94,8 @@ def read():
         all_data[simname]=[]
         
       else: # regular data
-        if simname=='ASW0001gae': # we don't like this sim
+        if simname=='ASW0001gae' or row[28]!='1': # we don't like this sim (all flag not set)
+          #print 'skipping'          
           continue
   
         # recover the fractions in row 12 and 13 as such..
@@ -117,8 +119,9 @@ def read():
           'rightOrder' : t if row[14]=='1' else f,
           'errors'     : '',
           'use'        : True if row[29]=='1' else False,
-          'fav1'       : True if row[30]=='1' else False,
-          'fav2'       : True if row[31]=='1' else False,
+          'acc'        : True if row[30]=='1' else False,
+          'fav1'       : True if row[31]=='1' else False,
+          'fav2'       : True if row[32]=='1' else False,
         }
         
         # parse the errors into a string of error number
@@ -130,7 +133,9 @@ def read():
             s.append(str(j+1))
         er = ','.join(s) #final string ,-sep ids
         data['errors'] = er
+        data['errors2'] = a # in form of bool array
         
+        all_mods[data['id']] = data
         all_data[simname].append(data)
 
 
@@ -278,55 +283,59 @@ def genStatsTable():
   # lookUptable for entries
   # which lines display in what order (order, 'descr'), 0=skip
   lue = {
-    'n'          : (00, 'n points'),
-    'npnt'       : (00, 'correct amount of images'),
-    'approxPlace': (01, 'R1: images approx. on right location'),
-    'approx'     : (00, 'setup approx. correct'),
-    'veryRight'  : (00, 'setup correct'),
-    'rightPlace' : (00, 'fraction of images on right place'),
-    'rightType'  : (00, 'fraction of images right type'),
-    'rightOrder' : (02, 'R2: images with correct parity'),
-    'err'        : (10, r'\hline', (
-      (20 , 'E01: inaccurate in arc'),
[…]

### 2013-11-12 text/tex/results.tex
rearranged figure placement

@@
 
-\subsection{First test T1} \label{sec:tests.t1}
+\subsection{Test of image identification} \label{sec:tests.t1}
 
@@
 \Figref{output_compare} shows the setup used to evaluate the models.
-
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/006917/input.png}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/006917/spaghetti.png}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/ASW0001hpf/arriv.png}}
-%  \subfigure{\includegraphics[width=0.45\textwidth]
-%             {fig/sims/ASW0001hpf/extr_points.png}}
-  \caption{Examples of Spaghetti input (left top) and modeled arrival time surface (right top) vs reconstructed arrival time surface from simulation parameters (bottom).}
-  \label{fig:output_compare}
-\end{figure}
 
@@
 
-\input{tab/auto/_stats}
-
 In \tabref{stats} a summary of this evaluation is presented.
@@
 
-\subsection{Second test T2} \label{sec:tests.t2}
+\subsection{Test of mass-profile recovery} \label{sec:tests.t2}
 
@@
 
-\begin{figure}[htbp]
-  \centering
-    \includegraphics[width=0.80\textwidth]{fig/sims/eR_1.png}
-  \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
-  \label{fig:ER_all_models}
-\end{figure}
-
 In \figref{ER_per_sim} can be seen, that the calculated \ERf of the models tend to be too high.
@@
 An offset leads to a a flatter mass profile for the model compared to the simulation.
-
-\begin{figure}[htbp]
-  \centering
-    \includegraphics[width=0.80\textwidth]{fig/sims/eR_4.png}
-  \caption{relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
-  \label{fig:ER_per_sim}
-\end{figure}
 
@@
 
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/007020/kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/007021/kappa_encl}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/007024/kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/007025/kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/007022/kappa_encl}}
-  \caption{\kenc for models of \asw{0h2m}: 7020 (top left), 7021 (top right), 7024 (mid left), 7025 (mid right) by volunteers and a correct model 7022 (bottom) by an expert.}
-  \label{fig:kapenc_compare_faulty}
-\end{figure}
-
 Comparing the models of volunteers and experts can be done in \figref{kapenc_compare_faulty}, where only the

### 2013-11-12 text/tex/theory.tex
rearranged figure placement

@@
 observed, but several observable quantities can be derived from it.
-
-\begin{figure}
-\centering
-\includegraphics[height=.3\vsize]{fig/arriv1.png}
-\includegraphics[height=.3\vsize]{fig/arriv2.png}
-\includegraphics[height=.3\vsize]{fig/arriv3.png}
-\caption{Arrival-time surfaces, with contour levels of equal arrival
-  time.  The upper surface is with no lens.  The middle surface is
-  when a circular lensing mass (offset from the source) is added; a
-  maximum, a minimum and a saddle point can be seen.  The last surface
-  is the result of an elongated lensing mass; a maximum, two minima
-  and two saddle points can be seen.
-}
-\label{fig:arriv}
-\end{figure}
 

### 2013-11-16 text/tex/intro.tex
small fixes on text, tables, plots
workshop corrections

--- 2013-08-26 text/tex/intro.tex
@@
 
-Spacewarps is the future.
+Gravitational lenses (GLs), predicted by General Relativity (GR), can be used as a tool by astronomers to examine many properties of the cosmos.
+They allow for example to estimate the masses and their distribution of galaxies, automatically including the hard to grasp dark matter.
+Further more, they allow an estimate on cosmological parameters like the Hubble constant \citep{Saha2006} and the mass density \needcite.
 
-Moreover, that lens modelling is an essential accompaniment to lens
-searches has been clear since the early days \citep{1981ApJ...244..723Y}.
+%To accheve this, one needs to find / identify lenses and in a second step model those lenses to get involved parameters.
 
-Why Fermat's principle?  Details in \ref{sec:Fermat} and
-\ref{sec:more-theory}.
+How to find gravitational lenses? Huge amount of image data from surveys that needs to be processed.
+Even more coming in the future with new surveys \needcite
+Robotic procession has been suggested \needcite, and tested \needcite, but has failed so far to be convincing \needcite.
+Another suggestion involves humans, volountiers \footnote{A volounteer is considered everybody that has no background in astro physics.}.
+\sw uses this approach with great succes so far.\needcite
 
-SpaghettiLens details in \ref{sec:SpaghettiLens}.
+Next step involves modeling. that needs advanced knolegde and takes a lot of time.
+Too much time to be done by astronomers them selves as the results from new surveys and identifiers like \sw come in.
+So there is a demand of a means to modell a great amount of identified lens data, that scales with increasing data amount.
 
-Sims in \ref{sec:sims}.
-
-Results in \ref{sec:results}.
-
-Things to do in \ref{sec:todo}.
+The purpose of this study was to provide a means to model a large amount of gravitational lenses by showing that gravitational lens modelling can be learned / done by volunteers.
+We suggest, that volunteers will be as successful as professionals with modelling if provided with an easy to use tool with visual feedback (What you see is what you get, WYSIWYG) and a minimal set of instructions.
+Volunteers will then crowd work (using buzz word here ;) ) / work collaborative on modelling lenses from several sources / groups at a central place.
+Since this is a iterative learning process, the more involved volunteers will quickly gain knowledge that can be passed down to new volunteers.
+That creates a social structure that scales well with the number of volunteers, as other projects\needcite have already shown.
+Finding people working as volunteers has been shown to be successful last but not least by \sw and the whole galaxy zoo project.
+To test the peoples abilities, we investigated the performance of a first set of volunteers modelling a set of simulated lenses.
+We tested the ability to correctly identify lensed images and reproduce similar mass maps of the lens.

### 2013-11-16 plots/gen_plots.py

@@ -93,8 +93,8 @@ from numpy import pi
 # set to false to do a dry run in /plots/res
 write_to_tex_folder = True
 #image extension (png or pdf)
-ext = 'png'
-
+#ext = 'png'
+ext = 'pdf'
 
 # realtive to plots dir
 outdir = 'figs'
@@ -901,7 +901,7 @@ def plotAllRE():
     pl.xlabel('model id')
     pl.ylabel(r'Einstrin radius $\Theta_\text{E}$ [pixel]')
     
-    pl.savefig(os.path.join(path, 'eR_1.png'))
+    pl.savefig(os.path.join(path, 'eR_1.'+ext))
     print '.',
     #pl.show()
   
@@ -910,7 +910,7 @@ def plotAllRE():
     pl.errorbar(xi, re_rel, ee_rel, marker='s', mfc='blue', ls='' ,ecolor='blue')
     pl.plot(xi, xi*0+1, '-r')
     
-    pl.savefig(os.path.join(path, 'eR_2.png'))
+    pl.savefig(os.path.join(path, 'eR_2.'+ext))
     print '.',
     #pl.show()
   
@@ -938,7 +938,7 @@ def plotAllRE():
     pl.xlim([-.5, i+0.5])
     pl.ylabel(r'Einstein Radius $\Theta_{\text{E}}$')
 
-    pl.savefig(os.path.join(path, 'eR_3.png'))
+    pl.savefig(os.path.join(path, 'eR_3.'+ext))
     print '.',
     #pl.show()
     
@@ -972,7 +972,7 @@ def plotAllRE():
     pl.xlabel('Simulation id')
     pl.ylabel(r'rel Einstein Radius $\Theta_{\text{E}}$/$\Theta_{\text{E, sim}}$')
 
-    pl.savefig(os.path.join(path, 'eR_4.png'))
+    pl.savefig(os.path.join(path, 'eR_4.'+ext))
     print '.',
     #pl.show()
     

### 2013-11-16 text/tex/intro.tex
save, go on sec 2

@@
 Robotic procession has been suggested \needcite, and tested \needcite, but has failed so far to be convincing \needcite.
-Another suggestion involves humans, volountiers \footnote{A volounteer is considered everybody that has no background in astro physics.}.
-\sw uses this approach with great succes so far.\needcite
+Another suggestion involves humans, volunteers \footnote{A volunteer is considered everybody that has no background in astro physics.}.
+\sw uses this approach with great success so far.\needcite
 

### 2013-11-17 text/tex/intro.tex
testing 2
Update intro.tex
spelling modifications

@@
 \section{Introduction}
+
+This is my new Text
 
@@
 Even more coming in the future with new surveys \needcite
-Robotic procession has been suggested \needcite, and tested \needcite, but has failed so far to be convincing \needcite.
-Another suggestion involves humans, volunteers \footnote{A volunteer is considered everybody that has no background in astro physics.}.
+Robotic processing has been suggested \needcite, and tested \needcite, but has failed so far to be convincing \needcite.
+Another suggestion involves humans, volunteers \footnote{A volunteer is considered everyone that has no background in astro physics.}.
 \sw uses this approach with great success so far.\needcite
 
-Next step involves modeling. that needs advanced knolegde and takes a lot of time.
-Too much time to be done by astronomers them selves as the results from new surveys and identifiers like \sw come in.
-So there is a demand of a means to modell a great amount of identified lens data, that scales with increasing data amount.
+Next step involves modeling. that needs advanced knowlegde and takes a lot of time.
+Too much time to be done by astronomers themselves as the results from new surveys and identifiers like \sw come in.
+So there is a demand of a means to model a great number of identified lens data, that scales with increasing volume of data.
 
-The purpose of this study was to provide a means to model a large amount of gravitational lenses by showing that gravitational lens modelling can be learned / done by volunteers.
+The purpose of this study was to provide a means to model a large number of gravitational lenses by showing that gravitational lens modeling can be learned / done by volunteers.
 We suggest, that volunteers will be as successful as professionals with modelling if provided with an easy to use tool with visual feedback (What you see is what you get, WYSIWYG) and a minimal set of instructions.
-Volunteers will then crowd work (using buzz word here ;) ) / work collaborative on modelling lenses from several sources / groups at a central place.
-Since this is a iterative learning process, the more involved volunteers will quickly gain knowledge that can be passed down to new volunteers.
+Volunteers will then crowd work (using buzz word here ;) ) / work collaborative on modeling lenses from several sources / groups at a central place.
+Since this is an iterative learning process, the more involved volunteers will quickly gain knowledge that can be passed down to new volunteers.
 That creates a social structure that scales well with the number of volunteers, as other projects\needcite have already shown.
 Finding people working as volunteers has been shown to be successful last but not least by \sw and the whole galaxy zoo project.
-To test the peoples abilities, we investigated the performance of a first set of volunteers modelling a set of simulated lenses.
-We tested the ability to correctly identify lensed images and reproduce similar mass maps of the lens.
+To test the people's abilities, we investigated the performance of a few volunteers modeling a set of simulated lenses.
+We tested the ability to correctly identify lensed images and reproduce similar mass map of the lens.

### 2013-11-18 text/tex/results.tex
now tracking figures
Update results.tex
typos

@@
 
-To estimate the performance of the volunteers and the quality of the generated models, two test T1 and T2 were done.
+To estimate the performance of the volunteers and the quality of the generated models, two tests T1 and T2 were done.
 
@@
 
-T1 was also designed to give some feedback on the difficulties volunteers encounter, to further imrove the tutorial materials.
+T1 was also designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
 

### 2013-11-18 text/tex/results.tex
Update results.tex
typos

@@
 \spl through a video tutorial and by videocon.  After this
-introductory stage, a modelling challenge was presented.  This
+introductory stage, a modeling challenge was presented.  This
 consisted of 29 simulated lenses (sims) covering a range of lensing
@@
 Singular isothermal is in equations (33-35) of
-\cite{2001astro.ph..2341K} with core radius set to zero. NFW is
-equations (48,50) of the same paper, and shear is the $\gamma$ term in
+\cite{2001astro.ph..2341K} with core radius set to zero. NFW (Navarro, Frenk & White 1996, 1997) is
+in equations (48,50) of the same paper, and shear is the $\gamma$ term in
 equation (76) of that work.
@@
 First, the correct identification and location of lensed images (T1a).
-Second to find the correct ordering in respect of the arrival time for the identified lensed images (T1b).
+Second the correct ordering for the identified lensed images in respect of the arrival time (T1b).
 
@@
 Most of the problems where due to unclear arc-like structures (E01, p=0.18; E04, p=0.03; E08, p=0.04).
-Critical errors like the failure to identify all five images in an five image system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happened.
+Critical errors like the failure to identify all five images in a five images system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happen.
 From this we conclude that the introduction materials was adequate and the volunteers understand the basics of gravitational lensing.
 
-\todo{add more detail?} Add more details like total amount of images detected / tot am images (rightPlace fraction)??
+\todo{add more detail?} Add more details like total number of images detected / tot am images (rightPlace fraction)??
 
@@
 These results can be seen in \figref{ER_all_models}.
-This figure shows that this technique of estimating the error is underestimating the error significantly and should be improved for further analysis.
+This figure shows that this technique of estimating the error minimizes the error significantly and should be improved for further analysis.
 
 In \figref{ER_per_sim} can be seen, that the calculated \ERf of the models tend to be too high.
-The overshoot varies from around 0.2 to 0.4 for for good models.
+The overshoot varies from around 0.2 to 0.4 for good models.
 One of the reasons for this is, that it's hard to get the center of the lens on spot.
@@
 This errors change the orientation of the mass distribution \kap[x,y],
-but \kenc and thus the \ERf is not influence that much, and thus the final result is still a valid model.
+but \kenc and thus the \ERf is not influenced that much, and thus the final result is still a valid model.
 This can be seen by comparing the results for \asw{0h2m}, shown in \figref{kapenc_compare_faulty}: 
@@
 
-Comparing the models of volunteers and experts can be done in \figref{kapenc_compare_faulty}, where only the
+Comparing the models from volunteers and experts can be done in \figref{kapenc_compare_faulty}, where only the
 expert got the right configuration, but all the resulting models are comparable, besides rotation.

### 2013-11-25 text/tex/outlook.tex

@@
 As seen in \secref{tests.t1}, volunteers fail mostly in two situations:
-when to identify an arclike structre while placing the points
+when to identify an arclike structure while placing the points
 and in 5 image configurations, that are quite symmetric, then to identify the correct ordering of the points.
@@
 
-The second can be made less by better training (introduce the ruler), maybe more filters help too?
+The second can be made less by better training (introduce the ruler), may be more filters help too?
 
-We hope that those problems are also taken care of when staring collaborative modelling. (Already in progress)
+We hope that those problems are also taken care of when staring collaborative modeling. (Already in progress)
 Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models ect...)

### 2013-11-25 text/tex/results.tex
spellings

@@
   \item {\em Galaxy:\/} Lens is circular NFW plus one dominant
-    elliptical SIE and perturbating elliptical SIEs, source as with
+    elliptical SIE and perturbed elliptical SIEs, source as with
     galaxy.
@@
 \Figref{6975} shows an example where substructure introduces a
-complication, but model ok.
+complication, but model OK.
 
@@
 For an example of E06, see \figref{7022}.
-It basically describes a situation, where the minima and saddle points of a five image configuration were exchanged (rotated by $\pm90\dgr$).
+It basically describes a situation, where the minima and saddle points of a five images configuration were exchanged (rotated by $\pm90\dgr$).
 E05 describes the situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$).

### 2013-11-25 text/tex/abstract.tex

@@
 \begin{abstract}
-In Spacewarps, volunteers are invited to search sky surveys for lens
+In Spacewarps **, volunteers are invited to search sky surveys for lens
 candidates.  Here we report on a system that allows experienced
 volunteers to model lens candidates.  A sample of 29 simulated lenses
-were modelled, each multiple times.  The quality of the models was
+were modeled, each multiple times.  The quality of the models was
 then examined in different ways.  In particular, (i)~the image

### 2013-11-25 text/tex/intro.tex

@@
 \section{Introduction}
-
-This is my new Text
-
 Gravitational lenses (GLs), predicted by General Relativity (GR), can be used as a tool by astronomers to examine many properties of the cosmos.
@@
 
-%To accheve this, one needs to find / identify lenses and in a second step model those lenses to get involved parameters.
+%To achieve this, one needs to find / identify lenses and in a second step model those lenses to get involved parameters.
 
@@
 Robotic processing has been suggested \needcite, and tested \needcite, but has failed so far to be convincing \needcite.
-Another suggestion involves humans, volunteers \footnote{A volunteer is considered everyone that has no background in astro physics.}.
+Another suggestion involves humans, volunteers \footnote{A volunteer is considered everyone that has no background in astrophysics.}.
 \sw uses this approach with great success so far.\needcite
 
-Next step involves modeling. that needs advanced knowlegde and takes a lot of time.
+Next step involves modeling. that needs advanced knowledge and takes a lot of time.
 Too much time to be done by astronomers themselves as the results from new surveys and identifiers like \sw come in.
@@
 The purpose of this study was to provide a means to model a large number of gravitational lenses by showing that gravitational lens modeling can be learned / done by volunteers.
-We suggest, that volunteers will be as successful as professionals with modelling if provided with an easy to use tool with visual feedback (What you see is what you get, WYSIWYG) and a minimal set of instructions.
+We suggest, that volunteers will be as successful as professionals with modeling if provided with an easy to use tool with visual feedback (What you see is what you get, WYSIWYG) and a minimal set of instructions.
 Volunteers will then crowd work (using buzz word here ;) ) / work collaborative on modeling lenses from several sources / groups at a central place.

### 2013-11-25 text/tex/outlook.tex

@@
 As seen in \secref{tests.t1}, volunteers fail mostly in two situations:
-when to identify an arclike structure while placing the points
+when to identify an arc like structure while placing the points
 and in 5 image configurations, that are quite symmetric, then to identify the correct ordering of the points.
@@
 We hope that those problems are also taken care of when staring collaborative modeling. (Already in progress)
-Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models ect...)
+Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models etc...)
 

### 2013-11-26 text/tex/appendix.tex
editing results section

@@
 explanation, we suppressed some constant factors in
-equations \eqref{eq:Ageom}, \eqref{eq:Ageom} and \eqref{eq:Poisson}.
+equations \eqref{eq:Ageom}, \eqref{eq:Aarriv} and \eqref{eq:Poisson}.
 Here we fill in the details.
 
-First we define comoving distances.
+First let us recall the formulas for comoving distances.
 \begin{equation}
@@
 \end{equation}
-Can replace comoving distances $D$ with angular-diameter distances
-$d$, and so on:
+Comoving distances $D$ with angular-diameter distances $d$, as follows.
 \begin{equation}
@@
 \end{equation}
-Angular-diameter distances are always smaller than comoving distances.
 Locations on the lens plane can be replaced angular coordinates on the

### 2013-11-26 text/tex/results.tex
editing results section

--- 2013-11-18 text/tex/results.tex
@@
 
+To estimate the performance of the volunteers and the quality of the generated models, two tests T1 and T2 were done.
+
+The second test T2 was to compare the desired results of a modeling process, the mass distribution of the lens $\kappa(x, y)$.
+
+The \sw sims were generated by AM, in consultation with PM and AV.
+
 \subsection{The simulated lenses} \label{sec:sims}
 
-The sims were generated by AM, in consultation with PM and AV.  In the
-interest of blind testing, the information in this section was
-revealed neither to RK, while choosing the challenge set, nor to the
-people making models (EB, CC, CM, JO, PS and JW) until they were done.
+In the interest of blind testing, the information in this section was
+revealed neither to RK, while choosing the challenge set of 29 sims,
+nor to the modellers (volunteers EB, CC, CM, JO, JW and `expert
+modeller' PS) until the modelling stage was done.
 
-The sims produced using {\tt gravlens}
-\citep{2001astro.ph..2341K,2001astro.ph..2340K} and were of three
+The sims were produced using {\tt gravlens}
+\citep{2001astro.ph..2341K,2001astro.ph..2340K}.  The were of three
 kinds, as follows.
 
-\begin{itemize}
-  \item {\em Quasar:\/} A singular elliptical isothermal plus constant
-    external shear, with a circular Gaussian source.
-  \item {\em Galaxy:\/} Lens as with quasar, but elliptical de
-    Vaucouleurs source.
-  \item {\em Galaxy:\/} Lens is circular NFW plus one dominant
-    elliptical SIE and perturbating elliptical SIEs, source as with
-    galaxy.
-\end{itemize}
+\begin{enumerate}
+  \item Imitating lensed quasars: having a singular elliptical
+    isothermal lens (SIE) plus constant external shear, and a circular
+    Gaussian source.
+  \item Emulating lensed galaxies: similar to the above, but with an
+    elliptical de Vaucouleurs source.
+  \item Resembling cluster lenses: having a source as above, but a
+    more complicated lens, with one dominant elliptical SIE and and
+    one or more perturbing elliptical SIEs, plus a circular NFW
+    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
+    the intergalactic mass.
+\end{enumerate}
 
-Singular isothermal is in equations (33-35) of
-\cite{2001astro.ph..2341K} with core radius set to zero. NFW (Navarro, Frenk & White 1996, 1997) is
-in equations (48,50) of the same paper, and shear is the $\gamma$ term in
-equation (76) of that work.
+Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
+lenses follow equations (33--35) of that work, with core radius set to
+zero.  The NFW lens is in equations (48) and (50), while shear is the
+$\gamma$ term in equation (76).
 
@@
 
-\Figref{6941} shows a simple example.
+The modellers proffered a total of 129 models for the 29 sims in the
+challenge.  Figures \ref{fig:6941} to \ref{fig:6915} show eight of
+models in some detail.
 
-\Figref{6975} shows an example where substructure introduces a
+Figure \ref{fig:6941} shows a simple example.  The top two panels show
+contours of convergence $\kappa(x,y)$, and of the arrival-time for a
+point source at the brightest point of the sim
+source.\footnote{Although the arrival-time contours represent an
+  abstract quantity that is not directly observable, the contour map
+  may actually resemble the appearance of the lensed arcs.  The
+  resemblance is due to a serendipitous computer-graphical effect
+  \citep{2001AJ....122..585S}.} The spatial scale is in pixels.  The
+information in these two panels was, of course, kept secret during the
+modelling challenge.  The panels in the middle row and at the lower
+right are the results that \spl returns to the modeller, in order for
+them to assess the model.  These derive from the mean of an ensemble
+of 200 models generated by \spl.  The model mass distribution is
+returned as a contour map superimposed on an intensity map. A fairly
+smooth mass distribution, as here, is a good sign.  An irregular or
+checkboard pattern in the mass map usually indicates a bad model.  The
+model arrival-time contours.  The special contour passing through a
+saddle point is shown in black: this is the model version of the
+spaghetti sketch provided by the user.  Also provided is a synthetic
+image of the lensed source.  For this, \spl assumes a simple conical
+source profile.  The user can change the contrast level on the image,
+which (though it is not saved) amounts to to adjusting the width of
+the cone.  Finally, to the lower left of the figure, we have a
+comparison of the input and recovered mass profiles.  The panel shows
+a average $\kappa$ within a given radius, as a function of radius.
+The red curve is the true value, and where it crosses unity (dotted
+horizonal line) is the notional Einstein radius.  The two blue curves
+are the minimal and maximal mean enclosed $\kappa$ from the internal
+ensemble in \spl.  The region between the blue curves is shaded
+between the radii of the innermost and the outermost images --- this
+is the confidence region from the modelling.  As we see, the shaded
+blue is slightly above the red curve.  In summary: the identification
+of minimum and saddle point is correct, but the estimated Einstein
+radius is a little too high.
+
+Figure \ref{fig:6975} shows an example where substructure introduces a
 complication, but model ok.
 
-\Figref{6937} shows a case where substructure leads to a
+Figure \ref{fig:6937} shows a case where substructure leads to a
 poorer model.
 
-\Figref{7022} shows incorrect identification, but the enclosed
+Figure \ref{fig:7022} shows incorrect identification, but the enclosed
 mass still quite well recovered.
 
-\Figref{6975} shows a nice symmetric quad.
+Figure \ref{fig:6975} shows a nice symmetric quad.
 
-\Figref{6990} shows a long-axis quad.
+Figure \ref{fig:6990} shows a long-axis quad.
 
-\Figref{6919} shows a short-axis quad.
+Figure \ref{fig:6919} shows a short-axis quad.
 
-\Figref{6915} shows an inclined quad.
+Figure \ref{fig:6915} shows an inclined quad.
 
-\subsection{The test setup} \label{sec:testsetup}
-
-To estimate the performance of the volunteers and the quality of the generated models, two tests T1 and T2 were done.
+\subsection{Test of image identification} \label{sec:tests.t1}
 
@@
 
-The second test T2 was to compare the desired results of a modeling process, the mass distribution of the lens $\kappa(x, y)$.
-To get a means of comparing the simulated data (sims) to the generated modeled data (models), the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ for both was calculated.
-The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between the sims and the models.
-We also let an expert model three selected systems to compare the results from volunteers to those of a professional.
-
-\todo{!} Should I write down prior expectation for T2 too?
-
-\subsection{Test of image identification} \label{sec:tests.t1}
-
 The evaluation of the volunteers performance for T1 was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot generated using simulation parameters.
-\Figref{output_compare} shows the setup used to evaluate the models.
+% \Figref{output_compare} shows the setup used to evaluate the models.
 
@@
 Additionally, ten types of errors (E01 -- E10, listed in \tabref{stats}) that could occur were analyzed, where each model could contain multiple of those.
-The complete table with the results for each model can be found in the appendix, \tabref{detail_results}.
+
+% The complete table with the results for each model can be found in
+% the appendix, \tabref{detail_results}.
 
@@
 
+To get a means of comparing the simulated data (sims) to the generated modeled data (models), the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ for both was calculated.
+The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between the sims and the models.
+We also let an expert model three selected systems to compare the results from volunteers to those of a professional.
+
+\todo{!} Should I write down prior expectation for T2 too?
+
 To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] directly generated in the modeling process.

### 2013-11-28 text/tex/results.tex

@@
 
-Figure \ref{fig:6975} shows an example where substructure introduces a
-complication, but model ok.
+Figure \ref{fig:6975} shows a lens with substructure in the form of a
+smaller secondary galaxy.  The galaxies in such group or cluster sims
+were, in fact, based on galaxies visible in the images --- but the
+modellers were not told in advance whether this was the case.  The
+model does not include any substructure, but otherwise is not bad.
+The minimum and saddle point are correct, and the Einstein radius is
+only a little underestimated.
 
 Figure \ref{fig:6937} shows a case where substructure leads to a
-poorer model.
+poor model.
 
-Figure \ref{fig:7022} shows incorrect identification, but the enclosed
-mass still quite well recovered.
-
-Figure \ref{fig:6975} shows a nice symmetric quad.
+Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
+saddle points are correctly identified, and the orientation of the
+ellipticity of the mass distribution is correctly reproduced.  The
+Einstein radius is somewhat overestimated.  Figure \ref{fig:7022}
+shows another model of the same system.  In this one, the
+identification of the minima and saddle points was incorrect, and mass
+distribution comes out elongated East-West instead of North-South.
+The mass distribution also appears somewhat jagged and the
+saddle-point contours are not as clean as in the previous examples;
+these are often indicators of a problem with the model.  The enclosed
+mass is, however, none the worse --- the reason is probably that in a
+relatively symmetric image configuration, the Einstein radius is quite
+well constrained by the images in a fairly model-independent way.
 

### 2013-11-28 text/tex/spaghetti.tex

@@
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/006941/input.png}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/007022/input.png}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/sims/006919/input.png}}
-  \caption{Examples of Spaghetti input.  The models from these appear
-    later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
-  }
-  \label{fig:input-spag}
-\end{figure}
-
 \subsection{\spl} \label{sec:SpaghettiLens}
@@
 \spl, as implemented so far, assumes that the lens is dominated by a
-single galaxy, and that the centre of that galaxy is the sole maximum.
+single galaxy, and that the center of that galaxy is the sole maximum.
 Once the maximum has been identified, we wish to characterize the rest
@@
 separate them from other background light sources, and (b)~classify
-and order images accoring to arrival time.  But the computer helps by
+and order images according to arrival time.  But the computer helps by
 allowing only valid lensing configurations to be entered, and ensuring
@@
 
-In the second modelling step, once the user has sketched a proposed
+In the second modeling step, once the user has sketched a proposed
 spaghetti configurations, \spl sends the input to its server-side
-modelling engine, called GLASS.  The task of GLASS is to find a mass
+modeling engine, called GLASS.  The task of GLASS is to find a mass
 distribution $\kappa(x,y)$ that exactly reproduces the locations of
 the maximum, minima and saddle points.  Now, this criterion by itself
-is extremely underdetermined --- there are infinitely many mass
+is extremely under-determined --- there are infinitely many mass
 distributions that will reproduce a given set of maxima, minima and
@@
   larger.)
-\item There is a notional lens centre, say $(x_0,y_0)$ which is
+\item There is a notional lens center, say $(x_0,y_0)$ which is
   identified with the maximum of the arrival time.  The source can
-  have an arbitrary offset with respect to the lens centre.
+  have an arbitrary offset with respect to the lens center.
 \item The mass distribution must be centrally concentrated, in two
@@
 
-In the third modelling step, having generated a model ensemble, \spl
-postprocesses it to present results and diagnostics to the user for
+In the third modeling step, having generated a model ensemble, \spl
+post-processes it to present results and diagnostics to the user for
 inspection.  This takes the form of three figures.
 \begin{enumerate}
-\item A grayscale plus contour map of the mass distribution.
+\item A gray scale plus contour map of the mass distribution.
 \item A contour map of the arrival-time surface.
@@
 
-The fourth modelling step is discussion among modellers and iteration
+The fourth modeling step is discussion among modelers and iteration
 on the model.  Any archived model can be revised by another user: they
@@
 interesting lens candidates lead to trees of models in this way.
-Forum discussion then prunes the tree, focussing attention on one or a
+Forum discussion then prunes the tree, focusing attention on one or a
 few models.

### 2013-11-28 text/tex/stats.tex

\begin{table}\centering\begin{tabular}{llcc}\hline
 & & n & p \\
\hline
 N: & Total number of models & 119 & 1.00\\
\hline
 R1: & images approx. on right location & 110 & 0.92\\
 R2: & images with correct parity & 70 & 0.59\\
\hline
 E01: & inaccurate in arc & 21 & 0.18\\
 E02: & wrong parity in 3 lens conf. & 2 & 0.02\\
 E03: & identified 3 of 5 imgs. & 5 & 0.04\\
 E04: & modeled arc with single img. & 4 & 0.03\\
 E05: & $\pi$ rotated parity & 7 & 0.06\\
 E06: & $\pi$/2 rotated parity & 38 & 0.32\\
 E07: & missed faint img. & 1 & 0.01\\
 E08: & too many imgs in arc. & 5 & 0.04\\
 E09: & missed double img & 3 & 0.03\\
 E10: & too many imgs. & 1 & 0.01\\
\end{tabular}

\caption{1. col: id of model; 2.col: places of imgs idendited
  correctly; 3. col: more or less correct identification of extr
  points. 4.col: exact identification of extremal points. 5.col:
  type(s) of errors ocured....  1) inaccurate placement in an extended
  arc; 2) wrongly identified sad and min in 3 image configuration; 3)
  identified only 3 instead of 5 images; 4) tried to model an arc with
  a min instead of min-sad-min; 5) PI-err (rotation by 180 degrees; in
  5 image configuration, exchanged the ordering of the two saddle
  points); 6) PI/2-err (rotation by 90 degree; sad-min-sad-min-sad);
  7) missed faint image(s); 8) tried to model an arc with min-sad-min
  instead of only min; 9) did identify two close by images as one; 10)
  used 7 or more image to model a 5 image system. }

\label{tab:stats}
\end{table}

### 2013-11-28 text/tex/theory.tex
save, go on sec 2

@@
 originally introduced by \cite{1986ApJ...310..568B}, based on Fermat's
-principle.  They key to this approach is an abstract construct called
+principle.  The key to this approach is an abstract construct called
 the {\em arrival-time surface.}  This surface cannot itself be
@@
 lens plane $(x,y)$ and then changing its path to reach the observer.
-The observer would see the source apparently behind $(x,y)$ on the
+The observer would see the source apparently behind $(x,y)$ the
 lens.  If this happened, the light ray would have taken a longer path
@@
 \end{equation}
-As written, this looks is an area.  But we can also think of it as a
+As written, this looks an area.  But we can also think of it as a
 time delay introduced by the light ray bending --- with an unwritten
-constant factor.  That factor is basically the lens distance times the
+constant factor.  That factor is basically the lens distance divided by the
 speed of light; the precise expression has to take the expansion of
@@
 kg\,m^{-2}$, in special units (given in the Appendix) such that the
-surface density becomes a dimensional field $\kappa(x,y)$.  Then
+surface density becomes a dimensionless field $\kappa(x,y)$.  Then
 \begin{equation} \label{eq:Poisson}
@@
 Equations of the type \eqref{eq:Poisson} are, however, well known (as
-Poisson equations in two dimensions), and there are techniques for
+Poisson's equation in two dimensions), and there are techniques for
 solving them.  Incidentally, $\kappa$ has a second meaning, as well as
@@
 a peak at the centre of that circle.  But if the mass is not precisely
-in front of the mass, the arrival-time surface will not be circular
+in front of the source, the arrival-time surface will not be circular
 any more.  The result is illustrated in the middle part of
@@
 minimum and a saddle point.  According to Fermat's principle, images
-appear at maxima and saddle points as well as minima.  If the mass
+appear at maximum and saddle points as well as minimum.  If the mass
 distribution is non-circular, more images can appear.  The lower
@@
 arrival-time surface is nearly flat.  As is evident from
-\figref{arriv}, this is the region where the minima and saddle
+\figref{arriv}, this is the region where the minimum and saddle
 points lie, or near the images.  In this region, points on the source
@@
 demagnified.  We see from \figref{arriv} that the arrival-time
-surface tends to be highly curved near the maximum.  Hence maxima tend
+surface tends to be highly curved near the maximum.  Hence maximum tend
 to be demagnified.  In practice, maxima of the arrival time are nearly

### 2013-12-01 text/tex/appendix.tex
tentative correction wording formulas

--- 2013-09-11 text/tex/appendix.tex
@@
 \begin{aligned}
-&D_S    &                                &\int_0^{a_S} \\
+&D_S    &                                &\int_0^{z_S} \\
 \noalign{\medskip}
@@
 \end{equation}
-Can replace comoving distances $D$ with angular-diameter distances
-$d$, and so on:
+
+% wording and equations to be revised
+Comoving distances $D$ are defined in angular-diameter distances
+$d$ by:
 \begin{equation}
@@
 Angular-diameter distances are always smaller than comoving distances.
-Locations on the lens plane can be replaced angular coordinates on the
-sky, as
+Locations on the lens plane in angular coordinates on the
+sky, are expressed as
 \begin{equation}
@@
 
-The $A$ times and the $\kappa$ density are related to physical arrival
+The $A$ areas (with the implicit
+constant factor that makes them time delays) and the $\kappa$ density are related to physical arrival
 time $t$ and density $\Sigma$ as
@@
 A           &= \frac{cD_L}{(1+z_L)^2} \frac{D_{LS}}{D_S} \times t \\
-\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L}{1+z_L} \frac{D_{LS}}{D_S}
+\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{{D_L}{D_{LS}}}{D_S}
                \times \Sigma(x,y)

### 2013-12-02 text/tex/appendix.tex
modifications suggested
Please review wording and formulas

@@
 
-First we define comoving distances.
+First we define comoving distances:
+The scale factor is
 \begin{equation}
@@
 \end{equation}
+The total line-of-sight comoving distance is then given by
 \begin{equation}
@@
 % wording and equations to be revised
-Comoving distances $D$ are defined in angular-diameter distances
+Then the comoving distances $D$ can be written as function of the angular-diameter distances
 $d$ by:
@@
 \begin{aligned}
-A           &= \frac{cD_L}{(1+z_L)^2} \frac{D_{LS}}{D_S} \times t \\
-\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{{D_L}{D_{LS}}}{D_S}
-               \times \Sigma(x,y)
+A           &= \frac{c}{(1+z_L)^2} \frac{D_L D_{LS}}{D_S} \times t \\
+\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L D_{LS}}{D_S}
+               \times \Sigma(x,y)\end{aligned}
 \end{aligned}
 \end{equation}
-Letting the source be behind $(s_x,s_y)$ rather than behind the
+Letting the source be offset at $(s_x,s_y)$ rather than at the
 origin, we have
@@
 \begin{aligned}
-t_{\rm geom} &= \frac{(1+z_L)^2}{2cD_L} \frac{D_S}{D_{LS}}
-\left( (x-s_x)^2 + (y-s_y)^2 \right)
+t_{\rm geom} &= \frac{(1+z_L)^2}{2c} \frac{D_S}{D_L D_{LS}}
+\left( (x-s_x)^2 + (y-s_y)^2 \right) \\
 \nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
@@
 from \cite{1986ApJ...310..568B}.
+%% to be explicited

### 2013-12-03 text/tex/appendix.tex

@@
 explanation, we suppressed some constant factors in
-equations \eqref{eq:Ageom}, \eqref{eq:Ageom} and \eqref{eq:Poisson}.
+equations \eqref{eq:Ageom}, \eqref{eq:Aarriv} and \eqref{eq:Poisson}.
 Here we fill in the details.
 
-First we define comoving distances:
-The scale factor is
+First let us recall the formulas for comoving distances.
 \begin{equation}
-\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
+D = \int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
 \end{equation}
-The total line-of-sight comoving distance is then given by
+The comoving distances from observer to lens, from lens to source, and
+from observer to source are given as follows.
 \begin{equation}
@@
 \end{equation}
-
-% wording and equations to be revised
-Then the comoving distances $D$ can be written as function of the angular-diameter distances
-$d$ by:
+Angular-diameter distances $d$ can be related to comoving distances
+$D$ by:
 \begin{equation}
@@
 Angular-diameter distances are always smaller than comoving distances.
-Locations on the lens plane in angular coordinates on the
-sky, are expressed as
+
+Locations on the lens plane in angular coordinates on the sky, are
+expressed as
 \begin{equation}
@@
 \kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L D_{LS}}{D_S}
-               \times \Sigma(x,y)\end{aligned}
+               \times \Sigma(x,y)
 \end{aligned}

### 2013-12-05 text/tex/appendix-formalism.tex
appendix "Gravitational lensing formalism" from Wikipedia

%% http://latex.informatik.uni-halle.de/latex-online/latex.php

\documentclass[12pt,preprint]{aastex}

%%% PACKAGES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath}
\usepackage{xspace}   % fixes spaces after custom commands
\usepackage{float}    % exact placement of figures (using \begin{figure}[H])
\usepackage{color}  % well... color...
\usepackage{subfigure}

\usepackage{longtable}   % those two to generate the 2 ol table
\usepackage{multicol}

%%% DEFINITIONS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\half{{\textstyle\frac12}}
% makes it easier to have consistent writing, and easy change if needed
\newcommand{\spl}{SpaghettiLens\xspace}
\newcommand{\sw}{SpaceWarps\xspace}

% shortcut for einstein radius (Text Greek Full Math)
\newcommand{\ER}{Einstein radius\xspace} % text
\newcommand{\ERg}[1][]{$\Theta_\text{E#1}$\xspace} % er in textmode with greek
\newcommand{\ERf}[1][]{Einstein radius $\Theta_\text{E#1}$\xspace} % full output
\newcommand{\ERm}[1][]{\Theta_\text{E#1}} % mathmode with greek symbol

%shortcuts for kappa
\newcommand{\kenc}[1][r]{$\kappa_\text{encl}(#1)$\xspace}
\newcommand{\kap}[1][r]{$\kappa(#1)$\xspace}

% shorcuts for refs (use capital for beginning of sentence)
% first 3 are for the real layz people..
\newcommand{\fref}[1]{\ref{fig:#1}}
\newcommand{\sref}[1]{\ref{sec:#1}}
\newcommand{\tref}[1]{\ref{tab:#1}}
\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\secref}[1]{Section~\ref{sec:#1}}
\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\Figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\Secref}[1]{Section~\ref{sec:#1}}
\newcommand{\Tabref}[1]{Table~\ref{tab:#1}}

% shortcut for ASW000xxxx
\newcommand{\asw}[1]{ASW000#1\xspace}

%shorcut for model (maybe link later to appendix)
% use \model{4356} for with text and
% \model[]{3456} for only number
\newcommand{\model}[2][Model~]{#1#2\xspace}

%degrees
\newcommand{\dgr}{^{\circ}}

% todo annotations on side in red. easy to search for at the end
\newcommand{\todo}[2][red]{%\marginpar{\raggedright\scriptsize\sffamily\textcolor{red}{#1}}}
%{\marginpar{\raggedright\sffamily\textcolor{red}\footnotesize
%\setstretch{1.025}%
%#1}}}
\textcolor{#1}{\textbullet}%
\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}

\newcommand{\needfig}[1][]{%
\colorbox{yellow}{\parbox{0.5\textwidth}{%
\vspace{0.5cm}\setstretch{0.5}\textcolor{black}{\scriptsize{fig #1}}\vspace{0.5cm}}}%
\marginpar{\colorbox{yellow}{\parbox{\marginparwidth}{%
\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{fig}}}}}}

\setlength{\marginparsep}{0mm}
\setlength{\marginparwidth}{2.2cm}

% quick todo annotations
\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
\newcommand{\needchk}[1][]{\todo[red]{chk #1}}

\newcommand{\hr}{\vspace{5mm}\noindent\rule{0.8\textwidth}{0.4pt}\vspace{5mm}}

%%% DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\section{\Large Gravitational Lensing formalism }

In general relativity, a point mass deflects a light ray with impact parameter \(b\) by an angle \[\hat{\alpha} = \frac{4GM}{c^2b}\] where G is the Gravitational constant, M the mass of the deflecting object  and c the speed of light. A naive application of Newtonian gravity can yield exactly half this value, where the light ray is assumed as a massed particle and scattered by the gravitational potential well.
In situations where General Relativity can be approximated by linearized gravity, the deflection due to a spatially extended mass can be written simply as a vector sum over point masses.  In the continuum limit, this becomes an integral over the density \(\rho \), and if the deflection is small we can approximate the gravitational potential along the deflected trajectory by the potential along the undeflected trajectory, as in the Born approximation in Quantum Mechanics.  The deflection is then
\[\vec{\hat{\alpha}}(\vec{\xi})=\frac{4 G}{c^2} \int d^2\xi^{\prime} \int dz \rho(\vec{\xi}^{\prime},z)  \frac{\vec{b}}{|\vec{b}|^2}, ~ \vec{b} \equiv \vec{\xi}  - \xi^{\prime} \]
where \(z\) is the line-of-sight coordinate, and \( \vec{b} \) is the vector impact parameter of the actual ray path from the infinitesimal mass \[  d^2\xi^{\prime}  dz\rho(\vec{\xi}^{\prime},z) \] located at the coordinates \[(\vec{\xi}^{\prime}, z)\]
% cite >{{ cite journal | last = Bartelmann| first = M. | coauthors = Schneider, P. | year = 2001 | month = January | title=Weak Gravitational Lensing| journal = Physics Reports D | volume=340 | issue = 4–5 |pages = 291–472 | bibcode = 2001PhR...340..291B | doi = 10.1016/S0370-1573(00)00082-X|arxiv = astro-ph/9912508 }}

\textbf{Thin lens approximation}

In the limit of a "thin lens", where the distances between the source, lens, and observer are much larger than the size of the lens (this is almost always true for astronomical objects), we can define the projected mass density

\[\Sigma(\vec{\xi})=\int \rho(\vec{\xi},z) dz \]

where \(\vec{\xi}\) is a vector in the plane of the sky.  The deflection angle is then

\[\vec{\hat{\alpha}}= 
\frac{4 G}{c^2} \int \frac{(\vec{\xi}-\vec{\xi}^{\prime})\Sigma(\vec{\xi}^{\prime})}{|\vec{\xi}-\vec{\xi}^{\prime}|^2}d^2 \xi^{\prime 2}\]

\includegraphics[scale=0.5]{../fig/lens-sketch.png}
\caption{\\fig. Sketch of a typical gravitational lens system \footnotesize [Bartelmann, Schneider 2001] }

As shown in the diagram, the difference between the unlensed angular position \(\vec{\beta}\) and the observed position \(\vec{\theta}\) is this deflection angle, reduced by a ratio of distances, described as the lens equation

\[\vec{\beta}=\vec{\theta}-\vec{\alpha}(\vec{\theta}) = \vec{\theta} - \frac{D_{ds}}{D_s} \vec{\hat{\alpha}}(\vec{D_d\theta})\]

where \(D_{ds}\) is the distance from the lens to the source \(D_s\) is the distance from the observer to the source, and \(D_d\) is the distance from the observer to the lens.  For extragalactic lenses, these must be angular diameter distances.

In strong gravitational lensing, this equation can have multiple solutions, because a single source at \(\vec{\beta}\) can be lensed into multiple images.

\pagebreak
\textbf{Convergence and deflection potential}

It is convenient to define the "convergence"

\[
\kappa(\vec{\theta}) = \frac{\Sigma(D_d\vec{\theta})}{\Sigma_{cr}}
\]

and the "critical surface density" (not to be confused with the Critical density (cosmology) of the universe)

\[
\Sigma_{cr} = \frac{c^2 D_s}{4\pi G D_{ds}D_d}
\]

The reduced deflection angle \(\vec{\alpha}(\vec{\theta})\) can now be written as

\[
\vec{\alpha}(\vec{\theta}) = \frac{1}{\pi}\int d^2 \theta^{\prime} \frac{(\vec{\theta}-\vec{\theta}^{\prime})\kappa(\vec{\theta}^{\prime})}{|\vec{\theta}-\vec{\theta}^{\prime}|^2}
\]

We can also define the "`deflection potential"'

\[
\psi(\vec{\theta}) = \frac{1}{\pi}\int d^2 \theta^{\prime} \kappa(\vec{\theta}^{\prime}) \ln |\vec{\theta}-\vec{\theta}^{\prime}|
\]

such that the scaled deflection angle is just the gradient of the potential and the convergence is half the Laplacian of the potential:

\[
\vec{\alpha}(\vec{\theta}) = \vec{\nabla} \psi(\vec{\theta})
\]

\[
\kappa(\vec{\theta}) = \frac{1}{2} \nabla^2 \psi(\vec{\theta})
\]

The deflection potential can also be written as a scaled projection of the Newtonian gravitational potential \(\Phi\) of the lens

%%% {{ cite journal | last = Narayan| first = R. | coauthors = Bartelmann, M. | year = 1996 | month = June | title=Lectures on Gravitational Lensing| journal = Eprint arXiv:astro-ph/9606001 | bibcode =1996astro.ph..6001N |arxiv = astro-ph/9606001 | last2 = Bartelmann | pages = 6001 }}

\[
\psi(\vec{\theta}) = \frac{2 D_{ds}}{D_d D_s c^2} \int \Phi(D_d\vec{\theta},z) dz
\]

\pagebreak
\textbf{Lensing Jacobian}

The Jacobian matrix and determinant between the unlensed and lensed coordinate systems is

\[A_{ij}=\frac{\partial \beta_i}{\partial \theta_j}=\delta_{ij} - \frac{\partial \alpha_i}{\partial \theta_j}
= \delta_{ij} - \frac{\partial^2 \psi}{\partial \theta_i \partial \theta_j}\]

where \(\delta_{ij}\) is the Kronecker delta.  Because the matrix of second derivatives must be symmetric, the Jacobian can be decomposed into a diagonal term involving the convergence and a matrix trace-free term involving the "shear" \(\gamma \)

\[A=(1-\kappa)\left[\begin{array}{ c c } 1 & 0 \\ 0 & 1 \end{array}\right]-\gamma\left[\begin{array}{ c c } \cos 2\phi & \sin 2\phi \\ \sin 2\phi & -\cos 2\phi \end{array}\right]\]

where \(\phi\) is the angle between \(\vec{\alpha}\) and the x-axis.  The term involving the convergence magnifies the image by increasing its size while conserving surface brightness.  The term involving the shear stretches the image tangentially around the lens, as discussed in Weak lensing observables.

The shear defined here is 'not' equivalent to the shear mapping traditionally defined in mathematics, though both stretch an image non-uniformly.

\textbf{Fermat Surface}

There is an alternative way of deriving the lens equation, starting from the photon arrival time (Fermat surface)

\[
t =  \int_{0}^{z_s}   { n dz \over c \cos \alpha(z) } 
\]

where \( dz/c \) is the time to travel an infinitesimal line element along the source-observer straight line in vacuum, which is 
then corrected by the factor

\[
1/\cos(\alpha(z))  \approx 1 + {\alpha(z)^2 \over 2}
\]

to get the line element along the bended path \[ dl = {dz \over c \cos \alpha(z) } \]  with a varying small pitch angle \( \alpha(z)\), and the refraction index  
\(n \) for the "aether", i.e., the gravitational field.  The last can be obtained from the fact that a photon travels on a null geodesic of a weakly perturbed static Minkowski universe

\[
ds^2 = 0  = c^2 dt^2 \left(1 + {2 \Phi \over c^2} \right) -  \left(1 + {2 \Phi \over c^2} \right)^{-1} dl^2
\]

where  the uneven gravitational potential \[ \Phi \ll c^2 \] drives a changing the speed of light

\[
c' = {dl/dt} = \left(1 + {2 \Phi \over c^2} \right)  c
\]

So the refraction index

\[
 n \equiv {c \over c'} \approx  \left(1 -  {2 \Phi \over c^2} \right) 
\]

The refraction index greater than unity because of the negative gravitational potential \( \Phi \).

Put these together and keep the leading terms we have the time arrival surface

\[ 
t \approx \int_0^{z_s} {dz \over c}  +  \int_0^{z_s} {dz \over c} {\alpha(z)^2 \over 2} -  \int_0^{z_s} {dz \over c} {2 \Phi \over c^2} 
\]

The first term is the straight path travel time, the second term is the extra geometric path, and the third is the gravitational delay. 
Make the triangle approximation that \[ \alpha(z) = \theta - \beta \] for the path between the observer and the lens, 
and \[ \alpha(z) \approx (\theta - \beta) {D_d \over D_{ds}} \] for the path between the lens and the source.  
The geometric delay term becomes

\[
    {D_d \over c}   { (\vec{\theta} - \vec{\beta})^2 \over 2} 
+ {D_{ds} \over c}  { \left[ (\vec{\theta} - \vec{\beta} )  {D_d \over D_{ds}} \right]^2 \over 2}
= {D_d D_s \over D_{ds} }    { ( \vec{\theta} - \vec{\beta} )^2 \over 2}
\]

So the Fermat surface becomes

\[
 t = constant + {D_d D_s \over D_{ds} c} \tau, ~ \tau \equiv \left[  { (\vec{\theta}-\vec{\beta})^2 \over 2} -  \psi \right] 
\]

where \( \tau \) is so-called dimensionless time delay, and the 2D lensing potential

\[
\psi(\vec{\theta}) = \frac{2 D_{ds}}{D_d D_s c^2} \int \Phi(D_d\vec{\theta},z) dz
\]

The images lie at the extrema of this surface, so the variation of \(t \) with \(\vec{\theta} \) is zero,

\[
0 = \nabla_{\vec{\theta}} \tau  =  \vec{\theta} - \vec{\beta}  - \nabla_{\vec{\theta}} \psi(\vec{\theta}) 
\]

which is the lens equation.  Take the Poisson's equation for 3D potential 
\[
\Phi(\vec{\xi})  = - \int  \frac{d^3\xi^{\prime} \rho(\vec{\xi}^{\prime})}{|\vec{\xi}-\vec{\xi}^{\prime}|} 
\]

and we find the 2D lensing potential

\[\psi(\vec{\theta})   = - \frac{2 G D_{ds}}{D_d D_s c^2}   \int dz  \int  \frac{d^3\xi^{\prime} \rho(\vec{\xi}^{\prime})}{|\vec{\xi}-\vec{\xi}^{\prime}|}
 =  -  \sum_i \frac{2 G M_i D_{is} }{D_s D_i c^2}   \left[  \sinh^{-1}  { |z -D_i| \over D_i |\vec{\theta}-\vec{\theta}_i |  }  \right ] |_{D_i}^{D_s}  + |_{D_i}^{0}
\]

Here we assumed the lens is a collection of point masses \( M_i \) at angular coordinates \( \vec{\theta}_i \) and distances \[ z=D_i\]  
Use \[ \sinh^{-1} 1/x = \ln(1/x + \sqrt{1/x^2+1}) \approx -\ln(x/2) \] for very small \(x \) we find

\[
\psi(\vec{\theta})   \approx   \sum_i  \frac{2 GM_i D_{is} }{D_s D_i c^2}   \left[   \ln\left( { |\vec{\theta}-\vec{\theta}_i |^2 \over 4}  { D_i \over D_{is} } \right)     \right]
\]

One can  compute the ''convergence'' by applying the 2D Laplacian of the 2D lensing potential

\[
\kappa(\vec{\theta}) = \frac{1}{2} \nabla_{\vec{\theta}}^2 \psi(\vec{\theta}) =  \frac{4\pi G D_{ds}D_d} {c^2 D_s} \int dz \rho( D_d \vec{\theta},z) 
== {\Sigma \over \Sigma_{cr} } == \sum_i { 4\pi G M_i D_{is} \over c^2 D_i D_s}  \delta(\vec{\theta}-\vec{\theta}_i)
\]

in agreement with earlier definition \[ \kappa(\vec{\theta}) =  {\Sigma \over \Sigma_{cr} }\] as the ratio of projected density with the critical density.  
Here we used \[ \nabla^2 1/r = - 4 \pi \delta(r) \] and \[  \nabla_{\vec{\theta}} = D_d \nabla \]

We can also confirm the previously defined reduced deflection angle

\[
\vec{\theta} -\vec{\beta}   =  \nabla_{\vec{\theta}} \psi(\vec{\theta}) =  \sum_i   {  \theta_{Ei}^2  \over |\vec{\theta}-\vec{\theta}_i |} , ~ 
\pi \theta_{Ei}^2 \equiv  {4 \pi GM_i D_{is}  \over c^2 D_s  D_i } 
\]

where \( \theta_{Ei} \) is the so-called Einstein angular radius of a point lens \(Mi \).   For a single point lens at the origin we recover the standard result 
that there will be two images at the two solutions of the essentially quadratic equation

\[ \vec{\theta} -\vec{\beta}   =   {\theta_{E}^2  \over |\vec{\theta} |} \]

The amplification matrix can be obtained by double derivatives of the dimensionless time delay

\[
A_{ij} = {\partial \beta_j \over \partial \theta_i} = {\partial \tau \over \partial \theta_i \partial \theta_j } = \delta_{ij} -  {\partial \psi \over \partial \theta_i \partial \theta_j } 
= \left[\begin{array}{ c c } 1-\kappa -\gamma_1 & \gamma_2 \\ \gamma_2 & 1-\kappa +\gamma_1 \end{array}\right]  \]

where we have define the derivatives

\[ \kappa = {\partial \psi \over 2 \partial \theta_1 \partial \theta_1 } +  {\partial \psi \over 2\partial \theta_2 \partial \theta_2 } , 
~ \gamma_1 \equiv  {\partial \psi \over 2 \partial \theta_1 \partial \theta_1 } -  {\partial \psi \over 2\partial \theta_2 \partial \theta_2 } ,   
~ \gamma_2 \equiv  {\partial \psi \over \partial \theta_1 \partial \theta_2 }    \]

which takes the meaning of convergence and shear.   The amplification is the inverse of the Jacobian

\[   A = 1/det(A_{ij}) = {1 \over (1-\kappa)^2 -\gamma_1^2 -\gamma_2^2}  \]

where a positive A means either a maxima or a minima, and a negative A means a saddle point in the arrival surface.

For a single point lens, one can show (albeit a lengthy calculation) that

\[ \kappa =0, ~ \gamma = \sqrt{\gamma_1^2 + \gamma_2^2} = {\theta_E^2 \over |\theta|^2}, ~ \theta_E^2= {4GM D_{ds} \over c^2 D_dD_s}
\]

So the amplification of a point lens is given by

\[
A = \left( 1 - {\theta_E^4 \over \theta^4} \right)^{-1}
\]

Note A diverges for images at the Einstein radius \( \theta_E \).

In cases there are multiple point lenses plus a smooth background of (dark) particles of surface density \(\Sigma_{\rm cr} \kappa_{\rm smooth}\), the time arrival surface is

\[
\psi(\vec{\theta})   \approx   {1 \over 2} \kappa_{\rm smooth} |\theta|^2 +  \sum_i  \theta_E^2  \left[   \ln\left( { |\vec{\theta}-\vec{\theta}_i |^2 \over 4}  { D_d \over D_{ds} } \right)     \right] 
\]

To compute the amplification, e.g., at the origin (0,0), due to identical point masses distributed at \((\theta_{xi},\theta_{yi} ) \) 
we have to add up the total shear, and include a convergence of the smooth background, 
\[
A = \left[ (1 - \kappa_{\rm smooth})^2  
                  - \left( \sum_i {  (\theta_{xi}^2 - \theta_{yi}^2 ) \theta_E^2 \over (\theta_{xi}^2 + \theta_{yi}^2)^2 }\right) ^2  
                  - \left( \sum_i {  (2 \theta_{xi} \theta_{yi}) \theta_E^2 \over (\theta_{xi}^2 + \theta_{yi}^2)^2 }  \right)^2   \right] ^{-1}
\]

This generally creates a network of critical curves, lines connecting image points of infinite amplification.

\end{document}

### 2013-12-12 text/tex/intro.tex
editing text

@@
 \section{Introduction}
-Gravitational lenses (GLs), predicted by General Relativity (GR), can be used as a tool by astronomers to examine many properties of the cosmos.
-They allow for example to estimate the masses and their distribution of galaxies, automatically including the hard to grasp dark matter.
-Further more, they allow an estimate on cosmological parameters like the Hubble constant \citep{Saha2006} and the mass density \needcite.
 
-%To achieve this, one needs to find / identify lenses and in a second step model those lenses to get involved parameters.
+Our present picture of the formation of the largest structures in the
+Universe --- galaxies and clusters of galaxies --- is that
+gravitational instabilities (whose early stages are observable as
+fluctuations of the microwave background) caused collapse into
+gravitationally bound structures.  Dark matter would have dominated
+the initial collapse, but once deeps wells in the gravitational
+potential appeared, gas would have fallen into them, while radiating
+away its gravitational potential energy through atomic processes, to
+form the first stars and galaxies.  With time, galaxies and clusters
+would continue to merge and change their form.  This rough outline is
+well-established now, but many puzzles remain to be solved before our
+understanding is satisfactory.  To progress towards solving these, we
+need more knowledge of dark matter.  And to study dark matter,
+gravitational lensing is attractive, because it depends only on mass.
 
-How to find gravitational lenses? Huge amount of image data from surveys that needs to be processed.
-Even more coming in the future with new surveys \needcite
-Robotic processing has been suggested \needcite, and tested \needcite, but has failed so far to be convincing \needcite.
-Another suggestion involves humans, volunteers \footnote{A volunteer is considered everyone that has no background in astrophysics.}.
-\sw uses this approach with great success so far.\needcite
+There are many manifestions of gravitational lensing in astrophysics,
+but the most spectacular are multiple images.  Gravitational lensing
+depends on the sky-projection mass density, and this fact has a
+counter-intuitive consequence: nearby galaxies do not produce multiple
+images, because their projected density is not high enough, but beyond
+100~Mpc or so, most galaxies are potential lenses.  Even so, for a
+potential lens to really lens, a background source in the lensed
+region is required.  So, in practice galaxies are very rarely observed
+to lens.  The same tendency applies to clusters of galaxies, but since
+clusters are bigger, they are more likely to lens.
 
-Next step involves modeling. that needs advanced knowledge and takes a lot of time.
-Too much time to be done by astronomers themselves as the results from new surveys and identifiers like \sw come in.
-So there is a demand of a means to model a great number of identified lens data, that scales with increasing volume of data.
+In lensing galaxies, the observable lensing usually involves a single
+object multiply imaged.  It is therefore desirable to have large
+numbers of lenses, and this is the strategy researchers follow.  The
+largest single study so far uses 58 lenses \citep{2009ApJ...703L..51K}.
+Other studies have compared lens models in more detail with stellar
+mass \citep{2011ApJ...740...97L} or used time-delay information from
+lensed quasars to infer the Hubble time
+\citep{2008ApJ...679...17C,2010ApJ...712.1378P}.
+
+The trend is clear: researchers want to find many more lenses and
+model them in more sophisticated ways.
+
+A good summary of the observational situation is offered by
+Figure~\ref{fig:masterlens}, which shows known secure multiply-imaging
+lenses.  The non-uniformity on the sky is not intrinsic, it just
+indicates the density of deep surveys up to 2013.  At HST resolution,
+of order 1~square degree of sky must be searched to find a lens.
+Older ground-based surveys yield a lens per roughly 10~square degree
+of survey area.  The large surveys starting now
+(DES\footnote{\tt http://pan-starrs.ifa.hawaii.edu} or
+PanStarrs\footnote{\tt http://www.darkenergysurvey.org}) the expected
+yield is something in between.  Given their unprecedented survey area,
+these are likely to yield thousands of lenses.  Over the 2020s,
+LSST\footnote{\tt http://www.lsst.org/} can be expected to give us ten
+thousand lenses.
+
+The question now arises: can lenses be found and modelled
+automatically in large surveys?  Work using software robots
+\citep{2009ApJ...694..924M} gives some interesting and unexpected
+results.  In clean lensing system in uncrowded fields with high
+signal-to-noise, the robots to very well.  In most situations,
+however, robots miss lenses (low completeness) or contaminate the
+results with non-lenses (low purity).  Robots can be made to
+prioritize completeness or purity, but they cannot deliver both.
+
+In response to the lessons learned from automated lens searches, the
+{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org} has
+(Saha is a co-investigator) has been launched.  {\em Spacewarps\/} is
+part of the {\em Zooniverse\/} family of citizen-science
+projects,\footnote{\tt http://www.zooniverse.org} where members of the
+public are invited to analyze different kinds of scientific data which
+are too difficult for robots and too large for specialists.  In {\em
+  Spacewarps\/} itself, since the launch in May 2013, $\sim100$ square
+degrees (or a quarter of a percent of the sky) has been examined.  The
+survey is divided into patches of $400\times 400$ pixels, each of
+which is seen by ten volunteers.  Simulated lenses are mixed in with
+the data, both to help train volunteers on what to look for, and to
+estimate completeness and purity.  A list of candidates is being
+processed, but intermediate results show some very good candidates, as
+well re-discoveries of a few known lenses.
+
+These encouraging results now raise the question: could modelling of
+the lenses also be done by volunteers?  Modelling is a much much more
+difficult task than searching for lens candidates, as it requires some
+expert knowledge.  Nonetheless, a subset of the {\em Spacewarps\/}
+volunteers are quite experienced from earlier projects, having spent a
+thousand hours or more with data.  Some of these experienced
+volunteers are very interested in more demanding projects.  This also
+is a general trend \citep[cf.][]{Khatib22112011}.  The present project
+then suggests itself.
 

### 2013-12-12 text/tex/outlook.tex
editing text

@@
-\section{Extensions needed} \label{sec:todo}
+\section{Outlook} \label{sec:todo}
+
+The lens-modelling challenge indicates that the Spacewarps collective
+is good, not only for identifying lens candidates for follow-up, but
+modelling candidates as well.
+
+There is, however, plenty of room for improvement.
+
+First, the particular modelling strategy implemented is not the only
+one possible.  \spl requires modellers to characterise the overall
+image structure in abstract terms based on Fermat's principle, and the
+placement of a mass distributions is done by the computer.  In other
+modelling tools, the user puts down a trial mass distribution and has
+the machine refine it.  A few of these modelling programs have also
+been designed with citizen science in mind, and would be interesting
+the in Spacewarps environment.
+
+Second, \spl needs some enhancements.
+
+One, model the source as well.
+
+Another, lenses with more than one maximum can be allowed for.  See
+Figure 5c in \citep{2001ApJ...557..594R} and Figure 4b in
+\cite{2003ApJ...590...39K}.
+
+Other modelling approaches also possible.
 
@@
 Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models etc...)
-
-Lenses with more than one maximum can be allowed for.  See Figure 5c
-in \citep{2001ApJ...557..594R} and Figure 4b in
-\cite{2003ApJ...590...39K}.
-
-Supply several images from several bands.

### 2013-12-12 text/tex/results.tex

@@
 The modellers proffered a total of 129 models for the 29 sims in the
-challenge.  Figures \ref{fig:6941} to \ref{fig:6915} show eight of
+challenge.  Figures \ref{fig:6941} to \ref{fig:6919} show eight of
 models in some detail.
@@
 
+Figure \ref{fig:6990} shows an example of an arc that has split into
+three images.  This kind of configuration, with a counter-image close
+to the lensing galaxy and a more distant arc/triplet on the other
+side, generically arises from an elongated mass distribution when the
+source is displaced along the elongated direction.
+
+Figure \ref{fig:6915} shows another quad.  This kind of configuration
+arises when the mass is elongated and the source is displaced at an
+angle to the elongation.
+
 Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
@@
 
-Figure \ref{fig:6990} shows a long-axis quad.
-
-Figure \ref{fig:6919} shows a short-axis quad.
-
-Figure \ref{fig:6915} shows an inclined quad.
+To conclude this set of examples, Figure \ref{fig:6919} shows another
+type of quad.  Actually, in this case the brightest part of the source
+is only doubly imaged, but the source extends into a region that
+produces four images.  We can also see from the real arrival-time
+surface that a point source is a double on the verge of splitting into
+a quad.  The modeller interpreted the system as a quad.  The
+appearance of the arcs, shown in the bottom panel of Figure
+\ref{fig:input-spag}, looks like an arc and counter-image such as
+discussed with Figure \ref{fig:6990} above.  But there is an important
+difference: the long arc is closer to the galaxy, as if the arc and
+counter-image have swapped roles.  This configuration arises if the
+source displacement is perpendicular to the long axis of the lensing
+mass.
 
@@
 
+The second test T2 was to compare the desired results of a modeling process, the mass distribution of the lens $\kappa(x, y)$.
+To get a means of comparing the simulated data (sims) to the generated modeled data (models), the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ for both was calculated.
+The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between the sims and the models.
+We also let an expert model three selected systems to compare the results from volunteers to those of a professional.
+
+\todo{!} Should I write down prior expectation for T2 too?
+
 To get a means of comparing the simulated data (sims) to the generated modeled data (models), the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ for both was calculated.

### 2013-12-12 text/tex/results.tex
editing sec3

@@
 
-To estimate the performance of the volunteers and the quality of the generated models, two tests T1 and T2 were done.
-
-The second test T2 was to compare the desired results of a modeling process, the mass distribution of the lens $\kappa(x, y)$.
-
 The \sw sims were generated by AM, in consultation with PM and AV.
+To estimate the performance of the volunteers and the quality of the generated models, two analyses  were done.
+The first tested the correct identification and ordering of lensed images.
+The second compared the mass distribution of the lens $\kappa(x, y)$ of the generated models to the mass distribution of the simulations.
 

### 2013-12-12 text/tex/abstract.tex
more edits

@@
 \begin{abstract}
-In Spacewarps **, volunteers are invited to search sky surveys for lens
-candidates.  Here we report on a system that allows experienced
+In Spacewarps **, volunteers are invited to search sky surveys for
+lens candidates.  Here we report on a system that allows experienced
 volunteers to model lens candidates.  A sample of 29 simulated lenses
-were modeled, each multiple times.  The quality of the models was
-then examined in different ways.  In particular, (i)~the image
-parities and time orderings were identified correctly in most cases,
-but not always; (ii)~the mean convergence (equivalent to the enclosed
-mass) was quite well constrained in the image region; (iii)~the
-performance of experienced volunteers was comparable to professionals.
+were modeled, each multiple times.  The quality of the models was then
+examined in different ways.  In particular: (i)~the image parities and
+time orderings were identified correctly in most cases, but not
+always; (ii)~the mean convergence (equivalent to the enclosed mass)
+was quite well constrained in the image region; (iii)~the performance
+of experienced volunteers was comparable to professionals.
 \end{abstract}

### 2013-12-12 text/tex/outlook.tex
more edits

@@
 
-One, model the source as well.
+\begin{itemize}
+\item Currently, \spl does not attempt to model the source shape; the
+  user identifies the brightest points on the image, and these are
+  taken as the images of a point-like source, whose positions must be
+  reproduced exactly.  For generating a synthetic image, a conical
+  source profile is assumed.  Fitting for the source profile to
+  optimize resemblance to the observed lensed image after the lens
+  model has been generated, is algorithmically straightforward and
+  planned to be implemented.  This would alleviate another problem
+  with \spl, which is that there is no numerical figure of merit, and
+  assessment of a model is a judgement call based on the synthetic
+  image, and on whether the mass distribution and the arrival-time
+  surface show suspect features.
+\item Another current limitation in \spl is that the lens is assumed
+  to be dominated by one galaxy, which puts most galaxy-group lenses
+  beyond the reach of the modeller.  Since complicated group lenses
+  are some of the most interesting candidates present, removing this
+  limitation is most desirable.  From the users' point of view, it
+  would mean that spaghetti contours with more than one maximum can be
+  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
+  and Figure 4b in \cite{2003ApJ...590...39K}.
+\item Currently, a single false-colour composite is used as the data.
+  An option could be added to use all available filters, individually
+  or in cmbination, as the user wishes.
+\end{itemize}
 
-Another, lenses with more than one maximum can be allowed for.  See
-Figure 5c in \citep{2001ApJ...557..594R} and Figure 4b in
-\cite{2003ApJ...590...39K}.
+The third desirable avenue of improvement is to facilitate
+collaborative work.
 
-Other modelling approaches also possible.
-
-As seen in \secref{tests.t1}, volunteers fail mostly in two situations:
-when to identify an arc like structure while placing the points
-and in 5 image configurations, that are quite symmetric, then to identify the correct ordering of the points.
-
-The first can be improved by additional images from different filters (?)
-Will be implemented with a next major \spl upgrade.
-
-The second can be made less by better training (introduce the ruler), may be more filters help too?
-
-We hope that those problems are also taken care of when staring collaborative modeling. (Already in progress)
-Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models etc...)
+\begin{itemize}
+\item As mentioned above, the option of revising an already-archived
+  model is already available.  Desired now are tools for comparing
+  different models of a given system, both visually and through
+  different statistical measures.  As evidenced by a current
+  collaborative modelling effort, a particularly interesting candidate
+  can lead to an extended discussion and dozens of models.
+\item Better tutorial materials are also needed, and this would
+  address some of the problem areas found in the modelling challenge.
+  For example, we saw in \secref{tests.t1} that volunteers are most
+  prone to making errors in two situations: when to identify an arc-
+  like structure while placing the points, and in identifying the
+  correct ordering of the points in nearly-symmetric configurations.
+\end{itemize}

### 2013-12-12 text/tex/results.tex
editing sec3
editing sec3

more edits

editing sec3 - added proper def of theta_e_sim, theta_e

editing sec3

@@
 The sims were produced using {\tt gravlens}
-\citep{2001astro.ph..2341K,2001astro.ph..2340K}.  The were of three
+\citep{2001astro.ph..2341K,2001astro.ph..2340K}.  There were of three
 kinds, as follows.
@@
   \item Resembling cluster lenses: having a source as above, but a
-    more complicated lens, with one dominant elliptical SIE and and
+    more complicated lens, with one dominant elliptical SIE and
     one or more perturbing elliptical SIEs, plus a circular NFW
@@
 them to assess the model.  These derive from the mean of an ensemble
-of 200 models generated by \spl.  The model mass distribution is
+of 200 models generated by \spl.  The model mass distribution (left mid) is
 returned as a contour map superimposed on an intensity map. A fairly
@@
 checkboard pattern in the mass map usually indicates a bad model.  The
-model arrival-time contours.  The special contour passing through a
+models arrival-time contours are shown in the right middle panel.
+  The special contour passing through a
 saddle point is shown in black: this is the model version of the
 spaghetti sketch provided by the user.  Also provided is a synthetic
-image of the lensed source.  For this, \spl assumes a simple conical
+image of the lensed source in the bottom right panel.
+  For this, \spl assumes a simple conical
 source profile.  The user can change the contrast level on the image,
-which (though it is not saved) amounts to to adjusting the width of
+which (though it is not saved) amounts to adjusting the width of
 the cone.  Finally, to the lower left of the figure, we have a
@@
 The red curve is the true value, and where it crosses unity (dotted
-horizonal line) is the notional Einstein radius.  The two blue curves
+horizonal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
+  The two blue curves
 are the minimal and maximal mean enclosed $\kappa$ from the internal
@@
 is the confidence region from the modelling.  As we see, the shaded
-blue is slightly above the red curve.  In summary: the identification
+blue is slightly above the red curve. 
+The Einstein Radius $\Theta_\text{E}$ of the model is estimated crossing the
+mean enclosed $\kappa$ (not plotted) with unity.
+  In summary: the identification
 of minimum and saddle point is correct, but the estimated Einstein
@@
 
-T1 tested the volunteers ability to reconstruct the arrival time surface given a survey image containing a sim.
+A first manual evaluation tested the volunteers ability to reconstruct the arrival time surface given a survey image containing a sim.
 This task consists of two parts.
-First, the correct identification and location of lensed images (T1a).
-Second the correct ordering for the identified lensed images in respect of the arrival time (T1b).
+First to correctly identify and locate the lensed images.
+Second the correct ordering for the identified lensed images in respect of the arrival time.
 
-While we expected T1a to be trivial, given the nature of the survey images and the success of \sw, we expect T1b to be more difficult.
-T1b tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
-While we can provide the volunteers with some general rules of thumb, T1b involves imagination and guessing and therefore training could improve their skills in a later stage.
+While we expected the identification of lensed images to be trivial, given the nature of the survey images and the success of \sw, we expect the correct ordering to be more difficult.
+This tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
+While we can provide the volunteers with some general rules of thumb, ordering involves imagination and guessing and therefore training could improve results in a later stage.
 
-T1 was also designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
+This first test was designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
 
-The evaluation of the volunteers performance for T1 was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot generated using simulation parameters.
+The evaluation of the volunteers performance was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot generated using simulation parameters.
 % \Figref{output_compare} shows the setup used to evaluate the models.
 
-To evaluate T1a and T1b, each model was evaluated with two binary tests.
-The first is passed, if all the images have been identified and are approx. within $\pm0.05\cdot\text{imgage width}$.
-The second test is passed, if those identified points have the right parity, ordering with respect to arrival time.
+The images of the system are considered to be identified correcly, if all the images have been identified and are approx. within $\pm0.05\cdot\text{imgage width}$.
+The parity is considered correct, if those identified points have the right ordering with respect to arrival time.
 
-Additionally, ten types of errors (E01 -- E10, listed in \tabref{stats}) that could occur were analyzed, where each model could contain multiple of those.
+Additionally, ten types of errors (labeled E01 -- E10, listed in \tabref{stats}) that occurred were identified.
+Each generated model could contain more than one error.
 
@@
 
-In \tabref{stats} a summary of this evaluation is presented.
-We conclude that the volunteers are performing very well identifying and positioning images (T1a), with a performance of 92\% (R1, p=0.92).
+\tabref{stats} presents a summary of this evaluation.
+We conclude that the volunteers are performing very well identifying and positioning images, with a performance of 92\% (R1, p=0.92).
 Most of the problems where due to unclear arc-like structures (E01, p=0.18; E04, p=0.03; E08, p=0.04).
@@
 
-\todo{add more detail?} Add more details like total number of images detected / tot am images (rightPlace fraction)??
-
-The assignment of the parity of the images (T1b) was a more difficult task.
+The assignment of the parity of the images was a more difficult task.
 In 59\% (R2, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
-Most of the failures are due to E06, with N=38, p=32\%, followed by E05 (PI) with N=7, p=6\%.
-For an example of E06, see \figref{7022}.
-It basically describes a situation, where the minima and saddle points of a five image configuration were exchanged (rotated by $\pm90\dgr$).
-E05 describes the situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$).
+Most of the failures are due to E06 (N=38, p=32\%), followed by E05 (N=7, p=6\%).
+E06 describes a situation, where the minima and saddle points of a five image configuration were exchanged (rotated by $\pm90\dgr$), see \figref{7022} for an example.
+E05 describes a situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$).
 While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
@@
 
-The second test T2 was to compare the desired results of a modeling process, the mass distribution of the lens $\kappa(x, y)$.
-To get a means of comparing the simulated data (sims) to the generated modeled data (models), the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ for both was calculated.
-The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between the sims and the models.
-We also let an expert model three selected systems to compare the results from volunteers to those of a professional.
+The second test was to compare the mass distribution of the lens $\kappa(x, y)$ of the models and the sims.
+To get a means of comparing the sims to the models, the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ was calculated for both.
+The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between a sim and a model.
+We also let an expert (PS) model three selected systems to compare the results from volunteers to those of a professional.
 
-\todo{!} Should I write down prior expectation for T2 too?
-
-To get a means of comparing the simulated data (sims) to the generated modeled data (models), the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ for both was calculated.
-The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between the sims and the models.
-We also let an expert model three selected systems to compare the results from volunteers to those of a professional.
-
-\todo{!} Should I write down prior expectation for T2 too?
-
-To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] directly generated in the modeling process.
+To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
 From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate $\theta_E$.
@@
 One of the reasons for this is, that it's hard to get the center of the lens on spot.
-An offset leads to a a flatter mass profile for the model compared to the simulation.
+An offset leads to a flatter mass profile for the model compared to the simulation.
 

### 2013-12-12 text/tex/theory.tex
editing sec3
editing sec3

more edits

editing sec3 - added proper def of theta_e_sim, theta_e

editing sec3

@@
 solving them.  Incidentally, $\kappa$ has a second meaning, as well as
-being a dimensionless form of the the projected density: it also
+being a dimensionless form of the projected density: it also
 measures how a bundle of light rays is brought together by

### 2013-12-12 text/tex/results.tex
final changes

@@
 
-The second test was to compare the mass distribution of the lens $\kappa(x, y)$ of the models and the sims.
-To get a means of comparing the sims to the models, the total convergence, also called enclosed mass $\kappa_{\text{encl}}(r)$ was calculated for both.
-The Einstein radius $\Theta_E$ is defined by $\kappa_{\text{encl}}(\Theta_E)=1$ and gives a number that allows the comparison between a sim and a model.
+The second test was to compare the mass distributions of the lens $\kappa(x, y)$ given by the sims and generated by the volunteers.
+To get a means of comparing the sims to the models, the total convergence\footnote{often called enclosed mass} $\kappa_{\text{encl}}(r)$ was calculated for both.
+The Einstein radius $\Theta_\text{E}$ is defined by $\kappa_{\text{encl}}(\Theta_\text{E})=1$ and gives a number that allows a rough comparison between a sim and a model.
 We also let an expert (PS) model three selected systems to compare the results from volunteers to those of a professional.
@@
 To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
-From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate $\theta_E$.
-To estimate the errors, also the extremal models are used to estimate a lower and upper limit for $\theta_E$.
+From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate \ERg.
+To estimate the errors, the extremal models are used to estimate a lower and upper limit for \ERg.
 These results can be seen in \figref{ER_all_models}.
-This figure shows that this technique of estimating the error minimizes the error significantly and should be improved for further analysis.
+This figure shows that this technique of estimating the error using the ensemble of all models underestimates the error significantly and should be improved for further analysis.
 
-In \figref{ER_per_sim} can be seen, that the calculated \ERf of the models tend to be too high.
+\Figref{ER_per_sim} shows that the calculated \ERf of the models tend to be too high.
 The overshoot varies from around 0.2 to 0.4 for good models.
-One of the reasons for this is, that it's hard to get the center of the lens on spot.
+One of the reasons for this is that it is hard to get the center of the lens on spot.
 An offset leads to a flatter mass profile for the model compared to the simulation.
@@
 Looking at the results for \ERg for those models in \figref{ER_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses, offset) is comparable.
-Note that the models of \asw{0vqg} with \ERg[, rel] around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
+%Note that the models of \asw{0vqg} with \ERg[, rel] around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
 

### 2013-12-14 text/tex/abstract.tex
editing typos
typos in abstract, theory, spaghetti, outlook, appendix

@@
 \begin{abstract}
-In Spacewarps **, volunteers are invited to search sky surveys for
+In Spacewarps **, volunteers are invited to search CFHTLS sky survey for
 lens candidates.  Here we report on a system that allows experienced
@@
 examined in different ways.  In particular: (i)~the image parities and
-time orderings were identified correctly in most cases, but not
+time arrivals were identified correctly in most cases, but not
 always; (ii)~the mean convergence (equivalent to the enclosed mass)

### 2013-12-14 text/tex/outlook.tex
editing typos
typos in abstract, theory, spaghetti, outlook, appendix

--- 2013-11-25 text/tex/outlook.tex
@@
-\section{Extensions needed} \label{sec:todo}
+\section{Outlook} \label{sec:todo}
 
-As seen in \secref{tests.t1}, volunteers fail mostly in two situations:
-when to identify an arc like structure while placing the points
-and in 5 image configurations, that are quite symmetric, then to identify the correct ordering of the points.
+The lens-modeling challenge indicates that the Spacewarps collective
+is good, not only for identifying lens candidates for follow-up, but
+modeling candidates as well.
 
-The first can be improved by additional images from different filters (?)
-Will be implemented with a next major \spl upgrade.
+There is, however, plenty of room for improvement.
 
-The second can be made less by better training (introduce the ruler), may be more filters help too?
+First, the particular modeling strategy implemented is not the only
+one possible.  \spl requires modelers to characterize the overall
+image structure in abstract terms based on Fermat's principle, and the
+placement of mass distributions is done by the computer.  In other
+modeling tools, the user puts down a trial mass distribution and has
+the machine refine it.  A few of these modeling programs have also
+been designed with citizen science in mind, and would be interesting
+in the Spacewarps environment.
 
-We hope that those problems are also taken care of when staring collaborative modeling. (Already in progress)
-Basics are available ('revise functionality') but better community tools would help (having overview, comparison between different models etc...)
+Second, \spl needs some enhancements.
 
-Lenses with more than one maximum can be allowed for.  See Figure 5c
-in \citep{2001ApJ...557..594R} and Figure 4b in
-\cite{2003ApJ...590...39K}.
+\begin{itemize}
+\item Currently, \spl does not attempt to model the source shape; the
+  user identifies the brightest points on the image, and these are
+  taken as images of a point-like source, whose positions must be
+  reproduced exactly. For generating a synthetic image, a conical
+  source profile is assumed. Fitting for the source profile to
+  optimize resemblance to the observed lensed image after the lens
+  model has been generated, is algorithmically straightforward and
+  planned to be implemented.  This would alleviate another problem
+  with \spl, which is that there is no numerical figure of merit, and
+  assessment of a model is a judgment call based on the synthetic
+  image, and on whether the mass distribution and the arrival-time
+  surface show suspect features.
+\item Another current limitation in \spl is that the lens is assumed
+  to be dominated by one galaxy, which puts most galaxy-group lenses
+  beyond the reach of the modeler. Since complicated group lenses
+  are some of the most interesting candidates present, removing this
+  limitation is most desirable.  From the users' point of view, it
+  would mean that spaghetti contours with more than one maximum can be
+  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
+  and Figure 4b in \cite{2003ApJ...590...39K}.
+\item Currently, a single false-color composite is used as the data.
+  An option could be added to use all available filters, individually
+  or in combination, at the user wishes.
+\end{itemize}
 
-Supply several images from several bands.
+The third desirable avenue of improvement is to facilitate
+collaborative work.
+
+\begin{itemize}
+\item As mentioned above, the option of revising an already-archived
+  model is already available.  Desired now are tools for comparing
+  different models of a given system, both visually and through
+  different statistical measures.  As evidenced by a current
+  collaborative modeling effort, a particularly interesting candidate
+  can lead to an extended discussion and dozens of models.
+\item Better tutorial materials are also needed, and this would
+  address some of the problem areas found in the modeling challenge.
+  For example, we saw in \secref{tests.t1} that volunteers are most
+  prone to making errors in two situations: when in identifying an arc-
+  like structure while placing the points, and in identifying the
+  correct ordering of the points in nearly-symmetric configurations.
+\end{itemize}

### 2013-12-14 text/tex/theory.tex
editing typos
typos in abstract, theory, spaghetti, outlook, appendix

@@
-\section{A lens modelling program}
+\section{A lens modeling program}
 
@@
 motivation cited by citizen-science volunteers.  So for a lens
-modeller in a citizen-science project, nobody wants that it should
+modeler in a citizen-science project, nobody wants that it should
 sacrifice scientific usefulness in order to be fun to use.  That said,
@@
 plane.  Let $x,y$ be coordinates on this plane.  These coordinates
-measure ordinary physical lengths (metres, etc) on the lens plane.
+measure ordinary physical lengths (meters, etc) on the lens plane.
 Let there be a point source of light at $x=0,y=0$ but far behind the
@@
 distribution $\kappa(x,y)$ is circular, $A_{\rm grav}(x,y)$ will have
-a peak at the centre of that circle.  But if the mass is not precisely
+a peak at the center of that circle.  But if the mass is not precisely
 in front of the source, the arrival-time surface will not be circular
@@
 original point source behind $x=0,y=0$ and move the lens slightly.
-The contours of constant arrival time will, naturally, move slightly,
+The contours of constant arrival time will naturally move slightly,
 and so will the images.  The movement of the contours will be most
-noticeable where the contours are far apart, that is, where the
+noticeable where the contours are far apart, that is where the
 arrival-time surface is nearly flat.  As is evident from
@@
 apart.  In other words, the image is highly magnified.  In summary,
-lower curvature in the arrival-time surface for a point source implied
+lower curvature in the arrival-time surface for a point source implies
 larger magnification of an extended source.  Conversely, where the
@@
 is evident.  In other cases the X contour passing through the saddle
-point is not itself plotted, but we can tell the other contours where
+point is not itself plotted, but we can tell by the other contours where
 the X would be.  These self-crossing contours will play a special role

### 2013-12-14 text/tex/intro.tex
editing sec3
editing sec3

more edits

editing sec3 - added proper def of theta_e_sim, theta_e

editing sec3

@@
 gravitational instabilities (whose early stages are observable as
-fluctuations of the microwave background) caused collapse into
+fluctuations of the microwave background) caused a collapse into
 gravitationally bound structures.  Dark matter would have dominated
-the initial collapse, but once deeps wells in the gravitational
-potential appeared, gas would have fallen into them, while radiating
-away its gravitational potential energy through atomic processes, to
+the initial collapse, but once deep wells in the gravitational
+potential appeared, gas would have fallen into them while radiating
+away its gravitational potential energy through atomic processes to
 form the first stars and galaxies.  With time, galaxies and clusters
@@
 
-There are many manifestions of gravitational lensing in astrophysics,
+There are many manifestations of gravitational lensing in astrophysics,
 but the most spectacular are multiple images.  Gravitational lensing
@@
 100~Mpc or so, most galaxies are potential lenses.  Even so, for a
-potential lens to really lens, a background source in the lensed
+potential lens to really act as a lens, a background source in the close-by
 region is required.  So, in practice galaxies are very rarely observed
-to lens.  The same tendency applies to clusters of galaxies, but since
-clusters are bigger, they are more likely to lens.
+as lens.  The same tendency applies to clusters of galaxies, but since
+clusters are bigger, they are more likely lens-like.
 
-In lensing galaxies, the observable lensing usually involves a single
+In lensing galaxies, the observable features usually involve a single
 object multiply imaged.  It is therefore desirable to have large
-numbers of lenses, and this is the strategy researchers follow.  The
+numbers of lenses, and this is the strategy researchers are following.  The
 largest single study so far uses 58 lenses \citep{2009ApJ...703L..51K}.
@@
 Older ground-based surveys yield a lens per roughly 10~square degree
-of survey area.  The large surveys starting now
+of survey area.  By the large surveys starting now
 (DES\footnote{\tt http://pan-starrs.ifa.hawaii.edu} or
@@
 
-The question now arises: can lenses be found and modelled
+The question now arises: can lenses be found and modeled
 automatically in large surveys?  Work using software robots
@@
 results.  In clean lensing system in uncrowded fields with high
-signal-to-noise, the robots to very well.  In most situations,
+signal-to-noise, the robots do very well.  In most situations,
 however, robots miss lenses (low completeness) or contaminate the
@@
 In response to the lessons learned from automated lens searches, the
-{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org} has
+{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org} 
 (Saha is a co-investigator) has been launched.  {\em Spacewarps\/} is
@@
 degrees (or a quarter of a percent of the sky) has been examined.  The
-survey is divided into patches of $400\times 400$ pixels, each of
+survey is divided into patches of $440\times 440$ pixels, each of
 which is seen by ten volunteers.  Simulated lenses are mixed in with
@@
 
-These encouraging results now raise the question: could modelling of
-the lenses also be done by volunteers?  Modelling is a much much more
+These encouraging results now raise the question: could modeling of
+the lenses also be done by volunteers?  Modeling is a much more
 difficult task than searching for lens candidates, as it requires some
-expert knowledge.  Nonetheless, a subset of the {\em Spacewarps\/}
-volunteers are quite experienced from earlier projects, having spent a
+expert knowledge.  Nonetheless, a panel of the {\em Spacewarps\/}
+volunteers is quite experienced from earlier projects, having spent a
 thousand hours or more with data.  Some of these experienced

### 2013-12-14 text/tex/results.tex

@@
 The \sw sims were generated by AM, in consultation with PM and AV.
-To estimate the performance of the volunteers and the quality of the generated models, two analyses  were done.
-The first tested the correct identification and ordering of lensed images.
-The second compared the mass distribution of the lens $\kappa(x, y)$ of the generated models to the mass distribution of the simulations.
+To estimate the performance of the volunteers and the quality of the generated models, two analysis  were done.
+The first analysis tested the correct identification and ordering of lensed images.
+The second one compared the mass distribution of the lens $\kappa(x, y)$ of the generated models to the mass distribution of the simulations.
 
@@
 
-In the interest of blind testing, the information in this section was
+For the sake of blind testing, the information in this section was
 revealed neither to RK, while choosing the challenge set of 29 sims,
-nor to the modellers (volunteers EB, CC, CM, JO, JW and `expert
-modeller' PS) until the modelling stage was done.
+nor to the modelers (volunteers EB, CC, CM, JO, JW and `expert
+modeler' PS) until the modeling stage was done.
 
 The sims were produced using {\tt gravlens}
-\citep{2001astro.ph..2341K,2001astro.ph..2340K}.  There were of three
+\citep{2001astro.ph..2341K,2001astro.ph..2340K}. They were of three
 kinds, as follows.
@@
 
-The modellers proffered a total of 129 models for the 29 sims in the
+The modelers proposed a total of 129 models for the 29 sims in the
 challenge.  Figures \ref{fig:6941} to \ref{fig:6919} show eight of
@@
 information in these two panels was, of course, kept secret during the
-modelling challenge.  The panels in the middle row and at the lower
-right are the results that \spl returns to the modeller, in order for
+modeling challenge.  The panels in the middle row and at the lower
+right are the results that \spl returns to the modeler, in order for
 them to assess the model.  These derive from the mean of an ensemble
@@
 The red curve is the true value, and where it crosses unity (dotted
-horizonal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
+horizontal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
   The two blue curves
@@
 between the radii of the innermost and the outermost images --- this
-is the confidence region from the modelling.  As we see, the shaded
+is the confidence region from the modeling.  As we see, the shaded
 blue is slightly above the red curve. 
@@
 were, in fact, based on galaxies visible in the images --- but the
-modellers were not told in advance whether this was the case.  The
+modelers were not told in advance whether this was the case.  The
 model does not include any substructure, but otherwise is not bad.
@@
 mass is, however, none the worse --- the reason is probably that in a
-relatively symmetric image configuration, the Einstein radius is quite
+relatively symmetrical image configuration, the Einstein radius is quite
 well constrained by the images in a fairly model-independent way.
@@
 surface that a point source is a double on the verge of splitting into
-a quad.  The modeller interpreted the system as a quad.  The
+a quad.  The modeler interpreted the system as a quad.  The
 appearance of the arcs, shown in the bottom panel of Figure
@@
 While we expected the identification of lensed images to be trivial, given the nature of the survey images and the success of \sw, we expect the correct ordering to be more difficult.
-This tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
+This part tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
 While we can provide the volunteers with some general rules of thumb, ordering involves imagination and guessing and therefore training could improve results in a later stage.
@@
 
-The images of the system are considered to be identified correcly, if all the images have been identified and are approx. within $\pm0.05\cdot\text{imgage width}$.
+The images of the system are considered to be identified correctly, if all the images have been identified and are approximated within $\pm0.05\cdot\text{imgage width}$.
 The parity is considered correct, if those identified points have the right ordering with respect to arrival time.

### 2013-12-14 text/tex/spaghetti.tex
editing sec3
editing sec3

more edits

editing sec3 - added proper def of theta_e_sim, theta_e

editing sec3

@@
 In the second modeling step, once the user has sketched a proposed
-spaghetti configurations, \spl sends the input to its server-side
+spaghetti configuration, \spl sends the input to its server-side
 modeling engine, called GLASS.  The task of GLASS is to find a mass
@@
   Second, the direction of increasing density at any $(x,y)$ can point
-  at most $45^\circ$ away from from $(x_0,y_0)$.
-\item The lens must symmetric with respect to $180^\circ$ rotations
+  at most $45^\circ$ away from $(x_0,y_0)$.
+\item The lens must be symmetrical with respect to $180^\circ$ rotations
   about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the

### 2013-12-16 text/tex/outlook.tex
another small edit

@@
   surface show suspect features.
-\item Another current limitation in \spl is that the lens is assumed
-  to be dominated by one galaxy, which puts most galaxy-group lenses
-  beyond the reach of the modeler. Since complicated group lenses
-  are some of the most interesting candidates present, removing this
+\item \spl has a tendency to somewhat overestimate the Einstein radius
+  (evident from Figure \ref{fig:ER_per_sim}), and it is also apparent
+  that the models tend to be too shallow.  This possible explanation
+  is that, while the sims are steeply peaked at the centre, the
+  pixellated mass model fixes a comparatively large area near the
+  central at constant density.  One way to solve this problem would be
+  to introduce smaller pixels in the central region, thus enabled a
+  steeper centre.
+\item Another limitation so far in \spl is that the lens is assumed to
+  be dominated by one galaxy, which puts most galaxy-group lenses
+  beyond the reach of the modeler. Since complicated group lenses are
+  some of the most interesting candidates present, removing this
   limitation is most desirable.  From the users' point of view, it
@@
   and Figure 4b in \cite{2003ApJ...590...39K}.
-\item Currently, a single false-color composite is used as the data.
+\item At present, a single false-color composite is used as the data.
   An option could be added to use all available filters, individually

### 2014-01-12 text/ms.tex
finished setup
plaing around with contour plots
layout of intro, changes to scirpts, made class copatible to setlength
a few words on sl, added bib entries
plot gen script working
added results for test1, inkl autogenerated tables
ideas for results section
t1
further work
moar results
minor ajustments to plots and tables
small fixes on text, tables, plots
rearranged figure placement
suppressing figures
added pdf
Commented out images
testing 2
editing sec3
editing sec3

more edits

editing sec3 - added proper def of theta_e_sim, theta_e

editing sec3

@@
 \documentclass[12pt,preprint]{aastex}
-
-%%% PACKAGES %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \usepackage{float}    % exact placement of figures (using \begin{figure}[H])
-%\usepackage{caption}  % better captions for floats
+\usepackage{color}  % well... color...
+\usepackage{setspace}    % remove this pgk for final compile!! 
+\usepackage{subfigure}
 
-%%% DEFINITIONS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{longtable}   % those two to generate the 2 ol table
+\usepackage{multicol}
 
 \def\half{{\textstyle\frac12}}
+% makes it easier to have consistent writing, and easy change if needed
 \newcommand{\spl}{SpaghettiLens\xspace}
@@
 
-%%% DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% shortcut for einstein radius (Text Greek Full Math)
+\newcommand{\ER}{Einstein radius\xspace} % text
+\newcommand{\ERg}[1][]{$\Theta_{\text{E#1}}$\xspace} % er in textmode with greek
+\newcommand{\ERf}[1][]{Einstein radius $\Theta_\text{E#1}$\xspace} % full output
+\newcommand{\ERm}[1][]{\Theta_\text{E#1}} % mathmode with greek symbol
+
+%shortcuts for kappa
+\newcommand{\kenc}[1][r]{$\kappa_\text{encl}(#1)$\xspace}
+\newcommand{\kap}[1][r]{$\kappa(#1)$\xspace}
+
+% shorcuts for refs (use capital for beginning of sentence)
+% first 3 are for the real layz people..
+\newcommand{\fref}[1]{\ref{fig:#1}}
+\newcommand{\sref}[1]{\ref{sec:#1}}
+\newcommand{\tref}[1]{\ref{tab:#1}}
+\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
+\newcommand{\secref}[1]{Section~\ref{sec:#1}}
+\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
+\newcommand{\Figref}[1]{Figure~\ref{fig:#1}}
+\newcommand{\Secref}[1]{Section~\ref{sec:#1}}
+\newcommand{\Tabref}[1]{Table~\ref{tab:#1}}
+
+% shortcut for ASW000xxxx
+\newcommand{\asw}[1]{ASW000#1\xspace}
+
+\newcommand{\model}[2][Model~]{#1#2\xspace}
+
+%degrees
+\newcommand{\dgr}{^{\circ}}
+
+% todo annotations on side in red. easy to search for at the end
+\newcommand{\todo}[2][red]{%\marginpar{\raggedright\scriptsize\sffamily\textcolor{red}{#1}}}
+
+\textcolor{#1}{\textbullet}%
+\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
+\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}
+
+\newcommand{\needfig}[1][]{%
+\colorbox{yellow}{\parbox{0.5\textwidth}{%
+\vspace{0.5cm}\setstretch{0.5}\textcolor{black}{\scriptsize{fig #1}}\vspace{0.5cm}}}%
+\marginpar{\colorbox{yellow}{\parbox{\marginparwidth}{%
+\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{fig}}}}}}
+
+\setlength{\marginparsep}{0mm}
+\setlength{\marginparwidth}{2.2cm}
+
+\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
+\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
+\newcommand{\needchk}[1][]{\todo[red]{chk #1}}
+
+\newcommand{\hr}{\vspace{5mm}\noindent\rule{0.8\textwidth}{0.4pt}\vspace{5mm}}
 
@@
 
-\input{tex/abstract}
-\input{tex/intro}
-\input{tex/theory}
-\input{tex/spaghetti}
-\input{tex/results}
-\input{tex/outlook}
+\begin{abstract}
+In Spacewarps, volunteers are invited to search CFHTLS sky survey for
+lens candidates.  Here we report on a system that allows experienced
+volunteers to model lens candidates.  A sample of 29 simulated lenses
+were modeled, each multiple times.  The quality of the models was then
+examined in different ways.  In particular: (i)~the image parities and
+time arrivals were identified correctly in most cases, but not
+always; (ii)~the mean convergence (equivalent to the enclosed mass)
+was quite well constrained in the image region; (iii)~the performance
+of experienced volunteers was comparable to professionals.
+\end{abstract}
 
-\input{tex/appendix}
+\keywords{}
+\section{Introduction}
+
+Our present picture of the formation of the largest structures in the
+Universe --- galaxies and clusters of galaxies --- is that
+gravitational instabilities (whose early stages are observable as
+fluctuations of the microwave background) caused a collapse into
+gravitationally bound structures.  Dark matter would have dominated
+the initial collapse, but once deep wells in the gravitational
+potential appeared, gas would have fallen into them while radiating
+away its gravitational potential energy through atomic processes to
+form the first stars and galaxies.  With time, galaxies and clusters
+would continue to merge and change their form.  This rough outline is
+well-established now, but many puzzles remain to be solved before our
+understanding is satisfactory.  To progress towards solving these, we
+need more knowledge of dark matter.  And to study dark matter,
+gravitational lensing is attractive, because it depends only on mass.
+
+There are many manifestations of gravitational lensing in astrophysics,
+but the most spectacular are multiple images.  Gravitational lensing
+depends on the sky-projection mass density, and this fact has a
+counter-intuitive consequence: nearby galaxies do not produce multiple
+images, because their projected density is not high enough, but beyond
+100~Mpc or so, most galaxies are potential lenses.  Even so, for a
+potential lens to really act as a lens, a background source in the close-by
+region is required.  So, in practice galaxies are very rarely observed
+as lens.  The same tendency applies to clusters of galaxies, but since
+clusters are bigger, they are more likely lens-like.
+
+In lensing galaxies, the observable features usually involve a single
+object multiply imaged.  It is therefore desirable to have large
+numbers of lenses, and this is the strategy researchers are following.  The
+largest single study so far uses 58 lenses \citep{2009ApJ...703L..51K}.
+Other studies have compared lens models in more detail with stellar
+mass \citep{2011ApJ...740...97L} or used time-delay information from
+lensed quasars to infer the Hubble time
+\citep{2008ApJ...679...17C,2010ApJ...712.1378P}.
+
+The trend is clear: researchers want to find many more lenses and
+model them in more sophisticated ways.
+
+\begin{figure}
+\centering
+\includegraphics[width=.8\hsize]{fig/lenssky.png}
+\caption{Sky distribution of 398 published lenses considered secure as
+  of September 2013 (from the Masterlens catalog at the University of
+  Utah, maintained by Joel Brownstein and Leonidas Moustakas.) The map
+  is in Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
+  middle.  The empty swathes to the left and right of center are the
+  Milky Way.}
+\label{fig:masterlens}
+\end{figure}
+
+A good summary of the observational situation is offered by
+Figure~\ref{fig:masterlens}, which shows known secure multiply-imaging
+lenses.  The non-uniformity on the sky is not intrinsic, it just
+indicates the density of deep surveys up to 2013.  At HST resolution,
+of order 1~square degree of sky must be searched to find a lens.
+Older ground-based surveys yield a lens per roughly 10~square degree
+of survey area.  By the large surveys starting now
+(DES\footnote{\tt http://pan-starrs.ifa.hawaii.edu} or
+PanStarrs\footnote{\tt http://www.darkenergysurvey.org}) the expected
+yield is something in between.  Given their unprecedented survey area,
+these are likely to yield thousands of lenses.  Over the 2020s,
+LSST\footnote{\tt http://www.lsst.org/} can be expected to give us ten
+thousand lenses.
+
+The question now arises: can lenses be found and modeled
+automatically in large surveys?  Work using software robots
+\citep{2009ApJ...694..924M} gives some interesting and unexpected
+results.  In clean lensing system in uncrowded fields with high
+signal-to-noise, the robots do very well.  In most situations,
+however, robots miss lenses (low completeness) or contaminate the
+results with non-lenses (low purity).  Robots can be made to
+prioritize completeness or purity, but they cannot deliver both.
+
+In response to the lessons learned from automated lens searches, the
+{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org} 
+(Saha is a co-investigator) has been launched.  {\em Spacewarps\/} is
+part of the {\em Zooniverse\/} family of citizen-science
+projects,\footnote{\tt http://www.zooniverse.org} where members of the
+public are invited to analyze different kinds of scientific data which
+are too difficult for robots and too large for specialists.  In {\em
+ Spacewarps\/} itself, since the launch in May 2013, $\sim100$ square
+degrees (or a quarter of a percent of the sky) has been examined.  The
+survey is divided into patches of $440\times 440$ pixels, each of
+which is seen by ten volunteers.  Simulated lenses are mixed in with
+the data, both to help train volunteers on what to look for, and to
+estimate completeness and purity.  A list of candidates is being
+processed, but intermediate results show some very good candidates, as
+well re-discoveries of a few known lenses.
+
+These encouraging results now raise the question: could modeling of
+the lenses also be done by volunteers?  Modeling is a much more
+difficult task than searching for lens candidates, as it requires some
+expert knowledge.  Nonetheless, a panel of the {\em Spacewarps\/}
+volunteers is quite experienced from earlier projects, having spent a
+thousand hours or more with data.  Some of these experienced
+volunteers are very interested in more demanding projects.  This also
+is a general trend \citep[cf.][]{Khatib22112011}.  The present project
+then suggests itself.
+
+The purpose of this study was to provide a means to model a large number of gravitational lenses by showing that gravitational lens modeling can be learned / done by volunteers.
+We suggest, that volunteers will be as successful as professionals with modeling if provided with an easy to use tool with visual feedback (What you see is what you get, WYSIWYG) and a minimal set of instructions.
+Volunteers will then crowd work (using buzz word here ;) ) / work collaborative on modeling lenses from several sources / groups at a central place.
+Since this is an iterative learning process, the more involved volunteers will quickly gain knowledge that can be passed down to new volunteers.
+That creates a social structure that scales well with the number of volunteers, as other projects\needcite have already shown.
+Finding people working as volunteers has been shown to be successful last but not least by \sw and the whole galaxy zoo project.
+To test the people's abilities, we investigated the performance of a few volunteers modeling a set of simulated lenses.
+We tested the ability to correctly identify lensed images and reproduce similar mass map of the lens.
+
+\section{A lens modeling program}
+
+It is well known that making a scientific contribution is a central
+motivation cited by citizen-science volunteers.  So for a lens
+modeler in a citizen-science project, nobody wants that it should
+sacrifice scientific usefulness in order to be fun to use.  That said,
+being aesthetically pleasing and having a short initial learning curve
+are also essential qualities.  Also desirable is that the user is
+encouraged/challenged to go deeper and wider into the subject.
+Yet another aspect is enabling small incremental contributions
+by different people.  \spl tries to address all of these.
+
+\subsection{Theory: Fermat's Principle} \label{sec:Fermat}
+
+There are several ways to understand the formation of arcs and
+multiple images in gravitational lensing.  We will follow some ideas
+originally introduced by \cite{1986ApJ...310..568B}, based on Fermat's
+principle.  The key to this approach is an abstract construct called
+the {\em arrival-time surface.}  This surface cannot itself be
+observed, but several observable quantities can be derived from it.
+
+Consider a gravitational lens.  As in most astrophysical lensing, this
+lens is `thin' along the line of sight, and effectively lies on a
+plane.  Let $x,y$ be coordinates on this plane.  These coordinates
+measure ordinary physical lengths (meters, etc) on the lens plane.
+Let there be a point source of light at $x=0,y=0$ but far behind the
+lens plane.  Now imagine a light ray coming from the source to the
+lens plane $(x,y)$ and then changing its path to reach the observer.
+The observer would see the source apparently behind $(x,y)$ the
+lens.  If this happened, the light ray would have taken a longer path
+than having come directly to the observer.  The geometrical path
+difference would be $\propto x^2 + y^2$ for small angles.  Let us
+define
+\begin{equation} \label{eq:Ageom}
+A_{\rm geom}(x,y) = \half(x^2 + y^2) \,.
+\end{equation}
+As written, this looks an area.  But we can also think of it as a
+time delay introduced by the light ray bending --- with an unwritten
+constant factor.  That factor is basically the lens distance divided by the
+speed of light; the precise expression has to take the expansion of
+the Universe into account, and is given in the Appendix.  The top part
+of \figref{arriv} shows what $A_{\rm geom}$ looks like: it is
+just a paraboloid.  The minimum of the surface is at the source, and
+that is where, according to Fermat's principle, the image will be.
+The rest of the arrival-time surface is just an imaginary surface.
+
+\begin{figure}
+\centering
+\includegraphics[height=.3\vsize]{fig/arriv1.png}
+\includegraphics[height=.3\vsize]{fig/arriv2.png}
+\includegraphics[height=.3\vsize]{fig/arriv3.png}
+\caption{Arrival-time surfaces, with contour levels of equal arrival
+  time.  The upper surface is with no lens.  The middle surface is
+  when a circular lensing mass (offset from the source) is added; a
+  maximum, a minimum and a saddle point can be seen.  The last surface
+  is the result of an elongated lensing mass; a maximum, two minima
+  and two saddle points can be seen.
+}
+\label{fig:arriv}
+\end{figure}
+
+The geometrical time delay \eqref{eq:Ageom} is not the only one.  The
+warping of space by a gravitational field introduces a further time
+delay $A_{\rm grav}$.  The arrival time of the light ray, compared to
+what it would have been with no lensing, is
+\begin{equation} \label{eq:Aarriv}
+\hbox{arrival time} = A_{\rm geom} + A_{\rm grav} \,.
+\end{equation}
+We can also express $A_{\rm grav}$ as an area, with the same implicit
+constant factor that makes it a time.  The expression, however, is
+more complicated.  We first introduce the notation $\nabla^2 f(x,y)$
+to denote
+\begin{equation}
+ \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
+        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
+      {\Delta x \; \Delta y}
+\end{equation}
+in the limit of $\Delta x,\Delta y$ small.  Then we write the
+projected mass density of the lens, which would be in $\rm
+kg\,m^{-2}$, in special units (given in the Appendix) such that the
+surface density becomes a dimensionless field $\kappa(x,y)$.  Then
+\begin{equation} \label{eq:Poisson}
+\nabla^2 A_{\rm grav}(x,y) = -2\kappa(x,y) \,.
+\end{equation}
+Note that $A_{\rm grav}(x,y)$ is not given explicitly in terms of
+$\kappa(x,y)$, but as a differential equation that must be solved.
+Equations of the type \eqref{eq:Poisson} are, however, well known (as
+Poisson's equation in two dimensions), and there are techniques for
+solving them.  Incidentally, $\kappa$ has a second meaning, as well as
+being a dimensionless form of the projected density: it also
+measures how a bundle of light rays is brought together by
+gravitational lensing, and is called the {\em convergence}.
+
+What is the effect of mass on the arrival-time surface?  If the mass
+distribution $\kappa(x,y)$ is circular, $A_{\rm grav}(x,y)$ will have
+a peak at the center of that circle.  But if the mass is not precisely
+in front of the source, the arrival-time surface will not be circular
+any more.  The result is illustrated in the middle part of
+\figref{arriv}.  In addition to the maximum, there is now a
+minimum and a saddle point.  According to Fermat's principle, images
+appear at maximum and saddle points as well as minimum.  If the mass
+distribution is non-circular, more images can appear.  The lower
+example in \figref{arriv} shows the arrival-time surface due
+to an elongated mass.  There is still a maximum, part of an elongated
+hill, and around it there are two minima and two saddle points.
+
+So far we have considered a single point source.  What if there is an
+extended source?  To understand this, let us consider what happens if
+we move the original point source slightly, or equivalently, keep the
+original point source behind $x=0,y=0$ and move the lens slightly.
+The contours of constant arrival time will naturally move slightly,
+and so will the images.  The movement of the contours will be most
+noticeable where the contours are far apart, that is where the
+arrival-time surface is nearly flat.  As is evident from
+\figref{arriv}, this is the region where the minimum and saddle
+points lie, or near the images.  In this region, points on the source
+that are close together produce images that are comparatively far
+apart.  In other words, the image is highly magnified.  In summary,
+lower curvature in the arrival-time surface for a point source implies
+larger magnification of an extended source.  Conversely, where the
+arrival-time surface is strongly curved, the image will be
+demagnified.  We see from \figref{arriv} that the arrival-time
+surface tends to be highly curved near the maximum.  Hence maximum tend
+to be demagnified.  In practice, maxima of the arrival time are nearly
+always too faint to see. The minima and saddle points dominate.
+
+In the lensed examples in \figref{arriv}, the maximum in each
+case is evident, but how can we distinguish the minima from the saddle
+points?  The key is to examine the contours of equal arrival time.
+Contour curves loop around a minimum or a maximum, but near a saddle
+point, the contours do not loop around, instead they approach and then
+pull away.  The contour curve precisely level with a saddle point
+forms an X at the saddle point.  In \figref{arriv}, one such X
+is evident.  In other cases the X contour passing through the saddle
+point is not itself plotted, but we can tell by the other contours where
+the X would be.  These self-crossing contours will play a special role
+in the following.
+
+\subsection{\spl} \label{sec:SpaghettiLens}
+
+\spl implements the ideas of the arrival-time surface into a modeling
+program.
+
+\begin{figure}
+  \centering
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/006941_input.png}} \\
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/007022_input.png}} \\
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/006919_input.png}}
+  \caption{Examples of Spaghetti input.  The models from these appear
+    later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
+  }
+  \label{fig:input-spag}
+\end{figure}
+
+The first modeling step with \spl is for the user to call up the \sw
+of interest, and make an educated guess for the topography of the
+arrival-time surface --- specifically, the locations of the maxima,
+minima and saddle points that correspond to images of the brightest
+part of the source.  Input is given by tracing a proposal for the
+saddle-point contours.  \Figref{input-spag} shows some example
+inputs --- and incidentally also indicates the origin of the name
+\spl.
+
+\spl, as implemented so far, assumes that the lens is dominated by a
+single galaxy, and that the center of that galaxy is the sole maximum.
+Once the maximum has been identified, we wish to characterize the rest
+of the arrival-time surface in relation to it.  From geometry, any
+surface that has a central maximum and goes high at the edges must
+have a lima\c con-like contour, meaning a two-looped curve with one
+loop inside the other.  The maximum lies within the inner loop, a
+minimum lies between the two loops, and a saddle point lies at the
+self-intersection of the contour.  The top panel of
+\figref{input-spag} shows an example: red marks a suggested
+maximum, green a saddle point, and blue a minimum.  (The small pink
+dots are just help sketch the proposed contour.)  The middle and lower
+panels of \figref{input-spag} show a more complex situation.
+The minimum has split into three images: two new minima with a saddle
+point in between.  In this situation the contour through the new
+saddle point forms a figure of eight, its two loops enclosing the two
+new minima.  The configuration of a lima\c con, possibly with a figure
+of eight inside, suffices to give a preliminary description of the
+arrival-time surface for nearly all cases where the lens is dominated
+by one galaxy, and the source is a single small object.
+
+The exact placement of the pink spaghetti loops shown in
+\Figref{input-spag} has no significance.  Their utility is
+just to help mark up the image as maximum, minima and saddle points.
+As the figure suggests, the marking up can be done very easily with a
+mouse.  Human intuition is required to (a)~identify lensed images and
+separate them from other background light sources, and (b)~classify
+and order images according to arrival time.  But the computer helps by
+allowing only valid lensing configurations to be entered, and ensuring
+that the odd image theorem\needcite is taken care of.
+
+In the second modeling step, once the user has sketched a proposed
+spaghetti configuration, \spl sends the input to its server-side
+modeling engine, called GLASS.  The task of GLASS is to find a mass
+distribution $\kappa(x,y)$ that exactly reproduces the locations of
+the maximum, minima and saddle points. Now, this criterion by itself
+is extremely under-determined --- there are infinitely many mass
+distributions that will reproduce a given set of maxima, minima and
+saddle points, but typically they (a)~produce lots of extra images,
+and (b)~look very unlike galaxies.  Additional assumptions (a prior)
+are necessary.  GLASS uses a prior based on suggestions by
+\cite{1997MNRAS.292..148S}.
+\begin{enumerate}
+\item The mass distribution is built out of non-negative tiles of
+  mass.  (Sometimes these tiles are called mass pixels, but we should
+  emphasize that they are unrelated to image pixels, and are much
+  larger.)
+\item There is a notional lens center, say $(x_0,y_0)$ which is
+  identified with the maximum of the arrival time.  The source can
+  have an arbitrary offset with respect to the lens center.
+\item The mass distribution must be centrally concentrated, in two
+  respects.  First, the circularly averaged density must fall away
+  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
+  Second, the direction of increasing density at any $(x,y)$ can point
+  at most $45^\circ$ away from $(x_0,y_0)$.
+\item The lens must be symmetrical with respect to $180^\circ$ rotations
+  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
+  user wishes.
+\end{enumerate}
+There are still infinitely many models that satisfy both data and
+prior constraints, but now they are more credible as galaxy lenses.
+It is then possible to generate an ensemble of models.  The sampling
+technique used by GLASS is described in \citep{Lubini2012}, and
+improves upon earlier techniques \citep{2000AJ....119..439W,Saha2004}.
+Typically, ensembles of 200 models are used.  That is to say, what we
+call a \spl model is really the mean of an ensemble of 200 models, and
+its estimated uncertainty is the range covered by the whole ensemble.
+
+In the third modeling step, having generated a model ensemble, \spl
+post-processes it to present results and diagnostics to the user for
+inspection. This takes the form of three figures.
+\begin{enumerate}
+\item A gray scale plus contour map of the mass distribution.
+\item A contour map of the arrival-time surface.
+\item A synthetic image of the lensed features.
+\end{enumerate}
+After examining this feedback, the user can archive the results for
+discussion, or modify their input and try again, or discard the
+attempt altogether.
+
+The fourth modeling step is discussion among modelers and iteration
+on the model.  Any archived model can be revised by another user: they
+can modify the spaghetti configuration slightly or drastically, or
+change options like the size of the mass tiles.  Particularly
+interesting lens candidates lead to trees of models in this way.
+Forum discussion then prunes the tree, focusing attention on one or a
+few models.
+
+\section{A lens modeling challenge} \label{sec:mod_challenge}
+
+% \subsection{\sw} \label{sec:SpaceWarps} some words about \sw
+
+Interested volunteers from the \sw forum were initially introduced to
+\spl through a video tutorial and by videocon.  After this
+introductory stage, a modeling challenge was presented.  This
+consisted of 29 simulated lenses (sims) covering a range of lensing
+configurations.
+
+The \sw sims were generated by AM, in consultation with PM and AV.
+To estimate the performance of the volunteers and the quality of the generated models, two analysis  were done.
+The first analysis tested the correct identification and ordering of lensed images.
+The second one compared the mass distribution of the lens $\kappa(x, y)$ of the generated models to the mass distribution of the simulations.
+
+\subsection{The simulated lenses} \label{sec:sims}
+
+For the sake of blind testing, the information in this section was
+revealed neither to RK, while choosing the challenge set of 29 sims,
+nor to the modelers (volunteers EB, CC, CM, JO, JW and `expert
+modeler' PS) until the modeling stage was done.
+
+The sims were produced using {\tt gravlens}
+\citep{2001astro.ph..2341K,2001astro.ph..2340K}. They were of three
+kinds, as follows.
+
+\begin{enumerate}
+  \item Imitating lensed quasars: having a singular elliptical
+    isothermal lens (SIE) plus constant external shear, and a circular
+    Gaussian source.
+  \item Emulating lensed galaxies: similar to the above, but with an
+    elliptical de Vaucouleurs source.
+  \item Resembling cluster lenses: having a source as above, but a
+    more complicated lens, with one dominant elliptical SIE and
+    one or more perturbing elliptical SIEs, plus a circular NFW
+    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
+    the intergalactic mass.
+\end{enumerate}
+
+Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
+lenses follow equations (33--35) of that work, with core radius set to
+zero.  The NFW lens is in equations (48) and (50), while shear is the
+$\gamma$ term in equation (76).
+
+\subsection{Some example models} \label{sec:example_models}
+
+The modelers proposed a total of 129 models for the 29 sims in the
+challenge.  Figures \ref{fig:6941} to \ref{fig:6919} show eight of
+models in some detail.
+
+\begin{figure}
+  \centering
+  \subfigure[real mass distribution]{
+    \label{fig:6941_sim_mass}
+    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa.png}
+  }
+  \subfigure[real arrival-time surface]{
+    \label{fig:6941_sim_arr}
+    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv.png}
+  }
+  \subfigure[model mass distribution]{
+    \label{fig:6941_mass}
+    \includegraphics[width=0.45\textwidth]{fig/006941_mass.png}
+  }
+  \subfigure[model arrival-time surface]{
+    \label{fig:6941_cont}
+    \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti.png}
+  }
+  \subfigure[real vs model enclosed mass]{
+    \label{fig:6941_kappa}
+    \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl.png}
+  }
+  \subfigure[model lensed image]{
+    \label{fig:6941_atime}
+    \includegraphics[width=0.45\textwidth]{fig/006941_arr_time.png}
+  }
+  \caption[result 6941 (ASW000102p)]{Sim ASW000102p and a model.  A
+    simple configuration; the model input is shown in the top panel of
+    Figure~\ref{fig:input-spag}.  For details of the individual panels
+    here, see section \ref{sec:example_models}.}
+  \label{fig:6941}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfigure[real mass distribution]{
+    \label{fig:6975_sim_mass}
+    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_kappa.png}
+  }
+  \subfigure[real arrival-time surface]{
+    \label{fig:6975_sim_arr}
+    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_arriv.png}
+  }
+  \subfigure[model mass distribution]{
+    \label{fig:6975_mass}
+    \includegraphics[width=0.45\textwidth]{fig/006975_mass.png}
+  }
+  \subfigure[model arrival-time surface]{
+    \label{fig:6975_cont}
+    \includegraphics[width=0.45\textwidth]{fig/006975_spaghetti.png}
+  }
+  \subfigure[real vs model enclosed mass]{
+    \label{fig:6975_kappa}
+    \includegraphics[width=0.45\textwidth]{fig/006975_kappa_encl.png}
+  }
+  \subfigure[model lensed image]{
+    \label{fig:6975_atime}
+    \includegraphics[width=0.45\textwidth]{fig/006975_arr_time.png}
+  }
+  \caption[result 6975 (ASW000195x)]{Sim ASW000195x and a
+    model.}
+  \label{fig:6975}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfigure[real mass distribution]{
+    \label{fig:6937_sim_mass}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_kappa.png}
+  }
+  \subfigure[real arrival-time surface]{
+    \label{fig:6937_sim_arr}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_arriv.png}
+  }
+  \subfigure[model mass distribution]{
+    \label{fig:6937_mass}
+    \includegraphics[width=0.45\textwidth]{fig/006937_mass.png}
+  }
+  \subfigure[model arrival-time surface]{
+    \label{fig:6937_cont}
+    \includegraphics[width=0.45\textwidth]{fig/006937_spaghetti.png}
+  }
+  \subfigure[real vs model enclosed mass]{
+    \label{fig:6937_kappa}
+    \includegraphics[width=0.45\textwidth]{fig/006937_kappa_encl.png}
+  }
+  \subfigure[model lensed image]{
+    \label{fig:6937_atime}
+    \includegraphics[width=0.45\textwidth]{fig/006937_arr_time.png}
+  }
+  \caption[result 6937 (ASW0000vqg)]{Sim ASW0000vqg and a model.}
+  \label{fig:6937}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfigure[real mass distribution]{
+    \label{fig:6990_sim_mass}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_kappa.png}
+  }
+  \subfigure[real arrival-time surface]{
+    \label{fig:6990_sim_arr}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_arriv.png}
+  }
+  \subfigure[model mass distribution]{
+    \label{fig:6990_mass}
+    \includegraphics[width=0.45\textwidth]{fig/006990_mass.png}
+  }
+  \subfigure[model arrival-time surface]{
+    \label{fig:6990_cont}
+    \includegraphics[width=0.45\textwidth]{fig/006990_spaghetti.png}
+  }
+  \subfigure[real vs model enclosed mass]{
+    \label{fig:6990_kappa}
+    \includegraphics[width=0.45\textwidth]{fig/006990_kappa_encl.png}
+  }
+  \subfigure[model lensed image]{
+    \label{fig:6990_atime}
+    \includegraphics[width=0.45\textwidth]{fig/006990_arr_time.png}
+  }
+
+  \caption[result 6990 (ASW0004oux)]{Sim ASW0004oux and a model.}
+  \label{fig:6990}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfigure[real mass distribution]{
+    \label{fig:6915_sim_mass}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_kappa.png}
+  }
+  \subfigure[real arrival-time surface]{
+    \label{fig:6915_sim_arr}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_arriv.png}
+  }
+  \subfigure[model mass distribution]{
+    \label{fig:6915_mass}
+    \includegraphics[width=0.45\textwidth]{fig/006915_mass.png}
+  }
+  \subfigure[model arrival-time surface]{
+    \label{fig:6915_cont}
+    \includegraphics[width=0.45\textwidth]{fig/006915_spaghetti.png}
+  }
+  \subfigure[real vs model enclosed mass]{
+    \label{fig:6915_kappa}
+    \includegraphics[width=0.45\textwidth]{fig/006915_kappa_encl.png}
+  }
+  \subfigure[model lensed image]{
+    \label{fig:6915_atime}
+    \includegraphics[width=0.45\textwidth]{fig/006915_arr_time.png}
+  }
+
+  \caption[result 6915 (ASW0001hpf)]{Sim ASW0001hpf and a model. Inclined quad.}
+  \label{fig:6915}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfigure[real mass distribution]{
+    \label{fig:7022_sim_mass}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa.png}
+  }
+  \subfigure[real arrival-time surface]{
+    \label{fig:7022_sim_arr}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv.png}
+  }
+  \subfigure[model mass distribution]{
+    \label{fig:7022_mass}
+    \includegraphics[width=0.45\textwidth]{fig/007022_mass.png}
+  }
+  \subfigure[model arrival-time surface]{
+    \label{fig:7022_cont}
+    \includegraphics[width=0.45\textwidth]{fig/007022_spaghetti.png}
+  }
+  \subfigure[real vs model enclosed mass]{
+    \label{fig:7022_kappa}
+    \includegraphics[width=0.45\textwidth]{fig/007022_kappa_encl.png}
+  }
+  \subfigure[model lensed image]{
+    \label{fig:7022_atime}
+    \includegraphics[width=0.45\textwidth]{fig/007022_arr_time.png}
+  }
+
+  \caption[result 7022 (ASW0000h2m)]{Sim ASW0000h2m and a model.  The
+    model input appears in the middle panel of Figure
+    \ref{fig:input-spag}.}
+  \label{fig:7022}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfigure[real mass distribution]{
+    \label{fig:7025_sim_mass}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa.png}
+  }
+  \subfigure[real arrival-time surface]{
+    \label{fig:7025_sim_arr}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv.png}
+  }
+  \subfigure[model mass distribution]{
+    \label{fig:7025_mass}
+    \includegraphics[width=0.45\textwidth]{fig/007025_mass.png}
+  }
+  \subfigure[model arrival-time surface]{
+    \label{fig:7025_cont}
+    \includegraphics[width=0.45\textwidth]{fig/007025_spaghetti.png}
+  }
+  \subfigure[real vs model enclosed mass]{
+    \label{fig:7025_kappa}
+    \includegraphics[width=0.45\textwidth]{fig/007025_kappa_encl.png}
+  }
+  \subfigure[model lensed image]{
+    \label{fig:7025_atime}
+    \includegraphics[width=0.45\textwidth]{fig/007025_arr_time.png}
+  }
+
+  \caption[result 7025 (ASW0000h2m)]{Sim ASW0000h2m and another model.  The image parities in this case were incorrectly identified.}
+  \label{fig:7025}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfigure[real mass distribution]{
+    \label{fig:6919_sim_mass}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_kappa.png}
+  }
+  \subfigure[real arrival-time surface]{
+    \label{fig:6919_sim_arr}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_arriv.png}
+  }
+  \subfigure[model mass distribution]{
+    \label{fig:6919_mass}
+    \includegraphics[width=0.45\textwidth]{fig/006919_mass.png}
+  }
+  \subfigure[model arrival-time surface]{
+    \label{fig:6919_cont}
+    \includegraphics[width=0.45\textwidth]{fig/006919_spaghetti.png}
+  }
+  \subfigure[real vs model enclosed mass]{
+    \label{fig:6919_kappa}
+    \includegraphics[width=0.45\textwidth]{fig/006919_kappa_encl.png}
+  }
+  \subfigure[model lensed image]{
+    \label{fig:6919_atime}
+    \includegraphics[width=0.45\textwidth]{fig/006919_arr_time.png}
+  }
+  \caption[result 6919 (ASW0002z6f)]{Sim ASW0002z6f and a model.  The
+    model input appears in the bottom panel of Figure
+    \ref{fig:input-spag}. An incipient quad}
+  \label{fig:6919}
+\end{figure}
+
+Figure \ref{fig:6941} shows a simple example.  The top two panels show
+contours of convergence $\kappa(x,y)$, and of the arrival-time for a
+point source at the brightest point of the sim
+source.\footnote{Although the arrival-time contours represent an
+  abstract quantity that is not directly observable, the contour map
+  may actually resemble the appearance of the lensed arcs.  The
+  resemblance is due to a serendipitous computer-graphical effect
+  \citep{2001AJ....122..585S}.} The spatial scale is in pixels.  The
+information in these two panels was, of course, kept secret during the
+modeling challenge.  The panels in the middle row and at the lower
+right are the results that \spl returns to the modeler, in order for
+them to assess the model.  These derive from the mean of an ensemble
+of 200 models generated by \spl.  The model mass distribution (left mid) is
+returned as a contour map superimposed on an intensity map. A fairly
+smooth mass distribution, as here, is a good sign.  An irregular or
+checkboard pattern in the mass map usually indicates a bad model.  The
+models arrival-time contours are shown in the right middle panel.
+  The special contour passing through a
+saddle point is shown in black: this is the model version of the
+spaghetti sketch provided by the user.  Also provided is a synthetic
+image of the lensed source in the bottom right panel.
+  For this, \spl assumes a simple conical
+source profile.  The user can change the contrast level on the image,
+which (though it is not saved) amounts to adjusting the width of
+the cone.  Finally, to the lower left of the figure, we have a
+comparison of the input and recovered mass profiles.  The panel shows
+a average $\kappa$ within a given radius, as a function of radius.
+The red curve is the true value, and where it crosses unity (dotted
+horizontal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
+  The two blue curves
+are the minimal and maximal mean enclosed $\kappa$ from the internal
+ensemble in \spl.  The region between the blue curves is shaded
+between the radii of the innermost and the outermost images --- this
+is the confidence region from the modeling.  As we see, the shaded
+blue is slightly above the red curve. 
+The Einstein Radius $\Theta_\text{E}$ of the model is estimated crossing the
+mean enclosed $\kappa$ (not plotted) with unity.
+  In summary: the identification
+of minimum and saddle point is correct, but the estimated Einstein
+radius is a little too high.
+
+Figure \ref{fig:6975} shows a lens with substructure in the form of a
+smaller secondary galaxy.  The galaxies in such group or cluster sims
+were, in fact, based on galaxies visible in the images --- but the
+modelers were not told in advance whether this was the case.  The
+model does not include any substructure, but otherwise is not bad.
+The minimum and saddle point are correct, and the Einstein radius is
+only a little underestimated.
+
+Figure \ref{fig:6937} shows a case where substructure leads to a
+poor model.
+
+Figure \ref{fig:6990} shows an example of an arc that has split into
+three images.  This kind of configuration, with a counter-image close
+to the lensing galaxy and a more distant arc/triplet on the other
+side, generically arises from an elongated mass distribution when the
+source is displaced along the elongated direction.
+
+Figure \ref{fig:6915} shows another quad.  This kind of configuration
+arises when the mass is elongated and the source is displaced at an
+angle to the elongation.
+
+Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
+saddle points are correctly identified, and the orientation of the
+ellipticity of the mass distribution is correctly reproduced.  The
+Einstein radius is somewhat overestimated.  Figure \ref{fig:7022}
+shows another model of the same system.  In this one, the
+identification of the minima and saddle points was incorrect, and mass
+distribution comes out elongated East-West instead of North-South.
+The mass distribution also appears somewhat jagged and the
+saddle-point contours are not as clean as in the previous examples;
+these are often indicators of a problem with the model.  The enclosed
+mass is, however, none the worse --- the reason is probably that in a
+relatively symmetrical image configuration, the Einstein radius is quite
+well constrained by the images in a fairly model-independent way.
+
+To conclude this set of examples, Figure \ref{fig:6919} shows another
+type of quad.  Actually, in this case the brightest part of the source
+is only doubly imaged, but the source extends into a region that
+produces four images.  We can also see from the real arrival-time
+surface that a point source is a double on the verge of splitting into
+a quad.  The modeler interpreted the system as a quad.  The
+appearance of the arcs, shown in the bottom panel of Figure
+\ref{fig:input-spag}, looks like an arc and counter-image such as
+discussed with Figure \ref{fig:6990} above.  But there is an important
+difference: the long arc is closer to the galaxy, as if the arc and
+counter-image have swapped roles.  This configuration arises if the
+source displacement is perpendicular to the long axis of the lensing
+mass.
+
+\subsection{Test of image identification} \label{sec:tests.t1}
+
+A first manual evaluation tested the volunteers ability to reconstruct the arrival time surface given a survey image containing a sim.
+This task consists of two parts.
+First to correctly identify and locate the lensed images.
+Second the correct ordering for the identified lensed images in respect of the arrival time.
+
+While we expected the identification of lensed images to be trivial, given the nature of the survey images and the success of \sw, we expect the correct ordering to be more difficult.
+This part tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
+While we can provide the volunteers with some general rules of thumb, ordering involves imagination and guessing and therefore training could improve results in a later stage.
+
+This first test was designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
+
+The evaluation of the volunteers performance was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot generated using simulation parameters.
+% \Figref{output_compare} shows the setup used to evaluate the models.
+
+The images of the system are considered to be identified correctly, if all the images have been identified and are approximated within $\pm0.05\cdot\text{imgage width}$.
+The parity is considered correct, if those identified points have the right ordering with respect to arrival time.
+
+Additionally, ten types of errors (labeled E01 -- E10, listed in \tabref{stats}) that occurred were identified.
+Each generated model could contain more than one error.
+
+\begin{table}\centering\begin{tabular}{llcc}\hline
+ & & n & p \\
+\hline
+ N: & Total number of models & 119 & 1.00\\
+\hline
+ R1: & images approx. on right location & 110 & 0.92\\
+ R2: & images with correct parity & 70 & 0.59\\
+\hline
+ E01: & inaccurate in arc & 21 & 0.18\\
+ E02: & wrong parity in 3 lens conf. & 2 & 0.02\\
+ E03: & identified 3 of 5 imgs. & 5 & 0.04\\
+ E04: & modeled arc with single img. & 4 & 0.03\\
+ E05: & $\pi$ rotated parity & 7 & 0.06\\
+ E06: & $\pi$/2 rotated parity & 38 & 0.32\\
+ E07: & missed faint img. & 1 & 0.01\\
+ E08: & too many imgs in arc. & 5 & 0.04\\
+ E09: & missed double img & 3 & 0.03\\
+ E10: & too many imgs. & 1 & 0.01\\
+\end{tabular}
+
+\caption{1. col: id of model; 2.col: places of imgs idendited
+  correctly; 3. col: more or less correct identification of extr
+  points. 4.col: exact identification of extremal points. 5.col:
+  type(s) of errors ocured....  1) inaccurate placement in an extended
+  arc; 2) wrongly identified sad and min in 3 image configuration; 3)
+  identified only 3 instead of 5 images; 4) tried to model an arc with
+  a min instead of min-sad-min; 5) PI-err (rotation by 180 degrees; in
+  5 image configuration, exchanged the ordering of the two saddle
+  points); 6) PI/2-err (rotation by 90 degree; sad-min-sad-min-sad);
+  7) missed faint image(s); 8) tried to model an arc with min-sad-min
+  instead of only min; 9) did identify two close by images as one; 10)
+  used 7 or more image to model a 5 image system. }
+
+\label{tab:stats}
+\end{table}
+
+\tabref{stats} presents a summary of this evaluation.
+We conclude that the volunteers are performing very well identifying and positioning images, with a performance of 92\% (R1, p=0.92).
+Most of the problems where due to unclear arc-like structures (E01, p=0.18; E04, p=0.03; E08, p=0.04).
+Critical errors like the failure to identify all five images in a five images system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happen.
+From this we conclude that the introduction materials was adequate and the volunteers understand the basics of gravitational lensing.
+
+The assignment of the parity of the images was a more difficult task.
+In 59\% (R2, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
+Most of the failures are due to E06 (N=38, p=32\%), followed by E05 (N=7, p=6\%).
+E06 describes a situation, where the minima and saddle points of a five image configuration were exchanged (rotated by $\pm90\dgr$), see \figref{7022} for an example.
+E05 describes a situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$).
+While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
+For more challenging cases, like very symmetrical distribution of the lensed images (for example model 7022, \figref{7022}), those errors should still produce plausible results, as will be explained in the next section.
+
+\subsection{Test of mass-profile recovery} \label{sec:tests.t2}
+
+The second test was to compare the mass distributions of the lens $\kappa(x, y)$ given by the sims and generated by the volunteers.
+To get a means of comparing the sims to the models, the total convergence\footnote{often called enclosed mass} $\kappa_{\text{encl}}(r)$ was calculated for both.
+The Einstein radius $\Theta_\text{E}$ is defined by $\kappa_{\text{encl}}(\Theta_\text{E})=1$ and gives a number that allows a rough comparison between a sim and a model.
+We also let an expert (PS) model three selected systems to compare the results from volunteers to those of a professional.
+
+\begin{figure}[htbp]
+  \centering
+    \includegraphics[width=0.80\textwidth]{fig/eR_1.png}
+  \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
+  \label{fig:ER_all_models}
+\end{figure}
+
+To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
+From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate \ERg.
+To estimate the errors, the extremal models are used to estimate a lower and upper limit for \ERg.
+These results can be seen in \figref{ER_all_models}.
+This figure shows that this technique of estimating the error using the ensemble of all models underestimates the error significantly and should be improved for further analysis.
+
+\begin{figure}[htbp]
+  \centering
+    \includegraphics[width=0.80\textwidth]{fig/eR_4.png}
+  \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
+  \label{fig:ER_per_sim}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/007020_kappa_encl}}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/007021_kappa_encl}} \\
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/007024_kappa_encl}}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/007025_kappa_encl}}
+  \subfigure{\includegraphics[width=0.45\textwidth]
+             {fig/007022_kappa_encl}}
+  \caption{\kenc for models of \asw{0h2m}: 7020 (top left), 7021 (top right), 7024 (mid left), 7025 (mid right) by volunteers and a correct model 7022 (bottom) by an expert.}
+  \label{fig:kapenc_compare_faulty}
+\end{figure}
+
+\Figref{ER_per_sim} shows that the calculated \ERf of the models tend to be too high.
+The overshoot varies from around 0.2 to 0.4 for good models.
+One of the reasons for this is that it is hard to get the center of the lens on spot.
+An offset leads to a flatter mass profile for the model compared to the simulation.
+
+E05 and E06 happen mostly in very symmetric images, that are hard to come up with a unique valid solution.
+This errors change the orientation of the mass distribution \kap[x,y],
+but \kenc and thus the \ERf is not influenced that much, and thus the final result is still a valid model.
+This can be seen by comparing the results for \asw{0h2m}, shown in \figref{kapenc_compare_faulty}: 
+The correct model 7022 ($\ERm=10.76px$), which was done by an expert, and
+7020 ($\ERm=10.72px$),
+7024 ($\ERm=10.80px$),
+7025 ($\ERm=11.16px$),
+7021 ($\ERm=11.04px$).
+The first three are of type E06, while the last is of type E05.
+This can be easily corrected, if further analysis is done for the modeled system and time delays are known.
+
+Comparing the models from volunteers and experts can be done in \figref{kapenc_compare_faulty}, where only the
+expert got the right configuration, but all the resulting models are comparable, besides rotation.
+
+Two additional sims were modeled by an expert, \asw{1hpf} and \asw{0vqg}.
+Looking at the results for \ERg for those models in \figref{ER_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses, offset) is comparable.
+%Note that the models of \asw{0vqg} with \ERg[, rel] around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
+
+\clearpage
+\section{Outlook} \label{sec:todo}
+
+The lens-modeling challenge indicates that the Spacewarps collective
+is good, not only for identifying lens candidates for follow-up, but
+modeling candidates as well.
+
+There is, however, plenty of room for improvement.
+
+First, the particular modeling strategy implemented is not the only
+one possible.  \spl requires modelers to characterize the overall
+image structure in abstract terms based on Fermat's principle, and the
+placement of mass distributions is done by the computer.  In other
+modeling tools, the user puts down a trial mass distribution and has
+the machine refine it.  A few of these modeling programs have also
+been designed with citizen science in mind, and would be interesting
+in the Spacewarps environment.
+
+Second, \spl needs some enhancements.
+
+\begin{itemize}
+\item Currently, \spl does not attempt to model the source shape; the
+  user identifies the brightest points on the image, and these are
+  taken as images of a point-like source, whose positions must be
+  reproduced exactly. For generating a synthetic image, a conical
+  source profile is assumed. Fitting for the source profile to
+  optimize resemblance to the observed lensed image after the lens
+  model has been generated, is algorithmically straightforward and
+  planned to be implemented.  This would alleviate another problem
+  with \spl, which is that there is no numerical figure of merit, and
+  assessment of a model is a judgment call based on the synthetic
+  image, and on whether the mass distribution and the arrival-time
+  surface show suspect features.
+\item \spl has a tendency to somewhat overestimate the Einstein radius
+  (evident from Figure \ref{fig:ER_per_sim}), and it is also apparent
+  that the models tend to be too shallow.  This possible explanation
+  is that, while the sims are steeply peaked at the centre, the
+  pixellated mass model fixes a comparatively large area near the
+  central at constant density.  One way to solve this problem would be
+  to introduce smaller pixels in the central region, thus enabled a
+  steeper centre.
+\item Another limitation so far in \spl is that the lens is assumed to
+  be dominated by one galaxy, which puts most galaxy-group lenses
+  beyond the reach of the modeler. Since complicated group lenses are
+  some of the most interesting candidates present, removing this
+  limitation is most desirable.  From the users' point of view, it
+  would mean that spaghetti contours with more than one maximum can be
+  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
+  and Figure 4b in \cite{2003ApJ...590...39K}.
+\item At present, a single false-color composite is used as the data.
+  An option could be added to use all available filters, individually
+  or in combination, at the user wishes.
+\end{itemize}
+
+The third desirable avenue of improvement is to facilitate
+collaborative work.
+
+\begin{itemize}
+\item As mentioned above, the option of revising an already-archived
+  model is already available.  Desired now are tools for comparing
+  different models of a given system, both visually and through
+  different statistical measures.  As evidenced by a current
+  collaborative modeling effort, a particularly interesting candidate
+  can lead to an extended discussion and dozens of models.
+\item Better tutorial materials are also needed, and this would
+  address some of the problem areas found in the modeling challenge.
+  For example, we saw in \secref{tests.t1} that volunteers are most
+  prone to making errors in two situations: when in identifying an arc-
+  like structure while placing the points, and in identifying the
+  correct ordering of the points in nearly-symmetric configurations.
+\end{itemize}
+
+\clearpage
+
+\appendix
+
+\section{More on Lensing Theory} \label{sec:more-theory}
+
+In \secref{Fermat}, for the sake of a more intuitive
+explanation, we suppressed some constant factors in
+equations \eqref{eq:Ageom}, \eqref{eq:Aarriv} and \eqref{eq:Poisson}.
+Here we fill in the details.
+
+First let us recall the formulas for comoving distances.
+\begin{equation}
+D = \int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
+\end{equation}
+The comoving distances from observer to lens, from lens to source, and
+from observer to source are given as follows:
+\begin{equation}
+\begin{aligned}
+&D_S    &                                &\int_0^{z_S} \\
+\noalign{\medskip}
+&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
+\noalign{\medskip}
+&D_L    &                                &\int_0^{z_L}
+\end{aligned}
+\end{equation}
+Angular-diameter distances $d$ can be related to comoving distances
+$D$ by:
+\begin{equation}
+\begin{aligned}
+D_S &= (1+z_S) \, d_S \\
+D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
+D_L &= (1+z_L) \, d_L
+\end{aligned}
+\end{equation}
+Angular-diameter distances are always smaller than comoving distances.
+
+Locations on the lens plane in angular coordinates on the sky, are
+expressed as
+\begin{equation}
+(x,y) = d_L (\theta_x,\theta_y) \,.
+\end{equation}
+
+The $A$ areas (with the implicit
+constant factor that makes them time delays) and the $\kappa$ density are related to physical arrival
+time $t$ and density $\Sigma$ as
+\begin{equation}
+\begin{aligned}
+A           &= \frac{c}{(1+z_L)^2} \frac{D_L D_{LS}}{D_S} \times t \\
+\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L D_{LS}}{D_S}
+               \times \Sigma(x,y)
+\end{aligned}
+\end{equation}
+Letting the source be offset at $(s_x,s_y)$ rather than at the
+origin, we have
+\begin{equation}
+\begin{aligned}
+t_{\rm geom} &= \frac{(1+z_L)^2}{2c} \frac{D_S}{D_L D_{LS}}
+\left( (x-s_x)^2 + (y-s_y)^2 \right) \\
+\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
+\end{aligned}
+\end{equation}
+We can now compare with equations (2.1) to (2.6)
+from \cite{1986ApJ...310..568B}.
+%% comparison to be explicited
 
@@
 
-\newpage
-
-\input{tex/figures}
-
 \end{document}

### 2014-01-12 text/ms.tex
Test

@@
 
+xxx
 Our present picture of the formation of the largest structures in the

### 2014-01-16 text/ms.tex
editing appendix

--- 2014-01-12 text/ms.tex
@@
 
-First let us recall the formulas for comoving distances.
+First let us recall various distance formulas. Comoving distances
+involve integrals of the type
 \begin{equation}
 D = \int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
+%\raise.5ex\hbox{.}
 \end{equation}
-The comoving distances from observer to lens, from lens to source, and
-from observer to source are given as follows:
+Specifically, the comoving distance $D_S$ from observer to source,
+$D_{LS}$ from lens to source, and $D_L$ from observer to lens are
+given by the above integral, with limits as follows.
 \begin{equation}
@@
 Angular-diameter distances $d$ can be related to comoving distances
-$D$ by:
+$D$ by scale factors, as follows.
 \begin{equation}
@@
 \end{equation}
-Angular-diameter distances are always smaller than comoving distances.
+Consequently, angular-diameter distances are always smaller than
+comoving distances.  Moreover, while comoving distances can be simply
+added, angular distances cannot:
+\begin{equation}
+d_S \neq d_L + d_{LS}
+\end{equation}
 
@@
 \begin{equation}
-(x,y) = d_L (\theta_x,\theta_y) \,.
+(x,y) = d_L (\theta_x,\theta_y)
 \end{equation}

### 2014-01-18 text/ms.tex
Correcting grammatical errors and adding some further clarification to the abstract.

@@
 \begin{abstract}
-In Spacewarps, volunteers are invited to search CFHTLS sky survey for
-lens candidates.  Here we report on a system that allows experienced
-volunteers to model lens candidates.  A sample of 29 simulated lenses
-were modeled, each multiple times.  The quality of the models was then
-examined in different ways.  In particular: (i)~the image parities and
+In Spacewarps, volunteers are invited to search the CFHTLS sky survey for
+gravitational lens candidates.  Here we report on a web interface    \todo[red]{or should that be application?} that allows experienced
+volunteers to model lens candidates found within the classification phase of the project. A sample of 29 simulated lenses
+were modeled multiple times by different users.  The quality of the models were then
+examined in several ways.  Particular attention was paid to the following: (i)~the image parities and
 time arrivals were identified correctly in most cases, but not
@@
 {\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org}
-(Saha is a co-investigator) has been launched.  {\em Spacewarps\/} is
+ has been launched.  {\em Spacewarps\/} is
 part of the {\em Zooniverse\/} family of citizen-science
@@
 are too difficult for robots and too large for specialists.  In {\em
- Spacewarps\/} itself, since the launch in May 2013, $\sim100$ square
-degrees (or a quarter of a percent of the sky) has been examined.  The
-survey is divided into patches of $440\times 440$ pixels, each of
+ Spacewarps\/} itself, survey sky images are presented for volunteers to identify lens candidates.  The first tranche of survey data, introduced from May 2013, is an area of $\sim172$ square
+degrees (or a quarter of a percent of the sky) from the Canada-France-Hawaii Telescope Legacy Survey.
+ The survey is divided into patches of $440\times 440$ pixels, each of
 which is seen by ten volunteers.  Simulated lenses are mixed in with
 the data, both to help train volunteers on what to look for, and to
-estimate completeness and purity.  A list of candidates is being
+estimate completeness and purity.
+These data had previously been analyzed by robots \cite{Gavazzi2012, More2012ApJ}.  
+  A list of candidates is being
 processed, but intermediate results show some very good candidates, as
-well re-discoveries of a few known lenses.
+well as re-discoveries of candidates previously found robotically.
 
@@
 The sims were produced using {\tt gravlens}
-\citep{2001astro.ph..2341K,2001astro.ph..2340K}. They were of three
+\citep{2001astro.ph..2341K,2001astro.ph..2340K} using the CFHTLS survey data
+and catalogues \cite{Coupon2009}. They were of three
 kinds, as follows.
@@
     \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
-    the intergalactic mass.
+    the underlying dark matter distribution.
 \end{enumerate}
@@
 
-\begin{figure}[htbp]
-  \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_1.png}
-  \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
-  \label{fig:ER_all_models}
-\end{figure}
-
 To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
@@
 To estimate the errors, the extremal models are used to estimate a lower and upper limit for \ERg.
-These results can be seen in \figref{ER_all_models}.
+These results can be seen in \figref{ER_per_sim}.
 This figure shows that this technique of estimating the error using the ensemble of all models underestimates the error significantly and should be improved for further analysis.
@@
   \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007021_kappa_encl}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
              {fig/007024_kappa_encl}}
   \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007025_kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007022_kappa_encl}}
-  \caption{\kenc for models of \asw{0h2m}: 7020 (top left), 7021 (top right), 7024 (mid left), 7025 (mid right) by volunteers and a correct model 7022 (bottom) by an expert.}
+             {fig/007021_kappa_encl}} \\
+  \caption{Enclosed mass \kenc for models of \asw{0h2m} with incorrect image identifications.
+\figref{7022} shows a model (by an expert) with the correct identifications, while \figref{7025} shows a model with error E06 (parities rotated by $\pi/2$).
+Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).}
 \label{fig:kapenc_compare_faulty}
@@
 
-E05 and E06 happen mostly in very symmetric images, that are hard to come up with a unique valid solution.
-This errors change the orientation of the mass distribution \kap[x,y],
-but \kenc and thus the \ERf is not influenced that much, and thus the final result is still a valid model.
-This can be seen by comparing the results for \asw{0h2m}, shown in \figref{kapenc_compare_faulty}: 
-The correct model 7022 ($\ERm=10.76px$), which was done by an expert, and
-7020 ($\ERm=10.72px$),
-7024 ($\ERm=10.80px$),
-7025 ($\ERm=11.16px$),
-7021 ($\ERm=11.04px$).
-The first three are of type E06, while the last is of type E05.
-This can be easily corrected, if further analysis is done for the modeled system and time delays are known.
-
-Comparing the models from volunteers and experts can be done in \figref{kapenc_compare_faulty}, where only the
-expert got the right configuration, but all the resulting models are comparable, besides rotation.
+Comparing the models from volunteers and experts can be done in Figures \ref{fig:7022}, \ref{fig:7025} and \ref{fig:kapenc_compare_faulty}, where only the
+expert got the right configuration (\figref{7022}) but enclosed masses are all fairly similar.
+In cases with nearly symmetric image configurations, like this example, it is more difficult to identify the image parities correctly.
+Incorrect image-parity identification changes the orientation of the mass distribution \kap[x,y],
+but \kenc and thus the \ERf are not influenced that much.
 
@@
 \begin{equation}
-D = \int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
+\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
 %\raise.5ex\hbox{.}
@@
 
+\section{Miscellaneous}
+
+Maybe discard \figref{ER_all_models}, since the essential information is there in \figref{ER_per_sim}.
+
+\begin{figure}[htbp]
+  \centering
+    \includegraphics[width=0.80\textwidth]{fig/eR_1.png}
+  \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
+  \label{fig:ER_all_models}
+\end{figure}
+
 \newpage

### 2014-01-18 text/ms.tex
Another edit of the Abstract.

@@
 \begin{abstract}
-In Spacewarps, volunteers are invited to search the CFHTLS sky survey for
-gravitational lens candidates.  Here we report on a web interface    \todo[red]{or should that be application?} that allows experienced
-volunteers to model lens candidates found within the classification phase of the project. A sample of 29 simulated lenses
-were modeled multiple times by different users.  The quality of the models were then
-examined in several ways.  Particular attention was paid to the following: (i)~the image parities and
-time arrivals were identified correctly in most cases, but not
-always; (ii)~the mean convergence (equivalent to the enclosed mass)
-was quite well constrained in the image region; (iii)~the performance
-of experienced volunteers was comparable to professionals.
+In the initial phase of Spacewarps (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses. Here we report on a web interface    
+that gives experienced volunteers the opportunity to model the candidates that have been identified.
+These volunteers, were also invited to model a sample of 29 simulated lenses, to order to gauge the quality of the models that were being generated.
+These were then examined in greater detail, with particular attention being paid to the following: (i)~identification of image parities and time arrivals; (ii)~the mean convergence (equivalent to the enclosed mass), and finally;
+ (iii)~ the performance of a volunteer vs.
+a professional.
+In most cases, the volunteers were found to correctly identify the image parities and time arrivals, along with a mean convergence that was contained within the image region; In all, the results could be comparable to that of professionals.
+
+\todo[red]{or should that be web application?
+Also, I change the tense of the opening statement.
+Not too sure of the correct way of introducing an abbreviation (SW1 in this case), and I was a bit liberal with my editing - Sorry.
+Another comment - is 'Well contained' better than 'contained' or is it just a figure of speech?}
+
 \end{abstract}
@@
 
-xxx
 Our present picture of the formation of the largest structures in the

### 2014-01-20 text/ms.tex
only reformatting orginial changes

@@
 \usepackage{float}      % exact placement of figures (using \begin{figure}[H])
-\usepackage{color}  % well... color...
-\usepackage{setspace}    % remove this pgk for final compile!! 
-\usepackage{subfigure}
-
-\usepackage{longtable}   % those two to generate the 2 ol table
-\usepackage{multicol}
+\usepackage{subfigure}  % does what it says
+
+%\usepackage{longtable}   % those two to generate the 2 ol table
+%\usepackage{multicol}
 
@@
 \newcommand{\asw}[1]{ASW000#1\xspace}
-
 \newcommand{\model}[2][Model~]{#1#2\xspace}
@@
 % todo annotations on side in red. easy to search for at the end
-\newcommand{\todo}[2][red]{%\marginpar{\raggedright\scriptsize\sffamily\textcolor{red}{#1}}}
-
+% !!! REMOVE THIS LINES FOR FINAL COMPILE
+%
+\usepackage{setspace}    % remove this pgk for final compile!! 
+\usepackage{color}    % well... color...
+
+\newcommand{\todo}[2][red]{%
 \textcolor{#1}{\textbullet}%
@@
 \newcommand{\needchk}[1][]{\todo[red]{chk #1}}
+
+% \\\ END OF REMOVAL SECION
 
@@
 \begin{abstract}
-In the initial phase of Spacewarps (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses. Here we report on a web interface    
-that gives experienced volunteers the opportunity to model the candidates that have been identified.
+In the initial phase of Spacewarps (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses.
+Here we report on a web interface that gives experienced volunteers the opportunity to model the candidates that have been identified.
 These volunteers, were also invited to model a sample of 29 simulated lenses, to order to gauge the quality of the models that were being generated.
-These were then examined in greater detail, with particular attention being paid to the following: (i)~identification of image parities and time arrivals; (ii)~the mean convergence (equivalent to the enclosed mass), and finally;
- (iii)~ the performance of a volunteer vs.
-a professional.
-In most cases, the volunteers were found to correctly identify the image parities and time arrivals, along with a mean convergence that was contained within the image region; In all, the results could be comparable to that of professionals.
-
-\todo[red]{or should that be web application?
-Also, I change the tense of the opening statement.
-Not too sure of the correct way of introducing an abbreviation (SW1 in this case), and I was a bit liberal with my editing - Sorry.
-Another comment - is 'Well contained' better than 'contained' or is it just a figure of speech?}
+These were then examined in greater detail, with particular attention being paid to the following:
+(i)~identification of image parities and time arrivals;
+(ii)~the mean convergence (equivalent to the enclosed mass), and finally;
+(iii)~the performance of a volunteer vs. a professional.
+In most cases, the volunteers were found to correctly identify the image parities and time arrivals, along with a mean convergence that was contained within the image region;
+In all, the results could be comparable to that of professionals.
+
+\todo{see code}
+% or should that be web application?
+% Also, I change the tense of the opening statement.
+% Not too sure of the correct way of introducing an abbreviation (SW1 in this case), and I was a bit liberal with my editing - Sorry.
+% Another comment - is 'Well contained' better than 'contained' or is it just a figure of speech?}
 

### 2014-01-20 text/ms.tex
minor changes

@@
 \begin{abstract}
-In the initial phase of Spacewarps (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses.
-Here we report on a web interface that gives experienced volunteers the opportunity to model the candidates that have been identified.
-These volunteers, were also invited to model a sample of 29 simulated lenses, to order to gauge the quality of the models that were being generated.
+In the initial phase of \sw (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses.
+Here we report on a web application that gives experienced volunteers the opportunity to model the candidates that have been identified.
+These volunteers were also invited to model a sample of 29 simulated lenses in order to gauge the quality of the models that were being generated.
 These were then examined in greater detail, with particular attention being paid to the following:
@@
 (iii)~the performance of a volunteer vs. a professional.
-In most cases, the volunteers were found to correctly identify the image parities and time arrivals, along with a mean convergence that was contained within the image region;
+In most cases, the volunteers were found to correctly identify the image parities and time arrivals, along with a mean convergence that was contained within the image region.
 In all, the results could be comparable to that of professionals.
-
-\todo{see code}
-% or should that be web application?
-% Also, I change the tense of the opening statement.
-% Not too sure of the correct way of introducing an abbreviation (SW1 in this case), and I was a bit liberal with my editing - Sorry.
-% Another comment - is 'Well contained' better than 'contained' or is it just a figure of speech?}
 

### 2014-01-20 text/ms.tex
increase resolution of allsky, and add vector format

@@
 \begin{abstract}
-In the initial phase of \sw (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses.
-Here we report on a web application that gives experienced volunteers the opportunity to model the candidates that have been identified.
-These volunteers were also invited to model a sample of 29 simulated lenses in order to gauge the quality of the models that were being generated.
-These were then examined in greater detail, with particular attention being paid to the following:
-(i)~identification of image parities and time arrivals;
-(ii)~the mean convergence (equivalent to the enclosed mass), and finally;
-(iii)~the performance of a volunteer vs. a professional.
-In most cases, the volunteers were found to correctly identify the image parities and time arrivals, along with a mean convergence that was contained within the image region.
-In all, the results could be comparable to that of professionals.
-
+In Spacewarps, volunteers are invited to search CFHTLS sky survey for
+lens candidates.  Here we report on a system that allows experienced
+volunteers to model lens candidates.  A sample of 29 simulated lenses
+were modeled, each multiple times.  The quality of the models was then
+examined in different ways.  In particular: (i)~the image parities and
+time arrivals were identified correctly in most cases, but not
+always; (ii)~the mean convergence (equivalent to the enclosed mass)
+was quite well constrained in the image region; (iii)~the performance
+of experienced volunteers was comparable to professionals.
 \end{abstract}
@@
 
-\begin{figure}
-\centering
-\includegraphics[width=.8\hsize]{fig/lenssky.png}
+\begin{figure}[h]
+\centering
+\includegraphics[width=.7\hsize]{fig/lenssky}
 \caption{Sky distribution of 398 published lenses considered secure as

### 2014-01-20 allsky/map.py

@@ -1,5 +1,5 @@
 from numpy import isnan, pi, cos, sin, sqrt, linspace
-from pylab import figure, show
+from pylab import figure, show, savefig, subplots_adjust
 
 
 lis = open('alist.csv').readlines()[2:]
@@ -48,10 +48,9 @@ panel.xaxis.set_visible(False)
 panel.yaxis.set_visible(False)
 panel.set_xlim(-1,1)
 panel.set_ylim(-0.5,0.5)
-show()
+#show()
 
-
-
-
-
-
+# patch to fill whole image, remove white border, use vecort graphics
+subplots_adjust(left=0.0, right=1.0, bottom=0.0, top=1.0)
+savefig("lenssky.png", dpi=200, bbox_inches='tight', pad_inches=0)
+savefig("lenssky.pdf", dpi=100, bbox_inches='tight', pad_inches=0)

### 2014-01-20 plots/gen_plots.py
k_encl added asw id and vecorised

@@ -97,7 +97,7 @@ write_to_tex_folder = True
 ext = 'pdf'
 
 # realtive to plots dir
-outdir = 'figs'
+outdir = 'figs_new'
 
 simdir = 'sim'
 moddir = 'mod'
@@ -815,13 +815,29 @@ def draw_mod(mid, elem, data, sims):
   pl.ylabel(r'mean convergance [1]')  
   
   pl.xlim([0,np.max(elem['x'])])
+
+  ax = pl.gca()
+  pl.text(0.95, 0.95,elem['name'],
+    horizontalalignment='right',
+    verticalalignment='top',
+    fontsize=14,
+    transform = ax.transAxes)
+
   
   if show:
     #print 'show'
     pl.show()
   else:
-    imgname = ('kappa_encl.%s'%ext)
-    pl.savefig(os.path.join(save_fig_path, imgname))
+    #imgname = ('kappa_encl.%s'%ext)
+    #pl.savefig(os.path.join(save_fig_path, imgname))
+    
+    # new style direct image names \figs_new\mod\001234_kappa_encl.ext
+    # save as png and pdf anyways
+    imgname1 = ('_kappa_encl.%s'%'png')
+    imgname2 = ('_kappa_encl.%s'%'pdf')
+    imgname = ('_kappa_encl.%s'%'pdf')
+    pl.savefig(os.path.join(save_fig_path + imgname1))
+    pl.savefig(os.path.join(save_fig_path + imgname2))
     pass
   
   

### 2014-01-20 plots/gen_plots.py
vecotrized ASW plots

@@ -616,7 +616,9 @@ def draw_sim(asw, sim):
     lev = np.linspace(0,10,41)
     pc = panel.contour(x,y,kappa,lev)
     panel.clabel(pc, inline=1, fontsize=10)
-    pl.savefig(os.path.join(path, '%s.%s'%(filenames[1],ext)))
+    #pl.savefig(os.path.join(path, '%s.%s'%(filenames[1],ext)))
+    pl.savefig(os.path.join(path + '_%s.%s'%(filenames[1],'pdf')))
+    pl.savefig(os.path.join(path + '_%s.%s'%(filenames[1],'png')))
 
 
     fig = pl.figure()
@@ -644,7 +646,9 @@ def draw_sim(asw, sim):
     panel.scatter(rad,sum)
     panel.set_xlabel('radius [pixels]')
     panel.set_ylabel('average interior \textkappa [1]')
-    pl.savefig(os.path.join(path, '%s.%s'%(filenames[2],ext)))
+    #pl.savefig(os.path.join(path, '%s.%s'%(filenames[2],ext)))
+    pl.savefig(os.path.join(path + '_%s.%s'%(filenames[2],'png')))
+    pl.savefig(os.path.join(path + '_%s.%s'%(filenames[2],'pdf')))
     
     
     fig = pl.figure()
@@ -653,7 +657,9 @@ def draw_sim(asw, sim):
     lo,hi = np.amin(arriv), np.amax(arriv)
     lev = np.linspace(lo,lo+.2*(hi-lo),100)
     panel.contour(x,y,arriv,lev)
-    pl.savefig(os.path.join(path, '%s.%s'%(filenames[0],ext)))
+    #pl.savefig(os.path.join(path, '%s.%s'%(filenames[0],ext)))
+    pl.savefig(os.path.join(path + '_%s.%s'%(filenames[0],'pdf')))
+    pl.savefig(os.path.join(path + '_%s.%s'%(filenames[0],'png')))
 
     #restore stdout
     sys.stdout = old_stdout

### 2014-01-20 text/ms.tex
k_encl added asw id and vecorised
vecotrized ASW plots
made tex extension independant
it prefers now .pdf if both available, otherwise use .png

@@
 \usepackage{subfigure}  % does what it says
+
+\usepackage{graphicx}
+% prefere pdf over png (better quality, lossless compression)
+\DeclareGraphicsExtensions{%
+  .pdf,.PDF,%
+  .png,.PNG,%
+  .jpg,.mps,.jpeg,.jbig2,.jb2,.JPG,.JPEG,.JBIG2,.JB2}
 
@@
 \centering
-\includegraphics[height=.3\vsize]{fig/arriv1.png}
-\includegraphics[height=.3\vsize]{fig/arriv2.png}
-\includegraphics[height=.3\vsize]{fig/arriv3.png}
+\includegraphics[height=.3\vsize]{fig/arriv1}
+\includegraphics[height=.3\vsize]{fig/arriv2}
+\includegraphics[height=.3\vsize]{fig/arriv3}
 \caption{Arrival-time surfaces, with contour levels of equal arrival
@@
   \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/006941_input.png}} \\
+             {fig/006941_input}} \\
   \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007022_input.png}} \\
+             {fig/007022_input}} \\
   \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/006919_input.png}}
+             {fig/006919_input}}
   \caption{Examples of Spaghetti input.  The models from these appear
@@
     \label{fig:6941_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa}
   }
@@
     \label{fig:6941_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv}
   }
@@
     \label{fig:6941_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006941_mass.png}
+    \includegraphics[width=0.45\textwidth]{fig/006941_mass}
   }
@@
     \label{fig:6941_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti.png}
+    \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti}
   }
@@
     \label{fig:6941_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl.png}
+    \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl}
   }
@@
     \label{fig:6941_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006941_arr_time.png}
+    \includegraphics[width=0.45\textwidth]{fig/006941_arr_time}
   }
@@
     \label{fig:6975_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_kappa.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_kappa}
   }
@@
     \label{fig:6975_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_arriv.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_arriv}
   }
@@
     \label{fig:6975_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006975_mass.png}
+    \includegraphics[width=0.45\textwidth]{fig/006975_mass}
   }
@@
     \label{fig:6975_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006975_spaghetti.png}
+    \includegraphics[width=0.45\textwidth]{fig/006975_spaghetti}
   }
@@
     \label{fig:6975_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006975_kappa_encl.png}
+    \includegraphics[width=0.45\textwidth]{fig/006975_kappa_encl}
   }
@@
     \label{fig:6975_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006975_arr_time.png}
+    \includegraphics[width=0.45\textwidth]{fig/006975_arr_time}
   }
@@
     \label{fig:6937_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_kappa.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_kappa}
   }
@@
     \label{fig:6937_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_arriv.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_arriv}
   }
@@
     \label{fig:6937_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006937_mass.png}
+    \includegraphics[width=0.45\textwidth]{fig/006937_mass}
   }
@@
     \label{fig:6937_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006937_spaghetti.png}
+    \includegraphics[width=0.45\textwidth]{fig/006937_spaghetti}
   }
@@
     \label{fig:6937_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006937_kappa_encl.png}
+    \includegraphics[width=0.45\textwidth]{fig/006937_kappa_encl}
   }
@@
     \label{fig:6937_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006937_arr_time.png}
+    \includegraphics[width=0.45\textwidth]{fig/006937_arr_time}
   }
@@
     \label{fig:6990_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_kappa.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_kappa}
   }
@@
     \label{fig:6990_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_arriv.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_arriv}
   }
@@
     \label{fig:6990_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006990_mass.png}
+    \includegraphics[width=0.45\textwidth]{fig/006990_mass}
   }
@@
     \label{fig:6990_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006990_spaghetti.png}
+    \includegraphics[width=0.45\textwidth]{fig/006990_spaghetti}
   }
@@
     \label{fig:6990_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006990_kappa_encl.png}
+    \includegraphics[width=0.45\textwidth]{fig/006990_kappa_encl}
   }
@@
     \label{fig:6990_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006990_arr_time.png}
+    \includegraphics[width=0.45\textwidth]{fig/006990_arr_time}
   }
@@
     \label{fig:6915_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_kappa.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_kappa}
   }
@@
     \label{fig:6915_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_arriv.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_arriv}
   }
@@
     \label{fig:6915_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006915_mass.png}
+    \includegraphics[width=0.45\textwidth]{fig/006915_mass}
   }
@@
     \label{fig:6915_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006915_spaghetti.png}
+    \includegraphics[width=0.45\textwidth]{fig/006915_spaghetti}
   }
@@
     \label{fig:6915_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006915_kappa_encl.png}
+    \includegraphics[width=0.45\textwidth]{fig/006915_kappa_encl}
   }
@@
     \label{fig:6915_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006915_arr_time.png}
+    \includegraphics[width=0.45\textwidth]{fig/006915_arr_time}
   }
@@
     \label{fig:7022_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa}
   }
@@
     \label{fig:7022_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv}
   }
@@
     \label{fig:7022_mass}
-    \includegraphics[width=0.45\textwidth]{fig/007022_mass.png}
+    \includegraphics[width=0.45\textwidth]{fig/007022_mass}
   }
@@
     \label{fig:7022_cont}
-    \includegraphics[width=0.45\textwidth]{fig/007022_spaghetti.png}
+    \includegraphics[width=0.45\textwidth]{fig/007022_spaghetti}
   }
@@
     \label{fig:7022_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/007022_kappa_encl.png}
+    \includegraphics[width=0.45\textwidth]{fig/007022_kappa_encl}
   }
@@
     \label{fig:7022_atime}
-    \includegraphics[width=0.45\textwidth]{fig/007022_arr_time.png}
+    \includegraphics[width=0.45\textwidth]{fig/007022_arr_time}
   }
@@
     \label{fig:7025_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa}
   }
@@
     \label{fig:7025_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv}
   }
@@
     \label{fig:7025_mass}
-    \includegraphics[width=0.45\textwidth]{fig/007025_mass.png}
+    \includegraphics[width=0.45\textwidth]{fig/007025_mass}
   }
@@
     \label{fig:7025_cont}
-    \includegraphics[width=0.45\textwidth]{fig/007025_spaghetti.png}
+    \includegraphics[width=0.45\textwidth]{fig/007025_spaghetti}
   }
@@
     \label{fig:7025_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/007025_kappa_encl.png}
+    \includegraphics[width=0.45\textwidth]{fig/007025_kappa_encl}
   }
@@
     \label{fig:7025_atime}
-    \includegraphics[width=0.45\textwidth]{fig/007025_arr_time.png}
+    \includegraphics[width=0.45\textwidth]{fig/007025_arr_time}
   }
@@
     \label{fig:6919_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_kappa.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_kappa}
   }
@@
     \label{fig:6919_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_arriv.png}
+    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_arriv}
   }
@@
     \label{fig:6919_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006919_mass.png}
+    \includegraphics[width=0.45\textwidth]{fig/006919_mass}
   }
@@
     \label{fig:6919_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006919_spaghetti.png}
+    \includegraphics[width=0.45\textwidth]{fig/006919_spaghetti}
   }
@@
     \label{fig:6919_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006919_kappa_encl.png}
+    \includegraphics[width=0.45\textwidth]{fig/006919_kappa_encl}
   }
@@
     \label{fig:6919_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006919_arr_time.png}
+    \includegraphics[width=0.45\textwidth]{fig/006919_arr_time}
   }
@@
   \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_4.png}
+    \includegraphics[width=0.80\textwidth]{fig/eR_4}
   \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
@@
   \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_1.png}
+    \includegraphics[width=0.80\textwidth]{fig/eR_1}
   \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}

### 2014-01-22 text/ms.tex

--- 2014-01-20 text/ms.tex
@@
 (iii)~the performance of a volunteer vs. a professional.
-In most cases, the volunteers were found to correctly identify the image parities and time arrivals, along with a mean convergence that was contained within the image region.
+In most cases, the volunteers were found to correctly identify the image parities and time arrivals.
+% Einstein radius +~20%
+ along with a mean convergence that was contained within the image region.
 In all, the results could be comparable to that of professionals.

### 2014-01-22 text/ms.tex

--- 2014-01-20 text/ms.tex
@@
 \begin{abstract}
-In Spacewarps, volunteers are invited to search CFHTLS sky survey for
-lens candidates.  Here we report on a system that allows experienced
-volunteers to model lens candidates.  A sample of 29 simulated lenses
-were modeled, each multiple times.  The quality of the models was then
-examined in different ways.  In particular: (i)~the image parities and
-time arrivals were identified correctly in most cases, but not
-always; (ii)~the mean convergence (equivalent to the enclosed mass)
-was quite well constrained in the image region; (iii)~the performance
-of experienced volunteers was comparable to professionals.
+In the initial phase of \sw (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses.
+Here we report on a web application that gives experienced volunteers the opportunity to model the candidates that have been identified.
+These volunteers were also invited to model a sample of 29 simulated lenses in order to gauge the quality of the models that were being generated.
+These were then examined in greater detail, with particular attention being paid to the following:
+(i)~identification of image parities and time arrivals;
+(ii)~the mean convergence (equivalent to the enclosed mass), and finally;
+(iii)~the performance of a volunteer vs. a professional.
+In most cases, the volunteers were found to correctly identify the image parities and time arrivals.
+% Einstein radius +~20%
+ along with a mean convergence that was contained within the image region.
+In all, the results could be comparable to that of professionals.
+
 \end{abstract}

### 2014-01-23 text/ms.tex

@@
 Here we report on a web application that gives experienced volunteers the opportunity to model the candidates that have been identified.
-These volunteers were also invited to model a sample of 29 simulated lenses in order to gauge the quality of the models that were being generated.
-These were then examined in greater detail, with particular attention being paid to the following:
+In order to gauge the quality of the models that were being rendered, these same volunteers were invited to model a sample of 29 simulated lenses.
+The models were then examined in greater detail, with particular attention being paid to the following:
 (i)~identification of image parities and time arrivals;
@@
 % Einstein radius +~20%
- along with a mean convergence that was contained within the image region.
-In all, the results could be comparable to that of professionals.
+ along with a mean convergence that was well constrained within the image region;
+In all, the results could be comparable to that of a professional.
 

### 2014-01-24 text/ms.tex

@@
 
-Our present picture of the formation of the largest structures in the
-Universe --- galaxies and clusters of galaxies --- is that
-gravitational instabilities (whose early stages are observable as
-fluctuations of the microwave background) caused a collapse into
-gravitationally bound structures.  Dark matter would have dominated
-the initial collapse, but once deep wells in the gravitational
-potential appeared, gas would have fallen into them while radiating
-away its gravitational potential energy through atomic processes to
-form the first stars and galaxies.  With time, galaxies and clusters
-would continue to merge and change their form.  This rough outline is
-well-established now, but many puzzles remain to be solved before our
-understanding is satisfactory.  To progress towards solving these, we
-need more knowledge of dark matter.  And to study dark matter,
-gravitational lensing is attractive, because it depends only on mass.
+Galaxies and clusters of galaxies are considered to be the largest structures in the known Universe.
+When one questions their formation,  the present picture suggests that gravitational instabilities  (observable as fluctuations of the microwave background)
+ \todo[red]{not too sure of the relevance of this statement} 
+ - caused a collapse into gravitationally bound structures. 
+ \todo[yellow]{clarification needed: are they becoming gravitationally bound structures, or are they collapsing into objects that are already gravitationally bound?
+I know it is most probably the later, but it is a bit ambiguous}
+The initial collapse would have been dominated by dark matter, but once deep wells had appeared in the gravitational potential, gas would have started falling into those wells. 
+\todo[red]{not too sure of this wording}
+Through atomic processes,
+\todo[yellow]{perhaps we need to clarify those processes or are they generally know?}
+ the gravitational potential energy of this gas would of been radiated away, leading to the formation of the largest stars. 
+\todo[red]{or perhaps just stars?}
+In turn, galaxies and clusters of galaxies would form, all the time merging and changing their form.
+This rough outline is now well established, but in order to satisfy our understanding of this process, there are many puzzles  that remain to be solved.
+For progress to be made, we need to extend our knowledge of dark matter.
+Gravitational lensing has proven as a particularly attractive way of studying dark matter as it is only dependant on mass.
+\todo[yellow]{silly question, but shouldn't we briefly explain why it is only dependant on mass? or should everyone already reading this paper know that? or is it explained later?}
 

### 2014-01-24 text/ms.tex
added bylines

@@
 \title{Lens Modeling in \sw}
-\author{First Author,\altaffilmark{1}
-Second Author,\altaffilmark{2} and
-Third Author\altaffilmark{3}}
-\altaffiltext{1}{First place}
-\altaffiltext{2}{Second place}
-\altaffiltext{3}{Third place}
+\author{Rafael Kueng,\altaffilmark{1}
+Phil Marshall,\altaffilmark{2}
+Anupreeta More,\altaffilmark{3}
+Aprajita Verma,\altaffilmark{4}
+Amit Kapadia,\altaffilmark{5}
+Elisabeth Baeten,\altaffilmark{6}
+Claude Cornen,\altaffilmark{6}
+Christine McMillan,\altaffilmark{6}
+Julianne Wilcox,\altaffilmark{6}
+Jonas Odermatt,\altaffilmark{7}
+Jonathan Coles,\altaffilmark{8}
+Surhud More\altaffilmark{3}
+and Prasenjit Saha\altaffilmark{1}
+}
+\altaffiltext{1}{Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland}
+\altaffiltext{2}{Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA}
+\altaffiltext{3}{Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan}
+\altaffiltext{4}{Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK}
+\altaffiltext{5}{Adler Planetarium, 1300 S. Lake Shore Drive,
+  Chicago, IL 60605, USA}
+\altaffiltext{6}{Spacewarps}
+\altaffiltext{7}{Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland}
+\altaffiltext{8}{Exascale
+  Research Computing Lab, Versailles Saint-Quentin-en-Yvelines
+  University, 55 Avenue des Etats Unis, 78000, Versailles, France}
 

### 2014-01-24 text/ms.tex
edits

--- 2014-01-24 text/ms.tex
@@
 \newcommand{\spl}{SpaghettiLens\xspace}
-\newcommand{\sw}{SpaceWarps\xspace}
+\newcommand{\sw}{Spacewarps\xspace}
 
@@
 \includegraphics[width=.7\hsize]{fig/lenssky}
-\caption{Sky distribution of 398 published lenses considered secure as
-  of September 2013 (from the Masterlens catalog at the University of
+\caption{Sky distribution of 423 published lenses considered secure as
+  of January 2014 (from the Masterlens catalog at the University of
   Utah, maintained by Joel Brownstein and Leonidas Moustakas.) The map
@@
 
-% \subsection{\sw} \label{sec:SpaceWarps} some words about \sw
-
 Interested volunteers from the \sw forum were initially introduced to
@@
   }
-  \caption[result 6941 (ASW000102p)]{Sim ASW000102p and a model.  A
-    simple configuration; the model input is shown in the top panel of
+  \caption[result 6941 (ASW000102p)]{Model results for a simple
+    simulated lens.  (The label ASW000102p in the lower-left panel is
+    the \sw identifier for the sky patch where this sim was placed.)
+    The model input is shown in the top panel of
     Figure~\ref{fig:input-spag}.  For details of the individual panels
@@
   }
-  \caption[result 6975 (ASW000195x)]{Sim ASW000195x and a
-    model.}
+  \caption[result 6975 (ASW000195x)]{A model for a sim with some mass
+    substructure. (See Section \ref{sec:example_models} for details.)}
   \label{fig:6975}
@@
   }
-  \caption[result 6937 (ASW0000vqg)]{Sim ASW0000vqg and a model.}
+  \caption[result 6937 (ASW0000vqg)]{A sim with substructure,
+    resulting in a poor mass model. (See Section
+    \ref{sec:example_models} for details.)}
   \label{fig:6937}
@@
 
-  \caption[result 6990 (ASW0004oux)]{Sim ASW0004oux and a model.}
+  \caption[result 6990 (ASW0004oux)]{Model of a four-image
+    system. (See Section \ref{sec:example_models} for details.)}
   \label{fig:6990}
@@
 
-  \caption[result 6915 (ASW0001hpf)]{Sim ASW0001hpf and a model. Inclined quad.}
+  \caption[result 6915 (ASW0001hpf)]{Model of another four-image
+    system. (See Section \ref{sec:example_models} for details.)}
   \label{fig:6915}
@@
 
-  \caption[result 7022 (ASW0000h2m)]{Sim ASW0000h2m and a model.  The
+  \caption[result 7022 (ASW0000h2m)]{A further four-image system. The
     model input appears in the middle panel of Figure
-    \ref{fig:input-spag}.}
+    \ref{fig:input-spag}. (See Section \ref{sec:example_models} for
+    details.)}
   \label{fig:7022}
@@
 
-  \caption[result 7025 (ASW0000h2m)]{Sim ASW0000h2m and another model.  The image parities in this case were incorrectly identified.}
+  \caption[result 7025 (ASW0000h2m)]{Another model for the lens
+    appearing in \figref{fig:7022}.  The image parities in this case
+    were incorrectly identified, resulting in a rather poor model, but
+    the estimated Einstein radius is not bad. (See Section
+    \ref{sec:example_models} for details.)}
   \label{fig:7025}
@@
   }
-  \caption[result 6919 (ASW0002z6f)]{Sim ASW0002z6f and a model.  The
-    model input appears in the bottom panel of Figure
-    \ref{fig:input-spag}. An incipient quad}
+  \caption[result 6919 (ASW0002z6f)]{A sim on the border case of
+    double vs quad; modelled as a quad.  The model input appears in
+    the bottom panel of Figure \ref{fig:input-spag}. (See Section
+    \ref{sec:example_models} for details.)}
   \label{fig:6919}
@@
              {fig/007021_kappa_encl}} \\
-  \caption{Enclosed mass \kenc for models of \asw{0h2m} with incorrect image identifications.
+  \caption{Enclosed mass \kenc from three models of a system with incorrect image identifications.
 \figref{7022} shows a model (by an expert) with the correct identifications, while \figref{7025} shows a model with error E06 (parities rotated by $\pi/2$).
@@
 
+\section{Acknowledgments}
+
+We thank the Swiss Society for Astrophysics and Astronomy and the
+Swiss Academy of Sciences.
+
 \clearpage

### 2014-02-21 text/ms.tex
removed needcitation marks from display in pdf

@@
 
-\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
-\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
-\newcommand{\needchk}[1][]{\todo[red]{chk #1}}
+%\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
+%\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
+%\newcommand{\needchk}[1][]{\todo[red]{chk #1}}
 
@@
 Since this is an iterative learning process, the more involved volunteers will quickly gain knowledge that can be passed down to new volunteers.
-That creates a social structure that scales well with the number of volunteers, as other projects\needcite have already shown.
+That creates a social structure that scales well with the number of volunteers, as other projects have already shown. %TODO \needcite
 Finding people working as volunteers has been shown to be successful last but not least by \sw and the whole galaxy zoo project.
@@
 allowing only valid lensing configurations to be entered, and ensuring
-that the odd image theorem\needcite is taken care of.
+that the odd image theorem is taken care of.%TODO \needcite
 

### 2014-02-21 text/ms.tex
moved comments to code for nice print
note they are NOT YET resolved, just not anymore displayed...

@@
 
-\newcommand{\todo}[2][red]{%
-\textcolor{#1}{\textbullet}%
-\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
-\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}
+%\newcommand{\todo}[2][red]{%
+%\textcolor{#1}{\textbullet}%
+%\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
+%\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}
 
@@
 
-%\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
-%\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
-%\newcommand{\needchk}[1][]{\todo[red]{chk #1}}
+\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
+\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
+\newcommand{\needchk}[1][]{\todo[red]{chk #1}}
 
@@
 Galaxies and clusters of galaxies are considered to be the largest structures in the known Universe.
-When one questions their formation,  the present picture suggests that gravitational instabilities  (observable as fluctuations of the microwave background)
- \todo[red]{not too sure of the relevance of this statement} 
+When one questions their formation,  the present picture suggests that gravitational instabilities  (observable as fluctuations of the microwave background) %TODO \todo[red]{not too sure of the relevance of this statement}
  - caused a collapse into gravitationally bound structures.
- \todo[yellow]{clarification needed: are they becoming gravitationally bound structures, or are they collapsing into objects that are already gravitationally bound?
+%TODO\todo[yellow]{clarification needed: are they becoming gravitationally bound structures, or are they collapsing into objects that are already gravitationally bound?
 I know it is most probably the later, but it is a bit ambiguous}
 The initial collapse would have been dominated by dark matter, but once deep wells had appeared in the gravitational potential, gas would have started falling into those wells. 
-\todo[red]{not too sure of this wording}
+%TODO \todo[red]{not too sure of this wording}
 Through atomic processes,
-\todo[yellow]{perhaps we need to clarify those processes or are they generally know?}
+%TODO \todo[yellow]{perhaps we need to clarify those processes or are they generally know?}
  the gravitational potential energy of this gas would of been radiated away, leading to the formation of the largest stars. 
-\todo[red]{or perhaps just stars?}
+%TODO \todo[red]{or perhaps just stars?}
 In turn, galaxies and clusters of galaxies would form, all the time merging and changing their form.
@@
 Gravitational lensing has proven as a particularly attractive way of studying dark matter as it is only dependant on mass.
-\todo[yellow]{silly question, but shouldn't we briefly explain why it is only dependant on mass? or should everyone already reading this paper know that? or is it explained later?}
+%TODO \todo[yellow]{silly question, but shouldn't we briefly explain why it is only dependant on mass? or should everyone already reading this paper know that? or is it explained later?}
 

### 2014-02-24 text/ms.tex
edits

@@
 spaghetti configuration, \spl sends the input to its server-side
-modeling engine, called GLASS.  The task of GLASS is to find a mass
-distribution $\kappa(x,y)$ that exactly reproduces the locations of
-the maximum, minima and saddle points. Now, this criterion by itself
-is extremely under-determined --- there are infinitely many mass
-distributions that will reproduce a given set of maxima, minima and
-saddle points, but typically they (a)~produce lots of extra images,
-and (b)~look very unlike galaxies.  Additional assumptions (a prior)
-are necessary.  GLASS uses a prior based on suggestions by
-\cite{1997MNRAS.292..148S}.
+modeling engine, called GLASS \citep{2014arXiv1401.7990C}.  The task
+of GLASS is to find a mass distribution $\kappa(x,y)$ that exactly
+reproduces the locations of the maximum, minima and saddle
+points. Now, this criterion by itself is extremely under-determined
+--- there are infinitely many mass distributions that will reproduce a
+given set of maxima, minima and saddle points, but typically they
+(a)~produce lots of extra images, and (b)~look very unlike galaxies.
+Additional assumptions (a prior) are necessary.  GLASS uses a prior
+based on suggestions by \cite{1997MNRAS.292..148S}.
 \begin{enumerate}
@@
   \caption[result 7025 (ASW0000h2m)]{Another model for the lens
-    appearing in \figref{fig:7022}.  The image parities in this case
+    appearing in \figref{7022}.  The image parities in this case
     were incorrectly identified, resulting in a rather poor model, but

### 2014-04-02 text/ms.tex
changed to mn2e format

--- 2014-02-21 text/ms.tex
@@
-\documentclass[12pt,preprint]{aastex}
+\documentclass[usenatbib,onecolumn]{mn2e}
 
@@
 
-%\usepackage{longtable}   % those two to generate the 2 ol table
-%\usepackage{multicol}
+\newcommand{\apj}{ApJ}
+\newcommand{\apjs}{ApJS}
+\newcommand{\apjl}{ApJL}
+\newcommand{\aap}{A{\&}A}
+\newcommand{\aaps}{A{\&}AS}
+\newcommand{\mnras}{MNRAS}
+\newcommand{\aj}{AJ}
+\newcommand{\araa}{ARAA}
+\newcommand{\pasp}{PASP}
+\newcommand{\pasj}{PASJ}
 
@@
 \title{Lens Modeling in \sw}
-\author{First Author,\altaffilmark{1}
-Second Author,\altaffilmark{2} and
-Third Author\altaffilmark{3}}
-\altaffiltext{1}{First place}
-\altaffiltext{2}{Second place}
-\altaffiltext{3}{Third place}
+
+\author[Kueng et al]{Rafael Kueng,$^{1}$
+Phil Marshall,$^{2}$
+Anupreeta More,$^{3}$
+Aprajita Verma,$^{4}$
+Amit Kapadia,$^{5}$
+\newauthor
+Elisabeth Baeten,$^{6}$
+Claude Cornen,$^{6}$
+Christine McMillan,$^{6}$
+Julianne Wilcox,$^{6}$
+\newauthor
+Jonas Odermatt,$^{7}$
+Jonathan Coles,$^{8}$
+Surhud More$^{3}$
+and Prasenjit Saha$^{1}$ \\
+$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
+$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
+$^{3}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
+$^{4}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
+$^{5}$Adler Planetarium, 1300 S. Lake Shore Drive, Chicago, IL 60605, USA\\
+$^{6}$Spacewarps\\
+$^{7}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
+$^{8}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France}
+
+\maketitle
 
@@
 In all, the results could be comparable to that of a professional.
-
 \end{abstract}
 
-\keywords{}
+\begin{keywords}
+\end{keywords}
+
 \section{Introduction}
@@
 
-\begin{figure}[h]
+\begin{figure}
 \centering
@@
 
-\bibliographystyle{apj}
+\bibliographystyle{mn2e}
 \bibliography{bib/ms}

### 2014-05-18 plots/gen_plots.py
updated plot style
added input plots, made mass distr and arr time plots look similar

@@ -96,8 +96,10 @@ write_to_tex_folder = True
 #ext = 'png'
 ext = 'pdf'
 
+figsize=(10,10)
+
 # realtive to plots dir
-outdir = 'figs_new'
+outdir = 'figs_new2'
 
 simdir = 'sim'
 moddir = 'mod'
@@ -126,7 +128,7 @@ sys.path.append(os.path.join(pardir, 'systems'))
 import gen_kappa_encl_plots as spg
 import many
 
-debug = False
+debug = True
 
 
 
@@ -579,7 +581,7 @@ def get_sim_adv_data(asw):
 def draw_sim(asw, sim):
   
     path = os.path.join(simdir, asw)
-#    try:
+#    try: 
 #      path = os.path.join(simdir, asw)
 #      #os.makedirs(path)
 #    except OSError as e:
@@ -589,6 +591,10 @@ def draw_sim(asw, sim):
     filenames = ['arriv', 'kappa', 'm_encl']
     
     print '> drawing sim %s'%asw,    
+    
+    path = os.path.join(simdir, asw)
+    if not os.path.isdir(path):
+      os.makedirs(path)  
     
     #prevent submodules prints..
     old_stdout = sys.stdout
@@ -610,18 +616,30 @@ def draw_sim(asw, sim):
     kappa,arriv = many.grids(asw,x,y)
 
 
-    fig = pl.figure()
+    #
+    # KAPPA - MASS MAP
+    #
+    fig = pl.figure(figsize=figsize)
     panel = fig.add_subplot(1,1,1)
     panel.set_aspect('equal')
-    lev = np.linspace(0,10,41)
-    pc = panel.contour(x,y,kappa,lev)
-    panel.clabel(pc, inline=1, fontsize=10)
+    
+    #lev = np.linspace(0,10,41)
+    #pc = panel.contour(x,y,kappa,lev, colors=0.8)
+    
+    pc = panel.contour(x+.25,y+.25,np.log(kappa),colors=0.9)
+    
+    #panel.clabel(pc, inline=1, fontsize=10)
+    panel.pcolormesh(x,y,np.sqrt(kappa), edgecolors="None", cmap='bone', shading="flat")
+    
     #pl.savefig(os.path.join(path, '%s.%s'%(filenames[1],ext)))
     pl.savefig(os.path.join(path + '_%s.%s'%(filenames[1],'pdf')))
[…]

### 2014-05-19 plots/gen_plots.py
finetune plots

@@ -60,12 +60,56 @@ import os
 import shutil
 
 import requests as rq
+import csv
 
 import numpy as np
+from numpy import pi
 import matplotlib as mpl
 
-mpl.rc('text', usetex=True)
-mpl.rc('font', family='serif')
+
+debug = True
+# set to false to do a dry run in /plots/[outdir]
+write_to_tex_folder = not debug
+
+
+#image extension (png or pdf)
+exts = ['png', 'pdf']
+
+figsize=(10,10) 
+
+# realtive to plots dir, incase of debug
+outdir = 'figs_new2'
+
+#select sims to produce plots for (using sel_sim_plots())
+sel_sim = [
+    'ASW000102p',
+    'ASW000195x',
+    'ASW0000vqg',
+    'ASW0004oux',
+    'ASW0001hpf',
+    'ASW0000h2m',
+    'ASW0002z6f',
+]
+
+#select models to produce pots for (using sel_mod_plots())
+sel_mod = [
+    '006915',
+    '006919',
+    '006937',
+    '006941',
+    '006975',
+    '006990',
+    '007020',
+    '007022',
+    '007024',
+    '007025',
+]
+
+#enable ER plots to produce
+ERplots = [True, True, True, True, True]
+if debug: ERplots = [0, True, 0, 0, True]
+
+
 
 params = {
 # see http://matplotlib.org/users/customizing.html
@@ -86,49 +130,46 @@ params = {
   }
 mpl.rcParams.update(params)
 
+mpl.rc('text', usetex=True)
+mpl.rc('font', family='serif')
+
+#############################################################
+# END SETTINGS
+#############################################################
+
+
+
+
 import matplotlib.pylab as pl
+import gen_kappa_encl_plots as spg
+import many
+# import after, to be sure mpl.rc changes are valid
 
-from numpy import pi
-
-# set to false to do a dry run in /plots/res
-write_to_tex_folder = True
-#image extension (png or pdf)
-#ext = 'png'
-ext = 'pdf'
-
-figsize=(10,10)
-
-# realtive to plots dir
-outdir = 'figs_new2'
-
+# subdirs, this should be fixed..
 simdir = 'sim'
 moddir = 'mod'
-
 resdir = 'res'
-# relative to root of git repro
-texdir = 'text/fig/sims'
[…]

### 2014-05-20 plots/gen_plots.py
crop plots

@@ -67,9 +67,9 @@ from numpy import pi
 import matplotlib as mpl
 
 
-debug = True
+debug = False
 # set to false to do a dry run in /plots/[outdir]
-write_to_tex_folder = not debug
+write_to_tex_folder = False
 
 
 #image extension (png or pdf)
@@ -93,16 +93,16 @@ sel_sim = [
 
 #select models to produce pots for (using sel_mod_plots())
 sel_mod = [
-    '006915',
-    '006919',
-    '006937',
-    '006941',
-    '006975',
-    '006990',
-    '007020',
-    '007022',
-    '007024',
-    '007025',
+    6915,
+    6919,
+    6937,
+    6941,
+    6975,
+    6990,
+    7020,
+    7022,
+    7024,
+    7025,
 ]
 
 #enable ER plots to produce
@@ -137,13 +137,15 @@ mpl.rc('font', family='serif')
 # END SETTINGS
 #############################################################
 
-
-
+pardir = os.path.abspath(os.path.join(os.path.dirname(__file__), os.path.pardir))
+sys.path.append(os.path.join(pardir, 'spaghetti'))
+sys.path.append(os.path.join(pardir, 'systems'))
 
 import matplotlib.pylab as pl
 import gen_kappa_encl_plots as spg
 import many
 # import after, to be sure mpl.rc changes are valid
+
 
 # subdirs, this should be fixed..
 simdir = 'sim'
@@ -158,7 +160,6 @@ simdir = os.path.abspath(os.path.join(outdir, simdir))
 moddir = os.path.abspath(os.path.join(outdir, moddir))
 resdir = os.path.abspath(os.path.join(outdir, resdir))
 
-pardir = os.path.abspath(os.path.join(os.path.dirname(__file__), os.path.pardir))
 
 if write_to_tex_folder:
   resdir = os.path.abspath(os.path.join(pardir, texdir))
@@ -168,8 +169,6 @@ if write_to_tex_folder:
   
 
   
-sys.path.append(os.path.join(pardir, 'spaghetti'))
-sys.path.append(os.path.join(pardir, 'systems'))
 
 
 
@@ -183,6 +182,11 @@ def test():
     break
 
 
+def plot_sel():
+# only plot selected plots, no tex output
+  sel_sim_plots()
+  sel_mod_plots()
+  plotAllRE()
 
 def run():
   all_sim_plots()
[…]

### 2014-05-20 plots/gen_plots.py
updated images

@@ -75,7 +75,9 @@ write_to_tex_folder = False
 #image extension (png or pdf)
 exts = ['png', 'pdf']
 
-figsize=(10,10) 
+figsize   = (10,10) 
+figsizeER = (10,8)
+figsizeKE = (10,8)
 
 # realtive to plots dir, incase of debug
 outdir = 'figs_new2'
@@ -100,6 +102,7 @@ sel_mod = [
     6975,
     6990,
     7020,
+    7021,
     7022,
     7024,
     7025,
@@ -656,7 +659,7 @@ def get_mod_adv_data(mid):
     im = Image.open(os.path.join(path, iname))
     w,h = im.size
     b = (w-h)/2
-    im.crop((b, 0, w-b, h)).save(os.path.join(path, saveas2[j]))
+    im.crop((b, 0, w-b, h)).save(os.path.join(path, saveas[j]))
     print '  -',saveas2[j]
   
 
@@ -926,8 +929,8 @@ def draw_mod(mid, elem, data, sims):
     
     
   pl.ioff()
-  pl.figure(figsize=figsize)
-  #panel = fig.add_subplot(1,1,1)
+  fig = pl.figure(figsize=figsizeKE)
+  ax = fig.add_subplot(1,1,1)
   
   x  = elem['x']
   #y  = elem['y']
@@ -1026,6 +1029,7 @@ def draw_mod(mid, elem, data, sims):
     fontsize=14,
     transform = ax.transAxes)
 
+  #ax.set_yscale('log')
   
   if show:
     #print 'show'
@@ -1110,7 +1114,7 @@ def plotAllRE():
   ac = np.array(ac, dtype=bool)
   
   if ERplots[0]:
-    pl.figure(figsize=figsize)
+    pl.figure(figsize=figsizeER)
     #paint accepted
     pl.errorbar(xi[ac], re[ac], [ep[ac], em[ac]], marker='s', mfc='blue', ls='' ,ecolor='blue')
     #paint rejected
@@ -1127,7 +1131,7 @@ def plotAllRE():
     #pl.show()
   
   if ERplots[1]:  
-    pl.figure(figsize=figsize)
+    pl.figure(figsize=figsizeER)
     pl.errorbar(xi, re_rel, ee_rel, marker='s', mfc='blue', ls='' ,ecolor='blue')
     pl.plot(xi, xi*0+1, '-r')
     
@@ -1137,7 +1141,7 @@ def plotAllRE():
     #pl.show()
   
   if ERplots[2]:
[…]

### 2014-05-23 plots/gen_plots.py
temp commit

@@ -656,7 +656,7 @@ def draw_sim(asw, sim):
     panel.set_aspect('equal')
     lo,hi = np.amin(arriv), np.amax(arriv)
     lev = np.linspace(lo,lo+.2*(hi-lo),100)
-    panel.contour(x,y,arriv,lev)
+    panel.contour(x,y,arriv,lev, cmap=mpl.cm.gist_rainbow, linewidths=5)
     #pl.savefig(os.path.join(path, '%s.%s'%(filenames[0],ext)))
     pl.savefig(os.path.join(path + '_%s.%s'%(filenames[0],'pdf')))
     pl.savefig(os.path.join(path + '_%s.%s'%(filenames[0],'png')))

### 2014-05-23 surf/topos0.py

@@ -96,7 +96,7 @@ def Maskb(c):
     return a    
     
 
-spag = 2
+spag = 0
 
 ma = np.zeros((100,100), float)
 levs = [1620]
@@ -119,7 +119,7 @@ for i in range(len(h)):
             h[i,j] = side
         
 
-bgcol = '0.25'
+bgcol = '0.5'
 
 fig = plt.figure(facecolor=bgcol)
 ax = fig.add_subplot(1, 1, 1, projection='3d', axisbg=bgcol)  # 3D
@@ -145,7 +145,7 @@ while True:
         break
     levs = levs + [v]
 
-surf = plt.contour(X, Y, h, levs, cmap=cm.gist_rainbow, linewidths=5)
+surf = plt.contour(X, Y, h, levs, cmap=cm.gist_rainbow)
 
 plt.show()
 

### 2014-05-23 surf/topos1.py

@@ -96,7 +96,7 @@ def Maskb(c):
     return a    
     
 
-spag = 2
+spag = 1
 
 ma = np.zeros((100,100), float)
 levs = [1620]
@@ -119,7 +119,7 @@ for i in range(len(h)):
             h[i,j] = side
         
 
-bgcol = '0.25'
+bgcol = '0.5'
 
 fig = plt.figure(facecolor=bgcol)
 ax = fig.add_subplot(1, 1, 1, projection='3d', axisbg=bgcol)  # 3D
@@ -145,7 +145,7 @@ while True:
         break
     levs = levs + [v]
 
-surf = plt.contour(X, Y, h, levs, cmap=cm.gist_rainbow, linewidths=5)
+surf = plt.contour(X, Y, h, levs, cmap=cm.gist_rainbow)
 
 plt.show()
 

### 2014-05-28 surf/topos.py
changed topos imgs

@@ -96,7 +96,7 @@ def Maskb(c):
     return a    
     
 
-spag = 2
+spag = 1
 
 ma = np.zeros((100,100), float)
 levs = [1620]

### 2014-05-28 surf/all_topos.py
added new topos files

@@ -4,6 +4,18 @@ import matplotlib.pyplot as plt
 from mpl_toolkits.mplot3d.axes3d import get_test_data
 
 
+
+exts = ['png', 'pdf']
+fname = 'arriv_'
+savepath = ''
+savepath = '../text/fig/'
+figs = [0,0,0]
+elevs = [  70,  73,  75]
+azims = [-134,-147,-141]
+bgcol = '0.25'
+fccol = 'white'
+
+
 def Mask(c):
     
     a = 0*c
@@ -95,57 +107,62 @@ def Maskb(c):
                 
     return a    
     
+for ii in range(3):
 
-spag = 0
+    spag = ii
 
-ma = np.zeros((100,100), float)
-levs = [1620]
-dt = 55
-if spag==1:
-    ma[45: 52, 45: 52] = 25
-    levs = [1211]
-if spag==2:
-    ma[45: 50, 30: 70] = 10
+    ma = np.zeros((100,100), float)
+    levs = [1620]
+    dt = 55
+    if spag==1:
+        ma[45: 52, 45: 52] = 25
+        levs = [1211]
+    if spag==2:
+        ma[45: 50, 30: 70] = 10
 
 
-        
-h = Phermat(ma)
+            
+    h = Phermat(ma)
 
-side = h[0,50]
-for i in range(len(h)):
-    for j in range(len(h)):
-        dh = h[i,j] - side
-        if dh > 0:
-            h[i,j] = side
-        
+    side = h[0,50]
+    for i in range(len(h)):
+        for j in range(len(h)):
+            dh = h[i,j] - side
+            if dh > 0:
+                h[i,j] = side
+     
+    fig = plt.figure(facecolor=fccol)
+    figs[ii]=fig
+    ax = fig.add_subplot(1, 1, 1, projection='3d', axisbg=bgcol)  # 3D
+    ax.axis('off')
 
-bgcol = '0.5'
+    X = np.arange(100)
+    Y = np.arange(100)
+    X, Y = np.meshgrid(X, Y)
 
-fig = plt.figure(facecolor=bgcol)
-ax = fig.add_subplot(1, 1, 1, projection='3d', axisbg=bgcol)  # 3D
-ax.axis('off')
+    ax.axis('equal')
 
-X = np.arange(100)
-Y = np.arange(100)
-X, Y = np.meshgrid(X, Y)
+    lo = np.amin(h)
+    hi = np.amax(h)
 
-ax.axis('equal')
+    while True:
[…]

### 2014-05-28 plots/gen_plots.py
proper scaling

@@ -173,6 +173,34 @@ if write_to_tex_folder:
 
   
 
+#manual scaling factors
+
+scales = {}
+scales['ASW000102p'] = {}
+scales['ASW000102p'][6941] = {
+#    Map radius           = 0.1851 [arcsec] Distance to center of outer pixel.
+#    Map Extent           = 0.1929 [arcsec] Distance to outer edge of outer pixel.
+#    top_level_cell_size  = 0.0154 [arcsec]
+#    Map radius           = 1.1168 [kpc]    H0inv=13.7
+#    Map Extent           = 1.1633 [kpc]    H0inv=13.7
+    'map_r'           : {'arcsec':0.1851, 'kpc': 1.1168},
+    'map_e'           : {'arcsec':0.1929, 'kpc': 1.1633},
+}
+  
+scales['ASW000195x'] = {}
+scales['ASW000195x'][6975] = {
+    # Pixel radius         = 12
+    # Map radius           = 0.4118 [arcsec] Distance to center of outer pixel.
+    # Map Extent           = 0.4290 [arcsec] Distance to outer edge of outer pixel.
+    # top_level_cell_size  = 0.0343 [arcsec]
+    # Map radius           = 2.1836 [kpc]    H0inv=13.7
+    # Map Extent           = 2.2746 [kpc]    H0inv=13.7
+    # top_level_cell       = 0.1820 [kpc]    H0inv=13.7
+    'map_r'           : {'arcsec':0.4118, 'kpc': 2.1836},
+    'map_e'           : {'arcsec':0.4290, 'kpc': 2.2746},
+}
+  
+  
 
 
 def test():

### 2014-06-01 plots/gen_plots.py
cleanup

@@ -67,7 +67,7 @@ from numpy import pi
 import matplotlib as mpl
 
 
-debug = True
+debug = False
 # set to false to do a dry run in /plots/[outdir]
 write_to_tex_folder = False
 
@@ -817,7 +817,7 @@ def draw_sim(asw, sim):
     
     pc = panel.contour(x_cut,y_cut, 2.5*np.log10(kappa_cut+eps), n_contours, colors='0.9' )
     
-    panel.clabel(pc, inline=1, fontsize=10)
+    #panel.clabel(pc, inline=1, fontsize=10)
     panel.pcolormesh(x_cm, y_cm, np.log10(kappa_cut+eps), vmin=np.log10(eps), vmax=np.log10(np.amax(kappa_cut)), edgecolors="None", cmap='bone', shading="flat")
     
     panel.tick_params(

### 2014-06-01 plots/gen_plots.py
many2 prints arriv contour levels

@@ -67,7 +67,7 @@ from numpy import pi
 import matplotlib as mpl
 
 
-debug = False
+debug = True
 # set to false to do a dry run in /plots/[outdir]
 write_to_tex_folder = False
 
@@ -80,7 +80,7 @@ figsizeER = (10,8)
 figsizeKE = (10,8)
 
 # realtive to plots dir, incase of debug
-outdir = 'figs_new2'
+outdir = 'figs_new3'
 
 #select sims to produce plots for (using sel_sim_plots())
 sel_sim = [
@@ -206,7 +206,11 @@ scales['ASW000195x'][6975] = {
     'map_e'           : {'arcsec':0.4290, 'kpc': 2.2746},
 }
   
+
   
+levels = {
+    'ASW000195x': [-682.16]
+}
 
 
 def test():
@@ -890,7 +894,11 @@ def draw_sim(asw, sim):
     lev = np.linspace(lo,lo+.2*(hi-lo),30)
     
     mpl.rcParams['contour.negative_linestyle'] = 'solid'
-    panel.contour(x,y,arriv,lev, cmap=mpl.cm.gist_rainbow, linewidths=3)
+    #panel.contour(x,y,arriv,lev, cmap=mpl.cm.gist_rainbow, linewidths=3)
+    panel.contour(x,y,arriv,lev, colors='magenta', linewidths=2)
+    
+    lev = levels[asw]
+    panel.contour(x,y,arriv,lev, colors='black', linewidths=4)
 
     
     # hide axis
@@ -1356,7 +1364,8 @@ def getModelError():
     return err_data
     
   
-#if __name__ == "__main__":
-#  run()
+if __name__ == "__main__":
+  #sel_sim_plots()
+  pass
   
   

### 2014-06-01 systems/many2.py

@@ -37,9 +37,9 @@ def shear(xs,x,y):
 def poten_gal(gal,x,y):
     z = zeros((len(y),len(x)))
     re,ell,ell_pa = float(gal[1]), float(gal[2]), float(gal[3])
-    print 'Einstein radius',re
-    print 'ellipticity',ell
-    print 'pos ang',ell_pa
+    #print 'Einstein radius',re
+    #print 'ellipticity',ell
+    #print 'pos ang',ell_pa
     for i in range(len(x)):
         for j in range(len(y)):
             z[j,i] = poten_SIE(x[i],y[j],re,ell,ell_pa)
@@ -51,7 +51,7 @@ def poten_clus(clus,zsrc,x,y):
     rsc = float(clus[0][2])
     zlens = float(clus[1][3])
     fudge = dratio(zlens,zsrc)
-    print 'kaps,rsc',kaps,rsc
+    #print 'kaps,rsc',kaps,rsc
     for i in range(len(x)):
         for j in range(len(y)):
             z[j,i] = fudge * poten_NFW(x[i],y[j],kaps,rsc)
@@ -61,11 +61,11 @@ def poten_clus(clus,zsrc,x,y):
         zlens = float(gal[3])
         fudge = dratio(zlens,zsrc)
         sig = float(gal[4])
-        print 'sigma', sig
+        #print 'sigma', sig
         ell,pa = float(gal[5]),float(gal[6])
         reinst = 4*pi*(sig/3e5)**2 * 206265/.186 * fudge
-        print 'pos',gx,gy
-        print 'reinst,ell,pa',reinst,ell,pa
+        #print 'pos',gx,gy
+        #print 'reinst,ell,pa',reinst,ell,pa
         for i in range(len(x)):
             for j in range(len(y)):
                 z[j,i] += poten_SIE(x[i]-gx,y[j]-gy,reinst,ell,pa)
@@ -91,8 +91,8 @@ def grids(asw,x,y):
     return (kappa,basic-poten)
 
 def draw(asw):
-    print asw
-    N,R = 100,50
+    print "\n    '%s':["%asw
+    N,R = 200,50
     if sim[asw][0] == 'quasar':
         flag,R = 'Q',20
     if sim[asw][0] == 'galaxy':
@@ -102,15 +102,15 @@ def draw(asw):
     x = linspace(-R,R,N)
     y = 1*x
     kappa,arriv = grids(asw,x,y)
-    fig = figure()
-    panel = fig.add_subplot(1,1,1)
[…]

### 2014-06-03 plots/conv_inp.py
new input and convert script

@@ -1,0 +1,34 @@
+# -*- coding: utf-8 -*-
+"""
+
+"""
+
+import numpy as np
+from PIL import Image as img
+
+fact = 750./480
+
+org_dim = np.array([600,480,70,50,60,60])
+new_dim = np.int16(np.round(org_dim * fact))
+
+sel_mod = [
+    6915,
+    6919,
+    6937,
+    6941,
+    6975,
+    6990,
+    7020,
+    7021,
+    7022,
+    7024,
+    7025,
+]
+
+for i in sel_mod:
+    nimg = img.new(img.MODES[6], (new_dim[0],new_dim[0]), color=(255,255,255))
+    oimg = img.open('new_inputs\\%06i.png' % i)
+    
+    nimg.paste(oimg, (new_dim[2], new_dim[4], 750+new_dim[2],750+new_dim[4]))
+    
+    nimg.save('new_inputs\\%06i_input.png' % i)

### 2014-06-03 plots/gen_plots.py

@@ -76,7 +76,7 @@ write_to_tex_folder = False
 fetch_onlinedata = False
 
 # do which plots of sims?
-simplots = {'arriv': True, 'kappa':False, 'kappaenc':False}
+simplots = {'arriv': True, 'kappa':True, 'kappaenc':False}
 
 #image extension (png or pdf)
 #exts = ['png', 'pdf']

### 2014-06-03 text/ms.tex
2 column, example model figures one per row
updated plot style
added input plots, made mass distr and arr time plots look similar
finetune plots
tex new plot layout
edits
updated images
editing theory part
temp commit
changed topos imgs
cleanup

@@
-\documentclass[usenatbib,onecolumn]{mn2e}
+% \documentclass[usenatbib,onecolumn]{mn2e}
+\documentclass[usenatbib]{mn2e}
 
@@
 
+%more abbrevs
+\newcommand{\tgeom}{t_{\rm geom}}
+\newcommand{\tgrav}{t_{\rm grav}}
+
 % todo annotations on side in red. easy to search for at the end
@@
 % \\\ END OF REMOVAL SECION
+
+\newlength{\myplotswidth}
+\setlength{\myplotswidth}{0.4\linewidth}
 
@@
 \centering
-\includegraphics[width=.7\hsize]{fig/lenssky}
+\includegraphics[width=0.9\hsize]{fig/lenssky}
 \caption{Sky distribution of 423 published lenses considered secure as
@@
 
+\begin{equation} \label{eq:tgeom}
+\tgeom(x,y) \propto x^2 + y^2 \,.
+\end{equation}
+
+\begin{equation} \label{eq:tgrav}
+\tgrav = \langle \tgrav^{\textstyle\circ} \rangle
+         + (1+z_L) \frac{2G}{c^3} M_{\textstyle\bullet}
+\end{equation}
+
+\begin{equation} \label{eq:kappa}
+\kappa(x,y) \propto \Sigma(x,y) \,.
+\end{equation}
+
+Incidentally, $\kappa$ has a second meaning, as well as
+being a dimensionless form of the projected density: it also
+measures how a bundle of light rays is brought together by
+gravitational lensing, and is called the {\em convergence}.
+
 Consider a gravitational lens.  As in most astrophysical lensing, this
@@
 \centering
-\includegraphics[height=.3\vsize]{fig/arriv1}
-\includegraphics[height=.3\vsize]{fig/arriv2}
-\includegraphics[height=.3\vsize]{fig/arriv3}
+%\includegraphics[height=.3\vsize]{fig/arriv_0}
+%\includegraphics[height=.3\vsize]{fig/arriv_1}
+%\includegraphics[height=.3\vsize]{fig/arriv_2}
+\includegraphics[width=\columnwidth]{fig/arriv_0}
+\includegraphics[width=\columnwidth]{fig/arriv_1}
+\includegraphics[width=\columnwidth]{fig/arriv_2}
 \caption{Arrival-time surfaces, with contour levels of equal arrival
@@
 \begin{equation} \label{eq:Aarriv}
-\hbox{arrival time} = A_{\rm geom} + A_{\rm grav} \,.
+\hbox{arrival time} = t_{\rm geom} + t_{\rm grav} \,.
 \end{equation}
@@
 Poisson's equation in two dimensions), and there are techniques for
-solving them.  Incidentally, $\kappa$ has a second meaning, as well as
-being a dimensionless form of the projected density: it also
-measures how a bundle of light rays is brought together by
-gravitational lensing, and is called the {\em convergence}.
+solving them.
 
@@
   \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
+  \subfigure{\includegraphics[width=0.9\hsize]
              {fig/006941_input}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
+  \subfigure{\includegraphics[width=0.9\hsize]
              {fig/007022_input}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
+  \subfigure{\includegraphics[width=0.9\hsize]
              {fig/006919_input}}
@@
 
+% Rafi's original one-coumn design:
+%
+% \begin{figure}
+%   \centering
+%   \subfigure[real mass distribution]{
+%     \label{fig:6941_sim_mass}
+%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa}
+%   }
+%   \subfigure[real arrival-time surface]{
+%     \label{fig:6941_sim_arr}
+%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv}
+%   }
+%   \subfigure[model mass distribution]{
+%     \label{fig:6941_mass}
+%     \includegraphics[width=0.45\textwidth]{fig/006941_mass}
+%   }
+%   \subfigure[model arrival-time surface]{
+%     \label{fig:6941_cont}
+%     \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti}
+%   }
+%   \subfigure[real vs model enclosed mass]{
+%     \label{fig:6941_kappa}
+%     \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl}
+%   }
+%   \subfigure[model lensed image]{
+%     \label{fig:6941_atime}
+%     \includegraphics[width=0.45\textwidth]{fig/006941_arr_time}
+%   }
+%   \caption[result 6941 (ASW000102p)]{Model results for a simple
+%     simulated lens.  (The label ASW000102p in the lower-left panel is
+%     the \sw identifier for the sky patch where this sim was placed.)
+%     The model input is shown in the top panel of
+%     Figure~\ref{fig:input-spag}.  For details of the individual panels
+%     here, see section \ref{sec:example_models}.}
+%   \label{fig:6941}
+% \end{figure}
+
+% Rafik's two row design
 \begin{figure}
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6941_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6941_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6941_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006941_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6941_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6941_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6941_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006941_arr_time}
-  }
+  \includegraphics[width=\myplotswidth]{fig/006941_input}
+  \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000102p_arriv} 
+  \includegraphics[width=\myplotswidth]{fig/006941_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000102p_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006941_mass}
+
   \caption[result 6941 (ASW000102p)]{Model results for a simple
@@
 
+% Convert all to Phil's scheme:
 \begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6975_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6975_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6975_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006975_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6975_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006975_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6975_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006975_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6975_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006975_arr_time}
-  }
+  \includegraphics[width=\myplotswidth]{fig/006975_input}
+  \includegraphics[width=\myplotswidth]{fig/006975_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_arriv.png}
+  \includegraphics[width=\myplotswidth]{fig/006975_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_kappa.png}
+  \includegraphics[width=\myplotswidth]{fig/006975_mass}
+
   \caption[result 6975 (ASW000195x)]{A model for a sim with some mass
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6937_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6937_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6937_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006937_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6937_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006937_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6937_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006937_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6937_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006937_arr_time}
-  }
+
+  \includegraphics[width=\myplotswidth]{fig/006937_input}
+  \includegraphics[width=\myplotswidth]{fig/006937_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006937_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006937_mass}
+
   \caption[result 6937 (ASW0000vqg)]{A sim with substructure,
@@
 
+  %\includegraphics[width=\myplotswidth]{fig/00_input}
+  %\includegraphics[width=\myplotswidth]{fig/00_arr_time} \\
+  %\includegraphics[width=\myplotswidth]{fig/ASW000_arriv}
+  %\includegraphics[width=\myplotswidth]{fig/00_spaghetti} \\
+  %\includegraphics[width=\myplotswidth]{fig/ASW000_kappa}
+  %\includegraphics[width=\myplotswidth]{fig/00_mass}
+
 \begin{figure}
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6990_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6990_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6990_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006990_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6990_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006990_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6990_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006990_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6990_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006990_arr_time}
-  }
+
+  \includegraphics[width=\myplotswidth]{fig/006990_input}
+  \includegraphics[width=\myplotswidth]{fig/006990_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006990_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006990_mass}
 
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6915_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6915_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6915_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006915_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6915_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006915_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6915_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006915_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6915_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006915_arr_time}
-  }
+
+  \includegraphics[width=\myplotswidth]{fig/006915_input}
+  \includegraphics[width=\myplotswidth]{fig/006915_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006915_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006915_mass}
 
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:7022_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:7022_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:7022_mass}
-    \includegraphics[width=0.45\textwidth]{fig/007022_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:7022_cont}
-    \includegraphics[width=0.45\textwidth]{fig/007022_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:7022_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/007022_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:7022_atime}
-    \includegraphics[width=0.45\textwidth]{fig/007022_arr_time}
-  }
+
+  \includegraphics[width=\myplotswidth]{fig/007022_input}
+  \includegraphics[width=\myplotswidth]{fig/007022_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_arriv}
+  \includegraphics[width=\myplotswidth]{fig/007022_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_kappa}
+  \includegraphics[width=\myplotswidth]{fig/007022_mass}
 
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:7025_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:7025_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:7025_mass}
-    \includegraphics[width=0.45\textwidth]{fig/007025_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:7025_cont}
-    \includegraphics[width=0.45\textwidth]{fig/007025_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:7025_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/007025_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:7025_atime}
-    \includegraphics[width=0.45\textwidth]{fig/007025_arr_time}
-  }
+
+  \includegraphics[width=\myplotswidth]{fig/007025_input}
+  \includegraphics[width=\myplotswidth]{fig/007025_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_arriv}
+  \includegraphics[width=\myplotswidth]{fig/007025_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_kappa}
+  \includegraphics[width=\myplotswidth]{fig/007025_mass}
 
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6919_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6919_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6919_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006919_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6919_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006919_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6919_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006919_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6919_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006919_arr_time}
-  }
+
+  \includegraphics[width=\myplotswidth]{fig/006919_input}
+  \includegraphics[width=\myplotswidth]{fig/006919_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006919_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006919_mass}
+
   \caption[result 6919 (ASW0002z6f)]{A sim on the border case of
@@
 
-\begin{figure}[htbp]
-  \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_4}
+\begin{figure*}
+  \centering
+    \includegraphics[width=0.90\linewidth]{fig/eR_5}
   \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
   \label{fig:ER_per_sim}
-\end{figure}
+\end{figure*}
 
@@
   \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
+  \subfigure{\includegraphics[width=0.9\linewidth]
              {fig/007020_kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
+  \subfigure{\includegraphics[width=0.9\linewidth]
              {fig/007024_kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
+  \subfigure{\includegraphics[width=0.9\linewidth]
              {fig/007021_kappa_encl}} \\
@@
 
-\section{More on Lensing Theory} \label{sec:more-theory}
+\section{Relation to standard lensing formalism} \label{sec:more-theory}
+
+The presentation of the arrival-time surface in \secref{Fermat}
+omitted details, for the sake of a more intuitive explanation.
+Specifically, the geometric time delay $t_{\rm geom}$ and the
+convergence $\kappa$ were left as proportionalities (equations
+\ref{eq:tgeom} and \ref{eq:kappa}), and the gravitational time delay
+$t_{\rm grav}$ was given in an implicit form (equation
+\ref{eq:tgrav}).  We now fill in the details.
+
+The original formulation of the arrival-time surface appears in
+equations (2.1) to (2.6) of \cite{1986ApJ...310..568B}.  Their
+equations can be rearranged as follows.
+\begin{equation} \label{cfBN86}
+\begin{aligned}
+t_{\rm geom} &= \frac{(1+z_L)}{2c} \frac{d_S}{d_L d_{LS}}
+\left[ (x-s_x)^2 + (y-s_y)^2 \right] \\
+\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y) \\
+\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{d_L d_{LS}}{d_S}
+               \times \Sigma(x,y)
+\end{aligned}
+\end{equation}
+We symbols $d_L,d_S$ and $d_{LS}$ are angular-diameter distances,
+respectively from observer to lens, observer to source, and lens to
+source.  In the concordance cosmology, we have
+\begin{equation}
+d_{LS} = \frac{c}{H_0}\,\frac1{1+z_S} \,
+\int_{z_L}^{z_S} \!\! \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
+%\raise.5ex\hbox{.}
+\end{equation}
+and similarly $d_S$ and $d_{LS}$.  We have replaced angular positions
+on the sky with positions on the lens plane as
+\begin{equation}
+(x,y) = d_L (\theta_x,\theta_y) \,.
+\end{equation}
+
+The first line of equation \eqref{cfBN86} is $t_{\rm geom}$ from
+equation \eqref{eq:tgeom} with the proportionality filled in, and with
+the source offset at $(s_x,s_y)$ rather than at the origin.
+
+The last line of equation \eqref{cfBN86} fills in the proportionality
+factor in equation \eqref{eq:kappa} for the convergence (or
+dimensionless surface density) $\kappa$.
+
+The middle line of equation \eqref{cfBN86}, though very different in
+form to the expression \eqref{eq:tgrav} for the gravitational time
+delay, is in fact equivalent.  This can be shown by considering a
+multipole expansions for the solution of the latter equation.  Another
+way to see the equivalence is to write a discrete version of
+\eqref{eq:tgrav}.  Introducing a small distance $\Delta$, we write
+\begin{equation} \label{eq:relax}
+\begin{aligned}
+\tgrav(x,y) &=
+\frac14 \Big[ \tgrav(x+\Delta,y) + \tgrav(x-\Delta,y) + {}\\
+&\kern24pt    \tgrav(x,y+\Delta) + \tgrav(x,y-\Delta) \Big] \\
+&+ (1+z_L)\frac{2G}{c^3} \, \pi\Delta^2 \, \Sigma(x,y) \,.
+\end{aligned}
+\end{equation}
+The circular average $\langle \tgrav^{\textstyle\circ} \rangle$ has
+been replaced by the average over four neighbouring points, and the
+enclosed mass $M_{\textstyle\bullet}$ has been replaced by
+$\pi\Delta^2\,\Sigma(x,y)$.  Now let us read equation \eqref{eq:relax}
+as an iterative formula for $t_{\rm grav}(x,y)$.  It is then
+recognisable as a formula for solving the two-dimensional Poisson
+equation from \eqref{cfBN86} by relaxation.
+
+\newpage
+
+\bibliographystyle{mn2e}
+\bibliography{bib/ms}
+
+\end{document}
 
@@
   \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_1}
+    \includegraphics[width=0.90\linewidth]{fig/eR_1}
   \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
@@
 \end{figure}
-
-\newpage
-
-\bibliographystyle{mn2e}
-\bibliography{bib/ms}
-
-\end{document}

### 2014-06-03 plots/gen_plots.py
forgott kappa plots

@@ -891,7 +891,7 @@ def draw_sim(asw, sim):
         
         n_contours=15
         
-        pc = panel.contour(x_cut,y_cut, 2.5*np.log10(kappa_cut+eps), n_contours, colors='0.9' )
+        pc = panel.contour(x_cut,y_cut, 2.5*np.log10(kappa_cut+eps), n_contours, colors='0.9', linewidths=2)
         
         #panel.clabel(pc, inline=1, fontsize=10)
         panel.pcolormesh(x_cm, y_cm, np.log10(kappa_cut+eps), vmin=np.log10(eps), vmax=np.log10(np.amax(kappa_cut)), edgecolors="None", cmap='bone', shading="flat")

### 2014-06-04 text/ms.tex
edits

@@
 
-\subsection{Theory: Fermat's Principle} \label{sec:Fermat}
+\subsection{Fermat's Principle and ``spaghetti'' contours} \label{sec:Fermat}
 
@@
 \centering
-%\includegraphics[height=.3\vsize]{fig/arriv_0}
-%\includegraphics[height=.3\vsize]{fig/arriv_1}
-%\includegraphics[height=.3\vsize]{fig/arriv_2}
 \includegraphics[width=\columnwidth]{fig/arriv_0}
@@
 \includegraphics[width=\columnwidth]{fig/arriv_2}
-\caption{Arrival-time surfaces, with contour levels of equal arrival
-  time.  The upper surface is with no lens.  The middle surface is
-  when a circular lensing mass (offset from the source) is added; a
-  maximum, a minimum and a saddle point can be seen.  The last surface
-  is the result of an elongated lensing mass; a maximum, two minima
-  and two saddle points can be seen.
-}
+\caption{Perspective views and contour maps of example arrival time
+  surfaces.  Contours are coloured in rainbow order (red: least delay,
+  violet: highest delay).  {\em Upper panel:\/} No lens, hence showing
+  the parabolic shape of the geometrical time delay.  The image would
+  be at the bottom, coinciding with the source. {\em Middle panel:\/}
+  A circular lensing mass (offset from the source) has been added,
+  which has pushed the minimum to one side and introduced a maximum
+  and saddle point, each corresponding to an image.  The saddle-point
+  is characterized by a self-crossing ``spaghetti'' contour. {\em
+    Lower panel:\/} An elongated lensing mass has been added.  There
+  are now two minima, two saddle points, and a maximum, each
+  corresponding to an image.  Note the two self-crossing spaghetti
+  contours associated with the saddle points.}
 \label{fig:arriv}

### 2014-06-04 text/ms.tex
finishing plots and cleaning up part 1

--- 2014-06-03 text/ms.tex
@@
 \usepackage{float}      % exact placement of figures (using \begin{figure}[H])
+\usepackage[section,below]{placeins}   % gives a \FloatBarrier to force all floats to appear
 \usepackage{subfigure}  % does what it says
@@
 
+\FloatBarrier
 \subsection{\spl} \label{sec:SpaghettiLens}
@@
 
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.9\hsize]
-             {fig/006941_input}} \\
-  \subfigure{\includegraphics[width=0.9\hsize]
-             {fig/007022_input}} \\
-  \subfigure{\includegraphics[width=0.9\hsize]
-             {fig/006919_input}}
-  \caption{Examples of Spaghetti input.  The models from these appear
-    later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
-  }
-  \label{fig:input-spag}
-\end{figure}
+%\begin{figure}
+  %\centering
+  %
+  %\includegraphics[width=0.8\hsize]{fig/006941_input} \\
+  %
+  %\includegraphics[width=0.8\hsize]{fig/007022_input} \\
+  %
+  %\includegraphics[width=0.8\hsize]{fig/006919_input}
+%
+  %\caption{Examples of Spaghetti input.  The models from these appear
+    %later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
+  %}
+  %\label{fig:input-spag}
+%\end{figure}
 
@@
   \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000102p_arriv} 
+  \vskip-20pt
+  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 
   \includegraphics[width=\myplotswidth]{fig/006941_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000102p_kappa}
+  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_kappa}
   \includegraphics[width=\myplotswidth]{fig/006941_mass}
@@
 \begin{figure}
+  \centering
   \includegraphics[width=\myplotswidth]{fig/006975_input}
   \includegraphics[width=\myplotswidth]{fig/006975_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000195x_arriv.png}
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_arriv.png}
   \includegraphics[width=\myplotswidth]{fig/006975_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000195x_kappa.png}
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_kappa.png}
   \includegraphics[width=\myplotswidth]{fig/006975_mass}
@@
   \includegraphics[width=\myplotswidth]{fig/006937_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_arriv}
+  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_arriv}
   \includegraphics[width=\myplotswidth]{fig/006937_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_kappa}
+  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_kappa}
   \includegraphics[width=\myplotswidth]{fig/006937_mass}
@@
   \includegraphics[width=\myplotswidth]{fig/006990_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_arriv}
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_arriv}
   \includegraphics[width=\myplotswidth]{fig/006990_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_kappa}
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_kappa}
   \includegraphics[width=\myplotswidth]{fig/006990_mass}
@@
   \includegraphics[width=\myplotswidth]{fig/006915_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_arriv}
+  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_arriv}
   \includegraphics[width=\myplotswidth]{fig/006915_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_kappa}
+  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_kappa}
   \includegraphics[width=\myplotswidth]{fig/006915_mass}
@@
   \includegraphics[width=\myplotswidth]{fig/007022_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_arriv}
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_arriv}
   \includegraphics[width=\myplotswidth]{fig/007022_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_kappa}
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_kappa}
   \includegraphics[width=\myplotswidth]{fig/007022_mass}
@@
   \includegraphics[width=\myplotswidth]{fig/007025_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_arriv}
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_arriv}
   \includegraphics[width=\myplotswidth]{fig/007025_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_kappa}
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_kappa}
   \includegraphics[width=\myplotswidth]{fig/007025_mass}
@@
   \includegraphics[width=\myplotswidth]{fig/006919_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_arriv}
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_arriv}
   \includegraphics[width=\myplotswidth]{fig/006919_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_kappa}
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_kappa}
   \includegraphics[width=\myplotswidth]{fig/006919_mass}
@@
   \subfigure{\includegraphics[width=0.9\linewidth]
-             {fig/007020_kappa_encl}}
+             {fig/007022_kappa_encl}}
   \subfigure{\includegraphics[width=0.9\linewidth]
-             {fig/007024_kappa_encl}}
+             {fig/007025_kappa_encl}}
   \subfigure{\includegraphics[width=0.9\linewidth]
              {fig/007021_kappa_encl}} \\
-  \caption{Enclosed mass \kenc from three models of a system with incorrect image identifications.
+  \caption{
+  %Enclosed mass \kenc from three models of a system with incorrect image identifications.
 \figref{7022} shows a model (by an expert) with the correct identifications, while \figref{7025} shows a model with error E06 (parities rotated by $\pi/2$).
-Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).}
+Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).
+  }
 \label{fig:kapenc_compare_faulty}

### 2014-06-04 plots/gen_plots.py

@@ -214,6 +214,7 @@ scales['ASW000195x'][6975] = {
 }
 
 
+scales['ASW0001hpf'] = {}
 scales['ASW0001hpf'][6915] = {
     # Map radius           = 0.1213 [arcsec] Distance to center of outer pixel.
     # Map Extent           = 0.1264 [arcsec] Distance to outer edge of outer pixel.
@@ -225,7 +226,8 @@ scales['ASW0001hpf'][6915] = {
     'map_e'           : {'arcsec':0.1264, 'kpc': 0.7623},
 }
 
-scales['ASW0002y6f'][6919] = {
+scales['ASW0002z6f'] = {}
+scales['ASW0002z6f'][6919] = {
     # Map radius           = 0.3153 [arcsec] Distance to center of outer pixel.
     # Map Extent           = 0.3284 [arcsec] Distance to outer edge of outer pixel.
     # top_level_cell_size  = 0.0263 [arcsec]
@@ -235,6 +237,8 @@ scales['ASW0002y6f'][6919] = {
     'map_r'           : {'arcsec':0.3153, 'kpc':1.9018 },
     'map_e'           : {'arcsec':0.3284, 'kpc':1.9810 },
 }
+
+scales['ASW0000vqg'] = {}
 scales['ASW0000vqg'][6937] = {
     # Map radius           = 0.2163 [arcsec] Distance to center of outer pixel.
     # Map Extent           = 0.2253 [arcsec] Distance to outer edge of outer pixel.
@@ -245,6 +249,8 @@ scales['ASW0000vqg'][6937] = {
     'map_r'           : {'arcsec':0.2163, 'kpc':0.7024 },
     'map_e'           : {'arcsec':0.2253, 'kpc':0.7316 },
 }
+
+scales['ASW000102p'] = {}
 scales['ASW000102p'][6941] = {
     # Map radius           = 0.1851 [arcsec] Distance to center of outer pixel.
     # Map Extent           = 0.1929 [arcsec] Distance to outer edge of outer pixel.
@@ -255,6 +261,8 @@ scales['ASW000102p'][6941] = {
     'map_r'           : {'arcsec':0.1851, 'kpc':1.1168 },
     'map_e'           : {'arcsec':0.1929, 'kpc':1.1633 },
 }
+
+scales['ASW000195x'] = {}
 scales['ASW000195x'][6975] = {
     # Pixel radius         = 12
[…]

### 2014-06-04 systems/many.py

@@ -45,7 +45,7 @@ def poten_clus(clus,zsrc,x,y):
     rsc = float(clus[0][2])
     zlens = float(clus[1][3])
     fudge = dratio(zlens,zsrc)
-    print 'kaps,rsc',kaps,rsc
+    print 'kaps,rsc,fudge',kaps,rsc,fudge
     for i in range(len(x)):
         for j in range(len(y)):
             z[j,i] = fudge * poten_NFW(x[i],y[j],kaps,rsc)

### 2014-06-05 text/ms.tex
editing

@@
 \newcommand{\tgrav}{t_{\rm grav}}
+\newcommand{\subcirc}{{\lower.33ex\hbox{$\circ$}}}
+\newcommand{\subbullet}{{\lower.33ex\hbox{$\bullet$}}}
 
@@
 \begin{equation} \label{eq:tgeom}
-\tgeom(x,y) \propto x^2 + y^2 \,.
-\end{equation}
-
+\tgeom(x,y) \propto (x-x_s)^2 + (y-s_y)^2 \,.
+\end{equation}
+
+The gravitational time delay can be expressed in an implicit form: the
+value of $\tgrav$ through $(x,y)$ equals its average value on the
+circumference of a small circle centred at $(x,y)$, plus a constant
+times the mass within that circle.  The proportionality constant is
+$2G/c^3$ times the cosmological expansion factor $(1+z_L)$. Thus
 \begin{equation} \label{eq:tgrav}
-\tgrav = \langle \tgrav^{\textstyle\circ} \rangle
-         + (1+z_L) \frac{2G}{c^3} M_{\textstyle\bullet}
-\end{equation}
+\tgrav(x,y) = \left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle
+              + (1+z_L) \frac{2G}{c^3} M(x_\subbullet,y_\subbullet) \,.
+\end{equation}
+We have used $(x_\subcirc,y_\subcirc)$ to denote the circumference of
+a circle, and $(x_\subbullet,y_\subbullet)$ to indicate the integrated
+mass within the circle.
 

### 2014-06-05 text/ms.tex
partA partB partC

--- 2014-06-04 text/ms.tex
@@
 
-\begin{abstract}
-In the initial phase of \sw (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses.
-Here we report on a web application that gives experienced volunteers the opportunity to model the candidates that have been identified.
-In order to gauge the quality of the models that were being rendered, these same volunteers were invited to model a sample of 29 simulated lenses.
-The models were then examined in greater detail, with particular attention being paid to the following:
-(i)~identification of image parities and time arrivals;
-(ii)~the mean convergence (equivalent to the enclosed mass), and finally;
-(iii)~the performance of a volunteer vs. a professional.
-In most cases, the volunteers were found to correctly identify the image parities and time arrivals.
-% Einstein radius +~20%
- along with a mean convergence that was well constrained within the image region;
-In all, the results could be comparable to that of a professional.
-\end{abstract}
+\input partA
 
-\begin{keywords}
-\end{keywords}
+\input partB
 
-\section{Introduction}
-
-Galaxies and clusters of galaxies are considered to be the largest structures in the known Universe.
-When one questions their formation,  the present picture suggests that gravitational instabilities  (observable as fluctuations of the microwave background) %TODO \todo[red]{not too sure of the relevance of this statement}
- - caused a collapse into gravitationally bound structures.
-%TODO\todo[yellow]{clarification needed: are they becoming gravitationally bound structures, or are they collapsing into objects that are already gravitationally bound?
-I know it is most probably the later, but it is a bit ambiguous}
-The initial collapse would have been dominated by dark matter, but once deep wells had appeared in the gravitational potential, gas would have started falling into those wells. 
-%TODO \todo[red]{not too sure of this wording}
-Through atomic processes,
-%TODO \todo[yellow]{perhaps we need to clarify those processes or are they generally know?}
- the gravitational potential energy of this gas would of been radiated away, leading to the formation of the largest stars. 
-%TODO \todo[red]{or perhaps just stars?}
-In turn, galaxies and clusters of galaxies would form, all the time merging and changing their form.
-This rough outline is now well established, but in order to satisfy our understanding of this process, there are many puzzles  that remain to be solved.
-For progress to be made, we need to extend our knowledge of dark matter.
-Gravitational lensing has proven as a particularly attractive way of studying dark matter as it is only dependant on mass.
-%TODO \todo[yellow]{silly question, but shouldn't we briefly explain why it is only dependant on mass? or should everyone already reading this paper know that? or is it explained later?}
-
-There are many manifestations of gravitational lensing in astrophysics,
-but the most spectacular are multiple images.  Gravitational lensing
-depends on the sky-projection mass density, and this fact has a
-counter-intuitive consequence: nearby galaxies do not produce multiple
-images, because their projected density is not high enough, but beyond
-100~Mpc or so, most galaxies are potential lenses.  Even so, for a
-potential lens to really act as a lens, a background source in the close-by
-region is required.  So, in practice galaxies are very rarely observed
-as lens.  The same tendency applies to clusters of galaxies, but since
-clusters are bigger, they are more likely lens-like.
-
-In lensing galaxies, the observable features usually involve a single
-object multiply imaged.  It is therefore desirable to have large
-numbers of lenses, and this is the strategy researchers are following.  The
-largest single study so far uses 58 lenses \citep{2009ApJ...703L..51K}.
-Other studies have compared lens models in more detail with stellar
-mass \citep{2011ApJ...740...97L} or used time-delay information from
-lensed quasars to infer the Hubble time
-\citep{2008ApJ...679...17C,2010ApJ...712.1378P}.
-
-The trend is clear: researchers want to find many more lenses and
-model them in more sophisticated ways.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.9\hsize]{fig/lenssky}
-\caption{Sky distribution of 423 published lenses considered secure as
-  of January 2014 (from the Masterlens catalog at the University of
-  Utah, maintained by Joel Brownstein and Leonidas Moustakas.) The map
-  is in Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
-  middle.  The empty swathes to the left and right of center are the
-  Milky Way.}
-\label{fig:masterlens}
-\end{figure}
-
-A good summary of the observational situation is offered by
-Figure~\ref{fig:masterlens}, which shows known secure multiply-imaging
-lenses.  The non-uniformity on the sky is not intrinsic, it just
-indicates the density of deep surveys up to 2013.  At HST resolution,
-of order 1~square degree of sky must be searched to find a lens.
-Older ground-based surveys yield a lens per roughly 10~square degree
-of survey area.  By the large surveys starting now
-(DES\footnote{\tt http://pan-starrs.ifa.hawaii.edu} or
-PanStarrs\footnote{\tt http://www.darkenergysurvey.org}) the expected
-yield is something in between.  Given their unprecedented survey area,
-these are likely to yield thousands of lenses.  Over the 2020s,
-LSST\footnote{\tt http://www.lsst.org/} can be expected to give us ten
-thousand lenses.
-
-The question now arises: can lenses be found and modeled
-automatically in large surveys?  Work using software robots
-\citep{2009ApJ...694..924M} gives some interesting and unexpected
-results.  In clean lensing system in uncrowded fields with high
-signal-to-noise, the robots do very well.  In most situations,
-however, robots miss lenses (low completeness) or contaminate the
-results with non-lenses (low purity).  Robots can be made to
-prioritize completeness or purity, but they cannot deliver both.
-
-In response to the lessons learned from automated lens searches, the
-{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org}
- has been launched.  {\em Spacewarps\/} is
-part of the {\em Zooniverse\/} family of citizen-science
-projects,\footnote{\tt http://www.zooniverse.org} where members of the
-public are invited to analyze different kinds of scientific data which
-are too difficult for robots and too large for specialists.  In {\em
- Spacewarps\/} itself, survey sky images are presented for volunteers to identify lens candidates.  The first tranche of survey data, introduced from May 2013, is an area of $\sim172$ square
-degrees (or a quarter of a percent of the sky) from the Canada-France-Hawaii Telescope Legacy Survey.
- The survey is divided into patches of $440\times 440$ pixels, each of
-which is seen by ten volunteers.  Simulated lenses are mixed in with
-the data, both to help train volunteers on what to look for, and to
-estimate completeness and purity.
-These data had previously been analyzed by robots \cite{Gavazzi2012, More2012ApJ}.  
-  A list of candidates is being
-processed, but intermediate results show some very good candidates, as
-well as re-discoveries of candidates previously found robotically.
-
-These encouraging results now raise the question: could modeling of
-the lenses also be done by volunteers?  Modeling is a much more
-difficult task than searching for lens candidates, as it requires some
-expert knowledge.  Nonetheless, a panel of the {\em Spacewarps\/}
-volunteers is quite experienced from earlier projects, having spent a
-thousand hours or more with data.  Some of these experienced
-volunteers are very interested in more demanding projects.  This also
-is a general trend \citep[cf.][]{Khatib22112011}.  The present project
-then suggests itself.
-
-The purpose of this study was to provide a means to model a large number of gravitational lenses by showing that gravitational lens modeling can be learned / done by volunteers.
-We suggest, that volunteers will be as successful as professionals with modeling if provided with an easy to use tool with visual feedback (What you see is what you get, WYSIWYG) and a minimal set of instructions.
-Volunteers will then crowd work (using buzz word here ;) ) / work collaborative on modeling lenses from several sources / groups at a central place.
-Since this is an iterative learning process, the more involved volunteers will quickly gain knowledge that can be passed down to new volunteers.
-That creates a social structure that scales well with the number of volunteers, as other projects have already shown. %TODO \needcite
-Finding people working as volunteers has been shown to be successful last but not least by \sw and the whole galaxy zoo project.
-To test the people's abilities, we investigated the performance of a few volunteers modeling a set of simulated lenses.
-We tested the ability to correctly identify lensed images and reproduce similar mass map of the lens.
-
-\section{A lens modeling program}
-
-It is well known that making a scientific contribution is a central
-motivation cited by citizen-science volunteers.  So for a lens
-modeler in a citizen-science project, nobody wants that it should
-sacrifice scientific usefulness in order to be fun to use.  That said,
-being aesthetically pleasing and having a short initial learning curve
-are also essential qualities.  Also desirable is that the user is
-encouraged/challenged to go deeper and wider into the subject.
-Yet another aspect is enabling small incremental contributions
-by different people.  \spl tries to address all of these.
-
-\subsection{Theory: Fermat's Principle} \label{sec:Fermat}
-
-There are several ways to understand the formation of arcs and
-multiple images in gravitational lensing.  We will follow some ideas
-originally introduced by \cite{1986ApJ...310..568B}, based on Fermat's
-principle.  The key to this approach is an abstract construct called
-the {\em arrival-time surface.}  This surface cannot itself be
-observed, but several observable quantities can be derived from it.
-
-\begin{equation} \label{eq:tgeom}
-\tgeom(x,y) \propto x^2 + y^2 \,.
-\end{equation}
-
-\begin{equation} \label{eq:tgrav}
-\tgrav = \langle \tgrav^{\textstyle\circ} \rangle
-         + (1+z_L) \frac{2G}{c^3} M_{\textstyle\bullet}
-\end{equation}
-
-\begin{equation} \label{eq:kappa}
-\kappa(x,y) \propto \Sigma(x,y) \,.
-\end{equation}
-
-Incidentally, $\kappa$ has a second meaning, as well as
-being a dimensionless form of the projected density: it also
-measures how a bundle of light rays is brought together by
-gravitational lensing, and is called the {\em convergence}.
-
-Consider a gravitational lens.  As in most astrophysical lensing, this
-lens is `thin' along the line of sight, and effectively lies on a
-plane.  Let $x,y$ be coordinates on this plane.  These coordinates
-measure ordinary physical lengths (meters, etc) on the lens plane.
-Let there be a point source of light at $x=0,y=0$ but far behind the
-lens plane.  Now imagine a light ray coming from the source to the
-lens plane $(x,y)$ and then changing its path to reach the observer.
-The observer would see the source apparently behind $(x,y)$ the
-lens.  If this happened, the light ray would have taken a longer path
-than having come directly to the observer.  The geometrical path
-difference would be $\propto x^2 + y^2$ for small angles.  Let us
-define
-\begin{equation} \label{eq:Ageom}
-A_{\rm geom}(x,y) = \half(x^2 + y^2) \,.
-\end{equation}
-As written, this looks an area.  But we can also think of it as a
-time delay introduced by the light ray bending --- with an unwritten
-constant factor.  That factor is basically the lens distance divided by the
-speed of light; the precise expression has to take the expansion of
-the Universe into account, and is given in the Appendix.  The top part
-of \figref{arriv} shows what $A_{\rm geom}$ looks like: it is
-just a paraboloid.  The minimum of the surface is at the source, and
-that is where, according to Fermat's principle, the image will be.
-The rest of the arrival-time surface is just an imaginary surface.
-
-\begin{figure}
-\centering
-%\includegraphics[height=.3\vsize]{fig/arriv_0}
-%\includegraphics[height=.3\vsize]{fig/arriv_1}
-%\includegraphics[height=.3\vsize]{fig/arriv_2}
-\includegraphics[width=\columnwidth]{fig/arriv_0}
-\includegraphics[width=\columnwidth]{fig/arriv_1}
-\includegraphics[width=\columnwidth]{fig/arriv_2}
-\caption{Arrival-time surfaces, with contour levels of equal arrival
-  time.  The upper surface is with no lens.  The middle surface is
-  when a circular lensing mass (offset from the source) is added; a
-  maximum, a minimum and a saddle point can be seen.  The last surface
-  is the result of an elongated lensing mass; a maximum, two minima
-  and two saddle points can be seen.
-}
-\label{fig:arriv}
-\end{figure}
-
-The geometrical time delay \eqref{eq:Ageom} is not the only one.  The
-warping of space by a gravitational field introduces a further time
-delay $A_{\rm grav}$.  The arrival time of the light ray, compared to
-what it would have been with no lensing, is
-\begin{equation} \label{eq:Aarriv}
-\hbox{arrival time} = t_{\rm geom} + t_{\rm grav} \,.
-\end{equation}
-We can also express $A_{\rm grav}$ as an area, with the same implicit
-constant factor that makes it a time.  The expression, however, is
-more complicated.  We first introduce the notation $\nabla^2 f(x,y)$
-to denote
-\begin{equation}
- \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
-        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
-      {\Delta x \; \Delta y}
-\end{equation}
-in the limit of $\Delta x,\Delta y$ small.  Then we write the
-projected mass density of the lens, which would be in $\rm
-kg\,m^{-2}$, in special units (given in the Appendix) such that the
-surface density becomes a dimensionless field $\kappa(x,y)$.  Then
-\begin{equation} \label{eq:Poisson}
-\nabla^2 A_{\rm grav}(x,y) = -2\kappa(x,y) \,.
-\end{equation}
-Note that $A_{\rm grav}(x,y)$ is not given explicitly in terms of
-$\kappa(x,y)$, but as a differential equation that must be solved.
-Equations of the type \eqref{eq:Poisson} are, however, well known (as
-Poisson's equation in two dimensions), and there are techniques for
-solving them.
-
-What is the effect of mass on the arrival-time surface?  If the mass
-distribution $\kappa(x,y)$ is circular, $A_{\rm grav}(x,y)$ will have
-a peak at the center of that circle.  But if the mass is not precisely
-in front of the source, the arrival-time surface will not be circular
-any more.  The result is illustrated in the middle part of
-\figref{arriv}.  In addition to the maximum, there is now a
-minimum and a saddle point.  According to Fermat's principle, images
-appear at maximum and saddle points as well as minimum.  If the mass
-distribution is non-circular, more images can appear.  The lower
-example in \figref{arriv} shows the arrival-time surface due
-to an elongated mass.  There is still a maximum, part of an elongated
-hill, and around it there are two minima and two saddle points.
-
-So far we have considered a single point source.  What if there is an
-extended source?  To understand this, let us consider what happens if
-we move the original point source slightly, or equivalently, keep the
-original point source behind $x=0,y=0$ and move the lens slightly.
-The contours of constant arrival time will naturally move slightly,
-and so will the images.  The movement of the contours will be most
-noticeable where the contours are far apart, that is where the
-arrival-time surface is nearly flat.  As is evident from
-\figref{arriv}, this is the region where the minimum and saddle
-points lie, or near the images.  In this region, points on the source
-that are close together produce images that are comparatively far
-apart.  In other words, the image is highly magnified.  In summary,
-lower curvature in the arrival-time surface for a point source implies
-larger magnification of an extended source.  Conversely, where the
-arrival-time surface is strongly curved, the image will be
-demagnified.  We see from \figref{arriv} that the arrival-time
-surface tends to be highly curved near the maximum.  Hence maximum tend
-to be demagnified.  In practice, maxima of the arrival time are nearly
-always too faint to see. The minima and saddle points dominate.
-
-In the lensed examples in \figref{arriv}, the maximum in each
-case is evident, but how can we distinguish the minima from the saddle
-points?  The key is to examine the contours of equal arrival time.
-Contour curves loop around a minimum or a maximum, but near a saddle
-point, the contours do not loop around, instead they approach and then
-pull away.  The contour curve precisely level with a saddle point
-forms an X at the saddle point.  In \figref{arriv}, one such X
-is evident.  In other cases the X contour passing through the saddle
-point is not itself plotted, but we can tell by the other contours where
-the X would be.  These self-crossing contours will play a special role
-in the following.
-
-\FloatBarrier
-\subsection{\spl} \label{sec:SpaghettiLens}
-
-\spl implements the ideas of the arrival-time surface into a modeling
-program.
-
-%\begin{figure}
-  %\centering
-  %
-  %\includegraphics[width=0.8\hsize]{fig/006941_input} \\
-  %
-  %\includegraphics[width=0.8\hsize]{fig/007022_input} \\
-  %
-  %\includegraphics[width=0.8\hsize]{fig/006919_input}
-%
-  %\caption{Examples of Spaghetti input.  The models from these appear
-    %later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
-  %}
-  %\label{fig:input-spag}
-%\end{figure}
-
-The first modeling step with \spl is for the user to call up the \sw
-of interest, and make an educated guess for the topography of the
-arrival-time surface --- specifically, the locations of the maxima,
-minima and saddle points that correspond to images of the brightest
-part of the source.  Input is given by tracing a proposal for the
-saddle-point contours.  \Figref{input-spag} shows some example
-inputs --- and incidentally also indicates the origin of the name
-\spl.
-
-\spl, as implemented so far, assumes that the lens is dominated by a
-single galaxy, and that the center of that galaxy is the sole maximum.
-Once the maximum has been identified, we wish to characterize the rest
-of the arrival-time surface in relation to it.  From geometry, any
-surface that has a central maximum and goes high at the edges must
-have a lima\c con-like contour, meaning a two-looped curve with one
-loop inside the other.  The maximum lies within the inner loop, a
-minimum lies between the two loops, and a saddle point lies at the
-self-intersection of the contour.  The top panel of
-\figref{input-spag} shows an example: red marks a suggested
-maximum, green a saddle point, and blue a minimum.  (The small pink
-dots are just help sketch the proposed contour.)  The middle and lower
-panels of \figref{input-spag} show a more complex situation.
-The minimum has split into three images: two new minima with a saddle
-point in between.  In this situation the contour through the new
-saddle point forms a figure of eight, its two loops enclosing the two
-new minima.  The configuration of a lima\c con, possibly with a figure
-of eight inside, suffices to give a preliminary description of the
-arrival-time surface for nearly all cases where the lens is dominated
-by one galaxy, and the source is a single small object.
-
-The exact placement of the pink spaghetti loops shown in
-\Figref{input-spag} has no significance.  Their utility is
-just to help mark up the image as maximum, minima and saddle points.
-As the figure suggests, the marking up can be done very easily with a
-mouse.  Human intuition is required to (a)~identify lensed images and
-separate them from other background light sources, and (b)~classify
-and order images according to arrival time.  But the computer helps by
-allowing only valid lensing configurations to be entered, and ensuring
-that the odd image theorem is taken care of.%TODO \needcite
-
-In the second modeling step, once the user has sketched a proposed
-spaghetti configuration, \spl sends the input to its server-side
-modeling engine, called GLASS \citep{2014arXiv1401.7990C}.  The task
-of GLASS is to find a mass distribution $\kappa(x,y)$ that exactly
-reproduces the locations of the maximum, minima and saddle
-points. Now, this criterion by itself is extremely under-determined
---- there are infinitely many mass distributions that will reproduce a
-given set of maxima, minima and saddle points, but typically they
-(a)~produce lots of extra images, and (b)~look very unlike galaxies.
-Additional assumptions (a prior) are necessary.  GLASS uses a prior
-based on suggestions by \cite{1997MNRAS.292..148S}.
-\begin{enumerate}
-\item The mass distribution is built out of non-negative tiles of
-  mass.  (Sometimes these tiles are called mass pixels, but we should
-  emphasize that they are unrelated to image pixels, and are much
-  larger.)
-\item There is a notional lens center, say $(x_0,y_0)$ which is
-  identified with the maximum of the arrival time.  The source can
-  have an arbitrary offset with respect to the lens center.
-\item The mass distribution must be centrally concentrated, in two
-  respects.  First, the circularly averaged density must fall away
-  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
-  Second, the direction of increasing density at any $(x,y)$ can point
-  at most $45^\circ$ away from $(x_0,y_0)$.
-\item The lens must be symmetrical with respect to $180^\circ$ rotations
-  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
-  user wishes.
-\end{enumerate}
-There are still infinitely many models that satisfy both data and
-prior constraints, but now they are more credible as galaxy lenses.
-It is then possible to generate an ensemble of models.  The sampling
-technique used by GLASS is described in \citep{Lubini2012}, and
-improves upon earlier techniques \citep{2000AJ....119..439W,Saha2004}.
-Typically, ensembles of 200 models are used.  That is to say, what we
-call a \spl model is really the mean of an ensemble of 200 models, and
-its estimated uncertainty is the range covered by the whole ensemble.
-
-In the third modeling step, having generated a model ensemble, \spl
-post-processes it to present results and diagnostics to the user for
-inspection. This takes the form of three figures.
-\begin{enumerate}
-\item A gray scale plus contour map of the mass distribution.
-\item A contour map of the arrival-time surface.
-\item A synthetic image of the lensed features.
-\end{enumerate}
-After examining this feedback, the user can archive the results for
-discussion, or modify their input and try again, or discard the
-attempt altogether.
-
-The fourth modeling step is discussion among modelers and iteration
-on the model.  Any archived model can be revised by another user: they
-can modify the spaghetti configuration slightly or drastically, or
-change options like the size of the mass tiles.  Particularly
-interesting lens candidates lead to trees of models in this way.
-Forum discussion then prunes the tree, focusing attention on one or a
-few models.
-
-\section{A lens modeling challenge} \label{sec:mod_challenge}
-
-Interested volunteers from the \sw forum were initially introduced to
-\spl through a video tutorial and by videocon.  After this
-introductory stage, a modeling challenge was presented.  This
-consisted of 29 simulated lenses (sims) covering a range of lensing
-configurations.
-
-The \sw sims were generated by AM, in consultation with PM and AV.
-To estimate the performance of the volunteers and the quality of the generated models, two analysis  were done.
-The first analysis tested the correct identification and ordering of lensed images.
-The second one compared the mass distribution of the lens $\kappa(x, y)$ of the generated models to the mass distribution of the simulations.
-
-\subsection{The simulated lenses} \label{sec:sims}
-
-For the sake of blind testing, the information in this section was
-revealed neither to RK, while choosing the challenge set of 29 sims,
-nor to the modelers (volunteers EB, CC, CM, JO, JW and `expert
-modeler' PS) until the modeling stage was done.
-
-The sims were produced using {\tt gravlens}
-\citep{2001astro.ph..2341K,2001astro.ph..2340K} using the CFHTLS survey data
-and catalogues \cite{Coupon2009}. They were of three
-kinds, as follows.
-
-\begin{enumerate}
-  \item Imitating lensed quasars: having a singular elliptical
-    isothermal lens (SIE) plus constant external shear, and a circular
-    Gaussian source.
-  \item Emulating lensed galaxies: similar to the above, but with an
-    elliptical de Vaucouleurs source.
-  \item Resembling cluster lenses: having a source as above, but a
-    more complicated lens, with one dominant elliptical SIE and
-    one or more perturbing elliptical SIEs, plus a circular NFW
-    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
-    the underlying dark matter distribution.
-\end{enumerate}
-
-Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
-lenses follow equations (33--35) of that work, with core radius set to
-zero.  The NFW lens is in equations (48) and (50), while shear is the
-$\gamma$ term in equation (76).
-
-\subsection{Some example models} \label{sec:example_models}
-
-The modelers proposed a total of 129 models for the 29 sims in the
-challenge.  Figures \ref{fig:6941} to \ref{fig:6919} show eight of
-models in some detail.
-
-% Rafi's original one-coumn design:
-%
-% \begin{figure}
-%   \centering
-%   \subfigure[real mass distribution]{
-%     \label{fig:6941_sim_mass}
-%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa}
-%   }
-%   \subfigure[real arrival-time surface]{
-%     \label{fig:6941_sim_arr}
-%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv}
-%   }
-%   \subfigure[model mass distribution]{
-%     \label{fig:6941_mass}
-%     \includegraphics[width=0.45\textwidth]{fig/006941_mass}
-%   }
-%   \subfigure[model arrival-time surface]{
-%     \label{fig:6941_cont}
-%     \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti}
-%   }
-%   \subfigure[real vs model enclosed mass]{
-%     \label{fig:6941_kappa}
-%     \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl}
-%   }
-%   \subfigure[model lensed image]{
-%     \label{fig:6941_atime}
-%     \includegraphics[width=0.45\textwidth]{fig/006941_arr_time}
-%   }
-%   \caption[result 6941 (ASW000102p)]{Model results for a simple
-%     simulated lens.  (The label ASW000102p in the lower-left panel is
-%     the \sw identifier for the sky patch where this sim was placed.)
-%     The model input is shown in the top panel of
-%     Figure~\ref{fig:input-spag}.  For details of the individual panels
-%     here, see section \ref{sec:example_models}.}
-%   \label{fig:6941}
-% \end{figure}
-
-% Rafik's two row design
-\begin{figure}
-  \centering
-  \includegraphics[width=\myplotswidth]{fig/006941_input}
-  \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
-  \vskip-20pt
-  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 
-  \includegraphics[width=\myplotswidth]{fig/006941_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_kappa}
-  \includegraphics[width=\myplotswidth]{fig/006941_mass}
-
-  \caption[result 6941 (ASW000102p)]{Model results for a simple
-    simulated lens.  (The label ASW000102p in the lower-left panel is
-    the \sw identifier for the sky patch where this sim was placed.)
-    The model input is shown in the top panel of
-    Figure~\ref{fig:input-spag}.  For details of the individual panels
-    here, see section \ref{sec:example_models}.}
-  \label{fig:6941}
-\end{figure}
-
-% Convert all to Phil's scheme:
-\begin{figure}
-  \centering
-  \includegraphics[width=\myplotswidth]{fig/006975_input}
-  \includegraphics[width=\myplotswidth]{fig/006975_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_arriv.png}
-  \includegraphics[width=\myplotswidth]{fig/006975_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_kappa.png}
-  \includegraphics[width=\myplotswidth]{fig/006975_mass}
-
-  \caption[result 6975 (ASW000195x)]{A model for a sim with some mass
-    substructure. (See Section \ref{sec:example_models} for details.)}
-  \label{fig:6975}
-\end{figure}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\myplotswidth]{fig/006937_input}
-  \includegraphics[width=\myplotswidth]{fig/006937_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_arriv}
-  \includegraphics[width=\myplotswidth]{fig/006937_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_kappa}
-  \includegraphics[width=\myplotswidth]{fig/006937_mass}
-
-  \caption[result 6937 (ASW0000vqg)]{A sim with substructure,
-    resulting in a poor mass model. (See Section
-    \ref{sec:example_models} for details.)}
-  \label{fig:6937}
-\end{figure}
-
-  %\includegraphics[width=\myplotswidth]{fig/00_input}
-  %\includegraphics[width=\myplotswidth]{fig/00_arr_time} \\
-  %\includegraphics[width=\myplotswidth]{fig/ASW000_arriv}
-  %\includegraphics[width=\myplotswidth]{fig/00_spaghetti} \\
-  %\includegraphics[width=\myplotswidth]{fig/ASW000_kappa}
-  %\includegraphics[width=\myplotswidth]{fig/00_mass}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\myplotswidth]{fig/006990_input}
-  \includegraphics[width=\myplotswidth]{fig/006990_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_arriv}
-  \includegraphics[width=\myplotswidth]{fig/006990_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_kappa}
-  \includegraphics[width=\myplotswidth]{fig/006990_mass}
-
-  \caption[result 6990 (ASW0004oux)]{Model of a four-image
-    system. (See Section \ref{sec:example_models} for details.)}
-  \label{fig:6990}
-\end{figure}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\myplotswidth]{fig/006915_input}
-  \includegraphics[width=\myplotswidth]{fig/006915_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_arriv}
-  \includegraphics[width=\myplotswidth]{fig/006915_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_kappa}
-  \includegraphics[width=\myplotswidth]{fig/006915_mass}
-
-  \caption[result 6915 (ASW0001hpf)]{Model of another four-image
-    system. (See Section \ref{sec:example_models} for details.)}
-  \label{fig:6915}
-\end{figure}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\myplotswidth]{fig/007022_input}
-  \includegraphics[width=\myplotswidth]{fig/007022_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_arriv}
-  \includegraphics[width=\myplotswidth]{fig/007022_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_kappa}
-  \includegraphics[width=\myplotswidth]{fig/007022_mass}
-
-  \caption[result 7022 (ASW0000h2m)]{A further four-image system. The
-    model input appears in the middle panel of Figure
-    \ref{fig:input-spag}. (See Section \ref{sec:example_models} for
-    details.)}
-  \label{fig:7022}
-\end{figure}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\myplotswidth]{fig/007025_input}
-  \includegraphics[width=\myplotswidth]{fig/007025_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_arriv}
-  \includegraphics[width=\myplotswidth]{fig/007025_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_kappa}
-  \includegraphics[width=\myplotswidth]{fig/007025_mass}
-
-  \caption[result 7025 (ASW0000h2m)]{Another model for the lens
-    appearing in \figref{7022}.  The image parities in this case
-    were incorrectly identified, resulting in a rather poor model, but
-    the estimated Einstein radius is not bad. (See Section
-    \ref{sec:example_models} for details.)}
-  \label{fig:7025}
-\end{figure}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\myplotswidth]{fig/006919_input}
-  \includegraphics[width=\myplotswidth]{fig/006919_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_arriv}
-  \includegraphics[width=\myplotswidth]{fig/006919_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_kappa}
-  \includegraphics[width=\myplotswidth]{fig/006919_mass}
-
-  \caption[result 6919 (ASW0002z6f)]{A sim on the border case of
-    double vs quad; modelled as a quad.  The model input appears in
-    the bottom panel of Figure \ref{fig:input-spag}. (See Section
-    \ref{sec:example_models} for details.)}
-  \label{fig:6919}
-\end{figure}
-
-Figure \ref{fig:6941} shows a simple example.  The top two panels show
-contours of convergence $\kappa(x,y)$, and of the arrival-time for a
-point source at the brightest point of the sim
-source.\footnote{Although the arrival-time contours represent an
-  abstract quantity that is not directly observable, the contour map
-  may actually resemble the appearance of the lensed arcs.  The
-  resemblance is due to a serendipitous computer-graphical effect
-  \citep{2001AJ....122..585S}.} The spatial scale is in pixels.  The
-information in these two panels was, of course, kept secret during the
-modeling challenge.  The panels in the middle row and at the lower
-right are the results that \spl returns to the modeler, in order for
-them to assess the model.  These derive from the mean of an ensemble
-of 200 models generated by \spl.  The model mass distribution (left mid) is
-returned as a contour map superimposed on an intensity map. A fairly
-smooth mass distribution, as here, is a good sign.  An irregular or
-checkboard pattern in the mass map usually indicates a bad model.  The
-models arrival-time contours are shown in the right middle panel.
-  The special contour passing through a
-saddle point is shown in black: this is the model version of the
-spaghetti sketch provided by the user.  Also provided is a synthetic
-image of the lensed source in the bottom right panel.
-  For this, \spl assumes a simple conical
-source profile.  The user can change the contrast level on the image,
-which (though it is not saved) amounts to adjusting the width of
-the cone.  Finally, to the lower left of the figure, we have a
-comparison of the input and recovered mass profiles.  The panel shows
-a average $\kappa$ within a given radius, as a function of radius.
-The red curve is the true value, and where it crosses unity (dotted
-horizontal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
-  The two blue curves
-are the minimal and maximal mean enclosed $\kappa$ from the internal
-ensemble in \spl.  The region between the blue curves is shaded
-between the radii of the innermost and the outermost images --- this
-is the confidence region from the modeling.  As we see, the shaded
-blue is slightly above the red curve. 
-The Einstein Radius $\Theta_\text{E}$ of the model is estimated crossing the
-mean enclosed $\kappa$ (not plotted) with unity.
-  In summary: the identification
-of minimum and saddle point is correct, but the estimated Einstein
-radius is a little too high.
-
-Figure \ref{fig:6975} shows a lens with substructure in the form of a
-smaller secondary galaxy.  The galaxies in such group or cluster sims
-were, in fact, based on galaxies visible in the images --- but the
-modelers were not told in advance whether this was the case.  The
-model does not include any substructure, but otherwise is not bad.
-The minimum and saddle point are correct, and the Einstein radius is
-only a little underestimated.
-
-Figure \ref{fig:6937} shows a case where substructure leads to a
-poor model.
-
-Figure \ref{fig:6990} shows an example of an arc that has split into
-three images.  This kind of configuration, with a counter-image close
-to the lensing galaxy and a more distant arc/triplet on the other
-side, generically arises from an elongated mass distribution when the
-source is displaced along the elongated direction.
-
-Figure \ref{fig:6915} shows another quad.  This kind of configuration
-arises when the mass is elongated and the source is displaced at an
-angle to the elongation.
-
-Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
-saddle points are correctly identified, and the orientation of the
-ellipticity of the mass distribution is correctly reproduced.  The
-Einstein radius is somewhat overestimated.  Figure \ref{fig:7022}
-shows another model of the same system.  In this one, the
-identification of the minima and saddle points was incorrect, and mass
-distribution comes out elongated East-West instead of North-South.
-The mass distribution also appears somewhat jagged and the
-saddle-point contours are not as clean as in the previous examples;
-these are often indicators of a problem with the model.  The enclosed
-mass is, however, none the worse --- the reason is probably that in a
-relatively symmetrical image configuration, the Einstein radius is quite
-well constrained by the images in a fairly model-independent way.
-
-To conclude this set of examples, Figure \ref{fig:6919} shows another
-type of quad.  Actually, in this case the brightest part of the source
-is only doubly imaged, but the source extends into a region that
-produces four images.  We can also see from the real arrival-time
-surface that a point source is a double on the verge of splitting into
-a quad.  The modeler interpreted the system as a quad.  The
-appearance of the arcs, shown in the bottom panel of Figure
-\ref{fig:input-spag}, looks like an arc and counter-image such as
-discussed with Figure \ref{fig:6990} above.  But there is an important
-difference: the long arc is closer to the galaxy, as if the arc and
-counter-image have swapped roles.  This configuration arises if the
-source displacement is perpendicular to the long axis of the lensing
-mass.
-
-\subsection{Test of image identification} \label{sec:tests.t1}
-
-A first manual evaluation tested the volunteers ability to reconstruct the arrival time surface given a survey image containing a sim.
-This task consists of two parts.
-First to correctly identify and locate the lensed images.
-Second the correct ordering for the identified lensed images in respect of the arrival time.
-
-While we expected the identification of lensed images to be trivial, given the nature of the survey images and the success of \sw, we expect the correct ordering to be more difficult.
-This part tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
-While we can provide the volunteers with some general rules of thumb, ordering involves imagination and guessing and therefore training could improve results in a later stage.
-
-This first test was designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
-
-The evaluation of the volunteers performance was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot generated using simulation parameters.
-% \Figref{output_compare} shows the setup used to evaluate the models.
-
-The images of the system are considered to be identified correctly, if all the images have been identified and are approximated within $\pm0.05\cdot\text{imgage width}$.
-The parity is considered correct, if those identified points have the right ordering with respect to arrival time.
-
-Additionally, ten types of errors (labeled E01 -- E10, listed in \tabref{stats}) that occurred were identified.
-Each generated model could contain more than one error.
-
-\begin{table}\centering\begin{tabular}{llcc}\hline
- & & n & p \\
-\hline
- N: & Total number of models & 119 & 1.00\\
-\hline
- R1: & images approx. on right location & 110 & 0.92\\
- R2: & images with correct parity & 70 & 0.59\\
-\hline
- E01: & inaccurate in arc & 21 & 0.18\\
- E02: & wrong parity in 3 lens conf. & 2 & 0.02\\
- E03: & identified 3 of 5 imgs. & 5 & 0.04\\
- E04: & modeled arc with single img. & 4 & 0.03\\
- E05: & $\pi$ rotated parity & 7 & 0.06\\
- E06: & $\pi$/2 rotated parity & 38 & 0.32\\
- E07: & missed faint img. & 1 & 0.01\\
- E08: & too many imgs in arc. & 5 & 0.04\\
- E09: & missed double img & 3 & 0.03\\
- E10: & too many imgs. & 1 & 0.01\\
-\end{tabular}
-
-\caption{1. col: id of model; 2.col: places of imgs idendited
-  correctly; 3. col: more or less correct identification of extr
-  points. 4.col: exact identification of extremal points. 5.col:
-  type(s) of errors ocured....  1) inaccurate placement in an extended
-  arc; 2) wrongly identified sad and min in 3 image configuration; 3)
-  identified only 3 instead of 5 images; 4) tried to model an arc with
-  a min instead of min-sad-min; 5) PI-err (rotation by 180 degrees; in
-  5 image configuration, exchanged the ordering of the two saddle
-  points); 6) PI/2-err (rotation by 90 degree; sad-min-sad-min-sad);
-  7) missed faint image(s); 8) tried to model an arc with min-sad-min
-  instead of only min; 9) did identify two close by images as one; 10)
-  used 7 or more image to model a 5 image system. }
-
-\label{tab:stats}
-\end{table}
-
-\tabref{stats} presents a summary of this evaluation.
-We conclude that the volunteers are performing very well identifying and positioning images, with a performance of 92\% (R1, p=0.92).
-Most of the problems where due to unclear arc-like structures (E01, p=0.18; E04, p=0.03; E08, p=0.04).
-Critical errors like the failure to identify all five images in a five images system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happen.
-From this we conclude that the introduction materials was adequate and the volunteers understand the basics of gravitational lensing.
-
-The assignment of the parity of the images was a more difficult task.
-In 59\% (R2, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
-Most of the failures are due to E06 (N=38, p=32\%), followed by E05 (N=7, p=6\%).
-E06 describes a situation, where the minima and saddle points of a five image configuration were exchanged (rotated by $\pm90\dgr$), see \figref{7022} for an example.
-E05 describes a situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$).
-While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
-For more challenging cases, like very symmetrical distribution of the lensed images (for example model 7022, \figref{7022}), those errors should still produce plausible results, as will be explained in the next section.
-
-\subsection{Test of mass-profile recovery} \label{sec:tests.t2}
-
-The second test was to compare the mass distributions of the lens $\kappa(x, y)$ given by the sims and generated by the volunteers.
-To get a means of comparing the sims to the models, the total convergence\footnote{often called enclosed mass} $\kappa_{\text{encl}}(r)$ was calculated for both.
-The Einstein radius $\Theta_\text{E}$ is defined by $\kappa_{\text{encl}}(\Theta_\text{E})=1$ and gives a number that allows a rough comparison between a sim and a model.
-We also let an expert (PS) model three selected systems to compare the results from volunteers to those of a professional.
-
-To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
-From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate \ERg.
-To estimate the errors, the extremal models are used to estimate a lower and upper limit for \ERg.
-These results can be seen in \figref{ER_per_sim}.
-This figure shows that this technique of estimating the error using the ensemble of all models underestimates the error significantly and should be improved for further analysis.
-
-\begin{figure*}
-  \centering
-    \includegraphics[width=0.90\linewidth]{fig/eR_5}
-  \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
-  \label{fig:ER_per_sim}
-\end{figure*}
-
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.9\linewidth]
-             {fig/007022_kappa_encl}}
-  \subfigure{\includegraphics[width=0.9\linewidth]
-             {fig/007025_kappa_encl}}
-  \subfigure{\includegraphics[width=0.9\linewidth]
-             {fig/007021_kappa_encl}} \\
-  \caption{
-  %Enclosed mass \kenc from three models of a system with incorrect image identifications.
-\figref{7022} shows a model (by an expert) with the correct identifications, while \figref{7025} shows a model with error E06 (parities rotated by $\pi/2$).
-Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).
-  }
-\label{fig:kapenc_compare_faulty}
-\end{figure}
-
-\Figref{ER_per_sim} shows that the calculated \ERf of the models tend to be too high.
-The overshoot varies from around 0.2 to 0.4 for good models.
-One of the reasons for this is that it is hard to get the center of the lens on spot.
-An offset leads to a flatter mass profile for the model compared to the simulation.
-
-Comparing the models from volunteers and experts can be done in Figures \ref{fig:7022}, \ref{fig:7025} and \ref{fig:kapenc_compare_faulty}, where only the
-expert got the right configuration (\figref{7022}) but enclosed masses are all fairly similar.
-In cases with nearly symmetric image configurations, like this example, it is more difficult to identify the image parities correctly.
-Incorrect image-parity identification changes the orientation of the mass distribution \kap[x,y],
-but \kenc and thus the \ERf are not influenced that much.
-
-Two additional sims were modeled by an expert, \asw{1hpf} and \asw{0vqg}.
-Looking at the results for \ERg for those models in \figref{ER_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses, offset) is comparable.
-%Note that the models of \asw{0vqg} with \ERg[, rel] around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
-
-\clearpage
-\section{Outlook} \label{sec:todo}
-
-The lens-modeling challenge indicates that the Spacewarps collective
-is good, not only for identifying lens candidates for follow-up, but
-modeling candidates as well.
-
-There is, however, plenty of room for improvement.
-
-First, the particular modeling strategy implemented is not the only
-one possible.  \spl requires modelers to characterize the overall
-image structure in abstract terms based on Fermat's principle, and the
-placement of mass distributions is done by the computer.  In other
-modeling tools, the user puts down a trial mass distribution and has
-the machine refine it.  A few of these modeling programs have also
-been designed with citizen science in mind, and would be interesting
-in the Spacewarps environment.
-
-Second, \spl needs some enhancements.
-
-\begin{itemize}
-\item Currently, \spl does not attempt to model the source shape; the
-  user identifies the brightest points on the image, and these are
-  taken as images of a point-like source, whose positions must be
-  reproduced exactly. For generating a synthetic image, a conical
-  source profile is assumed. Fitting for the source profile to
-  optimize resemblance to the observed lensed image after the lens
-  model has been generated, is algorithmically straightforward and
-  planned to be implemented.  This would alleviate another problem
-  with \spl, which is that there is no numerical figure of merit, and
-  assessment of a model is a judgment call based on the synthetic
-  image, and on whether the mass distribution and the arrival-time
-  surface show suspect features.
-\item \spl has a tendency to somewhat overestimate the Einstein radius
-  (evident from Figure \ref{fig:ER_per_sim}), and it is also apparent
-  that the models tend to be too shallow.  This possible explanation
-  is that, while the sims are steeply peaked at the centre, the
-  pixellated mass model fixes a comparatively large area near the
-  central at constant density.  One way to solve this problem would be
-  to introduce smaller pixels in the central region, thus enabled a
-  steeper centre.
-\item Another limitation so far in \spl is that the lens is assumed to
-  be dominated by one galaxy, which puts most galaxy-group lenses
-  beyond the reach of the modeler. Since complicated group lenses are
-  some of the most interesting candidates present, removing this
-  limitation is most desirable.  From the users' point of view, it
-  would mean that spaghetti contours with more than one maximum can be
-  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
-  and Figure 4b in \cite{2003ApJ...590...39K}.
-\item At present, a single false-color composite is used as the data.
-  An option could be added to use all available filters, individually
-  or in combination, at the user wishes.
-\end{itemize}
-
-The third desirable avenue of improvement is to facilitate
-collaborative work.
-
-\begin{itemize}
-\item As mentioned above, the option of revising an already-archived
-  model is already available.  Desired now are tools for comparing
-  different models of a given system, both visually and through
-  different statistical measures.  As evidenced by a current
-  collaborative modeling effort, a particularly interesting candidate
-  can lead to an extended discussion and dozens of models.
-\item Better tutorial materials are also needed, and this would
-  address some of the problem areas found in the modeling challenge.
-  For example, we saw in \secref{tests.t1} that volunteers are most
-  prone to making errors in two situations: when in identifying an arc-
-  like structure while placing the points, and in identifying the
-  correct ordering of the points in nearly-symmetric configurations.
-\end{itemize}
-
-\section{Acknowledgments}
-
-We thank the Swiss Society for Astrophysics and Astronomy and the
-Swiss Academy of Sciences.
-
-\clearpage
-
-\appendix
-
-\section{Relation to standard lensing formalism} \label{sec:more-theory}
-
-The presentation of the arrival-time surface in \secref{Fermat}
-omitted details, for the sake of a more intuitive explanation.
-Specifically, the geometric time delay $t_{\rm geom}$ and the
-convergence $\kappa$ were left as proportionalities (equations
-\ref{eq:tgeom} and \ref{eq:kappa}), and the gravitational time delay
-$t_{\rm grav}$ was given in an implicit form (equation
-\ref{eq:tgrav}).  We now fill in the details.
-
-The original formulation of the arrival-time surface appears in
-equations (2.1) to (2.6) of \cite{1986ApJ...310..568B}.  Their
-equations can be rearranged as follows.
-\begin{equation} \label{cfBN86}
-\begin{aligned}
-t_{\rm geom} &= \frac{(1+z_L)}{2c} \frac{d_S}{d_L d_{LS}}
-\left[ (x-s_x)^2 + (y-s_y)^2 \right] \\
-\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y) \\
-\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{d_L d_{LS}}{d_S}
-               \times \Sigma(x,y)
-\end{aligned}
-\end{equation}
-We symbols $d_L,d_S$ and $d_{LS}$ are angular-diameter distances,
-respectively from observer to lens, observer to source, and lens to
-source.  In the concordance cosmology, we have
-\begin{equation}
-d_{LS} = \frac{c}{H_0}\,\frac1{1+z_S} \,
-\int_{z_L}^{z_S} \!\! \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
-%\raise.5ex\hbox{.}
-\end{equation}
-and similarly $d_S$ and $d_{LS}$.  We have replaced angular positions
-on the sky with positions on the lens plane as
-\begin{equation}
-(x,y) = d_L (\theta_x,\theta_y) \,.
-\end{equation}
-
-The first line of equation \eqref{cfBN86} is $t_{\rm geom}$ from
-equation \eqref{eq:tgeom} with the proportionality filled in, and with
-the source offset at $(s_x,s_y)$ rather than at the origin.
-
-The last line of equation \eqref{cfBN86} fills in the proportionality
-factor in equation \eqref{eq:kappa} for the convergence (or
-dimensionless surface density) $\kappa$.
-
-The middle line of equation \eqref{cfBN86}, though very different in
-form to the expression \eqref{eq:tgrav} for the gravitational time
-delay, is in fact equivalent.  This can be shown by considering a
-multipole expansions for the solution of the latter equation.  Another
-way to see the equivalence is to write a discrete version of
-\eqref{eq:tgrav}.  Introducing a small distance $\Delta$, we write
-\begin{equation} \label{eq:relax}
-\begin{aligned}
-\tgrav(x,y) &=
-\frac14 \Big[ \tgrav(x+\Delta,y) + \tgrav(x-\Delta,y) + {}\\
-&\kern24pt    \tgrav(x,y+\Delta) + \tgrav(x,y-\Delta) \Big] \\
-&+ (1+z_L)\frac{2G}{c^3} \, \pi\Delta^2 \, \Sigma(x,y) \,.
-\end{aligned}
-\end{equation}
-The circular average $\langle \tgrav^{\textstyle\circ} \rangle$ has
-been replaced by the average over four neighbouring points, and the
-enclosed mass $M_{\textstyle\bullet}$ has been replaced by
-$\pi\Delta^2\,\Sigma(x,y)$.  Now let us read equation \eqref{eq:relax}
-as an iterative formula for $t_{\rm grav}(x,y)$.  It is then
-recognisable as a formula for solving the two-dimensional Poisson
-equation from \eqref{cfBN86} by relaxation.
-
-\newpage
+\input partC
 

### 2014-06-05 text/partA.tex
partA partB partC

--- 2014-01-20 text/ms.tex
@@
-\documentclass[12pt,preprint]{aastex}
-
-\usepackage{amsmath}
-\usepackage{xspace}     % fixes spaces after custom commands
-\usepackage{float}      % exact placement of figures (using \begin{figure}[H])
-\usepackage{subfigure}  % does what it says
-
-%\usepackage{longtable}   % those two to generate the 2 ol table
-%\usepackage{multicol}
-
-\def\half{{\textstyle\frac12}}
-% makes it easier to have consistent writing, and easy change if needed
-\newcommand{\spl}{SpaghettiLens\xspace}
-\newcommand{\sw}{SpaceWarps\xspace}
-
-% shortcut for einstein radius (Text Greek Full Math)
-\newcommand{\ER}{Einstein radius\xspace} % text
-\newcommand{\ERg}[1][]{$\Theta_{\text{E#1}}$\xspace} % er in textmode with greek
-\newcommand{\ERf}[1][]{Einstein radius $\Theta_\text{E#1}$\xspace} % full output
-\newcommand{\ERm}[1][]{\Theta_\text{E#1}} % mathmode with greek symbol
-
-%shortcuts for kappa
-\newcommand{\kenc}[1][r]{$\kappa_\text{encl}(#1)$\xspace}
-\newcommand{\kap}[1][r]{$\kappa(#1)$\xspace}
-
-% shorcuts for refs (use capital for beginning of sentence)
-% first 3 are for the real layz people..
-\newcommand{\fref}[1]{\ref{fig:#1}}
-\newcommand{\sref}[1]{\ref{sec:#1}}
-\newcommand{\tref}[1]{\ref{tab:#1}}
-\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
-\newcommand{\secref}[1]{Section~\ref{sec:#1}}
-\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
-\newcommand{\Figref}[1]{Figure~\ref{fig:#1}}
-\newcommand{\Secref}[1]{Section~\ref{sec:#1}}
-\newcommand{\Tabref}[1]{Table~\ref{tab:#1}}
-
-% shortcut for ASW000xxxx
-\newcommand{\asw}[1]{ASW000#1\xspace}
-\newcommand{\model}[2][Model~]{#1#2\xspace}
-
-%degrees
-\newcommand{\dgr}{^{\circ}}
-
-% todo annotations on side in red. easy to search for at the end
-% !!! REMOVE THIS LINES FOR FINAL COMPILE
-%
-\usepackage{setspace}    % remove this pgk for final compile!! 
-\usepackage{color}    % well... color...
-
-\newcommand{\todo}[2][red]{%
-\textcolor{#1}{\textbullet}%
-\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
-\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}
-
-\newcommand{\needfig}[1][]{%
-\colorbox{yellow}{\parbox{0.5\textwidth}{%
-\vspace{0.5cm}\setstretch{0.5}\textcolor{black}{\scriptsize{fig #1}}\vspace{0.5cm}}}%
-\marginpar{\colorbox{yellow}{\parbox{\marginparwidth}{%
-\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{fig}}}}}}
-
-\setlength{\marginparsep}{0mm}
-\setlength{\marginparwidth}{2.2cm}
-
-\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
-\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
-\newcommand{\needchk}[1][]{\todo[red]{chk #1}}
-
-% \\\ END OF REMOVAL SECION
-
-\newcommand{\hr}{\vspace{5mm}\noindent\rule{0.8\textwidth}{0.4pt}\vspace{5mm}}
-
-\begin{document}
-
-\title{Lens Modeling in \sw}
-\author{First Author,\altaffilmark{1}
-Second Author,\altaffilmark{2} and
-Third Author\altaffilmark{3}}
-\altaffiltext{1}{First place}
-\altaffiltext{2}{Second place}
-\altaffiltext{3}{Third place}
-
 \begin{abstract}
@@
 (iii)~the performance of a volunteer vs. a professional.
-In most cases, the volunteers were found to correctly identify the image parities and time arrivals, along with a mean convergence that was contained within the image region.
-In all, the results could be comparable to that of professionals.
-
+In most cases, the volunteers were found to correctly identify the image parities and time arrivals.
+% Einstein radius +~20%
+ along with a mean convergence that was well constrained within the image region;
+In all, the results could be comparable to that of a professional.
 \end{abstract}
 
-\keywords{}
+\begin{keywords}
+\end{keywords}
+
 \section{Introduction}
 
-Our present picture of the formation of the largest structures in the
-Universe --- galaxies and clusters of galaxies --- is that
-gravitational instabilities (whose early stages are observable as
-fluctuations of the microwave background) caused a collapse into
-gravitationally bound structures.  Dark matter would have dominated
-the initial collapse, but once deep wells in the gravitational
-potential appeared, gas would have fallen into them while radiating
-away its gravitational potential energy through atomic processes to
-form the first stars and galaxies.  With time, galaxies and clusters
-would continue to merge and change their form.  This rough outline is
-well-established now, but many puzzles remain to be solved before our
-understanding is satisfactory.  To progress towards solving these, we
-need more knowledge of dark matter.  And to study dark matter,
-gravitational lensing is attractive, because it depends only on mass.
+Galaxies and clusters of galaxies are considered to be the largest structures in the known Universe.
+When one questions their formation,  the present picture suggests that gravitational instabilities  (observable as fluctuations of the microwave background) %TODO \todo[red]{not too sure of the relevance of this statement}
+ - caused a collapse into gravitationally bound structures.
+%TODO\todo[yellow]{clarification needed: are they becoming gravitationally bound structures, or are they collapsing into objects that are already gravitationally bound?
+I know it is most probably the later, but it is a bit ambiguous}
+The initial collapse would have been dominated by dark matter, but once deep wells had appeared in the gravitational potential, gas would have started falling into those wells. 
+%TODO \todo[red]{not too sure of this wording}
+Through atomic processes,
+%TODO \todo[yellow]{perhaps we need to clarify those processes or are they generally know?}
+ the gravitational potential energy of this gas would of been radiated away, leading to the formation of the largest stars. 
+%TODO \todo[red]{or perhaps just stars?}
+In turn, galaxies and clusters of galaxies would form, all the time merging and changing their form.
+This rough outline is now well established, but in order to satisfy our understanding of this process, there are many puzzles  that remain to be solved.
+For progress to be made, we need to extend our knowledge of dark matter.
+Gravitational lensing has proven as a particularly attractive way of studying dark matter as it is only dependant on mass.
+%TODO \todo[yellow]{silly question, but shouldn't we briefly explain why it is only dependant on mass? or should everyone already reading this paper know that? or is it explained later?}
 
@@
 \centering
-\includegraphics[width=.8\hsize]{fig/lenssky.png}
-\caption{Sky distribution of 398 published lenses considered secure as
-  of September 2013 (from the Masterlens catalog at the University of
+\includegraphics[width=0.9\hsize]{fig/lenssky}
+\caption{Sky distribution of 423 published lenses considered secure as
+  of January 2014 (from the Masterlens catalog at the University of
   Utah, maintained by Joel Brownstein and Leonidas Moustakas.) The map
@@
 observed, but several observable quantities can be derived from it.
+
+\begin{equation} \label{eq:tgeom}
+\tgeom(x,y) \propto (x-x_s)^2 + (y-s_y)^2 \,.
+\end{equation}
+
+The gravitational time delay can be expressed in an implicit form: the
+value of $\tgrav$ through $(x,y)$ equals its average value on the
+circumference of a small circle centred at $(x,y)$, plus a constant
+times the mass within that circle.  The proportionality constant is
+$2G/c^3$ times the cosmological expansion factor $(1+z_L)$. Thus
+\begin{equation} \label{eq:tgrav}
+\tgrav(x,y) = \left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle
+              + (1+z_L) \frac{2G}{c^3} M(x_\subbullet,y_\subbullet) \,.
+\end{equation}
+We have used $(x_\subcirc,y_\subcirc)$ to denote the circumference of
+a circle, and $(x_\subbullet,y_\subbullet)$ to indicate the integrated
+mass within the circle.
+
+\begin{equation} \label{eq:kappa}
+\kappa(x,y) \propto \Sigma(x,y) \,.
+\end{equation}
+
+Incidentally, $\kappa$ has a second meaning, as well as
+being a dimensionless form of the projected density: it also
+measures how a bundle of light rays is brought together by
+gravitational lensing, and is called the {\em convergence}.
 
@@
 \centering
-\includegraphics[height=.3\vsize]{fig/arriv1.png}
-\includegraphics[height=.3\vsize]{fig/arriv2.png}
-\includegraphics[height=.3\vsize]{fig/arriv3.png}
-\caption{Arrival-time surfaces, with contour levels of equal arrival
-  time.  The upper surface is with no lens.  The middle surface is
-  when a circular lensing mass (offset from the source) is added; a
-  maximum, a minimum and a saddle point can be seen.  The last surface
-  is the result of an elongated lensing mass; a maximum, two minima
-  and two saddle points can be seen.
-}
+\includegraphics[width=\columnwidth]{fig/arriv_0}
+\includegraphics[width=\columnwidth]{fig/arriv_1}
+\includegraphics[width=\columnwidth]{fig/arriv_2}
+\caption{Perspective views and contour maps of example arrival time
+  surfaces.  Contours are coloured in rainbow order (red: least delay,
+  violet: highest delay).  {\em Upper panel:\/} No lens, hence showing
+  the parabolic shape of the geometrical time delay.  The image would
+  be at the bottom, coinciding with the source. {\em Middle panel:\/}
+  A circular lensing mass (offset from the source) has been added,
+  which has pushed the minimum to one side and introduced a maximum
+  and saddle point, each corresponding to an image.  The saddle-point
+  is characterized by a self-crossing ``spaghetti'' contour. {\em
+    Lower panel:\/} An elongated lensing mass has been added.  There
+  are now two minima, two saddle points, and a maximum, each
+  corresponding to an image.  Note the two self-crossing spaghetti
+  contours associated with the saddle points.}
 \label{fig:arriv}
@@
 program.
-
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/006941_input.png}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007022_input.png}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/006919_input.png}}
-  \caption{Examples of Spaghetti input.  The models from these appear
-    later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
-  }
-  \label{fig:input-spag}
-\end{figure}
 
@@
 
+% Rafi's original one-coumn design:
+%
+% \begin{figure}
+%   \centering
+%   \subfigure[real mass distribution]{
+%     \label{fig:6941_sim_mass}
+%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa}
+%   }
+%   \subfigure[real arrival-time surface]{
+%     \label{fig:6941_sim_arr}
+%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv}
+%   }
+%   \subfigure[model mass distribution]{
+%     \label{fig:6941_mass}
+%     \includegraphics[width=0.45\textwidth]{fig/006941_mass}
+%   }
+%   \subfigure[model arrival-time surface]{
+%     \label{fig:6941_cont}
+%     \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti}
+%   }
+%   \subfigure[real vs model enclosed mass]{
+%     \label{fig:6941_kappa}
+%     \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl}
+%   }
+%   \subfigure[model lensed image]{
+%     \label{fig:6941_atime}
+%     \includegraphics[width=0.45\textwidth]{fig/006941_arr_time}
+%   }
+%   \caption[result 6941 (ASW000102p)]{Model results for a simple
+%     simulated lens.  (The label ASW000102p in the lower-left panel is
+%     the \sw identifier for the sky patch where this sim was placed.)
+%     The model input is shown in the top panel of
+%     Figure~\ref{fig:input-spag}.  For details of the individual panels
+%     here, see section \ref{sec:example_models}.}
+%   \label{fig:6941}
+% \end{figure}
+
+% Rafik's two row design
+
+\newpage
+
 \begin{figure}
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6941_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa.png}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6941_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv.png}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6941_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006941_mass.png}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6941_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti.png}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6941_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl.png}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6941_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006941_arr_time.png}
-  }
-  \caption[result 6941 (ASW000102p)]{Sim ASW000102p and a model.  A
-    simple configuration; the model input is shown in the top panel of
-    Figure~\ref{fig:input-spag}.  For details of the individual panels
-    here, see section \ref{sec:example_models}.}
+  \includegraphics[width=\myplotswidth]{fig/006941_input}
+  \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
+  \vskip-20pt
+  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 
+  \includegraphics[width=\myplotswidth]{fig/006941_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006941_mass}
+  \caption[result 6941 (ASW000102p)]{A simulated lens that mimics a
+    lensed quasar, and model results.  The left panels derive from the
+    simulation, and the right panels are \sl output.  Details of
+    individual panels are in \ref{sec:example_models}.}
+
   \label{fig:6941}
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6975_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_kappa.png}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6975_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_arriv.png}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6975_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006975_mass.png}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6975_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006975_spaghetti.png}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6975_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006975_kappa_encl.png}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6975_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006975_arr_time.png}
-  }
-  \caption[result 6975 (ASW000195x)]{Sim ASW000195x and a
-    model.}
+
+  \includegraphics[width=\myplotswidth]{fig/006915_input}
+  \includegraphics[width=\myplotswidth]{fig/006915_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006915_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006915_mass}
+
+  \caption[result 6915 (ASW0001hpf)]{A four-image configuration
+    typical of lensed quasars. (See Section \ref{sec:example_models}
+    for details.)}
+  \label{fig:6915}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/006990_input}
+  \includegraphics[width=\myplotswidth]{fig/006990_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006990_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006990_mass}
+
+  \caption[result 6990 (ASW0004oux)]{Results from a system with an arc
+    plus a counter-image, typical of lensed galaxies. (See Section
+    \ref{sec:example_models} for details.)}
+  \label{fig:6990}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/006919_input}
+  \includegraphics[width=\myplotswidth]{fig/006919_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006919_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006919_mass}
+
+  \caption[result 6919 (ASW0002z6f)]{Another configuration of arc plus
+    counter-image; here the arc is closer to the lensing galaxy than
+    the counter-image. (See Section \ref{sec:example_models} for
+    details.)}
+  \label{fig:6919}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\myplotswidth]{fig/006975_input}
+  \includegraphics[width=\myplotswidth]{fig/006975_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_arriv.png}
+  \includegraphics[width=\myplotswidth]{fig/006975_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_kappa.png}
+  \includegraphics[width=\myplotswidth]{fig/006975_mass}
+  \caption[result 6975 (ASW000195x)]{A lens with unrecovered mass
+    substructure. (See Section \ref{sec:example_models} for details.)}
   \label{fig:6975}
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6937_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_kappa.png}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6937_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_arriv.png}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6937_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006937_mass.png}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6937_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006937_spaghetti.png}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6937_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006937_kappa_encl.png}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6937_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006937_arr_time.png}
-  }
-  \caption[result 6937 (ASW0000vqg)]{Sim ASW0000vqg and a model.}
+
+  \includegraphics[width=\myplotswidth]{fig/006937_input}
+  \includegraphics[width=\myplotswidth]{fig/006937_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006937_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006937_mass}
+
+  \caption[result 6937 (ASW0000vqg)]{A sim with unrecovered
+    substructure, resulting in a poor mass model. (See Section
+    \ref{sec:example_models} for details.)}
   \label{fig:6937}
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6990_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_kappa.png}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6990_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_arriv.png}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6990_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006990_mass.png}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6990_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006990_spaghetti.png}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6990_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006990_kappa_encl.png}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6990_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006990_arr_time.png}
-  }
 
-  \caption[result 6990 (ASW0004oux)]{Sim ASW0004oux and a model.}
-  \label{fig:6990}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6915_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_kappa.png}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6915_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_arriv.png}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6915_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006915_mass.png}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6915_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006915_spaghetti.png}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6915_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006915_kappa_encl.png}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6915_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006915_arr_time.png}
-  }
-
-  \caption[result 6915 (ASW0001hpf)]{Sim ASW0001hpf and a model. Inclined quad.}
-  \label{fig:6915}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:7022_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa.png}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:7022_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv.png}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:7022_mass}
-    \includegraphics[width=0.45\textwidth]{fig/007022_mass.png}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:7022_cont}
-    \includegraphics[width=0.45\textwidth]{fig/007022_spaghetti.png}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:7022_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/007022_kappa_encl.png}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:7022_atime}
-    \includegraphics[width=0.45\textwidth]{fig/007022_arr_time.png}
-  }
-
-  \caption[result 7022 (ASW0000h2m)]{Sim ASW0000h2m and a model.  The
-    model input appears in the middle panel of Figure
-    \ref{fig:input-spag}.}
-  \label{fig:7022}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:7025_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa.png}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:7025_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv.png}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:7025_mass}
-    \includegraphics[width=0.45\textwidth]{fig/007025_mass.png}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:7025_cont}
-    \includegraphics[width=0.45\textwidth]{fig/007025_spaghetti.png}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:7025_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/007025_kappa_encl.png}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:7025_atime}
-    \includegraphics[width=0.45\textwidth]{fig/007025_arr_time.png}
-  }
-
-  \caption[result 7025 (ASW0000h2m)]{Sim ASW0000h2m and another model.  The image parities in this case were incorrectly identified.}
+  \includegraphics[width=\myplotswidth]{fig/007025_input}
+  \includegraphics[width=\myplotswidth]{fig/007025_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_arriv}
+  \includegraphics[width=\myplotswidth]{fig/007025_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_kappa}
+  \includegraphics[width=\myplotswidth]{fig/007025_mass}
+  \caption[result 7025 (ASW0000h2m)]{A four-image system with image
+    parities incorrectly identified.  The model is poor, but the
+    estimated Einstein radius is not bad. (See Section
+    \ref{sec:example_models} for details.)}
   \label{fig:7025}
@@
   \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6919_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_kappa.png}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6919_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_arriv.png}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6919_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006919_mass.png}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6919_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006919_spaghetti.png}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6919_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006919_kappa_encl.png}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6919_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006919_arr_time.png}
-  }
-  \caption[result 6919 (ASW0002z6f)]{Sim ASW0002z6f and a model.  The
-    model input appears in the bottom panel of Figure
-    \ref{fig:input-spag}. An incipient quad}
-  \label{fig:6919}
+
+  \includegraphics[width=\myplotswidth]{fig/007022_input}
+  \includegraphics[width=\myplotswidth]{fig/007022_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_arriv}
+  \includegraphics[width=\myplotswidth]{fig/007022_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_kappa}
+  \includegraphics[width=\myplotswidth]{fig/007022_mass}
+
+  \caption[result 7022 (ASW0000h2m)]{The same system as in Figure
+    \ref{fig:7025}, this time with image parities correctly
+    identified. (See Section \ref{sec:example_models} for details.)}
+  \label{fig:7022}
 \end{figure}
+
+\newpage
 
@@
 mass.
-
-\subsection{Test of image identification} \label{sec:tests.t1}
-
-A first manual evaluation tested the volunteers ability to reconstruct the arrival time surface given a survey image containing a sim.
-This task consists of two parts.
-First to correctly identify and locate the lensed images.
-Second the correct ordering for the identified lensed images in respect of the arrival time.
-
-While we expected the identification of lensed images to be trivial, given the nature of the survey images and the success of \sw, we expect the correct ordering to be more difficult.
-This part tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
-While we can provide the volunteers with some general rules of thumb, ordering involves imagination and guessing and therefore training could improve results in a later stage.
-
-This first test was designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
-
-The evaluation of the volunteers performance was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot generated using simulation parameters.
-% \Figref{output_compare} shows the setup used to evaluate the models.
-
-The images of the system are considered to be identified correctly, if all the images have been identified and are approximated within $\pm0.05\cdot\text{imgage width}$.
-The parity is considered correct, if those identified points have the right ordering with respect to arrival time.
-
-Additionally, ten types of errors (labeled E01 -- E10, listed in \tabref{stats}) that occurred were identified.
-Each generated model could contain more than one error.
-
-\begin{table}\centering\begin{tabular}{llcc}\hline
- & & n & p \\
-\hline
- N: & Total number of models & 119 & 1.00\\
-\hline
- R1: & images approx. on right location & 110 & 0.92\\
- R2: & images with correct parity & 70 & 0.59\\
-\hline
- E01: & inaccurate in arc & 21 & 0.18\\
- E02: & wrong parity in 3 lens conf. & 2 & 0.02\\
- E03: & identified 3 of 5 imgs. & 5 & 0.04\\
- E04: & modeled arc with single img. & 4 & 0.03\\
- E05: & $\pi$ rotated parity & 7 & 0.06\\
- E06: & $\pi$/2 rotated parity & 38 & 0.32\\
- E07: & missed faint img. & 1 & 0.01\\
- E08: & too many imgs in arc. & 5 & 0.04\\
- E09: & missed double img & 3 & 0.03\\
- E10: & too many imgs. & 1 & 0.01\\
-\end{tabular}
-
-\caption{1. col: id of model; 2.col: places of imgs idendited
-  correctly; 3. col: more or less correct identification of extr
-  points. 4.col: exact identification of extremal points. 5.col:
-  type(s) of errors ocured....  1) inaccurate placement in an extended
-  arc; 2) wrongly identified sad and min in 3 image configuration; 3)
-  identified only 3 instead of 5 images; 4) tried to model an arc with
-  a min instead of min-sad-min; 5) PI-err (rotation by 180 degrees; in
-  5 image configuration, exchanged the ordering of the two saddle
-  points); 6) PI/2-err (rotation by 90 degree; sad-min-sad-min-sad);
-  7) missed faint image(s); 8) tried to model an arc with min-sad-min
-  instead of only min; 9) did identify two close by images as one; 10)
-  used 7 or more image to model a 5 image system. }
-
-\label{tab:stats}
-\end{table}
-
-\tabref{stats} presents a summary of this evaluation.
-We conclude that the volunteers are performing very well identifying and positioning images, with a performance of 92\% (R1, p=0.92).
-Most of the problems where due to unclear arc-like structures (E01, p=0.18; E04, p=0.03; E08, p=0.04).
-Critical errors like the failure to identify all five images in a five images system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happen.
-From this we conclude that the introduction materials was adequate and the volunteers understand the basics of gravitational lensing.
-
-The assignment of the parity of the images was a more difficult task.
-In 59\% (R2, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
-Most of the failures are due to E06 (N=38, p=32\%), followed by E05 (N=7, p=6\%).
-E06 describes a situation, where the minima and saddle points of a five image configuration were exchanged (rotated by $\pm90\dgr$), see \figref{7022} for an example.
-E05 describes a situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$).
-While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
-For more challenging cases, like very symmetrical distribution of the lensed images (for example model 7022, \figref{7022}), those errors should still produce plausible results, as will be explained in the next section.
-
-\subsection{Test of mass-profile recovery} \label{sec:tests.t2}
-
-The second test was to compare the mass distributions of the lens $\kappa(x, y)$ given by the sims and generated by the volunteers.
-To get a means of comparing the sims to the models, the total convergence\footnote{often called enclosed mass} $\kappa_{\text{encl}}(r)$ was calculated for both.
-The Einstein radius $\Theta_\text{E}$ is defined by $\kappa_{\text{encl}}(\Theta_\text{E})=1$ and gives a number that allows a rough comparison between a sim and a model.
-We also let an expert (PS) model three selected systems to compare the results from volunteers to those of a professional.
-
-To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
-From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate \ERg.
-To estimate the errors, the extremal models are used to estimate a lower and upper limit for \ERg.
-These results can be seen in \figref{ER_per_sim}.
-This figure shows that this technique of estimating the error using the ensemble of all models underestimates the error significantly and should be improved for further analysis.
-
-\begin{figure}[htbp]
-  \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_4.png}
-  \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
-  \label{fig:ER_per_sim}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007020_kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007024_kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007021_kappa_encl}} \\
-  \caption{Enclosed mass \kenc for models of \asw{0h2m} with incorrect image identifications.
-\figref{7022} shows a model (by an expert) with the correct identifications, while \figref{7025} shows a model with error E06 (parities rotated by $\pi/2$).
-Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).}
-\label{fig:kapenc_compare_faulty}
-\end{figure}
-
-\Figref{ER_per_sim} shows that the calculated \ERf of the models tend to be too high.
-The overshoot varies from around 0.2 to 0.4 for good models.
-One of the reasons for this is that it is hard to get the center of the lens on spot.
-An offset leads to a flatter mass profile for the model compared to the simulation.
-
-Comparing the models from volunteers and experts can be done in Figures \ref{fig:7022}, \ref{fig:7025} and \ref{fig:kapenc_compare_faulty}, where only the
-expert got the right configuration (\figref{7022}) but enclosed masses are all fairly similar.
-In cases with nearly symmetric image configurations, like this example, it is more difficult to identify the image parities correctly.
-Incorrect image-parity identification changes the orientation of the mass distribution \kap[x,y],
-but \kenc and thus the \ERf are not influenced that much.
-
-Two additional sims were modeled by an expert, \asw{1hpf} and \asw{0vqg}.
-Looking at the results for \ERg for those models in \figref{ER_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses, offset) is comparable.
-%Note that the models of \asw{0vqg} with \ERg[, rel] around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
-
-\clearpage
-\section{Outlook} \label{sec:todo}
-
-The lens-modeling challenge indicates that the Spacewarps collective
-is good, not only for identifying lens candidates for follow-up, but
-modeling candidates as well.
-
-There is, however, plenty of room for improvement.
-
-First, the particular modeling strategy implemented is not the only
-one possible.  \spl requires modelers to characterize the overall
-image structure in abstract terms based on Fermat's principle, and the
-placement of mass distributions is done by the computer.  In other
-modeling tools, the user puts down a trial mass distribution and has
-the machine refine it.  A few of these modeling programs have also
-been designed with citizen science in mind, and would be interesting
-in the Spacewarps environment.
-
-Second, \spl needs some enhancements.
-
-\begin{itemize}
-\item Currently, \spl does not attempt to model the source shape; the
-  user identifies the brightest points on the image, and these are
-  taken as images of a point-like source, whose positions must be
-  reproduced exactly. For generating a synthetic image, a conical
-  source profile is assumed. Fitting for the source profile to
-  optimize resemblance to the observed lensed image after the lens
-  model has been generated, is algorithmically straightforward and
-  planned to be implemented.  This would alleviate another problem
-  with \spl, which is that there is no numerical figure of merit, and
-  assessment of a model is a judgment call based on the synthetic
-  image, and on whether the mass distribution and the arrival-time
-  surface show suspect features.
-\item \spl has a tendency to somewhat overestimate the Einstein radius
-  (evident from Figure \ref{fig:ER_per_sim}), and it is also apparent
-  that the models tend to be too shallow.  This possible explanation
-  is that, while the sims are steeply peaked at the centre, the
-  pixellated mass model fixes a comparatively large area near the
-  central at constant density.  One way to solve this problem would be
-  to introduce smaller pixels in the central region, thus enabled a
-  steeper centre.
-\item Another limitation so far in \spl is that the lens is assumed to
-  be dominated by one galaxy, which puts most galaxy-group lenses
-  beyond the reach of the modeler. Since complicated group lenses are
-  some of the most interesting candidates present, removing this
-  limitation is most desirable.  From the users' point of view, it
-  would mean that spaghetti contours with more than one maximum can be
-  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
-  and Figure 4b in \cite{2003ApJ...590...39K}.
-\item At present, a single false-color composite is used as the data.
-  An option could be added to use all available filters, individually
-  or in combination, at the user wishes.
-\end{itemize}
-
-The third desirable avenue of improvement is to facilitate
-collaborative work.
-
-\begin{itemize}
-\item As mentioned above, the option of revising an already-archived
-  model is already available.  Desired now are tools for comparing
-  different models of a given system, both visually and through
-  different statistical measures.  As evidenced by a current
-  collaborative modeling effort, a particularly interesting candidate
-  can lead to an extended discussion and dozens of models.
-\item Better tutorial materials are also needed, and this would
-  address some of the problem areas found in the modeling challenge.
-  For example, we saw in \secref{tests.t1} that volunteers are most
-  prone to making errors in two situations: when in identifying an arc-
-  like structure while placing the points, and in identifying the
-  correct ordering of the points in nearly-symmetric configurations.
-\end{itemize}
-
-\clearpage
-
-\appendix
-
-\section{More on Lensing Theory} \label{sec:more-theory}
-
-In \secref{Fermat}, for the sake of a more intuitive
-explanation, we suppressed some constant factors in
-equations \eqref{eq:Ageom}, \eqref{eq:Aarriv} and \eqref{eq:Poisson}.
-Here we fill in the details.
-
-First let us recall various distance formulas. Comoving distances
-involve integrals of the type
-\begin{equation}
-\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
-%\raise.5ex\hbox{.}
-\end{equation}
-Specifically, the comoving distance $D_S$ from observer to source,
-$D_{LS}$ from lens to source, and $D_L$ from observer to lens are
-given by the above integral, with limits as follows.
-\begin{equation}
-\begin{aligned}
-&D_S    &                                &\int_0^{z_S} \\
-\noalign{\medskip}
-&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
-\noalign{\medskip}
-&D_L    &                                &\int_0^{z_L}
-\end{aligned}
-\end{equation}
-Angular-diameter distances $d$ can be related to comoving distances
-$D$ by scale factors, as follows.
-\begin{equation}
-\begin{aligned}
-D_S &= (1+z_S) \, d_S \\
-D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
-D_L &= (1+z_L) \, d_L
-\end{aligned}
-\end{equation}
-Consequently, angular-diameter distances are always smaller than
-comoving distances.  Moreover, while comoving distances can be simply
-added, angular distances cannot:
-\begin{equation}
-d_S \neq d_L + d_{LS}
-\end{equation}
-
-Locations on the lens plane in angular coordinates on the sky, are
-expressed as
-\begin{equation}
-(x,y) = d_L (\theta_x,\theta_y)
-\end{equation}
-
-The $A$ areas (with the implicit
-constant factor that makes them time delays) and the $\kappa$ density are related to physical arrival
-time $t$ and density $\Sigma$ as
-\begin{equation}
-\begin{aligned}
-A           &= \frac{c}{(1+z_L)^2} \frac{D_L D_{LS}}{D_S} \times t \\
-\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L D_{LS}}{D_S}
-               \times \Sigma(x,y)
-\end{aligned}
-\end{equation}
-Letting the source be offset at $(s_x,s_y)$ rather than at the
-origin, we have
-\begin{equation}
-\begin{aligned}
-t_{\rm geom} &= \frac{(1+z_L)^2}{2c} \frac{D_S}{D_L D_{LS}}
-\left( (x-s_x)^2 + (y-s_y)^2 \right) \\
-\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
-\end{aligned}
-\end{equation}
-We can now compare with equations (2.1) to (2.6)
-from \cite{1986ApJ...310..568B}.
-%% comparison to be explicited
-
-\section{Miscellaneous}
-
-Maybe discard \figref{ER_all_models}, since the essential information is there in \figref{ER_per_sim}.
-
-\begin{figure}[htbp]
-  \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_1.png}
-  \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
-  \label{fig:ER_all_models}
-\end{figure}
-
-\newpage
-
-\bibliographystyle{apj}
-\bibliography{bib/ms}
-
-\end{document}

### 2014-06-05 text/partB.tex
partA partB partC

--- 2014-01-20 text/ms.tex
@@
-\documentclass[12pt,preprint]{aastex}
-
-\usepackage{amsmath}
-\usepackage{xspace}     % fixes spaces after custom commands
-\usepackage{float}      % exact placement of figures (using \begin{figure}[H])
-\usepackage{subfigure}  % does what it says
-
-\usepackage{graphicx}
-% prefere pdf over png (better quality, lossless compression)
-\DeclareGraphicsExtensions{%
-  .pdf,.PDF,%
-  .png,.PNG,%
-  .jpg,.mps,.jpeg,.jbig2,.jb2,.JPG,.JPEG,.JBIG2,.JB2}
-
-%\usepackage{longtable}   % those two to generate the 2 ol table
-%\usepackage{multicol}
-
-\def\half{{\textstyle\frac12}}
-% makes it easier to have consistent writing, and easy change if needed
-\newcommand{\spl}{SpaghettiLens\xspace}
-\newcommand{\sw}{SpaceWarps\xspace}
-
-% shortcut for einstein radius (Text Greek Full Math)
-\newcommand{\ER}{Einstein radius\xspace} % text
-\newcommand{\ERg}[1][]{$\Theta_{\text{E#1}}$\xspace} % er in textmode with greek
-\newcommand{\ERf}[1][]{Einstein radius $\Theta_\text{E#1}$\xspace} % full output
-\newcommand{\ERm}[1][]{\Theta_\text{E#1}} % mathmode with greek symbol
-
-%shortcuts for kappa
-\newcommand{\kenc}[1][r]{$\kappa_\text{encl}(#1)$\xspace}
-\newcommand{\kap}[1][r]{$\kappa(#1)$\xspace}
-
-% shorcuts for refs (use capital for beginning of sentence)
-% first 3 are for the real layz people..
-\newcommand{\fref}[1]{\ref{fig:#1}}
-\newcommand{\sref}[1]{\ref{sec:#1}}
-\newcommand{\tref}[1]{\ref{tab:#1}}
-\newcommand{\figref}[1]{Figure~\ref{fig:#1}}
-\newcommand{\secref}[1]{Section~\ref{sec:#1}}
-\newcommand{\tabref}[1]{Table~\ref{tab:#1}}
-\newcommand{\Figref}[1]{Figure~\ref{fig:#1}}
-\newcommand{\Secref}[1]{Section~\ref{sec:#1}}
-\newcommand{\Tabref}[1]{Table~\ref{tab:#1}}
-
-% shortcut for ASW000xxxx
-\newcommand{\asw}[1]{ASW000#1\xspace}
-\newcommand{\model}[2][Model~]{#1#2\xspace}
-
-%degrees
-\newcommand{\dgr}{^{\circ}}
-
-% todo annotations on side in red. easy to search for at the end
-% !!! REMOVE THIS LINES FOR FINAL COMPILE
-%
-\usepackage{setspace}    % remove this pgk for final compile!! 
-\usepackage{color}    % well... color...
-
-\newcommand{\todo}[2][red]{%
-\textcolor{#1}{\textbullet}%
-\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
-\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}
-
-\newcommand{\needfig}[1][]{%
-\colorbox{yellow}{\parbox{0.5\textwidth}{%
-\vspace{0.5cm}\setstretch{0.5}\textcolor{black}{\scriptsize{fig #1}}\vspace{0.5cm}}}%
-\marginpar{\colorbox{yellow}{\parbox{\marginparwidth}{%
-\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{fig}}}}}}
-
-\setlength{\marginparsep}{0mm}
-\setlength{\marginparwidth}{2.2cm}
-
-\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
-\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
-\newcommand{\needchk}[1][]{\todo[red]{chk #1}}
-
-% \\\ END OF REMOVAL SECION
-
-\newcommand{\hr}{\vspace{5mm}\noindent\rule{0.8\textwidth}{0.4pt}\vspace{5mm}}
-
-\begin{document}
-
-\title{Lens Modeling in \sw}
-\author{First Author,\altaffilmark{1}
-Second Author,\altaffilmark{2} and
-Third Author\altaffilmark{3}}
-\altaffiltext{1}{First place}
-\altaffiltext{2}{Second place}
-\altaffiltext{3}{Third place}
-
-\begin{abstract}
-In Spacewarps, volunteers are invited to search CFHTLS sky survey for
-lens candidates.  Here we report on a system that allows experienced
-volunteers to model lens candidates.  A sample of 29 simulated lenses
-were modeled, each multiple times.  The quality of the models was then
-examined in different ways.  In particular: (i)~the image parities and
-time arrivals were identified correctly in most cases, but not
-always; (ii)~the mean convergence (equivalent to the enclosed mass)
-was quite well constrained in the image region; (iii)~the performance
-of experienced volunteers was comparable to professionals.
-\end{abstract}
-
-\keywords{}
-\section{Introduction}
-
-Our present picture of the formation of the largest structures in the
-Universe --- galaxies and clusters of galaxies --- is that
-gravitational instabilities (whose early stages are observable as
-fluctuations of the microwave background) caused a collapse into
-gravitationally bound structures.  Dark matter would have dominated
-the initial collapse, but once deep wells in the gravitational
-potential appeared, gas would have fallen into them while radiating
-away its gravitational potential energy through atomic processes to
-form the first stars and galaxies.  With time, galaxies and clusters
-would continue to merge and change their form.  This rough outline is
-well-established now, but many puzzles remain to be solved before our
-understanding is satisfactory.  To progress towards solving these, we
-need more knowledge of dark matter.  And to study dark matter,
-gravitational lensing is attractive, because it depends only on mass.
-
-There are many manifestations of gravitational lensing in astrophysics,
-but the most spectacular are multiple images.  Gravitational lensing
-depends on the sky-projection mass density, and this fact has a
-counter-intuitive consequence: nearby galaxies do not produce multiple
-images, because their projected density is not high enough, but beyond
-100~Mpc or so, most galaxies are potential lenses.  Even so, for a
-potential lens to really act as a lens, a background source in the close-by
-region is required.  So, in practice galaxies are very rarely observed
-as lens.  The same tendency applies to clusters of galaxies, but since
-clusters are bigger, they are more likely lens-like.
-
-In lensing galaxies, the observable features usually involve a single
-object multiply imaged.  It is therefore desirable to have large
-numbers of lenses, and this is the strategy researchers are following.  The
-largest single study so far uses 58 lenses \citep{2009ApJ...703L..51K}.
-Other studies have compared lens models in more detail with stellar
-mass \citep{2011ApJ...740...97L} or used time-delay information from
-lensed quasars to infer the Hubble time
-\citep{2008ApJ...679...17C,2010ApJ...712.1378P}.
-
-The trend is clear: researchers want to find many more lenses and
-model them in more sophisticated ways.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=.7\hsize]{fig/lenssky}
-\caption{Sky distribution of 398 published lenses considered secure as
-  of September 2013 (from the Masterlens catalog at the University of
-  Utah, maintained by Joel Brownstein and Leonidas Moustakas.) The map
-  is in Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
-  middle.  The empty swathes to the left and right of center are the
-  Milky Way.}
-\label{fig:masterlens}
-\end{figure}
-
-A good summary of the observational situation is offered by
-Figure~\ref{fig:masterlens}, which shows known secure multiply-imaging
-lenses.  The non-uniformity on the sky is not intrinsic, it just
-indicates the density of deep surveys up to 2013.  At HST resolution,
-of order 1~square degree of sky must be searched to find a lens.
-Older ground-based surveys yield a lens per roughly 10~square degree
-of survey area.  By the large surveys starting now
-(DES\footnote{\tt http://pan-starrs.ifa.hawaii.edu} or
-PanStarrs\footnote{\tt http://www.darkenergysurvey.org}) the expected
-yield is something in between.  Given their unprecedented survey area,
-these are likely to yield thousands of lenses.  Over the 2020s,
-LSST\footnote{\tt http://www.lsst.org/} can be expected to give us ten
-thousand lenses.
-
-The question now arises: can lenses be found and modeled
-automatically in large surveys?  Work using software robots
-\citep{2009ApJ...694..924M} gives some interesting and unexpected
-results.  In clean lensing system in uncrowded fields with high
-signal-to-noise, the robots do very well.  In most situations,
-however, robots miss lenses (low completeness) or contaminate the
-results with non-lenses (low purity).  Robots can be made to
-prioritize completeness or purity, but they cannot deliver both.
-
-In response to the lessons learned from automated lens searches, the
-{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org}
- has been launched.  {\em Spacewarps\/} is
-part of the {\em Zooniverse\/} family of citizen-science
-projects,\footnote{\tt http://www.zooniverse.org} where members of the
-public are invited to analyze different kinds of scientific data which
-are too difficult for robots and too large for specialists.  In {\em
- Spacewarps\/} itself, survey sky images are presented for volunteers to identify lens candidates.  The first tranche of survey data, introduced from May 2013, is an area of $\sim172$ square
-degrees (or a quarter of a percent of the sky) from the Canada-France-Hawaii Telescope Legacy Survey.
- The survey is divided into patches of $440\times 440$ pixels, each of
-which is seen by ten volunteers.  Simulated lenses are mixed in with
-the data, both to help train volunteers on what to look for, and to
-estimate completeness and purity.
-These data had previously been analyzed by robots \cite{Gavazzi2012, More2012ApJ}.  
-  A list of candidates is being
-processed, but intermediate results show some very good candidates, as
-well as re-discoveries of candidates previously found robotically.
-
-These encouraging results now raise the question: could modeling of
-the lenses also be done by volunteers?  Modeling is a much more
-difficult task than searching for lens candidates, as it requires some
-expert knowledge.  Nonetheless, a panel of the {\em Spacewarps\/}
-volunteers is quite experienced from earlier projects, having spent a
-thousand hours or more with data.  Some of these experienced
-volunteers are very interested in more demanding projects.  This also
-is a general trend \citep[cf.][]{Khatib22112011}.  The present project
-then suggests itself.
-
-The purpose of this study was to provide a means to model a large number of gravitational lenses by showing that gravitational lens modeling can be learned / done by volunteers.
-We suggest, that volunteers will be as successful as professionals with modeling if provided with an easy to use tool with visual feedback (What you see is what you get, WYSIWYG) and a minimal set of instructions.
-Volunteers will then crowd work (using buzz word here ;) ) / work collaborative on modeling lenses from several sources / groups at a central place.
-Since this is an iterative learning process, the more involved volunteers will quickly gain knowledge that can be passed down to new volunteers.
-That creates a social structure that scales well with the number of volunteers, as other projects\needcite have already shown.
-Finding people working as volunteers has been shown to be successful last but not least by \sw and the whole galaxy zoo project.
-To test the people's abilities, we investigated the performance of a few volunteers modeling a set of simulated lenses.
-We tested the ability to correctly identify lensed images and reproduce similar mass map of the lens.
-
-\section{A lens modeling program}
-
-It is well known that making a scientific contribution is a central
-motivation cited by citizen-science volunteers.  So for a lens
-modeler in a citizen-science project, nobody wants that it should
-sacrifice scientific usefulness in order to be fun to use.  That said,
-being aesthetically pleasing and having a short initial learning curve
-are also essential qualities.  Also desirable is that the user is
-encouraged/challenged to go deeper and wider into the subject.
-Yet another aspect is enabling small incremental contributions
-by different people.  \spl tries to address all of these.
-
-\subsection{Theory: Fermat's Principle} \label{sec:Fermat}
-
-There are several ways to understand the formation of arcs and
-multiple images in gravitational lensing.  We will follow some ideas
-originally introduced by \cite{1986ApJ...310..568B}, based on Fermat's
-principle.  The key to this approach is an abstract construct called
-the {\em arrival-time surface.}  This surface cannot itself be
-observed, but several observable quantities can be derived from it.
-
-Consider a gravitational lens.  As in most astrophysical lensing, this
-lens is `thin' along the line of sight, and effectively lies on a
-plane.  Let $x,y$ be coordinates on this plane.  These coordinates
-measure ordinary physical lengths (meters, etc) on the lens plane.
-Let there be a point source of light at $x=0,y=0$ but far behind the
-lens plane.  Now imagine a light ray coming from the source to the
-lens plane $(x,y)$ and then changing its path to reach the observer.
-The observer would see the source apparently behind $(x,y)$ the
-lens.  If this happened, the light ray would have taken a longer path
-than having come directly to the observer.  The geometrical path
-difference would be $\propto x^2 + y^2$ for small angles.  Let us
-define
-\begin{equation} \label{eq:Ageom}
-A_{\rm geom}(x,y) = \half(x^2 + y^2) \,.
-\end{equation}
-As written, this looks an area.  But we can also think of it as a
-time delay introduced by the light ray bending --- with an unwritten
-constant factor.  That factor is basically the lens distance divided by the
-speed of light; the precise expression has to take the expansion of
-the Universe into account, and is given in the Appendix.  The top part
-of \figref{arriv} shows what $A_{\rm geom}$ looks like: it is
-just a paraboloid.  The minimum of the surface is at the source, and
-that is where, according to Fermat's principle, the image will be.
-The rest of the arrival-time surface is just an imaginary surface.
-
-\begin{figure}
-\centering
-\includegraphics[height=.3\vsize]{fig/arriv1}
-\includegraphics[height=.3\vsize]{fig/arriv2}
-\includegraphics[height=.3\vsize]{fig/arriv3}
-\caption{Arrival-time surfaces, with contour levels of equal arrival
-  time.  The upper surface is with no lens.  The middle surface is
-  when a circular lensing mass (offset from the source) is added; a
-  maximum, a minimum and a saddle point can be seen.  The last surface
-  is the result of an elongated lensing mass; a maximum, two minima
-  and two saddle points can be seen.
-}
-\label{fig:arriv}
-\end{figure}
-
-The geometrical time delay \eqref{eq:Ageom} is not the only one.  The
-warping of space by a gravitational field introduces a further time
-delay $A_{\rm grav}$.  The arrival time of the light ray, compared to
-what it would have been with no lensing, is
-\begin{equation} \label{eq:Aarriv}
-\hbox{arrival time} = A_{\rm geom} + A_{\rm grav} \,.
-\end{equation}
-We can also express $A_{\rm grav}$ as an area, with the same implicit
-constant factor that makes it a time.  The expression, however, is
-more complicated.  We first introduce the notation $\nabla^2 f(x,y)$
-to denote
-\begin{equation}
- \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
-        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
-      {\Delta x \; \Delta y}
-\end{equation}
-in the limit of $\Delta x,\Delta y$ small.  Then we write the
-projected mass density of the lens, which would be in $\rm
-kg\,m^{-2}$, in special units (given in the Appendix) such that the
-surface density becomes a dimensionless field $\kappa(x,y)$.  Then
-\begin{equation} \label{eq:Poisson}
-\nabla^2 A_{\rm grav}(x,y) = -2\kappa(x,y) \,.
-\end{equation}
-Note that $A_{\rm grav}(x,y)$ is not given explicitly in terms of
-$\kappa(x,y)$, but as a differential equation that must be solved.
-Equations of the type \eqref{eq:Poisson} are, however, well known (as
-Poisson's equation in two dimensions), and there are techniques for
-solving them.  Incidentally, $\kappa$ has a second meaning, as well as
-being a dimensionless form of the projected density: it also
-measures how a bundle of light rays is brought together by
-gravitational lensing, and is called the {\em convergence}.
-
-What is the effect of mass on the arrival-time surface?  If the mass
-distribution $\kappa(x,y)$ is circular, $A_{\rm grav}(x,y)$ will have
-a peak at the center of that circle.  But if the mass is not precisely
-in front of the source, the arrival-time surface will not be circular
-any more.  The result is illustrated in the middle part of
-\figref{arriv}.  In addition to the maximum, there is now a
-minimum and a saddle point.  According to Fermat's principle, images
-appear at maximum and saddle points as well as minimum.  If the mass
-distribution is non-circular, more images can appear.  The lower
-example in \figref{arriv} shows the arrival-time surface due
-to an elongated mass.  There is still a maximum, part of an elongated
-hill, and around it there are two minima and two saddle points.
-
-So far we have considered a single point source.  What if there is an
-extended source?  To understand this, let us consider what happens if
-we move the original point source slightly, or equivalently, keep the
-original point source behind $x=0,y=0$ and move the lens slightly.
-The contours of constant arrival time will naturally move slightly,
-and so will the images.  The movement of the contours will be most
-noticeable where the contours are far apart, that is where the
-arrival-time surface is nearly flat.  As is evident from
-\figref{arriv}, this is the region where the minimum and saddle
-points lie, or near the images.  In this region, points on the source
-that are close together produce images that are comparatively far
-apart.  In other words, the image is highly magnified.  In summary,
-lower curvature in the arrival-time surface for a point source implies
-larger magnification of an extended source.  Conversely, where the
-arrival-time surface is strongly curved, the image will be
-demagnified.  We see from \figref{arriv} that the arrival-time
-surface tends to be highly curved near the maximum.  Hence maximum tend
-to be demagnified.  In practice, maxima of the arrival time are nearly
-always too faint to see. The minima and saddle points dominate.
-
-In the lensed examples in \figref{arriv}, the maximum in each
-case is evident, but how can we distinguish the minima from the saddle
-points?  The key is to examine the contours of equal arrival time.
-Contour curves loop around a minimum or a maximum, but near a saddle
-point, the contours do not loop around, instead they approach and then
-pull away.  The contour curve precisely level with a saddle point
-forms an X at the saddle point.  In \figref{arriv}, one such X
-is evident.  In other cases the X contour passing through the saddle
-point is not itself plotted, but we can tell by the other contours where
-the X would be.  These self-crossing contours will play a special role
-in the following.
-
-\subsection{\spl} \label{sec:SpaghettiLens}
-
-\spl implements the ideas of the arrival-time surface into a modeling
-program.
-
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/006941_input}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007022_input}} \\
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/006919_input}}
-  \caption{Examples of Spaghetti input.  The models from these appear
-    later in Figures \ref{fig:6941}, \ref{fig:7022} and \ref{fig:6919}.
-  }
-  \label{fig:input-spag}
-\end{figure}
-
-The first modeling step with \spl is for the user to call up the \sw
-of interest, and make an educated guess for the topography of the
-arrival-time surface --- specifically, the locations of the maxima,
-minima and saddle points that correspond to images of the brightest
-part of the source.  Input is given by tracing a proposal for the
-saddle-point contours.  \Figref{input-spag} shows some example
-inputs --- and incidentally also indicates the origin of the name
-\spl.
-
-\spl, as implemented so far, assumes that the lens is dominated by a
-single galaxy, and that the center of that galaxy is the sole maximum.
-Once the maximum has been identified, we wish to characterize the rest
-of the arrival-time surface in relation to it.  From geometry, any
-surface that has a central maximum and goes high at the edges must
-have a lima\c con-like contour, meaning a two-looped curve with one
-loop inside the other.  The maximum lies within the inner loop, a
-minimum lies between the two loops, and a saddle point lies at the
-self-intersection of the contour.  The top panel of
-\figref{input-spag} shows an example: red marks a suggested
-maximum, green a saddle point, and blue a minimum.  (The small pink
-dots are just help sketch the proposed contour.)  The middle and lower
-panels of \figref{input-spag} show a more complex situation.
-The minimum has split into three images: two new minima with a saddle
-point in between.  In this situation the contour through the new
-saddle point forms a figure of eight, its two loops enclosing the two
-new minima.  The configuration of a lima\c con, possibly with a figure
-of eight inside, suffices to give a preliminary description of the
-arrival-time surface for nearly all cases where the lens is dominated
-by one galaxy, and the source is a single small object.
-
-The exact placement of the pink spaghetti loops shown in
-\Figref{input-spag} has no significance.  Their utility is
-just to help mark up the image as maximum, minima and saddle points.
-As the figure suggests, the marking up can be done very easily with a
-mouse.  Human intuition is required to (a)~identify lensed images and
-separate them from other background light sources, and (b)~classify
-and order images according to arrival time.  But the computer helps by
-allowing only valid lensing configurations to be entered, and ensuring
-that the odd image theorem\needcite is taken care of.
-
-In the second modeling step, once the user has sketched a proposed
-spaghetti configuration, \spl sends the input to its server-side
-modeling engine, called GLASS.  The task of GLASS is to find a mass
-distribution $\kappa(x,y)$ that exactly reproduces the locations of
-the maximum, minima and saddle points. Now, this criterion by itself
-is extremely under-determined --- there are infinitely many mass
-distributions that will reproduce a given set of maxima, minima and
-saddle points, but typically they (a)~produce lots of extra images,
-and (b)~look very unlike galaxies.  Additional assumptions (a prior)
-are necessary.  GLASS uses a prior based on suggestions by
-\cite{1997MNRAS.292..148S}.
-\begin{enumerate}
-\item The mass distribution is built out of non-negative tiles of
-  mass.  (Sometimes these tiles are called mass pixels, but we should
-  emphasize that they are unrelated to image pixels, and are much
-  larger.)
-\item There is a notional lens center, say $(x_0,y_0)$ which is
-  identified with the maximum of the arrival time.  The source can
-  have an arbitrary offset with respect to the lens center.
-\item The mass distribution must be centrally concentrated, in two
-  respects.  First, the circularly averaged density must fall away
-  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
-  Second, the direction of increasing density at any $(x,y)$ can point
-  at most $45^\circ$ away from $(x_0,y_0)$.
-\item The lens must be symmetrical with respect to $180^\circ$ rotations
-  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
-  user wishes.
-\end{enumerate}
-There are still infinitely many models that satisfy both data and
-prior constraints, but now they are more credible as galaxy lenses.
-It is then possible to generate an ensemble of models.  The sampling
-technique used by GLASS is described in \citep{Lubini2012}, and
-improves upon earlier techniques \citep{2000AJ....119..439W,Saha2004}.
-Typically, ensembles of 200 models are used.  That is to say, what we
-call a \spl model is really the mean of an ensemble of 200 models, and
-its estimated uncertainty is the range covered by the whole ensemble.
-
-In the third modeling step, having generated a model ensemble, \spl
-post-processes it to present results and diagnostics to the user for
-inspection. This takes the form of three figures.
-\begin{enumerate}
-\item A gray scale plus contour map of the mass distribution.
-\item A contour map of the arrival-time surface.
-\item A synthetic image of the lensed features.
-\end{enumerate}
-After examining this feedback, the user can archive the results for
-discussion, or modify their input and try again, or discard the
-attempt altogether.
-
-The fourth modeling step is discussion among modelers and iteration
-on the model.  Any archived model can be revised by another user: they
-can modify the spaghetti configuration slightly or drastically, or
-change options like the size of the mass tiles.  Particularly
-interesting lens candidates lead to trees of models in this way.
-Forum discussion then prunes the tree, focusing attention on one or a
-few models.
-
-\section{A lens modeling challenge} \label{sec:mod_challenge}
-
-% \subsection{\sw} \label{sec:SpaceWarps} some words about \sw
-
-Interested volunteers from the \sw forum were initially introduced to
-\spl through a video tutorial and by videocon.  After this
-introductory stage, a modeling challenge was presented.  This
-consisted of 29 simulated lenses (sims) covering a range of lensing
-configurations.
-
-The \sw sims were generated by AM, in consultation with PM and AV.
-To estimate the performance of the volunteers and the quality of the generated models, two analysis  were done.
-The first analysis tested the correct identification and ordering of lensed images.
-The second one compared the mass distribution of the lens $\kappa(x, y)$ of the generated models to the mass distribution of the simulations.
-
-\subsection{The simulated lenses} \label{sec:sims}
-
-For the sake of blind testing, the information in this section was
-revealed neither to RK, while choosing the challenge set of 29 sims,
-nor to the modelers (volunteers EB, CC, CM, JO, JW and `expert
-modeler' PS) until the modeling stage was done.
-
-The sims were produced using {\tt gravlens}
-\citep{2001astro.ph..2341K,2001astro.ph..2340K} using the CFHTLS survey data
-and catalogues \cite{Coupon2009}. They were of three
-kinds, as follows.
-
-\begin{enumerate}
-  \item Imitating lensed quasars: having a singular elliptical
-    isothermal lens (SIE) plus constant external shear, and a circular
-    Gaussian source.
-  \item Emulating lensed galaxies: similar to the above, but with an
-    elliptical de Vaucouleurs source.
-  \item Resembling cluster lenses: having a source as above, but a
-    more complicated lens, with one dominant elliptical SIE and
-    one or more perturbing elliptical SIEs, plus a circular NFW
-    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
-    the underlying dark matter distribution.
-\end{enumerate}
-
-Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
-lenses follow equations (33--35) of that work, with core radius set to
-zero.  The NFW lens is in equations (48) and (50), while shear is the
-$\gamma$ term in equation (76).
-
-\subsection{Some example models} \label{sec:example_models}
-
-The modelers proposed a total of 129 models for the 29 sims in the
-challenge.  Figures \ref{fig:6941} to \ref{fig:6919} show eight of
-models in some detail.
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6941_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6941_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6941_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006941_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6941_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6941_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6941_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006941_arr_time}
-  }
-  \caption[result 6941 (ASW000102p)]{Sim ASW000102p and a model.  A
-    simple configuration; the model input is shown in the top panel of
-    Figure~\ref{fig:input-spag}.  For details of the individual panels
-    here, see section \ref{sec:example_models}.}
-  \label{fig:6941}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6975_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6975_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW000195x_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6975_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006975_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6975_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006975_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6975_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006975_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6975_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006975_arr_time}
-  }
-  \caption[result 6975 (ASW000195x)]{Sim ASW000195x and a
-    model.}
-  \label{fig:6975}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6937_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6937_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000vqg_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6937_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006937_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6937_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006937_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6937_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006937_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6937_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006937_arr_time}
-  }
-  \caption[result 6937 (ASW0000vqg)]{Sim ASW0000vqg and a model.}
-  \label{fig:6937}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6990_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6990_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0004oux_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6990_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006990_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6990_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006990_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6990_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006990_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6990_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006990_arr_time}
-  }
-
-  \caption[result 6990 (ASW0004oux)]{Sim ASW0004oux and a model.}
-  \label{fig:6990}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6915_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6915_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0001hpf_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6915_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006915_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6915_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006915_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6915_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006915_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6915_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006915_arr_time}
-  }
-
-  \caption[result 6915 (ASW0001hpf)]{Sim ASW0001hpf and a model. Inclined quad.}
-  \label{fig:6915}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:7022_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:7022_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:7022_mass}
-    \includegraphics[width=0.45\textwidth]{fig/007022_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:7022_cont}
-    \includegraphics[width=0.45\textwidth]{fig/007022_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:7022_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/007022_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:7022_atime}
-    \includegraphics[width=0.45\textwidth]{fig/007022_arr_time}
-  }
-
-  \caption[result 7022 (ASW0000h2m)]{Sim ASW0000h2m and a model.  The
-    model input appears in the middle panel of Figure
-    \ref{fig:input-spag}.}
-  \label{fig:7022}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:7025_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:7025_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0000h2m_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:7025_mass}
-    \includegraphics[width=0.45\textwidth]{fig/007025_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:7025_cont}
-    \includegraphics[width=0.45\textwidth]{fig/007025_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:7025_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/007025_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:7025_atime}
-    \includegraphics[width=0.45\textwidth]{fig/007025_arr_time}
-  }
-
-  \caption[result 7025 (ASW0000h2m)]{Sim ASW0000h2m and another model.  The image parities in this case were incorrectly identified.}
-  \label{fig:7025}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \subfigure[real mass distribution]{
-    \label{fig:6919_sim_mass}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_kappa}
-  }
-  \subfigure[real arrival-time surface]{
-    \label{fig:6919_sim_arr}
-    \includegraphics[width=0.45\textwidth]{fig/ASW0002z6f_arriv}
-  }
-  \subfigure[model mass distribution]{
-    \label{fig:6919_mass}
-    \includegraphics[width=0.45\textwidth]{fig/006919_mass}
-  }
-  \subfigure[model arrival-time surface]{
-    \label{fig:6919_cont}
-    \includegraphics[width=0.45\textwidth]{fig/006919_spaghetti}
-  }
-  \subfigure[real vs model enclosed mass]{
-    \label{fig:6919_kappa}
-    \includegraphics[width=0.45\textwidth]{fig/006919_kappa_encl}
-  }
-  \subfigure[model lensed image]{
-    \label{fig:6919_atime}
-    \includegraphics[width=0.45\textwidth]{fig/006919_arr_time}
-  }
-  \caption[result 6919 (ASW0002z6f)]{Sim ASW0002z6f and a model.  The
-    model input appears in the bottom panel of Figure
-    \ref{fig:input-spag}. An incipient quad}
-  \label{fig:6919}
-\end{figure}
-
-Figure \ref{fig:6941} shows a simple example.  The top two panels show
-contours of convergence $\kappa(x,y)$, and of the arrival-time for a
-point source at the brightest point of the sim
-source.\footnote{Although the arrival-time contours represent an
-  abstract quantity that is not directly observable, the contour map
-  may actually resemble the appearance of the lensed arcs.  The
-  resemblance is due to a serendipitous computer-graphical effect
-  \citep{2001AJ....122..585S}.} The spatial scale is in pixels.  The
-information in these two panels was, of course, kept secret during the
-modeling challenge.  The panels in the middle row and at the lower
-right are the results that \spl returns to the modeler, in order for
-them to assess the model.  These derive from the mean of an ensemble
-of 200 models generated by \spl.  The model mass distribution (left mid) is
-returned as a contour map superimposed on an intensity map. A fairly
-smooth mass distribution, as here, is a good sign.  An irregular or
-checkboard pattern in the mass map usually indicates a bad model.  The
-models arrival-time contours are shown in the right middle panel.
-  The special contour passing through a
-saddle point is shown in black: this is the model version of the
-spaghetti sketch provided by the user.  Also provided is a synthetic
-image of the lensed source in the bottom right panel.
-  For this, \spl assumes a simple conical
-source profile.  The user can change the contrast level on the image,
-which (though it is not saved) amounts to adjusting the width of
-the cone.  Finally, to the lower left of the figure, we have a
-comparison of the input and recovered mass profiles.  The panel shows
-a average $\kappa$ within a given radius, as a function of radius.
-The red curve is the true value, and where it crosses unity (dotted
-horizontal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
-  The two blue curves
-are the minimal and maximal mean enclosed $\kappa$ from the internal
-ensemble in \spl.  The region between the blue curves is shaded
-between the radii of the innermost and the outermost images --- this
-is the confidence region from the modeling.  As we see, the shaded
-blue is slightly above the red curve. 
-The Einstein Radius $\Theta_\text{E}$ of the model is estimated crossing the
-mean enclosed $\kappa$ (not plotted) with unity.
-  In summary: the identification
-of minimum and saddle point is correct, but the estimated Einstein
-radius is a little too high.
-
-Figure \ref{fig:6975} shows a lens with substructure in the form of a
-smaller secondary galaxy.  The galaxies in such group or cluster sims
-were, in fact, based on galaxies visible in the images --- but the
-modelers were not told in advance whether this was the case.  The
-model does not include any substructure, but otherwise is not bad.
-The minimum and saddle point are correct, and the Einstein radius is
-only a little underestimated.
-
-Figure \ref{fig:6937} shows a case where substructure leads to a
-poor model.
-
-Figure \ref{fig:6990} shows an example of an arc that has split into
-three images.  This kind of configuration, with a counter-image close
-to the lensing galaxy and a more distant arc/triplet on the other
-side, generically arises from an elongated mass distribution when the
-source is displaced along the elongated direction.
-
-Figure \ref{fig:6915} shows another quad.  This kind of configuration
-arises when the mass is elongated and the source is displaced at an
-angle to the elongation.
-
-Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
-saddle points are correctly identified, and the orientation of the
-ellipticity of the mass distribution is correctly reproduced.  The
-Einstein radius is somewhat overestimated.  Figure \ref{fig:7022}
-shows another model of the same system.  In this one, the
-identification of the minima and saddle points was incorrect, and mass
-distribution comes out elongated East-West instead of North-South.
-The mass distribution also appears somewhat jagged and the
-saddle-point contours are not as clean as in the previous examples;
-these are often indicators of a problem with the model.  The enclosed
-mass is, however, none the worse --- the reason is probably that in a
-relatively symmetrical image configuration, the Einstein radius is quite
-well constrained by the images in a fairly model-independent way.
-
-To conclude this set of examples, Figure \ref{fig:6919} shows another
-type of quad.  Actually, in this case the brightest part of the source
-is only doubly imaged, but the source extends into a region that
-produces four images.  We can also see from the real arrival-time
-surface that a point source is a double on the verge of splitting into
-a quad.  The modeler interpreted the system as a quad.  The
-appearance of the arcs, shown in the bottom panel of Figure
-\ref{fig:input-spag}, looks like an arc and counter-image such as
-discussed with Figure \ref{fig:6990} above.  But there is an important
-difference: the long arc is closer to the galaxy, as if the arc and
-counter-image have swapped roles.  This configuration arises if the
-source displacement is perpendicular to the long axis of the lensing
-mass.
-
 \subsection{Test of image identification} \label{sec:tests.t1}
@@
 
-\begin{figure}[htbp]
+\begin{figure*}
   \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_4}
+    \includegraphics[width=0.90\linewidth]{fig/eR_5}
   \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
   \label{fig:ER_per_sim}
-\end{figure}
+\end{figure*}
 
@@
   \centering
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007020_kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
-             {fig/007024_kappa_encl}}
-  \subfigure{\includegraphics[width=0.45\textwidth]
+  \subfigure{\includegraphics[width=0.9\linewidth]
+             {fig/007022_kappa_encl}}
+  \subfigure{\includegraphics[width=0.9\linewidth]
+             {fig/007025_kappa_encl}}
+  \subfigure{\includegraphics[width=0.9\linewidth]
              {fig/007021_kappa_encl}} \\
-  \caption{Enclosed mass \kenc for models of \asw{0h2m} with incorrect image identifications.
+  \caption{
+  %Enclosed mass \kenc from three models of a system with incorrect image identifications.
 \figref{7022} shows a model (by an expert) with the correct identifications, while \figref{7025} shows a model with error E06 (parities rotated by $\pi/2$).
-Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).}
+Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).
+  }
 \label{fig:kapenc_compare_faulty}
@@
 
+\section{Acknowledgments}
+
+We thank the Swiss Society for Astrophysics and Astronomy and the
+Swiss Academy of Sciences.
+
 \clearpage
-
-\appendix
-
-\section{More on Lensing Theory} \label{sec:more-theory}
-
-In \secref{Fermat}, for the sake of a more intuitive
-explanation, we suppressed some constant factors in
-equations \eqref{eq:Ageom}, \eqref{eq:Aarriv} and \eqref{eq:Poisson}.
-Here we fill in the details.
-
-First let us recall various distance formulas. Comoving distances
-involve integrals of the type
-\begin{equation}
-\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
-%\raise.5ex\hbox{.}
-\end{equation}
-Specifically, the comoving distance $D_S$ from observer to source,
-$D_{LS}$ from lens to source, and $D_L$ from observer to lens are
-given by the above integral, with limits as follows.
-\begin{equation}
-\begin{aligned}
-&D_S    &                                &\int_0^{z_S} \\
-\noalign{\medskip}
-&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
-\noalign{\medskip}
-&D_L    &                                &\int_0^{z_L}
-\end{aligned}
-\end{equation}
-Angular-diameter distances $d$ can be related to comoving distances
-$D$ by scale factors, as follows.
-\begin{equation}
-\begin{aligned}
-D_S &= (1+z_S) \, d_S \\
-D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
-D_L &= (1+z_L) \, d_L
-\end{aligned}
-\end{equation}
-Consequently, angular-diameter distances are always smaller than
-comoving distances.  Moreover, while comoving distances can be simply
-added, angular distances cannot:
-\begin{equation}
-d_S \neq d_L + d_{LS}
-\end{equation}
-
-Locations on the lens plane in angular coordinates on the sky, are
-expressed as
-\begin{equation}
-(x,y) = d_L (\theta_x,\theta_y)
-\end{equation}
-
-The $A$ areas (with the implicit
-constant factor that makes them time delays) and the $\kappa$ density are related to physical arrival
-time $t$ and density $\Sigma$ as
-\begin{equation}
-\begin{aligned}
-A           &= \frac{c}{(1+z_L)^2} \frac{D_L D_{LS}}{D_S} \times t \\
-\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L D_{LS}}{D_S}
-               \times \Sigma(x,y)
-\end{aligned}
-\end{equation}
-Letting the source be offset at $(s_x,s_y)$ rather than at the
-origin, we have
-\begin{equation}
-\begin{aligned}
-t_{\rm geom} &= \frac{(1+z_L)^2}{2c} \frac{D_S}{D_L D_{LS}}
-\left( (x-s_x)^2 + (y-s_y)^2 \right) \\
-\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
-\end{aligned}
-\end{equation}
-We can now compare with equations (2.1) to (2.6)
-from \cite{1986ApJ...310..568B}.
-%% comparison to be explicited
-
-\section{Miscellaneous}
-
-Maybe discard \figref{ER_all_models}, since the essential information is there in \figref{ER_per_sim}.
-
-\begin{figure}[htbp]
-  \centering
-    \includegraphics[width=0.80\textwidth]{fig/eR_1}
-  \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
-  \label{fig:ER_all_models}
-\end{figure}
-
-\newpage
-
-\bibliographystyle{apj}
-\bibliography{bib/ms}
-
-\end{document}

### 2014-06-05 text/partC.tex
partA partB partC

\appendix

\section{Relation to standard lensing formalism} \label{sec:more-theory}

The presentation of the arrival-time surface in \secref{Fermat}
omitted details, for the sake of a more intuitive explanation.
Specifically, the geometric time delay $t_{\rm geom}$ and the
convergence $\kappa$ were left as proportionalities (equations
\ref{eq:tgeom} and \ref{eq:kappa}), and the gravitational time delay
$t_{\rm grav}$ was given in an implicit form (equation
\ref{eq:tgrav}).  We now fill in the details.

The original formulation of the arrival-time surface appears in
equations (2.1) to (2.6) of \cite{1986ApJ...310..568B}.  Their
equations can be rearranged as follows.
\begin{equation} \label{cfBN86}
\begin{aligned}
t_{\rm geom} &= \frac{(1+z_L)}{2c} \frac{d_S}{d_L d_{LS}}
\left[ (x-s_x)^2 + (y-s_y)^2 \right] \\
\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y) \\
\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{d_L d_{LS}}{d_S}
               \times \Sigma(x,y)
\end{aligned}
\end{equation}
We symbols $d_L,d_S$ and $d_{LS}$ are angular-diameter distances,
respectively from observer to lens, observer to source, and lens to
source.  In the concordance cosmology, we have
\begin{equation}
d_{LS} = \frac{c}{H_0}\,\frac1{1+z_S} \,
\int_{z_L}^{z_S} \!\! \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
%\raise.5ex\hbox{.}
\end{equation}
and similarly $d_S$ and $d_{LS}$.  We have replaced angular positions
on the sky with positions on the lens plane as
\begin{equation}
(x,y) = d_L (\theta_x,\theta_y) \,.
\end{equation}

The first line of equation \eqref{cfBN86} is $t_{\rm geom}$ from
equation \eqref{eq:tgeom} with the proportionality filled in, and with
the source offset at $(s_x,s_y)$ rather than at the origin.

The last line of equation \eqref{cfBN86} fills in the proportionality
factor in equation \eqref{eq:kappa} for the convergence (or
dimensionless surface density) $\kappa$.

The middle line of equation \eqref{cfBN86}, though very different in
form to the expression \eqref{eq:tgrav} for the gravitational time
delay, is in fact equivalent.  This can be shown by considering a
multipole expansions for the solution of the latter equation.  Another
way to see the equivalence is to write a discrete version of
\eqref{eq:tgrav}.  Introducing a small distance $\Delta$, we write
\begin{equation} \label{eq:relax}
\begin{aligned}
\tgrav(x,y) &=
\frac14 \Big[ \tgrav(x+\Delta,y) + \tgrav(x-\Delta,y) + {}\\
&\kern24pt    \tgrav(x,y+\Delta) + \tgrav(x,y-\Delta) \Big] \\
&+ (1+z_L)\frac{2G}{c^3} \, \pi\Delta^2 \, \Sigma(x,y) \,.
\end{aligned}
\end{equation}
The circular average $\langle \tgrav^{\textstyle\circ} \rangle$ has
been replaced by the average over four neighbouring points, and the
enclosed mass $M_{\textstyle\bullet}$ has been replaced by
$\pi\Delta^2\,\Sigma(x,y)$.  Now let us read equation \eqref{eq:relax}
as an iterative formula for $t_{\rm grav}(x,y)$.  It is then
recognisable as a formula for solving the two-dimensional Poisson
equation from \eqref{cfBN86} by relaxation.

\newpage

### 2014-06-05 text/partA.tex
edits

@@
 
+A challenge set of 29 was selected, representing the different image
+morphologies among the spacewarps sims.  Modellers then contributed
+129 models for these sims (at least for each sim).  Models were
+collecting in the same forum used to model real candidate lenses, and
+under the same conditions --- modellers were free to consult and
+refine each other's models, but had no information on how the sims
+were generated.
+
 \subsection{The simulated lenses} \label{sec:sims}
-
-For the sake of blind testing, the information in this section was
-revealed neither to RK, while choosing the challenge set of 29 sims,
-nor to the modelers (volunteers EB, CC, CM, JO, JW and `expert
-modeler' PS) until the modeling stage was done.
 
@@
 
+The information in this section was not revealed to the person
+choosing the challenge set (RK) or to the modellers (EB, CC, CM, PS,
+JO and JW) until the modeling stage was done.
+
 \subsection{Some example models} \label{sec:example_models}
+
+The community of modellers proposed a total of 129 models for the 29 sims in the challenge.
+Typically XX models were produced per simulation, with XX systems having only XX models (corresponding tot he most simplest simulations) to XX having XX models (a complex and contrversial lens system).
+In the following we discuss eight example lenses in detail with individual notes per system and corresponding plots are shown in figures 4-11 (or whatever compact version you end up making).
+These eight examples were chosen on the basis of …??
 
@@
 
-% Rafi's original one-coumn design:
-%
-% \begin{figure}
-%   \centering
-%   \subfigure[real mass distribution]{
-%     \label{fig:6941_sim_mass}
-%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_kappa}
-%   }
-%   \subfigure[real arrival-time surface]{
-%     \label{fig:6941_sim_arr}
-%     \includegraphics[width=0.45\textwidth]{fig/ASW000102p_arriv}
-%   }
-%   \subfigure[model mass distribution]{
-%     \label{fig:6941_mass}
-%     \includegraphics[width=0.45\textwidth]{fig/006941_mass}
-%   }
-%   \subfigure[model arrival-time surface]{
-%     \label{fig:6941_cont}
-%     \includegraphics[width=0.45\textwidth]{fig/006941_spaghetti}
-%   }
-%   \subfigure[real vs model enclosed mass]{
-%     \label{fig:6941_kappa}
-%     \includegraphics[width=0.45\textwidth]{fig/006941_kappa_encl}
-%   }
-%   \subfigure[model lensed image]{
-%     \label{fig:6941_atime}
-%     \includegraphics[width=0.45\textwidth]{fig/006941_arr_time}
-%   }
-%   \caption[result 6941 (ASW000102p)]{Model results for a simple
-%     simulated lens.  (The label ASW000102p in the lower-left panel is
-%     the \sw identifier for the sky patch where this sim was placed.)
-%     The model input is shown in the top panel of
-%     Figure~\ref{fig:input-spag}.  For details of the individual panels
-%     here, see section \ref{sec:example_models}.}
-%   \label{fig:6941}
-% \end{figure}
-
-% Rafik's two row design
-
-\newpage
+\vfill\penalty-2000
 
@@
 
-\newpage
+\vfill\penalty-2000
 
@@
 
-Figure \ref{fig:6975} shows a lens with substructure in the form of a
-smaller secondary galaxy.  The galaxies in such group or cluster sims
-were, in fact, based on galaxies visible in the images --- but the
-modelers were not told in advance whether this was the case.  The
-model does not include any substructure, but otherwise is not bad.
-The minimum and saddle point are correct, and the Einstein radius is
-only a little underestimated.
-
-Figure \ref{fig:6937} shows a case where substructure leads to a
-poor model.
+Figure \ref{fig:6915} shows another quad.  This kind of configuration
+arises when the mass is elongated and the source is displaced at an
+angle to the elongation.
 
@@
 source is displaced along the elongated direction.
-
-Figure \ref{fig:6915} shows another quad.  This kind of configuration
-arises when the mass is elongated and the source is displaced at an
-angle to the elongation.
-
-Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
-saddle points are correctly identified, and the orientation of the
-ellipticity of the mass distribution is correctly reproduced.  The
-Einstein radius is somewhat overestimated.  Figure \ref{fig:7022}
-shows another model of the same system.  In this one, the
-identification of the minima and saddle points was incorrect, and mass
-distribution comes out elongated East-West instead of North-South.
-The mass distribution also appears somewhat jagged and the
-saddle-point contours are not as clean as in the previous examples;
-these are often indicators of a problem with the model.  The enclosed
-mass is, however, none the worse --- the reason is probably that in a
-relatively symmetrical image configuration, the Einstein radius is quite
-well constrained by the images in a fairly model-independent way.
 
@@
 mass.
+
+Figure \ref{fig:6975} shows a lens with substructure in the form of a
+smaller secondary galaxy.  The galaxies in such group or cluster sims
+were, in fact, based on galaxies visible in the images --- but the
+modelers were not told in advance whether this was the case.  The
+model does not include any substructure, but otherwise is not bad.
+The minimum and saddle point are correct, and the Einstein radius is
+only a little underestimated.
+
+Figure \ref{fig:6937} shows a case where substructure leads to a
+poor model.
+
+Figure \ref{fig:7022} shows another model of the same system.  In this
+one, the identification of the minima and saddle points was incorrect,
+and mass distribution comes out elongated East-West instead of
+North-South.  The mass distribution also appears somewhat jagged and
+the saddle-point contours are not as clean as in the previous
+examples; these are often indicators of a problem with the model.  The
+enclosed mass is, however, none the worse --- the reason is probably
+that in a relatively symmetrical image configuration, the Einstein
+radius is quite well constrained by the images in a fairly
+model-independent way.
+
+Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
+saddle points are correctly identified, and the orientation of the
+ellipticity of the mass distribution is correctly reproduced.  The
+Einstein radius is somewhat overestimated.

### 2014-06-05 text/partC.tex
edits

@@
 \end{equation}
-The circular average $\langle \tgrav^{\textstyle\circ} \rangle$ has
+The circular average
+$\left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle$ has
 been replaced by the average over four neighbouring points, and the
-enclosed mass $M_{\textstyle\bullet}$ has been replaced by
+enclosed mass $M(x_\subbullet,y_\subbullet)$ has been replaced by
 $\pi\Delta^2\,\Sigma(x,y)$.  Now let us read equation \eqref{eq:relax}

### 2014-06-06 text/partA.tex
edits

@@
 
-\section{A lens modeling program}
-
 It is well known that making a scientific contribution is a central
@@
 by different people.  \spl tries to address all of these.
+
+\section{A lens modeling program}
+
+\spl is an extension of the GLASS framework for modelling lenses
+\citep{2014arXiv1401.7990C}, and introduces two main innovations.
 
@@
 
-In the lensed examples in \figref{arriv}, the maximum in each
-case is evident, but how can we distinguish the minima from the saddle
-points?  The key is to examine the contours of equal arrival time.
-Contour curves loop around a minimum or a maximum, but near a saddle
-point, the contours do not loop around, instead they approach and then
-pull away.  The contour curve precisely level with a saddle point
-forms an X at the saddle point.  In \figref{arriv}, one such X
-is evident.  In other cases the X contour passing through the saddle
-point is not itself plotted, but we can tell by the other contours where
-the X would be.  These self-crossing contours will play a special role
-in the following.
+In the lensed examples in \figref{arriv}, the maximum in each case is
+evident, but how can we distinguish the minima from the saddle points?
+The key is to examine the contours of equal arrival time.  Contour
+curves loop around a minimum or a maximum, but near a saddle point,
+the contours do not loop around, instead they approach and then pull
+away.  The contour curve precisely level with a saddle point forms an
+X at the saddle point.  In \figref{arriv}, one such X is evident.  In
+other cases the X contour passing through the saddle point is not
+itself plotted, but we can tell by the other contours where the X
+would be.  These self-crossing contours will play a special role in
+the following.
+
+Although the arrival-time contours represent an abstract quantity that
+is not directly observable, the contour map may actually resemble the
+appearance of the lensed arcs.  The resemblance is due to a
+serendipitous computer-graphical effect \citep{2001AJ....122..585S}.
 
@@
 \begin{enumerate}
+\item A synthetic image of the lensed features.
+\item A contour map of the arrival-time surface.
 \item A gray scale plus contour map of the mass distribution.
-\item A contour map of the arrival-time surface.
-\item A synthetic image of the lensed features.
 \end{enumerate}
@@
 
-The community of modellers proposed a total of 129 models for the 29 sims in the challenge.
-Typically XX models were produced per simulation, with XX systems having only XX models (corresponding tot he most simplest simulations) to XX having XX models (a complex and contrversial lens system).
-In the following we discuss eight example lenses in detail with individual notes per system and corresponding plots are shown in figures 4-11 (or whatever compact version you end up making).
-These eight examples were chosen on the basis of …??
+The modellers proposed a total of 129 models for the 29 sims in the
+challenge. Typically XX models were produced per simulation, with XX
+systems having only XX models (corresponding tot he most simplest
+simulations) to XX having XX models (a complex and contrversial lens
+system). In the following we discuss eight example lenses in detail
+with individual notes per system and corresponding plots are shown in
+figures 4-11 (or whatever compact version you end up making). These
+eight examples were chosen on the basis of …??
 
@@
 
-\vfill\penalty-2000
+\FloatBarrier
 
@@
     lensed quasar, and model results.  The left panels derive from the
-    simulation, and the right panels are \sl output.  Details of
+    simulation, and the right panels are \spl output.  Details of
     individual panels are in \ref{sec:example_models}.}
@@
 
-\vfill\penalty-2000
+\FloatBarrier
 
-Figure \ref{fig:6941} shows a simple example.  The top two panels show
-contours of convergence $\kappa(x,y)$, and of the arrival-time for a
-point source at the brightest point of the sim
-source.\footnote{Although the arrival-time contours represent an
-  abstract quantity that is not directly observable, the contour map
-  may actually resemble the appearance of the lensed arcs.  The
-  resemblance is due to a serendipitous computer-graphical effect
-  \citep{2001AJ....122..585S}.} The spatial scale is in pixels.  The
-information in these two panels was, of course, kept secret during the
-modeling challenge.  The panels in the middle row and at the lower
-right are the results that \spl returns to the modeler, in order for
-them to assess the model.  These derive from the mean of an ensemble
-of 200 models generated by \spl.  The model mass distribution (left mid) is
-returned as a contour map superimposed on an intensity map. A fairly
+Figures \ref{fig:6941}--\ref{fig:7022} show details of eight models.
+The first four of these show the most common image morphologies, the
+other four explain some problem cases.  Each of the figures has the
+following layout.
+$$ \begin{matrix}
+\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
+t(x,y)                        &\hbox{model } t(x,y) \\
+\kappa(x,y)                   &\hbox{model } \kappa(x,y)
+\end{matrix} $$
+The marked-up image is a zoom of the lensed image on Spacewarps,
+marked up with one or more spaghetti contours; this is the modeller's
+input to \spl.  The three panels on the right show the graphical
+output returned to the modeller by \spl, listed as (i), (ii), (iii) in
+\S\ref{sec:SpaghettiLens}).  These derive from the mean of an ensemble
+of 200 models generated by \spl.  To generate the synthetic image,
+\spl assumes a simple conical source profile.  The user can change the
+contrast level on the image, which (though it is not saved) amounts to
+adjusting the width of the cone.  These synthetic images are still
+very crude and not useful for assessing models.  The best indicator of
+whether the modelling was successful is the arrival-time surface
+$t(x,y)$, shown as contour maps. Contours passing through saddle
+points are shown in black: these are the model versions of the
+spaghetti sketch provided by the user.  The model $\kappa(x,y)$ is
+shown as a contour map superimposed on an intensity map. A fairly
 smooth mass distribution, as here, is a good sign.  An irregular or
-checkboard pattern in the mass map usually indicates a bad model.  The
-models arrival-time contours are shown in the right middle panel.
-  The special contour passing through a
-saddle point is shown in black: this is the model version of the
-spaghetti sketch provided by the user.  Also provided is a synthetic
-image of the lensed source in the bottom right panel.
-  For this, \spl assumes a simple conical
-source profile.  The user can change the contrast level on the image,
-which (though it is not saved) amounts to adjusting the width of
-the cone.  Finally, to the lower left of the figure, we have a
-comparison of the input and recovered mass profiles.  The panel shows
-a average $\kappa$ within a given radius, as a function of radius.
-The red curve is the true value, and where it crosses unity (dotted
-horizontal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
-  The two blue curves
-are the minimal and maximal mean enclosed $\kappa$ from the internal
-ensemble in \spl.  The region between the blue curves is shaded
-between the radii of the innermost and the outermost images --- this
-is the confidence region from the modeling.  As we see, the shaded
-blue is slightly above the red curve. 
-The Einstein Radius $\Theta_\text{E}$ of the model is estimated crossing the
-mean enclosed $\kappa$ (not plotted) with unity.
-  In summary: the identification
-of minimum and saddle point is correct, but the estimated Einstein
-radius is a little too high.
+checkboard pattern in the mass map usually indicates a bad model.
+
+Figure \ref{fig:6941} shows a simple example.  In summary: the
+identification of minimum and saddle point is correct, but the
+estimated Einstein radius is a little too high.
 
@@
 Einstein radius is somewhat overestimated.
+
+Figure \ref{fig:kapenc_compare_faulty} shows a comparison of input and
+recovered mass profiles.  The panel shows a average $\kappa$ within a
+given radius, as a function of radius.  The red curve is the true
+value, and where it crosses unity (dotted horizontal line) is the
+notional Einstein radius $\Theta_{\text{E, sim}}$.  The two blue
+curves are the minimal and maximal mean enclosed $\kappa$ from the
+internal ensemble in \spl.  The region between the blue curves is
+shaded between the radii of the innermost and the outermost images ---
+this is the confidence region from the modeling.  As we see, the
+shaded blue is slightly above the red curve.  The Einstein Radius
+$\Theta_\text{E}$ of the model is estimated crossing the mean enclosed
+$\kappa$ (not plotted) with unity.
+
+The panel shows
+a average $\kappa$ within a given radius, as a function of radius.
+The red curve is the true value, and where it crosses unity (dotted
+horizontal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
+  The two blue curves
+are the minimal and maximal mean enclosed $\kappa$ from the internal
+ensemble in \spl.  The region between the blue curves is shaded
+between the radii of the innermost and the outermost images --- this
+is the confidence region from the modeling.  As we see, the shaded
+blue is slightly above the red curve. 
+The Einstein Radius $\Theta_\text{E}$ of the model is estimated crossing the
+mean enclosed $\kappa$ (not plotted) with unity.

### 2014-06-06 text/partC.tex
edits

--- 2014-06-05 text/partC.tex
@@
 
-The middle line of equation \eqref{cfBN86}, though very different in
-form to the expression \eqref{eq:tgrav} for the gravitational time
-delay, is in fact equivalent.  This can be shown by considering a
-multipole expansions for the solution of the latter equation.  Another
-way to see the equivalence is to write a discrete version of
+The middle line of equation \eqref{cfBN86} is a Poisson equation for
+the gravitational time delay, and is equivalent to the implicit
+expression \eqref{eq:tgrav}.  One way to verify the equivalence is to
+consider the small circle in equation \eqref{eq:tgrav} as a region
+where $\Sigma$ is constant, and approximate $\tgrav$ by its Taylor
+expansion to $O(x^2,y^2)$.  Substituting in equations \eqref{cfBN86}
+gives the Taylor coefficients in terms of $\Sigma$, and result
+satisfies the expression \eqref{eq:tgrav}.  Alternatively, we can
+proceed with a discrete form of the Poission equation \eqref{cfBN86}.
+Discretising on a grid with spacing $\Delta$, we have
 \eqref{eq:tgrav}.  Introducing a small distance $\Delta$, we write
@@
 \end{equation}
-The circular average $\langle \tgrav^{\textstyle\circ} \rangle$ has
-been replaced by the average over four neighbouring points, and the
-enclosed mass $M_{\textstyle\bullet}$ has been replaced by
-$\pi\Delta^2\,\Sigma(x,y)$.  Now let us read equation \eqref{eq:relax}
-as an iterative formula for $t_{\rm grav}(x,y)$.  It is then
-recognisable as a formula for solving the two-dimensional Poisson
-equation from \eqref{cfBN86} by relaxation.
+This is recognisable as a formula for solving the two-dimensional
+Poisson equation from \eqref{cfBN86} by relaxation.  Let us now
+replace the average over four neighbouring points by the circular
+average $\left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle$ and
+replace $\pi\Delta^2\,\Sigma(x,y)$ by the enclosed mass
+$M(x_\subbullet,y_\subbullet)$.  These replacements are valid in the
+limit of a small grid.  The result is the implicit
+equation \eqref{eq:tgrav}.
 

### 2014-06-06 text/partA.tex
edits

@@
 
-\begin{figure}
-\centering
-\includegraphics[width=0.9\hsize]{fig/lenssky}
-\caption{Sky distribution of 423 published lenses considered secure as
-  of January 2014 (from the Masterlens catalog at the University of
-  Utah, maintained by Joel Brownstein and Leonidas Moustakas.) The map
-  is in Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
-  middle.  The empty swathes to the left and right of center are the
-  Milky Way.}
-\label{fig:masterlens}
-\end{figure}
-
-A good summary of the observational situation is offered by
-Figure~\ref{fig:masterlens}, which shows known secure multiply-imaging
-lenses.  The non-uniformity on the sky is not intrinsic, it just
-indicates the density of deep surveys up to 2013.  At HST resolution,
-of order 1~square degree of sky must be searched to find a lens.
-Older ground-based surveys yield a lens per roughly 10~square degree
-of survey area.  By the large surveys starting now
-(DES\footnote{\tt http://pan-starrs.ifa.hawaii.edu} or
-PanStarrs\footnote{\tt http://www.darkenergysurvey.org}) the expected
-yield is something in between.  Given their unprecedented survey area,
-these are likely to yield thousands of lenses.  Over the 2020s,
-LSST\footnote{\tt http://www.lsst.org/} can be expected to give us ten
-thousand lenses.
-
 The question now arises: can lenses be found and modeled
@@
 encouraged/challenged to go deeper and wider into the subject.
-Yet another aspect is enabling small incremental contributions
-by different people.  \spl tries to address all of these.
+  \spl tries to address all of these.
 
@@
 \spl is an extension of the GLASS framework for modelling lenses
-\citep{2014arXiv1401.7990C}, and introduces two main innovations.
+\citep{2014arXiv1401.7990C}, with two innovations.
+\begin{itemize}
+\item \spl is friendly to the forum style of citizen-science projects,
+  without sacrificing any features of GLASS.  The program is invoked
+  online, and the modelling history is logged in a transparent way,
+  enabling forum discussions and incremental improvement by different
+  people.  Moreover, any result can be revisited and post-processed if
+  desired.  These features are of general interest for citizen-science
+  projects, but somewhat outside the scope of the present paper.
+\item The innovation specific to lens modelling is the use of
+  ``spaghetti diagrams'' or saddle-point contours.  These were
+  originally introduced to gravitational lensing by
+  \cite{1986ApJ...310..568B} as a way of classifying lensed images,
+  and are sometimes shown as part of the output of lens models
+  \citep[for
+    example][]{2001ApJ...557..594R,2003ApJ...590...39K,Lubini2012}.
+  In \spl, however, spaghetti diagrams are the {\em input\/} through
+  which the modeller tells the program what to do.
+\end{itemize}
 
-\subsection{Fermat's Principle and ``spaghetti'' contours} \label{sec:Fermat}
+We now explain spaghetti diagrams and then outline what \spl does with
+them.
 
-There are several ways to understand the formation of arcs and
-multiple images in gravitational lensing.  We will follow some ideas
-originally introduced by \cite{1986ApJ...310..568B}, based on Fermat's
-principle.  The key to this approach is an abstract construct called
-the {\em arrival-time surface.}  This surface cannot itself be
-observed, but several observable quantities can be derived from it.
+\subsection{Fermat's Principle} \label{sec:Fermat}
 
+Let us recall the formulation of gravitational lensing by
+\cite{1986ApJ...310..568B} in terms of Fermat's principle .  Consider
+a lens at some redshift $z_L$ and let $(x,y)$ be planar coordinates at
+the lens, transverse to the line of sight.  Let $\Sigma(x,y)$ be the
+mass per unit area, projecting along the line of sight.  This surface
+density is often given in a dimensional form
+\begin{equation} \label{eq:kappa}
+\kappa(x,y) \propto \Sigma(x,y) \,.
+\end{equation}
+Consider also a more distant light source, at redshift $z_S$, behind
+point $(x_s,y_s)$ on the lens.
+
+Now imagine a virtual photon coming from the source to some $(x,y)$ on
+the lens, then changing direction and coming to the observer.  Such a
+direction change would increase the light travel time compared to
+coming through $(x_s,y_s)$.  The increased light travel time from the
+geometry of deflection would be
 \begin{equation} \label{eq:tgeom}
@@
 \end{equation}
+assuming it is small compared to the total light travel time.
 
-The gravitational time delay can be expressed in an implicit form: the
+An additional delay of the photon comes from travelling through the
+curved spacetime at the lens, and can be related to the mass
+distribution of the lens by an implicit relation, as follows.  The
 value of $\tgrav$ through $(x,y)$ equals its average value on the
@@
 
-\begin{equation} \label{eq:kappa}
-\kappa(x,y) \propto \Sigma(x,y) \,.
+The light travel time of a virtual photon is therefore longer by
+\begin{equation}  \label{eq:tarriv}
+t(x,y) = t_{\rm geom} + t_{\rm grav} \,.
 \end{equation}
+than it would have been with no lens present.  Real photons take paths
+that make $t(x,y)$ extremal, that is, having a minimum, maximum or
+saddle point (Fermat's principle).
 
-Incidentally, $\kappa$ has a second meaning, as well as
-being a dimensionless form of the projected density: it also
-measures how a bundle of light rays is brought together by
-gravitational lensing, and is called the {\em convergence}.
-
-Consider a gravitational lens.  As in most astrophysical lensing, this
-lens is `thin' along the line of sight, and effectively lies on a
-plane.  Let $x,y$ be coordinates on this plane.  These coordinates
-measure ordinary physical lengths (meters, etc) on the lens plane.
-Let there be a point source of light at $x=0,y=0$ but far behind the
-lens plane.  Now imagine a light ray coming from the source to the
-lens plane $(x,y)$ and then changing its path to reach the observer.
-The observer would see the source apparently behind $(x,y)$ the
-lens.  If this happened, the light ray would have taken a longer path
-than having come directly to the observer.  The geometrical path
-difference would be $\propto x^2 + y^2$ for small angles.  Let us
-define
-\begin{equation} \label{eq:Ageom}
-A_{\rm geom}(x,y) = \half(x^2 + y^2) \,.
-\end{equation}
-As written, this looks an area.  But we can also think of it as a
-time delay introduced by the light ray bending --- with an unwritten
-constant factor.  That factor is basically the lens distance divided by the
-speed of light; the precise expression has to take the expansion of
-the Universe into account, and is given in the Appendix.  The top part
-of \figref{arriv} shows what $A_{\rm geom}$ looks like: it is
-just a paraboloid.  The minimum of the surface is at the source, and
-that is where, according to Fermat's principle, the image will be.
-The rest of the arrival-time surface is just an imaginary surface.
+The proportionality factors in \eqref{eq:kappa} and \eqref{eq:tgeom}
+depend on the redshifts and cosmological parameters and are given in
+the Appendix~\ref{more-theory}.
 
@@
 
-The geometrical time delay \eqref{eq:Ageom} is not the only one.  The
-warping of space by a gravitational field introduces a further time
-delay $A_{\rm grav}$.  The arrival time of the light ray, compared to
-what it would have been with no lensing, is
-\begin{equation} \label{eq:Aarriv}
-\hbox{arrival time} = t_{\rm geom} + t_{\rm grav} \,.
-\end{equation}
-We can also express $A_{\rm grav}$ as an area, with the same implicit
-constant factor that makes it a time.  The expression, however, is
-more complicated.  We first introduce the notation $\nabla^2 f(x,y)$
-to denote
-\begin{equation}
- \frac{ f(x+\Delta x, y) + f(x-\Delta x, y) +
-        f(x, y+\Delta y) + f(x, y-\Delta y) - 4 f(x,y) }
-      {\Delta x \; \Delta y}
-\end{equation}
-in the limit of $\Delta x,\Delta y$ small.  Then we write the
-projected mass density of the lens, which would be in $\rm
-kg\,m^{-2}$, in special units (given in the Appendix) such that the
-surface density becomes a dimensionless field $\kappa(x,y)$.  Then
-\begin{equation} \label{eq:Poisson}
-\nabla^2 A_{\rm grav}(x,y) = -2\kappa(x,y) \,.
-\end{equation}
-Note that $A_{\rm grav}(x,y)$ is not given explicitly in terms of
-$\kappa(x,y)$, but as a differential equation that must be solved.
-Equations of the type \eqref{eq:Poisson} are, however, well known (as
-Poisson's equation in two dimensions), and there are techniques for
-solving them.
+\subsection{``Spaghetti'' diagrams} \label{sec:arriv}
 
-What is the effect of mass on the arrival-time surface?  If the mass
-distribution $\kappa(x,y)$ is circular, $A_{\rm grav}(x,y)$ will have
-a peak at the center of that circle.  But if the mass is not precisely
-in front of the source, the arrival-time surface will not be circular
-any more.  The result is illustrated in the middle part of
-\figref{arriv}.  In addition to the maximum, there is now a
-minimum and a saddle point.  According to Fermat's principle, images
-appear at maximum and saddle points as well as minimum.  If the mass
-distribution is non-circular, more images can appear.  The lower
-example in \figref{arriv} shows the arrival-time surface due
-to an elongated mass.  There is still a maximum, part of an elongated
-hill, and around it there are two minima and two saddle points.
+The full function $t(x,y)$, also known as the arrival-time surface,
+applies to virtual photons.  In other words, it is an abstract
+construct and not itself observable.  But observable image positions
+can be derived from the arrival-time surface, so visualising it is
+useful.  Figure~\figref{arriv} does so.  In this figure, a maximum,
+where present, is easy to see.  To locate mimima and saddle-points,
+however, one needs to examine the contours of equal arrival time.  A
+saddle point is characterised by a contour crossing itself, forming
+an~X.  Mimima, on the other hand, have contours looping around them,
+as do maxima.
 
-So far we have considered a single point source.  What if there is an
-extended source?  To understand this, let us consider what happens if
-we move the original point source slightly, or equivalently, keep the
-original point source behind $x=0,y=0$ and move the lens slightly.
-The contours of constant arrival time will naturally move slightly,
-and so will the images.  The movement of the contours will be most
-noticeable where the contours are far apart, that is where the
+The saddle-point contours which form an X are especially interesting,
+because they set the overall topography of the arrival-time surface.
+The obviously give the locations of the saddle points, and roughly
+localise the minima and maxima as well.  If more precise locations of
+the minima and maxima are added, the whole arrival-time surface is
+already approximately known.  Since the arrival-time surface has an
+exact relation to the lens-mass distribution and the source position,
+in effect the mass distribution is also automatically specified.  In
+other words, a simple sketch of saddle-point contours along with
+locations of minima and maxima ---which we call a spaghetti diagram---
+is implicitly already an approximation to a lens-mass distribution.
+This idea is the basis of \spl.
+
+The preceding assumes a point source.  To get an idea of what an
+extended source would do, let us imagine moving the original source
+slightly.  The contours of constant arrival time will naturally move
+slightly, and so will the images.  The movement of the contours will
+be most noticeable where the contours are far apart, that is where the
 arrival-time surface is nearly flat.  As is evident from
-\figref{arriv}, this is the region where the minimum and saddle
-points lie, or near the images.  In this region, points on the source
-that are close together produce images that are comparatively far
-apart.  In other words, the image is highly magnified.  In summary,
-lower curvature in the arrival-time surface for a point source implies
+\figref{arriv}, this is the region where the minimum and saddle points
+lie, or near the images.  In this region, points on the source that
+are close together produce images that are comparatively far apart.
+In other words, the image is highly magnified.  In summary, lower
+curvature in the arrival-time surface for a point source implies
 larger magnification of an extended source.  Conversely, where the
 arrival-time surface is strongly curved, the image will be
-demagnified.  We see from \figref{arriv} that the arrival-time
-surface tends to be highly curved near the maximum.  Hence maximum tend
-to be demagnified.  In practice, maxima of the arrival time are nearly
+demagnified.  We see from \figref{arriv} that the arrival-time surface
+tends to be highly curved near the maximum.  Hence maximum tend to be
+demagnified.  In practice, maxima of the arrival time are nearly
 always too faint to see. The minima and saddle points dominate.
 
-In the lensed examples in \figref{arriv}, the maximum in each case is
-evident, but how can we distinguish the minima from the saddle points?
-The key is to examine the contours of equal arrival time.  Contour
-curves loop around a minimum or a maximum, but near a saddle point,
-the contours do not loop around, instead they approach and then pull
-away.  The contour curve precisely level with a saddle point forms an
-X at the saddle point.  In \figref{arriv}, one such X is evident.  In
-other cases the X contour passing through the saddle point is not
-itself plotted, but we can tell by the other contours where the X
-would be.  These self-crossing contours will play a special role in
-the following.
-
-Although the arrival-time contours represent an abstract quantity that
-is not directly observable, the contour map may actually resemble the
-appearance of the lensed arcs.  The resemblance is due to a
-serendipitous computer-graphical effect \citep{2001AJ....122..585S}.
-
-\FloatBarrier
 \subsection{\spl} \label{sec:SpaghettiLens}
@@
 
-The fourth modeling step is discussion among modelers and iteration
-on the model.  Any archived model can be revised by another user: they
+The fourth modeling step is discussion among modelers and iteration on
+the model.  Any archived model can be revised by another user: they
 can modify the spaghetti configuration slightly or drastically, or
@@
 Forum discussion then prunes the tree, focusing attention on one or a
-few models.
+few models.\footnote{See ``Collaborative gravitational lens
+  modelling\dots'' in {\tt http://letters.zooniverse.org} for an
+  example.}
 

### 2014-06-06 text/partC.tex
edits

@@
 
-\section{Relation to standard lensing formalism} \label{sec:more-theory}
+\section{Relation to standard lensing formalism}\label{more-theory}
 
-The presentation of the arrival-time surface in \secref{Fermat}
-omitted details, for the sake of a more intuitive explanation.
-Specifically, the geometric time delay $t_{\rm geom}$ and the
-convergence $\kappa$ were left as proportionalities (equations
+The presentation of the arrival-time surface in \S\secref{Fermat}
+and \secref{arriv} omitted some details for the sake of a more
+intuitive explanation.  Specifically, the geometric time delay $t_{\rm
+geom}$ and the convergence $\kappa$ were left as proportionalities
+(equations
 \ref{eq:tgeom} and \ref{eq:kappa}), and the gravitational time delay

### 2014-06-11 plots/conv_inp.py
updated plot generation scripts input: kernel to inclease size of magenta parts mod: cut white border sim: plot without white border

@@ -25,10 +25,60 @@ sel_mod = [
     7025,
 ]
 
-for i in sel_mod:
+#sel_mod = [6915]
+
+for mnr in sel_mod:
     nimg = img.new(img.MODES[6], (new_dim[0],new_dim[0]), color=(255,255,255))
-    oimg = img.open('new_inputs\\%06i.png' % i)
+    
+    oimg = img.open('inputs\\%06i.png' % mnr)
     
     nimg.paste(oimg, (new_dim[2], new_dim[4], 750+new_dim[2],750+new_dim[4]))
+    nimg.save('new_inputs\\%06i_input_old.png' % mnr)
     
-    nimg.save('new_inputs\\%06i_input.png' % i)
+    #set source and target
+    src = oimg    
+    trg = img.new(img.MODES[6], src.size, color=(255,255,255))
+    trg.paste(src, (0, 0, src.size[0],src.size[1]))
+    srcpix = src.load()
+    trgpix = trg.load()
+    sx, sy = src.size
+    
+    #kernel = [(-1,0),(0,-1),(1,0),(0,1)]
+    kernel = [  (-1, 0),( 0,-1),( 1, 0),( 0, 1),
+                (-1,-1),(-1,+1),(+1,-1),(+1,+1)]
+    
+    lernel = [
+    [ 0.0 , 0.2 , 0.5 , 0.2 , 0.0],
+    [ 0.2 , 0.8 , 1.0 , 0.8 , 0.2],
+    [ 0.5 , 1.0 , 1.0 , 1.0 , 0.5],
+    [ 0.2 , 0.8 , 1.0 , 0.8 , 0.2],
+    [ 0.0 , 0.2 , 0.5 , 0.2 , 0.0],
+    ]
+    
+    #lernel = [[1.0]]
+    
+    kernel = []
+    d=len(lernel)//2
+    for i, r in enumerate(lernel):
+        for j, a in enumerate(r):
+            kernel.append((i-d, j-d, a))
+    
+    cr, cg, cb = (255,0,255)
+    for x in range(sx):
+        for y in range(sy):
+            #print pix[x,y]
+            r,g,b = srcpix[x,y]
+            if r==255 and g==0 and b==255:
+                for dx, dy, aa in kernel:
+                    bb = 1.-aa
+                    aa=bb=0.5
+                    try:
+                        r,g,b = srcpix[x+dx,y+dy]
+                    except Exception as e:
+                        #print "error", x, y, dx, dy, e
[…]

### 2014-06-11 plots/gen_plots.py

@@ -66,6 +66,7 @@ import numpy as np
 from numpy import pi
 import matplotlib as mpl
 
+import PIL
 
 debug = False
 
@@ -73,7 +74,7 @@ debug = False
 write_to_tex_folder = False
 
 #timeintensive dl of online data. only neeeds to be done once
-fetch_onlinedata = False
+fetch_onlinedata = True
 
 # do which plots of sims?
 simplots = {'arriv': True, 'kappa':True, 'kappaenc':False}
@@ -87,7 +88,7 @@ figsizeER = (10,8)
 figsizeKE = (10,8)
 
 # realtive to plots dir, incase of debug
-outdir = 'figs_new3'
+outdir = 'figs_new4'
 
 #select sims to produce plots for (using sel_sim_plots())
 sel_sim = [
@@ -559,6 +560,32 @@ def sel_mod_plots():
 
     #if debug: break
   
+
+#==============================================================================
+#   
+# fetch online only used models plots
+#==============================================================================
+def fetch_sel_mod():
+  print '\nMODS (fetching only)'
+  print '============================================='  
+  iii=0
+  for idd, elem in enumerate(spg.data):
+    
+    mid = elem['id']
+    
+    if not mid in sel_mod:
+      print mid
+      continue
+    else:
+        iii=iii+1
+        print iii,'/',len(sel_mod)
+
+    path = os.path.join(moddir, '%06i'%mid)
+    if not os.path.isdir(path):
+      os.makedirs(path)    
+
+    get_mod_adv_data(mid) #download plots generated on server with glass
+  
   
 #==============================================================================
 #   # create tex folder / files for easy include
@@ -874,20 +901,35 @@ def copy_sim_files(asw):
 
 def get_mod_def_data(mid):
   """gets the deault data directly from mite web"""
+  '''
   baseurl = 'http://mite.physik.uzh.ch/result/%06i/' %mid
   
   imgs = ['img2.png', 'img3.png']
   saveas = ['mass.png', 'arr_time.png']
+  
+  # box to cut
+  x=71
+  y=61
+  dd=477
+
+  box_mod = (x,y,x+dd,y+dd)
 
   path = os.path.join(moddir, '%06i'%mid)
 
   for k, img in enumerate(imgs):
     r = rq.get(baseurl+img, stream=True)
[…]

### 2014-06-11 plots/crop_imgs.py
update figures

@@ -1,0 +1,35 @@
+# -*- coding: utf-8 -*-
+"""
+removes the white border from the SL png images
+"""
+
+import numpy as np
+from PIL import Image as img
+
+
+
+# dim of new imgs
+newdim_mod = (box_mod[2]-box_mod[0], box_mod[3]-box_mod[1])
+
+
+sel_mod = [
+    6915,
+    6919,
+    6937,
+    6941,
+    6975,
+    6990,
+    7020,
+    7021,
+    7022,
+    7024,
+    7025,
+]
+
+sel_mod = [6915]
+
+for i in sel_mod:
+    for t in ['arr_time', 'mass']:
+        oimg = img.open('%06i_%s.png' % (i,t))
+        nimg = oimg.crop(box_mod)
+        nimg.save('cropped\\%06i_%s.png' % (i,t))

### 2014-06-15 text/partA.tex
updated plot generation scripts input: kernel to inclease size of magenta parts mod: cut white border sim: plot without white border
editing

@@
 
-Galaxies and clusters of galaxies are considered to be the largest structures in the known Universe.
-When one questions their formation,  the present picture suggests that gravitational instabilities  (observable as fluctuations of the microwave background) %TODO \todo[red]{not too sure of the relevance of this statement}
- - caused a collapse into gravitationally bound structures.
-%TODO\todo[yellow]{clarification needed: are they becoming gravitationally bound structures, or are they collapsing into objects that are already gravitationally bound?
-I know it is most probably the later, but it is a bit ambiguous}
-The initial collapse would have been dominated by dark matter, but once deep wells had appeared in the gravitational potential, gas would have started falling into those wells. 
-%TODO \todo[red]{not too sure of this wording}
-Through atomic processes,
-%TODO \todo[yellow]{perhaps we need to clarify those processes or are they generally know?}
- the gravitational potential energy of this gas would of been radiated away, leading to the formation of the largest stars. 
-%TODO \todo[red]{or perhaps just stars?}
-In turn, galaxies and clusters of galaxies would form, all the time merging and changing their form.
-This rough outline is now well established, but in order to satisfy our understanding of this process, there are many puzzles  that remain to be solved.
-For progress to be made, we need to extend our knowledge of dark matter.
-Gravitational lensing has proven as a particularly attractive way of studying dark matter as it is only dependant on mass.
-%TODO \todo[yellow]{silly question, but shouldn't we briefly explain why it is only dependant on mass? or should everyone already reading this paper know that? or is it explained later?}
+The first discoveries of gravitational lenses, in each case a galaxy
+lensing a background quasar into multiple images
+\citep{1979Natur.279..381W,1980Natur.285..641W}, brought in their wake
+the first work on lens modelling
+\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.  For one of the
+lenses, mass modelling predicted that one of the lensed quasar images
+seen would split further into two at higher resolution, which was
+crucial in subsequently confirming that the object is indeed an
+example of lensing.  It had long been expected that galaxies must
+sometimes cause lensed images \citep{1937ApJ....86..217Z}, and it even
+been argued that the phenomenon could help measure cosmological
+parameters \citep{1964MNRAS.128..307R,1966MNRAS.132..101R}.  But
+apparently nobody was expecting that lenses would need detailed
+modelling.  In hindsight, the reason is not difficult to understand.
+The scale of image separations is set by the angular Einstein radius
+\begin{equation}
+\sim \left(\frac{4GM}{c^2 D_L}\right)^{1/2}
+\simeq 0.1'' \left(\frac{M}{M_\odot}\right)^{1/2}
+             \left(\frac{D_L}{\rm pc}\right)^{-1/2}
+\end{equation}
+where $D_L$ is the distance to the lens, and $M$ its mass.  For
+stellar-mass lenses, the image separation is much larger than the size
+of the lensing mass, so lenses are well approximated by point masses.
+In galaxy lenses, the image separation is comparable to the size of
+the galaxy; typically the lensed images are seen through the galaxy
+halo.  Consequently, the lensed images depend on the detailed mass
+distribution of the lensing galaxy.  Once observations of the lensed
+images became available, the need to model the lensing-mass
+distribution in detail became obvious.
 
-There are many manifestations of gravitational lensing in astrophysics,
-but the most spectacular are multiple images.  Gravitational lensing
-depends on the sky-projection mass density, and this fact has a
-counter-intuitive consequence: nearby galaxies do not produce multiple
-images, because their projected density is not high enough, but beyond
-100~Mpc or so, most galaxies are potential lenses.  Even so, for a
-potential lens to really act as a lens, a background source in the close-by
-region is required.  So, in practice galaxies are very rarely observed
-as lens.  The same tendency applies to clusters of galaxies, but since
-clusters are bigger, they are more likely lens-like.
-
-In lensing galaxies, the observable features usually involve a single
-object multiply imaged.  It is therefore desirable to have large
-numbers of lenses, and this is the strategy researchers are following.  The
-largest single study so far uses 58 lenses \citep{2009ApJ...703L..51K}.
-Other studies have compared lens models in more detail with stellar
-mass \citep{2011ApJ...740...97L} or used time-delay information from
-lensed quasars to infer the Hubble time
-\citep{2008ApJ...679...17C,2010ApJ...712.1378P}.
-
-The trend is clear: researchers want to find many more lenses and
-model them in more sophisticated ways.
+Nowadays, detailed modelling of the mass distribution is part of any
+research using multiply-imaging lenses.  Many works combine multiple
+lenses with different mass ranges and other properties.  The largest
+single study so far used 58 lenses \citep{2009ApJ...703L..51K} to
+uncover the distribution of dark matter around galaxies.
+\cite{2011ApJ...740...97L} combined lens models of 21 galaxies with
+models of their stellar populations, to study the relation between
+stars and dark matter in galaxies.  \cite{2014MNRAS.437..600S}
+modelled 18 time-delay lenses together to infer cosmological
+parameters.  But lens modelling has remained a separate process from
+lens discovery.
 
@@
   \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
-  \vskip-20pt
   \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 

### 2014-06-15 text/partA.tex
editing

@@
 the galaxy; typically the lensed images are seen through the galaxy
-halo.  Consequently, the lensed images depend on the detailed mass
+halo.  Hence the lensed images depend on the detailed mass
 distribution of the lensing galaxy.  Once observations of the lensed
-images became available, the need to model the lensing-mass
-distribution in detail became obvious.
+images became available, it was obvious that mass modelling was
+needed.
 
-Nowadays, detailed modelling of the mass distribution is part of any
-research using multiply-imaging lenses.  Many works combine multiple
-lenses with different mass ranges and other properties.  The largest
-single study so far used 58 lenses \citep{2009ApJ...703L..51K} to
-uncover the distribution of dark matter around galaxies.
+Since that early days, more than 400 secure lenses have been found
+(see Figure \ref{fig:masterlens}) and detailed modelling of the mass
+distribution is part of any research using multiply-imaging lenses.
+But so far no modelling study has used anything approaching 400
+lenses.  The largest single one \citep{2009ApJ...703L..51K} used 58
+lenses to uncover the distribution of dark matter around galaxies.
 \cite{2011ApJ...740...97L} combined lens models of 21 galaxies with
-models of their stellar populations, to study the relation between
+models of their stellar populations, to infer the relation between
 stars and dark matter in galaxies.  \cite{2014MNRAS.437..600S}
 modelled 18 time-delay lenses together to infer cosmological
-parameters.  But lens modelling has remained a separate process from
-lens discovery.
+parameters.
 
-The question now arises: can lenses be found and modeled
-automatically in large surveys?  Work using software robots
-\citep{2009ApJ...694..924M} gives some interesting and unexpected
-results.  In clean lensing system in uncrowded fields with high
-signal-to-noise, the robots do very well.  In most situations,
-however, robots miss lenses (low completeness) or contaminate the
-results with non-lenses (low purity).  Robots can be made to
-prioritize completeness or purity, but they cannot deliver both.
+\begin{figure}
+\centering
+\includegraphics[width=\hsize]{fig/lenssky}
+\caption{Sky distribution of 423 published lenses considered secure
+  (from the Masterlens catalog at the University of Utah, maintained
+  by Joel Brownstein and Leonidas Moustakas.) The map is in
+  Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
+  middle. The empty swathes to the left and right of center are the
+  Milky Way. {\em Should we drop this figure?  It's not relevant to
+    modelling, but as there seems to be no publicly available summary
+    of all known lenses, it does help set context.}}
+\label{fig:masterlens}
+\end{figure}
 
-In response to the lessons learned from automated lens searches, the
-{\em Spacewarps\/} project\footnote{\tt http://www.spacewarps.org}
- has been launched.  {\em Spacewarps\/} is
-part of the {\em Zooniverse\/} family of citizen-science
-projects,\footnote{\tt http://www.zooniverse.org} where members of the
-public are invited to analyze different kinds of scientific data which
-are too difficult for robots and too large for specialists.  In {\em
- Spacewarps\/} itself, survey sky images are presented for volunteers to identify lens candidates.  The first tranche of survey data, introduced from May 2013, is an area of $\sim172$ square
-degrees (or a quarter of a percent of the sky) from the Canada-France-Hawaii Telescope Legacy Survey.
- The survey is divided into patches of $440\times 440$ pixels, each of
-which is seen by ten volunteers.  Simulated lenses are mixed in with
-the data, both to help train volunteers on what to look for, and to
-estimate completeness and purity.
-These data had previously been analyzed by robots \cite{Gavazzi2012, More2012ApJ}.  
-  A list of candidates is being
-processed, but intermediate results show some very good candidates, as
-well as re-discoveries of candidates previously found robotically.
+Integrating modelling into the lens discovery process would be a great
+advantage.  In this paper, we develop a technique towards this goal,
+working with the \sw lens search (Marshall et al, in prep, More et al,
+in prep).  \sw is a citizen science project\footnote{\tt
+  http://www.spacewarps.org} in which volunteers are presented
+sky-survey images and invited to identify lens candidates.  Simulated
+lenses are mixed in with the data, both to help train volunteers on
+what to look for, and to estimate completeness and purity.  The
+motivation for \sw is to enable volunteers, some of whom had
+previously serendipitously identified lens candidates on earlier
+citizen-science surveys, to make discoveries missed in automatic
+searches by software robots.  Robots are very good with clean lensing
+system in uncrowded fields with high signal-to-noise, but in general
+test situations \citep{2009ApJ...694..924M} robots miss lenses (low
+completeness) or contaminate the results with non-lenses (low purity).
+
+The first tranche of survey data, introduced from May 2013, is an area
+of $\sim172$ square degrees (or a quarter of a percent of the sky)
+from the Canada-France-Hawaii Telescope Legacy Survey.  The survey is
+divided into patches of $440\times 440$ pixels, each of which is seen
+by ten volunteers.  These data had previously been analyzed by robots
+\cite{Gavazzi2012, More2012ApJ}.
 

### 2014-06-16 text/partA.tex
editing

@@
 
-Since that early days, more than 400 secure lenses have been found
-(see Figure \ref{fig:masterlens}) and detailed modelling of the mass
-distribution is part of any research using multiply-imaging lenses.
-But so far no modelling study has used anything approaching 400
-lenses.  The largest single one \citep{2009ApJ...703L..51K} used 58
-lenses to uncover the distribution of dark matter around galaxies.
+Since those first discoveries, more than 400 secure lenses are now
+known (see Figure \ref{fig:masterlens}).  Modelling of the mass
+distribution is part of any research using lenses, but so far no
+modelling study has spanned all known lenses.  The largest single one
+\citep{2009ApJ...703L..51K} models 58 lenses to infer the distribution
+of dark matter around galaxies.  In other work,
 \cite{2011ApJ...740...97L} combined lens models of 21 galaxies with
-models of their stellar populations, to infer the relation between
-stars and dark matter in galaxies.  \cite{2014MNRAS.437..600S}
-modelled 18 time-delay lenses together to infer cosmological
-parameters.
+models of their stellar populations, to find the relation between
+stars and dark matter, and \cite{2014MNRAS.437..600S} modelled 18
+time-delay lenses together to infer cosmological parameters.
 
@@
   (from the Masterlens catalog at the University of Utah, maintained
-  by Joel Brownstein and Leonidas Moustakas.) The map is in
+  by Joel Brownstein and Leonidas Moustakas).  The map is in
   Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
-  middle. The empty swathes to the left and right of center are the
+  middle. The empty swathes to the left and right of centre are the
   Milky Way. {\em Should we drop this figure?  It's not relevant to
@@
 
-Integrating modelling into the lens discovery process would be a great
-advantage.  In this paper, we develop a technique towards this goal,
-working with the \sw lens search (Marshall et al, in prep, More et al,
-in prep).  \sw is a citizen science project\footnote{\tt
+Surveys now under way aim to increase the inventory of lenses another
+ten or a hundred fold.  Among these is \sw (Marshall et al, in prep,
+More et al, in prep).  \sw is a citizen science project\footnote{\tt
   http://www.spacewarps.org} in which volunteers are presented
@@
 lenses are mixed in with the data, both to help train volunteers on
-what to look for, and to estimate completeness and purity.  The
-motivation for \sw is to enable volunteers, some of whom had
-previously serendipitously identified lens candidates on earlier
+what to look for, and to provide measures of the effectiveness of the
+search.  The motivation for \sw is to enable volunteers, some of whom
+had previously serendipitously identified lens candidates on earlier
 citizen-science surveys, to make discoveries missed in automatic
@@
 completeness) or contaminate the results with non-lenses (low purity).
+\sw introduced its first tranche of survey data, in May 2013.  The
+Canada-France-Hawaii Telescope Legacy Survey (CFHTLS), covering
+$\simeq172$ square degrees (0.4\% of the sky) was divided into tiles
+of $80''\times80''$ ($440\times 440$ pixels).  Each such tile was
+shown to ten volunteers.  Despite the rarity of detectable lenses
+($<10^{-3}$ per tile), the \sw collective soon found lens candidates,
+both rediscoveries already known from robots \citep{Gavazzi2012,
+  More2012ApJ} and possible new lenses.
 
-The first tranche of survey data, introduced from May 2013, is an area
-of $\sim172$ square degrees (or a quarter of a percent of the sky)
-from the Canada-France-Hawaii Telescope Legacy Survey.  The survey is
-divided into patches of $440\times 440$ pixels, each of which is seen
-by ten volunteers.  These data had previously been analyzed by robots
-\cite{Gavazzi2012, More2012ApJ}.
+The encouraging early results prompted the question: could modeling of
+the lenses also be done by volunteers?  There are several software
+packages available for lens modelling \citep[for a review,
+  see][]{2013NewAR..57....1L} and work has been done on generic
+interfaces \citep{2014A&C.....5...28L}, lens modelling has remained
+the domain of specialists.  An early proposal for what became \sw had
+included a prototype lens modelling tool \citep{2010AAS...21543527N},
+but the learning curve for modelling was subsequently considered too
+long to attract potential volunteers.  On the other hand, some {\em
+  Spacewarps\/} volunteers were quite experienced from earlier
+projects, having spent a thousand hours or more with data, and very
+interested in more demanding projects.  The interests of
+citizen-science communities are just beginning to be studied
+\citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that some
+volunteers welcome open-ended challenges, and sometimes these have led
+to new scientific results: one example is the discovery of an
+exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
+the development of new algorithms for protein folding
+\citep{Khatib22112011}.
 
-These encouraging results now raise the question: could modeling of
-the lenses also be done by volunteers?  Modeling is a much more
-difficult task than searching for lens candidates, as it requires some
-expert knowledge.  Nonetheless, a panel of the {\em Spacewarps\/}
-volunteers is quite experienced from earlier projects, having spent a
-thousand hours or more with data.  Some of these experienced
-volunteers are very interested in more demanding projects.  This also
-is a general trend \citep[cf.][]{Khatib22112011}.  The present project
-then suggests itself.
+A characteristic of strong gravitational lensing is that lensed images
+do not look much like the source, and still less do they resemble the
+mass distribution.  Modellers develop an intuition for image
+configurations, and it is necessary to do so before one can begin to
+model lenses.  Indeed, the modelling section of
+\cite{1981ApJ...244..723Y} can be read as researchers developing
+lens-modellers' intuition for the first time.
 

### 2014-06-16 text/partA.tex
editing

@@
 The encouraging early results prompted the question: could modeling of
-the lenses also be done by volunteers?  There are several software
-packages available for lens modelling \citep[for a review,
-  see][]{2013NewAR..57....1L} and work has been done on generic
-interfaces \citep{2014A&C.....5...28L}, lens modelling has remained
-the domain of specialists.  An early proposal for what became \sw had
-included a prototype lens modelling tool \citep{2010AAS...21543527N},
-but the learning curve for modelling was subsequently considered too
-long to attract potential volunteers.  On the other hand, some {\em
-  Spacewarps\/} volunteers were quite experienced from earlier
+the lenses also be done by volunteers?  An early proposal for what
+became \sw included a prototype lens modelling tool
+\citep{2010AAS...21543527N}.  There are several software tools for
+lensing modelling available, and work has been done on generic
+interfaces \citep{2013NewAR..57....1L,2014A&C.....5...28L}.  Moreover,
+some {\em Spacewarps\/} volunteers are quite experienced from earlier
 projects, having spent a thousand hours or more with data, and very
@@
 the development of new algorithms for protein folding
-\citep{Khatib22112011}.
+\citep{Khatib22112011}.  All these are grounds for optimism.  There
+is, however, a basic difficulty in strong gravitational lensing.
+Lensed images do not look much like their source, and still less do
+they resemble the lensing-mass distribution.  To model a lens, one
+needs to either do a lot of random guessing, or have a good intuition
+for what works.
 
-A characteristic of strong gravitational lensing is that lensed images
-do not look much like the source, and still less do they resemble the
-mass distribution.  Modellers develop an intuition for image
-configurations, and it is necessary to do so before one can begin to
-model lenses.  Indeed, the modelling section of
-\cite{1981ApJ...244..723Y} can be read as researchers developing
-lens-modellers' intuition for the first time.
+In this paper, we propose a way around the difficulty, and report on a
+modelling test on \sw using simulated lenses.  There are three aspects
+to this work, covered in the following three sections.
+
+In section \secref{Fermat} we introduces a theoretical innovation,
+which we call a ``spaghetti diagram''.  This is a cartoon-like markup
+of a lensed system, which implicitly encodes the basis of a mass
+model.  A spaghetti diagram is basically the saddle-point contours
+originally introduced to gravitational lensing by
+\cite{1986ApJ...310..568B} as a way of classifying lensed images.
+They are sometimes shown as part of the output of lens models
+\citep[for
+  example][]{2001ApJ...557..594R,2003ApJ...590...39K,Lubini2012}.  In
+the present work, however, spaghetti diagrams are the {\em input\/}
+through which the modeller tells the program what to do.
+
+In section \secref{SpaghettiLens} we describe the \spl program, which
+implements everything.  \spl is an extension of the GLASS framework
+for modelling lenses \citep{2014arXiv1401.7990C}.  \spl is designed to
+be friendly to the forum style of citizen-science projects, without
+sacrificing any features of GLASS.  The program is invoked online, and
+the modelling history is logged in a transparent way, enabling forum
+discussions and incremental improvement by different people.
+Moreover, any result can be revisited and post-processed if desired.
+These features are of general interest for citizen-science projects,
+but somewhat outside the scope of the present paper.
 
@@
 
-\section{A lens modeling program}
-
-\spl is an extension of the GLASS framework for modelling lenses
-\citep{2014arXiv1401.7990C}, with two innovations.
-\begin{itemize}
-\item \spl is friendly to the forum style of citizen-science projects,
-  without sacrificing any features of GLASS.  The program is invoked
-  online, and the modelling history is logged in a transparent way,
-  enabling forum discussions and incremental improvement by different
-  people.  Moreover, any result can be revisited and post-processed if
-  desired.  These features are of general interest for citizen-science
-  projects, but somewhat outside the scope of the present paper.
-\item The innovation specific to lens modelling is the use of
-  ``spaghetti diagrams'' or saddle-point contours.  These were
-  originally introduced to gravitational lensing by
-  \cite{1986ApJ...310..568B} as a way of classifying lensed images,
-  and are sometimes shown as part of the output of lens models
-  \citep[for
-    example][]{2001ApJ...557..594R,2003ApJ...590...39K,Lubini2012}.
-  In \spl, however, spaghetti diagrams are the {\em input\/} through
-  which the modeller tells the program what to do.
-\end{itemize}
+\section{Fermat's principle and spaghetti diagrams} \label{sec:Fermat}
 
@@
 
-\subsection{Fermat's Principle} \label{sec:Fermat}
+\subsection{Geometrical and gravitational time delays}
 
@@
 
-\subsection{``Spaghetti'' diagrams} \label{sec:arriv}
+\subsection{Arrival-time contours} \label{sec:arriv}
 
@@
 
-\subsection{\spl} \label{sec:SpaghettiLens}
+\section{A lens-modelling program} \label{sec:SpaghettiLens}
 

### 2014-06-17 text/partA.tex
editing

@@
 
-The first discoveries of gravitational lenses, in each case a galaxy
-lensing a background quasar into multiple images
-\citep{1979Natur.279..381W,1980Natur.285..641W}, brought in their wake
-the first work on lens modelling
-\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.  For one of the
-lenses, mass modelling predicted that one of the lensed quasar images
-seen would split further into two at higher resolution, which was
-crucial in subsequently confirming that the object is indeed an
-example of lensing.  It had long been expected that galaxies must
-sometimes cause lensed images \citep{1937ApJ....86..217Z}, and it even
-been argued that the phenomenon could help measure cosmological
-parameters \citep{1964MNRAS.128..307R,1966MNRAS.132..101R}.  But
-apparently nobody was expecting that lenses would need detailed
-modelling.  In hindsight, the reason is not difficult to understand.
-The scale of image separations is set by the angular Einstein radius
+The first two discoveries of gravitational lenses
+\citep{1979Natur.279..381W,1980Natur.285..641W}, where a galaxy images
+a background quasar into two (four), brought in their wake the first
+work on lens modelling
+\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.  For the second lens
+(PG1115+080), mass models scored an early success with the prediction
+that one of the lensed images seen would split further into two at
+higher resolution.  That galaxies must sometimes cause had long been
+expected \citep{1937ApJ....86..217Z}, and it had even been argued that
+the phenomenon could help measure cosmological parameters
+\citep{1964MNRAS.128..307R,1966MNRAS.132..101R}.  But apparently
+nobody was expecting that lenses would need detailed modelling.
+The first observations, however, immediately stimulated models.
+The reason for that lies in the image separation. Recall that image
+separations are of order the angular Einstein radius
 \begin{equation}
@@
 halo.  Hence the lensed images depend on the detailed mass
-distribution of the lensing galaxy.  Once observations of the lensed
-images became available, it was obvious that mass modelling was
-needed.
+distribution of the lensing galaxy.  Galaxy lenses demand modelling.
 
-Since those first discoveries, more than 400 secure lenses are now
-known (see Figure \ref{fig:masterlens}).  Modelling of the mass
-distribution is part of any research using lenses, but so far no
-modelling study has spanned all known lenses.  The largest single one
+Since the early discoveries, more than 400 secure lenses are now known
+(see Figure \ref{fig:masterlens}).  Modelling of the mass distribution
+is part of any research using lenses, but so far no modelling study
+has spanned all known lenses.  The largest single one
 \citep{2009ApJ...703L..51K} models 58 lenses to infer the distribution
@@
 completeness) or contaminate the results with non-lenses (low purity).
-\sw introduced its first tranche of survey data, in May 2013.  The
+\sw introduced its first tranche of survey data, in May 2013, from the
 Canada-France-Hawaii Telescope Legacy Survey (CFHTLS), covering
 $\simeq172$ square degrees (0.4\% of the sky) was divided into tiles
-of $80''\times80''$ ($440\times 440$ pixels).  Each such tile was
-shown to ten volunteers.  Despite the rarity of detectable lenses
+of $\simeq 80''\times80''$ ($440\times 440$ pixels).  Each such tile
+was shown to ten volunteers.  Despite the rarity of detectable lenses
 ($<10^{-3}$ per tile), the \sw collective soon found lens candidates,
@@
 The encouraging early results prompted the question: could modeling of
-the lenses also be done by volunteers?  An early proposal for what
-became \sw included a prototype lens modelling tool
-\citep{2010AAS...21543527N}.  There are several software tools for
-lensing modelling available, and work has been done on generic
-interfaces \citep{2013NewAR..57....1L,2014A&C.....5...28L}.  Moreover,
-some {\em Spacewarps\/} volunteers are quite experienced from earlier
-projects, having spent a thousand hours or more with data, and very
-interested in more demanding projects.  The interests of
-citizen-science communities are just beginning to be studied
-\citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that some
-volunteers welcome open-ended challenges, and sometimes these have led
-to new scientific results: one example is the discovery of an
-exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
-the development of new algorithms for protein folding
-\citep{Khatib22112011}.  All these are grounds for optimism.  There
-is, however, a basic difficulty in strong gravitational lensing.
-Lensed images do not look much like their source, and still less do
-they resemble the lensing-mass distribution.  To model a lens, one
-needs to either do a lot of random guessing, or have a good intuition
-for what works.
+the lenses also be done by volunteers?  There are several software
+tools for lensing modelling available, and work has been done on
+generic interfaces \citep{2013NewAR..57....1L,2014A&C.....5...28L}.
+Some early designs for \sw included a prototype lens modelling tool
+\citep{2010AAS...21543527N}.  Moreover, some {\em Spacewarps\/}
+volunteers are quite experienced from earlier projects, having spent a
+thousand hours or more with data, and are very interested in more
+demanding projects.  The interests of citizen-science communities are
+just beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but
+it is clear that some volunteers welcome open-ended challenges, and
+sometimes these have led to new scientific results: one example is the
+discovery of an exceptional extra-solar planet
+\citep{2013ApJ...768..127S}; another is the development of new
+algorithms for protein folding \citep{Khatib22112011}.  All these are
+grounds for optimism.  There is, however, a basic difficulty in strong
+gravitational lensing.  Lensed images do not look much like their
+source, and still less do they resemble the lensing-mass distribution.
+To model a lens, one needs to either do a lot of random guessing, or
+have a good intuition for what works.
 
 In this paper, we propose a way around the difficulty, and report on a
-modelling test on \sw using simulated lenses.  There are three aspects
-to this work, covered in the following three sections.
+modelling test on \sw using simulated lenses.  The three following
+sections are devoted to the concept, the implementation, and tests
+respectively.
 
-In section \secref{Fermat} we introduces a theoretical innovation,
-which we call a ``spaghetti diagram''.  This is a cartoon-like markup
-of a lensed system, which implicitly encodes the basis of a mass
-model.  A spaghetti diagram is basically the saddle-point contours
-originally introduced to gravitational lensing by
-\cite{1986ApJ...310..568B} as a way of classifying lensed images.
-They are sometimes shown as part of the output of lens models
-\citep[for
+In \secref{Fermat} we introduce a markup system for lensed images,
+which we call a ``spaghetti diagram''.  A spaghetti diagram resembles
+the visible image system, in a cartoon-like way, and at the same time
+it encodes the basis of a mass model.  This supplies an intuitive link
+between the image system and the mass distribution, which look
+frustratingly different from each other.  Spaghetti diagrams are
+essentially the saddle-point contours originally introduced to
+gravitational lensing by \cite{1986ApJ...310..568B} as a way of
+classifying lensed images.  They are sometimes shown as part of the
+output of lens models \citep[for
   example][]{2001ApJ...557..594R,2003ApJ...590...39K,Lubini2012}.  In
@@
 
-In section \secref{SpaghettiLens} we describe the \spl program, which
+In \secref{SpaghettiLens} we describe the \spl program, which
 implements everything.  \spl is an extension of the GLASS framework
-for modelling lenses \citep{2014arXiv1401.7990C}.  \spl is designed to
-be friendly to the forum style of citizen-science projects, without
-sacrificing any features of GLASS.  The program is invoked online, and
-the modelling history is logged in a transparent way, enabling forum
+for modelling lenses \citep{2014arXiv1401.7990C}.  We will not go into
+software details in this paper, concentrating instead on lens
+modelling per se, but we remark that \spl is designed to be friendly
+to the forum style of citizen-science projects, without sacrificing
+any features of GLASS.  The program is invoked online, and the
+modelling history is logged in a transparent way, enabling forum
 discussions and incremental improvement by different people.
 Moreover, any result can be revisited and post-processed if desired.
-These features are of general interest for citizen-science projects,
-but somewhat outside the scope of the present paper.
 
-The purpose of this study was to provide a means to model a large number of gravitational lenses by showing that gravitational lens modeling can be learned / done by volunteers.
-We suggest, that volunteers will be as successful as professionals with modeling if provided with an easy to use tool with visual feedback (What you see is what you get, WYSIWYG) and a minimal set of instructions.
-Volunteers will then crowd work (using buzz word here ;) ) / work collaborative on modeling lenses from several sources / groups at a central place.
-Since this is an iterative learning process, the more involved volunteers will quickly gain knowledge that can be passed down to new volunteers.
-That creates a social structure that scales well with the number of volunteers, as other projects have already shown. %TODO \needcite
-Finding people working as volunteers has been shown to be successful last but not least by \sw and the whole galaxy zoo project.
-To test the people's abilities, we investigated the performance of a few volunteers modeling a set of simulated lenses.
-We tested the ability to correctly identify lensed images and reproduce similar mass map of the lens.
+In \secref{mod_challenge} we describe a modelling challenge where a
+diverse sample of 29 simulated lenses was modelled multiple times
+using \spl.  The models are then examined in two ways.  One was
+whether the spaghetti diagram was correct.  The other was the recovery
+of the Einstein radius.  In addition, we show some visual comparisons
+of actual and recovered lens shape and radial profile, and identify
+some areas to improve, but do not compute statistics on these aspects.
+Profile and shape recovery with GLASS has been studied in more detail
+in \citep{2014arXiv1401.7990C}.
 
-It is well known that making a scientific contribution is a central
-motivation cited by citizen-science volunteers.  So for a lens
-modeler in a citizen-science project, nobody wants that it should
-sacrifice scientific usefulness in order to be fun to use.  That said,
-being aesthetically pleasing and having a short initial learning curve
-are also essential qualities.  Also desirable is that the user is
-encouraged/challenged to go deeper and wider into the subject.
-  \spl tries to address all of these.
+\secref{todo} concludes, with the general outlook and next steps.
 
 \section{Fermat's principle and spaghetti diagrams} \label{sec:Fermat}
-
-We now explain spaghetti diagrams and then outline what \spl does with
-them.
 
@@
 Let us recall the formulation of gravitational lensing by
-\cite{1986ApJ...310..568B} in terms of Fermat's principle .  Consider
-a lens at some redshift $z_L$ and let $(x,y)$ be planar coordinates at
+\cite{1986ApJ...310..568B} in terms of Fermat's principle.  Consider a
+lens at some redshift $z_L$ and let $(x,y)$ be planar coordinates at
 the lens, transverse to the line of sight.  Let $\Sigma(x,y)$ be the
-mass per unit area, projecting along the line of sight.  This surface
-density is often given in a dimensional form
+mass per unit area, i.e., density projected along the line of sight.
+This surface density is often given in a dimensionless form
 \begin{equation} \label{eq:kappa}
-\kappa(x,y) \propto \Sigma(x,y) \,.
+\kappa(x,y) \propto \Sigma(x,y)
 \end{equation}
-Consider also a more distant light source, at redshift $z_S$, behind
-point $(x_s,y_s)$ on the lens.
+called the convergence.  And let there be light, in the form of a more
+distant source, at redshift $z_S$, behind point $(x_s,y_s)$ on the
+lens.
 
-Now imagine a virtual photon coming from the source to some $(x,y)$ on
-the lens, then changing direction and coming to the observer.  Such a
-direction change would increase the light travel time compared to
+We now imagine a virtual photon flying from the source to some $(x,y)$
+on the lens, then changing direction and coming to the observer.  Such
+a direction change would increase the light travel time compared to
 coming through $(x_s,y_s)$.  The increased light travel time from the
@@
 \end{equation}
-assuming it is small compared to the total light travel time.
+assuming the delay is small compared to the total light travel time.
 
@@
 The proportionality factors in \eqref{eq:kappa} and \eqref{eq:tgeom}
-depend on the redshifts and cosmological parameters and are given in
+depend on the redshifts and cosmological parameters, and are given in
 the Appendix~\ref{more-theory}.
@@
 construct and not itself observable.  But observable image positions
-can be derived from the arrival-time surface, so visualising it is
-useful.  Figure~\figref{arriv} does so.  In this figure, a maximum,
-where present, is easy to see.  To locate mimima and saddle-points,
-however, one needs to examine the contours of equal arrival time.  A
-saddle point is characterised by a contour crossing itself, forming
-an~X.  Mimima, on the other hand, have contours looping around them,
-as do maxima.
+can be derived from the arrival-time surface, so visualising the
+surface is useful.  \figref{arriv} does so.  In this figure, a
+maximum, if present, is easy to see.  To locate mimima and
+saddle-points, however, one needs to examine the contours of equal
+arrival time.  A saddle point is characterised by a contour crossing
+itself, forming an~X.  Mimima, on the other hand, have contours
+looping around them, as do maxima.
 
@@
 is implicitly already an approximation to a lens-mass distribution.
-This idea is the basis of \spl.
 
@@
 part of the source.  Input is given by tracing a proposal for the
-saddle-point contours.  \Figref{input-spag} shows some example
-inputs --- and incidentally also indicates the origin of the name
-\spl.
+saddle-point contours.  Examples appear in Figures \ref{fig:6941} to
+\ref{fig:7022} below (upper-left panel in each case).
 
@@
 allowing only valid lensing configurations to be entered, and ensuring
-that the odd image theorem is taken care of.%TODO \needcite
+that the odd image theorem is taken care of.
 

### 2014-06-17 text/partA.tex
editing

@@
 
-\spl implements the ideas of the arrival-time surface into a modeling
-program.
+Modelling with \spl involves three stages, markup of the image,
+followed by intensive numerics in the background, followed by review
+of diagnostics and possible discussion.  We now describe each of
+these, at a conceptual level, without implementation details.
+
+\subsection{Image markup}
 
@@
 
+\subsection{Numerics}
+
 In the second modeling step, once the user has sketched a proposed
@@
 its estimated uncertainty is the range covered by the whole ensemble.
+
+\subsection{Diagnostics}
 

### 2014-06-18 text/partA.tex
a mess

@@
 of diagnostics and possible discussion.  We now describe each of
-these, at a conceptual level, without implementation details.
+these, but do not go into implementation details.
 
@@
 
-The first modeling step with \spl is for the user to call up the \sw
+One begins by going to the \spl page and entering the number of a \sw
+image tile.  \spl then presents it, along with zoom and pan options
+and a markup tool.  The human modeller then has to make an educated
+guess for the topography of the arrival-time surface, and input the
+locations time-ordering of the maxima, minima, and saddle-points.  The
+markup tool allows this information to be entered without the user
+having to think explicitly about image parities or time ordering, by
+tracing saddle-point contours (which the collaboration soon named
+``spaghetti'', leading to the name of the program).
+
+Examples appear in the upper-left panels of Figures \ref{fig:6941} to
+\ref{fig:7022}.
+
+  It
+also allows only valid lensing configurations to be entered, including
+
+ensuring that the odd image theorem is taken care of.
+
+1986ApJ...310..568B
+
+The modeller begins by going to the \spl page, and entering the \sw name
+
+The modeller proceeds by modeling step with \spl is for the user to call up the \sw
 of interest, and make an educated guess for the topography of the
-arrival-time surface --- specifically, the locations of the maxima,
+arrival-time surface, specifically, the locations of the maxima,
 minima and saddle points that correspond to images of the brightest
 part of the source.  Input is given by tracing a proposal for the
-saddle-point contours.  Examples appear in Figures \ref{fig:6941} to
-\ref{fig:7022} below (upper-left panel in each case).
+saddle-point contours.
 
@@
 separate them from other background light sources, and (b)~classify
-and order images according to arrival time.  But the computer helps by
-allowing only valid lensing configurations to be entered, and ensuring
-that the odd image theorem is taken care of.
+and order images according to arrival time.
 

### 2014-06-18 text/partA.tex
editing

--- 2014-06-17 text/partA.tex
@@
 
-The first modeling step with \spl is for the user to call up the \sw
-of interest, and make an educated guess for the topography of the
-arrival-time surface --- specifically, the locations of the maxima,
-minima and saddle points that correspond to images of the brightest
-part of the source.  Input is given by tracing a proposal for the
-saddle-point contours.  Examples appear in Figures \ref{fig:6941} to
-\ref{fig:7022} below (upper-left panel in each case).
+One begins by going to the \spl page and entering the number of a \sw
+image tile.  \spl presents the image, along with zoom and pan options
+and a markup tool to construct a spaghetti diagram.  The human
+modeller now has to make an educated guess for the topography of the
+arrival-time surface, and input the locations time-ordering of the
+maxima, minima, and saddle-points.  The markup tool \citep[which is
+  inspired by Figure~6 of][and is like that figure made interactive
+  and overlaid on data]{1986ApJ...310..568B} lets the modeller enter
+the information by sketching saddle-point contours.  Examples can be
+seen in the upper-left panels of Figures \ref{fig:6941} to
+\ref{fig:7022}.  The loops in the markup tool were the origin of the
+``spaghetti'' metaphor.
 
-\spl, as implemented so far, assumes that the lens is dominated by a
-single galaxy, and that the center of that galaxy is the sole maximum.
-Once the maximum has been identified, we wish to characterize the rest
-of the arrival-time surface in relation to it.  From geometry, any
-surface that has a central maximum and goes high at the edges must
-have a lima\c con-like contour, meaning a two-looped curve with one
-loop inside the other.  The maximum lies within the inner loop, a
-minimum lies between the two loops, and a saddle point lies at the
-self-intersection of the contour.  The top panel of
-\figref{input-spag} shows an example: red marks a suggested
-maximum, green a saddle point, and blue a minimum.  (The small pink
-dots are just help sketch the proposed contour.)  The middle and lower
-panels of \figref{input-spag} show a more complex situation.
-The minimum has split into three images: two new minima with a saddle
-point in between.  In this situation the contour through the new
-saddle point forms a figure of eight, its two loops enclosing the two
-new minima.  The configuration of a lima\c con, possibly with a figure
-of eight inside, suffices to give a preliminary description of the
-arrival-time surface for nearly all cases where the lens is dominated
-by one galaxy, and the source is a single small object.
+The exact placement of the loops in a spaghetti diagram has no
+significance, only the connections, and the heirarchy of which loop is
+inside which matters.  The loops are there simply to help modeller's
+intuition.  The user does not need to think explicitly the image
+parities (though the markup tool provides this information using
+colour codes) or about time-ordering, or worry about the odd-image
+theorem.  The markup tool allows only valid lensing configurations to
+be entered.
 
-The exact placement of the pink spaghetti loops shown in
-\Figref{input-spag} has no significance.  Their utility is
-just to help mark up the image as maximum, minima and saddle points.
-As the figure suggests, the marking up can be done very easily with a
-mouse.  Human intuition is required to (a)~identify lensed images and
-separate them from other background light sources, and (b)~classify
-and order images according to arrival time.  But the computer helps by
-allowing only valid lensing configurations to be entered, and ensuring
-that the odd image theorem is taken care of.
+As implemented so far, \spl assumes that the lens is dominated by a
+single galaxy.  Only one maximum is permitted, and it is taken as the
+centre of the main lensing galaxy.  The user can, however, mark
+additional minor galaxies.  The latter are modelled as point
+masses.
 
@@
 
-In the second modeling step, once the user has sketched a proposed
-spaghetti configuration, \spl sends the input to its server-side
-modeling engine, called GLASS \citep{2014arXiv1401.7990C}.  The task
-of GLASS is to find a mass distribution $\kappa(x,y)$ that exactly
-reproduces the locations of the maximum, minima and saddle
-points. Now, this criterion by itself is extremely under-determined
---- there are infinitely many mass distributions that will reproduce a
-given set of maxima, minima and saddle points, but typically they
-(a)~produce lots of extra images, and (b)~look very unlike galaxies.
-Additional assumptions (a prior) are necessary.  GLASS uses a prior
-based on suggestions by \cite{1997MNRAS.292..148S}.
+Having sketched a spaghetti diagram, the user orders a model, and \spl
+sends the input to GLASS, which runs server-side as it is
+compute-intensive.  The task of GLASS is to find a mass distribution
+$\kappa(x,y)$ that exactly reproduces the given locations of the
+maximum, minima and saddle points. Now, this criterion by itself is
+extremely under-determined --- there are infinitely many mass
+distributions that will reproduce a given set of maxima, minima and
+saddle points, but typically they (a)~produce lots of extra images,
+and (b)~look very unlike galaxies.  Additional assumptions (a prior)
+are necessary.  GLASS uses the following prior
+\citep[cf]{1997MNRAS.292..148S,2008ApJ...679...17C}.
 \begin{enumerate}
@@
 It is then possible to generate an ensemble of models.  The sampling
-technique used by GLASS is described in \citep{Lubini2012}, and
-improves upon earlier techniques \citep{2000AJ....119..439W,Saha2004}.
+technique used by GLASS is described in \citep{Lubini2012}.
 Typically, ensembles of 200 models are used.  That is to say, what we
@@
 
-In the third modeling step, having generated a model ensemble, \spl
-post-processes it to present results and diagnostics to the user for
-inspection. This takes the form of three figures.
+After the model ensemble has been generated, \spl post-processes it to
+present results and diagnostics to the user for inspection. This takes
+the form of three figures.
 \begin{enumerate}
@@
 
-The fourth modeling step is discussion among modelers and iteration on
-the model.  Any archived model can be revised by another user: they
-can modify the spaghetti configuration slightly or drastically, or
-change options like the size of the mass tiles.  Particularly
-interesting lens candidates lead to trees of models in this way.
-Forum discussion then prunes the tree, focusing attention on one or a
-few models.\footnote{See ``Collaborative gravitational lens
+After archiving, there can be discussion among modellers and revision
+of the model.  Any archived model can be revised by any user: they can
+modify the spaghetti configuration slightly or drastically, or change
+options like the size of the mass tiles.  Particularly interesting
+lens candidates lead to trees of models in this way.  Forum discussion
+then prunes the tree, focusing attention on one or a few
+models.\footnote{See ``Collaborative gravitational lens
   modelling\dots'' in {\tt http://letters.zooniverse.org} for an

### 2014-06-20 text/partA.tex
editing

@@
 are necessary.  GLASS uses the following prior
-\citep[cf]{1997MNRAS.292..148S,2008ApJ...679...17C}.
+\citep[cf.]{1997MNRAS.292..148S,2008ApJ...679...17C}.
 \begin{enumerate}
@@
 \item A synthetic image of the lensed features.
-\item A contour map of the arrival-time surface.
+\item A contour map of the arrival-time surface $t(x,y)$.
 \item A gray scale plus contour map of the mass distribution.
 \end{enumerate}
-After examining this feedback, the user can archive the results for
-discussion, or modify their input and try again, or discard the
-attempt altogether.
+The synthetic image generated by \spl assumes a simple conical source
+profile.  The user can change the contrast level on the image, which
+(though it is not saved) amounts to adjusting the width of the cone.
+These synthetic images are still very crude and not very useful for
+assessing models.  The best indicator, in practice, of whether the
+modelling was successful is contour map of $t(x,y)$, with saddle-point
+contours highlighted.  It is, in effect, the computer's refinement of
+the spaghetti diagram input by the user.  If the arrival-time surface
+looks qualitatively similar to the spaghetti diagram, that generally
+indicates a successful model.  The mass distribution also provides
+indications; successful models generally lead to smooth-looking mass
+distribution, whereas an irregular or checkboard pattern in the mass
+map signals a bad model.
 
-After archiving, there can be discussion among modellers and revision
-of the model.  Any archived model can be revised by any user: they can
-modify the spaghetti configuration slightly or drastically, or change
-options like the size of the mass tiles.  Particularly interesting
-lens candidates lead to trees of models in this way.  Forum discussion
-then prunes the tree, focusing attention on one or a few
-models.\footnote{See ``Collaborative gravitational lens
-  modelling\dots'' in {\tt http://letters.zooniverse.org} for an
-  example.}
+After examining this feedback, the user can choose to post the model
+on the \spl archive.  They can also modify their input and try again,
+or discard the attempt altogether.  After archiving, there can be
+discussion among modellers, through the \sw forum or by any other
+means, and revision of the model.  Any archived model can be revised
+by any user: they can modify the spaghetti configuration slightly or
+drastically, or change options like the size of the mass tiles.
+Particularly interesting lens candidates lead to trees of models in
+this way.  Discussion among modellers tends to prune a model tree,
+focusing attention on the most interesting models.\footnote{See
+  ``Collaborative gravitational lens modelling\dots'' in {\tt
+    http://letters.zooniverse.org} for an example.}
 
-\section{A lens modeling challenge} \label{sec:mod_challenge}
+\section{A lens modelling challenge} \label{sec:mod_challenge}
+
+Once an early version of \spl
+
+Through the SW discusison forum (TALK), we asked active members of the Space Warps Citizen community for interest in modelling lens systems.
+XX interested users volunteered to perform modelling and were introduce
 
@@
 \S\ref{sec:SpaghettiLens}).  These derive from the mean of an ensemble
-of 200 models generated by \spl.  To generate the synthetic image,
-\spl assumes a simple conical source profile.  The user can change the
-contrast level on the image, which (though it is not saved) amounts to
-adjusting the width of the cone.  These synthetic images are still
-very crude and not useful for assessing models.  The best indicator of
-whether the modelling was successful is the arrival-time surface
-$t(x,y)$, shown as contour maps. Contours passing through saddle
-points are shown in black: these are the model versions of the
-spaghetti sketch provided by the user.  The model $\kappa(x,y)$ is
-shown as a contour map superimposed on an intensity map. A fairly
-smooth mass distribution, as here, is a good sign.  An irregular or
-checkboard pattern in the mass map usually indicates a bad model.
+of 200 models generated by \spl.
 

### 2014-06-21 text/partA.tex
editing

@@
 modelling per se, but we remark that \spl is designed to be friendly
-to the forum style of citizen-science projects, without sacrificing
-any features of GLASS.  The program is invoked online, and the
-modelling history is logged in a transparent way, enabling forum
-discussions and incremental improvement by different people.
-Moreover, any result can be revisited and post-processed if desired.
+to the forum style of citizen-science projects, and enable incremental
+model refinement by different people, without sacrificing any features
+of GLASS.
 
@@
 
+We now explain the lensing theory relevant to \spl, following the
+formulation of gravitational lensing in terms of Fermat's principle by
+\cite{1986ApJ...310..568B}.
+
 \subsection{Geometrical and gravitational time delays}
 
-Let us recall the formulation of gravitational lensing by
-\cite{1986ApJ...310..568B} in terms of Fermat's principle.  Consider a
-lens at some redshift $z_L$ and let $(x,y)$ be planar coordinates at
-the lens, transverse to the line of sight.  Let $\Sigma(x,y)$ be the
-mass per unit area, i.e., density projected along the line of sight.
-This surface density is often given in a dimensionless form
+Consider a lens at some redshift $z_L$ and let $(x,y)$ be planar
+coordinates at the lens, transverse to the line of sight.  Let
+$\Sigma(x,y)$ be the mass per unit area, i.e., density projected along
+the line of sight.  This surface density is often given in a
+dimensionless form
 \begin{equation} \label{eq:kappa}
@@
 An additional delay of the photon comes from travelling through the
-curved spacetime at the lens, and can be related to the mass
-distribution of the lens by an implicit relation, as follows.  The
-value of $\tgrav$ through $(x,y)$ equals its average value on the
+curved spacetime at the lens.  This gravitational time delay $\tgrav$
+is related to the mass distribution of the lens.  The relation is
+generally written as a two-dimensional Poisson equation, but an
+alternative expression, avoiding calculus, is as follows.  The value
+of $\tgrav$ through $(x,y)$ equals its average value on the
 circumference of a small circle centred at $(x,y)$, plus a constant
-times the mass within that circle.  The proportionality constant is
-$2G/c^3$ times the cosmological expansion factor $(1+z_L)$. Thus
+times the mass within that circle.  The constant is $2G/c^3$ times the
+cosmological expansion factor $(1+z_L)$. Thus
 \begin{equation} \label{eq:tgrav}
@@
 
+\spl is not part of \sw, but it makes use of the \sw infrastructure,
+in particular, the image database and the forum.  The forum is
+essential essential for initial contact with interested members of the
+\sw community.  Lens modelling from \sw works, in practice, in the
+usual style of medium-sized astronomical collaborations, with
+videoconferencing and in-person meetings where possible, but is
+reported in summary on modelling threads on the forum, and anyone
+interested is welcome to join at any time.  Another aspect of \sw (and
+the Zooniverse family of projects in general) that \spl adopts, is to
+keep all user interaction in a web browser, with intensive computing
+done on a server.  Several interesting design and implementation
+issues arise from these considerations, but we will not go into them
+in this paper, concentrating instead on aspects specific to lens
+modelling.
+
 Modelling with \spl involves three stages, markup of the image,
@@
 of diagnostics and possible discussion.  We now describe each of
-these, but do not go into implementation details.
+these.
 
@@
 
-Once an early version of \spl
-
-Through the SW discusison forum (TALK), we asked active members of the Space Warps Citizen community for interest in modelling lens systems.
-XX interested users volunteered to perform modelling and were introduce
-
-Interested volunteers from the \sw forum were initially introduced to
-\spl through a video tutorial and by videocon.  After this
-introductory stage, a modeling challenge was presented.  This
+After an introductory stage, a modeling challenge was presented.  This
 consisted of 29 simulated lenses (sims) covering a range of lensing

### 2014-06-21 plots/restats.py
stats of RE

@@ -1,0 +1,24 @@
+# -*- coding: utf-8 -*-
+"""
+Created on Sat Jun 21 14:22:48 2014
+
+@author: psaha
+"""
+
+from numpy import mean, std
+
+full = []
+fil = open('new_data.csv')
+while True:
+    lyne = fil.readline()
+    if not lyne:
+        break
+    full.append(lyne.split(','))
+
+ratio = map(lambda p: [float(p[1])/float(p[0]),p[2]],full)
+
+expert = [p[0] for p in ratio if p[1][:6]=='expert']
+many = [p[0] for p in ratio if p[1][:6]!='reject']
+
+print mean(expert), std(expert)
+print mean(many), std(many)

### 2014-06-21 text/ms.tex
editing

@@
 \newcommand{\pasj}{PASJ}
+\newcommand{\nat}{Nature}
+\newcommand{\nar}{New Astr Rev}
 
@@
 \end{document}
-
-In \secref{Fermat}, for the sake of a more intuitive
-explanation, we suppressed some constant factors in
-equations \eqref{eq:Ageom}, \eqref{eq:Aarriv} and \eqref{eq:Poisson}.
-Here we fill in the details.
-
-First let us recall various distance formulas. Comoving distances
-involve integrals of the type
-\begin{equation}
-\int \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
-%\raise.5ex\hbox{.}
-\end{equation}
-Specifically, the comoving distance $D_S$ from observer to source,
-$D_{LS}$ from lens to source, and $D_L$ from observer to lens are
-given by the above integral, with limits as follows.
-\begin{equation}
-\begin{aligned}
-&D_S    &                                &\int_0^{z_S} \\
-\noalign{\medskip}
-&D_{LS} & = \frac c{H_0} \ \times \quad  &\int_{z_L}^{z_S} \\
-\noalign{\medskip}
-&D_L    &                                &\int_0^{z_L}
-\end{aligned}
-\end{equation}
-Angular-diameter distances $d$ can be related to comoving distances
-$D$ by scale factors, as follows.
-\begin{equation}
-\begin{aligned}
-D_S &= (1+z_S) \, d_S \\
-D_{LS} &= \frac{1+z_S}{1+z_L} \, d_{LS} \\
-D_L &= (1+z_L) \, d_L
-\end{aligned}
-\end{equation}
-Consequently, angular-diameter distances are always smaller than
-comoving distances.  Moreover, while comoving distances can be simply
-added, angular distances cannot:
-\begin{equation}
-d_S \neq d_L + d_{LS}
-\end{equation}
-
-Locations on the lens plane in angular coordinates on the sky, are
-expressed as
-\begin{equation}
-(x,y) = d_L (\theta_x,\theta_y)
-\end{equation}
-
-The $A$ areas (with the implicit
-constant factor that makes them time delays) and the $\kappa$ density are related to physical arrival
-time $t$ and density $\Sigma$ as
-\begin{equation}
-\begin{aligned}
-A           &= \frac{c}{(1+z_L)^2} \frac{D_L D_{LS}}{D_S} \times t \\
-\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{D_L D_{LS}}{D_S}
-               \times \Sigma(x,y)
-\end{aligned}
-\end{equation}
-Letting the source be offset at $(s_x,s_y)$ rather than at the
-origin, we have
-\begin{equation}
-\begin{aligned}
-t_{\rm geom} &= \frac{(1+z_L)^2}{2c} \frac{D_S}{D_L D_{LS}}
-\left( (x-s_x)^2 + (y-s_y)^2 \right) \\
-\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y)
-\end{aligned}
-\end{equation}
-We can now compare with equations (2.1) to (2.6)
-from \cite{1986ApJ...310..568B}.
-%% comparison to be explicited
-
-\section{Miscellaneous}
-
-Maybe discard \figref{ER_all_models}, since the essential information is there in \figref{ER_per_sim}.
-
-\begin{figure}[htbp]
-  \centering
-    \includegraphics[width=0.90\linewidth]{fig/eR_1}
-  \caption{\ERf for all models with estimated errors in blue squares, \ERg of simulation in red crosses}
-  \label{fig:ER_all_models}
-\end{figure}

### 2014-06-21 text/partA.tex
editing

@@
 morphologies among the spacewarps sims.  Modellers then contributed
-129 models for these sims (at least for each sim).  Models were
+119 models for these sims (at least for each sim).  Models were
 collecting in the same forum used to model real candidate lenses, and
@@
 
-The modellers proposed a total of 129 models for the 29 sims in the
+The modellers proposed a total of 119 models for the 29 sims in the
 challenge. Typically XX models were produced per simulation, with XX
@@
 
-The modelers proposed a total of 129 models for the 29 sims in the
+The modelers proposed a total of 119 models for the 29 sims in the
 challenge.  Figures \ref{fig:6941} to \ref{fig:6919} show eight of

### 2014-06-21 text/partB.tex
stats of RE
editing

@@
 
-\Figref{ER_per_sim} shows that the calculated \ERf of the models tend to be too high.
-The overshoot varies from around 0.2 to 0.4 for good models.
+\Figref{ER_per_sim} shows that the calculated \ERf of the models tend
+to be too high.  The overshoot varies from around 0.2 to 0.4 for good
+models.  Discounting the models which were flagged by the users as
+poor, \ERf{E,rec} is overestimated.  The mean and standard deviation
+of the overestimate are 25\% and 20\%.  The models by an expert also
+show a similar bias, having mean and standard deviation of the
+overestimate being 20\% and 8\%.
 One of the reasons for this is that it is hard to get the center of the lens on spot.
@@
 We thank the Swiss Society for Astrophysics and Astronomy and the
-Swiss Academy of Sciences.
+Swiss Academy of Sciences. AV is supported by a research fellowship
+from the Leverhulme Trust.
 

### 2014-06-21 plots/restats.py

@@ -18,7 +18,9 @@ while True:
 ratio = map(lambda p: [float(p[1])/float(p[0]),p[2]],full)
 
 expert = [p[0] for p in ratio if p[1][:6]=='expert']
+good = [p[0] for p in ratio if p[1][:4]=='good']
 many = [p[0] for p in ratio if p[1][:6]!='reject']
 
+print len(good), len(ratio)
 print mean(expert), std(expert)
 print mean(many), std(many)

### 2014-06-23 text/partA.tex
edits

@@
 \begin{abstract}
-In the initial phase of \sw (SW1), volunteers were invited to search the CFHTLS sky survey to look for gravitational lenses.
-Here we report on a web application that gives experienced volunteers the opportunity to model the candidates that have been identified.
-In order to gauge the quality of the models that were being rendered, these same volunteers were invited to model a sample of 29 simulated lenses.
-The models were then examined in greater detail, with particular attention being paid to the following:
-(i)~identification of image parities and time arrivals;
-(ii)~the mean convergence (equivalent to the enclosed mass), and finally;
-(iii)~the performance of a volunteer vs. a professional.
-In most cases, the volunteers were found to correctly identify the image parities and time arrivals.
-% Einstein radius +~20%
- along with a mean convergence that was well constrained within the image region;
-In all, the results could be comparable to that of a professional.
+We develop a method to enable collaborative modelling of gravitational
+lenses and lens candidates by experienced but non-professional lens
+enthusiasts.  It uses an existing free-form modelling program (GLASS)
+but provides the modelling input in a novel way, via a user-provided
+diagram that is essentially a sketch of an arrival-time surface.
+
+An implementation (\spl) is tested in a modelling challenge using 29
+simulated lenses out of a larger set created for a citizen-science
+lens search (\sw).  We find that amateur lensers inferred the image
+parities and time ordering consistently in some lenses, while tending
+to make errors in other lenses, depending on the image morphology.
+Errors in image parity and time ordering lead to large errors in the
+mass distribution, but the enclosed mass is more robust.  Shortcomings
+include (a)~lens profiles that are too shallow, (b)~over-estimation of
+the Einstein radius by $\sim25\%$, (c)~lack of detailed source
+modelling.  Ideas for improvement are discussed.
 \end{abstract}
@@
 \sw introduced its first tranche of survey data, in May 2013, from the
-Canada-France-Hawaii Telescope Legacy Survey (CFHTLS), covering
-$\simeq172$ square degrees (0.4\% of the sky) was divided into tiles
-of $\simeq 80''\times80''$ ($440\times 440$ pixels).  Each such tile
-was shown to ten volunteers.  Despite the rarity of detectable lenses
-($<10^{-3}$ per tile), the \sw collective soon found lens candidates,
-both rediscoveries already known from robots \citep{Gavazzi2012,
-  More2012ApJ} and possible new lenses.
+Canada-France-Hawaii Telescope Legacy Survey, or CFHTLS
+\citep{Coupon2009}, covering $\simeq172$ square degrees (0.4\% of the
+sky) was divided into tiles of $\simeq 80''\times80''$ ($440\times
+440$ pixels).  Each such tile was shown to ten volunteers.  Despite
+the rarity of detectable lenses ($<10^{-3}$ per tile), the \sw
+collective soon found lens candidates, both rediscoveries already
+known from robots \citep{Gavazzi2012, More2012ApJ} and possible new
+lenses.
 
@@
 
-After an introductory stage, a modeling challenge was presented.  This
-consisted of 29 simulated lenses (sims) covering a range of lensing
-configurations.
-
-The \sw sims were generated by AM, in consultation with PM and AV.
-To estimate the performance of the volunteers and the quality of the generated models, two analysis  were done.
-The first analysis tested the correct identification and ordering of lensed images.
-The second one compared the mass distribution of the lens $\kappa(x, y)$ of the generated models to the mass distribution of the simulations.
-
-A challenge set of 29 was selected, representing the different image
-morphologies among the spacewarps sims.  Modellers then contributed
-119 models for these sims (at least for each sim).  Models were
-collecting in the same forum used to model real candidate lenses, and
-under the same conditions --- modellers were free to consult and
+We now describe a test of the lens-modelling system, under conditions
+that mimic as closely as possible the modelling of real lens
+discoveries.  The lenses to be modelled were the simulated lenses
+(known as ``sims'') already sprinkled onto the \sw images.  Once a
+small user base had grown around \spl, a modelling challenge was
+announced through the \sw forum.  The challenge set consisted of 29
+sims, chosen to represent the different visual morphologies of \sw
+sims. Modellers then contributed 119 models for these sims (at least
+two for each sim).  Models were reported on the same forum used to
+model real candidate lenses.  Modellers were free to consult and
 refine each other's models, but had no information on how the sims
@@
 
+Once the modelling was complete, the models were compared with the
+originals.  There were two main tests: a check of whether the the
+spaghetti diagrams were correct for the lens in question, and a
+comparison of the effective Einstein radii of the sims and the models.
+
 \subsection{The simulated lenses} \label{sec:sims}
 
-The sims were produced using {\tt gravlens}
-\citep{2001astro.ph..2341K,2001astro.ph..2340K} using the CFHTLS survey data
-and catalogues \cite{Coupon2009}. They were of three
-kinds, as follows.
+The \sw sims are described in detail in More et al (in prep), but
+relevant here is that they were of three kinds, as follows.
 
@@
 \end{enumerate}
-
+The {\tt gravlens} program \citep{2001astro.ph..2340K} was used.
 Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
@@
 
-The information in this section was not revealed to the person
-choosing the challenge set (RK) or to the modellers (EB, CC, CM, PS,
-JO and JW) until the modeling stage was done.
+The information in this section was not revealed to the main developer
+of \spl (RK, who also chose the challenge set) or to the modellers
+(EB, CC, CM, JO, PSand JW) until the modelling stage was done.
 
@@
 
-The modellers proposed a total of 119 models for the 29 sims in the
-challenge. Typically XX models were produced per simulation, with XX
-systems having only XX models (corresponding tot he most simplest
-simulations) to XX having XX models (a complex and contrversial lens
-system). In the following we discuss eight example lenses in detail
-with individual notes per system and corresponding plots are shown in
-figures 4-11 (or whatever compact version you end up making). These
-eight examples were chosen on the basis of …??
+Of the 119 models proposed, we discuss now present eight examples in
+some detail.  Results from these are shown in Figures
+\ref{fig:6941}--\ref{fig:7022}.  The first four of these show the most
+common image morphologies, the other four explain some problem cases.
 
-The modelers proposed a total of 119 models for the 29 sims in the
-challenge.  Figures \ref{fig:6941} to \ref{fig:6919} show eight of
-models in some detail.
+Each of Figures \ref{fig:6941}--\ref{fig:7022} figures has the
+following layout.
+$$ \begin{matrix}
+\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
+t(x,y)                        &\hbox{model } t(x,y) \\
+\kappa(x,y)                   &\hbox{model } \kappa(x,y)
+\end{matrix} $$
+The marked-up image is a zoom of the lensed image on Spacewarps,
+marked up with one or more spaghetti contours; this is the modeller's
+input to \spl.  The three panels on the right show the graphical
+output returned to the modeller by \spl, listed as (i), (ii), (iii) in
+\secref{SpaghettiLens}.  These derive from the mean of an ensemble
+of 200 models generated by \spl.
 
@@
     simulation, and the right panels are \spl output.  Details of
-    individual panels are in \ref{sec:example_models}.}
+    individual panels are in \secref{example_models}.}
 
@@
   \caption[result 6915 (ASW0001hpf)]{A four-image configuration
-    typical of lensed quasars. (See Section \ref{sec:example_models}
-    for details.)}
+    typical of lensed quasars. (See \secref{example_models} for
+    details.)}
   \label{fig:6915}
@@
     counter-image; here the arc is closer to the lensing galaxy than
-    the counter-image. (See Section \ref{sec:example_models} for
-    details.)}
+    the counter-image. (See \secref{example_models} for details.)}
   \label{fig:6919}
@@
   \caption[result 6975 (ASW000195x)]{A lens with unrecovered mass
-    substructure. (See Section \ref{sec:example_models} for details.)}
+    substructure. (See \secref{example_models} for details.)}
   \label{fig:6975}
@@
     \ref{fig:7025}, this time with image parities correctly
-    identified. (See Section \ref{sec:example_models} for details.)}
+    identified. (See \secref{example_models} for details.)}
   \label{fig:7022}
@@
 \FloatBarrier
-
-Figures \ref{fig:6941}--\ref{fig:7022} show details of eight models.
-The first four of these show the most common image morphologies, the
-other four explain some problem cases.  Each of the figures has the
-following layout.
-$$ \begin{matrix}
-\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
-t(x,y)                        &\hbox{model } t(x,y) \\
-\kappa(x,y)                   &\hbox{model } \kappa(x,y)
-\end{matrix} $$
-The marked-up image is a zoom of the lensed image on Spacewarps,
-marked up with one or more spaghetti contours; this is the modeller's
-input to \spl.  The three panels on the right show the graphical
-output returned to the modeller by \spl, listed as (i), (ii), (iii) in
-\S\ref{sec:SpaghettiLens}).  These derive from the mean of an ensemble
-of 200 models generated by \spl.
 

### 2014-06-23 text/partB.tex
edits

@@
 
+The \spl program was developed by Kueng, with design suggestions from
+Coles, Cornen and Saha, and feedback from all co-authors.  The
+simulations were created by A.~More, in consultation with Marshall,
+S.~More and Verma.  Modelling was done by Baeten, Cornen, McMillan,
+Odermatt, Saha and Wilcox, with post-modelling analysis by Kueng and
+Saha.  All authors participated in writing and editing the manuscript.
+
 \section{Acknowledgments}

### 2014-06-23 text/partC.tex
edits

@@
 
-The presentation of the arrival-time surface in \S\secref{Fermat}
-and \secref{arriv} omitted some details for the sake of a more
-intuitive explanation.  Specifically, the geometric time delay $t_{\rm
-geom}$ and the convergence $\kappa$ were left as proportionalities
-(equations
+The presentation of the arrival-time surface in
+Sections~\ref{ref:Fermat} and \ref{sec:arriv} omitted some details for
+the sake of a more intuitive explanation.  Specifically, the geometric
+time delay $t_{\rm geom}$ and the convergence $\kappa$ were left as
+proportionalities (equations
 \ref{eq:tgeom} and \ref{eq:kappa}), and the gravitational time delay

### 2014-06-23 text/partB.tex
edits

@@
   optimize resemblance to the observed lensed image after the lens
-  model has been generated, is algorithmically straightforward and
-  planned to be implemented.  This would alleviate another problem
-  with \spl, which is that there is no numerical figure of merit, and
-  assessment of a model is a judgment call based on the synthetic
-  image, and on whether the mass distribution and the arrival-time
-  surface show suspect features.
+  model has been generated, is algorithmically straightforward
+  \citep[cf.][]{2003ApJ...590..673W,2006MNRAS.371..983S} and planned
+  to be implemented.  This would alleviate another problem with \spl,
+  which is that there is no numerical figure of merit, and assessment
+  of a model is a judgment call based on the synthetic image, and on
+  whether the mass distribution and the arrival-time surface show
+  suspect features.
 \item \spl has a tendency to somewhat overestimate the Einstein radius

### 2014-06-26 upd_plots/updateplots.py
working on updating the plots

@@ -1,0 +1,7 @@
+glass_basis('glass.basis.pixels', solver=None)
+
+#g = loadstate(path+'state.txt')
+#g.make_ensemble_average()
+#g.srcdiff_plot_adv(g.ensemble_average, night=True, upsample=8)
+#pl.savefig(path+'img3_ipol.png', facecolor='black', edgecolor='none')
+#pl.close()

### 2014-06-26 text/ms.tex

@@
 \usepackage{graphicx}
-% prefere pdf over png (better quality, lossless compression)
+% prefer pdf over png (better quality, lossless compression)
 \DeclareGraphicsExtensions{%
@@
 Aprajita Verma,$^{4}$
-Amit Kapadia,$^{5}$
 \newauthor
-Elisabeth Baeten,$^{6}$
-Claude Cornen,$^{6}$
-Christine McMillan,$^{6}$
-Julianne Wilcox,$^{6}$
+Elisabeth Baeten,$^{5}$
+Claude Cornen,$^{5}$
+Christine McMillan,$^{5}$
+Julianne Wilcox,$^{5}$
 \newauthor
-Jonas Odermatt,$^{7}$
-Jonathan Coles,$^{8}$
+Jonas Odermatt,$^{6}$
+Jonathan Coles,$^{7}$
 Surhud More$^{3}$
@@
 $^{4}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
-$^{5}$Adler Planetarium, 1300 S. Lake Shore Drive, Chicago, IL 60605, USA\\
-$^{6}$Spacewarps\\
-$^{7}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
-$^{8}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France}
+$^{5}$Spacewarps\\
+$^{6}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
+$^{7}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France}
+%$^{8}$Adler Planetarium, 1300 S. Lake Shore Drive, Chicago, IL 60605, USA\\
 

### 2014-06-26 text/partA.tex

@@
 
-The encouraging early results prompted the question: could modeling of
-the lenses also be done by volunteers?  There are several software
-tools for lensing modelling available, and work has been done on
-generic interfaces \citep{2013NewAR..57....1L,2014A&C.....5...28L}.
-Some early designs for \sw included a prototype lens modelling tool
-\citep{2010AAS...21543527N}.  Moreover, some {\em Spacewarps\/}
-volunteers are quite experienced from earlier projects, having spent a
-thousand hours or more with data, and are very interested in more
-demanding projects.  The interests of citizen-science communities are
-just beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but
-it is clear that some volunteers welcome open-ended challenges, and
-sometimes these have led to new scientific results: one example is the
-discovery of an exceptional extra-solar planet
-\citep{2013ApJ...768..127S}; another is the development of new
-algorithms for protein folding \citep{Khatib22112011}.  All these are
-grounds for optimism.  There is, however, a basic difficulty in strong
-gravitational lensing.  Lensed images do not look much like their
-source, and still less do they resemble the lensing-mass distribution.
-To model a lens, one needs to either do a lot of random guessing, or
-have a good intuition for what works.
+The encouraging early results prompted the question: could modelling
+of the lenses also be done by volunteers?  If so, modelling could help
+assess prioritise lens candidates at an early stage, which would be
+very useful with new wide-field and sensitive surveys, which will
+yield thousands of lens candidates. There are several software tools
+for lensing modelling available, and work has been done on generic
+interfaces \citep{2014A&C.....5...28L}.  Some early designs for \sw
+included a prototype lens modelling tool \citep{2010AAS...21543527N}.
+Moreover, some {\em Spacewarps\/} volunteers are quite experienced
+from earlier projects, having spent a thousand hours or more with
+data, and are very interested in more demanding projects.  The
+interests of citizen-science communities are just beginning to be
+studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that
+some volunteers welcome open-ended challenges, and sometimes these
+have led to new scientific results: one example is the discovery of an
+exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
+the development of new algorithms for protein folding
+\citep{Khatib22112011}.  All these are grounds for optimism.  There
+is, however, a basic difficulty in strong gravitational lensing.
+Lensed images do not look much like their source, and still less do
+they resemble the lensing-mass distribution.  To model a lens, one
+needs to either do a lot of random guessing, or have a good intuition
+for what works.
 
@@
 
-\subsection{Diagnostics}
+\subsection{Diagnostics} \label{sec:diag}
 
@@
 of \spl (RK, who also chose the challenge set) or to the modellers
-(EB, CC, CM, JO, PSand JW) until the modelling stage was done.
+(EB, CC, CM, JO, PS and JW) while modelling was in progress.  After
+the modelling stage, AM released the details of the sims for
+post-modelling analysis.  Results from the latter now follow.
 
@@
 
-Of the 119 models proposed, we discuss now present eight examples in
-some detail.  Results from these are shown in Figures
+Of the 119 models proposed, we now present eight examples in some
+detail.  Results from these are shown in Figures
 \ref{fig:6941}--\ref{fig:7022}.  The first four of these show the most
@@
 output returned to the modeller by \spl, listed as (i), (ii), (iii) in
-\secref{SpaghettiLens}.  These derive from the mean of an ensemble
-of 200 models generated by \spl.
+\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
+200 models generated by \spl.
+
+We should remark that the scales and alignments in the six panels are
+not completely consistent: the three panels on the right exactly the
+same coordinates.  The middle-left and lower-left panels are
+consistent with each other, but only approximately aligned with the
+right column, becuse the latter depends on inexact mouse placements.
+The upper-left panel is zoomed out slightly, to accommodate all the
+markup.  For qualitative features, which is all that
+Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, these small
+inconsistencies are harmless.
 
@@
 
-Figure \ref{fig:6941} shows a simple example.  In summary: the
-identification of minimum and saddle point is correct, but the
-estimated Einstein radius is a little too high.
+Figure \ref{fig:6941} shows the simplest case, with two clear images
+produced by a nearly-circular lens.  The centre of the lensing galaxy
+is a maximum, the image nearer to the galaxy is a saddle point, and
+image further away is a minimum, and these were correctly identified.
+The model synthetic images is very crude, because there is no fitting
+for the source light distribution, as already noted in \secref{diag}.
 

### 2014-06-27 text/partC.tex
typo fixes

--- 2014-06-06 text/partC.tex
@@
 
-The presentation of the arrival-time surface in \S\secref{Fermat}
-and \secref{arriv} omitted some details for the sake of a more
-intuitive explanation.  Specifically, the geometric time delay $t_{\rm
-geom}$ and the convergence $\kappa$ were left as proportionalities
-(equations
+The presentation of the arrival-time surface in
+Sections~\ref{sec:Fermat} and \ref{sec:arriv} omitted some details for
+the sake of a more intuitive explanation.  Specifically, the geometric
+time delay $t_{\rm geom}$ and the convergence $\kappa$ were left as
+proportionalities (equations
 \ref{eq:tgeom} and \ref{eq:kappa}), and the gravitational time delay

### 2014-06-27 text/partA.tex
typo fixes
small change
edits

@@
 of \spl (RK, who also chose the challenge set) or to the modellers
-(EB, CC, CM, JO, PS and JW) while modelling was in progress.  After
-the modelling stage, AM released the details of the sims for
-post-modelling analysis.  Results from the latter now follow.
+(EB, CC, CM, JO, PS and JW) while modelling was in progress.  That is,
+the modellers had no advance knowledge of what kind of
+parameterisation had been used to make the sims.  After the modelling
+stage, AM released the details of the sims for post-modelling
+analysis.  Results from the latter now follow.
 
@@
 
-Of the 119 models proposed, we now present eight examples in some
+Of the 119 models proposed, we now discuss eight examples in some
 detail.  Results from these are shown in Figures
@@
 \secref{SpaghettiLens}.  These derive from the mean of an ensemble of
-200 models generated by \spl.
-
-We should remark that the scales and alignments in the six panels are
-not completely consistent: the three panels on the right exactly the
-same coordinates.  The middle-left and lower-left panels are
-consistent with each other, but only approximately aligned with the
-right column, becuse the latter depends on inexact mouse placements.
-The upper-left panel is zoomed out slightly, to accommodate all the
-markup.  For qualitative features, which is all that
-Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, these small
-inconsistencies are harmless.
+200 models generated by \spl. The model synthetic image is too crude
+to be useful, because there is no fitting for the source light
+distribution, as already noted in \secref{diag}.  The quality of the
+synthetic images can be improved by interpolation, but as this feature
+was not available during the modelling challenge, we do not show the
+improved versions here.
 
@@
 
+In these figures, the scales and alignments in the six panels are not
+completely consistent: the three panels on the right exactly the same
+coordinates; the middle-left and lower-left panels are consistent with
+each other, but only approximately aligned with the right column,
+because the latter are affected by errors in mouse placements; the
+upper-left panel is zoomed out slightly, to accommodate the full
+markup.  For qualitative features, which is all that
+Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, the slightly
+varying scales and alignments are harmless.
+
 Figure \ref{fig:6941} shows the simplest case, with two clear images
@@
 image further away is a minimum, and these were correctly identified.
-The model synthetic images is very crude, because there is no fitting
-for the source light distribution, as already noted in \secref{diag}.
+  It shows wide arcs, almost a ring.
 
@@
 surface that a point source is a double on the verge of splitting into
-a quad.  The modeler interpreted the system as a quad.  The
-appearance of the arcs, shown in the bottom panel of Figure
-\ref{fig:input-spag}, looks like an arc and counter-image such as
-discussed with Figure \ref{fig:6990} above.  But there is an important
+a quad.  The modeler interpreted the system as a quad.  The appearance
+of the arcs, looks like an arc and counter-image such as discussed
+with Figure \ref{fig:6990} above.  But there is an important
 difference: the long arc is closer to the galaxy, as if the arc and

### 2014-06-27 text/partA.tex
edits

@@
 
-Figure \ref{fig:6941} shows the simplest case, with two clear images
-produced by a nearly-circular lens.  The centre of the lensing galaxy
-is a maximum, the image nearer to the galaxy is a saddle point, and
-image further away is a minimum, and these were correctly identified.
-  It shows wide arcs, almost a ring.
+Let us now consider these cases in turn.
 
-Figure \ref{fig:6915} shows another quad.  This kind of configuration
-arises when the mass is elongated and the source is displaced at an
-angle to the elongation.
+\begin{itemize}
 
-Figure \ref{fig:6990} shows an example of an arc that has split into
-three images.  This kind of configuration, with a counter-image close
-to the lensing galaxy and a more distant arc/triplet on the other
-side, generically arises from an elongated mass distribution when the
-source is displaced along the elongated direction.
+\item Figure \ref{fig:6941} shows the simplest case, with two clear
+  images produced by a nearly-circular lens.  The centre of the
+  lensing galaxy is a maximum, the image nearer to the galaxy is a
+  saddle point, and image further away is a minimum, and these were
+  correctly identified.
 
-To conclude this set of examples, Figure \ref{fig:6919} shows another
-type of quad.  Actually, in this case the brightest part of the source
-is only doubly imaged, but the source extends into a region that
-produces four images.  We can also see from the real arrival-time
-surface that a point source is a double on the verge of splitting into
-a quad.  The modeler interpreted the system as a quad.  The appearance
-of the arcs, looks like an arc and counter-image such as discussed
-with Figure \ref{fig:6990} above.  But there is an important
-difference: the long arc is closer to the galaxy, as if the arc and
-counter-image have swapped roles.  This configuration arises if the
-source displacement is perpendicular to the long axis of the lensing
-mass.
+\item Figure \ref{fig:6990} shows an example of an arc that has split
+  into three images.  This kind of configuration, with a counter-image
+  close to the lensing galaxy and a more distant arc/triplet on the
+  other side, generically arises from an elongated mass distribution
+  when the source is displaced along the elongated direction.
 
-Figure \ref{fig:6975} shows a lens with substructure in the form of a
-smaller secondary galaxy.  The galaxies in such group or cluster sims
-were, in fact, based on galaxies visible in the images --- but the
+\item Figure \ref{fig:6919} shows another example of an arc plus
+  counter-image, but (in contrast to Figure \ref{fig:6990}) the arc is
+  closer to the lens than the counter-image. This configuration arises
+  if the source displacement is perpendicular to the long axis of the
+  lensing mass.  Comparing the two panels in the middle row, we see
+  that the modeller interpreted arc as consisting of three images,
+  whereas the sim shows a single saddle point associated with the arc.
+  But the identification is not really erroneous --- we just need to
+  take into account that the source is extended.  In fact, in the sim
+  the brightest part of the source is only doubly imaged, but the
+  source extends into a region that produces four images.  In the
+  $t(x,y)$ of the sim, the hairpin-bending contours are typical of
+  double on the verge of splitting into a quad.
+
+\item Figure \ref{fig:6915} shows another quad.  This kind of
+  configuration arises when the mass is elongated and the source is
+  displaced at an angle to the elongation.  The minima and saddle
+  points are correctly identified, and the orientation of the
+  ellipticity of the mass distribution is correctly reproduced.
+
+\item Figure \ref{fig:6975} shows a lens with substructure in the form
+  of a smaller secondary galaxy.  The galaxies in such group or
+  cluster sims were based on galaxies visible in the images, but the
   modelers were not told in advance whether this was the case.  The
-model does not include any substructure, but otherwise is not bad.
-The minimum and saddle point are correct, and the Einstein radius is
-only a little underestimated.
+  minimum and saddle point are correctly identified.  The mass
+  distribution misses the substructure, but overall appears
+  reasonable.
 
-Figure \ref{fig:6937} shows a case where substructure leads to a
-poor model.
-
-Figure \ref{fig:7022} shows another model of the same system.  In this
-one, the identification of the minima and saddle points was incorrect,
-and mass distribution comes out elongated East-West instead of
+\item Figure \ref{fig:7025} shows a quad.  In this one, the
+  identification of the minima and saddle points was incorrect, and
+  mass distribution comes out elongated East-West instead of
   North-South.  The mass distribution also appears somewhat jagged and
   the saddle-point contours are not as clean as in the previous
-examples; these are often indicators of a problem with the model.  The
-enclosed mass is, however, none the worse --- the reason is probably
-that in a relatively symmetrical image configuration, the Einstein
-radius is quite well constrained by the images in a fairly
+  examples; these are often indicators of a problem with the model.
+  The enclosed mass is, however, none the worse --- the reason is
+  probably that in a relatively symmetrical image configuration, the
+  Einstein radius is quite well constrained by the images in a fairly
   model-independent way.
 
-Figure \ref{fig:6975} shows a fairly symmetric quad.  The minima and
-saddle points are correctly identified, and the orientation of the
-ellipticity of the mass distribution is correctly reproduced.  The
-Einstein radius is somewhat overestimated.
+\item Figure \ref{fig:7022} shows another model of the same system.
+
+\end{itemize}
+
+\begin{figure}
+  \centering
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007025_kappa_encl}}
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007022_kappa_encl}}
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007021_kappa_encl}} \\
+  \caption{Enclosed mass \kenc from three models of the same system.
+    The upper panel corresponds to the model shown in \figref{7025},
+    with the image-parities incorrect by a quarter-turn.  The middle
+    corresponds to \figref{7022}, where the image parities were
+    correct.  The lower panel corresponds to another model, where the
+    image parities were incorrect by a half-turn.}
+\label{fig:kapenc_compare_faulty}
+\end{figure}
 
@@
 
-The panel shows
-a average $\kappa$ within a given radius, as a function of radius.
-The red curve is the true value, and where it crosses unity (dotted
-horizontal line) is the notional Einstein radius $\Theta_{\text{E, sim}}$.
-  The two blue curves
-are the minimal and maximal mean enclosed $\kappa$ from the internal
-ensemble in \spl.  The region between the blue curves is shaded
-between the radii of the innermost and the outermost images --- this
-is the confidence region from the modeling.  As we see, the shaded
-blue is slightly above the red curve. 
-The Einstein Radius $\Theta_\text{E}$ of the model is estimated crossing the
-mean enclosed $\kappa$ (not plotted) with unity.
+The panel shows a average $\kappa$ within a given radius, as a
+function of radius.  The red curve is the true value, and where it
+crosses unity (dotted horizontal line) is the notional Einstein radius
+$\Theta_{\text{E, sim}}$.  The two blue curves are the minimal and
+maximal mean enclosed $\kappa$ from the internal ensemble in \spl.
+The region between the blue curves is shaded between the radii of the
+innermost and the outermost images --- this is the confidence region
+from the modeling.  As we see, the shaded blue is slightly above the
+red curve.  The Einstein Radius $\Theta_\text{E}$ of the model is
+estimated crossing the mean enclosed $\kappa$ (not plotted) with
+unity. \citep[see, e.g.,][]{2002LNP...608....1C}.

### 2014-06-28 text/partB.tex
edits

@@
 
-A first manual evaluation tested the volunteers ability to reconstruct the arrival time surface given a survey image containing a sim.
-This task consists of two parts.
-First to correctly identify and locate the lensed images.
-Second the correct ordering for the identified lensed images in respect of the arrival time.
+The first post-modelling test was a qualitative comparison of the
+original arrival-time surfaces and the input spaghetti diagrams given
+by the modeller.  This tested first, for correct identification and
+location of the lensed images, and second, the correct parities and
+ordering for the lensed images in respect of the arrival time.
 
-While we expected the identification of lensed images to be trivial, given the nature of the survey images and the success of \sw, we expect the correct ordering to be more difficult.
-This part tests the volunteers understandings of the theory of arrival time surfaces and the odd number theorem.
-While we can provide the volunteers with some general rules of thumb, ordering involves imagination and guessing and therefore training could improve results in a later stage.
+While we expected the identification of lensed images to be trivial,
+given the generally clean appearance of the sims in the test, we
+expected the parities and time-ordering to be more difficult.  While
+the \spl tutorials had provided some general rules of thumb, to be
+consistently correct with the time-ordering, a modeller needs to
+develop some intuition for arrival time surfaces.  This is an area
+where experience and tutorials training could improve results at a
+later stage, and correspondingly, feedback on the difficulties
+modellers encounter can help improve the tutorial materials.
 
-This first test was designed to give some feedback on the difficulties volunteers encounter, to further improve the tutorial materials.
+\tabref{stats} presents a summary of the test.  The evaluation was
+done manually, comparing the input to \spl with the actual
+arrival-time surface of the sim.  This amounts to comparing the
+upper-left and middle-left panels in each of
+Figures~\ref{fig:6941}--\ref{fig:7022}, and similaly the other 113
+models.
 
-The evaluation of the volunteers performance was done manually, comparing their input from \spl and the resulting reconstructed arrival time surface contour line plot (arrival plot) to the arrival plot generated using simulation parameters.
-% \Figref{output_compare} shows the setup used to evaluate the models.
+\begin{table}\centering\begin{tabular}{llcc}\hline
+     & Total & 119 \\
+\hline
+ R1: & images approximately on correct locations & 110 \\
+ R2: & images parities and time-ordering correct & 70 \\
+\hline
+ E01: & inaccurate image placement over an arc & 21 \\
+ E02: & swapped minimum and saddle in double  & 2 \\
+ E03: & identified two images of four & 5 \\
+ E04: & modelled a three-image arc as one image & 4 \\
+ E05: & swapped early and late saddles in a quad & 7 \\
+ E06: & swapped minima and saddles in quad & 38 \\
+ E07: & missed faint images & 1 \\
+ E08: & modelled one image as a three-image arc & 5 \\
+ E09: & identified two nearby images as one & 3 \\
+ E10: & proposed too many images & 1 \\
+\hline
+\end{tabular}
+\caption{Table of image-identification errors and the number of models
+  containing each.}
+\label{tab:stats}
+\end{table}
 
@@
 
-\begin{table}\centering\begin{tabular}{llcc}\hline
- & & n & p \\
-\hline
- N: & Total number of models & 119 & 1.00\\
-\hline
- R1: & images approx. on right location & 110 & 0.92\\
- R2: & images with correct parity & 70 & 0.59\\
-\hline
- E01: & inaccurate in arc & 21 & 0.18\\
- E02: & wrong parity in 3 lens conf. & 2 & 0.02\\
- E03: & identified 3 of 5 imgs. & 5 & 0.04\\
- E04: & modeled arc with single img. & 4 & 0.03\\
- E05: & $\pi$ rotated parity & 7 & 0.06\\
- E06: & $\pi$/2 rotated parity & 38 & 0.32\\
- E07: & missed faint img. & 1 & 0.01\\
- E08: & too many imgs in arc. & 5 & 0.04\\
- E09: & missed double img & 3 & 0.03\\
- E10: & too many imgs. & 1 & 0.01\\
-\end{tabular}
-
-\caption{1. col: id of model; 2.col: places of imgs idendited
-  correctly; 3. col: more or less correct identification of extr
-  points. 4.col: exact identification of extremal points. 5.col:
-  type(s) of errors ocured....  1) inaccurate placement in an extended
-  arc; 2) wrongly identified sad and min in 3 image configuration; 3)
-  identified only 3 instead of 5 images; 4) tried to model an arc with
-  a min instead of min-sad-min; 5) PI-err (rotation by 180 degrees; in
-  5 image configuration, exchanged the ordering of the two saddle
-  points); 6) PI/2-err (rotation by 90 degree; sad-min-sad-min-sad);
-  7) missed faint image(s); 8) tried to model an arc with min-sad-min
-  instead of only min; 9) did identify two close by images as one; 10)
-  used 7 or more image to model a 5 image system. }
-
-\label{tab:stats}
-\end{table}
-
-\tabref{stats} presents a summary of this evaluation.
 We conclude that the volunteers are performing very well identifying and positioning images, with a performance of 92\% (R1, p=0.92).
@@
 
-\begin{figure*}
-  \centering
-    \includegraphics[width=0.90\linewidth]{fig/eR_5}
-  \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
-  \label{fig:ER_per_sim}
-\end{figure*}
-
 \begin{figure}
@@
 \end{figure}
+
+\begin{figure*}
+  \centering
+    \includegraphics[width=0.90\linewidth]{fig/eR_5}
+  \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
+  \label{fig:ER_per_sim}
+\end{figure*}
 

### 2014-06-29 text/partA.tex

@@
     The upper panel corresponds to the model shown in \figref{7025},
-    with the image-parities incorrect by a quarter-turn.  The middle
-    corresponds to \figref{7022}, where the image parities were
-    correct.  The lower panel corresponds to another model, where the
-    image parities were incorrect by a half-turn.}
+    in which the minima and saddle-points have been incorrectly
+    swapped.  The middle corresponds to \figref{7022}, where the image
+    parities were correct.  The lower panel corresponds to another
+    model, where the image parities were correct but the time-ordering
+    was incorrect.  See text of Section~\ref{sec:example_models} for
+    details.}
 \label{fig:kapenc_compare_faulty}
@@
 
-Figure \ref{fig:kapenc_compare_faulty} shows a comparison of input and
-recovered mass profiles.  The panel shows a average $\kappa$ within a
-given radius, as a function of radius.  The red curve is the true
-value, and where it crosses unity (dotted horizontal line) is the
-notional Einstein radius $\Theta_{\text{E, sim}}$.  The two blue
-curves are the minimal and maximal mean enclosed $\kappa$ from the
-internal ensemble in \spl.  The region between the blue curves is
-shaded between the radii of the innermost and the outermost images ---
-this is the confidence region from the modeling.  As we see, the
-shaded blue is slightly above the red curve.  The Einstein Radius
-$\Theta_\text{E}$ of the model is estimated crossing the mean enclosed
-$\kappa$ (not plotted) with unity.
+\figref{kapenc_compare_faulty} compares the circularly
+averaged mass profiles for one input lens and three different models
+of it.  Each panel shows the mean convergence $\kappa$ within a
+circle of given radius, as a function of the radius.  The red curve is
+the true value.  The two blue curves are the minimal and maximal mean
+enclosed $\kappa$ from the internal ensemble in \spl.  Radial
+locations of the images are marked, along with the image parities.
+The region between the blue curves is shaded between the radii of the
+innermost and the outermost images: this is the confidence region from
+the modelling.
 
-The panel shows a average $\kappa$ within a given radius, as a
-function of radius.  The red curve is the true value, and where it
-crosses unity (dotted horizontal line) is the notional Einstein radius
-$\Theta_{\text{E, sim}}$.  The two blue curves are the minimal and
-maximal mean enclosed $\kappa$ from the internal ensemble in \spl.
-The region between the blue curves is shaded between the radii of the
-innermost and the outermost images --- this is the confidence region
-from the modeling.  As we see, the shaded blue is slightly above the
-red curve.  The Einstein Radius $\Theta_\text{E}$ of the model is
-estimated crossing the mean enclosed $\kappa$ (not plotted) with
-unity. \citep[see, e.g.,][]{2002LNP...608....1C}.
+In a circular lens, the Einstein radius has averaged enclosed $\kappa$
+of unity.  For a general, non-circular lens, there is no standard
+definition of an Einstein radius.  We adopt a notional Einstein radius
+$\Theta_E$ as the radius at which the mean enclosed $\kappa$ within a
+circle is unity.  In \figref{kapenc_compare_faulty}, that would be
+where the mean convergence crosses the dashed horizontal line.  Where
+the red curve crosses $10^0$ is the actual Einstein radius
+$\Theta_{E\rm,act}$.  The recovered Einstein radius
+$\Theta_{E\rm,rec}$ and its uncertainty are given by the blue curves
+crossing $10^0$.  We see that in all three panels, the blue curves are
+shallower than the red curve and $\Theta_{E\rm,rec}$ is more than
+$\Theta_{E\rm,act}$. by more than the model uncertainties.  Now,
+steeper mass profiles tend to give wider image separations --- recall
+that the image separation for a circular isothermal lens is
+$2\Theta_E$, whereas for a point mass it is more \citep[see,
+  e.g.,][]{2002LNP...608....1C} --- so $\Theta_{E\rm,rec}$ being too
+high is really a consequence of the \spl models being too shallow for
+the sims.

### 2014-06-30 text/partA.tex
more edits

--- 2014-06-27 text/partA.tex
@@
 mass distribution, but the enclosed mass is more robust.  Shortcomings
-include (a)~lens profiles that are too shallow, (b)~over-estimation of
-the Einstein radius by $\sim25\%$, (c)~lack of detailed source
-modelling.  Ideas for improvement are discussed.
+include (a)~lens profiles that tend to be systematically shallower
+than the lenses, and consequent over-estimation of the Einstein radius
+by $\sim25\%$, (c)~lack of detailed source modelling.  Ideas for
+improvement are discussed.
 \end{abstract}
@@
 \end{itemize}
-
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.8\linewidth]
-             {fig/007025_kappa_encl}}
-  \subfigure{\includegraphics[width=0.8\linewidth]
-             {fig/007022_kappa_encl}}
-  \subfigure{\includegraphics[width=0.8\linewidth]
-             {fig/007021_kappa_encl}} \\
-  \caption{Enclosed mass \kenc from three models of the same system.
-    The upper panel corresponds to the model shown in \figref{7025},
-    with the image-parities incorrect by a quarter-turn.  The middle
-    corresponds to \figref{7022}, where the image parities were
-    correct.  The lower panel corresponds to another model, where the
-    image parities were incorrect by a half-turn.}
-\label{fig:kapenc_compare_faulty}
-\end{figure}
-
-Figure \ref{fig:kapenc_compare_faulty} shows a comparison of input and
-recovered mass profiles.  The panel shows a average $\kappa$ within a
-given radius, as a function of radius.  The red curve is the true
-value, and where it crosses unity (dotted horizontal line) is the
-notional Einstein radius $\Theta_{\text{E, sim}}$.  The two blue
-curves are the minimal and maximal mean enclosed $\kappa$ from the
-internal ensemble in \spl.  The region between the blue curves is
-shaded between the radii of the innermost and the outermost images ---
-this is the confidence region from the modeling.  As we see, the
-shaded blue is slightly above the red curve.  The Einstein Radius
-$\Theta_\text{E}$ of the model is estimated crossing the mean enclosed
-$\kappa$ (not plotted) with unity.
-
-The panel shows a average $\kappa$ within a given radius, as a
-function of radius.  The red curve is the true value, and where it
-crosses unity (dotted horizontal line) is the notional Einstein radius
-$\Theta_{\text{E, sim}}$.  The two blue curves are the minimal and
-maximal mean enclosed $\kappa$ from the internal ensemble in \spl.
-The region between the blue curves is shaded between the radii of the
-innermost and the outermost images --- this is the confidence region
-from the modeling.  As we see, the shaded blue is slightly above the
-red curve.  The Einstein Radius $\Theta_\text{E}$ of the model is
-estimated crossing the mean enclosed $\kappa$ (not plotted) with
-unity. \citep[see, e.g.,][]{2002LNP...608....1C}.

### 2014-06-30 text/partB.tex
more edits

@@
 upper-left and middle-left panels in each of
-Figures~\ref{fig:6941}--\ref{fig:7022}, and similaly the other 113
+Figures~\ref{fig:6941}--\ref{fig:7022}, and similarly the other 111
 models.
 
-\begin{table}\centering\begin{tabular}{llcc}\hline
-     & Total & 119 \\
+\begin{table}\centering\begin{tabular}{lr}\hline
+    Total & 119 \\
 \hline
- R1: & images approximately on correct locations & 110 \\
- R2: & images parities and time-ordering correct & 70 \\
+    errors in image locations  & 9 \\
+    errors in image parities or time ordering & 49 \\
 \hline
- E01: & inaccurate image placement over an arc & 21 \\
- E02: & swapped minimum and saddle in double  & 2 \\
- E03: & identified two images of four & 5 \\
- E04: & modelled a three-image arc as one image & 4 \\
- E05: & swapped early and late saddles in a quad & 7 \\
- E06: & swapped minima and saddles in quad & 38 \\
- E07: & missed faint images & 1 \\
- E08: & modelled one image as a three-image arc & 5 \\
- E09: & identified two nearby images as one & 3 \\
- E10: & proposed too many images & 1 \\
+    inaccurate image placement over an arc & 21 \\
+    identified two images of four & 5 \\
+    identified two nearby images as one & 3 \\
+    missed faint images & 1 \\
+    proposed too many images & 1 \\
+    modelled a three-image arc as one image & 4 \\
+    modelled one image as a three-image arc & 5 \\
+    swapped minimum and saddle in double  & 2 \\
+    swapped minima and saddles in quad & 38 \\
+    swapped early and late saddles in a quad & 7 \\
 \hline
@@
 \caption{Table of image-identification errors and the number of models
-  containing each.}
+  containing each.  A model can contain more than one type of error.}
 \label{tab:stats}
@@
 
-The images of the system are considered to be identified correctly, if all the images have been identified and are approximated within $\pm0.05\cdot\text{imgage width}$.
-The parity is considered correct, if those identified points have the right ordering with respect to arrival time.
+The images of the system are considered to have been located
+correctly, if all the images have been identified and are approximated
+within about 5\% of \spl frame used to draw the spaghetti diagram.
+That frame size is adjustable by the user, but in practice it is
+somewhat larger than the spaghetti diagram.  Such image-placement
+errors were found in only 9 models.  That does not include inaccurate
+image placement over an arc, which was considered a separate category
+of error.
 
-Additionally, ten types of errors (labeled E01 -- E10, listed in \tabref{stats}) that occurred were identified.
-Each generated model could contain more than one error.
+In addition to simple image-placement errors, ten types of errors were
+recognised and are listed in \tabref{stats}.  Most of the problems
+where due to unclear arc-like structures.  Critical errors like the
+failure to identify all five images in a five images system, or to
+include too many images, were rare.
 
-We conclude that the volunteers are performing very well identifying and positioning images, with a performance of 92\% (R1, p=0.92).
-Most of the problems where due to unclear arc-like structures (E01, p=0.18; E04, p=0.03; E08, p=0.04).
-Critical errors like the failure to identify all five images in a five images system (E03, p=0.04) or to include too many images (E10, p=0.01) did almost never happen.
-From this we conclude that the introduction materials was adequate and the volunteers understand the basics of gravitational lensing.
+The assignment of the parity of the images was a more difficult task,
+and was successful in only 60\% of the cases.  The most common error
+was swapping of minima and saddle-points in a quad; \figref{7025}
+shows an example.  Another, less common, error was flipping the
+spaghetti diagram, this swapping the time-ordering of the two saddle
+points.
 
-The assignment of the parity of the images was a more difficult task.
-In 59\% (R2, p=0.59, N=70) of the cases the volunteers succeeded to identify the right configuration.
-Most of the failures are due to E06 (N=38, p=32\%), followed by E05 (N=7, p=6\%).
-E06 describes a situation, where the minima and saddle points of a five image configuration were exchanged (rotated by $\pm90\dgr$), see \figref{7022} for an example.
-E05 describes a situation, where the ordering of the saddle points was wrong (rotation by $180\dgr$).
-While these errors occurred, we suspect they can be avoided with better training material and some examples for the obvious cases.
-For more challenging cases, like very symmetrical distribution of the lensed images (for example model 7022, \figref{7022}), those errors should still produce plausible results, as will be explained in the next section.
+Incorrect image parities and time orderings tended to produce
+poorer-looking models, such as the checkboard patterns in the mass map
+in \figref{7025}.  Interestingly, however, the enclosed-mass profiles
+were quite robust.  We will consider this aspect in the next section.
+
+\FloatBarrier
+
+\begin{figure}
+  \centering
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007025_kappa_encl}}
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007022_kappa_encl}}
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007021_kappa_encl}} \\
+  \caption{Mean $\kappa$ inside a circle around the lens centre, as a
+    function of the radius of the circle.  (See \secref{tests.t2} for
+    details.) The upper panel corresponds to the model shown in
+    \figref{7025}, in which the minima and saddle-points have been
+    incorrectly swapped.  The middle corresponds to \figref{7022},
+    where the image parities were correct.  The lower panel
+    corresponds to another model, where the image parities were
+    correct but the time-ordering was incorrect.}
+\label{fig:kapenc_compare_faulty}
+\end{figure}
 
@@
 
-The second test was to compare the mass distributions of the lens $\kappa(x, y)$ given by the sims and generated by the volunteers.
-To get a means of comparing the sims to the models, the total convergence\footnote{often called enclosed mass} $\kappa_{\text{encl}}(r)$ was calculated for both.
-The Einstein radius $\Theta_\text{E}$ is defined by $\kappa_{\text{encl}}(\Theta_\text{E})=1$ and gives a number that allows a rough comparison between a sim and a model.
-We also let an expert (PS) model three selected systems to compare the results from volunteers to those of a professional.
+The second test was to compare the mass distributions $\kappa(x,y)$ of
+the sims and of the \spl models.  A visual comparison is presented for
+the eight models in Figures~\ref{fig:6941}--\ref{fig:7022}, in the
+lower-left versus lower-right panels.  We will summarize the mass
+distributions drastically in a single number, the effective Einstein
+radius.  Other measures for comparison of free-form lensing mass
+distributions appear in \cite{2014arXiv1401.7990C}, but comparing
+Einstein radii is already useful.
 
-To compare the enclosed mass profile and the Einstein radius of the simulation and the models, \kenc was calculated using the mass map \kap[x,y] generated in the modeling process.
-From the ensemble of models generated by one modeling process, the mean is taken as the resulting $\kappa(r)$ to calculate \ERg.
-To estimate the errors, the extremal models are used to estimate a lower and upper limit for \ERg.
-These results can be seen in \figref{ER_per_sim}.
-This figure shows that this technique of estimating the error using the ensemble of all models underestimates the error significantly and should be improved for further analysis.
+There is no standard way of defining the Einstein radius of a general
+non-circular lens.  We adopt the simple definition
+\begin{equation} \label{eq:effRE}
+ \langle\kappa\rangle_{\Theta_E} = 1,
+\end{equation}
+that is, the effective Einstein radius \ERf is such that the
+mean $\kappa$ is unity inside a circle of radius \ERf centred at
+the lens centre.
 
-\begin{figure}
-  \centering
-  \subfigure{\includegraphics[width=0.9\linewidth]
-             {fig/007022_kappa_encl}}
-  \subfigure{\includegraphics[width=0.9\linewidth]
-             {fig/007025_kappa_encl}}
-  \subfigure{\includegraphics[width=0.9\linewidth]
-             {fig/007021_kappa_encl}} \\
-  \caption{
-  %Enclosed mass \kenc from three models of a system with incorrect image identifications.
-\figref{7022} shows a model (by an expert) with the correct identifications, while \figref{7025} shows a model with error E06 (parities rotated by $\pi/2$).
-Shown here are the \kenc results for two models with error E06 and one model (bottom) with error E05 (parities rotated by $\pi$).
-  }
-\label{fig:kapenc_compare_faulty}
-\end{figure}
+To illustrate, \figref{kapenc_compare_faulty} compares the circularly
+averaged mass profiles three different models of one particular lens;
+two of the models are shown in Figures~\ref{fig:7025} and
+\ref{fig:7022}. Each panel in \figref{kapenc_compare_faulty} shows the
+mean $\kappa$ within a circle of given radius.  The red curve is the
+correct profile for the sim.  The two blue curves are the minimal and
+maximal mean enclosed $\kappa$ from the internal ensemble in \spl.
+Radial locations of the images are marked, along with the image
+parities.  The region between the blue curves is shaded between the
+radii of the innermost and the outermost images: this is the
+confidence region from the modelling.  The definition (\ref{eq:effRE})
+for \ERf corresponds to crossing the dashed horizontal line at
+$10^0$: the red curve crosses the dashed line at the actual Einstein
+radius \ERf[,act]; the recovered Einstein radius
+\ERf[,rec] and its uncertainty are given by the blue curves
+crossing the dashed line.  We see that in all three panels, the blue
+curves are shallower than the red curve and \ERf[,rec] is
+more than \ERf[,act], by more than the model uncertainties.
+Now, steeper mass profiles tend to give wider image separations ---
+recall that the image separation for a circular isothermal lens is
+$2\Theta_E$, whereas for a point mass it is more \citep[see,
+  e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
+high is really a consequence of the \spl models being too shallow for
+the sims.
 
@@
     \includegraphics[width=0.90\linewidth]{fig/eR_5}
-  \caption{Relative \ERf / \ERg[, sim] for models by volunteers (blue cross), models made by an expert (red cross with offset), including rejected models (green squares); binned per sim.}
+  \caption{Model-recovered versus actual Einstein radii \ERf[,rec] and
+    \ERf[,act].  Circles denote models where the image parities and
+    time-orderings were correct, with filled circles for models by an
+    expert.  Crosses mark models where the time-ordering was
+    incorrect.  Plus signs indicate models flagged by the modeller as
+    failures.}
   \label{fig:ER_per_sim}
@@
 
-\Figref{ER_per_sim} shows that the calculated \ERf of the models tend
-to be too high.  The overshoot varies from around 0.2 to 0.4 for good
-models.  Discounting the models which were flagged by the users as
-poor, \ERf{E,rec} is overestimated.  The mean and standard deviation
-of the overestimate are 25\% and 20\%.  The models by an expert also
-show a similar bias, having mean and standard deviation of the
-overestimate being 20\% and 8\%.
-One of the reasons for this is that it is hard to get the center of the lens on spot.
-An offset leads to a flatter mass profile for the model compared to the simulation.
+\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend
+to be too high.  Discounting the models which were flagged by the
+users as poor, the mean and standard deviation of the overestimate are
+25\% and 20\%.  The models by an expert (PS) also show a similar bias,
+having mean and standard deviation of the overestimate being 20\% and
+8\%.  One of the reasons for this is that it is hard to get the center
+of the lens on spot.  An offset leads to a flatter mass profile for
+the model compared to the simulation.
 
-Comparing the models from volunteers and experts can be done in Figures \ref{fig:7022}, \ref{fig:7025} and \ref{fig:kapenc_compare_faulty}, where only the
-expert got the right configuration (\figref{7022}) but enclosed masses are all fairly similar.
-In cases with nearly symmetric image configurations, like this example, it is more difficult to identify the image parities correctly.
-Incorrect image-parity identification changes the orientation of the mass distribution \kap[x,y],
-but \kenc and thus the \ERf are not influenced that much.
-
-Two additional sims were modeled by an expert, \asw{1hpf} and \asw{0vqg}.
-Looking at the results for \ERg for those models in \figref{ER_per_sim}, we conclude that the performance of volunteers (blue crosses) and experts (red crosses, offset) is comparable.
-%Note that the models of \asw{0vqg} with \ERg[, rel] around 0.25 (6935 -- 6937) are failed models that show the attempts of a single user, that came finally up with model 6938 as final result.\todo{remove?}
-
-\clearpage
 \section{Outlook} \label{sec:todo}
@@
 is good, not only for identifying lens candidates for follow-up, but
-modeling candidates as well.
+modelling candidates as well.
 

### 2014-06-30 text/ms.tex
partA partB partC back in ms.tex

@@
-% \documentclass[usenatbib,onecolumn]{mn2e}
 \documentclass[usenatbib]{mn2e}
@@
 % shortcut for einstein radius (Text Greek Full Math)
-\newcommand{\ER}{Einstein radius\xspace} % text
-\newcommand{\ERg}[1][]{$\Theta_{\text{E#1}}$\xspace} % er in textmode with greek
 \newcommand{\ERf}[1][]{Einstein radius $\Theta_\text{E#1}$\xspace} % full output
-\newcommand{\ERm}[1][]{\Theta_\text{E#1}} % mathmode with greek symbol
+
+%\newcommand{\ER}{Einstein radius\xspace} % text
+%\newcommand{\ERg}[1][]{$\Theta_{\text{E#1}}$\xspace} % er in textmode with greek
+%\newcommand{\ERm}[1][]{\Theta_\text{E#1}} % mathmode with greek symbol
 
 %shortcuts for kappa
-\newcommand{\kenc}[1][r]{$\kappa_\text{encl}(#1)$\xspace}
-\newcommand{\kap}[1][r]{$\kappa(#1)$\xspace}
+%\newcommand{\kenc}[1][r]{$\kappa_\text{encl}(#1)$\xspace}
+%\newcommand{\kap}[1][r]{$\kappa(#1)$\xspace}
 
@@
 
-\title{Lens Modeling in \sw}
+\title{Lens Modeling in a Citizen-Science context}
 
@@
 
-\input partA
+\begin{abstract}
+We develop a method to enable collaborative modelling of gravitational
+lenses and lens candidates by experienced but non-professional lens
+enthusiasts.  It uses an existing free-form modelling program (GLASS)
+but provides the modelling input in a novel way, via a user-provided
+diagram that is essentially a sketch of an arrival-time surface.
 
-\input partB
+An implementation (\spl) is tested in a modelling challenge using 29
+simulated lenses out of a larger set created for a citizen-science
+lens search (\sw).  We find that amateur lensers inferred the image
+parities and time ordering consistently in some lenses, while tending
+to make errors in other lenses, depending on the image morphology.
+Errors in image parity and time ordering lead to large errors in the
+mass distribution, but the enclosed mass is more robust.  Shortcomings
+include (a)~lens profiles that tend to be systematically shallower
+than the lenses, and consequent over-estimation of the Einstein radius
+by $\sim25\%$, and (b)~lack of detailed source modelling.  Ideas for
+improvement are discussed.
+\end{abstract}
 
-\input partC
+\begin{keywords}
+\end{keywords}
+
+\section{Introduction}
+
+The first two discoveries of gravitational lenses
+\citep{1979Natur.279..381W,1980Natur.285..641W}, where a galaxy images
+a background quasar into two (four), brought in their wake the first
+work on lens modelling
+\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.  For the second lens
+(PG1115+080), mass models scored an early success with the prediction
+that one of the lensed images seen would split further into two at
+higher resolution.  That galaxies must sometimes cause multiple images
+had long been expected \citep{1937ApJ....86..217Z}, and it had even
+been argued that the phenomenon could help measure cosmological
+parameters \citep{1964MNRAS.128..307R,1966MNRAS.132..101R}, but
+apparently nobody was expecting that lenses would need detailed
+modelling.  The first observations, however, immediately stimulated
+models.  The reason for that lies in the image separation. Recall that
+image separations are of order the angular Einstein radius
+\begin{equation}
+\sim \left(\frac{4GM}{c^2 D_L}\right)^{1/2}
+\simeq 0.1'' \left(\frac{M}{M_\odot}\right)^{1/2}
+             \left(\frac{D_L}{\rm pc}\right)^{-1/2}
+\end{equation}
+where $D_L$ is the distance to the lens, and $M$ its mass.  A lensing
+galaxy with $M\sim10^{11}M_\odot$ at $10^9\rm\,pc$ would cause image
+separations of $\sim1''$, which is comparable to the size of the
+galaxy; typically the lensed images are seen through the galaxy halo.
+Hence the lensed images depend on the detailed mass distribution of
+the lensing galaxy.  Galaxy lenses demand modelling.
+
+Since the early discoveries, more than 400 secure lenses are now known
+(see Figure \ref{fig:masterlens}).  Modelling of the mass distribution
+is part of any research using lenses, but so far no modelling study
+has spanned all known lenses.  The largest single one
+\citep{2009ApJ...703L..51K} models 58 lenses to infer the distribution
+of dark matter around galaxies.  In other work,
+\cite{2011ApJ...740...97L} combined lens models of 21 galaxies with
+models of their stellar populations, to find the relation between
+stars and dark matter, and \cite{2014MNRAS.437..600S} modelled 18
+time-delay lenses together to infer cosmological parameters.
+
+\begin{figure}
+\centering
+\includegraphics[width=\hsize]{fig/lenssky}
+\caption{Sky distribution of 423 published lenses considered secure
+  (from the Masterlens catalog at the University of Utah, maintained
+  by Joel Brownstein and Leonidas Moustakas).  The map is in
+  Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
+  middle. The empty swathes to the left and right of centre are the
+  Milky Way. {\em Should we drop this figure?  It's not relevant to
+    modelling, but as there seems to be no publicly available summary
+    of all known lenses, it does help set context.}}
+\label{fig:masterlens}
+\end{figure}
+
+Surveys now under way aim to increase the inventory of lenses another
+ten or a hundred fold.  Among these is \sw (Marshall et al, in prep,
+More et al, in prep).  \sw is a citizen science project\footnote{\tt
+  http://www.spacewarps.org} in which volunteers are presented
+sky-survey images and invited to identify lens candidates.  Simulated
+lenses are mixed in with the data, both to help train volunteers on
+what to look for, and to provide measures of the effectiveness of the
+search.  The motivation for \sw is to enable volunteers, some of whom
+had previously serendipitously identified lens candidates on earlier
+citizen-science surveys, to make discoveries missed in automatic
+searches by software robots.  Robots are very good with clean lensing
+system in uncrowded fields with high signal-to-noise, but in general
+test situations \citep{2009ApJ...694..924M} robots miss lenses (low
+completeness) or contaminate the results with non-lenses (low purity).
+\sw introduced its first tranche of survey data, in May 2013, from the
+Canada-France-Hawaii Telescope Legacy Survey, or CFHTLS
+\citep{Coupon2009}, covering $\simeq172$ square degrees (0.4\% of the
+sky) was divided into tiles of $\simeq 80''\times80''$ ($440\times
+440$ pixels).  Each such tile was shown to ten volunteers.  Despite
+the rarity of detectable lenses ($<10^{-3}$ per tile), the \sw
+collective soon found lens candidates, both rediscoveries already
+known from robots \citep{Gavazzi2012, More2012ApJ} and possible new
+lenses.
+
+The encouraging early results prompted the question: could modelling
+of the lenses also be done by volunteers?  If so, modelling could help
+assess prioritise lens candidates at an early stage, which would be
+very useful with new wide-field and sensitive surveys, which will
+yield thousands of lens candidates. There are several software tools
+for lensing modelling available, and work has been done on generic
+interfaces \citep{2014A&C.....5...28L}.  Some early designs for \sw
+included a prototype lens modelling tool \citep{2010AAS...21543527N}.
+Moreover, some {\em Spacewarps\/} volunteers are quite experienced
+from earlier projects, having spent a thousand hours or more with
+data, and are very interested in more demanding projects.  The
+interests of citizen-science communities are just beginning to be
+studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that
+some volunteers welcome open-ended challenges, and sometimes these
+have led to new scientific results: one example is the discovery of an
+exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
+the development of new algorithms for protein folding
+\citep{Khatib22112011}.  All these are grounds for optimism.  There
+is, however, a basic difficulty in strong gravitational lensing.
+Lensed images do not look much like their source, and still less do
+they resemble the lensing-mass distribution.  To model a lens, one
+needs to either do a lot of random guessing, or have a good intuition
+for what works.
+
+In this paper, we propose a way around the difficulty, and report on a
+modelling test on \sw using simulated lenses.  The three following
+sections are devoted to the concept, the implementation, and tests
+respectively.
+
+In \secref{Fermat} we introduce a markup system for lensed images,
+which we call a ``spaghetti diagram''.  A spaghetti diagram resembles
+the visible image system, in a cartoon-like way, and at the same time
+it encodes the basis of a mass model.  This supplies an intuitive link
+between the image system and the mass distribution, which look
+frustratingly different from each other.  Spaghetti diagrams are
+essentially the saddle-point contours originally introduced to
+gravitational lensing by \cite{1986ApJ...310..568B} as a way of
+classifying lensed images.  They are sometimes shown as part of the
+output of lens models \citep[for
+  example][]{2001ApJ...557..594R,2003ApJ...590...39K,Lubini2012}.  In
+the present work, however, spaghetti diagrams are the {\em input\/}
+through which the modeller tells the program what to do.
+
+In \secref{SpaghettiLens} we describe the \spl program, which
+implements everything.  \spl is an extension of the GLASS framework
+for modelling lenses \citep{2014arXiv1401.7990C}.  We will not go into
+software details in this paper, concentrating instead on lens
+modelling per se, but we remark that \spl is designed to be friendly
+to the forum style of citizen-science projects, and enable incremental
+model refinement by different people, without sacrificing any features
+of GLASS.
+
+In \secref{mod_challenge} we describe a modelling challenge where a
+diverse sample of 29 simulated lenses was modelled multiple times
+using \spl.  The models are then examined in two ways.  One was
+whether the spaghetti diagram was correct.  The other was the recovery
+of the Einstein radius.  In addition, we show some visual comparisons
+of actual and recovered lens shape and radial profile, and identify
+some areas to improve, but do not compute statistics on these aspects.
+Profile and shape recovery with GLASS has been studied in more detail
+in \citep{2014arXiv1401.7990C}.
+
+\secref{todo} concludes, with the general outlook and next steps.
+
+\section{Fermat's principle and spaghetti diagrams} \label{sec:Fermat}
+
+We now explain the lensing theory relevant to \spl, following the
+formulation of gravitational lensing in terms of Fermat's principle by
+\cite{1986ApJ...310..568B}.
+
+\subsection{Geometrical and gravitational time delays}
+
+Consider a lens at some redshift $z_L$ and let $(x,y)$ be planar
+coordinates at the lens, transverse to the line of sight.  Let
+$\Sigma(x,y)$ be the mass per unit area, i.e., density projected along
+the line of sight.  This surface density is often given in a
+dimensionless form
+\begin{equation} \label{eq:kappa}
+\kappa(x,y) \propto \Sigma(x,y)
+\end{equation}
+called the convergence.  And let there be light, in the form of a more
+distant source, at redshift $z_S$, behind point $(x_s,y_s)$ on the
+lens.
+
+We now imagine a virtual photon flying from the source to some $(x,y)$
+on the lens, then changing direction and coming to the observer.  Such
+a direction change would increase the light travel time compared to
+coming through $(x_s,y_s)$.  The increased light travel time from the
+geometry of deflection would be
+\begin{equation} \label{eq:tgeom}
+\tgeom(x,y) \propto (x-x_s)^2 + (y-s_y)^2 \,.
+\end{equation}
+assuming the delay is small compared to the total light travel time.
+
+An additional delay of the photon comes from travelling through the
+curved spacetime at the lens.  This gravitational time delay $\tgrav$
+is related to the mass distribution of the lens.  The relation is
+generally written as a two-dimensional Poisson equation, but an
+alternative expression, avoiding calculus, is as follows.  The value
+of $\tgrav$ through $(x,y)$ equals its average value on the
+circumference of a small circle centred at $(x,y)$, plus a constant
+times the mass within that circle.  The constant is $2G/c^3$ times the
+cosmological expansion factor $(1+z_L)$. Thus
+\begin{equation} \label{eq:tgrav}
+\tgrav(x,y) = \left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle
+              + (1+z_L) \frac{2G}{c^3} M(x_\subbullet,y_\subbullet) \,.
+\end{equation}
+We have used $(x_\subcirc,y_\subcirc)$ to denote the circumference of
+a circle, and $(x_\subbullet,y_\subbullet)$ to indicate the integrated
+mass within the circle.
+
+The light travel time of a virtual photon is therefore longer by
+\begin{equation}  \label{eq:tarriv}
+t(x,y) = t_{\rm geom} + t_{\rm grav} \,.
+\end{equation}
+than it would have been with no lens present.  Real photons take paths
+that make $t(x,y)$ extremal, that is, having a minimum, maximum or
+saddle point (Fermat's principle).
+
+The proportionality factors in \eqref{eq:kappa} and \eqref{eq:tgeom}
+depend on the redshifts and cosmological parameters, and are given in
+the Appendix~\ref{more-theory}.
+
+\begin{figure}
+\centering
+\includegraphics[width=\columnwidth]{fig/arriv_0}
+\includegraphics[width=\columnwidth]{fig/arriv_1}
+\includegraphics[width=\columnwidth]{fig/arriv_2}
+\caption{Perspective views and contour maps of example arrival time
+  surfaces.  Contours are coloured in rainbow order (red: least delay,
+  violet: highest delay).  The special contours that self-cross at
+  saddle points are the basis of spaphetti diagrams.  {\em Upper
+    panel:\/} No lens, hence showing the parabolic shape of the
+  geometrical time delay.  The image would be at the bottom,
+  coinciding with the source. {\em Middle panel:\/} A circular lensing
+  mass (offset from the source) has been added, which has pushed the
+  minimum to one side and introduced a maximum and saddle point, each
+  corresponding to an image.  The saddle-point is characterized by a
+  self-crossing ``spaghetti'' contour. {\em Lower panel:\/} An
+  elongated lensing mass has been added.  There are now two minima,
+  two saddle points, and a maximum, each corresponding to an image.}
+\label{fig:arriv}
+\end{figure}
+
+\subsection{Arrival-time contours} \label{sec:arriv}
+
+The full function $t(x,y)$, also known as the arrival-time surface,
+applies to virtual photons.  In other words, it is an abstract
+construct and not itself observable.  But observable image positions
+can be derived from the arrival-time surface, so visualising the
+surface is useful.  \figref{arriv} does so.  In this figure, a
+maximum, if present, is easy to see.  To locate mimima and
+saddle-points, however, one needs to examine the contours of equal
+arrival time.  A saddle point is characterised by a contour crossing
+itself, forming an~X.  Mimima, on the other hand, have contours
+looping around them, as do maxima.
+
+The saddle-point contours which form an X are especially interesting,
+because they set the overall topography of the arrival-time surface.
+The obviously give the locations of the saddle points, and roughly
+localise the minima and maxima as well.  If more precise locations of
+the minima and maxima are added, the whole arrival-time surface is
+already approximately known.  Since the arrival-time surface has an
+exact relation to the lens-mass distribution and the source position,
+in effect the mass distribution is also automatically specified.  In
+other words, a simple sketch of saddle-point contours along with
+locations of minima and maxima ---which we call a spaghetti diagram---
+is implicitly already an approximation to a lens-mass distribution.
+
+The preceding assumes a point source.  To get an idea of what an
+extended source would do, let us imagine moving the original source
+slightly.  The contours of constant arrival time will naturally move
+slightly, and so will the images.  The movement of the contours will
+be most noticeable where the contours are far apart, that is where the
+arrival-time surface is nearly flat.  As is evident from
+\figref{arriv}, this is the region where the minimum and saddle points
+lie, or near the images.  In this region, points on the source that
+are close together produce images that are comparatively far apart.
+In other words, the image is highly magnified.  In summary, lower
+curvature in the arrival-time surface for a point source implies
+larger magnification of an extended source.  Conversely, where the
+arrival-time surface is strongly curved, the image will be
+demagnified.  We see from \figref{arriv} that the arrival-time surface
+tends to be highly curved near the maximum.  Hence maximum tend to be
+demagnified.  In practice, maxima of the arrival time are nearly
+always too faint to see. The minima and saddle points dominate.
+
+\section{A lens-modelling program} \label{sec:SpaghettiLens}
+
+\spl is not part of \sw, but it makes use of the \sw infrastructure,
+in particular, the image database and the forum.  The forum is
+essential essential for initial contact with interested members of the
+\sw community.  Lens modelling from \sw works, in practice, in the
+usual style of medium-sized astronomical collaborations, with
+videoconferencing and in-person meetings where possible, but is
+reported in summary on modelling threads on the forum, and anyone
+interested is welcome to join at any time.  Another aspect of \sw (and
+the Zooniverse family of projects in general) that \spl adopts, is to
+keep all user interaction in a web browser, with intensive computing
+done on a server.  Several interesting design and implementation
+issues arise from these considerations, but we will not go into them
+in this paper, concentrating instead on aspects specific to lens
+modelling.
+
+Modelling with \spl involves three stages, markup of the image,
+followed by intensive numerics in the background, followed by review
+of diagnostics and possible discussion.  We now describe each of
+these.
+
+\subsection{Image markup}
+
+One begins by going to the \spl page and entering the number of a \sw
+image tile.  \spl presents the image, along with zoom and pan options
+and a markup tool to construct a spaghetti diagram.  The human
+modeller now has to make an educated guess for the topography of the
+arrival-time surface, and input the locations time-ordering of the
+maxima, minima, and saddle-points.  The markup tool \citep[which is
+  inspired by Figure~6 of][and is like that figure made interactive
+  and overlaid on data]{1986ApJ...310..568B} lets the modeller enter
+the information by sketching saddle-point contours.  Examples can be
+seen in the upper-left panels of Figures \ref{fig:6941} to
+\ref{fig:7022}.  The loops in the markup tool were the origin of the
+``spaghetti'' metaphor.
+
+The exact placement of the loops in a spaghetti diagram has no
+significance, only the connections, and the heirarchy of which loop is
+inside which matters.  The loops are there simply to help modeller's
+intuition.  The user does not need to think explicitly the image
+parities (though the markup tool provides this information using
+colour codes) or about time-ordering, or worry about the odd-image
+theorem.  The markup tool allows only valid lensing configurations to
+be entered.
+
+As implemented so far, \spl assumes that the lens is dominated by a
+single galaxy.  Only one maximum is permitted, and it is taken as the
+centre of the main lensing galaxy.  The user can, however, mark
+additional minor galaxies.  The latter are modelled as point
+masses.
+
+\subsection{Numerics}
+
+Having sketched a spaghetti diagram, the user orders a model, and \spl
+sends the input to GLASS, which runs server-side as it is
+compute-intensive.  The task of GLASS is to find a mass distribution
+$\kappa(x,y)$ that exactly reproduces the given locations of the
+maximum, minima and saddle points. Now, this criterion by itself is
+extremely under-determined --- there are infinitely many mass
+distributions that will reproduce a given set of maxima, minima and
+saddle points, but typically they (a)~produce lots of extra images,
+and (b)~look very unlike galaxies.  Additional assumptions (a prior)
+are necessary.  GLASS uses the following prior
+\citep[cf.]{1997MNRAS.292..148S,2008ApJ...679...17C}.
+\begin{enumerate}
+\item The mass distribution is built out of non-negative tiles of
+  mass.  (Sometimes these tiles are called mass pixels, but we should
+  emphasize that they are unrelated to image pixels, and are much
+  larger.)
+\item There is a notional lens center, say $(x_0,y_0)$ which is
+  identified with the maximum of the arrival time.  The source can
+  have an arbitrary offset with respect to the lens center.
+\item The mass distribution must be centrally concentrated, in two
+  respects.  First, the circularly averaged density must fall away
+  like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.
+  Second, the direction of increasing density at any $(x,y)$ can point
+  at most $45^\circ$ away from $(x_0,y_0)$.
+\item The lens must be symmetrical with respect to $180^\circ$ rotations
+  about $(x_0,y_0)$.  This symmetry assumption can be relaxed if the
+  user wishes.
+\end{enumerate}
+There are still infinitely many models that satisfy both data and
+prior constraints, but now they are more credible as galaxy lenses.
+It is then possible to generate an ensemble of models.  The sampling
+technique used by GLASS is described in \citep{Lubini2012}.
+Typically, ensembles of 200 models are used.  That is to say, what we
+call a \spl model is really the mean of an ensemble of 200 models, and
+its estimated uncertainty is the range covered by the whole ensemble.
+
+\subsection{Diagnostics} \label{sec:diag}
+
+After the model ensemble has been generated, \spl post-processes it to
+present results and diagnostics to the user for inspection. This takes
+the form of three figures.
+\begin{enumerate}
+\item A synthetic image of the lensed features.
+\item A contour map of the arrival-time surface $t(x,y)$.
+\item A gray scale plus contour map of the mass distribution.
+\end{enumerate}
+The synthetic image generated by \spl assumes a simple conical source
+profile.  The user can change the contrast level on the image, which
+(though it is not saved) amounts to adjusting the width of the cone.
+These synthetic images are still very crude and not very useful for
+assessing models.  The best indicator, in practice, of whether the
+modelling was successful is contour map of $t(x,y)$, with saddle-point
+contours highlighted.  It is, in effect, the computer's refinement of
+the spaghetti diagram input by the user.  If the arrival-time surface
+looks qualitatively similar to the spaghetti diagram, that generally
+indicates a successful model.  The mass distribution also provides
+indications; successful models generally lead to smooth-looking mass
+distribution, whereas an irregular or checkboard pattern in the mass
+map signals a bad model.
+
+After examining this feedback, the user can choose to post the model
+on the \spl archive.  They can also modify their input and try again,
+or discard the attempt altogether.  After archiving, there can be
+discussion among modellers, through the \sw forum or by any other
+means, and revision of the model.  Any archived model can be revised
+by any user: they can modify the spaghetti configuration slightly or
+drastically, or change options like the size of the mass tiles.
+Particularly interesting lens candidates lead to trees of models in
+this way.  Discussion among modellers tends to prune a model tree,
+focusing attention on the most interesting models.\footnote{See
+  ``Collaborative gravitational lens modelling\dots'' in {\tt
+    http://letters.zooniverse.org} for an example.}
+
+\section{A lens modelling challenge} \label{sec:mod_challenge}
+
+We now describe a test of the lens-modelling system, under conditions
+that mimic as closely as possible the modelling of real lens
+discoveries.  The lenses to be modelled were the simulated lenses
+(known as ``sims'') already sprinkled onto the \sw images.  Once a
+small user base had grown around \spl, a modelling challenge was
+announced through the \sw forum.  The challenge set consisted of 29
+sims, chosen to represent the different visual morphologies of \sw
+sims. Modellers then contributed 119 models for these sims (at least
+two for each sim).  Models were reported on the same forum used to
+model real candidate lenses.  Modellers were free to consult and
+refine each other's models, but had no information on how the sims
+were generated.
+
+Once the modelling was complete, the models were compared with the
+originals.  There were two main tests: a check of whether the the
+spaghetti diagrams were correct for the lens in question, and a
+comparison of the effective Einstein radii of the sims and the models.
+
+\subsection{The simulated lenses} \label{sec:sims}
+
+The \sw sims are described in detail in More et al (in prep), but
+relevant here is that they were of three kinds, as follows.
+
+\begin{enumerate}
+  \item Imitating lensed quasars: having a singular elliptical
+    isothermal lens (SIE) plus constant external shear, and a circular
+    Gaussian source.
+  \item Emulating lensed galaxies: similar to the above, but with an
+    elliptical de Vaucouleurs source.
+  \item Resembling cluster lenses: having a source as above, but a
+    more complicated lens, with one dominant elliptical SIE and
+    one or more perturbing elliptical SIEs, plus a circular NFW
+    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
+    the underlying dark matter distribution.
+\end{enumerate}
+The {\tt gravlens} program \citep{2001astro.ph..2340K} was used.
+Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
+lenses follow equations (33--35) of that work, with core radius set to
+zero.  The NFW lens is in equations (48) and (50), while shear is the
+$\gamma$ term in equation (76).
+
+The information in this section was not revealed to the main developer
+of \spl (RK, who also chose the challenge set) or to the modellers
+(EB, CC, CM, JO, PS and JW) while modelling was in progress.  That is,
+the modellers had no advance knowledge of what kind of
+parameterisation had been used to make the sims.  After the modelling
+stage, AM released the details of the sims for post-modelling
+analysis.  Results from the latter now follow.
+
+\subsection{Some example models} \label{sec:example_models}
+
+Of the 119 models proposed, we now discuss eight examples in some
+detail.  Results from these are shown in Figures
+\ref{fig:6941}--\ref{fig:7022}.  The first four of these show the most
+common image morphologies, the other four explain some problem cases.
+
+Each of Figures \ref{fig:6941}--\ref{fig:7022} figures has the
+following layout.
+$$ \begin{matrix}
+\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
+t(x,y)                        &\hbox{model } t(x,y) \\
+\kappa(x,y)                   &\hbox{model } \kappa(x,y)
+\end{matrix} $$
+The marked-up image is a zoom of the lensed image on Spacewarps,
+marked up with one or more spaghetti contours; this is the modeller's
+input to \spl.  The three panels on the right show the graphical
+output returned to the modeller by \spl, listed as (i), (ii), (iii) in
+\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
+200 models generated by \spl. The model synthetic image is too crude
+to be useful, because there is no fitting for the source light
+distribution, as already noted in \secref{diag}.  The quality of the
+synthetic images can be improved by interpolation, but as this feature
+was not available during the modelling challenge, we do not show the
+improved versions here.
+
+\FloatBarrier
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\myplotswidth]{fig/006941_input}
+  \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 
+  \includegraphics[width=\myplotswidth]{fig/006941_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006941_mass}
+  \caption[result 6941 (ASW000102p)]{A simulated lens that mimics a
+    lensed quasar, and model results.  The left panels derive from the
+    simulation, and the right panels are \spl output.  Details of
+    individual panels are in \secref{example_models}.}
+
+  \label{fig:6941}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/006915_input}
+  \includegraphics[width=\myplotswidth]{fig/006915_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006915_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006915_mass}
+
+  \caption[result 6915 (ASW0001hpf)]{A four-image configuration
+    typical of lensed quasars. (See \secref{example_models} for
+    details.)}
+  \label{fig:6915}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/006990_input}
+  \includegraphics[width=\myplotswidth]{fig/006990_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006990_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006990_mass}
+
+  \caption[result 6990 (ASW0004oux)]{Results from a system with an arc
+    plus a counter-image, typical of lensed galaxies. (See Section
+    \ref{sec:example_models} for details.)}
+  \label{fig:6990}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/006919_input}
+  \includegraphics[width=\myplotswidth]{fig/006919_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006919_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006919_mass}
+
+  \caption[result 6919 (ASW0002z6f)]{Another configuration of arc plus
+    counter-image; here the arc is closer to the lensing galaxy than
+    the counter-image. (See \secref{example_models} for details.)}
+  \label{fig:6919}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\myplotswidth]{fig/006975_input}
+  \includegraphics[width=\myplotswidth]{fig/006975_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_arriv.png}
+  \includegraphics[width=\myplotswidth]{fig/006975_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_kappa.png}
+  \includegraphics[width=\myplotswidth]{fig/006975_mass}
+  \caption[result 6975 (ASW000195x)]{A lens with unrecovered mass
+    substructure. (See \secref{example_models} for details.)}
+  \label{fig:6975}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/006937_input}
+  \includegraphics[width=\myplotswidth]{fig/006937_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006937_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006937_mass}
+
+  \caption[result 6937 (ASW0000vqg)]{A sim with unrecovered
+    substructure, resulting in a poor mass model. (See Section
+    \ref{sec:example_models} for details.)}
+  \label{fig:6937}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/007025_input}
+  \includegraphics[width=\myplotswidth]{fig/007025_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_arriv}
+  \includegraphics[width=\myplotswidth]{fig/007025_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_kappa}
+  \includegraphics[width=\myplotswidth]{fig/007025_mass}
+  \caption[result 7025 (ASW0000h2m)]{A four-image system with image
+    parities incorrectly identified.  The model is poor, but the
+    estimated Einstein radius is not bad. (See Section
+    \ref{sec:example_models} for details.)}
+  \label{fig:7025}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/007022_input}
+  \includegraphics[width=\myplotswidth]{fig/007022_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_arriv}
+  \includegraphics[width=\myplotswidth]{fig/007022_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_kappa}
+  \includegraphics[width=\myplotswidth]{fig/007022_mass}
+
+  \caption[result 7022 (ASW0000h2m)]{The same system as in Figure
+    \ref{fig:7025}, this time with image parities correctly
+    identified. (See \secref{example_models} for details.)}
+  \label{fig:7022}
+\end{figure}
+
+\FloatBarrier
+
+In these figures, the scales and alignments in the six panels are not
+completely consistent: the three panels on the right exactly the same
+coordinates; the middle-left and lower-left panels are consistent with
+each other, but only approximately aligned with the right column,
+because the latter are affected by errors in mouse placements; the
+upper-left panel is zoomed out slightly, to accommodate the full
+markup.  For qualitative features, which is all that
+Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, the slightly
+varying scales and alignments are harmless.
+
+Let us now consider these cases in turn.
+
+\begin{itemize}
+
+\item Figure \ref{fig:6941} shows the simplest case, with two clear
+  images produced by a nearly-circular lens.  The centre of the
+  lensing galaxy is a maximum, the image nearer to the galaxy is a
+  saddle point, and image further away is a minimum, and these were
+  correctly identified.
+
+\item Figure \ref{fig:6990} shows an example of an arc that has split
+  into three images.  This kind of configuration, with a counter-image
+  close to the lensing galaxy and a more distant arc/triplet on the
+  other side, generically arises from an elongated mass distribution
+  when the source is displaced along the elongated direction.
+
+\item Figure \ref{fig:6919} shows another example of an arc plus
+  counter-image, but (in contrast to Figure \ref{fig:6990}) the arc is
+  closer to the lens than the counter-image. This configuration arises
+  if the source displacement is perpendicular to the long axis of the
+  lensing mass.  Comparing the two panels in the middle row, we see
+  that the modeller interpreted arc as consisting of three images,
+  whereas the sim shows a single saddle point associated with the arc.
+  But the identification is not really erroneous --- we just need to
+  take into account that the source is extended.  In fact, in the sim
+  the brightest part of the source is only doubly imaged, but the
+  source extends into a region that produces four images.  In the
+  $t(x,y)$ of the sim, the hairpin-bending contours are typical of
+  double on the verge of splitting into a quad.
+
+\item Figure \ref{fig:6915} shows another quad.  This kind of
+  configuration arises when the mass is elongated and the source is
+  displaced at an angle to the elongation.  The minima and saddle
+  points are correctly identified, and the orientation of the
+  ellipticity of the mass distribution is correctly reproduced.
+
+\item Figure \ref{fig:6975} shows a lens with substructure in the form
+  of a smaller secondary galaxy.  The galaxies in such group or
+  cluster sims were based on galaxies visible in the images, but the
+  modelers were not told in advance whether this was the case.  The
+  minimum and saddle point are correctly identified.  The mass
+  distribution misses the substructure, but overall appears
+  reasonable.
+
+\item Figure \ref{fig:7025} shows a quad.  In this one, the
+  identification of the minima and saddle points was incorrect, and
+  mass distribution comes out elongated East-West instead of
+  North-South.  The mass distribution also appears somewhat jagged and
+  the saddle-point contours are not as clean as in the previous
+  examples; these are often indicators of a problem with the model.
+  The enclosed mass is, however, none the worse --- the reason is
+  probably that in a relatively symmetrical image configuration, the
+  Einstein radius is quite well constrained by the images in a fairly
+  model-independent way.
+
+\item Figure \ref{fig:7022} shows another model of the same system.
+
+\end{itemize}
+
+\subsection{Test of image identification} \label{sec:tests.t1}
+
+The first post-modelling test was a qualitative comparison of the
+original arrival-time surfaces and the input spaghetti diagrams given
+by the modeller.  This tested first, for correct identification and
+location of the lensed images, and second, the correct parities and
+ordering for the lensed images in respect of the arrival time.
+
+While we expected the identification of lensed images to be trivial,
+given the generally clean appearance of the sims in the test, we
+expected the parities and time-ordering to be more difficult.  While
+the \spl tutorials had provided some general rules of thumb, to be
+consistently correct with the time-ordering, a modeller needs to
+develop some intuition for arrival time surfaces.  This is an area
+where experience and tutorials training could improve results at a
+later stage, and correspondingly, feedback on the difficulties
+modellers encounter can help improve the tutorial materials.
+
+\tabref{stats} presents a summary of the test.  The evaluation was
+done manually, comparing the input to \spl with the actual
+arrival-time surface of the sim.  This amounts to comparing the
+upper-left and middle-left panels in each of
+Figures~\ref{fig:6941}--\ref{fig:7022}, and similarly the other 111
+models.
+
+\begin{table}\centering\begin{tabular}{lr}\hline
+    Total & 119 \\
+\hline
+    errors in image locations  & 9 \\
+    errors in image parities or time ordering & 49 \\
+\hline
+    inaccurate image placement over an arc & 21 \\
+    identified two images of four & 5 \\
+    identified two nearby images as one & 3 \\
+    missed faint images & 1 \\
+    proposed too many images & 1 \\
+    modelled a three-image arc as one image & 4 \\
+    modelled one image as a three-image arc & 5 \\
+    swapped minimum and saddle in double  & 2 \\
+    swapped minima and saddles in quad & 38 \\
+    swapped early and late saddles in a quad & 7 \\
+\hline
+\end{tabular}
+\caption{Table of image-identification errors and the number of models
+  containing each.  A model can contain more than one type of error.}
+\label{tab:stats}
+\end{table}
+
+The images of the system are considered to have been located
+correctly, if all the images have been identified and are approximated
+within about 5\% of \spl frame used to draw the spaghetti diagram.
+That frame size is adjustable by the user, but in practice it is
+somewhat larger than the spaghetti diagram.  Such image-placement
+errors were found in only 9 models.  That does not include inaccurate
+image placement over an arc, which was considered a separate category
+of error.
+
+In addition to simple image-placement errors, ten types of errors were
+recognised and are listed in \tabref{stats}.  Most of the problems
+where due to unclear arc-like structures.  Critical errors like the
+failure to identify all five images in a five images system, or to
+include too many images, were rare.
+
+The assignment of the parity of the images was a more difficult task,
+and was successful in only 60\% of the cases.  The most common error
+was swapping of minima and saddle-points in a quad; \figref{7025}
+shows an example.  Another, less common, error was flipping the
+spaghetti diagram, this swapping the time-ordering of the two saddle
+points.
+
+Incorrect image parities and time orderings tended to produce
+poorer-looking models, such as the checkboard patterns in the mass map
+in \figref{7025}.  Interestingly, however, the enclosed-mass profiles
+were quite robust.  We will consider this aspect in the next section.
+
+\FloatBarrier
+
+\begin{figure}
+  \centering
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007025_kappa_encl}}
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007022_kappa_encl}}
+  \subfigure{\includegraphics[width=0.8\linewidth]
+             {fig/007021_kappa_encl}} \\
+  \caption{Mean $\kappa$ inside a circle around the lens centre, as a
+    function of the radius of the circle.  (See \secref{tests.t2} for
+    details.) The upper panel corresponds to the model shown in
+    \figref{7025}, in which the minima and saddle-points have been
+    incorrectly swapped.  The middle corresponds to \figref{7022},
+    where the image parities were correct.  The lower panel
+    corresponds to another model, where the image parities were
+    correct but the time-ordering was incorrect.}
+\label{fig:kapenc_compare_faulty}
+\end{figure}
+
+\subsection{Test of mass-profile recovery} \label{sec:tests.t2}
+
+The second test was to compare the mass distributions $\kappa(x,y)$ of
+the sims and of the \spl models.  A visual comparison is presented for
+the eight models in Figures~\ref{fig:6941}--\ref{fig:7022}, in the
+lower-left versus lower-right panels.  We will summarize the mass
+distributions drastically in a single number, the effective Einstein
+radius.  Other measures for comparison of free-form lensing mass
+distributions appear in \cite{2014arXiv1401.7990C}, but comparing
+Einstein radii is already useful.
+
+There is no standard way of defining the Einstein radius of a general
+non-circular lens.  We adopt the simple definition
+\begin{equation} \label{eq:effRE}
+ \langle\kappa\rangle_{\Theta_E} = 1,
+\end{equation}
+that is, the effective Einstein radius \ERf is such that the
+mean $\kappa$ is unity inside a circle of radius \ERf centred at
+the lens centre.
+
+To illustrate, \figref{kapenc_compare_faulty} compares the circularly
+averaged mass profiles three different models of one particular lens;
+two of the models are shown in Figures~\ref{fig:7025} and
+\ref{fig:7022}. Each panel in \figref{kapenc_compare_faulty} shows the
+mean $\kappa$ within a circle of given radius.  The red curve is the
+correct profile for the sim.  The two blue curves are the minimal and
+maximal mean enclosed $\kappa$ from the internal ensemble in \spl.
+Radial locations of the images are marked, along with the image
+parities.  The region between the blue curves is shaded between the
+radii of the innermost and the outermost images: this is the
+confidence region from the modelling.  The definition (\ref{eq:effRE})
+for \ERf corresponds to crossing the dashed horizontal line at
+$10^0$: the red curve crosses the dashed line at the actual Einstein
+radius \ERf[,act]; the recovered Einstein radius
+\ERf[,rec] and its uncertainty are given by the blue curves
+crossing the dashed line.  We see that in all three panels, the blue
+curves are shallower than the red curve and \ERf[,rec] is
+more than \ERf[,act], by more than the model uncertainties.
+Now, steeper mass profiles tend to give wider image separations ---
+recall that the image separation for a circular isothermal lens is
+$2\Theta_E$, whereas for a point mass it is more \citep[see,
+  e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
+high is really a consequence of the \spl models being too shallow for
+the sims.
+
+\begin{figure*}
+  \centering
+    \includegraphics[width=0.90\linewidth]{fig/eR_5}
+  \caption{Model-recovered versus actual Einstein radii \ERf[,rec] and
+    \ERf[,act].  Circles denote models where the image parities and
+    time-orderings were correct, with filled circles for models by an
+    expert.  Crosses mark models where the time-ordering was
+    incorrect.  Plus signs indicate models flagged by the modeller as
+    failures.}
+  \label{fig:ER_per_sim}
+\end{figure*}
+
+\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend
+to be too high.  Discounting the models which were flagged by the
+users as poor, the mean and standard deviation of the overestimate are
+25\% and 20\%.  The models by an expert (PS) also show a similar bias,
+having mean and standard deviation of the overestimate being 20\% and
+8\%.  One of the reasons for this is that it is hard to get the center
+of the lens on spot.  An offset leads to a flatter mass profile for
+the model compared to the simulation.
+
+\section{Outlook} \label{sec:todo}
+
+The lens-modeling challenge indicates that the Spacewarps collective
+is good, not only for identifying lens candidates for follow-up, but
+modelling candidates as well.
+
+There is, however, plenty of room for improvement.
+
+First, the particular modeling strategy implemented is not the only
+one possible.  \spl requires modelers to characterize the overall
+image structure in abstract terms based on Fermat's principle, and the
+placement of mass distributions is done by the computer.  In other
+modeling tools, the user puts down a trial mass distribution and has
+the machine refine it.  A few of these modeling programs have also
+been designed with citizen science in mind, and would be interesting
+in the Spacewarps environment.
+
+Second, \spl needs some enhancements.
+
+\begin{itemize}
+\item Currently, \spl does not attempt to model the source shape; the
+  user identifies the brightest points on the image, and these are
+  taken as images of a point-like source, whose positions must be
+  reproduced exactly. For generating a synthetic image, a conical
+  source profile is assumed. Fitting for the source profile to
+  optimize resemblance to the observed lensed image after the lens
+  model has been generated, is algorithmically straightforward
+  \citep[cf.][]{2003ApJ...590..673W,2006MNRAS.371..983S} and planned
+  to be implemented.  This would alleviate another problem with \spl,
+  which is that there is no numerical figure of merit, and assessment
+  of a model is a judgment call based on the synthetic image, and on
+  whether the mass distribution and the arrival-time surface show
+  suspect features.
+\item \spl has a tendency to somewhat overestimate the Einstein radius
+  (evident from Figure \ref{fig:ER_per_sim}), and it is also apparent
+  that the models tend to be too shallow.  This possible explanation
+  is that, while the sims are steeply peaked at the centre, the
+  pixellated mass model fixes a comparatively large area near the
+  central at constant density.  One way to solve this problem would be
+  to introduce smaller pixels in the central region, thus enabled a
+  steeper centre.
+\item Another limitation so far in \spl is that the lens is assumed to
+  be dominated by one galaxy, which puts most galaxy-group lenses
+  beyond the reach of the modeler. Since complicated group lenses are
+  some of the most interesting candidates present, removing this
+  limitation is most desirable.  From the users' point of view, it
+  would mean that spaghetti contours with more than one maximum can be
+  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
+  and Figure 4b in \cite{2003ApJ...590...39K}.
+\item At present, a single false-color composite is used as the data.
+  An option could be added to use all available filters, individually
+  or in combination, at the user wishes.
+\end{itemize}
+
+The third desirable avenue of improvement is to facilitate
+collaborative work.
+
+\begin{itemize}
+\item As mentioned above, the option of revising an already-archived
+  model is already available.  Desired now are tools for comparing
+  different models of a given system, both visually and through
+  different statistical measures.  As evidenced by a current
+  collaborative modeling effort, a particularly interesting candidate
+  can lead to an extended discussion and dozens of models.
+\item Better tutorial materials are also needed, and this would
+  address some of the problem areas found in the modeling challenge.
+  For example, we saw in \secref{tests.t1} that volunteers are most
+  prone to making errors in two situations: when in identifying an arc-
+  like structure while placing the points, and in identifying the
+  correct ordering of the points in nearly-symmetric configurations.
+\end{itemize}
+
+The \spl program was developed by Kueng, with design suggestions from
+Coles, Cornen and Saha, and feedback from all co-authors.  The
+simulations were created by A.~More, in consultation with Marshall,
+S.~More and Verma.  Modelling was done by Baeten, Cornen, McMillan,
+Odermatt, Saha and Wilcox, with post-modelling analysis by Kueng and
+Saha.  All authors participated in writing and editing the manuscript.
+
+\section*{Acknowledgments}
+
+We thank the Swiss Society for Astrophysics and Astronomy and the
+Swiss Academy of Sciences. AV is supported by a research fellowship
+from the Leverhulme Trust.
+
+\appendix
+
+\section{Relation to standard lensing formalism}\label{more-theory}
+
+The description of the arrival-time surface in
+Sections~\ref{sec:Fermat} and \ref{sec:arriv} omitted some details for
+the sake of a more intuitive explanation.  The convergence $\kappa$
+and the geometric time delay $t_{\rm geom}$ were left as
+proportionalities (equations \ref{eq:kappa} and \ref{eq:tgeom}), and
+the gravitational time delay $t_{\rm grav}$ was given in an implicit
+form (equation \ref{eq:tgrav}).  Here we fill in the details.
+
+The original formulation of the arrival-time surface appears in
+equations (2.1) to (2.6) of \cite{1986ApJ...310..568B}.  Their
+equations can be rearranged as follows.
+\begin{equation} \label{cfBN86}
+\begin{aligned}
+t_{\rm geom} &= \frac{(1+z_L)}{2c} \frac{d_S}{d_L d_{LS}}
+\left[ (x-s_x)^2 + (y-s_y)^2 \right] \\
+\nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y) \\
+\kappa(x,y) &= \frac{4\pi G}{c^2} \frac{d_L d_{LS}}{d_S}
+               \times \Sigma(x,y)
+\end{aligned}
+\end{equation}
+The symbols $d_L,d_S$ and $d_{LS}$ are angular-diameter distances,
+respectively from observer to lens, observer to source, and lens to
+source. We have replaced angular positions on the sky with positions
+on the lens plane as
+\begin{equation}
+(x,y) = d_L (\theta_x,\theta_y) \,.
+\end{equation}
+In the concordance cosmology
+\begin{equation}
+d_{LS} = \frac{c}{H_0}\,\frac1{1+z_S} \,
+\int_{z_L}^{z_S} \!\! \frac{dz}{\sqrt{\Omega_m(1+z)^3 + \Omega_\Lambda}}
+%\raise.5ex\hbox{.}
+\end{equation}
+and similarly $d_L$ and $d_S$.
+
+The first line of equation \eqref{cfBN86} is $t_{\rm geom}$ from
+equation \eqref{eq:tgeom} with the proportionality filled in, and with
+the source offset at $(s_x,s_y)$ rather than at the origin.
+
+The last line of equation \eqref{cfBN86} fills in the proportionality
+factor in equation \eqref{eq:kappa} for the convergence (or
+dimensionless surface density) $\kappa$.
+
+The middle line of equation \eqref{cfBN86} is a Poisson equation for
+the gravitational time delay, and is equivalent to the implicit
+expression \eqref{eq:tgrav}.  One way to verify the equivalence is to
+consider the small circle in equation \eqref{eq:tgrav} as a region
+where $\Sigma$ is constant, and approximate $\tgrav$ by its Taylor
+expansion to $O(x^2,y^2)$.  Substituting in equations \eqref{cfBN86}
+gives the Taylor coefficients in terms of $\Sigma$, and result
+satisfies the expression \eqref{eq:tgrav}.  Alternatively, we can
+proceed with a discrete form of the Poission equation from
+\eqref{cfBN86}.  Discretising on a grid with spacing $\Delta$, we have
+\begin{equation} \label{eq:relax}
+\begin{aligned}
+\tgrav(x,y) &=
+\frac14 \Big[ \tgrav(x+\Delta,y) + \tgrav(x-\Delta,y) + {}\\
+&\kern24pt    \tgrav(x,y+\Delta) + \tgrav(x,y-\Delta) \Big] \\
+&+ (1+z_L)\frac{2G}{c^3} \, \pi\Delta^2 \, \Sigma(x,y) \,.
+\end{aligned}
+\end{equation}
+This is recognisable as a formula for solving the two-dimensional
+Poisson equation from \eqref{cfBN86} by relaxation.  Let us now
+replace the average over four neighbouring points by the circular
+average $\left\langle \tgrav(x_\subcirc,y_\subcirc) \right\rangle$ and
+replace $\pi\Delta^2\,\Sigma(x,y)$ by the enclosed mass
+$M(x_\subbullet,y_\subbullet)$.  These replacements are valid in the
+limit of a small grid.  The result is the implicit
+equation \eqref{eq:tgrav}.
 

### 2014-06-30 plots/gen_plots.py
new mod plots generation complete

@@ -1122,7 +1122,7 @@ def draw_sim(asw, sim):
             #lev = np.linspace(0,10,41)
             #pc = panel.contour(x,y,kappa,lev, colors=0.8)
 
-            eps = 0.1 #small offset to prevent div/0 in log()
+            eps = 1e-15 #small offset to prevent div/0 in log()
             
             # kappa(x,y) is value at (x,y), but pcolormesh needs edge coordinates of patches
             # kappa(xi, yj) => x[i]-d / y[j]-d ... x[i+1]-d / y[j+1]-d ; with d step/cell width

### 2014-06-30 upd_plots/plots.py

@@ -8,33 +8,80 @@ import matplotlib.pyplot as plt
 import matplotlib.cm as cm
 
 import numpy as np
+np.set_printoptions(precision=3)
 
 
 def new_kappaplot(model, obj_index=0, **kwargs):
     """basically copied from glass plot.py, but heavily modified"""
-  
-    obj, data = model['obj,data'][obj_index]
-    if not data: return
+    kw = kwargs
 
-    with_contours   = kwargs.pop('with_contours', False)
-    only_contours   = kwargs.pop('only_contours', False)
-    label_contours  = kwargs.pop('label_contours', False)
-    clevels         = kwargs.pop('clevels', 30)
-    with_colorbar   = kwargs.pop('with_colorbar', False) # colorbar is nor working
-    vmin            = kwargs.pop('vmin', None)
-    vmax            = kwargs.pop('vmax', None)
-    subtract        = kwargs.pop('subtract', 0)
-    xlabel          = kwargs.pop('xlabel', r'arcsec')
-    ylabel          = kwargs.pop('ylabel', r'arcsec')
+    obj, data = model['obj,data'][obj_index]
+    if not data: return None
+
+    #with_contours   = kwargs.pop('with_contours', True)
+    #only_contours   = kwargs.pop('only_contours', False)
+    #label_contours  = kwargs.pop('label_contours', False)
+    #clevels         = kwargs.pop('clevels', 30)
+    #with_colorbar   = kwargs.pop('with_colorbar', False) # colorbar is nor working
+    #vmin            = kwargs.pop('vmin', None)
+    #vmax            = kwargs.pop('vmax', None)
+    #subtract        = kwargs.pop('subtract', 0)
+    #xlabel          = kwargs.pop('xlabel', r'arcsec')
+    #ylabel          = kwargs.pop('ylabel', r'arcsec')
+    
+    #sl scaling factors
+    rscale          = kw.pop('rscale', 440./500*100) # default Spacewarps -> SL -> glass scaling of radius
+    dist_fact       = kw.pop('dist_fact', 0.428)     # because measurement at different redshiffts (we had 0.5/1 hardcoded in beginning)
+    delta           = kw['contour'].pop('cldelta', 0.1) # contour level spacing in log space
+
+
+
+
+    # SET UP DATA
+
[…]

### 2014-06-30 upd_plots/updateplots.py

@@ -9,27 +9,79 @@ if False: # get rid of undefined variables in ide
   glass_basis = None
   loadstate = None
 
-
+import numpy as np
 import matplotlib.pyplot as plt
 import os
 import sys
+import requests as rq
 
 # bugfix for interactive mode
 if not 'plots' in sys.modules:
-  import plots
+    import plots
 else:
-  reload(plots)
+    import plots
+    reload(plots)
+
+np.set_printoptions(precision=3)
 
 try: # check if glass already loaded (in interactive mode)
   a = loadstate
 except:
   glass_basis('glass.basis.pixels', solver=None)
+  
+
+class AttrDict(dict):
+    def __init__(self, *args, **kwargs):
+        super(AttrDict, self).__init__(*args, **kwargs)
+        self.__dict__ = self  
+  
+  
+
+
+figargs = AttrDict({
+    'figsize'   : (8,8),  # size in inches
+    'dpi'       : 80,       #
+    'facecolor' : 'w',
+    'edgecolor' : 'k',
+})
+
+contourargs = AttrDict({
+    'extend'        : 'both', # contour levels are automatically added to one or both ends of the range so that all data are included
+    'aspect'        : 'equal',
+    'origin'        : 'upper',
+    'colors'        : 'k',
+    'antialiased'   : True,
+    'cldelta'       : 0.05, # contour level spacing in log space
+})
+
+savefigargs = AttrDict({
+    'dpi'       : figargs.dpi
+})
+
+
+kw = AttrDict({'figure': figargs, 'contour': contourargs, 'savefig': savefigargs,})
+
 
 
 
 inppath = "states"
 outpath = "outp"
-exts = ["png"]
+outfn = "Image for %(fname)s.%(ext)s"
+exts = ["png",'pdf']
+
+sel_mod = [
+    6915,
+    6919,
+    6937,
+    6941,
+    6975,
+    6990,
+    7020,
+    7021,
+    7022,
+    7024,
+    7025,
+]
 
 
 files = [(os.path.splitext(_)) for _ in os.listdir(inppath) if not _.startswith('.')]
@@ -38,19 +90,46 @@ files = [(os.path.splitext(_)) for _ in os.listdir(inppath) if not _.startswith(
 print "starting.."
 
 for fn, fe in files: # filename fn, fileextension fe
-  ifpath = os.path.join(inppath,fn+fe)
+    kw.fname = fn
+    kw.fext  = fe 
+    ifpath = os.path.join(inppath,fn+fe)
+      
[…]

### 2014-06-30 text/ms.tex

@@
 % shortcut for einstein radius (Text Greek Full Math)
-\newcommand{\ERf}[1][]{Einstein radius $\Theta_\text{E#1}$\xspace} % full output
+\newcommand{\ERf}[1][]{$\Theta_\text{E#1}$\xspace} % full output
 
@@
 they resemble the lensing-mass distribution.  To model a lens, one
-needs to either do a lot of random guessing, or have a good intuition
-for what works.
+needs either to do a lot of random guessing, or to have a good
+intuition for what works.
 
@@
 
-\secref{todo} concludes, with the general outlook and next steps.
+\secref{todo} gives the general outlook and next steps.
 
@@
 \begin{equation}  \label{eq:tarriv}
-t(x,y) = t_{\rm geom} + t_{\rm grav} \,.
+t(x,y) = t_{\rm geom} + t_{\rm grav}
 \end{equation}
@@
 depend on the redshifts and cosmological parameters, and are given in
-the Appendix~\ref{more-theory}.
+Appendix~\ref{more-theory}.
 
@@
 exact relation to the lens-mass distribution and the source position,
-in effect the mass distribution is also automatically specified.  In
-other words, a simple sketch of saddle-point contours along with
-locations of minima and maxima ---which we call a spaghetti diagram---
-is implicitly already an approximation to a lens-mass distribution.
+in effect the mass distribution is also automatically approximately
+specified.  In other words, a simple sketch of saddle-point contours
+along with locations of minima and maxima ---which we call a
+``spaghetti diagram''--- is implicitly already an approximation to a
+lens-mass distribution.
 
@@
 usual style of medium-sized astronomical collaborations, with
-videoconferencing and in-person meetings where possible, but is
-reported in summary on modelling threads on the forum, and anyone
-interested is welcome to join at any time.  Another aspect of \sw (and
-the Zooniverse family of projects in general) that \spl adopts, is to
-keep all user interaction in a web browser, with intensive computing
-done on a server.  Several interesting design and implementation
-issues arise from these considerations, but we will not go into them
-in this paper, concentrating instead on aspects specific to lens
-modelling.
+video-conferencing and in-person meetings where possible. But
+preliminary results are immediately summarized on modelling threads on
+the forum, and anyone interested is welcome to join at any time.
+Another aspect of \sw (and of the Zooniverse family of projects in
+general) that \spl adopts, is to keep all user interaction in a web
+browser, with intensive computing done on a server.  Several
+interesting design and implementation issues arise from these
+considerations, but we will not go into them in this paper,
+concentrating instead on aspects specific to lens modelling.
 
@@
 
-The exact placement of the loops in a spaghetti diagram has no
-significance, only the connections, and the heirarchy of which loop is
-inside which matters.  The loops are there simply to help modeller's
-intuition.  The user does not need to think explicitly the image
+The markup tool allows only valid lensing configurations to be
+entered.  The user does not need to think explicitly the image
 parities (though the markup tool provides this information using
 colour codes) or about time-ordering, or worry about the odd-image
-theorem.  The markup tool allows only valid lensing configurations to
-be entered.
+theorem.  The exact placement of the loops in a spaghetti diagram has
+no significance.  Only the heirarchy of which loop is inside which is
+relevant.  The loops are there simply to help modeller's intuition.
 
@@
 
-Having sketched a spaghetti diagram, the user orders a model, and \spl
-sends the input to GLASS, which runs server-side as it is
-compute-intensive.  The task of GLASS is to find a mass distribution
-$\kappa(x,y)$ that exactly reproduces the given locations of the
-maximum, minima and saddle points. Now, this criterion by itself is
+Having sketched a spaghetti diagram, the user presses a button to
+initiate the next stage.  \spl then translates the spaghetti diagram
+into input for GLASS, and forwards this input.  The task of GLASS,
+which runs server-side as it is compute-intensive, is to find a mass
+distribution $\kappa(x,y)$ that exactly reproduces the given locations
+of the maximum, minima and saddle points. This criterion by itself is
 extremely under-determined --- there are infinitely many mass
@@
 are necessary.  GLASS uses the following prior
-\citep[cf.]{1997MNRAS.292..148S,2008ApJ...679...17C}.
+\citep[cf.][]{1997MNRAS.292..148S,2008ApJ...679...17C}.
 \begin{enumerate}
@@
 discoveries.  The lenses to be modelled were the simulated lenses
-(known as ``sims'') already sprinkled onto the \sw images.  Once a
+(known as ``sims'') already sprinkled onto the \sw field.  Once a
 small user base had grown around \spl, a modelling challenge was
@@
 Once the modelling was complete, the models were compared with the
-originals.  There were two main tests: a check of whether the the
+originals.  There were two main tests: a check of whether the
 spaghetti diagrams were correct for the lens in question, and a
@@
 The \sw sims are described in detail in More et al (in prep), but
-relevant here is that they were of three kinds, as follows.
+relevant here is that the sims were of three kinds, as follows.
 
@@
 
+  \includegraphics[width=\myplotswidth]{fig/006990_input}
+  \includegraphics[width=\myplotswidth]{fig/006990_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006990_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006990_mass}
+
+  \caption[result 6990 (ASW0004oux)]{Results from a system with an arc
+    plus a counter-image, typical of lensed galaxies. (See Section
+    \ref{sec:example_models} for details.)}
+  \label{fig:6990}
+\end{figure}
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=\myplotswidth]{fig/006919_input}
+  \includegraphics[width=\myplotswidth]{fig/006919_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_arriv}
+  \includegraphics[width=\myplotswidth]{fig/006919_spaghetti} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_kappa}
+  \includegraphics[width=\myplotswidth]{fig/006919_mass}
+
+  \caption[result 6919 (ASW0002z6f)]{Another configuration of arc plus
+    counter-image; here the arc is closer to the lensing galaxy than
+    the counter-image. (See \secref{example_models} for details.)}
+  \label{fig:6919}
+\end{figure}
+
+\begin{figure}
+  \centering
   \includegraphics[width=\myplotswidth]{fig/006915_input}
@@
   \label{fig:6915}
-\end{figure}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\myplotswidth]{fig/006990_input}
-  \includegraphics[width=\myplotswidth]{fig/006990_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_arriv}
-  \includegraphics[width=\myplotswidth]{fig/006990_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_kappa}
-  \includegraphics[width=\myplotswidth]{fig/006990_mass}
-
-  \caption[result 6990 (ASW0004oux)]{Results from a system with an arc
-    plus a counter-image, typical of lensed galaxies. (See Section
-    \ref{sec:example_models} for details.)}
-  \label{fig:6990}
-\end{figure}
-
-\begin{figure}
-  \centering
-
-  \includegraphics[width=\myplotswidth]{fig/006919_input}
-  \includegraphics[width=\myplotswidth]{fig/006919_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_arriv}
-  \includegraphics[width=\myplotswidth]{fig/006919_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_kappa}
-  \includegraphics[width=\myplotswidth]{fig/006919_mass}
-
-  \caption[result 6919 (ASW0002z6f)]{Another configuration of arc plus
-    counter-image; here the arc is closer to the lensing galaxy than
-    the counter-image. (See \secref{example_models} for details.)}
-  \label{fig:6919}
 \end{figure}
@@
 In these figures, the scales and alignments in the six panels are not
-completely consistent: the three panels on the right exactly the same
+completely consistent: the three panels on the right have the same
 coordinates; the middle-left and lower-left panels are consistent with
@@
   lensing galaxy is a maximum, the image nearer to the galaxy is a
-  saddle point, and image further away is a minimum, and these were
-  correctly identified.
+  saddle point, and the image further away is a minimum.  All these
+  were correctly identified.  As note above, in
+  \secref{SpaghettiLens}, the precise shape of the loops in the
+  spaghetti diagrams is unimportant, only the implied image locations,
+  parities and time-ordering matters.
 
@@
   other side, generically arises from an elongated mass distribution
-  when the source is displaced along the elongated direction.
+  when the source is displaced along the elongated direction.  The
+  spaghetti diagram in this case has another markup element, a grey
+  point and circle overlaid on a probable secondary lensing galaxy.
+  This is an instruction to \spl to allow a point mass at that
+  location, distinct from the main mass map.
 
 \item Figure \ref{fig:6919} shows another example of an arc plus
-  counter-image, but (in contrast to Figure \ref{fig:6990}) the arc is
-  closer to the lens than the counter-image. This configuration arises
-  if the source displacement is perpendicular to the long axis of the
-  lensing mass.  Comparing the two panels in the middle row, we see
-  that the modeller interpreted arc as consisting of three images,
-  whereas the sim shows a single saddle point associated with the arc.
-  But the identification is not really erroneous --- we just need to
-  take into account that the source is extended.  In fact, in the sim
-  the brightest part of the source is only doubly imaged, but the
-  source extends into a region that produces four images.  In the
-  $t(x,y)$ of the sim, the hairpin-bending contours are typical of
-  double on the verge of splitting into a quad.
+  counter-image, but (in contrast to \figref{6990}) the arc is closer
+  to the lens than the counter-image. This configuration arises if the
+  source displacement is perpendicular to the long axis of the lensing
+  mass.  Comparing the two panels in the middle row, we see that the
+  modeller interpreted arc as consisting of three images, whereas the
+  sim shows a single saddle point associated with the arc.  But the
+  identification is not really erroneous --- we just need to take into
+  account that the source is extended.  In fact, in the sim the
+  brightest part of the source is only doubly imaged, but the source
+  extends into a region that produces four images.  In the $t(x,y)$ of
+  the sim, the hairpin-bending contours are typical of double on the
+  verge of splitting into a quad.
 
@@
 \begin{equation} \label{eq:effRE}
- \langle\kappa\rangle_{\Theta_E} = 1,
+ \langle\kappa\rangle_{\Theta_{\rm E}} = 1,
 \end{equation}
@@
 radii of the innermost and the outermost images: this is the
-confidence region from the modelling.  The definition (\ref{eq:effRE})
-for \ERf corresponds to crossing the dashed horizontal line at
-$10^0$: the red curve crosses the dashed line at the actual Einstein
-radius \ERf[,act]; the recovered Einstein radius
-\ERf[,rec] and its uncertainty are given by the blue curves
-crossing the dashed line.  We see that in all three panels, the blue
-curves are shallower than the red curve and \ERf[,rec] is
-more than \ERf[,act], by more than the model uncertainties.
-Now, steeper mass profiles tend to give wider image separations ---
-recall that the image separation for a circular isothermal lens is
-$2\Theta_E$, whereas for a point mass it is more \citep[see,
-  e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
+confidence region from the modelling.  The definition \eqref{eq:effRE}
+for \ERf corresponds to crossing the dashed horizontal line at $10^0$:
+the red curve crosses the dashed line at the actual Einstein radius
+\ERf[,act]; the recovered Einstein radius \ERf[,rec] and its
+uncertainty are given by the blue curves crossing the dashed line.  We
+see that in all three panels, the blue curves are shallower than the
+red curve and \ERf[,rec] is more than \ERf[,act], by more than the
+model uncertainties.  Now, steeper mass profiles tend to give wider
+image separations --- recall that the image separation for a circular
+isothermal lens is 2\ERf, whereas for a point mass it is more
+\citep[see, e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
 high is really a consequence of the \spl models being too shallow for
@@
 
-The lens-modeling challenge indicates that the Spacewarps collective
-is good, not only for identifying lens candidates for follow-up, but
-modelling candidates as well.
+This work has developed the concept of saddle-point contours in the
+travel time of virtual photons, originally introduced by
+\cite{1986ApJ...310..568B} for understanding image structure in strong
+gravitational lenses, into a technique for mass-mapping lenses.
+Despite being highly abstract, saddle-point contours look like
+schematic arcs, and hence lend themselves to an intuitive markup tool
+for lenses or lens candidates, which we call a spaghetti diagram.  At
+the same time, saddle-point contours encode information about possible
+mass distributions, which can be translated into input for an existing
+lens-modelling engine \citep[GLASS, by][]{2014arXiv1401.7990C}.
+
+\spl is an implementation of these ideas, enabling experienced but
+non-professional lens enthusiasts to model newly-discovered lens
+candidates from the \sw citizen-science platform.  The tests in this
+paper indicate that such modelling would be both feasible and
+scientifically interesting.
 
@@
 
-First, the particular modeling strategy implemented is not the only
-one possible.  \spl requires modelers to characterize the overall
-image structure in abstract terms based on Fermat's principle, and the
-placement of mass distributions is done by the computer.  In other
-modeling tools, the user puts down a trial mass distribution and has
-the machine refine it.  A few of these modeling programs have also
-been designed with citizen science in mind, and would be interesting
-in the Spacewarps environment.
-
-Second, \spl needs some enhancements.
-
-\begin{itemize}
+\begin{enumerate}
 \item Currently, \spl does not attempt to model the source shape; the
@@
   suspect features.
-\item \spl has a tendency to somewhat overestimate the Einstein radius
-  (evident from Figure \ref{fig:ER_per_sim}), and it is also apparent
-  that the models tend to be too shallow.  This possible explanation
-  is that, while the sims are steeply peaked at the centre, the
-  pixellated mass model fixes a comparatively large area near the
-  central at constant density.  One way to solve this problem would be
-  to introduce smaller pixels in the central region, thus enabled a
-  steeper centre.
+\item \spl tends to overestimate the Einstein radius (evident from
+  Figure \ref{fig:ER_per_sim}), and it is also apparent that the
+  models tend to be too shallow.  The likely explanation is that while
+  the sims are steeply peaked at the centre, the pixellated mass model
+  fixes a comparatively large area near the central at constant
+  density.  One way to solve this problem would be to introduce
+  smaller pixels in the central region, thus enabled a steeper centre.
 \item Another limitation so far in \spl is that the lens is assumed to
@@
   or in combination, at the user wishes.
-\end{itemize}
-
-The third desirable avenue of improvement is to facilitate
-collaborative work.
-
-\begin{itemize}
 \item As mentioned above, the option of revising an already-archived
@@
   correct ordering of the points in nearly-symmetric configurations.
-\end{itemize}
+\end{enumerate}
 

### 2014-07-04 systems/many.py
plotsplotsplots

@@ -31,9 +31,9 @@ def shear(xs,x,y):
 def poten_gal(gal,x,y):
     z = zeros((len(y),len(x)))
     re,ell,ell_pa = float(gal[1]), float(gal[2]), float(gal[3])
-    print 'Einstein radius',re
-    print 'ellipticity',ell
-    print 'pos ang',ell_pa
+    #print 'Einstein radius',re
+    #print 'ellipticity',ell
+    #print 'pos ang',ell_pa
     for i in range(len(x)):
         for j in range(len(y)):
             z[j,i] = poten_SIE(x[i],y[j],re,ell,ell_pa)

### 2014-07-04 upd_plots/plots.py

@@ -13,12 +13,12 @@ import numpy as np
 np.set_printoptions(precision=3)
 
 sys.path.append(os.path.abspath("../plots"))
-from defaults import AttrDict
+#from defaults import AttrDict
 
 
 def new_kappaplot(model, obj_index=0, **kwargs):
     """basically copied from glass plot.py, but heavily modified"""
-    print type(kwargs), kwargs    
+    #print type(kwargs), kwargs    
     kw = kwargs
 
     obj, data = model['obj,data'][obj_index]
@@ -38,7 +38,7 @@ def new_kappaplot(model, obj_index=0, **kwargs):
     #sl scaling factors
     rscale          = kw.pop('rscale', 440./500*100) # default Spacewarps -> SL -> glass scaling of radius
     dist_fact       = kw.pop('dist_fact', 0.428)     # because measurement at different redshiffts (we had 0.5/1 hardcoded in beginning)
-    delta           = kw['contour'].pop('cldelta', 0.1) # contour level spacing in log space
+    delta           = kw['kappa'].pop('cldelta', 0.1) # contour level spacing in log space
 
 
 
@@ -47,7 +47,7 @@ def new_kappaplot(model, obj_index=0, **kwargs):
 
     # fix the different scaling of glass and SL
     R = obj.basis.mapextent
-    extent = np.array([-R,R,-R,R]) / rscale
+    extent = np.array([-R,R,-R,R]) * rscale
 
     # prepare data
     grid = obj.basis._to_grid(data['kappa'],1)
@@ -70,7 +70,7 @@ def new_kappaplot(model, obj_index=0, **kwargs):
     # PLOTTING
 
     # setup
-    kw['contour'].setdefault('extent', extent)
+    kw['kappa'].contour.update({'extent': extent})
 #    kw.setdefault('extend', 'both') # contour levels are automatically added to one or both ends of the range so that all data are included
 #    #kw.setdefault('interpolation', 'nearest')
 #    kw.setdefault('aspect', 'equal')
@@ -80,10 +80,11 @@ def new_kappaplot(model, obj_index=0, **kwargs):
 #    kw.setdefault('antialiased', True)
 
     # plot
-    fig = plt.figure(**kw['figure'])
+    fig = plt.figure(**kw['kappa'].figure)
     ax = fig.add_subplot(1,1,1)
-    C = ax.contour(grid, levels=clevels, **kw['contour'])
-    ax.set_aspect('equal')
[…]

### 2014-07-04 upd_plots/updateplots.py

@@ -79,7 +79,6 @@ except:
 args = defaults.args
 
 
-
 inppath = "states"
 outpath = "outp"
 outfn = "Image for %(name)s.%(ext)s"
@@ -113,7 +112,7 @@ for fn, fe in files: # filename fn, fileextension fe
         args.file.ext = ext
         ofpath = os.path.join(outpath,outfn % args.file)
         #fig.savefig(ofpath, facecolor='black', edgecolor='none')
-        fig.savefig(ofpath, **args.savefig)
+        fig.savefig(ofpath, **args.kappa.savefig)
     
         plt.close()
       

### 2014-07-04 plots/defaults.py
new plots finally ready

@@ -83,7 +83,7 @@ args.mplrc = AttrDict({
 
 args.figure = AttrDict({
     'figsize'   : (8,6),    # w,h tuple in inches
-    'dpi'       : 80,       # Dots per inch
+    'dpi'       : 300,       # Dots per inch
     'facecolor' : 'w',      # The figure patch facecolor; defaults to rc figure.facecolor
     'edgecolor' : 'k',      # The figure patch edge color; defaults to rc figure.edgecolor
 #    'linewidth' : 1,        # The figure patch edge linewidth; the default linewidth of the frame
@@ -279,8 +279,8 @@ def detailed_log_capt(fig=None, axes=None, lines=None):
 
 #collect the formatter
 args.formatter = AttrDict({
-#    'removeticks'   :   remove_all_ticks,
-    'removeticks'   :   lambda axes: None,
+    'removeticks'   :   remove_all_ticks,
+#    'removeticks'   :   lambda axes: None,
     'detailed_log_capt' : detailed_log_capt,   
 })
 

### 2014-07-04 plots/gen_plots.py

@@ -90,7 +90,7 @@ simplots = {'arriv': True, 'kappa':True, 'kappaenc':False}
 
 #image extension (png or pdf)
 #exts = ['png', 'pdf']
-exts = ['png']
+exts = kw.exts
 
 #figsize   = (10,10) 
 #figsizeER = (10,8)
@@ -1135,11 +1135,10 @@ def draw_sim(asw, sim):
         path = os.path.join(simdir, '%s_%06i' % (asw, modid))
         print "  making plots for %s with scales from %06i" % (asw, modid)
     
-        map_ext=scales[asw][modid]['map_e']['arcsec']
+        map_ext = scales[asw][modid]['map_e']['arcsec']
         R = map_ext * div_scale_factors
 
-        #print map_ext
-        #print R
+        print 'me', map_ext, R
 
         x = np.linspace(-R,R,N)
         y = 1*x
@@ -1163,21 +1162,22 @@ def draw_sim(asw, sim):
             
             # kappa(x,y) is value at (x,y), but pcolormesh needs edge coordinates of patches
             # kappa(xi, yj) => x[i]-d / y[j]-d ... x[i+1]-d / y[j+1]-d ; with d step/cell width
-            d=2.*R/(N-1)
-            x_cm=np.linspace(-R-d/2.,R+d/2.,N+1)
-            y_cm=1.*x_cm
+#            d=2.*R/(N-1)
+#            x_cm=np.linspace(-R-d/2.,R+d/2.,N+1)
+#            y_cm=1.*x_cm
             
             # cut border of 0 pixels, delete rrr rows on each side
             rr= kw.kappa.rmnrow
             kappa_cut = kappa[rr:-rr,rr:-rr]
             x_cut=x[rr:-rr] #original x,y will be used later
             y_cut=y[rr:-rr] 
-            x_cm=x_cm[rr:-rr] #x_cm not, just overwrite
-            y_cm=y_cm[rr:-rr]
+#            x_cm=x_cm[rr:-rr] #x_cm not, just overwrite
+#            y_cm=y_cm[rr:-rr]
             
             cldelta = kw.kappa.cldelta
             
-            vals = 2.5*np.log10(kappa_cut+eps)
+            #vals = 2.5*np.log10(kappa_cut+eps)
+            vals = np.log10(kappa_cut+eps)
             
             ma = np.nanmax(vals)
             mi = np.nanmin(vals)

### 2014-07-04 upd_plots/plots.py

@@ -84,7 +84,7 @@ def new_kappaplot(model, obj_index=0, **kwargs):
     ax = fig.add_subplot(1,1,1)
     ax.contour(grid, levels=clevels, **kw['kappa'].contour)
     kw['formatter'].removeticks(axes=ax)
-    #ax.set_aspect('equal')
+    ax.set_aspect('equal')
     return fig
 
 

### 2014-07-04 upd_plots/updateplots.py

@@ -81,7 +81,7 @@ args = defaults.args
 
 inppath = "states"
 outpath = "outp"
-outfn = "Image for %(name)s.%(ext)s"
+outfn = "%(name)s_mass.%(ext)s"
 exts = args.pop('exts', ['png'])
 
 sel_mod = args.pop('sel_mods', None)
@@ -117,7 +117,7 @@ for fn, fe in files: # filename fn, fileextension fe
         plt.close()
       
     print " \--- DONE"
-    break
+
   
 
 

### 2014-07-27 text/ms.tex
new plots finally ready
cleanup and regen of figs
Tidying prior to editing

@@
+% ====================================================================
+
 \documentclass[usenatbib]{mn2e}
@@
 
+\definecolor{purple}{RGB}{150,0,200}
+\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf Phil: #1}}
+
 % \\\ END OF REMOVAL SECION
@@
 
+% ====================================================================
+
 \begin{document}
@@
 \author[Kueng et al]{Rafael Kueng,$^{1}$
+Prasenjit Saha$^{1}$
+Elisabeth Baeten,$^{5}$
+Claude Cornen,$^{5}$
+\newauthor
+Christine McMillan,$^{5}$
+Julianne Wilcox,$^{5}$
+Jonas Odermatt,$^{6}$
+Jonathan Coles,$^{7}$
+\newauthor
 Phil Marshall,$^{2}$
@@
 Aprajita Verma,$^{4}$
-\newauthor
-Elisabeth Baeten,$^{5}$
-Claude Cornen,$^{5}$
-Christine McMillan,$^{5}$
-Julianne Wilcox,$^{5}$
-\newauthor
-Jonas Odermatt,$^{6}$
-Jonathan Coles,$^{7}$
-Surhud More$^{3}$
-and Prasenjit Saha$^{1}$ \\
+and Surhud More$^{3}$ \\
 $^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
@@
 \end{keywords}
+
+% ====================================================================
 
@@
 
+% ====================================================================
+
 \section{Fermat's principle and spaghetti diagrams} \label{sec:Fermat}
@@
 \cite{1986ApJ...310..568B}.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
@@
 
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
 \subsection{Arrival-time contours} \label{sec:arriv}
@@
 
+% ====================================================================
+
 \section{A lens-modelling program} \label{sec:SpaghettiLens}
@@
 
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
 \subsection{Image markup}
@@
 masses.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
@@
 
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
 \subsection{Diagnostics} \label{sec:diag}
@@
 
+% ====================================================================
+
 \section{A lens modelling challenge} \label{sec:mod_challenge}
@@
 comparison of the effective Einstein radii of the sims and the models.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
@@
 analysis.  Results from the latter now follow.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
@@
   \includegraphics[width=\myplotswidth]{fig/006941_input}
-  \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/006941_arr_time_ipol} \\
   \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 
@@
   \includegraphics[width=\myplotswidth]{fig/006990_input}
-  \includegraphics[width=\myplotswidth]{fig/006990_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/006990_arr_time_ipol} \\
   \includegraphics[width=\myplotswidth]{fig/ASW0004oux_006990_arriv}
@@
   \includegraphics[width=\myplotswidth]{fig/006919_input}
-  \includegraphics[width=\myplotswidth]{fig/006919_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/006919_arr_time_ipol} \\
   \includegraphics[width=\myplotswidth]{fig/ASW0002z6f_006919_arriv}
@@
   \includegraphics[width=\myplotswidth]{fig/006915_input}
-  \includegraphics[width=\myplotswidth]{fig/006915_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/006915_arr_time_ipol} \\
   \includegraphics[width=\myplotswidth]{fig/ASW0001hpf_006915_arriv}
@@
   \includegraphics[width=\myplotswidth]{fig/006975_input}
-  \includegraphics[width=\myplotswidth]{fig/006975_arr_time} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_arriv.png}
+  \includegraphics[width=\myplotswidth]{fig/006975_arr_time_ipol} \\
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_arriv}
   \includegraphics[width=\myplotswidth]{fig/006975_spaghetti} \\
-  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_kappa.png}
+  \includegraphics[width=\myplotswidth]{fig/ASW000195x_006975_kappa}
   \includegraphics[width=\myplotswidth]{fig/006975_mass}
@@
   \includegraphics[width=\myplotswidth]{fig/006937_input}
-  \includegraphics[width=\myplotswidth]{fig/006937_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/006937_arr_time_ipol} \\
   \includegraphics[width=\myplotswidth]{fig/ASW0000vqg_006937_arriv}
@@
   \includegraphics[width=\myplotswidth]{fig/007025_input}
-  \includegraphics[width=\myplotswidth]{fig/007025_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/007025_arr_time_ipol} \\
   \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007025_arriv}
@@
   \includegraphics[width=\myplotswidth]{fig/007022_input}
-  \includegraphics[width=\myplotswidth]{fig/007022_arr_time} \\
+  \includegraphics[width=\myplotswidth]{fig/007022_arr_time_ipol} \\
   \includegraphics[width=\myplotswidth]{fig/ASW0000h2m_007022_arriv}
@@
 \end{itemize}
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
@@
 \end{figure}
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
 
@@
 the model compared to the simulation.
+
+% ====================================================================
 
@@
 
+% ====================================================================
+
 \section*{Acknowledgments}
@@
 from the Leverhulme Trust.
+
+% ====================================================================
 
@@
 
+% ====================================================================
+
 \bibliographystyle{mn2e}
@@
 
+% ====================================================================
+
 \end{document}
+
+% ====================================================================

### 2014-07-27 text/ms.tex
Reordered authors to show properly Prasenjit's role, and to give the modellers full credit. Fixes #20. Also beefed up abstract a bit, addressing #19

@@
 \definecolor{purple}{RGB}{150,0,200}
-\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf Phil: #1}}
+\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}
 
@@
 
-\title{Lens Modeling in a Citizen-Science context}
+\title{Gravitational Lens Modeling in a Citizen Science Context}
 
 \author[Kueng et al]{Rafael Kueng,$^{1}$
-Prasenjit Saha$^{1}$
-Elisabeth Baeten,$^{5}$
-Claude Cornen,$^{5}$
+Prasenjit Saha,$^{1}$
+Elisabeth Baeten,$^{2}$
+Claude Cornen,$^{2}$
 \newauthor
-Christine McMillan,$^{5}$
-Julianne Wilcox,$^{5}$
-Jonas Odermatt,$^{6}$
-Jonathan Coles,$^{7}$
+Christine McMillan,$^{2}$
+Julianne Wilcox,$^{2}$
+Jonas Odermatt,$^{3}$
+Jonathan Coles,$^{4}$
 \newauthor
-Phil Marshall,$^{2}$
-Anupreeta More,$^{3}$
-Aprajita Verma,$^{4}$
-and Surhud More$^{3}$ \\
+Aprajita Verma,$^{5}$
+Anupreeta More,$^{6}$
+Surhud More,$^{6}$
+and Phil Marshall$^{7}$ \\
 $^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
-$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
-$^{3}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
-$^{4}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
-$^{5}$Spacewarps\\
-$^{6}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
-$^{7}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France}
-%$^{8}$Adler Planetarium, 1300 S. Lake Shore Drive, Chicago, IL 60605, USA\\
+$^{2}$The Space Warps Collaboration\\
+$^{3}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
+$^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
+$^{5}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
+$^{6}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
+$^{7}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
+}
 
@@
 \begin{abstract}
+
 We develop a method to enable collaborative modelling of gravitational
-lenses and lens candidates by experienced but non-professional lens
-enthusiasts.  It uses an existing free-form modelling program (GLASS)
-but provides the modelling input in a novel way, via a user-provided
-diagram that is essentially a sketch of an arrival-time surface.
-
-An implementation (\spl) is tested in a modelling challenge using 29
-simulated lenses out of a larger set created for a citizen-science
-lens search (\sw).  We find that amateur lensers inferred the image
-parities and time ordering consistently in some lenses, while tending
-to make errors in other lenses, depending on the image morphology.
-Errors in image parity and time ordering lead to large errors in the
-mass distribution, but the enclosed mass is more robust.  Shortcomings
-include (a)~lens profiles that tend to be systematically shallower
-than the lenses, and consequent over-estimation of the Einstein radius
-by $\sim25\%$, and (b)~lack of detailed source modelling.  Ideas for
-improvement are discussed.
+lenses, and lens candidates, by non-professional lens enthusiasts.  It
+uses an existing free-form modelling program (GLASS), but
+\phil{enables the input to this code to be provided in a novel way,}
+via a user-provided diagram that is essentially a sketch of an
+arrival-time surface.
+
+An implementation \phil{of this method, \spl} is tested in a modelling
+challenge using 29 simulated lenses \phil{drawn from the} set created
+for \phil{the \sw citizen-science lens search}.  We find that
+\phil{volunteers from this online community} asserted the image
+parities and time ordering consistently in some lenses, but made
+errors in other lenses depending on the image morphology. \phil{While}
+errors in image parity and time ordering lead to  large errors in the
+mass distribution, the enclosed mass \phil{was found to be more
+robust: the model-derived Einstein radii found by the volunteers were
+consistent with those produced by one of the professional team,
+suggesting that given the appropriate tools, gravitational lens
+modelling is a data analysis activity that can be crowd-sourced.}
+Shortcomings \phil{of the implementation itself include} (a)~lens
+profiles that tend to be systematically shallower than the lenses
+\phil{(leading to systematic over-estimation of the Einstein radius 
+by some $\sim25\%$}, and (b)~lack of detailed \phil{arc} modelling. 
+Ideas for improvement are discussed.
+
 \end{abstract}

### 2014-07-27 text/ms.tex
Edited introduction to remove detail on SW that will come in system papers, and to re-emphasise volunteer roles

@@
 \newcommand{\spl}{SpaghettiLens\xspace}
-\newcommand{\sw}{Spacewarps\xspace}
+\newcommand{\sw}{Space Warps\xspace}
 
@@
 \citep{1979Natur.279..381W,1980Natur.285..641W}, where a galaxy images
-a background quasar into two (four), brought in their wake the first
+a single background quasar into two (four), brought in their wake the
+first
 work on lens modelling
@@
 galaxy; typically the lensed images are seen through the galaxy halo.
-Hence the lensed images depend on the detailed mass distribution of
+Hence, the lensed images depend on the detailed mass distribution of
 the lensing galaxy.  Galaxy lenses demand modelling.
 
-Since the early discoveries, more than 400 secure lenses are now known
-(see Figure \ref{fig:masterlens}).  Modelling of the mass distribution
-is part of any research using lenses, but so far no modelling study
-has spanned all known lenses.  The largest single one
-\citep{2009ApJ...703L..51K} models 58 lenses to infer the distribution
-of dark matter around galaxies.  In other work,
-\cite{2011ApJ...740...97L} combined lens models of 21 galaxies with
-models of their stellar populations, to find the relation between
-stars and dark matter, and \cite{2014MNRAS.437..600S} modelled 18
-time-delay lenses together to infer cosmological parameters.
-
-\begin{figure}
-\centering
-\includegraphics[width=\hsize]{fig/lenssky}
-\caption{Sky distribution of 423 published lenses considered secure
-  (from the Masterlens catalog at the University of Utah, maintained
-  by Joel Brownstein and Leonidas Moustakas).  The map is in
-  Hammer-Aitoff projection, with North up and $\rm RA=0$ in the
-  middle. The empty swathes to the left and right of centre are the
-  Milky Way. {\em Should we drop this figure?  It's not relevant to
-    modelling, but as there seems to be no publicly available summary
-    of all known lenses, it does help set context.}}
-\label{fig:masterlens}
-\end{figure}
-
-Surveys now under way aim to increase the inventory of lenses another
-ten or a hundred fold.  Among these is \sw (Marshall et al, in prep,
-More et al, in prep).  \sw is a citizen science project\footnote{\tt
-  http://www.spacewarps.org} in which volunteers are presented
-sky-survey images and invited to identify lens candidates.  Simulated
-lenses are mixed in with the data, both to help train volunteers on
-what to look for, and to provide measures of the effectiveness of the
-search.  The motivation for \sw is to enable volunteers, some of whom
-had previously serendipitously identified lens candidates on earlier
-citizen-science surveys, to make discoveries missed in automatic
-searches by software robots.  Robots are very good with clean lensing
-system in uncrowded fields with high signal-to-noise, but in general
-test situations \citep{2009ApJ...694..924M} robots miss lenses (low
-completeness) or contaminate the results with non-lenses (low purity).
-\sw introduced its first tranche of survey data, in May 2013, from the
-Canada-France-Hawaii Telescope Legacy Survey, or CFHTLS
-\citep{Coupon2009}, covering $\simeq172$ square degrees (0.4\% of the
-sky) was divided into tiles of $\simeq 80''\times80''$ ($440\times
-440$ pixels).  Each such tile was shown to ten volunteers.  Despite
-the rarity of detectable lenses ($<10^{-3}$ per tile), the \sw
-collective soon found lens candidates, both rediscoveries already
-known from robots \citep{Gavazzi2012, More2012ApJ} and possible new
-lenses.
-
-The encouraging early results prompted the question: could modelling
-of the lenses also be done by volunteers?  If so, modelling could help
-assess prioritise lens candidates at an early stage, which would be
-very useful with new wide-field and sensitive surveys, which will
-yield thousands of lens candidates. There are several software tools
-for lensing modelling available, and work has been done on generic
-interfaces \citep{2014A&C.....5...28L}.  Some early designs for \sw
-included a prototype lens modelling tool \citep{2010AAS...21543527N}.
-Moreover, some {\em Spacewarps\/} volunteers are quite experienced
-from earlier projects, having spent a thousand hours or more with
-data, and are very interested in more demanding projects.  The
-interests of citizen-science communities are just beginning to be
-studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that
-some volunteers welcome open-ended challenges, and sometimes these
-have led to new scientific results: one example is the discovery of an
-exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
-the development of new algorithms for protein folding
-\citep{Khatib22112011}.  All these are grounds for optimism.  There
-is, however, a basic difficulty in strong gravitational lensing.
-Lensed images do not look much like their source, and still less do
-they resemble the lensing-mass distribution.  To model a lens, one
-needs either to do a lot of random guessing, or to have a good
-intuition for what works.
+Since the early discoveries, more than 400 secure lenses are now
+known. Modelling of the mass distribution is part of any research
+using lenses, but so far no modelling study has spanned all known
+lenses.  The largest single one \citep{2009ApJ...703L..51K} models 58
+lenses to infer the distribution of dark matter around galaxies.  In
+other work, \cite{2011ApJ...740...97L} combined lens models of 21
+galaxies with models of their stellar populations, to find the
+relation between stars and dark matter, and \cite{2014MNRAS.437..600S}
+modelled 18 time-delay lenses together to infer cosmological
+parameters.
+
+Imaging surveys now under way aim to increase the inventory of lenses
+another ten or a hundred fold \phil{\citep[see
+e.g.][]{Marshall2005,OguriMarshall2010}, with both automated and
+visual search techniques proposed
+\citep[e.g.][]{Marshall2009,More2012ApJ,Gavazzi2014}. For example, \sw
+(Marshall et al, in prep, More et al, in prep) is a citizen science
+project\footnote{\tt http://www.spacewarps.org} in which volunteers
+are presented sky-survey images and invited to identify lens
+candidates, by eye.}  Simulated lenses are mixed in with the data,
+both to help train volunteers on what to look for, and to provide
+measures of the effectiveness of the search.  The motivation for \sw
+is to enable volunteers, some of whom had previously serendipitously
+identified lens candidates on earlier citizen-science surveys,
+\phil{either} to make discoveries missed in automatic searches by
+software robots, \phil{or to perform the necessary inspection of an
+automatically-generated sample, for quality control}.  Robots can be
+built to be good at detecting clean lensing system in uncrowded fields
+with high signal-to-noise, but in more general test situations robots
+miss lenses (low completeness) or contaminate the results with
+non-lenses (low purity) \citep{Marshall2009}.
+
+\phil{The encouraging early results from the first \sw lens search,
+carried out on the  $\simeq172$ square degree Canada-France-Hawaii
+Telescope Legacy Survey (CFHTLS) imaging by over 30,000 volunteers
+(Marshall et al, in prep, More et al, in prep) prompted the question:}
+could the modelling of the lenses also be done by the volunteers?  If
+so, modelling could help assess prioritise lens candidates at an early
+stage, which would be very useful with new wide-field and sensitive
+surveys, which will yield thousands of lens candidates. There are
+several software tools for lensing modelling available, and work has
+been done on generic interfaces \citep[e.g.][]{2014A&C.....5...28L}. 
+Indeed, some early designs for \sw included a prototype lens modelling
+tool \citep{2010AAS...21543527N}. Moreover, some \sw volunteers are
+quite experienced from earlier projects, having spent a thousand hours
+or more with data, and are very interested in more demanding
+projects.  The interests of citizen-science communities are just
+beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it
+is clear that some volunteers welcome open-ended challenges, and
+sometimes these have led to new scientific results: one example is the
+discovery of an exceptional extra-solar planet
+\citep{2013ApJ...768..127S}; another is the development of new
+algorithms for protein folding \citep{Khatib22112011}.  All these are
+grounds for optimism.  There is, however, a basic difficulty in strong
+gravitational lensing. Lensed images do not look much like their
+source, and still less do they resemble the lensing-mass
+distribution.  To model a lens, one needs either to do a lot of random
+guessing, or to have a good intuition for what works.
 
@@
 In \secref{Fermat} we introduce a markup system for lensed images,
-which we call a ``spaghetti diagram''.  A spaghetti diagram resembles
+which we call a ``spaghetti diagram.''  A spaghetti diagram resembles
 the visible image system, in a cartoon-like way, and at the same time
@@
 In \secref{SpaghettiLens} we describe the \spl program, which
-implements everything.  \spl is an extension of the GLASS framework
-for modelling lenses \citep{2014arXiv1401.7990C}.  We will not go into
-software details in this paper, concentrating instead on lens
-modelling per se, but we remark that \spl is designed to be friendly
-to the forum style of citizen-science projects, and enable incremental
-model refinement by different people, without sacrificing any features
-of GLASS.
+implements \phil{the above scheme}.  \spl is an \phil{interface to and
+extension of} the GLASS framework for modelling lenses
+\citep{2014arXiv1401.7990C}.  We will not go into software details in
+this paper, but instead concentrate on lens modelling per se, but we
+remark that \spl is designed to be friendly to the forum style of
+citizen-science projects, and enable incremental \phil{collaborative}
+model refinement by different people, without sacrificing any of the
+technical features of GLASS.
 
@@
 diverse sample of 29 simulated lenses was modelled multiple times
-using \spl.  The models are then examined in two ways.  One was
-whether the spaghetti diagram was correct.  The other was the recovery
-of the Einstein radius.  In addition, we show some visual comparisons
-of actual and recovered lens shape and radial profile, and identify
-some areas to improve, but do not compute statistics on these aspects.
-Profile and shape recovery with GLASS has been studied in more detail
-in \citep{2014arXiv1401.7990C}.
+\phil{by a small number of \sw volunteers} using \spl.  The models
+were then examined in two ways.  One was whether the spaghetti diagram
+was correct.  The other was the recovery of the Einstein radius of the
+lens.  In addition, we show some visual comparisons of the actual and
+recovered lens shape and radial profile, and identify some areas to
+improve. Profile and shape recovery with GLASS has been studied in
+more detail in \citep{2014arXiv1401.7990C}.
 
@@
 
-We now explain the lensing theory relevant to \spl, following the
+We first explain the lensing theory relevant to \spl, following the
 formulation of gravitational lensing in terms of Fermat's principle by
@@
 \end{matrix} $$
-The marked-up image is a zoom of the lensed image on Spacewarps,
+The marked-up image is a zoom of the lensed image on \sw,
 marked up with one or more spaghetti contours; this is the modeller's

### 2014-07-27 text/ms.tex
Tightened final figure discussion, strengthened conclusions and abstract: fixed #18, fixed #19, fixed #21

@@
 suggesting that given the appropriate tools, gravitational lens
-modelling is a data analysis activity that can be crowd-sourced.}
-Shortcomings \phil{of the implementation itself include} (a)~lens
-profiles that tend to be systematically shallower than the lenses
-\phil{(leading to systematic over-estimation of the Einstein radius 
-by some $\sim25\%$}, and (b)~lack of detailed \phil{arc} modelling. 
-Ideas for improvement are discussed.
+modelling is a data analysis activity that can be crowd-sourced to
+good effect.} Shortcomings \phil{of the software implementation 
+include} (a)~lens profiles that tend to be systematically shallower
+than the lenses \phil{(leading to systematic over-estimation of the
+Einstein radius  by some $\sim25\%$}, and (b)~lack of detailed
+\phil{arc} modelling.  Ideas for improvement are discussed.
 
@@
 
-\section{A lens-modelling program} \label{sec:SpaghettiLens}
+\section{A lens-modelling program} 
+\label{sec:SpaghettiLens}
 
 \spl is not part of \sw, but it makes use of the \sw infrastructure,
-in particular, the image database and the forum.  The forum is
-essential essential for initial contact with interested members of the
-\sw community.  Lens modelling from \sw works, in practice, in the
-usual style of medium-sized astronomical collaborations, with
-video-conferencing and in-person meetings where possible. But
-preliminary results are immediately summarized on modelling threads on
-the forum, and anyone interested is welcome to join at any time.
-Another aspect of \sw (and of the Zooniverse family of projects in
-general) that \spl adopts, is to keep all user interaction in a web
-browser, with intensive computing done on a server.  Several
-interesting design and implementation issues arise from these
-considerations, but we will not go into them in this paper,
-concentrating instead on aspects specific to lens modelling.
-
+in particular, the image database and the \phil{discussion
+forum.\footnote{\texttt{http://talk.spacewarps.org}}  The forum is
+essential for establishing contact between interested members of the
+\sw community and the project science team, and then for enabling
+collaboration between them. We were able to collaborate together on 
+modelling objects from \sw in the usual style of medium-sized
+astronomical collaborations, with video-conferencing and in-person
+meetings where possible. Preliminary results were immediately
+summarized on modelling threads on the forum, and anyone interested
+was made welcome to join at any time.} 
+%
+% PJM: The following paragraph is not accurate: the manipulation of 
+%  FITS images in the Tools viewer, for example, is done client side.
+%  In future I very much hope to see lens modeling enabled in this
+%  mode as well!
+%
+% Another aspect of \sw (and of the Zooniverse family of projects in
+% general) that \spl adopts, is to keep all user interaction in a web
+% browser, with intensive computing done on a server.  Several
+% interesting design and implementation issues arise from these
+% considerations, but we will not go into them in this paper,
+% concentrating instead on aspects specific to lens modelling.
+%
 Modelling with \spl involves three stages, markup of the image,
-followed by intensive numerics in the background, followed by review
+followed by intensive \phil{numerical computation carried out on a
+server in the background,} followed by review
 of diagnostics and possible discussion.  We now describe each of
@@
 
-One begins by going to the \spl page and entering the number of a \sw
-image tile.  \spl presents the image, along with zoom and pan options
-and a markup tool to construct a spaghetti diagram.  The human
+One begins by going to the \spl \phil{web application\footnote{\bf Phil
+asks: should we provide a spl url here?}} and entering the number of a
+\sw image tile.  \spl then presents the image, along with zoom and pan
+options and a markup tool to construct a spaghetti diagram.  The human
 modeller now has to make an educated guess for the topography of the
@@
 maxima, minima, and saddle-points.  The markup tool \citep[which is
-  inspired by Figure~6 of][and is like that figure made interactive
-  and overlaid on data]{1986ApJ...310..568B} lets the modeller enter
-the information by sketching saddle-point contours.  Examples can be
-seen in the upper-left panels of Figures \ref{fig:6941} to
-\ref{fig:7022}.  The loops in the markup tool were the origin of the
-``spaghetti'' metaphor.
+inspired by Figure~6 of][and is like that figure made interactive and
+overlaid on data]{1986ApJ...310..568B} lets the modeller enter the
+information by sketching saddle-point contours.  Examples can be seen
+in the upper-left panels of Figures \ref{fig:6941} to \ref{fig:7022}. 
+The loops in the markup tool were the origin of the ``spaghetti''
+metaphor.
 
@@
 theorem.  The exact placement of the loops in a spaghetti diagram has
-no significance.  Only the heirarchy of which loop is inside which is
+no significance.  Only the hierarchy of which loop is inside which is
 relevant.  The loops are there simply to help modeller's intuition.
 
+{\bf Phil asks: how much training was necessary before the volunteers
+understood how to work the interface? Could this training be done with
+no input from the \spl developers, in future? It would be good to
+comment on how scalable this approach to modeling is.}
+
 As implemented so far, \spl assumes that the lens is dominated by a
-single galaxy.  Only one maximum is permitted, and it is taken as the
-centre of the main lensing galaxy.  The user can, however, mark
-additional minor galaxies.  The latter are modelled as point
-masses.
+single galaxy.  Only one maximum \phil{in the lens mass distrbution}
+is permitted, and it is taken to be the centre of the main lensing
+galaxy.  The user can, however, mark additional minor galaxies: these
+are modelled as point masses.  {\bf Phil asks: with what mass? Are
+these masses allowed to vary freely during the fit?}
 
@@
 
-\subsection{Diagnostics} \label{sec:diag}
+\subsection{Diagnostics} 
+\label{sec:diag}
 
@@
 \end{enumerate}
-The synthetic image generated by \spl assumes a simple conical source
-profile.  The user can change the contrast level on the image, which
-(though it is not saved) amounts to adjusting the width of the cone.
-These synthetic images are still very crude and not very useful for
-assessing models.  The best indicator, in practice, of whether the
-modelling was successful is contour map of $t(x,y)$, with saddle-point
-contours highlighted.  It is, in effect, the computer's refinement of
-the spaghetti diagram input by the user.  If the arrival-time surface
-looks qualitatively similar to the spaghetti diagram, that generally
-indicates a successful model.  The mass distribution also provides
-indications; successful models generally lead to smooth-looking mass
-distribution, whereas an irregular or checkboard pattern in the mass
-map signals a bad model.
-
-After examining this feedback, the user can choose to post the model
-on the \spl archive.  They can also modify their input and try again,
-or discard the attempt altogether.  After archiving, there can be
-discussion among modellers, through the \sw forum or by any other
-means, and revision of the model.  Any archived model can be revised
-by any user: they can modify the spaghetti configuration slightly or
-drastically, or change options like the size of the mass tiles.
-Particularly interesting lens candidates lead to trees of models in
-this way.  Discussion among modellers tends to prune a model tree,
-focusing attention on the most interesting models.\footnote{See
-  ``Collaborative gravitational lens modelling\dots'' in {\tt
-    http://letters.zooniverse.org} for an example.}
+The synthetic image generated by \spl assumes a simple
+\phil{circularly symmetric} source \phil{with linearly decreasing
+surface brightness profile}.  The user can change the contrast level
+on the image, which (though it is not saved) amounts to adjusting the
+\phil{size of the source}. These synthetic images are still very
+crude, and \phil{not always} useful for assessing models.  The best
+indicator, in practice, of whether the modelling was successful is
+contour map of $t(x,y)$, with saddle-point contours highlighted.  It
+is, in effect, the computer's refinement of the spaghetti diagram
+input by the user.  If the arrival-time surface looks qualitatively
+similar to the spaghetti diagram, that generally indicates a
+successful model.  The mass distribution also provides indications;
+successful models generally lead to smooth-looking mass distribution,
+whereas an irregular or checkboard pattern in the mass map signals a
+bad model.
+
+After examining this feedback, the user can choose to \phil{save the
+model to the \spl archive, at which point it is assigned a unique
+URL}.  They can also modify their input and try again, or discard the
+attempt altogether.  After archiving, there can be discussion among
+modellers, through the \sw forum or by any other means, and revision
+of the model.  \phil{This is achieved simply by sharing the model's
+URL; following its hyperlink takes one to the \spl app, pre-loaded
+with the correct data image and input spaghetti.} Any archived model
+can be revised by any user: they can modify the spaghetti
+configuration slightly or drastically, or change options like the size
+of the mass tiles. Particularly interesting lens candidates lead to
+trees of models in this way.  Discussion among modellers tends to
+prune a model tree, focusing attention on the most interesting
+models.
+% PJM: sadly this Letter is no longer available, Julianne took it
+% down. We can post the talk URL instead.
+% \footnote{See ``Collaborative gravitational lens
+% modelling\dots'' in {\tt http://letters.zooniverse.org} for an
+% example.}
+\footnote{See the ``collaborative modelling'' thread at 
+{\tt http://talk.spacewarps.org} for an example.}
+% TODO: fix url!
 
@@
 
+% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
+% THIS PARAGRAPH TO BE REMOVED WHEN THE FIGURES ARE FIXED!
+%
 In these figures, the scales and alignments in the six panels are not
@@
 varying scales and alignments are harmless.
+% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
 
@@
 \begin{table}\centering\begin{tabular}{lr}\hline
-    Total & 119 \\
+    Total & 119 (100\%)\\
 \hline
-    errors in image locations  & 9 \\
-    errors in image parities or time ordering & 49 \\
+    errors in image locations  & 9 (8\%)\\
+    errors in image parities or time ordering & 49 (41\%) \\
 \hline
-    inaccurate image placement over an arc & 21 \\
-    identified two images of four & 5 \\
-    identified two nearby images as one & 3 \\
-    missed faint images & 1 \\
-    proposed too many images & 1 \\
-    modelled a three-image arc as one image & 4 \\
-    modelled one image as a three-image arc & 5 \\
-    swapped minimum and saddle in double  & 2 \\
-    swapped minima and saddles in quad & 38 \\
-    swapped early and late saddles in a quad & 7 \\
+    inaccurate image placement over an arc & 21 (18\%) \\
+    identified two images of four & 5 (4\%) \\
+    identified two nearby images as one & 3 (3\%) \\
+    missed faint images & 1 (1\%) \\
+    proposed too many images & 1 (1\%) \\
+    modelled a three-image arc as one image & 4 (3\%) \\
+    modelled one image as a three-image arc & 5 (4\%) \\
+    swapped minimum and saddle in double  & 2 (2\%) \\
+    swapped minima and saddles in quad & 38 (32\%)\\
+    swapped early and late saddles in a quad & 7 (6\%) \\
 \hline
@@
 
-The images of the system are considered to have been located
-correctly, if all the images have been identified and are approximated
+The images of the system were considered to have been located
+correctly, if all the images were identified and were approximated
 within about 5\% of \spl frame used to draw the spaghetti diagram.
@@
 recognised and are listed in \tabref{stats}.  Most of the problems
-where due to unclear arc-like structures.  Critical errors like the
+were due to unclear arc-like structures.  Critical errors like the
 failure to identify all five images in a five images system, or to
@@
 The assignment of the parity of the images was a more difficult task,
-and was successful in only 60\% of the cases.  The most common error
-was swapping of minima and saddle-points in a quad; \figref{7025}
-shows an example.  Another, less common, error was flipping the
-spaghetti diagram, this swapping the time-ordering of the two saddle
-points.
+and was successful in only about 60\% of the cases.  The most common
+error was swapping of minima and saddle-points in a quad;
+\figref{7025} shows an example.  Another, less common, error was
+flipping the spaghetti diagram, this swapping the time-ordering of the
+two saddle points.
 
 Incorrect image parities and time orderings tended to produce
-poorer-looking models, such as the checkboard patterns in the mass map
-in \figref{7025}.  Interestingly, however, the enclosed-mass profiles
-were quite robust.  We will consider this aspect in the next section.
+poorer-looking models, such as the checkerboard patterns in the mass
+map in \figref{7025}.  Interestingly, however, the enclosed-mass
+profiles were quite robust.  We will consider this aspect in the next
+section.
 
@@
 \citep[see, e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
-high is really a consequence of the \spl models being too shallow for
+high is really a consequence of the \phil{GLASS} models being too shallow for
 the sims.
@@
 
-\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend
-to be too high.  Discounting the models which were flagged by the
-users as poor, the mean and standard deviation of the overestimate are
-25\% and 20\%.  The models by an expert (PS) also show a similar bias,
-having mean and standard deviation of the overestimate being 20\% and
-8\%.  One of the reasons for this is that it is hard to get the center
-of the lens on spot.  An offset leads to a flatter mass profile for
+% TODO: update figure, fix caption.
+
+\Figref{ER_per_sim} shows that \ERf[,rec] of the models tend to be too
+high.  \phil{However, this is entirely due to the GLASS model density
+profiles being too shallow, as illustrated above. We can separate out
+the performance of the \spl interface and its users by comparing their
+results with the  Einstein radii of \spl models made by an expert
+(PS). Discounting the models which were flagged by the volunteers as
+poor, the mean Einstein radius overestimate was 25\%, with a 20\%
+standard deviation (shown by the grey band in \Figref{ER_per_sim}). 
+The expert models show a statistically indistinguishable bias of 20\%,
+with standard deviation 8\% (the blue band in \Figref{ER_per_sim}).  }
+One source of this systematic error is that it is difficult to center
+of the lens accurately: an offset leads to a flatter mass profile for
 the model compared to the simulation.
@@
 paper indicate that such modelling would be both feasible and
-scientifically interesting.
-
-There is, however, plenty of room for improvement.
-
+scientifically interesting: \phil{given a suitable modelling tool, and
+appropriate guidance, a small team of non-professional volunteers was
+able to model a sample of 29 test lenses, and measure their Einstein
+radii with comparable accuracy to a professional expert.}
+
+There is, however, plenty of room for improvement:
 \begin{enumerate}
+\item \spl tends to overestimate the Einstein radius (evident from
+  \Figref{fig:ER_per_sim}), and it is also apparent that the model
+  density profiles tend to be too shallow.  The likely explanation is
+  that while the sims are steeply peaked at the centre, the pixellated
+  mass model fixes a comparatively large area near the central at
+  constant density.  One way to solve this problem would be to
+  introduce smaller pixels in the central region, thus enabling a
+  steeper centre. \phil{The use of simply parameterised, appropriately
+  steeply-profiled models would also avoid this problem.}
 \item Currently, \spl does not attempt to model the source shape; the
@@
   to be implemented.  This would alleviate another problem with \spl,
-  which is that there is no numerical figure of merit, and assessment
-  of a model is a judgment call based on the synthetic image, and on
-  whether the mass distribution and the arrival-time surface show
-  suspect features.
-\item \spl tends to overestimate the Einstein radius (evident from
-  Figure \ref{fig:ER_per_sim}), and it is also apparent that the
-  models tend to be too shallow.  The likely explanation is that while
-  the sims are steeply peaked at the centre, the pixellated mass model
-  fixes a comparatively large area near the central at constant
-  density.  One way to solve this problem would be to introduce
-  smaller pixels in the central region, thus enabled a steeper centre.
+  which is that there is as yet no quantitative figure of merit for 
+  any given model: assessment of each model is a judgment call based
+  on the synthetic image, and on whether the mass distribution and the
+  arrival-time surface show suspicious features.
 \item Another limitation so far in \spl is that the lens is assumed to
@@
   collaborative modeling effort, a particularly interesting candidate
-  can lead to an extended discussion and dozens of models.
+  can lead to an extended discussion and dozens of models, \phil{that
+  in some way sample the high likelihood region of model parameter
+  space}.
 \item Better tutorial materials are also needed, and this would
@@
 Swiss Academy of Sciences. AV is supported by a research fellowship
-from the Leverhulme Trust.
+from the Leverhulme Trust. This work was supported in part by the U.S.
+Department of Energy under contract number DE-AC02-76SF00515.
 

### 2014-07-29 text/ms.tex
editing

@@
 \newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}
+\let\phil\relax
 
@@
 
-One begins by going to the \spl \phil{web application\footnote{\bf Phil
-asks: should we provide a spl url here?}} and entering the number of a
+One begins by going to the \spl web application\footnote{\tt
+http://mite.physik.uzh.ch} and entering the number of a
 \sw image tile.  \spl then presents the image, along with zoom and pan
@@
 As implemented so far, \spl assumes that the lens is dominated by a
-single galaxy.  Only one maximum \phil{in the lens mass distrbution}
+single galaxy.  Only one maximum in the arrival-time surface
 is permitted, and it is taken to be the centre of the main lensing
 galaxy.  The user can, however, mark additional minor galaxies: these
-are modelled as point masses.  {\bf Phil asks: with what mass? Are
-these masses allowed to vary freely during the fit?}
+are modelled as point masses, the mass being fitted by the program along
+with the rest of the mass distribution.
 
@@
 \end{matrix} $$
-The marked-up image is a zoom of the lensed image on \sw,
-marked up with one or more spaghetti contours; this is the modeller's
-input to \spl.  The three panels on the right show the graphical
-output returned to the modeller by \spl, listed as (i), (ii), (iii) in
-\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
-200 models generated by \spl. The model synthetic image is too crude
-to be useful, because there is no fitting for the source light
-distribution, as already noted in \secref{diag}.  The quality of the
-synthetic images can be improved by interpolation, but as this feature
-was not available during the modelling challenge, we do not show the
-improved versions here.
 
@@
 \FloatBarrier
+
+\noindent The marked-up image is a zoom of the lensed image on \sw,
+marked up with one or more spaghetti contours; this is the modeller's
+input to \spl.  The three panels on the right show the graphical
+output returned to the modeller by \spl, listed as (i), (ii), (iii) in
+\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
+200 models generated by \spl. The model synthetic image is too crude
+to be useful, because there is no fitting for the source light
+distribution, as already noted in \secref{diag}.  The quality of the
+synthetic images can be improved by interpolation, but as this feature
+was not available during the modelling challenge, we do not show the
+improved versions here.
 
@@
 \item \spl tends to overestimate the Einstein radius (evident from
-  \Figref{fig:ER_per_sim}), and it is also apparent that the model
+  \Figref{ER_per_sim}), and it is also apparent that the model
   density profiles tend to be too shallow.  The likely explanation is

### 2014-07-29 text/ms.tex
filled in some references

@@
 \newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}
-\let\phil\relax
 
@@
 uses an existing free-form modelling program (GLASS), but
-\phil{enables the input to this code to be provided in a novel way,}
+enables the input to this code to be provided in a novel way,
 via a user-provided diagram that is essentially a sketch of an
@@
 
-An implementation \phil{of this method, \spl} is tested in a modelling
-challenge using 29 simulated lenses \phil{drawn from the} set created
-for \phil{the \sw citizen-science lens search}.  We find that
-\phil{volunteers from this online community} asserted the image
+An implementation of this method, \spl is tested in a modelling
+challenge using 29 simulated lenses drawn from the set created
+for the \sw citizen-science lens search.  We find that
+volunteers from this online community asserted the image
 parities and time ordering consistently in some lenses, but made
-errors in other lenses depending on the image morphology. \phil{While}
+errors in other lenses depending on the image morphology. While
 errors in image parity and time ordering lead to  large errors in the
-mass distribution, the enclosed mass \phil{was found to be more
+mass distribution, the enclosed mass was found to be more
 robust: the model-derived Einstein radii found by the volunteers were
@@
 modelling is a data analysis activity that can be crowd-sourced to
-good effect.} Shortcomings \phil{of the software implementation 
-include} (a)~lens profiles that tend to be systematically shallower
-than the lenses \phil{(leading to systematic over-estimation of the
-Einstein radius  by some $\sim25\%$}, and (b)~lack of detailed
-\phil{arc} modelling.  Ideas for improvement are discussed.
+good effect. Shortcomings of the software implementation 
+include (a)~lens profiles that tend to be systematically shallower
+than the lenses (leading to systematic over-estimation of the
+Einstein radius  by some $\sim25\%$, and (b)~lack of detailed
+arc modelling.  Ideas for improvement are discussed.
 
@@
 Imaging surveys now under way aim to increase the inventory of lenses
-another ten or a hundred fold \phil{\citep[see
+another ten or a hundred fold \citep[see
 e.g.][]{Marshall2005,OguriMarshall2010}, with both automated and
@@
 are presented sky-survey images and invited to identify lens
-candidates, by eye.}  Simulated lenses are mixed in with the data,
+candidates, by eye.  Simulated lenses are mixed in with the data,
 both to help train volunteers on what to look for, and to provide
@@
 identified lens candidates on earlier citizen-science surveys,
-\phil{either} to make discoveries missed in automatic searches by
-software robots, \phil{or to perform the necessary inspection of an
-automatically-generated sample, for quality control}.  Robots can be
+either to make discoveries missed in automatic searches by
+software robots, or to perform the necessary inspection of an
+automatically-generated sample, for quality control.  Robots can be
 built to be good at detecting clean lensing system in uncrowded fields
@@
 
-\phil{The encouraging early results from the first \sw lens search,
+The encouraging early results from the first \sw lens search,
 carried out on the  $\simeq172$ square degree Canada-France-Hawaii
 Telescope Legacy Survey (CFHTLS) imaging by over 30,000 volunteers
-(Marshall et al, in prep, More et al, in prep) prompted the question:}
+(Marshall et al, in prep, More et al, in prep) prompted the question:
 could the modelling of the lenses also be done by the volunteers?  If
@@
 In \secref{SpaghettiLens} we describe the \spl program, which
-implements \phil{the above scheme}.  \spl is an \phil{interface to and
-extension of} the GLASS framework for modelling lenses
+implements the above scheme.  \spl is an interface to and
+extension of the GLASS framework for modelling lenses
 \citep{2014arXiv1401.7990C}.  We will not go into software details in
@@
 remark that \spl is designed to be friendly to the forum style of
-citizen-science projects, and enable incremental \phil{collaborative}
+citizen-science projects, and enable incremental collaborative
 model refinement by different people, without sacrificing any of the
@@
 diverse sample of 29 simulated lenses was modelled multiple times
-\phil{by a small number of \sw volunteers} using \spl.  The models
+by a small number of \sw volunteers using \spl.  The models
 were then examined in two ways.  One was whether the spaghetti diagram
@@
 \spl is not part of \sw, but it makes use of the \sw infrastructure,
-in particular, the image database and the \phil{discussion
-forum.\footnote{\texttt{http://talk.spacewarps.org}}  The forum is
+in particular, the image database and the discussion
+forum.\footnote{\tt http://talk.spacewarps.org}  The forum is
 essential for establishing contact between interested members of the
@@
 summarized on modelling threads on the forum, and anyone interested
-was made welcome to join at any time.} 
-%
+was made welcome to join at any time.
+% {\bf Phil asks: how much training was necessary before the volunteers
+% understood how to work the interface? Could this training be done with
+% no input from the \spl developers, in future? It would be good to
+% comment on how scalable this approach to modeling is.}
+
 % PJM: The following paragraph is not accurate: the manipulation of 
@@
 %
+
 Modelling with \spl involves three stages, markup of the image,
-followed by intensive \phil{numerical computation carried out on a
-server in the background,} followed by review
+followed by intensive numerical computation carried out on a
+server in the background, followed by review
 of diagnostics and possible discussion.  We now describe each of
@@
 The markup tool allows only valid lensing configurations to be
-entered.  The user does not need to think explicitly the image
+entered.  The user does not need to think explicitly about the image
 parities (though the markup tool provides this information using
@@
 
-{\bf Phil asks: how much training was necessary before the volunteers
-understood how to work the interface? Could this training be done with
-no input from the \spl developers, in future? It would be good to
-comment on how scalable this approach to modeling is.}
-
 As implemented so far, \spl assumes that the lens is dominated by a
-single galaxy.  Only one maximum in the arrival-time surface
+single galaxy.  Accordingly, only one maximum in the arrival-time surface
 is permitted, and it is taken to be the centre of the main lensing
@@
   larger.)
-\item There is a notional lens center, say $(x_0,y_0)$ which is
+\item There is a notional lens centre, say $(x_0,y_0)$ which is
   identified with the maximum of the arrival time.  The source can
-  have an arbitrary offset with respect to the lens center.
+  have an arbitrary offset with respect to the lens centre.
 \item The mass distribution must be centrally concentrated, in two
@@
 The synthetic image generated by \spl assumes a simple
-\phil{circularly symmetric} source \phil{with linearly decreasing
-surface brightness profile}.  The user can change the contrast level
+circularly symmetric source with linearly decreasing
+surface brightness profile.  The user can change the contrast level
 on the image, which (though it is not saved) amounts to adjusting the
-\phil{size of the source}. These synthetic images are still very
-crude, and \phil{not always} useful for assessing models.  The best
+size of the source. These synthetic images are still very
+crude, and not always useful for assessing models.  The best
 indicator, in practice, of whether the modelling was successful is
@@
 
-After examining this feedback, the user can choose to \phil{save the
+After examining this feedback, the user can choose to save the
 model to the \spl archive, at which point it is assigned a unique
-URL}.  They can also modify their input and try again, or discard the
+URL.  They can also modify their input and try again, or discard the
 attempt altogether.  After archiving, there can be discussion among
 modellers, through the \sw forum or by any other means, and revision
-of the model.  \phil{This is achieved simply by sharing the model's
+of the model.  This is achieved simply by sharing the model's
 URL; following its hyperlink takes one to the \spl app, pre-loaded
-with the correct data image and input spaghetti.} Any archived model
+with the correct data image and input spaghetti. Any archived model
 can be revised by any user: they can modify the spaghetti
@@
 \end{matrix} $$
+The marked-up image is a zoom of the lensed image on \sw,
+marked up with one or more spaghetti contours; this is the modeller's
+input to \spl.  The three panels on the right show the graphical
+output returned to the modeller by \spl, listed as (i), (ii), (iii) in
+\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
+200 models generated by \spl. The model synthetic image is too crude
+to be useful, because there is no fitting for the source light
+distribution, as already noted in \secref{diag}.  The quality of the
+synthetic images can be improved by interpolation, but as this feature
+was not available during the modelling challenge, we do not show the
+improved versions here.
 
@@
 \FloatBarrier
-
-\noindent The marked-up image is a zoom of the lensed image on \sw,
-marked up with one or more spaghetti contours; this is the modeller's
-input to \spl.  The three panels on the right show the graphical
-output returned to the modeller by \spl, listed as (i), (ii), (iii) in
-\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
-200 models generated by \spl. The model synthetic image is too crude
-to be useful, because there is no fitting for the source light
-distribution, as already noted in \secref{diag}.  The quality of the
-synthetic images can be improved by interpolation, but as this feature
-was not available during the modelling challenge, we do not show the
-improved versions here.
 
@@
 \citep[see, e.g.,][]{2002LNP...608....1C} --- so \ERf[,rec] being too
-high is really a consequence of the \phil{GLASS} models being too shallow for
+high is really a consequence of the GLASS models being too shallow for
 the sims.
@@
 \Figref{ER_per_sim} shows that \ERf[,rec] of the models tend to be too
-high.  \phil{However, this is entirely due to the GLASS model density
+high.  However, this is entirely due to the GLASS model density
 profiles being too shallow, as illustrated above. We can separate out
@@
 The expert models show a statistically indistinguishable bias of 20\%,
-with standard deviation 8\% (the blue band in \Figref{ER_per_sim}).  }
-One source of this systematic error is that it is difficult to center
+with standard deviation 8\% (the blue band in \Figref{ER_per_sim}).
+One source of this systematic error is that it is difficult to centre
 of the lens accurately: an offset leads to a flatter mass profile for
@@
 paper indicate that such modelling would be both feasible and
-scientifically interesting: \phil{given a suitable modelling tool, and
+scientifically interesting: given a suitable modelling tool, and
 appropriate guidance, a small team of non-professional volunteers was
 able to model a sample of 29 test lenses, and measure their Einstein
-radii with comparable accuracy to a professional expert.}
+radii with comparable accuracy to a professional expert.
 
@@
   introduce smaller pixels in the central region, thus enabling a
-  steeper centre. \phil{The use of simply parameterised, appropriately
-  steeply-profiled models would also avoid this problem.}
+  steeper centre. The use of simply parameterised, appropriately
+  steeply-profiled models would also avoid this problem.
 \item Currently, \spl does not attempt to model the source shape; the
@@
   collaborative modeling effort, a particularly interesting candidate
-  can lead to an extended discussion and dozens of models, \phil{that
+  can lead to an extended discussion and dozens of models, that
   in some way sample the high likelihood region of model parameter
-  space}.
+  space.
 \item Better tutorial materials are also needed, and this would

### 2014-07-31 text/ms.tex
Added screenshot. Fixes #17

--- 2014-07-27 text/ms.tex
@@
 
-One begins by going to the \spl \phil{web application\footnote{\bf Phil
-asks: should we provide a spl url here?}} and entering the number of a
-\sw image tile.  \spl then presents the image, along with zoom and pan
-options and a markup tool to construct a spaghetti diagram.  The human
-modeller now has to make an educated guess for the topography of the
-arrival-time surface, and input the locations time-ordering of the
-maxima, minima, and saddle-points.  The markup tool \citep[which is
-inspired by Figure~6 of][and is like that figure made interactive and
-overlaid on data]{1986ApJ...310..568B} lets the modeller enter the
-information by sketching saddle-point contours.  Examples can be seen
-in the upper-left panels of Figures \ref{fig:6941} to \ref{fig:7022}. 
-The loops in the markup tool were the origin of the ``spaghetti''
-metaphor.
+One begins by going to the \spl \phil{web application\footnote{\bf
+Phil asks: should we provide a spl url here?}} and entering the number
+of a \sw image tile.  \spl then presents the image, along with zoom
+and pan options and a markup tool to construct a spaghetti diagram. 
+The human modeller now has to make an educated guess for the
+topography of the arrival-time surface, and input the locations
+time-ordering of the maxima, minima, and saddle-points.  The markup
+tool \citep[which is inspired by Figure~6 of][and is like that figure
+made interactive and overlaid on data]{1986ApJ...310..568B} lets the
+modeller enter the information by sketching saddle-point contours. 
+Examples can be seen in Figure~\ref{fig:screenshot} and the upper-left
+panels of Figures \ref{fig:6941} to \ref{fig:7022}.  The loops in the
+markup tool were the origin of the ``spaghetti'' metaphor.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.90\linewidth]{fig/screenshot}
+  \caption{\phil{Screen grab of \spl in action. A \sw image has been loaded
+  in, re-centred and zoomed, and 5 images and associated time delay
+  surface `spaghetti'' contours suggested, using the marking tools
+  associated with the buttons along the top of the panel. The mass
+  model is optimized server-side 
+  when the right-most button is pressed.}}
+  \label{fig:screenshot}
+\end{figure}
 

### 2014-07-31 text/ms.tex

--- 2014-07-29 text/ms.tex
@@
 overlaid on data]{1986ApJ...310..568B} lets the modeller enter the
-information by sketching saddle-point contours.  Examples can be seen
-in the upper-left panels of Figures \ref{fig:6941} to \ref{fig:7022}. 
-The loops in the markup tool were the origin of the ``spaghetti''
-metaphor.
+information by sketching saddle-point contours.  Examples can be seen in Figure~\ref{fig:screenshot} and the upper-left
+panels of Figures \ref{fig:6941} to \ref{fig:7022}.  The loops in the
+markup tool were the origin of the ``spaghetti'' metaphor.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=0.90\linewidth]{fig/screenshot}
+  \caption{\phil{Screen grab of \spl in action. A \sw image has been loaded
+  in, re-centred and zoomed, and 5 images and associated time delay
+  surface `spaghetti'' contours suggested, using the marking tools
+  associated with the buttons along the top of the panel. The mass
+  model is optimized server-side 
+  when the right-most button is pressed.}}
+  \label{fig:screenshot}
+\end{figure}
 

### 2014-08-04 plots/adv_inp/grab_svg.js
scaled input images

@@ -1,0 +1,129 @@
+/*
+ * Some function schnipplets that extract the working canvas
+ *
+ * intended to be run directly in the browser,
+ * using the debug console
+ */
+
+
+//extracts svg
+(function(){
+    $("svg").attr({xmlns:"http://www.w3.org/2000/svg"});
+    var b64 = btoa($("svg").parent().html());
+    var link = "<a href-lang='image/svg+xml' href='data:image/svg+xml;base64,\n"+b64+"' download='file.svg'>KLICK</a>";
+    $("<div>"+link+"</div>").dialog();
+})();
+
+
+
+    
+mids = "
+6915
+6919
+6937
+6941
+6975
+6990
+7022
+7025    
+"    
+    
+    
+//extracts png 
+(function(){
+
+    console.log("in fn");
+
+    var $parent = $('<div style="width:2048px;height:2048px;position:absolute;top:0;left:0;zindex:0;"></div>');
+
+    var canvas = document.createElement('canvas');
+    canvas.width = 2048;
+    canvas.height = 2048;
+    canvas.style.width  = '2048px';
+    canvas.style.height = '2048px';
+
+    var f = 2; //scale factor for line widths ect..
+    
+    //$(".contourpath").css({stroke:"blue"})
+    
+    $(".contourpath").each(function(){
+        var $t = $(this);
+        var sw = parseFloat($t.attr("stroke-width"));
+        $t.attr({"stroke-width": String(sw*f)});
+    });
+
+  
+    // first add css to the svg!!
+    // make a clone and add css to the clone
+    //  http://code.google.com/p/canvg/issues/detail?id=143
+    var clonedSVG = svg.root.cloneNode(true);
+
+    var style = document.createElementNS(svg.ns, "style");
+
+    //style.textContent += "<![CDATA[\n";
+  
+    //get stylesheet for svg
+    for (var i=0;i<document.styleSheets.length; i++) {
+    str = document.styleSheets[i].href;
+    if (str.indexOf('lmt.v')>-1 || str.indexOf('lmt.90.')>-1){
+        var rules = document.styleSheets[i].cssRules;
+        for (var j=0; j<rules.length;j++){
+            style.textContent += (rules[j].cssText + "\n");
+            }
+        break;
+        }
+    }
+  
+    //style.textContent += "]]>";
+
+    clonedSVG.getElementsByTagName("defs")[0].appendChild(style);
[…]

### 2014-08-04 plots/defaults.py

@@ -399,6 +399,12 @@ args.eR4.update({
 })
 
 
+
+#args.inp = AttrDict({
+#    'dim' : tmpl.squares.figure['figsize'][0] * tmpl.squares.figure['dpi']
+#})
+
+
 def set_mpl():
     #mpl.rcParams.update(args.mplrc)
     for k, v in args.mplrc.items():

### 2014-08-05 plots/defaults.py
bigger font on kappa_encl

@@ -249,6 +249,9 @@ tmpl.squares.contour = AttrDict({
 # the others have default values
 tmpl.rects = copyAD(args.default)
 
+tmpl.rects.figure.update({
+    'figsize'   : (6,4.5),
+})
 
 # Formatter Functions
 
@@ -335,6 +338,8 @@ args.kappaenc.update({
 
     'rEmod'     : AttrDict({'color': 'blue' , 'linestyle':'--' }),
     'rEsim'     : AttrDict({'color': 'red' ,       'linestyle':'--' }),
+
+
 })
 
 

### 2014-08-05 plots/defaults.py
updated final fig
changed to bw, added legend, incl contrast, incl markersize
fixes *16

@@ -250,7 +250,7 @@ tmpl.squares.contour = AttrDict({
 tmpl.rects = copyAD(args.default)
 
 tmpl.rects.figure.update({
-    'figsize'   : (6,4.5),
+    #'figsize'   : (6,4.5),
 })
 
 # Formatter Functions
@@ -323,6 +323,10 @@ args.arriv.minorc.update({
 
 # kappa enclosed for models
 args.kappaenc = copyAD(tmpl.rects)
+
+args.kappaenc.figure.update({
+    'figsize'   : (6,4.5),
+})
 args.kappaenc.addsub.update({
     'yscale'    : 'log',
 })
@@ -350,35 +354,39 @@ args.eR4.addsub.update({
     'yscale'    : 'log',
 })
 
-markersize = 6
+markersize = 7
 
 args.eR4.update({
     'expert'        : AttrDict({
-        'color'     :'blue',
+        'color'     :'black',
         'marker'    :'o',
-        'markerfacecolor': 'blue',
-        'markersize': markersize
+        'markerfacecolor': 'black',
+        'markersize': markersize,
+        'linestyle' : 'none',
         }),
 
     'rejected'      : AttrDict({
-        'color'     :'green',
+        'color'     :'black',
         'marker'    :'+',
 #        'markerfacecolor': 'green',
-        'markersize': markersize
+        'markersize': markersize*0.75,
+        'linestyle' : 'none',
         }),
 
     'regular'       : AttrDict({
         'color'     :'black',
         'marker'    :'o',
         'markerfacecolor': 'white',
-        'markersize': markersize
+        'markersize': markersize,
+        'linestyle' : 'none',
         }),
 
     'failed'        : AttrDict({
         'color'     :'black',
         'marker'    :'x',
 #        'markerfacecolor': 'white',
-        'markersize': markersize
+        'markersize': markersize,
+        'linestyle' : 'none',
         }),
 
     'unity'     : AttrDict({
@@ -387,10 +395,10 @@ args.eR4.update({
         }),
 
     'expertband'     : AttrDict({
-        'color': 'blue' ,
+        'color': 'black' ,
         'linestyle': 'solid',
         'linewidth' : 0,
-        'alpha': 0.5,
+        'alpha': 0.45,
         #'hatch': '/'
         }),
 
@@ -398,7 +406,14 @@ args.eR4.update({
         'color': 'black' ,
         'linestyle': 'solid',
         'linewidth' : 0,
[…]

### 2014-08-05 plots/gen_plots.py

@@ -1678,6 +1678,8 @@ def plotAllRE():
     draw_later2 = []
     flag = ''    
     new_data = []
+    leg_f = [None, None, None, None] # used for legend
+    leg_k = [None, None, None, None] # used for legend
     for i, dat in enumerate(spg.data):
       simn = dat['name']    
       #skip = False
@@ -1690,20 +1692,26 @@ def plotAllRE():
         style = kw.eR4.expert
         #skip=True
         draw_later.append([sims[simn], dat['rE_mean']])
-        flag = "expert"
+        flag = "modelled by expert"
+        nn = 1
       elif not tab.all_mods[str(dat['id'])]['acc']:
         style = kw.eR4.rejected
-        flag = "rejected"
+        flag = "rejected model"
+        nn = 3
       elif mod_err[dat['id']]['anyerr']:
         style = kw.eR4.failed
         draw_later2.append([sims[simn], dat['rE_mean']])
-        flag = "imgrecon wrong"
+        flag = "model with wrong ordering"
+        nn = 2
       else:
         style = kw.eR4.regular
-        flag = "good"
+        flag = "model with correct ordering"
+        nn = 0
       #xofs = 0.15 if dat['user']=='psaha' else -0.15
       #ax.plot(lu[simn]+xofs, dat['rE_mean']/sims[simn], style)
-      ax.plot(sims[simn], dat['rE_mean'], **style)
+      ap = ax.plot(sims[simn], dat['rE_mean'], **style)
+      leg_f[nn] = flag
+      leg_k[nn] = ap[0]
       new_data.append((sims[simn], dat['rE_mean'], flag))
     
     with open("new_data.csv", "w") as f:
@@ -1760,6 +1768,9 @@ def plotAllRE():
     plt.ylim([1, 40])
     plt.xlabel(r'actual Einstein Radius $\log(\Theta_{\text{E, act}})$')
     plt.ylabel(r'recovered Einstein Radius $\log(\Theta_{\text{E, rec}})$')
+    
+    
+    plt.legend(leg_k,leg_f, **kw.eR4.legend)
 
     
     for ext in exts:

### 2014-08-05 text/ms.tex
table to 3 col layout

@@
 
-\begin{table}\centering\begin{tabular}{lr}\hline
-    Total & 119 (100\%)\\
+\begin{table}\centering\begin{tabular}{lrr}\hline
+    Total & 119 &(100\%)\\
 \hline
-    errors in image locations  & 9 (8\%)\\
-    errors in image parities or time ordering & 49 (41\%) \\
+    errors in image locations  & 9 &(8\%)\\
+    errors in image parities or time ordering & 49 &(41\%) \\
 \hline
-    inaccurate image placement over an arc & 21 (18\%) \\
-    identified two images of four & 5 (4\%) \\
-    identified two nearby images as one & 3 (3\%) \\
-    missed faint images & 1 (1\%) \\
-    proposed too many images & 1 (1\%) \\
-    modelled a three-image arc as one image & 4 (3\%) \\
-    modelled one image as a three-image arc & 5 (4\%) \\
-    swapped minimum and saddle in double  & 2 (2\%) \\
-    swapped minima and saddles in quad & 38 (32\%)\\
-    swapped early and late saddles in a quad & 7 (6\%) \\
+    inaccurate image placement over an arc & 21 &(18\%) \\
+    identified two images of four & 5 &(4\%) \\
+    identified two nearby images as one & 3 &(3\%) \\
+    missed faint images & 1 &(1\%) \\
+    proposed too many images & 1 &(1\%) \\
+    modelled a three-image arc as one image & 4 &(3\%) \\
+    modelled one image as a three-image arc & 5 &(4\%) \\
+    swapped minimum and saddle in double  & 2 &(2\%) \\
+    swapped minima and saddles in quad & 38 &(32\%)\\
+    swapped early and late saddles in a quad & 7 &(6\%) \\
 \hline

### 2014-08-05 text/ms.tex
figure placement

--- 2014-07-31 text/ms.tex
@@
 
-One begins by going to the \spl \phil{web application\footnote{\bf
-Phil asks: should we provide a spl url here?}} and entering the number
-of a \sw image tile.  \spl then presents the image, along with zoom
-and pan options and a markup tool to construct a spaghetti diagram. 
-The human modeller now has to make an educated guess for the
-topography of the arrival-time surface, and input the locations
-time-ordering of the maxima, minima, and saddle-points.  The markup
-tool \citep[which is inspired by Figure~6 of][and is like that figure
-made interactive and overlaid on data]{1986ApJ...310..568B} lets the
-modeller enter the information by sketching saddle-point contours. 
-Examples can be seen in Figure~\ref{fig:screenshot} and the upper-left
-panels of Figures \ref{fig:6941} to \ref{fig:7022}.  The loops in the
-markup tool were the origin of the ``spaghetti'' metaphor.
-
-\begin{figure}
-  \centering
-    \includegraphics[width=0.90\linewidth]{fig/screenshot}
-  \caption{\phil{Screen grab of \spl in action. A \sw image has been loaded
-  in, re-centred and zoomed, and 5 images and associated time delay
-  surface `spaghetti'' contours suggested, using the marking tools
-  associated with the buttons along the top of the panel. The mass
-  model is optimized server-side 
-  when the right-most button is pressed.}}
-  \label{fig:screenshot}
-\end{figure}
+One begins by going to the \spl web application\footnote{\tt
+http://mite.physik.uzh.ch} and entering the number of a
+\sw image tile.  \spl then presents the image, along with zoom and pan
+options and a markup tool to construct a spaghetti diagram.  The human
+modeller now has to make an educated guess for the topography of the
+arrival-time surface, and input the locations time-ordering of the
+maxima, minima, and saddle-points.  The markup tool \citep[which is
+inspired by Figure~6 of][and is like that figure made interactive and
+overlaid on data]{1986ApJ...310..568B} lets the modeller enter the
+information by sketching saddle-point contours.  Examples can be seen
+in the upper-left panels of Figures \ref{fig:6941} to \ref{fig:7022}. 
+The loops in the markup tool were the origin of the ``spaghetti''
+metaphor.
 

### 2014-08-05 text/ms.tex

--- 2014-08-05 text/ms.tex
@@
     \includegraphics[width=0.90\linewidth]{fig/screenshot}
-  \caption{\phil{Screen grab of \spl in action. A \sw image has been loaded
+  \caption{Screen grab of \spl in action. A \sw image has been loaded
     in, re-centred and zoomed, and 5 images and associated time delay
-  surface `spaghetti'' contours suggested, using the marking tools
+    surface ``spaghetti'' contours suggested, using the marking tools
     associated with the buttons along the top of the panel. The mass
-  model is optimized server-side 
-  when the right-most button is pressed.}}
+    model is optimized server-side when the right-most button is
+    pressed.}
   \label{fig:screenshot}
@@
 
-The information in this section was not revealed to the main developer
-of \spl (RK, who also chose the challenge set) or to the modellers
-(EB, CC, CM, JO, PS and JW) while modelling was in progress.  That is,
-the modellers had no advance knowledge of what kind of
-parameterisation had been used to make the sims.  After the modelling
-stage, AM released the details of the sims for post-modelling
-analysis.  Results from the latter now follow.
-
-% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
-
-\subsection{Some example models} \label{sec:example_models}
-
-Of the 119 models proposed, we now discuss eight examples in some
-detail.  Results from these are shown in Figures
-\ref{fig:6941}--\ref{fig:7022}.  The first four of these show the most
-common image morphologies, the other four explain some problem cases.
-
-Each of Figures \ref{fig:6941}--\ref{fig:7022} figures has the
-following layout.
-$$ \begin{matrix}
-\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
-t(x,y)                        &\hbox{model } t(x,y) \\
-\kappa(x,y)                   &\hbox{model } \kappa(x,y)
-\end{matrix} $$
-The marked-up image is a zoom of the lensed image on \sw,
-marked up with one or more spaghetti contours; this is the modeller's
-input to \spl.  The three panels on the right show the graphical
-output returned to the modeller by \spl, listed as (i), (ii), (iii) in
-\secref{SpaghettiLens}.  These derive from the mean of an ensemble of
-200 models generated by \spl. The model synthetic image is too crude
-to be useful, because there is no fitting for the source light
-distribution, as already noted in \secref{diag}.  The quality of the
-synthetic images can be improved by interpolation, but as this feature
-was not available during the modelling challenge, we do not show the
-improved versions here.
-
 \FloatBarrier
@@
 
+The information in this section was not revealed to the main developer
+of \spl (RK, who also chose the challenge set) or to the modellers
+(EB, CC, CM, JO, PS and JW) while modelling was in progress.  That is,
+the modellers had no advance knowledge of what kind of
+parameterisation had been used to make the sims.  After the modelling
+stage, AM released the details of the sims for post-modelling
+analysis.  Results from the latter now follow.
+
+% - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - - -
+
+\subsection{Some example models} \label{sec:example_models}
+
+Of the 119 models proposed, we now discuss eight examples in some
+detail.  Results from these are shown in Figures
+\ref{fig:6941}--\ref{fig:7022}.  The first four of these show the most
+common image morphologies, the other four explain some problem cases.
+
+Each of Figures \ref{fig:6941}--\ref{fig:7022} figures has the
+following layout.
+$$ \begin{matrix}
+\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
+t(x,y)                        &\hbox{model } t(x,y) \\
+\kappa(x,y)                   &\hbox{model } \kappa(x,y)
+\end{matrix} $$
+
 % !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
@@
 %
-In these figures, the scales and alignments in the six panels are not
-completely consistent: the three panels on the right have the same
-coordinates; the middle-left and lower-left panels are consistent with
-each other, but only approximately aligned with the right column,
-because the latter are affected by errors in mouse placements; the
-upper-left panel is zoomed out slightly, to accommodate the full
-markup.  For qualitative features, which is all that
-Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, the slightly
-varying scales and alignments are harmless.
+% In these figures, the scales and alignments in the six panels are not
+% completely consistent: the three panels on the right have the same
+% coordinates; the middle-left and lower-left panels are consistent with
+% each other, but only approximately aligned with the right column,
+% because the latter are affected by errors in mouse placements; the
+% upper-left panel is zoomed out slightly, to accommodate the full
+% markup.  For qualitative features, which is all that
+% Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, the slightly
+% varying scales and alignments are harmless.
 % !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!

### 2014-08-06 systems/h2m.py
whatever

@@ -1,4 +1,4 @@
-from numpy import linspace, amin, amax, cos, arctan2
+from numpy import zeros, linspace, amin, amax, sqrt, pi, cos, sin, arctan, arctanh, arctan2
 from pylab import figure, show
 from many import grids
 
@@ -44,12 +44,31 @@ def alpha_SIE(x,y,reinst,ell,ell_pa):
 
 def alpha_shear(x,y,sh_str,sh_pa):
     thg = sh_pa*pi/180 + pi/2
-    rsq = x*x + y*y
+    r = sqrt(x*x + y*y)
     f = arctan2(y,x) - 2*thg
     return (-sh_str*r*cos(f),sh_str*r*sin(f))
 
-def disp(asw,x,y):
+def far(asw,x,y):
+    dx = zeros((len(y),len(x)))
+    dy = 1*dx
     obj = sim[asw]
+    src = obj[1]
+    bx,by = float(src[1]), float(src[2])
+    gal = obj[2]
+    reinst,ell,ell_pa = float(gal[1]), float(gal[2]), float(gal[3])
+    print reinst,ell,ell_pa
+    xs = obj[3]
+    sh_str,sh_pa = float(xs[1]), float(xs[2])
+    for i in range(len(x)):
+        for j in range(len(y)):
+            phix,phiy = alpha_SIE(x[i],y[j],reinst,ell,ell_pa)
+            shx,shy = alpha_shear(x[i],y[j],sh_str,sh_pa)
+            r = sqrt(x[i]**2+y[j]**2)
+            dx[j,i] = bx - x[i] + reinst*x[i]/r
+            dy[j,i] = by - y[j] + reinst*y[j]/r
+    return dx*dx + dy*dy
+
+
 
 sim = readq()
 
@@ -57,12 +76,15 @@ print sim['ASW0000h2m']
 N,R = 100,20
 x = linspace(-R,R,N)
 y = 1*x
-kappa,arriv = grids('ASW0000h2m',x,y)
+# kappa,arriv = grids('ASW0000h2m',x,y)
+lum = far('ASW0000h2m',x,y)
 fig = figure()
 panel = fig.add_subplot(1,1,1)
 panel.set_aspect('equal')
-lo,hi = amin(arriv), amax(arriv)
-lev = linspace(lo,lo+.2*(hi-lo),100)
-panel.contour(x,y,arriv,lev)
+# lo,hi = amin(lum), amax(lum)
[…]

### 2014-08-06 systems/h2m.py
revisited h2m

@@ -40,6 +40,7 @@ def alpha_SIE(x,y,reinst,ell,ell_pa):
     B = sqrt((1-q*q)/(q*q*x*x +y*y))
     phix = A*arctan(B*x)
     phiy = A*arctanh(B*y)
+    phix,phiy = cs*phix - sn*phiy, sn*phix + cs*phiy
     return (phix,phiy)
 
 def alpha_shear(x,y,sh_str,sh_pa):
@@ -64,9 +65,9 @@ def far(asw,x,y):
             phix,phiy = alpha_SIE(x[i],y[j],reinst,ell,ell_pa)
             shx,shy = alpha_shear(x[i],y[j],sh_str,sh_pa)
             r = sqrt(x[i]**2+y[j]**2)
-            dx[j,i] = bx - x[i] + reinst*x[i]/r
-            dy[j,i] = by - y[j] + reinst*y[j]/r
-    return dx*dx + dy*dy
+            dx[j,i] = bx - x[i] + phix + shx
+            dy[j,i] = by - y[j] + phiy + shy
+    return dx,dy
 
 
 
@@ -76,15 +77,16 @@ print sim['ASW0000h2m']
 N,R = 100,20
 x = linspace(-R,R,N)
 y = 1*x
-# kappa,arriv = grids('ASW0000h2m',x,y)
+kappa,arriv = grids('ASW0000h2m',x,y)
 lum = far('ASW0000h2m',x,y)
 fig = figure()
 panel = fig.add_subplot(1,1,1)
 panel.set_aspect('equal')
-# lo,hi = amin(lum), amax(lum)
+lo,hi = amin(arriv), amax(arriv)
+lev = linspace(lo,lo+.2*(hi-lo),100)
+panel.contour(x,y,arriv,lev)
 lev = linspace(-0.01,0.01,10)
-panel.contour(x,y,lum,lev)
-#panel.contour(x,y,lum[0],lev)
-#panel.contour(x,y,lum[1],lev)
+panel.contour(x,y,lum[0],lev)
+panel.contour(x,y,lum[1],lev)
 show()
 

### 2014-08-06 text/ms.tex
whatever
final changes for final draft
expanded fig description
corrected issue with colorspace
other minor cleanups in code
added utf8 encoding for my name

@@
 % ====================================================================
+% NOTES
+%
+% * in case of the following warning with png files:
+%     warning: pdflatex> libpng warning: iCCP: known incorrect sRGB profile
+%   do:
+%     find . -type f -name "*.png" -exec convert {} -strip {} \;
+%
+% * in case of multiple page group error:
+%    find . -type f -name "*.pdf" -exec sed -i 's/\/Group.*R//g' {} \;
+%    (http://tex.stackexchange.com/a/154416/37041)
+%
+% ====================================================================
 
 \documentclass[usenatbib]{mn2e}
+
+\usepackage[utf8]{inputenc}  % for my lastname :) allows umlauts in the input file without hacks
+\usepackage[T1]{fontenc}     % and a font that supports umlauts for output pdf
+\usepackage[english]{babel}  % just to be clear. should be the default anyways
 
@@
 
-\author[Kueng et al]{Rafael Kueng,$^{1}$
+\author[Küng et al]{Rafael Küng,$^{1}$
 Prasenjit Saha,$^{1}$
@@
 \centering
-\includegraphics[width=\columnwidth]{fig/arriv_0}
-\includegraphics[width=\columnwidth]{fig/arriv_1}
-\includegraphics[width=\columnwidth]{fig/arriv_2}
+\includegraphics[width=0.95\columnwidth]{fig/arriv_0}
+\includegraphics[width=0.95\columnwidth]{fig/arriv_1}
+\includegraphics[width=0.95\columnwidth]{fig/arriv_2}
 \caption{Perspective views and contour maps of example arrival time
@@
     counter-image; here the arc is closer to the lensing galaxy than
-    the counter-image. (See \secref{example_models} for details.)}
+    the counter-image. (See \secref{example_models} for details.)
+    \vspace{1em}}
   \label{fig:6919}
@@
   \caption[result 6975 (ASW000195x)]{A lens with unrecovered mass
-    substructure. (See \secref{example_models} for details.)}
+    substructure. (See \secref{example_models} for details.)
+    \vspace{1em}
+    }
   \label{fig:6975}
@@
 
+The model synthetic image presented in this paper is not the original one,
+but an interpolated version generated by an updated version of \spl. The image
+presented during the original experiment was more rough. The two plots in
+the middle showing $t(x,y)$ have uniform, but arbitrary spaced contour lines.
+The $\kappa(x,y)$ plots in the bottom row show solid lines for $\kappa>1$
+and dashed lines for $\kappa<1$. The spacing is logarithmic, with 10 contours
+for every decade --- that means contour spacing is a quarter magnitude in 
+optical terms.
+
 Let us now consider these cases in turn.
@@
 \item Figure \ref{fig:6941} shows the simplest case, with two clear
-  images produced by a nearly-circular lens.  The centre of the
+  images produced by a nearly-circular lens.  The center of the
   lensing galaxy is a maximum, the image nearer to the galaxy is a
@@
   spaghetti diagrams is unimportant, only the implied image locations,
-  parities and time-ordering matters.
+  parities and time-ordering matters. $\kappa(x,y)$ shows, that
+  the model has a more shallow mass distribution than the simulation.
+  This is a persistent issue throughout all models and is discussed in
+  \secref{tests.t2}.
 
@@
 
+\item Figure \ref{fig:6937} shows a sim with substructre, like
+  Figure \ref{fig:6937}. In comparison to the above, the resulting mass
+  model is poor.
+
 \item Figure \ref{fig:7025} shows a quad.  In this one, the
@@
 
-\item Figure \ref{fig:7022} shows another model of the same system.
+\item Figure \ref{fig:7022} shows another model of the same system, the
+  only one done by an expert in this sample. The image parities are
+  correct. The elongation has the right orientation, but is too shallow.
 
@@
 that is, the effective Einstein radius \ERf is such that the
-mean $\kappa$ is unity inside a circle of radius \ERf centred at
-the lens centre.
+mean $\kappa$ is unity inside a circle of radius \ERf centered at
+the lens center.
 
@@
   \caption{Model-recovered versus actual Einstein radii \ERf[,rec] and
-    \ERf[,act].  Circles denote models where the image parities and
-    time-orderings were correct, with filled circles for models by an
-    expert.  Crosses mark models where the time-ordering was
-    incorrect.  Plus signs indicate models flagged by the modeller as
-    failures.}
+    \ERf[,act].
+%    Circles denote models where the image parities and
+%    time-orderings were correct, with filled circles for models by an
+%    expert.  Crosses mark models where the time-ordering was
+%    incorrect.
+    Plus signs indicate models flagged by the modeller as
+    failures by commenting negatively about it in the forum.}
   \label{fig:ER_per_sim}
@@
 poor, the mean Einstein radius overestimate was 25\%, with a 20\%
-standard deviation (shown by the grey band in \Figref{ER_per_sim}). 
+standard deviation (shown by the light grey band in \Figref{ER_per_sim}). 
 The expert models show a statistically indistinguishable bias of 20\%,
-with standard deviation 8\% (the blue band in \Figref{ER_per_sim}).
-One source of this systematic error is that it is difficult to centre
+with standard deviation 8\% (the dark grey band in \Figref{ER_per_sim}).
+One source of this systematic error is that it is difficult to center
 of the lens accurately: an offset leads to a flatter mass profile for
@@
   correct ordering of the points in nearly-symmetric configurations.
+  Better and more detailed introductory materials would also allow
+  the community of modellers to grow faster and without individual
+  instructions by experts or experienced volunteers.
 \end{enumerate}

### 2014-08-06 systems/h2m.py
slight changes

@@ -62,31 +62,43 @@ def far(asw,x,y):
     sh_str,sh_pa = float(xs[1]), float(xs[2])
     for i in range(len(x)):
         for j in range(len(y)):
-            phix,phiy = alpha_SIE(x[i],y[j],reinst,ell,ell_pa)
-            shx,shy = alpha_shear(x[i],y[j],sh_str,sh_pa)
+            phix,phiy = alpha_SIE(x[i],-y[j],reinst,ell,ell_pa)
+            shx,shy = alpha_shear(x[i],-y[j],sh_str,sh_pa)
             r = sqrt(x[i]**2+y[j]**2)
             dx[j,i] = bx - x[i] + phix + shx
-            dy[j,i] = by - y[j] + phiy + shy
+            dy[j,i] = by + y[j] + phiy + shy
     return dx,dy
 
 
 
 sim = readq()
-
 print sim['ASW0000h2m']
-N,R = 100,20
-x = linspace(-R,R,N)
-y = 1*x
-kappa,arriv = grids('ASW0000h2m',x,y)
-lum = far('ASW0000h2m',x,y)
+
 fig = figure()
 panel = fig.add_subplot(1,1,1)
 panel.set_aspect('equal')
-lo,hi = amin(arriv), amax(arriv)
-lev = linspace(lo,lo+.2*(hi-lo),100)
-panel.contour(x,y,arriv,lev)
+# kappa,arriv = grids('ASW0000h2m',x,y)
+# lo,hi = amin(arriv), amax(arriv)
+# lev = linspace(lo,lo+.2*(hi-lo),100)
+# panel.contour(y+R,x+R,arriv,lev)
+N,R = 100,20
+x = linspace(-R,R,N)
+y = 1*x
+lum = far('ASW0000h2m',x,y)
 lev = linspace(-0.01,0.01,10)
-panel.contour(x,y,lum[0],lev)
-panel.contour(x,y,lum[1],lev)
+x = 1.15*x-1.5
+y = 1.15*y+.5
+panel.contour(x+R,y+R,lum[0],lev)
+panel.contour(x+R,y+R,lum[1],lev)
+
+from pylab import imshow, plot, show
+from matplotlib.image import imread
+
+full = imread('h2m.png')
+xc = 59
+yc = 67
+R = 20
+panel.imshow(full[yc-R:yc+R+1,xc-R:xc+R+1])
+
 show()
 

### 2014-08-13 text/ms.tex
slight changes
corrected text on RE bias

@@
 \newauthor
-Christine McMillan,$^{2}$
+Christine Macmillan,$^{2}$
 Julianne Wilcox,$^{2}$
@@
 include (a)~lens profiles that tend to be systematically shallower
-than the lenses (leading to systematic over-estimation of the
-Einstein radius  by some $\sim25\%$, and (b)~lack of detailed
+than the lenses, leading to systematic over-estimation of the
+Einstein radius by $\sim10\%$, and (b)~lack of detailed
 arc modelling.  Ideas for improvement are discussed.
@@
 
-% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-% THIS PARAGRAPH TO BE REMOVED WHEN THE FIGURES ARE FIXED!
-%
-% In these figures, the scales and alignments in the six panels are not
-% completely consistent: the three panels on the right have the same
-% coordinates; the middle-left and lower-left panels are consistent with
-% each other, but only approximately aligned with the right column,
-% because the latter are affected by errors in mouse placements; the
-% upper-left panel is zoomed out slightly, to accommodate the full
-% markup.  For qualitative features, which is all that
-% Figures~\ref{fig:6941}--\ref{fig:7022} try to convey, the slightly
-% varying scales and alignments are harmless.
-% !!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!!
-
 The model synthetic image presented in this paper is not the original one,
@@
 the lens center.
+
+\begin{figure*}
+  \centering
+    \includegraphics[width=0.90\linewidth]{fig/eR_5}
+  \caption{Model-recovered versus actual Einstein radii \ERf[,rec] and
+    \ERf[,act].
+    Plus signs indicate models flagged by the modeller as
+    failures by commenting negatively about it in the forum.}
+  \label{fig:ER_per_sim}
+\end{figure*}
 
@@
 
-\begin{figure*}
-  \centering
-    \includegraphics[width=0.90\linewidth]{fig/eR_5}
-  \caption{Model-recovered versus actual Einstein radii \ERf[,rec] and
-    \ERf[,act].
-%    Circles denote models where the image parities and
-%    time-orderings were correct, with filled circles for models by an
-%    expert.  Crosses mark models where the time-ordering was
-%    incorrect.
-    Plus signs indicate models flagged by the modeller as
-    failures by commenting negatively about it in the forum.}
-  \label{fig:ER_per_sim}
-\end{figure*}
-
-% TODO: update figure, fix caption.
-
 \Figref{ER_per_sim} shows that \ERf[,rec] of the models tend to be too
@@
 (PS). Discounting the models which were flagged by the volunteers as
-poor, the mean Einstein radius overestimate was 25\%, with a 20\%
+poor, the mean Einstein radius overestimate was 10\%, with a 15\%
 standard deviation (shown by the light grey band in \Figref{ER_per_sim}). 
-The expert models show a statistically indistinguishable bias of 20\%,
-with standard deviation 8\% (the dark grey band in \Figref{ER_per_sim}).
+The expert models show a similar bias,
+with standard deviation 10\% (the dark grey band in \Figref{ER_per_sim}).
 One source of this systematic error is that it is difficult to center
@@
 
-The \spl program was developed by Kueng, with design suggestions from
+The \spl program was developed by Küng, with design suggestions from
 Coles, Cornen and Saha, and feedback from all co-authors.  The

### 2014-08-13 plots/restats.py

@@ -17,9 +17,9 @@ while True:
 
 ratio = map(lambda p: [float(p[1])/float(p[0]),p[2]],full)
 
-expert = [p[0] for p in ratio if p[1][:-1]=='modelled by expert']
-good = [p[0] for p in ratio if p[1][:-1]=='model with correct ordering']
-many = [p[0] for p in ratio if p[1][:-1]!='rejected model']
+expert = [p[0] for p in ratio if p[1][:-2]=='modelled by expert']
+good = [p[0] for p in ratio if p[1][:-2]=='model with correct ordering']
+many = [p[0] for p in ratio if p[1][:-2]!='rejected model']
 
 print len(good), len(ratio), len(expert)
 print mean(expert), std(expert)

### 2014-09-08 text/ms.tex
edits from Jonathan

@@
 modelling is a data analysis activity that can be crowd-sourced to
-good effect. Shortcomings of the software implementation 
-include (a)~lens profiles that tend to be systematically shallower
-than the lenses, leading to systematic over-estimation of the
-Einstein radius by $\sim10\%$, and (b)~lack of detailed
-arc modelling.  Ideas for improvement are discussed.
+good effect. Ideas for improvement are discussed; these include
+(a)~overcoming the tendency of the models to be shallower than the
+correct answer in test cases, leading to systematic over-estimation of
+the Einstein radius by 10\% at present, and (b)~detailed modelling of
+arcs.
 
@@
 \item \spl tends to overestimate the Einstein radius (evident from
-  \Figref{ER_per_sim}), and it is also apparent that the model
-  density profiles tend to be too shallow.  The likely explanation is
-  that while the sims are steeply peaked at the centre, the pixellated
-  mass model fixes a comparatively large area near the central at
-  constant density.  One way to solve this problem would be to
-  introduce smaller pixels in the central region, thus enabling a
-  steeper centre. The use of simply parameterised, appropriately
-  steeply-profiled models would also avoid this problem.
+  \Figref{ER_per_sim}), and it is also apparent that the model density
+  profiles tend to be too shallow.  The likely explanation is that
+  while the sims are steeply peaked at the centre, the pixellated mass
+  model fixes a comparatively large area near the central at constant
+  density. Allowing smaller pixels in the centre region, thus enabling
+  a steeper centre \citep[similar to the ``high resolution'' feature
+    implemented in][]{2014arXiv1401.7990C} may remove this bias.  The
+  use of simply parameterised, appropriately steeply-profiled models
+  would also avoid the problem.
 \item Currently, \spl does not attempt to model the source shape; the

### 2014-09-16 text/ms.tex
some edits
bla
Christine Edits

@@
 
-% todo annotations on side in red. easy to search for at the end
-% !!! REMOVE THIS LINES FOR FINAL COMPILE
+\usepackage{color}
+
+% % todo annotations on side in red. easy to search for at the end
+% % !!! REMOVE THIS LINES FOR FINAL COMPILE
+% %
+% %\usepackage{setspace}    % remove this pgk for final compile!!
+% %\usepackage{color}    % well... color...
 % 
-\usepackage{setspace}    % remove this pgk for final compile!! 
-\usepackage{color}    % well... color...
-
-%\newcommand{\todo}[2][red]{%
-%\textcolor{#1}{\textbullet}%
-%\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
-%\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}
-
-\newcommand{\needfig}[1][]{%
-\colorbox{yellow}{\parbox{0.5\textwidth}{%
-\vspace{0.5cm}\setstretch{0.5}\textcolor{black}{\scriptsize{fig #1}}\vspace{0.5cm}}}%
-\marginpar{\colorbox{yellow}{\parbox{\marginparwidth}{%
-\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{fig}}}}}}
-
-\setlength{\marginparsep}{0mm}
-\setlength{\marginparwidth}{2.2cm}
-
-\newcommand{\needcite}[1][]{\todo[green]{cit #1}}
-\newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
-\newcommand{\needchk}[1][]{\todo[red]{chk #1}}
-
-\definecolor{purple}{RGB}{150,0,200}
-\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}
+% %\newcommand{\todo}[2][red]{%
+% %\textcolor{#1}{\textbullet}%
+% %\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
+% %\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}
+% 
+% \newcommand{\needfig}[1][]{%
+% \colorbox{yellow}{\parbox{0.5\textwidth}{%
+% \vspace{0.5cm}\setstretch{0.5}\textcolor{black}{\scriptsize{fig #1}}\vspace{0.5cm}}}%
+% \marginpar{\colorbox{yellow}{\parbox{\marginparwidth}{%
+% \setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{fig}}}}}}
+% 
+% \setlength{\marginparsep}{0mm}
+% \setlength{\marginparwidth}{2.2cm}
+% 
+% \newcommand{\needcite}[1][]{\todo[green]{cit #1}}
+% \newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
+% \newcommand{\needchk}[1][]{\todo[red]{chk #1}}
+% 
+% \definecolor{purple}{RGB}{150,0,200}
+% \newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}
 
@@
 built to be good at detecting clean lensing system in uncrowded fields
-with high signal-to-noise, but in more general test situations robots
+with high signal-to-noise, but in more general test situations, robots
 miss lenses (low completeness) or contaminate the results with
@@
 surveys, which will yield thousands of lens candidates. There are
-several software tools for lensing modelling available, and work has
+several software tools for lens modelling available, and work has
 been done on generic interfaces \citep[e.g.][]{2014A&C.....5...28L}.
@@
 because they set the overall topography of the arrival-time surface.
-The obviously give the locations of the saddle points, and roughly
+They obviously give the locations of the saddle points, and roughly
 localise the minima and maxima as well.  If more precise locations of
@@
 but an interpolated version generated by an updated version of \spl. The image
-presented during the original experiment was more rough. The two plots in
+presented during the original experiment was of lower resolution. The two plots in
 the middle showing $t(x,y)$ have uniform, but arbitrary spaced contour lines.
@@
 simulations were created by A.~More, in consultation with Marshall,
-S.~More and Verma.  Modelling was done by Baeten, Cornen, McMillan,
+S.~More and Verma.  Modelling was done by Baeten, Cornen, Macmillan,
 Odermatt, Saha and Wilcox, with post-modelling analysis by Kueng and

### 2014-09-16 text/ms.tex
more

@@
 \newcommand{\spl}{SpaghettiLens\xspace}
-\newcommand{\sw}{Space Warps\xspace}
+\newcommand{\sw}{SpaceWarps\xspace}
 
@@
 a single background quasar into two (four), brought in their wake the
-first
-work on lens modelling
-\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.  For the second lens
+first work on lens modelling
+\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.
+
+%CM:
+%The first work on lens modelling (Young et al. 1981a,b) was developed 
+%after the discovery of the first two gravitational lenses (Walsh et al. 
+% 1979; Weymann et al. 1980), where a massive galaxy causes a background 
+% quasar to be presented as two or four images.
+
+For the second lens
 (PG1115+080), mass models scored an early success with the prediction
@@
 automatically-generated sample, for quality control.  Robots can be
-built to be good at detecting clean lensing system in uncrowded fields
+built to be good at detecting lensing system in clean, uncrowded fields
 with high signal-to-noise, but in more general test situations, robots
@@
 \end{equation}
-called the convergence.  And let there be light, in the form of a more
+called the convergence.  Let there be light, in the form of a more
 distant source, at redshift $z_S$, behind point $(x_s,y_s)$ on the

### 2014-09-16 text/ms.tex
more edits, modeLLing

@@
 \usepackage{subfigure}  % does what it says
+\usepackage{hyperref}
 
@@
 
-\title{Gravitational Lens Modeling in a Citizen Science Context}
+\title{Gravitational Lens Modelling in a Citizen Science Context}
 
@@
 (Marshall et al, in prep, More et al, in prep) is a citizen science
-project\footnote{\tt http://www.spacewarps.org} in which volunteers
+project\footnote{\url{http://www.spacewarps.org}} in which volunteers
 are presented sky-survey images and invited to identify lens
@@
 in particular, the image database and the discussion
-forum.\footnote{\tt http://talk.spacewarps.org}  The forum is
+forum.\footnote{\url{ http://talk.spacewarps.org}}  The forum is
 essential for establishing contact between interested members of the
@@
 
-One begins by going to the \spl web application\footnote{\tt
-http://mite.physik.uzh.ch} and entering the number of a
+One begins by going to the \spl web application\footnote{\url{
+http://mite.physik.uzh.ch}} and entering the number of a
 \sw image tile.  \spl then presents the image, along with zoom and pan
@@
 \footnote{See the ``collaborative modelling'' thread at
-{\tt http://talk.spacewarps.org} for an example.}
-% TODO: fix url!
+\url{ http://talk.spacewarps.org} for an example.}
 
@@
   different statistical measures.  As evidenced by a current
-  collaborative modeling effort, a particularly interesting candidate
+  collaborative modelling effort, a particularly interesting candidate
   can lead to an extended discussion and dozens of models, that
@@
 \item Better tutorial materials are also needed, and this would
-  address some of the problem areas found in the modeling challenge.
+  address some of the problem areas found in the modelling challenge.
   For example, we saw in \secref{tests.t1} that volunteers are most

### 2014-09-16 text/ms.tex
authors

@@
 \usepackage{xspace}     % fixes spaces after custom commands
-\usepackage{float}      % exact placement of figures (using \begin{figure}[H])
+%\usepackage{float}      % exact placement of figures (using \begin{figure}[H])
 \usepackage[section,below]{placeins}   % gives a \FloatBarrier to force all floats to appear
@@
   .png,.PNG,%
-  .jpg,.mps,.jpeg,.jbig2,.jb2,.JPG,.JPEG,.JBIG2,.JB2}
+  .jpg,.mps,.jpeg,.jbig2,.jb2,.JPG,.JPEG,.JBIG2,.JB2 %
+}
 
@@
 
-\def\half{{\textstyle\frac12}}
+%\def\half{{\textstyle\frac12}}
 % makes it easier to have consistent writing, and easy change if needed
@@
 \newcommand{\ERf}[1][]{$\Theta_\text{E#1}$\xspace} % full output
-
 %\newcommand{\ER}{Einstein radius\xspace} % text
@@
 %\newcommand{\ERm}[1][]{\Theta_\text{E#1}} % mathmode with greek symbol
-
-%shortcuts for kappa
 %\newcommand{\kenc}[1][r]{$\kappa_\text{encl}(#1)$\xspace}
@@
 \newcommand{\dgr}{^{\circ}}
-
-%more abbrevs
 \newcommand{\tgeom}{t_{\rm geom}}
@@
 
-% % todo annotations on side in red. easy to search for at the end
-% % !!! REMOVE THIS LINES FOR FINAL COMPILE
-% %
-% %\usepackage{setspace}    % remove this pgk for final compile!!
-% %\usepackage{color}    % well... color...
-% 
-% %\newcommand{\todo}[2][red]{%
-% %\textcolor{#1}{\textbullet}%
-% %\marginpar{\colorbox{#1}{\parbox{\marginparwidth}{%
-% %\setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{#2}}}}}}
-% 
-% \newcommand{\needfig}[1][]{%
-% \colorbox{yellow}{\parbox{0.5\textwidth}{%
-% \vspace{0.5cm}\setstretch{0.5}\textcolor{black}{\scriptsize{fig #1}}\vspace{0.5cm}}}%
-% \marginpar{\colorbox{yellow}{\parbox{\marginparwidth}{%
-% \setstretch{0.4}\sffamily\textcolor{black}{\scriptsize{fig}}}}}}
-% 
-% \setlength{\marginparsep}{0mm}
-% \setlength{\marginparwidth}{2.2cm}
-% 
-% \newcommand{\needcite}[1][]{\todo[green]{cit #1}}
-% \newcommand{\needref}[1][]{\todo[cyan]{ref #1}}
-% \newcommand{\needchk}[1][]{\todo[red]{chk #1}}
-% 
-% \definecolor{purple}{RGB}{150,0,200}
-% \newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}
-
-% \\\ END OF REMOVAL SECION
-
 \newlength{\myplotswidth}
@@
 Elisabeth Baeten,$^{2}$
+Jonathan Coles,$^{3}$
+\newauthor
 Claude Cornen,$^{2}$
+Christine Macmillan,$^{2}$
+Phil Marshall$^{4}$ 
+Anupreeta More,$^{5}$
 \newauthor
-Christine Macmillan,$^{2}$
-Julianne Wilcox,$^{2}$
-Jonas Odermatt,$^{3}$
-Jonathan Coles,$^{4}$
-\newauthor
-Aprajita Verma,$^{5}$
-Anupreeta More,$^{6}$
-Surhud More,$^{6}$
-and Phil Marshall$^{7}$ \\
+Surhud More,$^{5}$
+Jonas Odermatt,$^{6}$
+Aprajita Verma,$^{7}$
+and Julianne Wilcox,$^{2}$
+%
+\\
+%
 $^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
 $^{2}$The Space Warps Collaboration\\
-$^{3}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
-$^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
-$^{5}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
-$^{6}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
-$^{7}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
+$^{3}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
+$^{4}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
+$^{5}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
+$^{6}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
+$^{7}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
 }

### 2014-09-16 plots/defaults.py
updated eR4 plot

@@ -391,7 +391,9 @@ args.eR4.update({
 
     'unity'     : AttrDict({
         'color': 'black' ,
-        'linestyle':':'
+        'linestyle':'--',
+        'linewidth' : 2,
+       
         }),
 
     'expertband'     : AttrDict({

### 2014-09-16 plots/gen_plots.py

@@ -104,7 +104,7 @@ exts = kw.exts
 #figsizeKE = (10,8)
 
 # realtive to plots dir, incase of debug
-outdir = 'figs_aug8'
+outdir = 'figs_09-16'
 
 #select sims to produce plots for (using sel_sim_plots())
 sel_sim = [

### 2014-09-17 text/ms.tex
minor typos / grammar

@@
 Prasenjit Saha,$^{1}$
-Elisabeth Baeten,$^{2}$
-Jonathan Coles,$^{3}$
+Anupreeta More,$^{2}$
+Elisabeth Baeten,$^{3}$
 \newauthor
-Claude Cornen,$^{2}$
-Christine Macmillan,$^{2}$
-Phil Marshall$^{4}$ 
-Anupreeta More,$^{5}$
+Jonathan Coles,$^{4}$
+Claude Cornen,$^{3}$
+Christine Macmillan,$^{3}$
+Phil Marshall,$^{5}$ 
 \newauthor
-Surhud More,$^{5}$
+Surhud More,$^{2}$
 Jonas Odermatt,$^{6}$
-Aprajita Verma,$^{7}$
-and Julianne Wilcox,$^{2}$
+Aprajita Verma$^{7}$
+and Julianne Wilcox$^{3}$
 %
@@
 %
-$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
-$^{2}$The Space Warps Collaboration\\
-$^{3}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
-$^{4}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
-$^{5}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
+$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057 Zurich, Switzerland\\
+$^{2}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
+$^{3}$The SpaceWarps Collaboration\\
+$^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
+$^{5}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
 $^{6}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
@@
 We develop a method to enable collaborative modelling of gravitational
-lenses, and lens candidates, by non-professional lens enthusiasts.  It
+lenses and lens candidates by non-professional lens enthusiasts.  It
 uses an existing free-form modelling program (GLASS), but
@@
 
-An implementation of this method, \spl is tested in a modelling
+An implementation of this method, \spl, is tested in a modelling
 challenge using 29 simulated lenses drawn from the set created
@@
 models.  The reason for that lies in the image separation. Recall that
-image separations are of order the angular Einstein radius
+image separations are of order of the angular Einstein radius
 \begin{equation}
@@
 \simeq 0.1'' \left(\frac{M}{M_\odot}\right)^{1/2}
-             \left(\frac{D_L}{\rm pc}\right)^{-1/2}
+             \left(\frac{D_L}{\rm pc}\right)^{-1/2},
 \end{equation}
@@
 \citep[e.g.][]{Marshall2009,More2012ApJ,Gavazzi2014}. For example, \sw
-(Marshall et al, in prep, More et al, in prep) is a citizen science
+(Marshall et al, in prep; More et al, in prep) is a citizen science
 project\footnote{\url{http://www.spacewarps.org}} in which volunteers
@@
 Telescope Legacy Survey (CFHTLS) imaging by over 30,000 volunteers
-(Marshall et al, in prep, More et al, in prep) prompted the question:
+(Marshall et al, in prep; More et al, in prep) prompted the question:
 could the modelling of the lenses also be done by the volunteers?  If
@@
 \begin{equation} \label{eq:tgeom}
-\tgeom(x,y) \propto (x-x_s)^2 + (y-s_y)^2 \,.
+\tgeom(x,y) \propto (x-x_s)^2 + (y-s_y)^2 ,
 \end{equation}
@@
 and (b)~look very unlike galaxies.  Additional assumptions (a prior)
-are necessary.  GLASS uses the following prior
+are necessary.  GLASS uses the following priors
 \citep[cf.][]{1997MNRAS.292..148S,2008ApJ...679...17C}.
@@
 After examining this feedback, the user can choose to save the
-model to the \spl archive, at which point it is assigned a unique
-URL.  They can also modify their input and try again, or discard the
+model to the \spl archive, at which point it is assigned an unique
+URL.  He can also modify his input and try again, or discard the
 attempt altogether.  After archiving, there can be discussion among

### 2014-09-22 text/ms.tex
typos and imgs

@@
 Aprajita Verma$^{7}$
-and Julianne Wilcox$^{3}$
+and Julianne K. Wilcox$^{3}$
 %
@@
 by the modeller.  This tested first, for correct identification and
-location of the lensed images, and second, the correct parities and
-ordering for the lensed images in respect of the arrival time.
+location of the lensed images, and second, for the correct parities and
+ordering of the lensed images in respect of the arrival time.
 
@@
 arrival-time surface of the sim.  This amounts to comparing the
-upper-left and middle-left panels in each of
-Figures~\ref{fig:6941}--\ref{fig:7022}, and similarly the other 111
+middle-left and middle-right panels in each of
+Figures~\ref{fig:6941}--\ref{fig:7022}, and similarly for the other 111
 models.
@@
 \begin{table}\centering\begin{tabular}{lrr}\hline
-    Total & 119 &(100\%)\\
+    Total                                                     & 119        & 100\%        \\
 \hline
-    errors in image locations  & 9 &(8\%)\\
-    errors in image parities or time ordering & 49 &(41\%) \\
+    errors in image locations          &     9         &    8\%            \\
+    errors in image parities or time ordering     &    49     &    41\%     \\
 \hline
-    inaccurate image placement over an arc & 21 &(18\%) \\
-    identified two images of four & 5 &(4\%) \\
-    identified two nearby images as one & 3 &(3\%) \\
-    missed faint images & 1 &(1\%) \\
-    proposed too many images & 1 &(1\%) \\
-    modelled a three-image arc as one image & 4 &(3\%) \\
-    modelled one image as a three-image arc & 5 &(4\%) \\
-    swapped minimum and saddle in double  & 2 &(2\%) \\
-    swapped minima and saddles in quad & 38 &(32\%)\\
-    swapped early and late saddles in a quad & 7 &(6\%) \\
+    inaccurate image placement over an arc     &    21    &    18\%     \\
+    identified two images of four                         &     5    &     4\% \\
+    identified two nearby images as one & 3  & 3\%     \\
+    missed faint images &   1   &  1\%     \\
+    proposed too many images &  1  &  1\%    \\
+    modelled a three-image arc as one image  &  4  &   3\%   \\
+    modelled one image as a three-image arc  &  5  &   4\%   \\
+    swapped minimum and saddle in double       &  2  &    2\%   \\
+    swapped minima and saddles in quad            & 38 & 32\%   \\
+    swapped early and late saddles in a quad &   7 &   6\%    \\
 \hline
@@
 \figref{7025} shows an example.  Another, less common, error was
-flipping the spaghetti diagram, this swapping the time-ordering of the
+flipping the spaghetti diagram, thus swapping the time-ordering of the
 two saddle points.
@@
 Incorrect image parities and time orderings tended to produce
-poorer-looking models, such as the checkerboard patterns in the mass
+poorer-looking models, such as the checker board patterns in the mass
 map in \figref{7025}.  Interestingly, however, the enclosed-mass
@@
     Plus signs indicate models flagged by the modeller as
-    failures by commenting negatively about it in the forum.}
+    failures by commenting negatively about it in the forum. Light and dark grey
+    bands show standard deviation of volunteers (15\%) and expert (10\%).}
   \label{fig:ER_per_sim}

### 2014-09-22 plots/defaults.py

@@ -322,6 +322,8 @@ args.arriv.minorc.update({
 
 
 # kappa enclosed for models
+fs  = 16
+fs2 = fs*0.8
 args.kappaenc = copyAD(tmpl.rects)
 
 args.kappaenc.figure.update({
@@ -337,13 +339,13 @@ args.kappaenc.update({
     'min'       : AttrDict({'color': 'black' , 'linestyle':':' }),
     'max'       : AttrDict({'color': 'black'  , 'linestyle':':' }),
     'sad'       : AttrDict({'color': 'black', 'linestyle':':' }),
-    'text'      : AttrDict({'ha': 'left', 'va':'bottom' }),
+    'text'      : AttrDict({'ha': 'left', 'va':'bottom', 'size': fs }),
     'unity'     : AttrDict({'color': 'black' , 'linestyle':':' }),
 
     'rEmod'     : AttrDict({'color': 'blue' , 'linestyle':'--' }),
     'rEsim'     : AttrDict({'color': 'red' ,       'linestyle':'--' }),
-
-
+    'label'     : AttrDict({'fontsize': fs }),
+    'ticklabel' : AttrDict({'labelsize': fs2 }),
 })
 
 
@@ -355,6 +357,8 @@ args.eR4.addsub.update({
 })
 
 markersize = 7
+fs  = 12
+fs2 = fs*0.8
 
 args.eR4.update({
     'expert'        : AttrDict({
@@ -392,7 +396,7 @@ args.eR4.update({
     'unity'     : AttrDict({
         'color': 'black' ,
         'linestyle':'--',
-        'linewidth' : 2,
+        'linewidth' : 1,
        
         }),
 
@@ -415,9 +419,11 @@ args.eR4.update({
         'loc': 'lower right' ,
         'numpoints': 1,
         'frameon' : True,
-        'fontsize': 12,
+        'fontsize': fs2,
         }),
         
+    'label'     : AttrDict({'fontsize': fs }),
+    'ticklabel' : AttrDict({'labelsize': fs2 }),
 })
 
 

### 2014-09-22 plots/gen_plots.py

@@ -171,6 +171,7 @@ sys.path.append(os.path.join(pardir, 'systems'))
 
 import matplotlib.pyplot as plt
 from matplotlib.ticker import LogFormatter
+import matplotlib.ticker
 
 plt.ioff()
 
@@ -1386,6 +1387,7 @@ def draw_sim(asw, sim):
     #for f in filenames: print '  - %s'%(f+'.'+repr(exts))
 
 
+
 def draw_mod(mid, elem, data, sims):
 #def plot2(idd, count):
   plt.close("all")
@@ -1442,6 +1444,9 @@ def draw_mod(mid, elem, data, sims):
   
   yvals_extr = np.logspace(np.log10(3.5),np.log10(1),8)
   ypos_theta = np.logspace(np.log10(7),np.log10(5),4)
+
+  yvals_extr = np.logspace(np.log10(3),np.log10(0.7),8)
+  ypos_theta = np.logspace(np.log10(5),np.log10(3.5),4)
   
   # text offsets and properties
   t_dx = 0.0
@@ -1453,7 +1458,7 @@ def draw_mod(mid, elem, data, sims):
     
   #pl.ioff()
   fig = plt.figure(**kw.kappaenc.figure)
-  fig.add_subplot(1,1,1, **kw.kappaenc.addsub)
+  ax = fig.add_subplot(1,1,1, **kw.kappaenc.addsub)
   
   x  = elem['x']
   #y  = elem['y']
@@ -1490,7 +1495,7 @@ def draw_mod(mid, elem, data, sims):
     #pl.text(p['d']+t_dx, yvals_extr[jj], p['t'], **t_props)
     
     plt.plot([p['d'], p['d']], [0.01,yvals_extr[jj]], **kw.kappaenc[p['t']])
-    plt.text(p['d']+t_dx, yvals_extr[jj], p['t'], **kw.kappaenc.text)
+    plt.text(p['d']+t_dx, yvals_extr[jj], r'$\text{%s}$'%p['t'], **kw.kappaenc.text)
 
   # plot simulation parameter data
   plt.plot(sims[name]['x'], sims[name]['y'], **kw.kappaenc.sim)
@@ -1527,7 +1532,7 @@ def draw_mod(mid, elem, data, sims):
     
     #new placement due to logscale
     plt.plot(a_re_mean, [0.001,ypos_theta[2]], **kw.kappaenc.rEmod)
[…]

### 2014-09-22 text/ms.tex
edits

@@
 
-We develop a method to enable collaborative modelling of gravitational
-lenses and lens candidates by non-professional lens enthusiasts.  It
-uses an existing free-form modelling program (GLASS), but
-enables the input to this code to be provided in a novel way,
-via a user-provided diagram that is essentially a sketch of an
-arrival-time surface.
+We develop a method to enable collaborative modelling of
+gravitational lenses and lens candidates, that could be used by
+non-professional lens enthusiasts.  It uses an existing free-form
+modelling program (GLASS), but enables the input to this code to be
+provided in a novel way, via a user-provided diagram that is
+essentially a sketch of an arrival-time surface.
 
 An implementation of this method, \spl, is tested in a modelling
-challenge using 29 simulated lenses drawn from the set created
-for the \sw citizen-science lens search.  We find that
-volunteers from this online community asserted the image
-parities and time ordering consistently in some lenses, but made
-errors in other lenses depending on the image morphology. While
-errors in image parity and time ordering lead to  large errors in the
-mass distribution, the enclosed mass was found to be more
-robust: the model-derived Einstein radii found by the volunteers were
-consistent with those produced by one of the professional team,
-suggesting that given the appropriate tools, gravitational lens
-modelling is a data analysis activity that can be crowd-sourced to
-good effect. Ideas for improvement are discussed; these include
-(a)~overcoming the tendency of the models to be shallower than the
-correct answer in test cases, leading to systematic over-estimation of
-the Einstein radius by 10\% at present, and (b)~detailed modelling of
-arcs.
+challenge that used 29 simulated lenses that had been drawn from the
+set created for the \sw citizen-science lens search.  We find that
+volunteers from this online community asserted the image parities and
+time ordering consistently in some lenses, but made errors in other
+lenses depending on the image morphology. While errors in image parity
+and time ordering lead to large errors in the mass distribution, the
+enclosed mass was found to be more robust: the model-derived Einstein
+radii found by the volunteers were consistent with those produced by
+one of the professional team, suggesting that given the appropriate
+tools, gravitational lens modelling is a data analysis activity that
+can be crowd-sourced to good effect. Ideas for improvement are
+discussed; these include (a)~overcoming the tendency of the models to
+be shallower than the correct answer in test cases, leading to
+systematic over-estimation of the Einstein radius by 10\% at present,
+and (b)~detailed modelling of arcs.
 
@@
 The first two discoveries of gravitational lenses
-\citep{1979Natur.279..381W,1980Natur.285..641W}, where a galaxy images
+, where a galaxy images
 a single background quasar into two (four), brought in their wake the
 first work on lens modelling
-\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.
+.
 
 %CM:
-%The first work on lens modelling (Young et al. 1981a,b) was developed 
-%after the discovery of the first two gravitational lenses (Walsh et al. 
-% 1979; Weymann et al. 1980), where a massive galaxy causes a background 
-% quasar to be presented as two or four images.
+The first work on lens modelling
+\citep{1981ApJ...244..723Y,1981ApJ...244..736Y} was developed after
+the discovery of the first two gravitational lenses
+\citep{1979Natur.279..381W,1980Natur.285..641W}, where a massive
+galaxy causes a background
+quasar to be presented as two or four images.
 
@@
 Hence, the lensed images depend on the detailed mass distribution of
-the lensing galaxy.  Galaxy lenses demand modelling.
-
-Since the early discoveries, more than 400 secure lenses are now
+the lensing galaxy, and to uncover that dependence, galaxy lenses
+require models.
+
+Since those early discoveries, more than 400 secure lenses are now
 known. Modelling of the mass distribution is part of any research
@@
 lenses.  The largest single one \citep{2009ApJ...703L..51K} models 58
-lenses to infer the distribution of dark matter around galaxies.  In
-other work, \cite{2011ApJ...740...97L} combined lens models of 21
-galaxies with models of their stellar populations, to find the
-relation between stars and dark matter, and \cite{2014MNRAS.437..600S}
-modelled 18 time-delay lenses together to infer cosmological
-parameters.
+separate lenses to infer the distribution of dark matter around
+galaxies.  In other work, \cite{2011ApJ...740...97L} combined lens
+models of 21 galaxies with models of their stellar populations, to
+find the relation between stars and dark matter, and
+\cite{2014MNRAS.437..600S} modelled 18 time-delay lenses together to
+infer cosmological parameters.
 
@@
 coordinates at the lens, transverse to the line of sight.  Let
-$\Sigma(x,y)$ be the mass per unit area, i.e., density projected along
-the line of sight.  This surface density is often given in a
-dimensionless form
+$\Sigma(x,y)$ be the mass distribution.  It is a mass per unit area,
+i.e., density projected along the line of sight.  It is often given in
+a dimensionless form
 \begin{equation} \label{eq:kappa}
@@
 prune a model tree, focusing attention on the most interesting
-models.
+models.\footnote{See ``Collaborative gravitational lens
+modelling\dots'' in {\tt http://letters.zooniverse.org} for an example.}
 % PJM: sadly this Letter is no longer available, Julianne took it
 % down. We can post the talk URL instead.
-% \footnote{See ``Collaborative gravitational lens
-% modelling\dots'' in {\tt http://letters.zooniverse.org} for an
-% example.}
-\footnote{See the ``collaborative modelling'' thread at
-\url{ http://talk.spacewarps.org} for an example.}
+% PS: Letter is available again now.
+%\footnote{See the ``collaborative modelling'' thread at
+% \url{ http://talk.spacewarps.org} for an example.}
 

### 2014-09-22 text/ms.tex
edits

@@
 
-The first two discoveries of gravitational lenses
-, where a galaxy images
-a single background quasar into two (four), brought in their wake the
-first work on lens modelling
-.
-
-%CM:
 The first work on lens modelling
@@
 \citep{1979Natur.279..381W,1980Natur.285..641W}, where a massive
-galaxy causes a background
-quasar to be presented as two or four images.
-
-For the second lens
-(PG1115+080), mass models scored an early success with the prediction
-that one of the lensed images seen would split further into two at
-higher resolution.  That galaxies must sometimes cause multiple images
-had long been expected \citep{1937ApJ....86..217Z}, and it had even
-been argued that the phenomenon could help measure cosmological
-parameters \citep{1964MNRAS.128..307R,1966MNRAS.132..101R}, but
-apparently nobody was expecting that lenses would need detailed
-modelling.  The first observations, however, immediately stimulated
-models.  The reason for that lies in the image separation. Recall that
-image separations are of order of the angular Einstein radius
+galaxy causes a background quasar to be presented as two or four
+images.  For the second lens to be discovered (PG1115+080), mass
+models scored an early success with the prediction that one of the
+lensed images seen would split further into two at higher resolution.
+That galaxies must sometimes cause multiple images had long been
+expected \citep{1937ApJ....86..217Z}, and it had even been argued that
+the phenomenon could help measure cosmological parameters
+\citep{1964MNRAS.128..307R,1966MNRAS.132..101R}, but apparently nobody
+was expecting that lenses would need detailed modelling.  The first
+observations, however, immediately stimulated models.  The reason for
+that lies in the image separation. Recall that image separations are
+of order of the angular Einstein radius
 \begin{equation}
@@
 $\Sigma(x,y)$ be the mass distribution.  It is a mass per unit area,
-i.e., density projected along the line of sight.  It is often given in
-a dimensionless form
+i.e., density projected along the line of sight.  The mass
+distribution is often given in a dimensionless form
 \begin{equation} \label{eq:kappa}

### 2014-09-22 text/ms.tex
edits

@@
 Hence, the lensed images depend on the detailed mass distribution of
-the lensing galaxy, and to uncover that dependence, galaxy lenses
-require models.
+the lensing galaxy.  Galaxy lenses therefore demand models of the mass
+distribution.
 

### 2014-09-23 text/ms.tex
edits
changed affiliation of zooites

@@
 $^{2}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
-$^{3}$The SpaceWarps Collaboration\\
+$^{3}$Zooniverse, c/o Astrophysics Department, University of Oxford, Oxford \\
 $^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
@@
 Hence, the lensed images depend on the detailed mass distribution of
-the lensing galaxy.  Galaxy lenses therefore demand models of the mass
-distribution.
+the lensing galaxy.  Galaxy lenses therefore require models of their
+mass distributions.
 
@@
 recovered lens shape and radial profile, and identify some areas to
-improve. Profile and shape recovery with GLASS has been studied in
+improve. Profile and shape recovery with GLASS has been studied inr
 more detail in \citep{2014arXiv1401.7990C}.
@@
 \item The mass distribution must be centrally concentrated, in two
-  respects.  First, the circularly averaged density must fall away
+  respects.  First, the circularly averaged density must fall awayr
   like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.

### 2014-09-30 text/ms.tex
unkown changes

@@
 recovered lens shape and radial profile, and identify some areas to
-improve. Profile and shape recovery with GLASS has been studied inr
+improve. Profile and shape recovery with GLASS has been studied in
 more detail in \citep{2014arXiv1401.7990C}.
@@
 \item The mass distribution must be centrally concentrated, in two
-  respects.  First, the circularly averaged density must fall awayr
+  respects.  First, the circularly averaged density must fall away
   like $$ \left[(x-x_0)^2+(y-y_0)^2\right]^{-1/2}$$ or more steeply.

### 2014-09-30 text/ms.tex
changed adress of jp

@@
 $^{3}$Zooniverse, c/o Astrophysics Department, University of Oxford, Oxford \\
-$^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
+$^{4}$Exascale Research Computing Lab, Campus Teratec, 2 Rue de la Piquetterie, 91680 Bruyeres-le-Chatel, France\\
 $^{5}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\

### 2014-10-01 text/ms.tex
edits following comments from AM

--- 2014-09-23 text/ms.tex
@@
 
-The encouraging early results from the first \sw lens search,
-carried out on the  $\simeq172$ square degree Canada-France-Hawaii
-Telescope Legacy Survey (CFHTLS) imaging by over 30,000 volunteers
-(Marshall et al, in prep; More et al, in prep) prompted the question:
-could the modelling of the lenses also be done by the volunteers?  If
-so, modelling could help assess prioritise lens candidates at an early
+The encouraging early results from the first \sw lens search, carried
+out on the $\simeq172$ square degree Canada-France-Hawaii Telescope
+Legacy Survey (CFHTLS) imaging by over 30,000 volunteers (Marshall et
+al, in prep; More et al, in prep) prompted the question: could the
+modelling of the lenses also be done by the volunteers?  If so,
+modelling could help assess prioritise lens candidates at an early
 stage, which would be very useful with new wide-field and sensitive
 surveys, which will yield thousands of lens candidates.  There are
-several software tools for lens modelling available, and work has
-been done on generic interfaces \citep[e.g.][]{2014A&C.....5...28L}.
-Indeed, some early designs for \sw included a prototype lens modelling
-tool \citep{2010AAS...21543527N}. Moreover, some \sw volunteers are
-quite experienced from earlier projects, having spent a thousand hours
-or more with data, and are very interested in more demanding
-projects.  The interests of citizen-science communities are just
-beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it
-is clear that some volunteers welcome open-ended challenges, and
-sometimes these have led to new scientific results: one example is the
-discovery of an exceptional extra-solar planet
-\citep{2013ApJ...768..127S}; another is the development of new
-algorithms for protein folding \citep{Khatib22112011}.  All these are
-grounds for optimism.  There is, however, a basic difficulty in strong
-gravitational lensing. Lensed images do not look much like their
-source, and still less do they resemble the lensing-mass
-distribution.  To model a lens, one needs either to do a lot of random
-guessing, or to have a good intuition for what works.
+several software tools for lens modelling available, and work has been
+done on generic interfaces \citep[e.g.][]{2014A&C.....5...28L}.  Some
+early designs for \sw included a prototype lens modelling tool
+\citep{2010AAS...21543527N}. Moreover, some \sw volunteers are quite
+experienced from earlier projects, having spent a thousand hours or
+more with data, and are very interested in more demanding projects.
+The interests of citizen-science communities are just beginning to be
+studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that
+some volunteers welcome open-ended challenges, and sometimes these
+have led to new scientific results: one example is the discovery of an
+exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
+the development of new algorithms for protein folding
+\citep{Khatib22112011}.  All these are grounds for optimism.  There
+is, however, a basic difficulty in strong gravitational
+lensing. Lensed images do not look much like their source, and still
+less do they resemble the lensing-mass distribution.  To model a lens,
+one needs either to do a lot of random guessing, or to have a good
+intuition for what works.
 
@@
 
-\spl is not part of \sw, but it makes use of the \sw infrastructure,
-in particular, the image database and the discussion
+\spl is a mass modelling program that makes use of the \sw
+infrastructure, in particular, the image database and the discussion
 forum.\footnote{\url{ http://talk.spacewarps.org}} The forum is
@@
 was made welcome to join at any time.
-% {\bf Phil asks: how much training was necessary before the volunteers
-% understood how to work the interface? Could this training be done with
-% no input from the \spl developers, in future? It would be good to
-% comment on how scalable this approach to modeling is.}
-
-% PJM: The following paragraph is not accurate: the manipulation of
-%  FITS images in the Tools viewer, for example, is done client side.
-%  In future I very much hope to see lens modeling enabled in this
-%  mode as well!
-%
-% Another aspect of \sw (and of the Zooniverse family of projects in
-% general) that \spl adopts, is to keep all user interaction in a web
-% browser, with intensive computing done on a server.  Several
-% interesting design and implementation issues arise from these
-% considerations, but we will not go into them in this paper,
-% concentrating instead on aspects specific to lens modelling.
-%
 
@@
   \caption{Screen grab of \spl in action. A \sw image has been loaded
-    in, re-centred and zoomed, and 5 images and associated time delay
-    surface ``spaghetti'' contours suggested, using the marking tools
-    associated with the buttons along the top of the panel. The mass
-    model is optimized server-side when the right-most button is
+    in, re-centred and zoomed.  Five images and the associated
+    ``spaghetti'' contours have then been suggested, using the marking
+    tools associated with the buttons along the top of the panel. The
+    mass model is generated server-side when the right-most button is
     pressed.}
@@
 modelling\dots'' in {\tt http://letters.zooniverse.org} for an example.}
-% PJM: sadly this Letter is no longer available, Julianne took it
-% down. We can post the talk URL instead.
-% PS: Letter is available again now.
-%\footnote{See the ``collaborative modelling'' thread at
-% \url{ http://talk.spacewarps.org} for an example.}
+% This Letter was inaccessible for a while is available again now.
 
@@
 relevant here is that the sims were of three kinds, as follows.
-
 \begin{enumerate}
-  \item Imitating lensed quasars: having a singular elliptical
-    isothermal lens (SIE) plus constant external shear, and a circular
-    Gaussian source.
-  \item Emulating lensed galaxies: similar to the above, but with an
-    elliptical de Vaucouleurs source.
-  \item Resembling cluster lenses: having a source as above, but a
-    more complicated lens, with one dominant elliptical SIE and
-    one or more perturbing elliptical SIEs, plus a circular NFW
-    \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent
-    the underlying dark matter distribution.
+\item Lensed quasars: The lens is modelled as a singular isothermal
+  ellipsoid (SIE) and a constant external shear whereas the quasar is
+  represented with a circular Gaussian source whose size is given by
+  the point spread function (PSF) in each imaging band.
+\item Galaxy-scale lenses: The lens model is the same as above whereas
+  the background galaxy is modelled as an elliptical de~Vaucouleurs.
+\item Groups-scale lenses: The lens model includes SIE models for the
+  central galaxy and the inner group members, plus a circular NFW
+  (ref) to represent the underlying dark matter distribution and the
+  background galaxy model stays the same as galaxy-scale lenses.
 \end{enumerate}
@@
   \caption[result 6919 (ASW0002z6f)]{Another configuration of arc plus
-    counter-image; here the arc is closer to the lensing galaxy than
-    the counter-image. (See \secref{example_models} for details.)
+    counter-image: an arc and counter image where the arc is closer to
+    the lensing galaxy than the counter image. (See
+    \secref{example_models} for details.)
     \vspace{1em}}
@@
 $$ \begin{matrix}
-\hbox{marked-up image} \qquad &\hbox{model synthetic image} \\
+\hbox{marked-up CFHTLS image} \qquad &\hbox{model synthetic image} \\
 t(x,y)                               &\hbox{model } t(x,y) \\
@@
 poor, the mean Einstein radius overestimate was 10\%, with a 15\%
-standard deviation (shown by the light grey band in \Figref{ER_per_sim}).
-The expert models show a similar bias,
-with standard deviation 10\% (the dark grey band in \Figref{ER_per_sim}).
-One source of this systematic error is that it is difficult to center
-of the lens accurately: an offset leads to a flatter mass profile for
-the model compared to the simulation.
+standard deviation (shown by the light grey band in
+\Figref{ER_per_sim}).  The expert models show a similar bias, with
+standard deviation 10\% (the dark grey band in \Figref{ER_per_sim}).
+One source of this systematic error is that it is difficult to centre
+the lens accurately: an offset leads to a flatter mass profile for the
+model compared to the simulation.
 
@@
 \item \spl tends to overestimate the Einstein radius (evident from
-  \Figref{ER_per_sim}), and it is also apparent that the model density
-  profiles tend to be too shallow.  The likely explanation is that
-  while the sims are steeply peaked at the centre, the pixellated mass
-  model fixes a comparatively large area near the central at constant
-  density. Allowing smaller pixels in the centre region, thus enabling
-  a steeper centre \citep[similar to the ``high resolution'' feature
-    implemented in][]{2014arXiv1401.7990C} may remove this bias.  The
-  use of simply parameterised, appropriately steeply-profiled models
-  would also avoid the problem.
+  \Figref{ER_per_sim}), and the model density profiles tend to be
+  shallower than the SIE model used for generating the sims.  The
+  likely explanation is that while the sims are steeply peaked at the
+  centre, the pixellated mass model fixes a comparatively large area
+  near the central at constant density. Allowing smaller pixels in the
+  centre region, thus enabling a steeper centre \citep[similar to the
+    ``high resolution'' feature implemented in][]{2014arXiv1401.7990C}
+  may remove this bias.  The use of simply parameterised,
+  appropriately steeply-profiled models would also avoid the problem.
 \item Currently, \spl does not attempt to model the source shape; the
@@
   on the synthetic image, and on whether the mass distribution and the
-  arrival-time surface show suspicious features.
+  arrival-time surface show suspicious features.  Another possibility
+  would be use the \spl models as a feeder to a different
+  lens-modelling program which already implements source-profile
+  fitting.
 \item Another limitation so far in \spl is that the lens is assumed to
@@
 S.~More and Verma.  Modelling was done by Baeten, Cornen, Macmillan,
-Odermatt, Saha and Wilcox, with post-modelling analysis by Kueng and
+Odermatt, Saha and Wilcox, with post-modelling analysis by Küng and
 Saha.  All authors participated in writing and editing the manuscript.
@@
 We thank the Swiss Society for Astrophysics and Astronomy and the
-Swiss Academy of Sciences. AV is supported by a research fellowship
+Swiss Academy of Sciences. The work of AM and SM was supported by
+World Premier International Research Center Initiative (WPI
+Initiative), MEXT, Japan. AV is supported by a research fellowship
 from the Leverhulme Trust. This work was supported in part by the U.S.

### 2014-10-01 text/ms.tex
edits following comments from AM

@@
   the background galaxy is modelled as an elliptical de~Vaucouleurs.
-\item Groups-scale lenses: The lens model includes SIE models for the
+\item Group-scale lenses: The lens model includes SIE models for the
   central galaxy and the inner group members, plus a circular NFW
-  (ref) to represent the underlying dark matter distribution and the
-  background galaxy model stays the same as galaxy-scale lenses.
+  \citep{1996ApJ...462..563N,1997ApJ...490..493N} to represent the
+  underlying dark matter distribution and the background galaxy model
+  stays the same as galaxy-scale lenses.
 \end{enumerate}

### 2014-10-06 text/ms.tex
added keywords
edits following comments by SM

@@
 \begin{equation}
+\Theta_{\rm E}
 \sim \left(\frac{4GM}{c^2 D_L}\right)^{1/2}
@@
 where $D_L$ is the distance to the lens, and $M$ its mass.  A lensing
-galaxy with $M\sim10^{11}M_\odot$ at $10^9\rm\,pc$ would cause image
+galaxy with $M\sim10^{11}M_\odot$ at $\sim1\rm\,Gpc$ would cause image
 separations of $\sim1''$, which is comparable to the size of the
@@
 modelling of the lenses also be done by the volunteers?  If so,
-modelling could help assess prioritise lens candidates at an early
-stage, which would be very useful with new wide-field and sensitive
-surveys, which will yield thousands of lens candidates.  There are
-several software tools for lens modelling available, and work has been
-done on generic interfaces \citep[e.g.][]{2014A&C.....5...28L}.  Some
-early designs for \sw included a prototype lens modelling tool
+modelling could help aprioritise lens candidates at an early stage,
+which would be very useful with new wide-field and sensitive surveys,
+which will yield thousands of lens candidates.  There are several
+software tools for lens modelling available, and work has been done on
+generic interfaces \citep[e.g.][]{2014A&C.....5...28L}.  Some early
+designs for \sw included a prototype lens modelling tool
 \citep{2010AAS...21543527N}. Moreover, some \sw volunteers are quite
-experienced from earlier projects, having spent a thousand hours or
-more with data, and are very interested in more demanding projects.
-The interests of citizen-science communities are just beginning to be
-studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but it is clear that
-some volunteers welcome open-ended challenges, and sometimes these
-have led to new scientific results: one example is the discovery of an
-exceptional extra-solar planet \citep{2013ApJ...768..127S}; another is
-the development of new algorithms for protein folding
-\citep{Khatib22112011}.  All these are grounds for optimism.  There
-is, however, a basic difficulty in strong gravitational
-lensing. Lensed images do not look much like their source, and still
-less do they resemble the lensing-mass distribution.  To model a lens,
-one needs either to do a lot of random guessing, or to have a good
-intuition for what works.
+experienced from earlier projects, having individually spent a
+thousand hours or more with data, and are very interested in more
+demanding projects.  The interests of citizen-science communities are
+just beginning to be studied \citep[e.g.,][]{2013AEdRv..12a0106J}, but
+it is clear that some volunteers welcome open-ended challenges, and
+sometimes these have led to new scientific results: one example is the
+discovery of an exceptional extra-solar planet
+\citep{2013ApJ...768..127S}; another is the development of new
+algorithms for protein folding \citep{Khatib22112011}.  All these are
+grounds for optimism.  There is, however, a basic difficulty in strong
+gravitational lensing. Lensed images do not look much like their
+source, and still less do they resemble the lensing-mass distribution.
+To model a lens, one needs either to do a lot of random guessing, or
+to have a good intuition for what works.
 
@@
 remark that \spl is designed to be friendly to the forum style of
-citizen-science projects, and enable incremental collaborative
+citizen-science projects, and enables incremental collaborative
 model refinement by different people, without sacrificing any of the
@@
 a circle, and $(x_\subbullet,y_\subbullet)$ to indicate the integrated
-mass within the circle.
+mass within the circle.  Appendix~\ref{more-theory} relates this
+expression to the better-known explicit form for the gravitational
+time delay.
 
@@
 model to the \spl archive, at which point it is assigned an unique
-URL.  He can also modify his input and try again, or discard the
+URL.  They can also modify the input and try again, or discard the
 attempt altogether.  After archiving, there can be discussion among

### 2014-10-06 text/ms.tex
edits following comments by SM

@@
 expected the parities and time-ordering to be more difficult.  While
-the \spl tutorials had provided some general rules of thumb, to be
-consistently correct with the time-ordering, a modeller needs to
-develop some intuition for arrival time surfaces.  This is an area
-where experience and tutorials training could improve results at a
-later stage, and correspondingly, feedback on the difficulties
-modellers encounter can help improve the tutorial materials.
+the \spl tutorials had provided general guidelines, to be consistently
+correct with the time-ordering, a modeller needs to develop some
+intuition for arrival time surfaces.  This is an area where experience
+and tutorials training could improve results at a later stage, and
+correspondingly, feedback on the difficulties modellers encounter can
+help improve the tutorial materials.
 

### 2014-10-08 text/ms.tex
unchecked edits

@@
 
-Modelling with \spl involves three stages, markup of the image,
-followed by intensive numerical computation carried out on a
-server in the background, followed by review
-of diagnostics and possible discussion.  We now describe each of
-these.
+Modelling with \spl involves three stages, (1)~markup of the image,
+followed by (2)~intensive numerical computation carried out on a
+server in the background, followed by (3)~review of diagnostics and
+possible discussion.  Human interaction is essential to the first and
+third stages, while stage~2 is completely automated. We now describe
+the three stages.
 

### 2014-10-08 text/ms.tex
edits following comments by SM

@@
 \citep{1979Natur.279..381W,1980Natur.285..641W}, where a massive
-galaxy causes a background quasar to be presented as two or four
+galaxy causes a background quasar to appear as two or four
 images.  For the second lens to be discovered (PG1115+080), mass
@@
 \end{enumerate}
-The {\tt gravlens} program \citep{2001astro.ph..2340K} was used.
-Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
-lenses follow equations (33--35) of that work, with core radius set to
-zero.  The NFW lens is in equations (48) and (50), while shear is the
-$\gamma$ term in equation (76).
 
@@
 \FloatBarrier
+
+The {\tt gravlens} program \citep{2001astro.ph..2340K} was used.
+Formulas for the lenses appear in \cite{2001astro.ph..2341K}. The SIE
+lenses follow equations (33--35) of that work, with core radius set to
+zero.  The NFW lens is in equations (48) and (50), while shear is the
+$\gamma$ term in equation (76).
 

### 2014-10-08 text/ms.tex
edit corrections following comments by SM

@@
 modelling of the lenses also be done by the volunteers?  If so,
-modelling could help aprioritise lens candidates at an early stage,
+modelling could help prioritise lens candidates at an early stage,
 which would be very useful with new wide-field and sensitive surveys,

### 2014-10-23 text/ms.tex
edits following AV

@@
 \newcommand{\spl}{SpaghettiLens\xspace}
-\newcommand{\sw}{SpaceWarps\xspace}
+\newcommand{\sw}{Space Warps\xspace}
 
@@
 $^{6}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
-$^{7}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
+$^{7}$Sub-department of Astrophysics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford, OX1 3RH, UK\\
 }
@@
 modelling program (GLASS), but enables the input to this code to be
-provided in a novel way, via a user-provided diagram that is
+provided in a novel way, via a user-generated diagram that is
 essentially a sketch of an arrival-time surface.
 
-An implementation of this method, \spl, is tested in a modelling
-challenge that used 29 simulated lenses that had been drawn from the
-set created for the \sw citizen-science lens search.  We find that
-volunteers from this online community asserted the image parities and
-time ordering consistently in some lenses, but made errors in other
-lenses depending on the image morphology. While errors in image parity
-and time ordering lead to large errors in the mass distribution, the
-enclosed mass was found to be more robust: the model-derived Einstein
-radii found by the volunteers were consistent with those produced by
-one of the professional team, suggesting that given the appropriate
-tools, gravitational lens modelling is a data analysis activity that
-can be crowd-sourced to good effect. Ideas for improvement are
-discussed; these include (a)~overcoming the tendency of the models to
-be shallower than the correct answer in test cases, leading to
-systematic over-estimation of the Einstein radius by 10\% at present,
-and (b)~detailed modelling of arcs.
+We report on an implementation of this method, \spl, that has been
+tested in a modelling challenge using 29 simulated lenses drawn from a
+larger set created for the \sw citizen science strong lens search. We
+find that volunteers from this online community asserted the image
+parities and time ordering consistently in some lenses, but made
+errors in other lenses depending on the image morphology. While errors
+in image parity and time ordering lead to large errors in the mass
+distribution, the enclosed mass was found to be more robust: the
+model-derived Einstein radii found by the volunteers were consistent
+with those produced by one of the professional team, suggesting that
+given the appropriate tools, gravitational lens modelling is a data
+analysis activity that can be crowd-sourced to good effect. Ideas for
+improvement are discussed; these include (a)~overcoming the tendency
+of the models to be shallower than the correct answer in test cases,
+leading to systematic over-estimation of the Einstein radius by 10\%
+at present, and (b)~detailed modelling of arcs.
 
@@
 \begin{keywords}
+gravitational lensing: strong, methods: numerical
 \end{keywords}
@@
 the present work, however, spaghetti diagrams are the {\em input\/}
-through which the modeller tells the program what to do.
+through which the modeller tells the \spl program what to do.
 
@@
   would be use the \spl models as a feeder to a different
-  lens-modelling program which already implements source-profile
+  lens-modelling program that already implements source-profile
   fitting.

### 2014-10-28 text/ms.tex

@@
 where $D_L$ is the distance to the lens, and $M$ its mass.  A lensing
-galaxy with $M\sim10^{11}M_\odot$ at $\sim1\rm\,Gpc$ would cause image
+galaxy with $M\sim10^{11}\rm\,M_\odot$ at $\sim1\rm\,Gpc$ would cause image
 separations of $\sim1''$, which is comparable to the size of the

### 2014-11-04 text/ms.tex
final edits

@@
 %
-$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057 Zurich, Switzerland\\
+$^{1}$Physik-Institut, University of Zurich, Winterthurerstrasse 190, 8057 Zurich, Switzerland\\
 $^{2}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
@@
 project\footnote{\url{http://www.spacewarps.org}} in which volunteers
-are presented sky-survey images and invited to identify lens
+are presented sky-survey images and are invited to identify lens
 candidates, by eye.  Simulated lenses are mixed in with the data,
@@
 \citep{2014arXiv1401.7990C}.  We will not go into software details in
-this paper, but instead concentrate on lens modelling per se, but we
+this paper, instead concentrate on lens modelling per se, but we
 remark that \spl is designed to be friendly to the forum style of
@@
 To illustrate, \figref{kapenc_compare_faulty} compares the circularly
-averaged mass profiles three different models of one particular lens;
+averaged mass profiles of three different models of one particular lens;
 two of the models are shown in Figures~\ref{fig:7025} and
@@
 Initiative), MEXT, Japan. AV is supported by a research fellowship
-from the Leverhulme Trust. This work was supported in part by the U.S.
+from the Leverhulme Trust. RK is supported by the Swiss National
+Science Foundation. This work was supported in part by the U.S.
 Department of Energy under contract number DE-AC02-76SF00515.

### 2014-11-04 text/compress.sh
compress cmd

@@ -1,0 +1,1 @@
+gs -dNOPAUSE -dBATCH -sDEVICE=pdfwrite -dCompatibilityLevel=1.4 -dPDFSETTINGS=/ebook -sOutputFile=ms_compressed.pdf ms.pdf

### 2015-01-30 text/ms.tex
applying corrections from mnras

@@
 $^{2}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
-$^{3}$Zooniverse, c/o Astrophysics Department, University of Oxford, Oxford \\
+$^{3}$Zooniverse, c/o Astrophysics Department, University of Oxford, Oxford OX1 3RH, UK \\
 $^{4}$Exascale Research Computing Lab, Campus Teratec, 2 Rue de la Piquetterie, 91680 Bruyeres-le-Chatel, France\\
@@
 \begin{equation} \label{eq:tgeom}
-\tgeom(x,y) \propto (x-x_s)^2 + (y-s_y)^2 ,
+\tgeom(x,y) \propto (x-x_s)^2 + (y-y_s)^2 ,
 \end{equation}
@@
 confidence region from the modelling.  The definition \eqref{eq:effRE}
-for \ERf corresponds to crossing the dashed horizontal line at $10^0$:
+for \ERf corresponds to crossing the dashed horizontal line at $1$:
 the red curve crosses the dashed line at the actual Einstein radius
@@
   would mean that spaghetti contours with more than one maximum can be
-  allowed.  For examples, see Figure 5c in \citep{2001ApJ...557..594R}
-  and Figure 4b in \cite{2003ApJ...590...39K}.
+  allowed.  For examples, see Fig. 5c in \citep{2001ApJ...557..594R}
+  and Fig. 4b in \cite{2003ApJ...590...39K}.
 \item At present, a single false-color composite is used as the data.
@@
 t_{\rm geom} &= \frac{(1+z_L)}{2c} \frac{d_S}{d_L d_{LS}}
-\left[ (x-s_x)^2 + (y-s_y)^2 \right] \\
+\left[ (x-x_s)^2 + (y-y_s)^2 \right] \\
 \nabla^2 t_{\rm grav} &= -(1+z_L)\frac{8\pi G}{c^3} \, \Sigma(x,y) \\
@@
 equation \eqref{eq:tgeom} with the proportionality filled in, and with
-the source offset at $(s_x,s_y)$ rather than at the origin.
+the source offset at $(x_s,y_s)$ rather than at the origin.
 
